### 2004-11-09 paper/trunk/systemfit.tex
Initial revision
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@555 591db2bd-e0a0-4d22-8c19-2192dbb7da31

\documentclass[article]{jss}

\author{Arne Henningsen\\University of Kiel\And 
        Jeff D. Hamann\\ ??? }

\title{systemfit: A Package to Estimate\\
       Simultaneous Equation Systems in \proglang{R}}

%% for pretty printing and a nice hypersummary also set:
\Plainauthor{Arne Henningsen, Jeff D. Hamann} %% comma-separated
\Plaintitle{systemfit: A Package to Estimate Simultaneous Equation Systems 
            in R} %% without formatting
\Shorttitle{systemfit} %% a short title (if necessary)

%% an abstract and keywords
\Abstract{
   Many statistical analysis are based on models that consist of 
   more than one equation.
   These equations should be fitted simultaneously rather than equation 
   by equation.
   The package \pkg{systemfit} provides the capability to estimate systems 
   of linear (and non-linear???) equations in \proglang{R}. 
}
\Keywords{R, equation system, seemingly unrelated regression}
%% at least one keyword must be supplied

%% publication information
%% NOTE: This needs to filled out ONLY IF THE PAPER WAS ACCEPTED.
%% If it was not (yet) accepted, leave them commented.
%% \Volume{13}
%% \Issue{9}
%% \Month{September}
%% \Year{2004}
%% \Submitdate{2004-09-29}
%% \Acceptdate{2004-09-29}

%% The address of (at least) one author should be given
%% in the following format:
\Address{
   Arne Henningsen\\
   Department of Agricultural Economics\\
   University of Kiel\\
   D-24098 Kiel, Germany\\
   E-mail: \email{ahenningsen@agric-econ.uni-kiel.de}\\
   URL: \url{http://www.uni-kiel.de/agrarpol/ahenningsen/}\\
   \\
   Jeff D. Hamann\\
   e-mail: \email{jeff.hamann@forestinformatics.com}\\
   URL: \url{http://www.forestinformatics.com}\\
}
%% It is also possible to add a telephone and fax number
%% before the e-mail in the following format:
%% Telephone: +43/1/31336-5053
%% Fax: +43/1/31336-734

%% for those who use Sweave please include the following line (with % symbols):
%% need no \usepackage{Sweave.sty}

%% end of declarations %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Many theoretical models consist of more than one equation.
If these models are econometrically estimated, 
the equations generally should not be estimated individually, 
but all together, because the disturbance terms of these 
equations are likely contemporaneously correlated.
Estimating all equations simultaneously and
taking the covariance structure of the residuals into account 
leads to efficient estimates, 
while ignoring this correlation leads to an inefficient 
parameter estimation \citep{zellner62}.

Another reason to estimate an equation system simultaneously are
cross-equation parameter restrictions. These restrictions can 
be tested and/or imposed only in a simultaneous estimation
approach. Especially the economic theory suggests many 
cross-equation restriction.

The \pkg{systemfit} package provides the capability to estimate 
linear (and non-linear???) equation systems in \proglang{R}.
Compared to other statistics and econometrics software 
the (advanced) user can control many important details of the 
estimation procedure. On the other hand these user options have
reasonable defaults that also beginners can easily use the
provided functions.

This paper is organized as follows: 
In section \ref{sec:Theory} we introduce the mathematics of estimating
equation systems and give a comprehensive overview over all
relevant issues regarding this topic.
Section \ref{sec:UsingSystemfit} demonstrates how to run \pkg{systemfit},
especially how the features presented in the previous chapter
can be used.
Finally, a summary and an outlook are given in section \ref{sec:Summmary}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Estimating equation systems}\label{sec:Theory}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Consider a system of $G$ equations, where the $i$th equation is of
the form 
\begin{equation}
   y_{i} = X_i \beta_i + u_i, \quad i = 1, 2, \ldots, G
\end{equation}
where $y_i$ is a vector of the dependent variable,
$X_i$ is a matrix of the exogenous variables,
$\beta_i$ is the coefficient vector and
$u_i$ is a vector of the disturbance terms of the $i$th equation.

We can write the 'stacked' system as
\begin{equation}
   \left[ \begin{array}{c}
      y_1 \\ y_2\\ \vdots\\ y_G
   \end{array} \right] = 
   \left[ \begin{array}{cccc}
      X_1 & 0 & \cdots & 0\\
      0 & X_2 & \cdots & 0\\
      \vdots & \vdots & \ddots & \vdots\\
      0 & 0 & \cdots & X_G
   \end{array}\right] \cdot 
   \left[ \begin{array}{c}
      \beta_1 \\ \beta_2 \\ \vdots\\ \beta_G
   \end{array} \right] +
   \left[ \begin{array}{c}
      u_1 \\ u_2 \\ \vdots\\ u_G
   \end{array} \right]
\end{equation}

or more simply as
\begin{equation}
   y = X \beta + u 
\end{equation}

We assume that there is no correlation of the disturbance terms 
across observations:
\begin{equation}
   E \left( u_{it} \, u_{jt^*} \right) = 0 
   \; \forall \; t \neq t^*
\end{equation}
where $i$ and $j$ indicate the equation number 
and $t$ and $t^*$ denote the observation number.

However, we explicitly allow for contemporaneous correlation:
\begin{equation}
   E \left( u_{it} \, u_{jt} \right) = \sigma_{ij}
\end{equation}

Thus, the covariance matrix of the total system is
\begin{equation}
   E \left( u \, u' \right) = \Omega = \Sigma \otimes I
\end{equation}
where $\Sigma = \left[ \sigma_{ij} \right]$ and $I$ is an
identity matrix.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Parameter estimation}

\subsubsection{Ordinary Least Squares estimation}

The Ordinary Least Squares (OLS) estimator of the system 
is obtained by
\begin{equation}
   \widehat{\beta} = \left( X'X \right)^{-1} X'y
\end{equation}

These estimates are efficient only if the disturbance terms are not 
contemporaneously correlated, which means 
\begin{equation}
   \sigma_{ij} = 0 \; \forall \; i \neq j
\end{equation}

If the whole system is treated as one single equation, 
the variance-covariance matrix of the estimated parameters is
\begin{equation}
   Var \left[ \widehat{\beta} \right] = \sigma^2 \left( X'X \right)^{-1}
\end{equation}
with $\sigma^2 = E \left( u' u \right)$.
This assumes that the disturbances of all equations have the
same variance.

If the disturbance terms of the individual equations 
are allowed to have different variances, 
the variance-covariance matrix of the estimated parameters is
\begin{equation}
   Var \left[ \widehat{\beta} \right] = \left( X' \Omega^{-1} X \right)^{-1}
   \label{eq:OLSvcm2}
\end{equation}
with $\Omega = \Sigma \otimes I$, 
$\sigma_{ij} = 0 \; \forall \; i \neq j$ and
$\sigma_{ii} = E \left( u_i' u_i \right)$.

If no cross-equation restrictions are imposed, the simultaneous 
OLS estimation of the system leads to the same parameter estimates 
as an equation-wise OLS estimation. The variance-covariance matrix 
of the parameters from an equation-wise OLS estimation is equal to 
the variance-covariance matrix obtained by equation 
(\ref{eq:OLSvcm2}).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Weighted least squares}

The Weighted Least Squares (WLS) estimator of the system 
is obtained by
\begin{equation}
   \widehat{\beta} = \left( X' \Omega^{-1} X \right)^{-1} X' \Omega^{-1} y
\end{equation}
with $\Omega = \Sigma \otimes I$, 
$\sigma_{ij} = 0 \; \forall \; i \neq j$ and
$\sigma_{ii} = E \left( u_i' u_i \right)$.

Like the OLS estimates these estimates are only efficient 
if the disturbance terms are not contemporaneously correlated.

The variance-covariance matrix of the estimated parameters is
\begin{equation}
   Var \left[ \widehat{\beta} \right] = \left( X' \Omega^{-1} X \right)^{-1}
\end{equation}

If no cross-equation restrictions are imposed, 
the parameter estimates and the variance-covariance matrix 
of the parameters are equal to the OLS estimates%
\footnote{The variance-covariance matrices of the 
estimated parameters might differ, 
if the WLS estimation is based on a different formula 
to calculate the (diagonal) variance-covariance 
matrix of the residuals (see section ????).}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Seemingly unrelated regression}

When the disturbances are contemporaneously correlated, a Generalized 
Least Squares (GLS) estimation leads to efficient parameter estimates.
In this case the GLS is generally called "Seemingly Unrelated Regression" 
(SUR) (see \citealp{zellner62}). 
Note, while an unbiased OLS or WLS estimation requires only that 
the regressors and the disturbance terms of each single 
equation are uncorrelated,
a consistent SUR estimation requires that all disturbance terms and all 
regressors are uncorrelated:
\begin{equation}
   E \left( u | X \right) = 0
\end{equation}

The SUR estimator can be obtained by:%
\footnote{To calculate $\Omega^{-1}$ it is not necessary to invert 
the (huge) $\Omega$ matrix, 
because it is sufficient to invert the (small) $\Sigma$ matrix and
calculate it by $\Omega^{-1} = \Sigma^{-1} \otimes I$.}

\begin{equation}
   \widehat{\beta} = \left( X' \Omega^{-1} X \right)^{-1} X' \Omega^{-1} y
\end{equation}
with $\Omega = \Sigma \otimes I$ and
$\sigma_{ij} = E \left( u_i' u_j \right)$.

And the variance-covariance matrix of the estimated parameters is
\begin{equation}
   Var \left[ \widehat{\beta} \right] = \left( X' \Omega^{-1} X \right)^{-1}
\end{equation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Two-stage least squares estimation}

If the regressors of one or more equations are correlated 
with the disturbances ($E \left( u_i | X_i \right) \neq 0$), 
the estimated coefficients are biased.
This can be circumvented by an instrumental variable (IV) estimation.
The instrumental variables for each equation $H_i$ 
can be either different or identical for all equations.
The instrumental variables of each equation may not be correlated with 
the disturbance terms of the corresponding equation 
($E \left( u_i | H_i \right) = 0$).

At the first stage new ('fitted') regressors are obtained by
\begin{equation}
   \widehat{X_i} = H_i \left( H_i' H_i \right)^{-1} H_i' X
\end{equation}

At the second stage the unbiased two-stage least squares estimates
of $\beta$ are obtained by:
\begin{equation}
   \widehat{\beta} = \left( \widehat{X}' \widehat{X} \right)^{-1} 
   \widehat{X}' y 
   \label{eq:beta2sls}
\end{equation}

If the whole system is treated as one single equation, 
the variance-covariance matrix of the estimated parameters is
\begin{equation}
   Var \left[ \widehat{\beta} \right] = \sigma^2 \left( \widehat{X}'
   \widehat{X} \right)^{-1}
\end{equation}
with $\sigma^2 = E \left( u' u \right)$.

If the disturbance terms of the individual equations 
are allowed to have different variances, 
the variance-covariance matrix of the estimated parameters is
\begin{equation}
   Var \left[ \widehat{\beta} \right] = \left( \widehat{X}' \Omega^{-1} 
   \widehat{X} \right)^{-1}
\end{equation}
with $\Omega = \Sigma \otimes I$, 
$\sigma_{ij} = 0 \; \forall \; i \neq j$ and
$\sigma_{ii} = E \left( u_i' u_i \right)$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Weighted two-stage least squares estimation}

The Weighted Two-Stage Least Squares (W2SLS) estimator of the system 
is obtained by
\begin{equation}
   \widehat{\beta} = \left( \widehat{X}' \Omega^{-1} \widehat{X} 
   \right)^{-1} \widehat{X}' \Omega^{-1} y
\end{equation}
with $\Omega = \Sigma \otimes I$, 
$\sigma_{ij} = 0 \; \forall \; i \neq j$ and
$\sigma_{ii} = E \left( u_i' u_i \right)$.

The variance-covariance matrix of the estimated parameters is
\begin{equation}
   Var \left[ \widehat{\beta} \right] = \left( \widehat{X}' \Omega^{-1} 
   \widehat{X} \right)^{-1}
\end{equation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Three-stage least squares estimation}

If the regressors are correlated with the disturbances 
($E \left( u | X \right) \neq 0$) and 
the disturbances are contemporaneously correlated, 
an Instrumental Variable (IV) Generalized Least Squares (GLS) 
estimation leads to consistent and efficient estimates. This estimation 
procedure is generally called "Three-stage Least Squares" (3SLS).

The standard 3SLS estimator can be obtained by:
\begin{equation}
   \widehat{\beta} = \left( \widehat{X}' \Omega^{-1} \widehat{X} 
   \right)^{-1} \widehat{X}' \Omega^{-1} y
\end{equation}
with $\Omega = \Sigma \otimes I$ and
$\sigma_{ij} = E \left( u_i' u_j \right)$.

Its variance-covariance matrix is:
\begin{equation}
   Var \left[ \widehat{\beta} \right] = \left( \widehat{X}' \Omega^{-1} 
   \widehat{X} \right)^{-1}
   \label{eq:cov3sls}
\end{equation}

While an unbiased 2SLS or W2SLS estimation requires only that 
the instrumental variables and the disturbance terms of each single 
equation are uncorrelated,
\cite{schmidt90} points out that this estimator is only consistent 
if all disturbance terms and all instrumental variables are uncorrelated:
\begin{equation}
   E \left( u | H \right) = 0
\end{equation}
with
\begin{equation}
   H =
   \left[ \begin{array}{cccc}
      H_1 & 0 & \cdots & 0\\
      0 & H_2 & \cdots & 0\\
      \vdots & \vdots & \ddots & \vdots\\
      0 & 0 & \cdots & H_G
   \end{array}\right] \cdot 
\end{equation}

Since there might be occasions where this cannot be avoided, 
\cite{schmidt90} analyses other approaches to obtain 3SLS estimators:

One of these approaches is based on instrumental variable estimation
(IV-3SLS):
\begin{equation}
   \widehat{\beta} = \left( \widehat{X}' \Omega^{-1} X 
   \right)^{-1} \widehat{X}' \Omega^{-1} y
\end{equation}

The variance-covariance matrix of this IV-3SLS estimator is:
\begin{equation}
   Var \left[ \widehat{\beta} \right] = \left( \widehat{X}' \Omega^{-1} 
   X \right)^{-1}
\end{equation}

Another approach is based on the Generalized Method of Moments (GMM) 
estimator (GMM-3SLS):
\begin{equation}
   \widehat{\beta} = \left( X' H \left( H' \Omega H \right)^{-1}
   H' X \right)^{-1} X' H \left( H' \Omega H \right)^{-1} H' y
   \label{eq:beta3slsGMM}
\end{equation}

The variance-covariance matrix of the GMM-3SLS estimator is:
\begin{equation}
   Var \left[ \widehat{\beta} \right] = 
   \left( X' H \left( H' \Omega H \right)^{-1} H' X \right)^{-1}
\end{equation}

A fourth approach developed by \cite{schmidt90} himself is:
\begin{equation}
   \widehat{\beta} = \left( \widehat{X}' \Omega^{-1} \widehat{X} 
   \right)^{-1} \widehat{X}' \Omega^{-1} 
   H \left( H' H \right)^{-1} H' y 
   \label{eq:beta3slsSchmidt}
\end{equation}

The variance-covariance matrix of this estimator is:
\begin{equation}
   Var \left[ \widehat{\beta} \right] = 
   \left( \widehat{X}' \Omega^{-1}  \widehat{X} \right)^{-1} 
   \widehat{X}' \Omega^{-1} H \left( H' H \right)^{-1} H' \Omega 
   H \left( H' H \right)^{-1} H' \Omega^{-1} \widehat{X}
   \left( \widehat{X}' \Omega^{-1}  \widehat{X} \right)^{-1}
\end{equation}

The econometrics software EViews uses following approach:
\begin{equation}
   \widehat{\beta} = \widehat{\beta}_{2SLS} + 
   \left( \widehat{X}' \Omega^{-1} \widehat{X} \right)^{-1} 
   \widehat{X}' \Omega^{-1} \left( y - X \widehat{\beta}_{2SLS} \right)
\end{equation}
with $\widehat{\beta}_{2SLS}$ is the two-stage least squares estimator
(equation \ref{eq:beta2sls}).
EViews uses the standard 3SLS formula (equation (\ref{eq:cov3sls})) to 
calculate the variance-covariance matrix of the 3SLS estimator.

If the same instrumental variables are used in all equations 
($H_1 = H_2 = \ldots = H_G$), 
all the above mentioned approaches lead to identical parameter estimates.
However, if this is not the case, the results depend on the 
equation used \citep{schmidt90}. 
The only reason to use different instruments in the different equations 
is that the instruments of one equation are correlated with the
disturbance terms in another equation. Otherwise, one would simply use
every instrument in every equation \citep{schmidt90}. 
In this case, only the GMM-3SLS (equation (\ref{eq:beta3slsGMM}))
and the 3SLS estimator developed by \cite{schmidt90} 
(equation (\ref{eq:beta3slsSchmidt})) are consistent.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Imposing linear restrictions}
There are two ways to impose linear parameter restrictions.
First, a matrix $T$ can be provided with
\begin{equation}
   \beta = T \cdot \beta^* \label{eq:T-restr} 
\end{equation}
where $\beta^*$ is a vector of linear independent coefficients.

To impose these restrictions the $X$ matrix is
(post-)multiplied by this $T$ matrix.
\begin{equation}
    X^* = X \cdot T
\end{equation}

Then, a standard estimation is done by substituting $X^*$ for $X$.
This results in the linear independent parameter estimates $\beta^*$ and
the variance-covariance matrix of these parameters. 
The original parameters can be obtained by equation (\ref{eq:T-restr})
and the variance-covariance matrix of the original parameters 
can be obtained by:
\begin{equation}
   Var \left[ \widehat{\beta} \right] = T \cdot Var \left[ \widehat{\beta^*} \right] \cdot T'
\end{equation}

The second way to impose linear parameter restrictions 
can be formulated by
\begin{equation}
   R \beta^0 = q
\end{equation}
where $\beta^0$ is the vector of the restricted coefficients, 
$R$ is a matrix and $q$ is a vector (see \citealp[p. 100]{greene02}). 
Each restriction belongs to one row of $R$ and the corresponding 
element of~$q$.

The first way is less flexible than this latter one%
\footnote{e.g. restrictions like $\beta_1 + \beta_2 = 4$ cannot be imposed
by the first method}, 
but the first way is preferable if equality constraints for coefficients
across many equations of the system are imposed. 
Of course, these restrictions can be also imposed using
the latter method.
However, while the latter method increases the dimension of the 
matrices to be inverted during estimation, the first reduces it. 
Thus, in some cases the latter way leads to estimation problems
(e.g. (near) singularity of the matrices to be inverted),
while the first doesn't.

These two methods can be combined. In this case the restrictions
imposed using the latter method are imposed on the linear independent 
parameters due to the restrictions imposed using the first method:
\begin{equation}
   R \beta^{*0} = q
\end{equation}
where $\beta^{*0}$ is the vector of the restricted $\beta^*$ coefficients.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Restricted OLS estimation}

The OLS estimator restricted by $R \beta^0 = q$ can be obtained by
\begin{equation}
   \left[ \begin{array}{c}
      \widehat{\beta^0} \\ \widehat{\lambda}
   \end{array} \right]
   =
   \left[ \begin{array}{cc}
      X' X & R' \\ 
      R & 0
   \end{array} \right]^{-1}
   \cdot
   \left[ \begin{array}{c}
      X' y \\ q 
   \end{array} \right]
\end{equation}
where $2\lambda$ is a vector of the Lagrangean multipliers of the restrictions.

If the whole system is treated as one single equation, 
the variance-covariance matrix of the estimated parameters is
\begin{equation}
   Var 
   \left[ \begin{array}{c}
      \widehat{\beta} \\ \widehat{\lambda}
   \end{array} \right] 
   = \sigma^2 
   \left[ \begin{array}{cc}
      X' X & R' \\ 
      R & 0
   \end{array} \right]^{-1}
\end{equation}
with $\sigma^2 = E \left( u' u \right)$.

If the disturbance terms of the individual equations 
are allowed to have different variances, 
the variance-covariance matrix of the estimated parameters is
\begin{equation}
   Var 
   \left[ \begin{array}{c}
      \widehat{\beta} \\ \widehat{\lambda}
   \end{array} \right] 
   = 
   \left[ \begin{array}{cc}
      X' \Omega^{-1} X & R' \\ 
      R & 0
   \end{array} \right]^{-1}
\end{equation}
with $\Omega = \Sigma \otimes I$, 
$\sigma_{ij} = 0 \; \forall \; i \neq j$ and
$\sigma_{ii} = E \left( u_i' u_i \right)$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Restricted WLS estimation}

The WLS estimator restricted by $R \beta^0 = q$ can be obtained by
\begin{equation}
   \left[ \begin{array}{c}
      \widehat{\beta^0} \\ \widehat{\lambda}
   \end{array} \right]
   =
   \left[ \begin{array}{cc}
      X' \Omega^{-1} X & R' \\ 
      R & 0
   \end{array} \right]^{-1}
   \cdot
   \left[ \begin{array}{c}
      X' \Omega^{-1} y \\ q 
   \end{array} \right]
\end{equation}
with $\Omega = \Sigma \otimes I$, 
$\sigma_{ij} = 0 \; \forall \; i \neq j$ and
$\sigma_{ii} = E \left( u_i' u_i \right)$.

The variance-covariance matrix of the estimated parameters is
\begin{equation}
   Var 
   \left[ \begin{array}{c}
      \widehat{\beta} \\ \widehat{\lambda}
   \end{array} \right] 
   = 
   \left[ \begin{array}{cc}
      X' \Omega^{-1} X & R' \\ 
      R & 0
   \end{array} \right]^{-1}
\end{equation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Restricted SUR estimation}

The SUR estimator restricted by $R \beta^0 = q$ can be obtained by
\begin{equation}
   \left[ \begin{array}{c}
      \widehat{\beta^0} \\ \widehat{\lambda}
   \end{array} \right]
   =
   \left[ \begin{array}{cc}
      X' \Omega^{-1} X & R' \\ 
      R & 0
   \end{array} \right]^{-1}
   \cdot
   \left[ \begin{array}{c}
      X' \Omega^{-1} y \\ q 
   \end{array} \right]
\end{equation}
with $\Omega = \Sigma \otimes I$ and
$\sigma_{ij} = E \left( u_i' u_j \right)$.

The variance-covariance matrix of the estimated parameters is
\begin{equation}
   Var 
   \left[ \begin{array}{c}
      \widehat{\beta} \\ \widehat{\lambda}
   \end{array} \right] 
   = 
   \left[ \begin{array}{cc}
      X' \Omega^{-1} X & R' \\ 
      R & 0
   \end{array} \right]^{-1}
\end{equation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Restricted 2SLS estimation}

The 2SLS estimator restricted by $R \beta^0 = q$ can be obtained by
\begin{equation}
   \left[ \begin{array}{c}
      \widehat{\beta^0} \\ \widehat{\lambda}
   \end{array} \right]
   =
   \left[ \begin{array}{cc}
      \widehat{X}' \widehat{X} & R' \\ 
      R & 0
   \end{array} \right]^{-1}
   \cdot
   \left[ \begin{array}{c}
      \widehat{X}' y \\ q 
   \end{array} \right]
   \label{eq:beta2SLSr}
\end{equation}

If the whole system is treated as one single equation, 
the variance-covariance matrix of the estimated parameters is
\begin{equation}
   Var 
   \left[ \begin{array}{c}
      \widehat{\beta} \\ \widehat{\lambda}
   \end{array} \right] 
   = \sigma^2 
   \left[ \begin{array}{cc}
      \widehat{X}' \widehat{X} & R' \\ 
      R & 0
   \end{array} \right]^{-1}
\end{equation}
with $\sigma^2 = E \left( u' u \right)$.

If the disturbance terms of the individual equations 
are allowed to have different variances, 
the variance-covariance matrix of the estimated parameters is
\begin{equation}
   Var 
   \left[ \begin{array}{c}
      \widehat{\beta} \\ \widehat{\lambda}
   \end{array} \right] 
   = 
   \left[ \begin{array}{cc}
      \widehat{X}' \Omega^{-1} \widehat{X} & R' \\ 
      R & 0
   \end{array} \right]^{-1}
\end{equation}
with $\Omega = \Sigma \otimes I$, 
$\sigma_{ij} = 0 \; \forall \; i \neq j$ and
$\sigma_{ii} = E \left( u_i' u_i \right)$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Restricted W2SLS estimation}

The W2SLS estimator restricted by $R \beta^0 = q$ can be obtained by
\begin{equation}
   \left[ \begin{array}{c}
      \widehat{\beta^0} \\ \widehat{\lambda}
   \end{array} \right]
   =
   \left[ \begin{array}{cc}
      \widehat{X}' \Omega^{-1} \widehat{X} & R' \\ 
      R & 0
   \end{array} \right]^{-1}
   \cdot
   \left[ \begin{array}{c}
      \widehat{X}' \Omega^{-1} y \\ q 
   \end{array} \right]
\end{equation}
with $\Omega = \Sigma \otimes I$, 
$\sigma_{ij} = 0 \; \forall \; i \neq j$ and
$\sigma_{ii} = E \left( u_i' u_i \right)$.

The variance-covariance matrix of the estimated parameters is
\begin{equation}
   Var 
   \left[ \begin{array}{c}
      \widehat{\beta} \\ \widehat{\lambda}
   \end{array} \right] 
   = 
   \left[ \begin{array}{cc}
      \widehat{X}' \Omega^{-1} \widehat{X} & R' \\ 
      R & 0
   \end{array} \right]^{-1}
\end{equation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Restricted 3SLS estimation}

The standard 3SLS estimator restricted by $R \beta^0 = q$ can be obtained by
\begin{equation}
   \left[ \begin{array}{c}
      \widehat{\beta^0} \\ \widehat{\lambda}
   \end{array} \right]
   =
   \left[ \begin{array}{cc}
      \widehat{X}' \Omega^{-1} \widehat{X} & R' \\ 
      R & 0
   \end{array} \right]^{-1}
   \cdot
   \left[ \begin{array}{c}
      \widehat{X}' \Omega^{-1} y \\ q 
   \end{array} \right]
\end{equation}
with $\Omega = \Sigma \otimes I$ and
$\sigma_{ij} = E \left( u_i' u_j \right)$.

The variance-covariance matrix of this estimator is
\begin{equation}
   Var 
   \left[ \begin{array}{c}
      \widehat{\beta^0} \\ \widehat{\lambda}
   \end{array} \right] 
   = 
   \left[ \begin{array}{cc}
      \widehat{X}' \Omega^{-1} \widehat{X} & R' \\ 
      R & 0
   \end{array} \right]^{-1}
   \label{eq:cov3slsr}
\end{equation}

The IV-3SLS estimator restricted by $R \beta^0 = q$ can be obtained by
\begin{equation}
   \left[ \begin{array}{c}
      \widehat{\beta^0} \\ \widehat{\lambda}
   \end{array} \right]
   =
   \left[ \begin{array}{cc}
      \widehat{X}' \Omega^{-1} X & R' \\ 
      R & 0
   \end{array} \right]^{-1}
   \cdot
   \left[ \begin{array}{c}
      \widehat{X}' \Omega^{-1} y \\ q 
   \end{array} \right]
\end{equation}

with
\begin{equation}
   Var 
   \left[ \begin{array}{c}
      \widehat{\beta^0} \\ \widehat{\lambda}
   \end{array} \right] 
   = 
   \left[ \begin{array}{cc}
      \widehat{X}' \Omega^{-1} \widehat{X} & R' \\ 
      R & 0
   \end{array} \right]^{-1}
\end{equation}

The restricted 3SLS-GMM estimator can be obtained by
\begin{equation}
   \left[ \begin{array}{c}
      \widehat{\beta^0} \\ \widehat{\lambda}
   \end{array} \right]
   =
   \left[ \begin{array}{cc}
      X' H \left( H' \Omega H \right)^{-1} H' X & R' \\ 
      R & 0
   \end{array} \right]^{-1}
   \cdot
   \left[ \begin{array}{c}
      X' H \left( H \Omega H \right)^{-1} H' y \\ q 
   \end{array} \right]
\end{equation}

with
\begin{equation}
   Var 
   \left[ \begin{array}{c}
      \widehat{\beta^0} \\ \widehat{\lambda}
   \end{array} \right] 
   = 
   \left[ \begin{array}{cc}
      X' H \left( H' \Omega H \right)^{-1} H' X & R' \\ 
      R & 0
   \end{array} \right]^{-1}
\end{equation}

The restricted 3SLS estimator based on the suggestion of 
\cite{schmidt90} is:
\begin{equation}
   \left[ \begin{array}{c}
      \widehat{\beta^0} \\ \widehat{\lambda}
   \end{array} \right]
   =
   \left[ \begin{array}{cc}
      \widehat{X}' \Omega^{-1} \widehat{X} & R' \\ 
      R & 0
   \end{array} \right]^{-1}
   \cdot
   \left[ \begin{array}{c}
      \widehat{X}' \Omega^{-1} H \left( H' H \right)^{-1} H' y \\ q 
   \end{array} \right]
\end{equation}

with
\begin{eqnarray}
   Var 
   \left[ \begin{array}{c}
      \widehat{\beta^0} \\ \widehat{\lambda}
   \end{array} \right] 
   & = & 
   \left[ \begin{array}{cc}
      \widehat{X}' \Omega^{-1} \widehat{X} & R' \\ 
      R & 0
   \end{array} \right]^{-1}
   \cdot
   \left[ \begin{array}{cc}
      \widehat{X}' \Omega^{-1} H \left( H' H \right)^{-1} H' \Omega
      H \left( H' H \right)^{-1} H' \Omega^{-1} \widehat{X} & 0' \\ 
      0 & 0
   \end{array} \right]^{-1}
   \nonumber \\
   & & \cdot
   \left[ \begin{array}{cc}
      \widehat{X}' \Omega^{-1} \widehat{X} & R' \\ 
      R & 0
   \end{array} \right]^{-1}
\end{eqnarray}

The econometrics software EViews calculates the restricted 3SLS estimator by:
\begin{equation}
   \left[ \begin{array}{c}
      \widehat{\beta^0} \\ \widehat{\lambda}
   \end{array} \right]
   =
   \left[ \begin{array}{cc}
      \widehat{X}' \Omega^{-1} \widehat{X} & R' \\ 
      R & 0
   \end{array} \right]^{-1}
   \cdot
   \left[ \begin{array}{c}
      \widehat{X}' \Omega^{-1} \left( y - X \widehat{\beta^0}_{2SLS} \right) 
      \\ q 
   \end{array} \right]
\end{equation}
where $\widehat{\beta^0}_{2SLS}$ is the restricted 2SLS estimator calculated
by equation (\ref{eq:beta2SLSr}). 
To calculate the variance-covariance matrix 
EViews uses the standard formula of the restricted 3SLS estimator
(equation (\ref{eq:cov3slsr})).

If the same instrumental variables are used in all equations 
($H_1 = H_2 = \ldots = H_G$), 
all the above mentioned approaches lead to identical parameter estimates
and identical variance-covariance matrices of the estimated parameters.

(***** This has to be checked!!!!! **************)

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Other issues}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Residual variance-covariance matrix}

Since the true residuals of the estimated equations are generally not known,
the true covariance matrix of the residuals cannot be determined.
Thus, the this covariance matrix must be calculated from the
\emph{estimated} residuals.

The estimated covariance matrix of the residuals ($\widehat{\Sigma} = 
\left[ \sigma \right]$) 
is often calculated by
\begin{equation}
   \widehat{\sigma}_{ij} = \frac{ \widehat{u}_i' \widehat{u}_j }{ T }
   \label{eq:rcov0}
\end{equation}
where $T$ is the number of observations in each equation.

However, this estimator is biased, because it is not corrected for the
degrees of freedom. 
The usual single-equation procedure to correct for degrees of freedom 
cannot always be applied, because the number of regressors in each equation 
might differ.
Two alternative approaches to calculate the residual variance-covariance 
matrix are
\begin{equation}
   \widehat{\sigma}_{ij} = \frac{ \widehat{u}_i' \widehat{u}_j }
   { \sqrt{ \left( T - K_i \right) \cdot \left( T - K_j \right) } }
   \label{eq:rcov1}
\end{equation}
and
\begin{equation}
   \widehat{\sigma}_{ij} = \frac{ \widehat{u}_i' \widehat{u}_j }
   { T - \max \left( K_i , K_j \right) }
\end{equation}
where $K_i$ and $K_j$ are the number of regressors in equation
$i$ and $j$, respectively.

However, this yields an unbiased estimator only if $K_i = K_j$ 
\citep[p. 469]{judge85}. 
% Greene (2002, p. 344) says that the second is unbiased if i = j or K_i = K_j,
% whereas the first is unbiased only if i = j. 
% However, if K_i = K_j the first and the second are equal.
% Why is the first biased if K_i = K_j ???????????

A further approach to obtain the estimated residual variance-covariance
matrix is
\begin{equation}
   \widehat{\sigma}_{ij} = \frac{ \widehat{u}_i' \widehat{u}_j } 
   { T - K_i - K_j + tr \left[ \left( X_i' X_i \right)^{-1}
   X_i' X_j \left( X_j' X_j \right)^{-1} X_j' X_i \right] }
   \label{eq:rcov2}
\end{equation} 
This yields an unbiased estimator for all elements of $\widehat{\Sigma}$,
but even if $\widehat{\Sigma}$ is an unbiased estimator of $\Sigma$, 
its inverse $\widehat{\Sigma}^{-1}$ is not an unbiased estimator 
of $\Sigma^{-1}$ \citep[p.322]{theil71}.
Furthermore, the variance-covariance matrix calculated by (\ref{eq:rcov2})
is not necessarily positive semidefinite \citep[p.322]{theil71}. 
Hence, "it is doubtful whether [this formula] is really superior to 
[equation (\ref{eq:rcov0})]" \citep[p.322]{theil71}.

The WLS, SUR, W2SLS and 3SLS parameter estimates are consistent,
if the estimated residual variance-covariance matrix is calculated
using the residuals from a first step OLS or 2SLS estimation.
It is also possible to iterate the estimation procedure and 
calculate the residual variance-covariance matrix from the
residuals of the previous iteration. 
If equation (\ref{eq:rcov0}) is applied to calculate the estimated
residual variance-covariance matrix, the results are maximum
likelihood \citep[p. 345]{greene02}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Degrees of freedom}

There exist different approaches to determine the degrees of
freedom for testing the estimated parameters with a t-test.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Goodness of fit}

The goodness of fit of each equation can be measured by the 
traditional $R^2$ values:
\begin{equation}
   R_i^2 = 1 - \frac{ \widehat{u}_i' \widehat{u}_i }
   { ( y_i - \overline{y_i} )' ( y_i - \overline{y_i} ) }
\end{equation}
where $R_i^2$ is the $R^2$ value of the $i$th equation
and $\overline{y_i}$ is the mean value of $y_i$.

The goodness of fit of the whole system can be measured by the
McElroy's $R^2$ value \citep{mcelroy77}: 
% also: \citep[p. 345]{greene02}
\begin{equation}
   R_*^2 = 1 - \frac{ \widehat{u}' \widehat{ \Omega }^{-1} \widehat{u} }
   { y' \left( \widehat{ \Sigma }^{-1} \otimes
   \left( I - \frac{i i'}{T} \right) \right) y }
\end{equation}
where $T$ is the number of observations in each equation,
$I$ is an $T \times T$ identity matrix and 
$i$ is a column vector of $T$ ones.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Using systemfit}\label{sec:UsingSystemfit}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Standard usage}

\pkg{systemfit} is generally called by

\code{
R> systemfit( method, eqns )
}

There are two mandatory arguments: \code{method} and \code{eqns}.

The argument \code{method} is a string determining the estimation method.
It must be one of "OLS", "WLS", "SUR", "2SLS", "W2SLS" or "3SLS".

The other mandatory argument \code{eqns} is a list of the equations 
to estimate. 
Each equation is a standard formula in \proglang{R}.
It starts with a dependent variable on the left hand side.
After a tilde ($~$) the regressors are listed%
\footnote{For Details see the \proglang{R} help files to \code{formula}}.

This is now demonstrated using an example: \\
\code{
R> library( systemfit ) \\
R> data( kmenta ) \\
R> attach( kmenta ) \\
R> fitsur <- systemfit( "SUR", list( q $\sim$ p + d, q $\sim$ p + f + a ) ) \\
}

The first line loads the \pkg{systemfit} package. 
The second line loads example data that are included in this package.
These data come from \cite{kmenta86}.
They are attached to the \proglang{R} search path in line three.
In the last line a seemingly unrelated regression is done.
The first equation represents the demand side of the food market.
The dependant variable is \code{q} (food consumption per capita). 
The regressors are \code{p} (ratio of food prices to general consumer prices)
and \code{disposable income} as well as a constant%
\footnote{a regression constant is always implied if not explicitly omitted.}.
The second equation represents the supply side.
Variable \code{q} (food consumption per capita) is also the dependant 
variable of this equation. 
The regressors are again \code{p} (ratio of food prices to general 
consumer prices) and a constant as well as 
\code{f} (ratio of preceding year's prices received by farmers) and 
\code{a} (a time trend in years).
The regression result is assigned to the variable \code{fitsur}.

Summary results can be printed by\\
\code{
R>~summary(~fitsur~)~\\
~\\
systemfit~results~\\
method:~SUR~\\
\\
\textcolor{white}{.}~~N~DF~~~~~~SSR~~~~~MSE~~~~RMSE~~~~~~~R2~~~Adj~R2 \\
1~20~17~~65.6829~3.86370~1.96563~0.755019~0.726198~\\
2~20~16~104.0584~6.50365~2.55023~0.611888~0.539117~\\
~\\
The~covariance~matrix~of~the~residuals~used~for~estimation\\
\textcolor{white}{.}~~~~~~~1~~~~~~~2~\\
1~3.72539~4.13696~\\
2~4.13696~5.78444~\\
~\\
The~covariance~matrix~of~the~residuals\\
\textcolor{white}{.}~~~~~~~1~~~~~~~2~\\
1~3.86370~4.92431~\\
2~4.92431~6.50365~\\
~\\
The~correlations~of~the~residuals\\
\textcolor{white}{.}~~~~~~~~1~~~~~~~~2~\\
1~1.000000~0.982348~\\
2~0.982348~1.000000~\\
~\\
The~determinant~of~the~residual~covariance~matrix:~0.879285~\\
OLS~R-squared~value~of~the~system:~0.683453~\\
McElroy's~R-squared~value~for~the~system:~0.788722~\\
~\\
SUR~estimates~for~1~~(equation~1~)~\\
Model~Formula:~q~~~p~+~d\\
\\
\textcolor{white}{.}~~~~~~~~~~~~Estimate~Std.~Error~~~t~value~Pr(>|t|)~\\
(Intercept)~99.332894~~~7.514452~13.218913~~~~~~~~0~***~\\
p~~~~~~~~~~~-0.275486~~~0.088509~-3.112513~0.006332~~**~\\
d~~~~~~~~~~~~~0.29855~~~0.041945~~7.117605~~~~2e-06~***~\\
---~\\
Signif.~codes:~~0~`***'~0.001~`**'~0.01~`*'~0.05~`.'~0.1~`~'~1~\\
~\\
Residual~standard~error:~1.96563~on~17~degrees~of~freedom~\\
Number~of~observations:~20~Degrees~of~Freedom:~17~\\
SSR:~65.682902~MSE:~3.8637~Root~MSE:~1.96563~\\
Multiple~R-Squared:~0.755019~Adjusted~R-Squared:~0.726198~\\
~\\
~\\
SUR~estimates~for~2~~(equation~2~)~\\
Model~Formula:~q~~~p~+~f~+~a\\
\\
\textcolor{white}{.}~~~~~~~~~~~~Estimate~Std.~Error~~t~value~Pr(>|t|)~\\
(Intercept)~61.966166~~~11.08079~5.592215~~~~4e-05~***~\\
p~~~~~~~~~~~~0.146884~~~0.094435~1.555397~0.139408~\\
f~~~~~~~~~~~~0.214004~~~0.039868~5.367761~~6.3e-05~***~\\
a~~~~~~~~~~~~0.339304~~~0.067911~4.996283~0.000132~***~\\
---~\\
Signif.~codes:~~0~`***'~0.001~`**'~0.01~`*'~0.05~`.'~0.1~`~'~1~\\
~\\
Residual~standard~error:~2.550226~on~16~degrees~of~freedom~\\
Number~of~observations:~20~Degrees~of~Freedom:~16~\\
SSR:~104.05843~MSE:~6.503652~Root~MSE:~2.550226~\\
Multiple~R-Squared:~0.611888~Adjusted~R-Squared:~0.539117~\\
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{User options}

Following additional options can be set by the user:

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Equation labels}
The optional argument \code{eqnlabels} allows the user to label the equations.
It has to be a vector of strings naming the equations.\\
\code{
R>~fitsur~<-~systemfit(~"SUR",~list(~q~~~p~+~d,~q~~~p~+~f~+~a~),\\~
R+~~~~eqnlabels~=~c(~"demand",~"supply"~)~)\\
R>~summary(~fitsur~)\\
systemfit~results\\
method:~SUR\\
\\
\textcolor{white}{.}~~~~~~~N~DF~~~~~~SSR~~~~~MSE~~~~RMSE~~~~~~~R2~~~Adj~R2\\
demand~20~17~~65.6829~3.86370~1.96563~0.755019~0.726198\\
supply~20~16~104.0584~6.50365~2.55023~0.611888~0.539117\\
\ldots\\
}
If no equation labels are provided, the equations are numbered.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Instrumental variables}   
\code{inst}one-sided model formula specifying instrumental variables
   or a list of one-sided model formulas if different instruments should
   be used for the different equations (only needed for 2SLS, W2SLS and
   3SLS estimations).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Data}   
\code{data} an optional data frame containing the variables in the model.
   By default the variables are taken from the environment from which
   systemfit is called.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Restrictions}   
\code{R.restr} an optional j x k matrix to impose linear
   restrictions on the parameters by \code{R.restr} * $\beta$ = \code{q.restr}
   (j = number of restrictions, k = number of all parameters,
   $\beta$ = vector of all parameters).

\code{q.restr} an optional j x 1 matrix to impose linear
   restrictions (see \code{R.restr}); default is a j x 1 matrix
   that contains only zeros.

\code{TX} an optional matrix to transform the regressor matrix and,
   hence, also the coefficient vector (see details).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Iteration control}   
\code{maxiter} maximum number of iterations for WLS, SUR, W2SLS and
   3SLS estimations.

\code{tol} tolerance level indicating when to stop the iteration (only
   WLS, SUR, W2SLS and 3SLS estimations).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Residual covariance matrix}   
\code{rcovformula} formula to calculate the estimated residual covariance
   matrix (see details).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{3SLS formula}   
\code{formula3sls} formula for calculating the 3SLS estimator,
   one of "GLS", "IV", "GMM", "Schmidt" or "EViews" (see details).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Degrees of freedom for t-tests}   
\code{probdfsys} use the degrees of freedom of the whole system
   (in place of the degrees of freedom of the single equation)
   to calculate prob values for the t-test of individual parameters.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Sigma squared}   
\code{single.eq.sigma} use different $\sigma^2$s for each
   single equation to calculate the covariance matrix and the
   standard errors of the coefficients (only OLS and 2SLS).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{System options}   
\code{solvetol} tolerance level for detecting linear dependencies
   when inverting a matrix or calculating a determinant (see
   \code{solve} and \code{det}).

\code{saveMemory} save memory by omitting some calculation that
   are not crucial for the basic estimation (e.g McElroy's
   $R^2$).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Other tools}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Likelihood ratio test}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Hausman test}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Summary and outlook}\label{sec:Summmary}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

not implemented yet: unequal numbers of observations

nlsystemfit finished?

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliography{/home/suapm095/Documents/Literatur/arne}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}

### 2004-11-09 paper/trunk/systemfit.tex
changed bibtex file to "references[.bib]"
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@561 591db2bd-e0a0-4d22-8c19-2192dbb7da31
changed ugly hack with "\textcolor{white}" in \code to an empty mbox
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@562 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 \\
-\textcolor{white}{.}~~N~DF~~~~~~SSR~~~~~MSE~~~~RMSE~~~~~~~R2~~~Adj~R2 \\
+\mbox{}~~~N~DF~~~~~~SSR~~~~~MSE~~~~RMSE~~~~~~~R2~~~Adj~R2 \\
 1~20~17~~65.6829~3.86370~1.96563~0.755019~0.726198~\\
@@
 The~covariance~matrix~of~the~residuals~used~for~estimation\\
-\textcolor{white}{.}~~~~~~~1~~~~~~~2~\\
+\mbox{}~~~~~~~~1~~~~~~~2~\\
 1~3.72539~4.13696~\\
@@
 The~covariance~matrix~of~the~residuals\\
-\textcolor{white}{.}~~~~~~~1~~~~~~~2~\\
+\mbox{}~~~~~~~~1~~~~~~~2~\\
 1~3.86370~4.92431~\\
@@
 The~correlations~of~the~residuals\\
-\textcolor{white}{.}~~~~~~~~1~~~~~~~~2~\\
+\mbox{}~~~~~~~~~1~~~~~~~~2~\\
 1~1.000000~0.982348~\\
@@
 \\
-\textcolor{white}{.}~~~~~~~~~~~~Estimate~Std.~Error~~~t~value~Pr(>|t|)~\\
+\mbox{}~~~~~~~~~~~~~Estimate~Std.~Error~~~t~value~Pr(>|t|)~\\
 (Intercept)~99.332894~~~7.514452~13.218913~~~~~~~~0~***~\\
@@
 \\
-\textcolor{white}{.}~~~~~~~~~~~~Estimate~Std.~Error~~t~value~Pr(>|t|)~\\
+\mbox{}~~~~~~~~~~~~~Estimate~Std.~Error~~t~value~Pr(>|t|)~\\
 (Intercept)~61.966166~~~11.08079~5.592215~~~~4e-05~***~\\
@@
 \\
-\textcolor{white}{.}~~~~~~~N~DF~~~~~~SSR~~~~~MSE~~~~RMSE~~~~~~~R2~~~Adj~R2\\
+\mbox{}~~~~~~~~N~DF~~~~~~SSR~~~~~MSE~~~~RMSE~~~~~~~R2~~~Adj~R2\\
 demand~20~17~~65.6829~3.86370~1.96563~0.755019~0.726198\\
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\bibliography{/home/suapm095/Documents/Literatur/arne}
+\bibliography{references} % a subset of my big bibtex file
+%\bibliography{/home/suapm095/Documents/Literatur/arne}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2004-11-16 paper/trunk/systemfit.tex
updated the systemfit.tex file to include the nlsystemfit sections. updated the bibtex file to match the filename and removed the systemfit.pdf from the repository
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@563 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 \author{Arne Henningsen\\University of Kiel\And 
-        Jeff D. Hamann\\ ??? }
+        Jeff D. Hamann\\ Forest Informatics, Inc. }
 
@@
 %% an abstract and keywords
-\Abstract{
-   Many statistical analysis are based on models that consist of 
-   more than one equation.
-   These equations should be fitted simultaneously rather than equation 
-   by equation.
-   The package \pkg{systemfit} provides the capability to estimate systems 
-   of linear (and non-linear???) equations in \proglang{R}. 
-}
-\Keywords{R, equation system, seemingly unrelated regression}
+\Abstract{ Many statistical analysis are based on models that consist
+  of  more than one equation.  These equations should be fitted
+  simultaneously rather than equation  by equation.  The package
+  \pkg{systemfit} provides the capability to estimate systems  of
+  linear and non-linear equations in \proglang{R}.  } \Keywords{R,
+  simultaneous equation system, seemingly unrelated regression}
+
 %% at least one keyword must be supplied
@@
    Jeff D. Hamann\\
+   Forest Informatics, Inc.\\
+   PO Box 1421\\
+   Corvallis, Oregon 97339-1421\\
    e-mail: \email{jeff.hamann@forestinformatics.com}\\
@@
 
-Many theoretical models consist of more than one equation.
-If these models are econometrically estimated, 
-the equations generally should not be estimated individually, 
-but all together, because the disturbance terms of these 
-equations are likely contemporaneously correlated.
-Estimating all equations simultaneously and
-taking the covariance structure of the residuals into account 
-leads to efficient estimates, 
-while ignoring this correlation leads to an inefficient 
+Many theoretical models consist of more than one equation.  If these
+models are econometrically estimated,  the equations generally should
+not be estimated individually,  but all together, because the
+disturbance terms of these  equations are likely contemporaneously
+correlated.  Estimating all equations simultaneously and taking the
+covariance structure of the residuals into account  leads to efficient
+estimates,  while ignoring this correlation leads to an inefficient
 parameter estimation \citep{zellner62}.
@@
 Another reason to estimate an equation system simultaneously are
-cross-equation parameter restrictions. These restrictions can 
-be tested and/or imposed only in a simultaneous estimation
-approach. Especially the economic theory suggests many 
-cross-equation restriction.
+cross-equation parameter restrictions. These restrictions can  be
+tested and/or imposed only in a simultaneous estimation
+approach. Especially the economic theory suggests many  cross-equation
+restriction.
 
 The \pkg{systemfit} package provides the capability to estimate
-linear (and non-linear???) equation systems in \proglang{R}.
-Compared to other statistics and econometrics software 
-the (advanced) user can control many important details of the 
-estimation procedure. On the other hand these user options have
-reasonable defaults that also beginners can easily use the
-provided functions.
-
-This paper is organized as follows: 
-In section \ref{sec:Theory} we introduce the mathematics of estimating
-equation systems and give a comprehensive overview over all
-relevant issues regarding this topic.
-Section \ref{sec:UsingSystemfit} demonstrates how to run \pkg{systemfit},
-especially how the features presented in the previous chapter
-can be used.
-Finally, a summary and an outlook are given in section \ref{sec:Summmary}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Estimating equation systems}\label{sec:Theory}
+linear and non-linear equation systems in \proglang{R}.  Compared to
+other statistics and econometrics software  the (advanced) user can
+control many important details of the  estimation procedure. On the
+other hand these user options have reasonable defaults that also
+beginners can easily use the provided functions.
+
+\textbf{I'll have to update this section to match the recent changes. JDH.}
+
+This paper is organized as follows:  In section \ref{sec:Theory} we
+introduce the mathematics of estimating equation systems and give a
+comprehensive overview over all relevant issues regarding this topic.
+Section \ref{sec:UsingSystemfit} demonstrates how to run
+\pkg{systemfit}, especially how the features presented in the previous
+chapter can be used.  Finally, a summary and an outlook are given in
+section \ref{sec:Summmary}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Estimating linear equation systems}\label{sec:Theory}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Parameter estimation}
+\subsection{Linear Parameter estimation}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Using systemfit}\label{sec:UsingSystemfit}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Standard usage}
+\subsection{Using systemfit}\label{sec:UsingSystemfit}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Standard usage}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{User options}
+\subsubsection{User options}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Equation labels}
+% \subsubsection{Equation labels}
+\paragraph{Equation labels}
 The optional argument \code{eqnlabels} allows the user to label the equations.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Instrumental variables}   
+%\subsubsection{Instrumental variables}   
+\paragraph{Instrumental variables}   
 \code{inst}one-sided model formula specifying instrumental variables
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Data}   
+%\subsubsection{Data}   
+\paragraph{Data}   
 \code{data} an optional data frame containing the variables in the model.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Restrictions}   
+%\subsubsection{Restrictions}   
+\paragraph{Restrictions}   
 \code{R.restr} an optional j x k matrix to impose linear
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Iteration control}   
+%\subsubsection{Iteration control}   
+\paragraph{Iteration control}
 \code{maxiter} maximum number of iterations for WLS, SUR, W2SLS and
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Residual covariance matrix}   
+%\subsubsection{Residual covariance matrix}   
+\paragraph{Residual covariance matrix}   
 \code{rcovformula} formula to calculate the estimated residual covariance
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{3SLS formula}   
+%\subsubsection{3SLS formula}   
+\paragraph{3SLS formula}   
 \code{formula3sls} formula for calculating the 3SLS estimator,
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Degrees of freedom for t-tests}   
+%\subsubsection{Degrees of freedom for t-tests}   
+\paragraph{Degrees of freedom for t-tests}   
 \code{probdfsys} use the degrees of freedom of the whole system
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Sigma squared}   
+%\subsubsection{Sigma squared}   
+\paragraph{Sigma squared}   
 \code{single.eq.sigma} use different $\sigma^2$s for each
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{System options}   
+%\subsubsection{System options}   
+\paragraph{System options}
+
 \code{solvetol} tolerance level for detecting linear dependencies
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Other tools}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Likelihood ratio test}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Hausman test}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Estimating nonlinear equation systems}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+The \pkg{systemfit} package also contains a method for fitting
+nonlinear systems of equations.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Nonlinear parameter estimation}
+\label{sec:nonlinear_estimation}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+A system of nonlinear equations can be written as:
+
+\begin{equation}
+  \label{eq:non_linear_eq_1}
+  \epsilon_{t} = q( y_t, x_t, \theta )
+\end{equation}
+
+\noindent and
+
+\begin{equation}
+  \label{eq:non_linear_eq_2}
+  z_{t} = Z( x_t )
+\end{equation}
+
+where $\epsilon_{t}$ are the residuals from the y observations and
+$z_{t}$ are the function evaluated at the parameter estimates.
+
+\textbf{you really need to check this!!!}
+
+The objective functions for the methods are:
+
+\begin{center}
+\begin{tabular}{|l|c|c|c|} \hline
+  Method & Instruments & Objective Function & Covariance of $\theta$ \\ \hline  
+  OLS & No & $r'r$ & $(X(diag(S)^{-1}\otimes I)X)^{-1}$ \\ \hline
+  SUR & No & $r'(diag(S)_{OLS}^{-1}\otimes I)r$ & $(X(S^{-1}\otimes I)X)^{-1}$ \\ \hline
+  2SLS & Yes & $r'(I \otimes W)r$ & $(X(diag(S)^{-1}\otimes I)X)^{-1}$ \\ \hline 
+  3SLS & Yes & $r'(S_{2SLS}^{-1} \otimes W)r$ & $(X(diag(S)^{-1}\otimes W)X)^{-1}$ \\ \hline
+\end{tabular}
+\end{center}
+
+where, $r$ is a column vector for the residuals for each equation, $X$
+is matrix of the partial derivatives with respect to the parameters,
+$W$ is a matrix of the instrument variables, $Z(Z'Z)^{-1}Z$, $Z$ is a
+matrix of the instrument variables, and $I$ is an $n \times n $
+identity matrix and $S$ is the variance-covariance matrix between the
+equations
+
+\begin{equation}
+  \label{eq:non-linear_varcov}
+  \hat{\sigma}_{ij} = (\hat{e}_i' \hat{e}_j) / \sqrt{(T - k_i)*(T- k_j)} 
+\end{equation}
+
+\textbf{You need to clear this up.}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Using nlsystemfit}\label{sec:UsingnlSystemfit}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\code {nlsystemfit} fits a set of structural nonlinear equations using
+Ordinary Least Squares (OLS), Seemingly Unrelated Regression (SUR),
+Two-Stage Least Squares (2SLS), Three-Stage Least Squares (3SLS) using
+the objective functions described in section
+\ref{sec:nonlinear_estimation}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Standard usage}
+
+Similar to calling the \code{systemfit} function, \code{nlsystemfit}
+is called with a minimum of three arguments,
+
+\code{
+R> nlsystemfit( method, eqns, start )
+}
+
+where \code{method} is one of the following estimation methods: "OLS",
+"SUR", "2SLS", or "3SLS", \code{eqns} is a list of equations similar
+to those described for \code{systemfit}, and \code{start} is a list of
+starting values of the parameter estimates.
+
+% The mandatory argument \code{eqns} is a list of the equations  to
+% estimate.  Each equation is a standard formula in \proglang{R}.  It
+% starts with a dependent variable on the left hand side.  After a tilde
+% ($~$) the regressors are listed  \footnote{For Details see the
+%   \proglang{R} help files to \code{formula}}.
+
+This is now demonstrated using an example: \\
+
+\code{
+R> library( systemfit ) \\
+R> data( ppine ) \\
+R> hg.formula <- hg $\sim$ exp( h0 + h1*log(tht) + h2*tht$\hat$2 + h3*elev + h4*cr) \\
+R> dg.formula <- dg $\sim$ exp( d0 + d1*log(dbh) + d2*hg + d3*cr + d4*ba  ) \\
+R> labels <- list( "height.growth", "diameter.growth" )\\
+R> inst <- $\sim$ tht + dbh + elev + cr + ba\\
+R> start.values <- c(h0=-0.5, h1=0.5, h2=-0.001, h3=0.0001, h4=0.08,\\
+R+   d0=-0.5, d1=0.009, d2=0.25, d3=0.005, d4=-0.02 )\\
+R> model <- list( hg.formula, dg.formula )\\
+R> model.3sls <- nlsystemfit( "3SLS", model, start.values, data=ppine,\\
+R+   eqnlabels=labels, inst=inst ) \\
+R> summary( model.3sls ) \\
+}
+
+The nlsystemfit function relies on \code{nlm} to perform the
+minimization of the objective functions and the \code{qr} set of
+functions. If the user does not have a set of estimates for the
+initial parameters, it is suggested using linearized forms in one of
+the linear methods, or simply using \code{nls} to obtain estimates for
+the equations.
+
+The outputs are similar to those for a systemfit object.
+
+\textbf{Describe the code and outputs as Arne did in the previous
+  section.}
+
+The first line loads the \pkg{systemfit} package. 
+The second line loads example data that are included in this package.
+These data come from \cite{kmenta86}.
+They are attached to the \proglang{R} search path in line three.
+In the last line a seemingly unrelated regression is done.
+The first equation represents the demand side of the food market.
+The dependant variable is \code{q} (food consumption per capita). 
+The regressors are \code{p} (ratio of food prices to general consumer prices)
+and \code{disposable income} as well as a constant%
+\footnote{a regression constant is always implied if not explicitly omitted.}.
+The second equation represents the supply side.
+Variable \code{q} (food consumption per capita) is also the dependant 
+variable of this equation. 
+The regressors are again \code{p} (ratio of food prices to general 
+consumer prices) and a constant as well as 
+\code{f} (ratio of preceding year's prices received by farmers) and 
+\code{a} (a time trend in years).
+The regression result is assigned to the variable \code{fitsur}.
+
+Summary results can be printed by\\
+\code{
+R>~summary(~fitsur~)~\\
+~\\
+systemfit~results~\\
+method:~SUR~\\
+\\
+\mbox{}~~~N~DF~~~~~~SSR~~~~~MSE~~~~RMSE~~~~~~~R2~~~Adj~R2 \\
+1~20~17~~65.6829~3.86370~1.96563~0.755019~0.726198~\\
+2~20~16~104.0584~6.50365~2.55023~0.611888~0.539117~\\
+~\\
+The~covariance~matrix~of~the~residuals~used~for~estimation\\
+\mbox{}~~~~~~~~1~~~~~~~2~\\
+1~3.72539~4.13696~\\
+2~4.13696~5.78444~\\
+~\\
+The~covariance~matrix~of~the~residuals\\
+\mbox{}~~~~~~~~1~~~~~~~2~\\
+1~3.86370~4.92431~\\
+2~4.92431~6.50365~\\
+~\\
+The~correlations~of~the~residuals\\
+\mbox{}~~~~~~~~~1~~~~~~~~2~\\
+1~1.000000~0.982348~\\
+2~0.982348~1.000000~\\
+~\\
+The~determinant~of~the~residual~covariance~matrix:~0.879285~\\
+OLS~R-squared~value~of~the~system:~0.683453~\\
+McElroy's~R-squared~value~for~the~system:~0.788722~\\
+~\\
+SUR~estimates~for~1~~(equation~1~)~\\
+Model~Formula:~q~~~p~+~d\\
+\\
+\mbox{}~~~~~~~~~~~~~Estimate~Std.~Error~~~t~value~Pr(>|t|)~\\
+(Intercept)~99.332894~~~7.514452~13.218913~~~~~~~~0~***~\\
+p~~~~~~~~~~~-0.275486~~~0.088509~-3.112513~0.006332~~**~\\
+d~~~~~~~~~~~~~0.29855~~~0.041945~~7.117605~~~~2e-06~***~\\
+---~\\
+Signif.~codes:~~0~`***'~0.001~`**'~0.01~`*'~0.05~`.'~0.1~`~'~1~\\
+~\\
+Residual~standard~error:~1.96563~on~17~degrees~of~freedom~\\
+Number~of~observations:~20~Degrees~of~Freedom:~17~\\
+SSR:~65.682902~MSE:~3.8637~Root~MSE:~1.96563~\\
+Multiple~R-Squared:~0.755019~Adjusted~R-Squared:~0.726198~\\
+~\\
+~\\
+SUR~estimates~for~2~~(equation~2~)~\\
+Model~Formula:~q~~~p~+~f~+~a\\
+\\
+\mbox{}~~~~~~~~~~~~~Estimate~Std.~Error~~t~value~Pr(>|t|)~\\
+(Intercept)~61.966166~~~11.08079~5.592215~~~~4e-05~***~\\
+p~~~~~~~~~~~~0.146884~~~0.094435~1.555397~0.139408~\\
+f~~~~~~~~~~~~0.214004~~~0.039868~5.367761~~6.3e-05~***~\\
+a~~~~~~~~~~~~0.339304~~~0.067911~4.996283~0.000132~***~\\
+---~\\
+Signif.~codes:~~0~`***'~0.001~`**'~0.01~`*'~0.05~`.'~0.1~`~'~1~\\
+~\\
+Residual~standard~error:~2.550226~on~16~degrees~of~freedom~\\
+Number~of~observations:~20~Degrees~of~Freedom:~16~\\
+SSR:~104.05843~MSE:~6.503652~Root~MSE:~2.550226~\\
+Multiple~R-Squared:~0.611888~Adjusted~R-Squared:~0.539117~\\
+}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Other issues}
+
+\subsubsection{nlsystemfit issue 1}
+% The user should be aware that the function is \bold{VERY} sensative to
+% the starting values and the nlm function may not converge. The nlm
+% function will be called with the \code{typsize} argument set the
+% absolute values of the starting values for the OLS and 2SLS
+% methods. For the SUR and 3SLS methods, the \code{typsize} argument is
+% set to the absolute values of the resulting OLS and 2SLS parameter
+% estimates from the nlm result structre. In addition, the starting
+% values for the SUR and 3SLS methods are obtained from the OLS and 2SLS
+% parameter estimates to shorten the number of iterations. The number of
+% iterations reported in the summary are only those used in the last
+% call to nlm, thus the number of iterations in the OLS portion of the
+% SUR fit and the 2SLS portion of the 3SLS fit are not included.  }
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Other tools}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Likelihood ratio test}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Hausman test}
+
+\textbf{Man, I really have to clean this up.}
+
+Hausman \citep{hausman1978} developed a test have power against
+alternatives for which $\hat{\theta}$ and $\tilde{\theta}$ diverge
+under misspecification.
+
+The Hausman result if the difference between the covariance of an
+efficient and the covariance of the inefficient estimator is zero, the
+model is correctly specified.
+
+The test is based on the idea that under the hypothesis of no
+correlation, both the limited information estimates and the
+full-information results are consistent, but only the full-information
+estimates are efficient and under the alternative, the limited
+information estimates are consistent, but the full information
+estimates are not.
+
+Since the models that are fit using the \pkg{systemfit} package are
+typically random effects models, the effects may be inconsistent
+between the introduced variables and the regressors.
 
@@
 
-not implemented yet: unequal numbers of observations
-
-nlsystemfit finished?
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\bibliography{references} % a subset of my big bibtex file
+What should we place in this section?
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Unbalanced Observations}\label{sec:summary_unbalanced_data}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+The \pkg{systemfit} package was originally developed to fit
+simultaneous equations for forestry datasets. Forestry datasets
+typically contain observations for many inexpensive observations
+(diameter at breast height) and very few expensive observations such
+as total tree height. While the package currently does not allow for
+unbalanced datasets, data where the equations contain different
+numbers of observations, future releases of the package will have that
+feature implemented. We are still trying to figure out how develop the
+code for instrumental variable methods (2SLS,3SLS) for the correct
+estimation methods if possible.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Additional Methods}\label{sec:summary_additional_methods}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsubsection{Full Information Maximum Likelihood}
+
+Describe, in a paragraph the process and give some reasons why you
+might want this in the package.
+
+\subsubsection{Generalized Method of Moments}
+
+Describe, in a paragraph the process and give some reasons why you
+might want this in the package.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\bibliography{systemfit} % a subset of my big bibtex file
 %\bibliography{/home/suapm095/Documents/Literatur/arne}

### 2004-11-16 paper/trunk/systemfit.tex
added cvs header and log section. to be removed when submitted.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@564 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
+%       $Id$
+
 \documentclass[article]{jss}
@@
 
-\code{
-R> library( systemfit ) \\
-R> data( ppine ) \\
-R> hg.formula <- hg $\sim$ exp( h0 + h1*log(tht) + h2*tht$\hat$2 + h3*elev + h4*cr) \\
-R> dg.formula <- dg $\sim$ exp( d0 + d1*log(dbh) + d2*hg + d3*cr + d4*ba  ) \\
-R> labels <- list( "height.growth", "diameter.growth" )\\
-R> inst <- $\sim$ tht + dbh + elev + cr + ba\\
-R> start.values <- c(h0=-0.5, h1=0.5, h2=-0.001, h3=0.0001, h4=0.08,\\
-R+   d0=-0.5, d1=0.009, d2=0.25, d3=0.005, d4=-0.02 )\\
-R> model <- list( hg.formula, dg.formula )\\
-R> model.3sls <- nlsystemfit( "3SLS", model, start.values, data=ppine,\\
-R+   eqnlabels=labels, inst=inst ) \\
-R> summary( model.3sls ) \\
-}
+% \code{
+% R> library( systemfit ) \\
+% R> data( ppine ) \\
+% R> hg.formula <- hg $\sim$ exp( h0 + h1*log(tht) + h2*tht$\hat$2 + h3*elev + h4*cr) \\
+% R> dg.formula <- dg $\sim$ exp( d0 + d1*log(dbh) + d2*hg + d3*cr + d4*ba  ) \\
+% R> labels <- list( "height.growth", "diameter.growth" )\\
+% R> inst <- $\sim$ tht + dbh + elev + cr + ba\\
+% R> start.values <- c(h0=-0.5, h1=0.5, h2=-0.001, h3=0.0001, h4=0.08,\\
+% R+   d0=-0.5, d1=0.009, d2=0.25, d3=0.005, d4=-0.02 )\\
+% R> model <- list( hg.formula, dg.formula )\\
+% R> model.3sls <- nlsystemfit( "3SLS", model, start.values, data=ppine,\\
+% R+   eqnlabels=labels, inst=inst ) \\
+% R> summary( model.3sls ) \\
+% }
 
@@
 
+\section{CVS Revision Log}
+\begin{verbatim}
+$Log$
+Revision 1.5  2004/11/16 05:31:19  hamannj
+added cvs header and log section. to be removed when submitted.
+
+\begin{verbatim}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2004-11-16 paper/trunk/systemfit.tex
made small changes because my code section wouldn't compile. argh.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@565 591db2bd-e0a0-4d22-8c19-2192dbb7da31
question about nlsystemfit code section (example)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@566 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 
+%% Arne -- Did you put the ~'s in by hand?
+
 Summary results can be printed by\\
@@
 $Log$
+Revision 1.7  2004/11/16 05:34:47  hamannj
+question about nlsystemfit code section (example)
+
+Revision 1.6  2004/11/16 05:32:46  hamannj
+made small changes because my code section wouldn't compile. argh.
+
 Revision 1.5  2004/11/16 05:31:19  hamannj
@@
 
-\begin{verbatim}
+\end{verbatim}
 

### 2004-11-16 paper/trunk/systemfit.tex
merged file and resolved conflict due to some minor changes I did while Jeff was adding nlsystemfit.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@567 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 to calculate the (diagonal) variance-covariance 
-matrix of the residuals (see section ????).}.
+matrix of the residuals (see section \ref{sec:residcov}).}.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Residual variance-covariance matrix}
+\subsubsection{Residual variance-covariance matrix}\label{sec:residcov}
 
@@
 the true covariance matrix of the residuals cannot be determined.
-Thus, the this covariance matrix must be calculated from the
+Thus, this covariance matrix must be calculated from the
 \emph{estimated} residuals.
@@
 The estimated covariance matrix of the residuals ($\widehat{\Sigma} = 
-\left[ \sigma \right]$) 
+\left[ \widehat{\sigma}_{ij} \right]$) 
 is often calculated by
@@
 
-However, this estimator is biased, because it is not corrected for the
+However, this estimator is biased, because it is not corrected for 
 degrees of freedom. 
@@
 A further approach to obtain the estimated residual variance-covariance
-matrix is
-\begin{equation}
-   \widehat{\sigma}_{ij} = \frac{ \widehat{u}_i' \widehat{u}_j } 
+matrix is \citep[p. 309]{zellner62c}
+\begin{eqnarray}
+   \widehat{\sigma}_{ij} & = & 
+   \frac{ \widehat{u}_i' \widehat{u}_j } 
+   { T - K_i - K_j + tr \left[ X_i \left( X_i' X_i \right)^{-1}
+   X_i' X_j \left( X_j' X_j \right)^{-1} X_j' \right] } \\
+   & = &
+   \frac{ \widehat{u}_i' \widehat{u}_j } 
    { T - K_i - K_j + tr \left[ \left( X_i' X_i \right)^{-1}
-   X_i' X_j \left( X_j' X_j \right)^{-1} X_j' X_i \right] }
+   X_i' X_j \left( X_j' X_j \right)^{-1} X_j' X_i \right] } \\
    \label{eq:rcov2}
-\end{equation} 
+\end{eqnarray} 
 This yields an unbiased estimator for all elements of $\widehat{\Sigma}$,
@@
 It starts with a dependent variable on the left hand side.
-After a tilde ($~$) the regressors are listed%
+After a tilde ($\sim$) the regressors are listed%
 \footnote{For Details see the \proglang{R} help files to \code{formula}}.
@@
 $Log$
+Revision 1.8  2004/11/16 08:55:53  henningsena
+merged file and resolved conflict due to some minor changes I did while Jeff was adding nlsystemfit.
+
 Revision 1.7  2004/11/16 05:34:47  hamannj

### 2004-11-16 paper/trunk/systemfit.tex
some minor changes and some comments on nlsystemfit
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@568 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
   \pkg{systemfit} provides the capability to estimate systems  of
-  linear and non-linear equations in \proglang{R}.  } \Keywords{R,
-  simultaneous equation system, seemingly unrelated regression}
+  linear and non-linear equations in \proglang{R}.
+}
+
+\Keywords{R, simultaneous equation system, seemingly unrelated regression}
 
@@
 
-This paper is organized as follows:  In section \ref{sec:Theory} we
+This paper is organized as follows: In section~\ref{sec:Theory} we
 introduce the mathematics of estimating equation systems and give a
 comprehensive overview over all relevant issues regarding this topic.
-Section \ref{sec:UsingSystemfit} demonstrates how to run
+Section~\ref{sec:UsingSystemfit} demonstrates how to run
 \pkg{systemfit}, especially how the features presented in the previous
 chapter can be used. Finally, a summary and an outlook are given in
-section \ref{sec:Summmary}.
+section~\ref{sec:Summmary}.
 
@@
 The regressors are \code{p} (ratio of food prices to general consumer prices)
-and \code{disposable income} as well as a constant%
+and \code{d} (disposable income) as well as a constant%
 \footnote{a regression constant is always implied if not explicitly omitted.}.
@@
 \end{equation}
+%%% It is not clear what these variable are. 
+%%% Adding an equation representing only a single equation
+%%% would make things clearer.
 
@@
 
-where, $r$ is a column vector for the residuals for each equation, $X$
-is matrix of the partial derivatives with respect to the parameters,
+where, $r$ is a column vector for the residuals for each equation,
+%% what is the difference to \epsilon?
+$X$ is matrix of the partial derivatives of the dependent variable 
+with respect to the parameters $\left( \frac{ \partial y }{ \theta} \right)$,
+%% is this correct?
+%% in the linear section X is used for the regressors, 
+%% thus we should use a different variable name here.
 $W$ is a matrix of the instrument variables, $Z(Z'Z)^{-1}Z$, $Z$ is a
 matrix of the instrument variables, and $I$ is an $n \times n $
-identity matrix and $S$ is the variance-covariance matrix between the
+identity matrix and $S$ is the estimated variance-covariance matrix between the
 equations
@@
 \end{equation}
+
+\textbf{The residual variance-covariance matrix can be calculated in 
+different ways (section~\ref{sec:residcov}). 
+It should be relatively easy to implement this also in nlsystemfit(), 
+e.g. the function 'calcRCov' in systemfit() could be moved outside
+systemfit() that it can be used also by nlsystemfit().}
 
@@
 %% Arne -- Did you put the ~'s in by hand?
+%% Yes, I did. I used find and replace to substitute the "~" for " ". 
+%% This is necessary to retain the format of the output.
 
@@
 % The user should be aware that the function is \bold{VERY} sensative to
-% the starting values and the nlm function may not converge. The nlm
+% the starting values and the nlm function may not converge.
+
+% did you try optim()? Especially if you provide (analytical) derivatives
+% it is very good in finding the optimum.
+
+% The nlm
 % function will be called with the \code{typsize} argument set the
@@
 $Log$
+Revision 1.9  2004/11/16 10:59:09  henningsena
+some minor changes and some comments on nlsystemfit
+
 Revision 1.8  2004/11/16 08:55:53  henningsena

### 2005-09-01 paper/trunk/nlsystemfit.tex
moved section about non-linear estimation from systemfit.tex into new file nlsystemfit.tex, because this section should not be included in (this version of) the article
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@570 591db2bd-e0a0-4d22-8c19-2192dbb7da31

--- 2004-11-16 paper/trunk/systemfit.tex
@@
-%       $Id$
-
-\documentclass[article]{jss}
-
-\author{Arne Henningsen\\University of Kiel\And 
-        Jeff D. Hamann\\ Forest Informatics, Inc. }
-
-\title{systemfit: A Package to Estimate\\
-       Simultaneous Equation Systems in \proglang{R}}
-
-%% for pretty printing and a nice hypersummary also set:
-\Plainauthor{Arne Henningsen, Jeff D. Hamann} %% comma-separated
-\Plaintitle{systemfit: A Package to Estimate Simultaneous Equation Systems 
-            in R} %% without formatting
-\Shorttitle{systemfit} %% a short title (if necessary)
-
-%% an abstract and keywords
-\Abstract{ Many statistical analysis are based on models that consist
-  of  more than one equation.  These equations should be fitted
-  simultaneously rather than equation  by equation.  The package
-  \pkg{systemfit} provides the capability to estimate systems  of
-  linear and non-linear equations in \proglang{R}.
-}
-
-\Keywords{R, simultaneous equation system, seemingly unrelated regression}
-
-%% at least one keyword must be supplied
-
-%% publication information
-%% NOTE: This needs to filled out ONLY IF THE PAPER WAS ACCEPTED.
-%% If it was not (yet) accepted, leave them commented.
-%% \Volume{13}
-%% \Issue{9}
-%% \Month{September}
-%% \Year{2004}
-%% \Submitdate{2004-09-29}
-%% \Acceptdate{2004-09-29}
-
-%% The address of (at least) one author should be given
-%% in the following format:
-\Address{
-   Arne Henningsen\\
-   Department of Agricultural Economics\\
-   University of Kiel\\
-   D-24098 Kiel, Germany\\
-   E-mail: \email{ahenningsen@agric-econ.uni-kiel.de}\\
-   URL: \url{http://www.uni-kiel.de/agrarpol/ahenningsen/}\\
-   \\
-   Jeff D. Hamann\\
-   Forest Informatics, Inc.\\
-   PO Box 1421\\
-   Corvallis, Oregon 97339-1421\\
-   e-mail: \email{jeff.hamann@forestinformatics.com}\\
-   URL: \url{http://www.forestinformatics.com}\\
-}
-%% It is also possible to add a telephone and fax number
-%% before the e-mail in the following format:
-%% Telephone: +43/1/31336-5053
-%% Fax: +43/1/31336-734
-
-%% for those who use Sweave please include the following line (with % symbols):
-%% need no \usepackage{Sweave.sty}
-
-%% end of declarations %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{document}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Introduction}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-Many theoretical models consist of more than one equation. If these
-models are econometrically estimated, the equations generally should
-not be estimated individually, but all together, because the
-disturbance terms of these equations are likely contemporaneously
-correlated. Estimating all equations simultaneously and taking the
-covariance structure of the residuals into account leads to efficient
-estimates, while ignoring this correlation leads to an inefficient
-parameter estimation \citep{zellner62}.
-
-Another reason to estimate an equation system simultaneously are
-cross-equation parameter restrictions. These restrictions can be
-tested and/or imposed only in a simultaneous estimation
-approach. Especially the economic theory suggests many cross-equation
-restriction.
-
-The \pkg{systemfit} package provides the capability to estimate
-linear and non-linear equation systems in \proglang{R}. Compared to
-other statistics and econometrics software the (advanced) user can
-control many important details of the estimation procedure. On the
-other hand these user options have reasonable defaults that also
-beginners can easily use the provided functions.
-
-\textbf{I'll have to update this section to match the recent changes. JDH.}
-
-This paper is organized as follows: In section~\ref{sec:Theory} we
-introduce the mathematics of estimating equation systems and give a
-comprehensive overview over all relevant issues regarding this topic.
-Section~\ref{sec:UsingSystemfit} demonstrates how to run
-\pkg{systemfit}, especially how the features presented in the previous
-chapter can be used. Finally, a summary and an outlook are given in
-section~\ref{sec:Summmary}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Estimating linear equation systems}\label{sec:Theory}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-Consider a system of $G$ equations, where the $i$th equation is of
-the form 
-\begin{equation}
-   y_{i} = X_i \beta_i + u_i, \quad i = 1, 2, \ldots, G
-\end{equation}
-where $y_i$ is a vector of the dependent variable,
-$X_i$ is a matrix of the exogenous variables,
-$\beta_i$ is the coefficient vector and
-$u_i$ is a vector of the disturbance terms of the $i$th equation.
-
-We can write the 'stacked' system as
-\begin{equation}
-   \left[ \begin{array}{c}
-      y_1 \\ y_2\\ \vdots\\ y_G
-   \end{array} \right] = 
-   \left[ \begin{array}{cccc}
-      X_1 & 0 & \cdots & 0\\
-      0 & X_2 & \cdots & 0\\
-      \vdots & \vdots & \ddots & \vdots\\
-      0 & 0 & \cdots & X_G
-   \end{array}\right] \cdot 
-   \left[ \begin{array}{c}
-      \beta_1 \\ \beta_2 \\ \vdots\\ \beta_G
-   \end{array} \right] +
-   \left[ \begin{array}{c}
-      u_1 \\ u_2 \\ \vdots\\ u_G
-   \end{array} \right]
-\end{equation}
-
-or more simply as
-\begin{equation}
-   y = X \beta + u 
-\end{equation}
-
-We assume that there is no correlation of the disturbance terms 
-across observations:
-\begin{equation}
-   E \left( u_{it} \, u_{jt^*} \right) = 0 
-   \; \forall \; t \neq t^*
-\end{equation}
-where $i$ and $j$ indicate the equation number 
-and $t$ and $t^*$ denote the observation number.
-
-However, we explicitly allow for contemporaneous correlation:
-\begin{equation}
-   E \left( u_{it} \, u_{jt} \right) = \sigma_{ij}
-\end{equation}
-
-Thus, the covariance matrix of the total system is
-\begin{equation}
-   E \left( u \, u' \right) = \Omega = \Sigma \otimes I
-\end{equation}
-where $\Sigma = \left[ \sigma_{ij} \right]$ and $I$ is an
-identity matrix.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Linear Parameter estimation}
-
-\subsubsection{Ordinary Least Squares estimation}
-
-The Ordinary Least Squares (OLS) estimator of the system 
-is obtained by
-\begin{equation}
-   \widehat{\beta} = \left( X'X \right)^{-1} X'y
-\end{equation}
-
-These estimates are efficient only if the disturbance terms are not 
-contemporaneously correlated, which means 
-\begin{equation}
-   \sigma_{ij} = 0 \; \forall \; i \neq j
-\end{equation}
-
-If the whole system is treated as one single equation, 
-the variance-covariance matrix of the estimated parameters is
-\begin{equation}
-   Var \left[ \widehat{\beta} \right] = \sigma^2 \left( X'X \right)^{-1}
-\end{equation}
-with $\sigma^2 = E \left( u' u \right)$.
-This assumes that the disturbances of all equations have the
-same variance.
-
-If the disturbance terms of the individual equations 
-are allowed to have different variances, 
-the variance-covariance matrix of the estimated parameters is
-\begin{equation}
-   Var \left[ \widehat{\beta} \right] = \left( X' \Omega^{-1} X \right)^{-1}
-   \label{eq:OLSvcm2}
-\end{equation}
-with $\Omega = \Sigma \otimes I$, 
-$\sigma_{ij} = 0 \; \forall \; i \neq j$ and
-$\sigma_{ii} = E \left( u_i' u_i \right)$.
-
-If no cross-equation restrictions are imposed, the simultaneous 
-OLS estimation of the system leads to the same parameter estimates 
-as an equation-wise OLS estimation. The variance-covariance matrix 
-of the parameters from an equation-wise OLS estimation is equal to 
-the variance-covariance matrix obtained by equation 
-(\ref{eq:OLSvcm2}).
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Weighted least squares}
-
-The Weighted Least Squares (WLS) estimator of the system 
-is obtained by
-\begin{equation}
-   \widehat{\beta} = \left( X' \Omega^{-1} X \right)^{-1} X' \Omega^{-1} y
-\end{equation}
-with $\Omega = \Sigma \otimes I$, 
-$\sigma_{ij} = 0 \; \forall \; i \neq j$ and
-$\sigma_{ii} = E \left( u_i' u_i \right)$.
-
-Like the OLS estimates these estimates are only efficient 
-if the disturbance terms are not contemporaneously correlated.
-
-The variance-covariance matrix of the estimated parameters is
-\begin{equation}
-   Var \left[ \widehat{\beta} \right] = \left( X' \Omega^{-1} X \right)^{-1}
-\end{equation}
-
-If no cross-equation restrictions are imposed, 
-the parameter estimates and the variance-covariance matrix 
-of the parameters are equal to the OLS estimates%
-\footnote{The variance-covariance matrices of the 
-estimated parameters might differ, 
-if the WLS estimation is based on a different formula 
-to calculate the (diagonal) variance-covariance 
-matrix of the residuals (see section \ref{sec:residcov}).}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Seemingly unrelated regression}
-
-When the disturbances are contemporaneously correlated, a Generalized 
-Least Squares (GLS) estimation leads to efficient parameter estimates.
-In this case the GLS is generally called "Seemingly Unrelated Regression" 
-(SUR) (see \citealp{zellner62}). 
-Note, while an unbiased OLS or WLS estimation requires only that 
-the regressors and the disturbance terms of each single 
-equation are uncorrelated,
-a consistent SUR estimation requires that all disturbance terms and all 
-regressors are uncorrelated:
-\begin{equation}
-   E \left( u | X \right) = 0
-\end{equation}
-
-The SUR estimator can be obtained by:%
-\footnote{To calculate $\Omega^{-1}$ it is not necessary to invert 
-the (huge) $\Omega$ matrix, 
-because it is sufficient to invert the (small) $\Sigma$ matrix and
-calculate it by $\Omega^{-1} = \Sigma^{-1} \otimes I$.}
-
-\begin{equation}
-   \widehat{\beta} = \left( X' \Omega^{-1} X \right)^{-1} X' \Omega^{-1} y
-\end{equation}
-with $\Omega = \Sigma \otimes I$ and
-$\sigma_{ij} = E \left( u_i' u_j \right)$.
-
-And the variance-covariance matrix of the estimated parameters is
-\begin{equation}
-   Var \left[ \widehat{\beta} \right] = \left( X' \Omega^{-1} X \right)^{-1}
-\end{equation}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Two-stage least squares estimation}
-
-If the regressors of one or more equations are correlated 
-with the disturbances ($E \left( u_i | X_i \right) \neq 0$), 
-the estimated coefficients are biased.
-This can be circumvented by an instrumental variable (IV) estimation.
-The instrumental variables for each equation $H_i$ 
-can be either different or identical for all equations.
-The instrumental variables of each equation may not be correlated with 
-the disturbance terms of the corresponding equation 
-($E \left( u_i | H_i \right) = 0$).
-
-At the first stage new ('fitted') regressors are obtained by
-\begin{equation}
-   \widehat{X_i} = H_i \left( H_i' H_i \right)^{-1} H_i' X
-\end{equation}
-
-At the second stage the unbiased two-stage least squares estimates
-of $\beta$ are obtained by:
-\begin{equation}
-   \widehat{\beta} = \left( \widehat{X}' \widehat{X} \right)^{-1} 
-   \widehat{X}' y 
-   \label{eq:beta2sls}
-\end{equation}
-
-If the whole system is treated as one single equation, 
-the variance-covariance matrix of the estimated parameters is
-\begin{equation}
-   Var \left[ \widehat{\beta} \right] = \sigma^2 \left( \widehat{X}'
-   \widehat{X} \right)^{-1}
-\end{equation}
-with $\sigma^2 = E \left( u' u \right)$.
-
-If the disturbance terms of the individual equations 
-are allowed to have different variances, 
-the variance-covariance matrix of the estimated parameters is
-\begin{equation}
-   Var \left[ \widehat{\beta} \right] = \left( \widehat{X}' \Omega^{-1} 
-   \widehat{X} \right)^{-1}
-\end{equation}
-with $\Omega = \Sigma \otimes I$, 
-$\sigma_{ij} = 0 \; \forall \; i \neq j$ and
-$\sigma_{ii} = E \left( u_i' u_i \right)$.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Weighted two-stage least squares estimation}
-
-The Weighted Two-Stage Least Squares (W2SLS) estimator of the system 
-is obtained by
-\begin{equation}
-   \widehat{\beta} = \left( \widehat{X}' \Omega^{-1} \widehat{X} 
-   \right)^{-1} \widehat{X}' \Omega^{-1} y
-\end{equation}
-with $\Omega = \Sigma \otimes I$, 
-$\sigma_{ij} = 0 \; \forall \; i \neq j$ and
-$\sigma_{ii} = E \left( u_i' u_i \right)$.
-
-The variance-covariance matrix of the estimated parameters is
-\begin{equation}
-   Var \left[ \widehat{\beta} \right] = \left( \widehat{X}' \Omega^{-1} 
-   \widehat{X} \right)^{-1}
-\end{equation}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Three-stage least squares estimation}
-
-If the regressors are correlated with the disturbances 
-($E \left( u | X \right) \neq 0$) and 
-the disturbances are contemporaneously correlated, 
-an Instrumental Variable (IV) Generalized Least Squares (GLS) 
-estimation leads to consistent and efficient estimates. This estimation 
-procedure is generally called "Three-stage Least Squares" (3SLS).
-
-The standard 3SLS estimator can be obtained by:
-\begin{equation}
-   \widehat{\beta} = \left( \widehat{X}' \Omega^{-1} \widehat{X} 
-   \right)^{-1} \widehat{X}' \Omega^{-1} y
-\end{equation}
-with $\Omega = \Sigma \otimes I$ and
-$\sigma_{ij} = E \left( u_i' u_j \right)$.
-
-Its variance-covariance matrix is:
-\begin{equation}
-   Var \left[ \widehat{\beta} \right] = \left( \widehat{X}' \Omega^{-1} 
-   \widehat{X} \right)^{-1}
-   \label{eq:cov3sls}
-\end{equation}
-
-While an unbiased 2SLS or W2SLS estimation requires only that 
-the instrumental variables and the disturbance terms of each single 
-equation are uncorrelated,
-\cite{schmidt90} points out that this estimator is only consistent 
-if all disturbance terms and all instrumental variables are uncorrelated:
-\begin{equation}
-   E \left( u | H \right) = 0
-\end{equation}
-with
-\begin{equation}
-   H =
-   \left[ \begin{array}{cccc}
-      H_1 & 0 & \cdots & 0\\
-      0 & H_2 & \cdots & 0\\
-      \vdots & \vdots & \ddots & \vdots\\
-      0 & 0 & \cdots & H_G
-   \end{array}\right] \cdot 
-\end{equation}
-
-Since there might be occasions where this cannot be avoided, 
-\cite{schmidt90} analyses other approaches to obtain 3SLS estimators:
-
-One of these approaches is based on instrumental variable estimation
-(IV-3SLS):
-\begin{equation}
-   \widehat{\beta} = \left( \widehat{X}' \Omega^{-1} X 
-   \right)^{-1} \widehat{X}' \Omega^{-1} y
-\end{equation}
-
-The variance-covariance matrix of this IV-3SLS estimator is:
-\begin{equation}
-   Var \left[ \widehat{\beta} \right] = \left( \widehat{X}' \Omega^{-1} 
-   X \right)^{-1}
-\end{equation}
-
-Another approach is based on the Generalized Method of Moments (GMM) 
-estimator (GMM-3SLS):
-\begin{equation}
-   \widehat{\beta} = \left( X' H \left( H' \Omega H \right)^{-1}
-   H' X \right)^{-1} X' H \left( H' \Omega H \right)^{-1} H' y
-   \label{eq:beta3slsGMM}
-\end{equation}
-
-The variance-covariance matrix of the GMM-3SLS estimator is:
-\begin{equation}
-   Var \left[ \widehat{\beta} \right] = 
-   \left( X' H \left( H' \Omega H \right)^{-1} H' X \right)^{-1}
-\end{equation}
-
-A fourth approach developed by \cite{schmidt90} himself is:
-\begin{equation}
-   \widehat{\beta} = \left( \widehat{X}' \Omega^{-1} \widehat{X} 
-   \right)^{-1} \widehat{X}' \Omega^{-1} 
-   H \left( H' H \right)^{-1} H' y 
-   \label{eq:beta3slsSchmidt}
-\end{equation}
-
-The variance-covariance matrix of this estimator is:
-\begin{equation}
-   Var \left[ \widehat{\beta} \right] = 
-   \left( \widehat{X}' \Omega^{-1}  \widehat{X} \right)^{-1} 
-   \widehat{X}' \Omega^{-1} H \left( H' H \right)^{-1} H' \Omega 
-   H \left( H' H \right)^{-1} H' \Omega^{-1} \widehat{X}
-   \left( \widehat{X}' \Omega^{-1}  \widehat{X} \right)^{-1}
-\end{equation}
-
-The econometrics software EViews uses following approach:
-\begin{equation}
-   \widehat{\beta} = \widehat{\beta}_{2SLS} + 
-   \left( \widehat{X}' \Omega^{-1} \widehat{X} \right)^{-1} 
-   \widehat{X}' \Omega^{-1} \left( y - X \widehat{\beta}_{2SLS} \right)
-\end{equation}
-with $\widehat{\beta}_{2SLS}$ is the two-stage least squares estimator
-(equation \ref{eq:beta2sls}).
-EViews uses the standard 3SLS formula (equation (\ref{eq:cov3sls})) to 
-calculate the variance-covariance matrix of the 3SLS estimator.
-
-If the same instrumental variables are used in all equations 
-($H_1 = H_2 = \ldots = H_G$), 
-all the above mentioned approaches lead to identical parameter estimates.
-However, if this is not the case, the results depend on the 
-equation used \citep{schmidt90}. 
-The only reason to use different instruments in the different equations 
-is that the instruments of one equation are correlated with the
-disturbance terms in another equation. Otherwise, one would simply use
-every instrument in every equation \citep{schmidt90}. 
-In this case, only the GMM-3SLS (equation (\ref{eq:beta3slsGMM}))
-and the 3SLS estimator developed by \cite{schmidt90} 
-(equation (\ref{eq:beta3slsSchmidt})) are consistent.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Imposing linear restrictions}
-There are two ways to impose linear parameter restrictions.
-First, a matrix $T$ can be provided with
-\begin{equation}
-   \beta = T \cdot \beta^* \label{eq:T-restr} 
-\end{equation}
-where $\beta^*$ is a vector of linear independent coefficients.
-
-To impose these restrictions the $X$ matrix is
-(post-)multiplied by this $T$ matrix.
-\begin{equation}
-    X^* = X \cdot T
-\end{equation}
-
-Then, a standard estimation is done by substituting $X^*$ for $X$.
-This results in the linear independent parameter estimates $\beta^*$ and
-the variance-covariance matrix of these parameters. 
-The original parameters can be obtained by equation (\ref{eq:T-restr})
-and the variance-covariance matrix of the original parameters 
-can be obtained by:
-\begin{equation}
-   Var \left[ \widehat{\beta} \right] = T \cdot Var \left[ \widehat{\beta^*} \right] \cdot T'
-\end{equation}
-
-The second way to impose linear parameter restrictions 
-can be formulated by
-\begin{equation}
-   R \beta^0 = q
-\end{equation}
-where $\beta^0$ is the vector of the restricted coefficients, 
-$R$ is a matrix and $q$ is a vector (see \citealp[p. 100]{greene02}). 
-Each restriction belongs to one row of $R$ and the corresponding 
-element of~$q$.
-
-The first way is less flexible than this latter one%
-\footnote{e.g. restrictions like $\beta_1 + \beta_2 = 4$ cannot be imposed
-by the first method}, 
-but the first way is preferable if equality constraints for coefficients
-across many equations of the system are imposed. 
-Of course, these restrictions can be also imposed using
-the latter method.
-However, while the latter method increases the dimension of the 
-matrices to be inverted during estimation, the first reduces it. 
-Thus, in some cases the latter way leads to estimation problems
-(e.g. (near) singularity of the matrices to be inverted),
-while the first doesn't.
-
-These two methods can be combined. In this case the restrictions
-imposed using the latter method are imposed on the linear independent 
-parameters due to the restrictions imposed using the first method:
-\begin{equation}
-   R \beta^{*0} = q
-\end{equation}
-where $\beta^{*0}$ is the vector of the restricted $\beta^*$ coefficients.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Restricted OLS estimation}
-
-The OLS estimator restricted by $R \beta^0 = q$ can be obtained by
-\begin{equation}
-   \left[ \begin{array}{c}
-      \widehat{\beta^0} \\ \widehat{\lambda}
-   \end{array} \right]
-   =
-   \left[ \begin{array}{cc}
-      X' X & R' \\ 
-      R & 0
-   \end{array} \right]^{-1}
-   \cdot
-   \left[ \begin{array}{c}
-      X' y \\ q 
-   \end{array} \right]
-\end{equation}
-where $2\lambda$ is a vector of the Lagrangean multipliers of the restrictions.
-
-If the whole system is treated as one single equation, 
-the variance-covariance matrix of the estimated parameters is
-\begin{equation}
-   Var 
-   \left[ \begin{array}{c}
-      \widehat{\beta} \\ \widehat{\lambda}
-   \end{array} \right] 
-   = \sigma^2 
-   \left[ \begin{array}{cc}
-      X' X & R' \\ 
-      R & 0
-   \end{array} \right]^{-1}
-\end{equation}
-with $\sigma^2 = E \left( u' u \right)$.
-
-If the disturbance terms of the individual equations 
-are allowed to have different variances, 
-the variance-covariance matrix of the estimated parameters is
-\begin{equation}
-   Var 
-   \left[ \begin{array}{c}
-      \widehat{\beta} \\ \widehat{\lambda}
-   \end{array} \right] 
-   = 
-   \left[ \begin{array}{cc}
-      X' \Omega^{-1} X & R' \\ 
-      R & 0
-   \end{array} \right]^{-1}
-\end{equation}
-with $\Omega = \Sigma \otimes I$, 
-$\sigma_{ij} = 0 \; \forall \; i \neq j$ and
-$\sigma_{ii} = E \left( u_i' u_i \right)$.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Restricted WLS estimation}
-
-The WLS estimator restricted by $R \beta^0 = q$ can be obtained by
-\begin{equation}
-   \left[ \begin{array}{c}
-      \widehat{\beta^0} \\ \widehat{\lambda}
-   \end{array} \right]
-   =
-   \left[ \begin{array}{cc}
-      X' \Omega^{-1} X & R' \\ 
-      R & 0
-   \end{array} \right]^{-1}
-   \cdot
-   \left[ \begin{array}{c}
-      X' \Omega^{-1} y \\ q 
-   \end{array} \right]
-\end{equation}
-with $\Omega = \Sigma \otimes I$, 
-$\sigma_{ij} = 0 \; \forall \; i \neq j$ and
-$\sigma_{ii} = E \left( u_i' u_i \right)$.
-
-The variance-covariance matrix of the estimated parameters is
-\begin{equation}
-   Var 
-   \left[ \begin{array}{c}
-      \widehat{\beta} \\ \widehat{\lambda}
-   \end{array} \right] 
-   = 
-   \left[ \begin{array}{cc}
-      X' \Omega^{-1} X & R' \\ 
-      R & 0
-   \end{array} \right]^{-1}
-\end{equation}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Restricted SUR estimation}
-
-The SUR estimator restricted by $R \beta^0 = q$ can be obtained by
-\begin{equation}
-   \left[ \begin{array}{c}
-      \widehat{\beta^0} \\ \widehat{\lambda}
-   \end{array} \right]
-   =
-   \left[ \begin{array}{cc}
-      X' \Omega^{-1} X & R' \\ 
-      R & 0
-   \end{array} \right]^{-1}
-   \cdot
-   \left[ \begin{array}{c}
-      X' \Omega^{-1} y \\ q 
-   \end{array} \right]
-\end{equation}
-with $\Omega = \Sigma \otimes I$ and
-$\sigma_{ij} = E \left( u_i' u_j \right)$.
-
-The variance-covariance matrix of the estimated parameters is
-\begin{equation}
-   Var 
-   \left[ \begin{array}{c}
-      \widehat{\beta} \\ \widehat{\lambda}
-   \end{array} \right] 
-   = 
-   \left[ \begin{array}{cc}
-      X' \Omega^{-1} X & R' \\ 
-      R & 0
-   \end{array} \right]^{-1}
-\end{equation}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Restricted 2SLS estimation}
-
-The 2SLS estimator restricted by $R \beta^0 = q$ can be obtained by
-\begin{equation}
-   \left[ \begin{array}{c}
-      \widehat{\beta^0} \\ \widehat{\lambda}
-   \end{array} \right]
-   =
-   \left[ \begin{array}{cc}
-      \widehat{X}' \widehat{X} & R' \\ 
-      R & 0
-   \end{array} \right]^{-1}
-   \cdot
-   \left[ \begin{array}{c}
-      \widehat{X}' y \\ q 
-   \end{array} \right]
-   \label{eq:beta2SLSr}
-\end{equation}
-
-If the whole system is treated as one single equation, 
-the variance-covariance matrix of the estimated parameters is
-\begin{equation}
-   Var 
-   \left[ \begin{array}{c}
-      \widehat{\beta} \\ \widehat{\lambda}
-   \end{array} \right] 
-   = \sigma^2 
-   \left[ \begin{array}{cc}
-      \widehat{X}' \widehat{X} & R' \\ 
-      R & 0
-   \end{array} \right]^{-1}
-\end{equation}
-with $\sigma^2 = E \left( u' u \right)$.
-
-If the disturbance terms of the individual equations 
-are allowed to have different variances, 
-the variance-covariance matrix of the estimated parameters is
-\begin{equation}
-   Var 
-   \left[ \begin{array}{c}
-      \widehat{\beta} \\ \widehat{\lambda}
-   \end{array} \right] 
-   = 
-   \left[ \begin{array}{cc}
-      \widehat{X}' \Omega^{-1} \widehat{X} & R' \\ 
-      R & 0
-   \end{array} \right]^{-1}
-\end{equation}
-with $\Omega = \Sigma \otimes I$, 
-$\sigma_{ij} = 0 \; \forall \; i \neq j$ and
-$\sigma_{ii} = E \left( u_i' u_i \right)$.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Restricted W2SLS estimation}
-
-The W2SLS estimator restricted by $R \beta^0 = q$ can be obtained by
-\begin{equation}
-   \left[ \begin{array}{c}
-      \widehat{\beta^0} \\ \widehat{\lambda}
-   \end{array} \right]
-   =
-   \left[ \begin{array}{cc}
-      \widehat{X}' \Omega^{-1} \widehat{X} & R' \\ 
-      R & 0
-   \end{array} \right]^{-1}
-   \cdot
-   \left[ \begin{array}{c}
-      \widehat{X}' \Omega^{-1} y \\ q 
-   \end{array} \right]
-\end{equation}
-with $\Omega = \Sigma \otimes I$, 
-$\sigma_{ij} = 0 \; \forall \; i \neq j$ and
-$\sigma_{ii} = E \left( u_i' u_i \right)$.
-
-The variance-covariance matrix of the estimated parameters is
-\begin{equation}
-   Var 
-   \left[ \begin{array}{c}
-      \widehat{\beta} \\ \widehat{\lambda}
-   \end{array} \right] 
-   = 
-   \left[ \begin{array}{cc}
-      \widehat{X}' \Omega^{-1} \widehat{X} & R' \\ 
-      R & 0
-   \end{array} \right]^{-1}
-\end{equation}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Restricted 3SLS estimation}
-
-The standard 3SLS estimator restricted by $R \beta^0 = q$ can be obtained by
-\begin{equation}
-   \left[ \begin{array}{c}
-      \widehat{\beta^0} \\ \widehat{\lambda}
-   \end{array} \right]
-   =
-   \left[ \begin{array}{cc}
-      \widehat{X}' \Omega^{-1} \widehat{X} & R' \\ 
-      R & 0
-   \end{array} \right]^{-1}
-   \cdot
-   \left[ \begin{array}{c}
-      \widehat{X}' \Omega^{-1} y \\ q 
-   \end{array} \right]
-\end{equation}
-with $\Omega = \Sigma \otimes I$ and
-$\sigma_{ij} = E \left( u_i' u_j \right)$.
-
-The variance-covariance matrix of this estimator is
-\begin{equation}
-   Var 
-   \left[ \begin{array}{c}
-      \widehat{\beta^0} \\ \widehat{\lambda}
-   \end{array} \right] 
-   = 
-   \left[ \begin{array}{cc}
-      \widehat{X}' \Omega^{-1} \widehat{X} & R' \\ 
-      R & 0
-   \end{array} \right]^{-1}
-   \label{eq:cov3slsr}
-\end{equation}
-
-The IV-3SLS estimator restricted by $R \beta^0 = q$ can be obtained by
-\begin{equation}
-   \left[ \begin{array}{c}
-      \widehat{\beta^0} \\ \widehat{\lambda}
-   \end{array} \right]
-   =
-   \left[ \begin{array}{cc}
-      \widehat{X}' \Omega^{-1} X & R' \\ 
-      R & 0
-   \end{array} \right]^{-1}
-   \cdot
-   \left[ \begin{array}{c}
-      \widehat{X}' \Omega^{-1} y \\ q 
-   \end{array} \right]
-\end{equation}
-
-with
-\begin{equation}
-   Var 
-   \left[ \begin{array}{c}
-      \widehat{\beta^0} \\ \widehat{\lambda}
-   \end{array} \right] 
-   = 
-   \left[ \begin{array}{cc}
-      \widehat{X}' \Omega^{-1} \widehat{X} & R' \\ 
-      R & 0
-   \end{array} \right]^{-1}
-\end{equation}
-
-The restricted 3SLS-GMM estimator can be obtained by
-\begin{equation}
-   \left[ \begin{array}{c}
-      \widehat{\beta^0} \\ \widehat{\lambda}
-   \end{array} \right]
-   =
-   \left[ \begin{array}{cc}
-      X' H \left( H' \Omega H \right)^{-1} H' X & R' \\ 
-      R & 0
-   \end{array} \right]^{-1}
-   \cdot
-   \left[ \begin{array}{c}
-      X' H \left( H \Omega H \right)^{-1} H' y \\ q 
-   \end{array} \right]
-\end{equation}
-
-with
-\begin{equation}
-   Var 
-   \left[ \begin{array}{c}
-      \widehat{\beta^0} \\ \widehat{\lambda}
-   \end{array} \right] 
-   = 
-   \left[ \begin{array}{cc}
-      X' H \left( H' \Omega H \right)^{-1} H' X & R' \\ 
-      R & 0
-   \end{array} \right]^{-1}
-\end{equation}
-
-The restricted 3SLS estimator based on the suggestion of 
-\cite{schmidt90} is:
-\begin{equation}
-   \left[ \begin{array}{c}
-      \widehat{\beta^0} \\ \widehat{\lambda}
-   \end{array} \right]
-   =
-   \left[ \begin{array}{cc}
-      \widehat{X}' \Omega^{-1} \widehat{X} & R' \\ 
-      R & 0
-   \end{array} \right]^{-1}
-   \cdot
-   \left[ \begin{array}{c}
-      \widehat{X}' \Omega^{-1} H \left( H' H \right)^{-1} H' y \\ q 
-   \end{array} \right]
-\end{equation}
-
-with
-\begin{eqnarray}
-   Var 
-   \left[ \begin{array}{c}
-      \widehat{\beta^0} \\ \widehat{\lambda}
-   \end{array} \right] 
-   & = & 
-   \left[ \begin{array}{cc}
-      \widehat{X}' \Omega^{-1} \widehat{X} & R' \\ 
-      R & 0
-   \end{array} \right]^{-1}
-   \cdot
-   \left[ \begin{array}{cc}
-      \widehat{X}' \Omega^{-1} H \left( H' H \right)^{-1} H' \Omega
-      H \left( H' H \right)^{-1} H' \Omega^{-1} \widehat{X} & 0' \\ 
-      0 & 0
-   \end{array} \right]^{-1}
-   \nonumber \\
-   & & \cdot
-   \left[ \begin{array}{cc}
-      \widehat{X}' \Omega^{-1} \widehat{X} & R' \\ 
-      R & 0
-   \end{array} \right]^{-1}
-\end{eqnarray}
-
-The econometrics software EViews calculates the restricted 3SLS estimator by:
-\begin{equation}
-   \left[ \begin{array}{c}
-      \widehat{\beta^0} \\ \widehat{\lambda}
-   \end{array} \right]
-   =
-   \left[ \begin{array}{cc}
-      \widehat{X}' \Omega^{-1} \widehat{X} & R' \\ 
-      R & 0
-   \end{array} \right]^{-1}
-   \cdot
-   \left[ \begin{array}{c}
-      \widehat{X}' \Omega^{-1} \left( y - X \widehat{\beta^0}_{2SLS} \right) 
-      \\ q 
-   \end{array} \right]
-\end{equation}
-where $\widehat{\beta^0}_{2SLS}$ is the restricted 2SLS estimator calculated
-by equation (\ref{eq:beta2SLSr}). 
-To calculate the variance-covariance matrix 
-EViews uses the standard formula of the restricted 3SLS estimator
-(equation (\ref{eq:cov3slsr})).
-
-If the same instrumental variables are used in all equations 
-($H_1 = H_2 = \ldots = H_G$), 
-all the above mentioned approaches lead to identical parameter estimates
-and identical variance-covariance matrices of the estimated parameters.
-
-(***** This has to be checked!!!!! **************)
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Other issues}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Residual variance-covariance matrix}\label{sec:residcov}
-
-Since the true residuals of the estimated equations are generally not known,
-the true covariance matrix of the residuals cannot be determined.
-Thus, this covariance matrix must be calculated from the
-\emph{estimated} residuals.
-
-The estimated covariance matrix of the residuals ($\widehat{\Sigma} = 
-\left[ \widehat{\sigma}_{ij} \right]$) 
-is often calculated by
-\begin{equation}
-   \widehat{\sigma}_{ij} = \frac{ \widehat{u}_i' \widehat{u}_j }{ T }
-   \label{eq:rcov0}
-\end{equation}
-where $T$ is the number of observations in each equation.
-
-However, this estimator is biased, because it is not corrected for 
-degrees of freedom. 
-The usual single-equation procedure to correct for degrees of freedom 
-cannot always be applied, because the number of regressors in each equation 
-might differ.
-Two alternative approaches to calculate the residual variance-covariance 
-matrix are
-\begin{equation}
-   \widehat{\sigma}_{ij} = \frac{ \widehat{u}_i' \widehat{u}_j }
-   { \sqrt{ \left( T - K_i \right) \cdot \left( T - K_j \right) } }
-   \label{eq:rcov1}
-\end{equation}
-and
-\begin{equation}
-   \widehat{\sigma}_{ij} = \frac{ \widehat{u}_i' \widehat{u}_j }
-   { T - \max \left( K_i , K_j \right) }
-\end{equation}
-where $K_i$ and $K_j$ are the number of regressors in equation
-$i$ and $j$, respectively.
-
-However, this yields an unbiased estimator only if $K_i = K_j$ 
-\citep[p. 469]{judge85}. 
-% Greene (2002, p. 344) says that the second is unbiased if i = j or K_i = K_j,
-% whereas the first is unbiased only if i = j. 
-% However, if K_i = K_j the first and the second are equal.
-% Why is the first biased if K_i = K_j ???????????
-
-A further approach to obtain the estimated residual variance-covariance
-matrix is \citep[p. 309]{zellner62c}
-\begin{eqnarray}
-   \widehat{\sigma}_{ij} & = & 
-   \frac{ \widehat{u}_i' \widehat{u}_j } 
-   { T - K_i - K_j + tr \left[ X_i \left( X_i' X_i \right)^{-1}
-   X_i' X_j \left( X_j' X_j \right)^{-1} X_j' \right] } \\
-   & = &
-   \frac{ \widehat{u}_i' \widehat{u}_j } 
-   { T - K_i - K_j + tr \left[ \left( X_i' X_i \right)^{-1}
-   X_i' X_j \left( X_j' X_j \right)^{-1} X_j' X_i \right] } \\
-   \label{eq:rcov2}
-\end{eqnarray} 
-This yields an unbiased estimator for all elements of $\widehat{\Sigma}$,
-but even if $\widehat{\Sigma}$ is an unbiased estimator of $\Sigma$, 
-its inverse $\widehat{\Sigma}^{-1}$ is not an unbiased estimator 
-of $\Sigma^{-1}$ \citep[p.322]{theil71}.
-Furthermore, the variance-covariance matrix calculated by (\ref{eq:rcov2})
-is not necessarily positive semidefinite \citep[p.322]{theil71}. 
-Hence, "it is doubtful whether [this formula] is really superior to 
-[equation (\ref{eq:rcov0})]" \citep[p.322]{theil71}.
-
-The WLS, SUR, W2SLS and 3SLS parameter estimates are consistent,
-if the estimated residual variance-covariance matrix is calculated
-using the residuals from a first step OLS or 2SLS estimation.
-It is also possible to iterate the estimation procedure and 
-calculate the residual variance-covariance matrix from the
-residuals of the previous iteration. 
-If equation (\ref{eq:rcov0}) is applied to calculate the estimated
-residual variance-covariance matrix, the results are maximum
-likelihood \citep[p. 345]{greene02}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Degrees of freedom}
-
-There exist different approaches to determine the degrees of
-freedom for testing the estimated parameters with a t-test.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Goodness of fit}
-
-The goodness of fit of each equation can be measured by the 
-traditional $R^2$ values:
-\begin{equation}
-   R_i^2 = 1 - \frac{ \widehat{u}_i' \widehat{u}_i }
-   { ( y_i - \overline{y_i} )' ( y_i - \overline{y_i} ) }
-\end{equation}
-where $R_i^2$ is the $R^2$ value of the $i$th equation
-and $\overline{y_i}$ is the mean value of $y_i$.
-
-The goodness of fit of the whole system can be measured by the
-McElroy's $R^2$ value \citep{mcelroy77}: 
-% also: \citep[p. 345]{greene02}
-\begin{equation}
-   R_*^2 = 1 - \frac{ \widehat{u}' \widehat{ \Omega }^{-1} \widehat{u} }
-   { y' \left( \widehat{ \Sigma }^{-1} \otimes
-   \left( I - \frac{i i'}{T} \right) \right) y }
-\end{equation}
-where $T$ is the number of observations in each equation,
-$I$ is an $T \times T$ identity matrix and 
-$i$ is a column vector of $T$ ones.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Using systemfit}\label{sec:UsingSystemfit}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Standard usage}
-
-\pkg{systemfit} is generally called by
-
-\code{
-R> systemfit( method, eqns )
-}
-
-There are two mandatory arguments: \code{method} and \code{eqns}.
-
-The argument \code{method} is a string determining the estimation method.
-It must be one of "OLS", "WLS", "SUR", "2SLS", "W2SLS" or "3SLS".
-
-The other mandatory argument \code{eqns} is a list of the equations 
-to estimate. 
-Each equation is a standard formula in \proglang{R}.
-It starts with a dependent variable on the left hand side.
-After a tilde ($\sim$) the regressors are listed%
-\footnote{For Details see the \proglang{R} help files to \code{formula}}.
-
-This is now demonstrated using an example: \\
-\code{
-R> library( systemfit ) \\
-R> data( kmenta ) \\
-R> attach( kmenta ) \\
-R> fitsur <- systemfit( "SUR", list( q $\sim$ p + d, q $\sim$ p + f + a ) ) \\
-}
-
-The first line loads the \pkg{systemfit} package. 
-The second line loads example data that are included in this package.
-These data come from \cite{kmenta86}.
-They are attached to the \proglang{R} search path in line three.
-In the last line a seemingly unrelated regression is done.
-The first equation represents the demand side of the food market.
-The dependant variable is \code{q} (food consumption per capita). 
-The regressors are \code{p} (ratio of food prices to general consumer prices)
-and \code{d} (disposable income) as well as a constant%
-\footnote{a regression constant is always implied if not explicitly omitted.}.
-The second equation represents the supply side.
-Variable \code{q} (food consumption per capita) is also the dependant 
-variable of this equation. 
-The regressors are again \code{p} (ratio of food prices to general 
-consumer prices) and a constant as well as 
-\code{f} (ratio of preceding year's prices received by farmers) and 
-\code{a} (a time trend in years).
-The regression result is assigned to the variable \code{fitsur}.
-
-Summary results can be printed by\\
-\code{
-R>~summary(~fitsur~)~\\
-~\\
-systemfit~results~\\
-method:~SUR~\\
-\\
-\mbox{}~~~N~DF~~~~~~SSR~~~~~MSE~~~~RMSE~~~~~~~R2~~~Adj~R2 \\
-1~20~17~~65.6829~3.86370~1.96563~0.755019~0.726198~\\
-2~20~16~104.0584~6.50365~2.55023~0.611888~0.539117~\\
-~\\
-The~covariance~matrix~of~the~residuals~used~for~estimation\\
-\mbox{}~~~~~~~~1~~~~~~~2~\\
-1~3.72539~4.13696~\\
-2~4.13696~5.78444~\\
-~\\
-The~covariance~matrix~of~the~residuals\\
-\mbox{}~~~~~~~~1~~~~~~~2~\\
-1~3.86370~4.92431~\\
-2~4.92431~6.50365~\\
-~\\
-The~correlations~of~the~residuals\\
-\mbox{}~~~~~~~~~1~~~~~~~~2~\\
-1~1.000000~0.982348~\\
-2~0.982348~1.000000~\\
-~\\
-The~determinant~of~the~residual~covariance~matrix:~0.879285~\\
-OLS~R-squared~value~of~the~system:~0.683453~\\
-McElroy's~R-squared~value~for~the~system:~0.788722~\\
-~\\
-SUR~estimates~for~1~~(equation~1~)~\\
-Model~Formula:~q~~~p~+~d\\
-\\
-\mbox{}~~~~~~~~~~~~~Estimate~Std.~Error~~~t~value~Pr(>|t|)~\\
-(Intercept)~99.332894~~~7.514452~13.218913~~~~~~~~0~***~\\
-p~~~~~~~~~~~-0.275486~~~0.088509~-3.112513~0.006332~~**~\\
-d~~~~~~~~~~~~~0.29855~~~0.041945~~7.117605~~~~2e-06~***~\\
----~\\
-Signif.~codes:~~0~`***'~0.001~`**'~0.01~`*'~0.05~`.'~0.1~`~'~1~\\
-~\\
-Residual~standard~error:~1.96563~on~17~degrees~of~freedom~\\
-Number~of~observations:~20~Degrees~of~Freedom:~17~\\
-SSR:~65.682902~MSE:~3.8637~Root~MSE:~1.96563~\\
-Multiple~R-Squared:~0.755019~Adjusted~R-Squared:~0.726198~\\
-~\\
-~\\
-SUR~estimates~for~2~~(equation~2~)~\\
-Model~Formula:~q~~~p~+~f~+~a\\
-\\
-\mbox{}~~~~~~~~~~~~~Estimate~Std.~Error~~t~value~Pr(>|t|)~\\
-(Intercept)~61.966166~~~11.08079~5.592215~~~~4e-05~***~\\
-p~~~~~~~~~~~~0.146884~~~0.094435~1.555397~0.139408~\\
-f~~~~~~~~~~~~0.214004~~~0.039868~5.367761~~6.3e-05~***~\\
-a~~~~~~~~~~~~0.339304~~~0.067911~4.996283~0.000132~***~\\
----~\\
-Signif.~codes:~~0~`***'~0.001~`**'~0.01~`*'~0.05~`.'~0.1~`~'~1~\\
-~\\
-Residual~standard~error:~2.550226~on~16~degrees~of~freedom~\\
-Number~of~observations:~20~Degrees~of~Freedom:~16~\\
-SSR:~104.05843~MSE:~6.503652~Root~MSE:~2.550226~\\
-Multiple~R-Squared:~0.611888~Adjusted~R-Squared:~0.539117~\\
-}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{User options}
-
-Following additional options can be set by the user:
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% \subsubsection{Equation labels}
-\paragraph{Equation labels}
-The optional argument \code{eqnlabels} allows the user to label the equations.
-It has to be a vector of strings naming the equations.\\
-\code{
-R>~fitsur~<-~systemfit(~"SUR",~list(~q~~~p~+~d,~q~~~p~+~f~+~a~),\\~
-R+~~~~eqnlabels~=~c(~"demand",~"supply"~)~)\\
-R>~summary(~fitsur~)\\
-systemfit~results\\
-method:~SUR\\
-\\
-\mbox{}~~~~~~~~N~DF~~~~~~SSR~~~~~MSE~~~~RMSE~~~~~~~R2~~~Adj~R2\\
-demand~20~17~~65.6829~3.86370~1.96563~0.755019~0.726198\\
-supply~20~16~104.0584~6.50365~2.55023~0.611888~0.539117\\
-\ldots\\
-}
-If no equation labels are provided, the equations are numbered.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\subsubsection{Instrumental variables}   
-\paragraph{Instrumental variables}   
-\code{inst}one-sided model formula specifying instrumental variables
-   or a list of one-sided model formulas if different instruments should
-   be used for the different equations (only needed for 2SLS, W2SLS and
-   3SLS estimations).
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\subsubsection{Data}   
-\paragraph{Data}   
-\code{data} an optional data frame containing the variables in the model.
-   By default the variables are taken from the environment from which
-   systemfit is called.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\subsubsection{Restrictions}   
-\paragraph{Restrictions}   
-\code{R.restr} an optional j x k matrix to impose linear
-   restrictions on the parameters by \code{R.restr} * $\beta$ = \code{q.restr}
-   (j = number of restrictions, k = number of all parameters,
-   $\beta$ = vector of all parameters).
-
-\code{q.restr} an optional j x 1 matrix to impose linear
-   restrictions (see \code{R.restr}); default is a j x 1 matrix
-   that contains only zeros.
-
-\code{TX} an optional matrix to transform the regressor matrix and,
-   hence, also the coefficient vector (see details).
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\subsubsection{Iteration control}   
-\paragraph{Iteration control}
-\code{maxiter} maximum number of iterations for WLS, SUR, W2SLS and
-   3SLS estimations.
-
-\code{tol} tolerance level indicating when to stop the iteration (only
-   WLS, SUR, W2SLS and 3SLS estimations).
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\subsubsection{Residual covariance matrix}   
-\paragraph{Residual covariance matrix}   
-\code{rcovformula} formula to calculate the estimated residual covariance
-   matrix (see details).
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\subsubsection{3SLS formula}   
-\paragraph{3SLS formula}   
-\code{formula3sls} formula for calculating the 3SLS estimator,
-   one of "GLS", "IV", "GMM", "Schmidt" or "EViews" (see details).
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\subsubsection{Degrees of freedom for t-tests}   
-\paragraph{Degrees of freedom for t-tests}   
-\code{probdfsys} use the degrees of freedom of the whole system
-   (in place of the degrees of freedom of the single equation)
-   to calculate prob values for the t-test of individual parameters.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\subsubsection{Sigma squared}   
-\paragraph{Sigma squared}   
-\code{single.eq.sigma} use different $\sigma^2$s for each
-   single equation to calculate the covariance matrix and the
-   standard errors of the coefficients (only OLS and 2SLS).
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\subsubsection{System options}   
-\paragraph{System options}
-
-\code{solvetol} tolerance level for detecting linear dependencies
-   when inverting a matrix or calculating a determinant (see
-   \code{solve} and \code{det}).
-
-\code{saveMemory} save memory by omitting some calculation that
-   are not crucial for the basic estimation (e.g McElroy's
-   $R^2$).
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 % SUR fit and the 2SLS portion of the 3SLS fit are not included.  }
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Other tools}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Likelihood ratio test}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Hausman test}
-
-\textbf{Man, I really have to clean this up.}
-
-Hausman \citep{hausman1978} developed a test have power against
-alternatives for which $\hat{\theta}$ and $\tilde{\theta}$ diverge
-under misspecification.
-
-The Hausman result if the difference between the covariance of an
-efficient and the covariance of the inefficient estimator is zero, the
-model is correctly specified.
-
-The test is based on the idea that under the hypothesis of no
-correlation, both the limited information estimates and the
-full-information results are consistent, but only the full-information
-estimates are efficient and under the alternative, the limited
-information estimates are consistent, but the full information
-estimates are not.
-
-Since the models that are fit using the \pkg{systemfit} package are
-typically random effects models, the effects may be inconsistent
-between the introduced variables and the regressors.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Summary and outlook}\label{sec:Summmary}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-What should we place in this section?
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Unbalanced Observations}\label{sec:summary_unbalanced_data}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-The \pkg{systemfit} package was originally developed to fit
-simultaneous equations for forestry datasets. Forestry datasets
-typically contain observations for many inexpensive observations
-(diameter at breast height) and very few expensive observations such
-as total tree height. While the package currently does not allow for
-unbalanced datasets, data where the equations contain different
-numbers of observations, future releases of the package will have that
-feature implemented. We are still trying to figure out how develop the
-code for instrumental variable methods (2SLS,3SLS) for the correct
-estimation methods if possible.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Additional Methods}\label{sec:summary_additional_methods}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsubsection{Full Information Maximum Likelihood}
-
-Describe, in a paragraph the process and give some reasons why you
-might want this in the package.
-
-\subsubsection{Generalized Method of Moments}
-
-Describe, in a paragraph the process and give some reasons why you
-might want this in the package.
-
-\section{CVS Revision Log}
-\begin{verbatim}
-$Log$
-Revision 1.9  2004/11/16 10:59:09  henningsena
-some minor changes and some comments on nlsystemfit
-
-Revision 1.8  2004/11/16 08:55:53  henningsena
-merged file and resolved conflict due to some minor changes I did while Jeff was adding nlsystemfit.
-
-Revision 1.7  2004/11/16 05:34:47  hamannj
-question about nlsystemfit code section (example)
-
-Revision 1.6  2004/11/16 05:32:46  hamannj
-made small changes because my code section wouldn't compile. argh.
-
-Revision 1.5  2004/11/16 05:31:19  hamannj
-added cvs header and log section. to be removed when submitted.
-
-\end{verbatim}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\bibliography{systemfit} % a subset of my big bibtex file
-%\bibliography{/home/suapm095/Documents/Literatur/arne}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\end{document}

### 2005-09-01 paper/trunk/systemfit.tex
moved section about non-linear estimation from systemfit.tex into new file nlsystemfit.tex, because this section should not be included in (this version of) the article
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@570 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Estimating nonlinear equation systems}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-The \pkg{systemfit} package also contains a method for fitting
-nonlinear systems of equations.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Nonlinear parameter estimation}
-\label{sec:nonlinear_estimation}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-A system of nonlinear equations can be written as:
-
-\begin{equation}
-  \label{eq:non_linear_eq_1}
-  \epsilon_{t} = q( y_t, x_t, \theta )
-\end{equation}
-%%% It is not clear what these variable are. 
-%%% Adding an equation representing only a single equation
-%%% would make things clearer.
-
-\noindent and
-
-\begin{equation}
-  \label{eq:non_linear_eq_2}
-  z_{t} = Z( x_t )
-\end{equation}
-
-where $\epsilon_{t}$ are the residuals from the y observations and
-$z_{t}$ are the function evaluated at the parameter estimates.
-
-\textbf{you really need to check this!!!}
-
-The objective functions for the methods are:
-
-\begin{center}
-\begin{tabular}{|l|c|c|c|} \hline
-  Method & Instruments & Objective Function & Covariance of $\theta$ \\ \hline  
-  OLS & No & $r'r$ & $(X(diag(S)^{-1}\otimes I)X)^{-1}$ \\ \hline
-  SUR & No & $r'(diag(S)_{OLS}^{-1}\otimes I)r$ & $(X(S^{-1}\otimes I)X)^{-1}$ \\ \hline
-  2SLS & Yes & $r'(I \otimes W)r$ & $(X(diag(S)^{-1}\otimes I)X)^{-1}$ \\ \hline 
-  3SLS & Yes & $r'(S_{2SLS}^{-1} \otimes W)r$ & $(X(diag(S)^{-1}\otimes W)X)^{-1}$ \\ \hline
-\end{tabular}
-\end{center}
-
-where, $r$ is a column vector for the residuals for each equation,
-%% what is the difference to \epsilon?
-$X$ is matrix of the partial derivatives of the dependent variable 
-with respect to the parameters $\left( \frac{ \partial y }{ \theta} \right)$,
-%% is this correct?
-%% in the linear section X is used for the regressors, 
-%% thus we should use a different variable name here.
-$W$ is a matrix of the instrument variables, $Z(Z'Z)^{-1}Z$, $Z$ is a
-matrix of the instrument variables, and $I$ is an $n \times n $
-identity matrix and $S$ is the estimated variance-covariance matrix between the
-equations
-
-\begin{equation}
-  \label{eq:non-linear_varcov}
-  \hat{\sigma}_{ij} = (\hat{e}_i' \hat{e}_j) / \sqrt{(T - k_i)*(T- k_j)} 
-\end{equation}
-
-\textbf{The residual variance-covariance matrix can be calculated in 
-different ways (section~\ref{sec:residcov}). 
-It should be relatively easy to implement this also in nlsystemfit(), 
-e.g. the function 'calcRCov' in systemfit() could be moved outside
-systemfit() that it can be used also by nlsystemfit().}
-
-\textbf{You need to clear this up.}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Using nlsystemfit}\label{sec:UsingnlSystemfit}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\code {nlsystemfit} fits a set of structural nonlinear equations using
-Ordinary Least Squares (OLS), Seemingly Unrelated Regression (SUR),
-Two-Stage Least Squares (2SLS), Three-Stage Least Squares (3SLS) using
-the objective functions described in section
-\ref{sec:nonlinear_estimation}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Standard usage}
-
-Similar to calling the \code{systemfit} function, \code{nlsystemfit}
-is called with a minimum of three arguments,
-
-\code{
-R> nlsystemfit( method, eqns, start )
-}
-
-where \code{method} is one of the following estimation methods: "OLS",
-"SUR", "2SLS", or "3SLS", \code{eqns} is a list of equations similar
-to those described for \code{systemfit}, and \code{start} is a list of
-starting values of the parameter estimates.
-
-% The mandatory argument \code{eqns} is a list of the equations  to
-% estimate.  Each equation is a standard formula in \proglang{R}.  It
-% starts with a dependent variable on the left hand side.  After a tilde
-% ($~$) the regressors are listed  \footnote{For Details see the
-%   \proglang{R} help files to \code{formula}}.
-
-This is now demonstrated using an example: \\
-
-% \code{
-% R> library( systemfit ) \\
-% R> data( ppine ) \\
-% R> hg.formula <- hg $\sim$ exp( h0 + h1*log(tht) + h2*tht$\hat$2 + h3*elev + h4*cr) \\
-% R> dg.formula <- dg $\sim$ exp( d0 + d1*log(dbh) + d2*hg + d3*cr + d4*ba  ) \\
-% R> labels <- list( "height.growth", "diameter.growth" )\\
-% R> inst <- $\sim$ tht + dbh + elev + cr + ba\\
-% R> start.values <- c(h0=-0.5, h1=0.5, h2=-0.001, h3=0.0001, h4=0.08,\\
-% R+   d0=-0.5, d1=0.009, d2=0.25, d3=0.005, d4=-0.02 )\\
-% R> model <- list( hg.formula, dg.formula )\\
-% R> model.3sls <- nlsystemfit( "3SLS", model, start.values, data=ppine,\\
-% R+   eqnlabels=labels, inst=inst ) \\
-% R> summary( model.3sls ) \\
-% }
-
-The nlsystemfit function relies on \code{nlm} to perform the
-minimization of the objective functions and the \code{qr} set of
-functions. If the user does not have a set of estimates for the
-initial parameters, it is suggested using linearized forms in one of
-the linear methods, or simply using \code{nls} to obtain estimates for
-the equations.
-
-The outputs are similar to those for a systemfit object.
-
-\textbf{Describe the code and outputs as Arne did in the previous
-  section.}
-
-The first line loads the \pkg{systemfit} package. 
-The second line loads example data that are included in this package.
-These data come from \cite{kmenta86}.
-They are attached to the \proglang{R} search path in line three.
-In the last line a seemingly unrelated regression is done.
-The first equation represents the demand side of the food market.
-The dependant variable is \code{q} (food consumption per capita). 
-The regressors are \code{p} (ratio of food prices to general consumer prices)
-and \code{disposable income} as well as a constant%
-\footnote{a regression constant is always implied if not explicitly omitted.}.
-The second equation represents the supply side.
-Variable \code{q} (food consumption per capita) is also the dependant 
-variable of this equation. 
-The regressors are again \code{p} (ratio of food prices to general 
-consumer prices) and a constant as well as 
-\code{f} (ratio of preceding year's prices received by farmers) and 
-\code{a} (a time trend in years).
-The regression result is assigned to the variable \code{fitsur}.
-
-%% Arne -- Did you put the ~'s in by hand?
-%% Yes, I did. I used find and replace to substitute the "~" for " ". 
-%% This is necessary to retain the format of the output.
-
-Summary results can be printed by\\
-\code{
-R>~summary(~fitsur~)~\\
-~\\
-systemfit~results~\\
-method:~SUR~\\
-\\
-\mbox{}~~~N~DF~~~~~~SSR~~~~~MSE~~~~RMSE~~~~~~~R2~~~Adj~R2 \\
-1~20~17~~65.6829~3.86370~1.96563~0.755019~0.726198~\\
-2~20~16~104.0584~6.50365~2.55023~0.611888~0.539117~\\
-~\\
-The~covariance~matrix~of~the~residuals~used~for~estimation\\
-\mbox{}~~~~~~~~1~~~~~~~2~\\
-1~3.72539~4.13696~\\
-2~4.13696~5.78444~\\
-~\\
-The~covariance~matrix~of~the~residuals\\
-\mbox{}~~~~~~~~1~~~~~~~2~\\
-1~3.86370~4.92431~\\
-2~4.92431~6.50365~\\
-~\\
-The~correlations~of~the~residuals\\
-\mbox{}~~~~~~~~~1~~~~~~~~2~\\
-1~1.000000~0.982348~\\
-2~0.982348~1.000000~\\
-~\\
-The~determinant~of~the~residual~covariance~matrix:~0.879285~\\
-OLS~R-squared~value~of~the~system:~0.683453~\\
-McElroy's~R-squared~value~for~the~system:~0.788722~\\
-~\\
-SUR~estimates~for~1~~(equation~1~)~\\
-Model~Formula:~q~~~p~+~d\\
-\\
-\mbox{}~~~~~~~~~~~~~Estimate~Std.~Error~~~t~value~Pr(>|t|)~\\
-(Intercept)~99.332894~~~7.514452~13.218913~~~~~~~~0~***~\\
-p~~~~~~~~~~~-0.275486~~~0.088509~-3.112513~0.006332~~**~\\
-d~~~~~~~~~~~~~0.29855~~~0.041945~~7.117605~~~~2e-06~***~\\
----~\\
-Signif.~codes:~~0~`***'~0.001~`**'~0.01~`*'~0.05~`.'~0.1~`~'~1~\\
-~\\
-Residual~standard~error:~1.96563~on~17~degrees~of~freedom~\\
-Number~of~observations:~20~Degrees~of~Freedom:~17~\\
-SSR:~65.682902~MSE:~3.8637~Root~MSE:~1.96563~\\
-Multiple~R-Squared:~0.755019~Adjusted~R-Squared:~0.726198~\\
-~\\
-~\\
-SUR~estimates~for~2~~(equation~2~)~\\
-Model~Formula:~q~~~p~+~f~+~a\\
-\\
-\mbox{}~~~~~~~~~~~~~Estimate~Std.~Error~~t~value~Pr(>|t|)~\\
-(Intercept)~61.966166~~~11.08079~5.592215~~~~4e-05~***~\\
-p~~~~~~~~~~~~0.146884~~~0.094435~1.555397~0.139408~\\
-f~~~~~~~~~~~~0.214004~~~0.039868~5.367761~~6.3e-05~***~\\
-a~~~~~~~~~~~~0.339304~~~0.067911~4.996283~0.000132~***~\\
----~\\
-Signif.~codes:~~0~`***'~0.001~`**'~0.01~`*'~0.05~`.'~0.1~`~'~1~\\
-~\\
-Residual~standard~error:~2.550226~on~16~degrees~of~freedom~\\
-Number~of~observations:~20~Degrees~of~Freedom:~16~\\
-SSR:~104.05843~MSE:~6.503652~Root~MSE:~2.550226~\\
-Multiple~R-Squared:~0.611888~Adjusted~R-Squared:~0.539117~\\
-}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Other issues}
-
-\subsubsection{nlsystemfit issue 1}
-% The user should be aware that the function is \bold{VERY} sensative to
-% the starting values and the nlm function may not converge.
-
-% did you try optim()? Especially if you provide (analytical) derivatives
-% it is very good in finding the optimum.
-
-% The nlm
-% function will be called with the \code{typsize} argument set the
-% absolute values of the starting values for the OLS and 2SLS
-% methods. For the SUR and 3SLS methods, the \code{typsize} argument is
-% set to the absolute values of the resulting OLS and 2SLS parameter
-% estimates from the nlm result structre. In addition, the starting
-% values for the SUR and 3SLS methods are obtained from the OLS and 2SLS
-% parameter estimates to shorten the number of iterations. The number of
-% iterations reported in the summary are only those used in the last
-% call to nlm, thus the number of iterations in the OLS portion of the
-% SUR fit and the 2SLS portion of the 3SLS fit are not included.  }
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 $Log$
+Revision 1.10  2005/09/01 11:49:02  henningsena
+moved section about non-linear estimation from systemfit.tex into new file nlsystemfit.tex, because this section should not be included in (this version of) the article
+
 Revision 1.9  2004/11/16 10:59:09  henningsena

### 2005-09-01 paper/trunk/systemfit.tex
revised structure of the document (sections, subsections, subsubsections) due to removal of nlsystemfit; moved each section into a separate file.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@571 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Introduction}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-Many theoretical models consist of more than one equation. If these
-models are econometrically estimated, the equations generally should
-not be estimated individually, but all together, because the
-disturbance terms of these equations are likely contemporaneously
-correlated. Estimating all equations simultaneously and taking the
-covariance structure of the residuals into account leads to efficient
-estimates, while ignoring this correlation leads to an inefficient
-parameter estimation \citep{zellner62}.
-
-Another reason to estimate an equation system simultaneously are
-cross-equation parameter restrictions. These restrictions can be
-tested and/or imposed only in a simultaneous estimation
-approach. Especially the economic theory suggests many cross-equation
-restriction.
-
-The \pkg{systemfit} package provides the capability to estimate
-linear and non-linear equation systems in \proglang{R}. Compared to
-other statistics and econometrics software the (advanced) user can
-control many important details of the estimation procedure. On the
-other hand these user options have reasonable defaults that also
-beginners can easily use the provided functions.
-
-\textbf{I'll have to update this section to match the recent changes. JDH.}
-
-This paper is organized as follows: In section~\ref{sec:Theory} we
-introduce the mathematics of estimating equation systems and give a
-comprehensive overview over all relevant issues regarding this topic.
-Section~\ref{sec:UsingSystemfit} demonstrates how to run
-\pkg{systemfit}, especially how the features presented in the previous
-chapter can be used. Finally, a summary and an outlook are given in
-section~\ref{sec:Summmary}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Estimating linear equation systems}\label{sec:Theory}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-Consider a system of $G$ equations, where the $i$th equation is of
-the form 
-\begin{equation}
-   y_{i} = X_i \beta_i + u_i, \quad i = 1, 2, \ldots, G
-\end{equation}
-where $y_i$ is a vector of the dependent variable,
-$X_i$ is a matrix of the exogenous variables,
-$\beta_i$ is the coefficient vector and
-$u_i$ is a vector of the disturbance terms of the $i$th equation.
-
-We can write the 'stacked' system as
-\begin{equation}
-   \left[ \begin{array}{c}
-      y_1 \\ y_2\\ \vdots\\ y_G
-   \end{array} \right] = 
-   \left[ \begin{array}{cccc}
-      X_1 & 0 & \cdots & 0\\
-      0 & X_2 & \cdots & 0\\
-      \vdots & \vdots & \ddots & \vdots\\
-      0 & 0 & \cdots & X_G
-   \end{array}\right] \cdot 
-   \left[ \begin{array}{c}
-      \beta_1 \\ \beta_2 \\ \vdots\\ \beta_G
-   \end{array} \right] +
-   \left[ \begin{array}{c}
-      u_1 \\ u_2 \\ \vdots\\ u_G
-   \end{array} \right]
-\end{equation}
-
-or more simply as
-\begin{equation}
-   y = X \beta + u 
-\end{equation}
-
-We assume that there is no correlation of the disturbance terms 
-across observations:
-\begin{equation}
-   E \left( u_{it} \, u_{jt^*} \right) = 0 
-   \; \forall \; t \neq t^*
-\end{equation}
-where $i$ and $j$ indicate the equation number 
-and $t$ and $t^*$ denote the observation number.
-
-However, we explicitly allow for contemporaneous correlation:
-\begin{equation}
-   E \left( u_{it} \, u_{jt} \right) = \sigma_{ij}
-\end{equation}
-
-Thus, the covariance matrix of the total system is
-\begin{equation}
-   E \left( u \, u' \right) = \Omega = \Sigma \otimes I
-\end{equation}
-where $\Sigma = \left[ \sigma_{ij} \right]$ and $I$ is an
-identity matrix.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Linear Parameter estimation}
-
-\subsubsection{Ordinary Least Squares estimation}
-
-The Ordinary Least Squares (OLS) estimator of the system 
-is obtained by
-\begin{equation}
-   \widehat{\beta} = \left( X'X \right)^{-1} X'y
-\end{equation}
-
-These estimates are efficient only if the disturbance terms are not 
-contemporaneously correlated, which means 
-\begin{equation}
-   \sigma_{ij} = 0 \; \forall \; i \neq j
-\end{equation}
-
-If the whole system is treated as one single equation, 
-the variance-covariance matrix of the estimated parameters is
-\begin{equation}
-   Var \left[ \widehat{\beta} \right] = \sigma^2 \left( X'X \right)^{-1}
-\end{equation}
-with $\sigma^2 = E \left( u' u \right)$.
-This assumes that the disturbances of all equations have the
-same variance.
-
-If the disturbance terms of the individual equations 
-are allowed to have different variances, 
-the variance-covariance matrix of the estimated parameters is
-\begin{equation}
-   Var \left[ \widehat{\beta} \right] = \left( X' \Omega^{-1} X \right)^{-1}
-   \label{eq:OLSvcm2}
-\end{equation}
-with $\Omega = \Sigma \otimes I$, 
-$\sigma_{ij} = 0 \; \forall \; i \neq j$ and
-$\sigma_{ii} = E \left( u_i' u_i \right)$.
-
-If no cross-equation restrictions are imposed, the simultaneous 
-OLS estimation of the system leads to the same parameter estimates 
-as an equation-wise OLS estimation. The variance-covariance matrix 
-of the parameters from an equation-wise OLS estimation is equal to 
-the variance-covariance matrix obtained by equation 
-(\ref{eq:OLSvcm2}).
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Weighted least squares}
-
-The Weighted Least Squares (WLS) estimator of the system 
-is obtained by
-\begin{equation}
-   \widehat{\beta} = \left( X' \Omega^{-1} X \right)^{-1} X' \Omega^{-1} y
-\end{equation}
-with $\Omega = \Sigma \otimes I$, 
-$\sigma_{ij} = 0 \; \forall \; i \neq j$ and
-$\sigma_{ii} = E \left( u_i' u_i \right)$.
-
-Like the OLS estimates these estimates are only efficient 
-if the disturbance terms are not contemporaneously correlated.
-
-The variance-covariance matrix of the estimated parameters is
-\begin{equation}
-   Var \left[ \widehat{\beta} \right] = \left( X' \Omega^{-1} X \right)^{-1}
-\end{equation}
-
-If no cross-equation restrictions are imposed, 
-the parameter estimates and the variance-covariance matrix 
-of the parameters are equal to the OLS estimates%
-\footnote{The variance-covariance matrices of the 
-estimated parameters might differ, 
-if the WLS estimation is based on a different formula 
-to calculate the (diagonal) variance-covariance 
-matrix of the residuals (see section \ref{sec:residcov}).}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Seemingly unrelated regression}
-
-When the disturbances are contemporaneously correlated, a Generalized 
-Least Squares (GLS) estimation leads to efficient parameter estimates.
-In this case the GLS is generally called "Seemingly Unrelated Regression" 
-(SUR) (see \citealp{zellner62}). 
-Note, while an unbiased OLS or WLS estimation requires only that 
-the regressors and the disturbance terms of each single 
-equation are uncorrelated,
-a consistent SUR estimation requires that all disturbance terms and all 
-regressors are uncorrelated:
-\begin{equation}
-   E \left( u | X \right) = 0
-\end{equation}
-
-The SUR estimator can be obtained by:%
-\footnote{To calculate $\Omega^{-1}$ it is not necessary to invert 
-the (huge) $\Omega$ matrix, 
-because it is sufficient to invert the (small) $\Sigma$ matrix and
-calculate it by $\Omega^{-1} = \Sigma^{-1} \otimes I$.}
-
-\begin{equation}
-   \widehat{\beta} = \left( X' \Omega^{-1} X \right)^{-1} X' \Omega^{-1} y
-\end{equation}
-with $\Omega = \Sigma \otimes I$ and
-$\sigma_{ij} = E \left( u_i' u_j \right)$.
-
-And the variance-covariance matrix of the estimated parameters is
-\begin{equation}
-   Var \left[ \widehat{\beta} \right] = \left( X' \Omega^{-1} X \right)^{-1}
-\end{equation}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Two-stage least squares estimation}
-
-If the regressors of one or more equations are correlated 
-with the disturbances ($E \left( u_i | X_i \right) \neq 0$), 
-the estimated coefficients are biased.
-This can be circumvented by an instrumental variable (IV) estimation.
-The instrumental variables for each equation $H_i$ 
-can be either different or identical for all equations.
-The instrumental variables of each equation may not be correlated with 
-the disturbance terms of the corresponding equation 
-($E \left( u_i | H_i \right) = 0$).
-
-At the first stage new ('fitted') regressors are obtained by
-\begin{equation}
-   \widehat{X_i} = H_i \left( H_i' H_i \right)^{-1} H_i' X
-\end{equation}
-
-At the second stage the unbiased two-stage least squares estimates
-of $\beta$ are obtained by:
-\begin{equation}
-   \widehat{\beta} = \left( \widehat{X}' \widehat{X} \right)^{-1} 
-   \widehat{X}' y 
-   \label{eq:beta2sls}
-\end{equation}
-
-If the whole system is treated as one single equation, 
-the variance-covariance matrix of the estimated parameters is
-\begin{equation}
-   Var \left[ \widehat{\beta} \right] = \sigma^2 \left( \widehat{X}'
-   \widehat{X} \right)^{-1}
-\end{equation}
-with $\sigma^2 = E \left( u' u \right)$.
-
-If the disturbance terms of the individual equations 
-are allowed to have different variances, 
-the variance-covariance matrix of the estimated parameters is
-\begin{equation}
-   Var \left[ \widehat{\beta} \right] = \left( \widehat{X}' \Omega^{-1} 
-   \widehat{X} \right)^{-1}
-\end{equation}
-with $\Omega = \Sigma \otimes I$, 
-$\sigma_{ij} = 0 \; \forall \; i \neq j$ and
-$\sigma_{ii} = E \left( u_i' u_i \right)$.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Weighted two-stage least squares estimation}
-
-The Weighted Two-Stage Least Squares (W2SLS) estimator of the system 
-is obtained by
-\begin{equation}
-   \widehat{\beta} = \left( \widehat{X}' \Omega^{-1} \widehat{X} 
-   \right)^{-1} \widehat{X}' \Omega^{-1} y
-\end{equation}
-with $\Omega = \Sigma \otimes I$, 
-$\sigma_{ij} = 0 \; \forall \; i \neq j$ and
-$\sigma_{ii} = E \left( u_i' u_i \right)$.
-
-The variance-covariance matrix of the estimated parameters is
-\begin{equation}
-   Var \left[ \widehat{\beta} \right] = \left( \widehat{X}' \Omega^{-1} 
-   \widehat{X} \right)^{-1}
-\end{equation}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Three-stage least squares estimation}
-
-If the regressors are correlated with the disturbances 
-($E \left( u | X \right) \neq 0$) and 
-the disturbances are contemporaneously correlated, 
-an Instrumental Variable (IV) Generalized Least Squares (GLS) 
-estimation leads to consistent and efficient estimates. This estimation 
-procedure is generally called "Three-stage Least Squares" (3SLS).
-
-The standard 3SLS estimator can be obtained by:
-\begin{equation}
-   \widehat{\beta} = \left( \widehat{X}' \Omega^{-1} \widehat{X} 
-   \right)^{-1} \widehat{X}' \Omega^{-1} y
-\end{equation}
-with $\Omega = \Sigma \otimes I$ and
-$\sigma_{ij} = E \left( u_i' u_j \right)$.
-
-Its variance-covariance matrix is:
-\begin{equation}
-   Var \left[ \widehat{\beta} \right] = \left( \widehat{X}' \Omega^{-1} 
-   \widehat{X} \right)^{-1}
-   \label{eq:cov3sls}
-\end{equation}
-
-While an unbiased 2SLS or W2SLS estimation requires only that 
-the instrumental variables and the disturbance terms of each single 
-equation are uncorrelated,
-\cite{schmidt90} points out that this estimator is only consistent 
-if all disturbance terms and all instrumental variables are uncorrelated:
-\begin{equation}
-   E \left( u | H \right) = 0
-\end{equation}
-with
-\begin{equation}
-   H =
-   \left[ \begin{array}{cccc}
-      H_1 & 0 & \cdots & 0\\
-      0 & H_2 & \cdots & 0\\
-      \vdots & \vdots & \ddots & \vdots\\
-      0 & 0 & \cdots & H_G
-   \end{array}\right] \cdot 
-\end{equation}
-
-Since there might be occasions where this cannot be avoided, 
-\cite{schmidt90} analyses other approaches to obtain 3SLS estimators:
-
-One of these approaches is based on instrumental variable estimation
-(IV-3SLS):
-\begin{equation}
-   \widehat{\beta} = \left( \widehat{X}' \Omega^{-1} X 
-   \right)^{-1} \widehat{X}' \Omega^{-1} y
-\end{equation}
-
-The variance-covariance matrix of this IV-3SLS estimator is:
-\begin{equation}
-   Var \left[ \widehat{\beta} \right] = \left( \widehat{X}' \Omega^{-1} 
-   X \right)^{-1}
-\end{equation}
-
-Another approach is based on the Generalized Method of Moments (GMM) 
-estimator (GMM-3SLS):
-\begin{equation}
-   \widehat{\beta} = \left( X' H \left( H' \Omega H \right)^{-1}
-   H' X \right)^{-1} X' H \left( H' \Omega H \right)^{-1} H' y
-   \label{eq:beta3slsGMM}
-\end{equation}
-
-The variance-covariance matrix of the GMM-3SLS estimator is:
-\begin{equation}
-   Var \left[ \widehat{\beta} \right] = 
-   \left( X' H \left( H' \Omega H \right)^{-1} H' X \right)^{-1}
-\end{equation}
-
-A fourth approach developed by \cite{schmidt90} himself is:
-\begin{equation}
-   \widehat{\beta} = \left( \widehat{X}' \Omega^{-1} \widehat{X} 
-   \right)^{-1} \widehat{X}' \Omega^{-1} 
-   H \left( H' H \right)^{-1} H' y 
-   \label{eq:beta3slsSchmidt}
-\end{equation}
-
-The variance-covariance matrix of this estimator is:
-\begin{equation}
-   Var \left[ \widehat{\beta} \right] = 
-   \left( \widehat{X}' \Omega^{-1}  \widehat{X} \right)^{-1} 
-   \widehat{X}' \Omega^{-1} H \left( H' H \right)^{-1} H' \Omega 
-   H \left( H' H \right)^{-1} H' \Omega^{-1} \widehat{X}
-   \left( \widehat{X}' \Omega^{-1}  \widehat{X} \right)^{-1}
-\end{equation}
-
-The econometrics software EViews uses following approach:
-\begin{equation}
-   \widehat{\beta} = \widehat{\beta}_{2SLS} + 
-   \left( \widehat{X}' \Omega^{-1} \widehat{X} \right)^{-1} 
-   \widehat{X}' \Omega^{-1} \left( y - X \widehat{\beta}_{2SLS} \right)
-\end{equation}
-with $\widehat{\beta}_{2SLS}$ is the two-stage least squares estimator
-(equation \ref{eq:beta2sls}).
-EViews uses the standard 3SLS formula (equation (\ref{eq:cov3sls})) to 
-calculate the variance-covariance matrix of the 3SLS estimator.
-
-If the same instrumental variables are used in all equations 
-($H_1 = H_2 = \ldots = H_G$), 
-all the above mentioned approaches lead to identical parameter estimates.
-However, if this is not the case, the results depend on the 
-equation used \citep{schmidt90}. 
-The only reason to use different instruments in the different equations 
-is that the instruments of one equation are correlated with the
-disturbance terms in another equation. Otherwise, one would simply use
-every instrument in every equation \citep{schmidt90}. 
-In this case, only the GMM-3SLS (equation (\ref{eq:beta3slsGMM}))
-and the 3SLS estimator developed by \cite{schmidt90} 
-(equation (\ref{eq:beta3slsSchmidt})) are consistent.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Imposing linear restrictions}
-There are two ways to impose linear parameter restrictions.
-First, a matrix $T$ can be provided with
-\begin{equation}
-   \beta = T \cdot \beta^* \label{eq:T-restr} 
-\end{equation}
-where $\beta^*$ is a vector of linear independent coefficients.
-
-To impose these restrictions the $X$ matrix is
-(post-)multiplied by this $T$ matrix.
-\begin{equation}
-    X^* = X \cdot T
-\end{equation}
-
-Then, a standard estimation is done by substituting $X^*$ for $X$.
-This results in the linear independent parameter estimates $\beta^*$ and
-the variance-covariance matrix of these parameters. 
-The original parameters can be obtained by equation (\ref{eq:T-restr})
-and the variance-covariance matrix of the original parameters 
-can be obtained by:
-\begin{equation}
-   Var \left[ \widehat{\beta} \right] = T \cdot Var \left[ \widehat{\beta^*} \right] \cdot T'
-\end{equation}
-
-The second way to impose linear parameter restrictions 
-can be formulated by
-\begin{equation}
-   R \beta^0 = q
-\end{equation}
-where $\beta^0$ is the vector of the restricted coefficients, 
-$R$ is a matrix and $q$ is a vector (see \citealp[p. 100]{greene02}). 
-Each restriction belongs to one row of $R$ and the corresponding 
-element of~$q$.
-
-The first way is less flexible than this latter one%
-\footnote{e.g. restrictions like $\beta_1 + \beta_2 = 4$ cannot be imposed
-by the first method}, 
-but the first way is preferable if equality constraints for coefficients
-across many equations of the system are imposed. 
-Of course, these restrictions can be also imposed using
-the latter method.
-However, while the latter method increases the dimension of the 
-matrices to be inverted during estimation, the first reduces it. 
-Thus, in some cases the latter way leads to estimation problems
-(e.g. (near) singularity of the matrices to be inverted),
-while the first doesn't.
-
-These two methods can be combined. In this case the restrictions
-imposed using the latter method are imposed on the linear independent 
-parameters due to the restrictions imposed using the first method:
-\begin{equation}
-   R \beta^{*0} = q
-\end{equation}
-where $\beta^{*0}$ is the vector of the restricted $\beta^*$ coefficients.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Restricted OLS estimation}
-
-The OLS estimator restricted by $R \beta^0 = q$ can be obtained by
-\begin{equation}
-   \left[ \begin{array}{c}
-      \widehat{\beta^0} \\ \widehat{\lambda}
-   \end{array} \right]
-   =
-   \left[ \begin{array}{cc}
-      X' X & R' \\ 
-      R & 0
-   \end{array} \right]^{-1}
-   \cdot
-   \left[ \begin{array}{c}
-      X' y \\ q 
-   \end{array} \right]
-\end{equation}
-where $2\lambda$ is a vector of the Lagrangean multipliers of the restrictions.
-
-If the whole system is treated as one single equation, 
-the variance-covariance matrix of the estimated parameters is
-\begin{equation}
-   Var 
-   \left[ \begin{array}{c}
-      \widehat{\beta} \\ \widehat{\lambda}
-   \end{array} \right] 
-   = \sigma^2 
-   \left[ \begin{array}{cc}
-      X' X & R' \\ 
-      R & 0
-   \end{array} \right]^{-1}
-\end{equation}
-with $\sigma^2 = E \left( u' u \right)$.
-
-If the disturbance terms of the individual equations 
-are allowed to have different variances, 
-the variance-covariance matrix of the estimated parameters is
-\begin{equation}
-   Var 
-   \left[ \begin{array}{c}
-      \widehat{\beta} \\ \widehat{\lambda}
-   \end{array} \right] 
-   = 
-   \left[ \begin{array}{cc}
-      X' \Omega^{-1} X & R' \\ 
-      R & 0
-   \end{array} \right]^{-1}
-\end{equation}
-with $\Omega = \Sigma \otimes I$, 
-$\sigma_{ij} = 0 \; \forall \; i \neq j$ and
-$\sigma_{ii} = E \left( u_i' u_i \right)$.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Restricted WLS estimation}
-
-The WLS estimator restricted by $R \beta^0 = q$ can be obtained by
-\begin{equation}
-   \left[ \begin{array}{c}
-      \widehat{\beta^0} \\ \widehat{\lambda}
-   \end{array} \right]
-   =
-   \left[ \begin{array}{cc}
-      X' \Omega^{-1} X & R' \\ 
-      R & 0
-   \end{array} \right]^{-1}
-   \cdot
-   \left[ \begin{array}{c}
-      X' \Omega^{-1} y \\ q 
-   \end{array} \right]
-\end{equation}
-with $\Omega = \Sigma \otimes I$, 
-$\sigma_{ij} = 0 \; \forall \; i \neq j$ and
-$\sigma_{ii} = E \left( u_i' u_i \right)$.
-
-The variance-covariance matrix of the estimated parameters is
-\begin{equation}
-   Var 
-   \left[ \begin{array}{c}
-      \widehat{\beta} \\ \widehat{\lambda}
-   \end{array} \right] 
-   = 
-   \left[ \begin{array}{cc}
-      X' \Omega^{-1} X & R' \\ 
-      R & 0
-   \end{array} \right]^{-1}
-\end{equation}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Restricted SUR estimation}
-
-The SUR estimator restricted by $R \beta^0 = q$ can be obtained by
-\begin{equation}
-   \left[ \begin{array}{c}
-      \widehat{\beta^0} \\ \widehat{\lambda}
-   \end{array} \right]
-   =
-   \left[ \begin{array}{cc}
-      X' \Omega^{-1} X & R' \\ 
-      R & 0
-   \end{array} \right]^{-1}
-   \cdot
-   \left[ \begin{array}{c}
-      X' \Omega^{-1} y \\ q 
-   \end{array} \right]
-\end{equation}
-with $\Omega = \Sigma \otimes I$ and
-$\sigma_{ij} = E \left( u_i' u_j \right)$.
-
-The variance-covariance matrix of the estimated parameters is
-\begin{equation}
-   Var 
-   \left[ \begin{array}{c}
-      \widehat{\beta} \\ \widehat{\lambda}
-   \end{array} \right] 
-   = 
-   \left[ \begin{array}{cc}
-      X' \Omega^{-1} X & R' \\ 
-      R & 0
-   \end{array} \right]^{-1}
-\end{equation}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Restricted 2SLS estimation}
-
-The 2SLS estimator restricted by $R \beta^0 = q$ can be obtained by
-\begin{equation}
-   \left[ \begin{array}{c}
-      \widehat{\beta^0} \\ \widehat{\lambda}
-   \end{array} \right]
-   =
-   \left[ \begin{array}{cc}
-      \widehat{X}' \widehat{X} & R' \\ 
-      R & 0
-   \end{array} \right]^{-1}
-   \cdot
-   \left[ \begin{array}{c}
-      \widehat{X}' y \\ q 
-   \end{array} \right]
-   \label{eq:beta2SLSr}
-\end{equation}
-
-If the whole system is treated as one single equation, 
-the variance-covariance matrix of the estimated parameters is
-\begin{equation}
-   Var 
-   \left[ \begin{array}{c}
-      \widehat{\beta} \\ \widehat{\lambda}
-   \end{array} \right] 
-   = \sigma^2 
-   \left[ \begin{array}{cc}
-      \widehat{X}' \widehat{X} & R' \\ 
-      R & 0
-   \end{array} \right]^{-1}
-\end{equation}
-with $\sigma^2 = E \left( u' u \right)$.
-
-If the disturbance terms of the individual equations 
-are allowed to have different variances, 
-the variance-covariance matrix of the estimated parameters is
-\begin{equation}
-   Var 
-   \left[ \begin{array}{c}
-      \widehat{\beta} \\ \widehat{\lambda}
-   \end{array} \right] 
-   = 
-   \left[ \begin{array}{cc}
-      \widehat{X}' \Omega^{-1} \widehat{X} & R' \\ 
-      R & 0
-   \end{array} \right]^{-1}
-\end{equation}
-with $\Omega = \Sigma \otimes I$, 
-$\sigma_{ij} = 0 \; \forall \; i \neq j$ and
-$\sigma_{ii} = E \left( u_i' u_i \right)$.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Restricted W2SLS estimation}
-
-The W2SLS estimator restricted by $R \beta^0 = q$ can be obtained by
-\begin{equation}
-   \left[ \begin{array}{c}
-      \widehat{\beta^0} \\ \widehat{\lambda}
-   \end{array} \right]
-   =
-   \left[ \begin{array}{cc}
-      \widehat{X}' \Omega^{-1} \widehat{X} & R' \\ 
-      R & 0
-   \end{array} \right]^{-1}
-   \cdot
-   \left[ \begin{array}{c}
-      \widehat{X}' \Omega^{-1} y \\ q 
-   \end{array} \right]
-\end{equation}
-with $\Omega = \Sigma \otimes I$, 
-$\sigma_{ij} = 0 \; \forall \; i \neq j$ and
-$\sigma_{ii} = E \left( u_i' u_i \right)$.
-
-The variance-covariance matrix of the estimated parameters is
-\begin{equation}
-   Var 
-   \left[ \begin{array}{c}
-      \widehat{\beta} \\ \widehat{\lambda}
-   \end{array} \right] 
-   = 
-   \left[ \begin{array}{cc}
-      \widehat{X}' \Omega^{-1} \widehat{X} & R' \\ 
-      R & 0
-   \end{array} \right]^{-1}
-\end{equation}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Restricted 3SLS estimation}
-
-The standard 3SLS estimator restricted by $R \beta^0 = q$ can be obtained by
-\begin{equation}
-   \left[ \begin{array}{c}
-      \widehat{\beta^0} \\ \widehat{\lambda}
-   \end{array} \right]
-   =
-   \left[ \begin{array}{cc}
-      \widehat{X}' \Omega^{-1} \widehat{X} & R' \\ 
-      R & 0
-   \end{array} \right]^{-1}
-   \cdot
-   \left[ \begin{array}{c}
-      \widehat{X}' \Omega^{-1} y \\ q 
-   \end{array} \right]
-\end{equation}
-with $\Omega = \Sigma \otimes I$ and
-$\sigma_{ij} = E \left( u_i' u_j \right)$.
-
-The variance-covariance matrix of this estimator is
-\begin{equation}
-   Var 
-   \left[ \begin{array}{c}
-      \widehat{\beta^0} \\ \widehat{\lambda}
-   \end{array} \right] 
-   = 
-   \left[ \begin{array}{cc}
-      \widehat{X}' \Omega^{-1} \widehat{X} & R' \\ 
-      R & 0
-   \end{array} \right]^{-1}
-   \label{eq:cov3slsr}
-\end{equation}
-
-The IV-3SLS estimator restricted by $R \beta^0 = q$ can be obtained by
-\begin{equation}
-   \left[ \begin{array}{c}
-      \widehat{\beta^0} \\ \widehat{\lambda}
-   \end{array} \right]
-   =
-   \left[ \begin{array}{cc}
-      \widehat{X}' \Omega^{-1} X & R' \\ 
-      R & 0
-   \end{array} \right]^{-1}
-   \cdot
-   \left[ \begin{array}{c}
-      \widehat{X}' \Omega^{-1} y \\ q 
-   \end{array} \right]
-\end{equation}
-
-with
-\begin{equation}
-   Var 
-   \left[ \begin{array}{c}
-      \widehat{\beta^0} \\ \widehat{\lambda}
-   \end{array} \right] 
-   = 
-   \left[ \begin{array}{cc}
-      \widehat{X}' \Omega^{-1} \widehat{X} & R' \\ 
-      R & 0
-   \end{array} \right]^{-1}
-\end{equation}
-
-The restricted 3SLS-GMM estimator can be obtained by
-\begin{equation}
-   \left[ \begin{array}{c}
-      \widehat{\beta^0} \\ \widehat{\lambda}
-   \end{array} \right]
-   =
-   \left[ \begin{array}{cc}
-      X' H \left( H' \Omega H \right)^{-1} H' X & R' \\ 
-      R & 0
-   \end{array} \right]^{-1}
-   \cdot
-   \left[ \begin{array}{c}
-      X' H \left( H \Omega H \right)^{-1} H' y \\ q 
-   \end{array} \right]
-\end{equation}
-
-with
-\begin{equation}
-   Var 
-   \left[ \begin{array}{c}
-      \widehat{\beta^0} \\ \widehat{\lambda}
-   \end{array} \right] 
-   = 
-   \left[ \begin{array}{cc}
-      X' H \left( H' \Omega H \right)^{-1} H' X & R' \\ 
-      R & 0
-   \end{array} \right]^{-1}
-\end{equation}
-
-The restricted 3SLS estimator based on the suggestion of 
-\cite{schmidt90} is:
-\begin{equation}
-   \left[ \begin{array}{c}
-      \widehat{\beta^0} \\ \widehat{\lambda}
-   \end{array} \right]
-   =
-   \left[ \begin{array}{cc}
-      \widehat{X}' \Omega^{-1} \widehat{X} & R' \\ 
-      R & 0
-   \end{array} \right]^{-1}
-   \cdot
-   \left[ \begin{array}{c}
-      \widehat{X}' \Omega^{-1} H \left( H' H \right)^{-1} H' y \\ q 
-   \end{array} \right]
-\end{equation}
-
-with
-\begin{eqnarray}
-   Var 
-   \left[ \begin{array}{c}
-      \widehat{\beta^0} \\ \widehat{\lambda}
-   \end{array} \right] 
-   & = & 
-   \left[ \begin{array}{cc}
-      \widehat{X}' \Omega^{-1} \widehat{X} & R' \\ 
-      R & 0
-   \end{array} \right]^{-1}
-   \cdot
-   \left[ \begin{array}{cc}
-      \widehat{X}' \Omega^{-1} H \left( H' H \right)^{-1} H' \Omega
-      H \left( H' H \right)^{-1} H' \Omega^{-1} \widehat{X} & 0' \\ 
-      0 & 0
-   \end{array} \right]^{-1}
-   \nonumber \\
-   & & \cdot
-   \left[ \begin{array}{cc}
-      \widehat{X}' \Omega^{-1} \widehat{X} & R' \\ 
-      R & 0
-   \end{array} \right]^{-1}
-\end{eqnarray}
-
-The econometrics software EViews calculates the restricted 3SLS estimator by:
-\begin{equation}
-   \left[ \begin{array}{c}
-      \widehat{\beta^0} \\ \widehat{\lambda}
-   \end{array} \right]
-   =
-   \left[ \begin{array}{cc}
-      \widehat{X}' \Omega^{-1} \widehat{X} & R' \\ 
-      R & 0
-   \end{array} \right]^{-1}
-   \cdot
-   \left[ \begin{array}{c}
-      \widehat{X}' \Omega^{-1} \left( y - X \widehat{\beta^0}_{2SLS} \right) 
-      \\ q 
-   \end{array} \right]
-\end{equation}
-where $\widehat{\beta^0}_{2SLS}$ is the restricted 2SLS estimator calculated
-by equation (\ref{eq:beta2SLSr}). 
-To calculate the variance-covariance matrix 
-EViews uses the standard formula of the restricted 3SLS estimator
-(equation (\ref{eq:cov3slsr})).
-
-If the same instrumental variables are used in all equations 
-($H_1 = H_2 = \ldots = H_G$), 
-all the above mentioned approaches lead to identical parameter estimates
-and identical variance-covariance matrices of the estimated parameters.
-
-(***** This has to be checked!!!!! **************)
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Other issues}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Residual variance-covariance matrix}\label{sec:residcov}
-
-Since the true residuals of the estimated equations are generally not known,
-the true covariance matrix of the residuals cannot be determined.
-Thus, this covariance matrix must be calculated from the
-\emph{estimated} residuals.
-
-The estimated covariance matrix of the residuals ($\widehat{\Sigma} = 
-\left[ \widehat{\sigma}_{ij} \right]$) 
-is often calculated by
-\begin{equation}
-   \widehat{\sigma}_{ij} = \frac{ \widehat{u}_i' \widehat{u}_j }{ T }
-   \label{eq:rcov0}
-\end{equation}
-where $T$ is the number of observations in each equation.
-
-However, this estimator is biased, because it is not corrected for 
-degrees of freedom. 
-The usual single-equation procedure to correct for degrees of freedom 
-cannot always be applied, because the number of regressors in each equation 
-might differ.
-Two alternative approaches to calculate the residual variance-covariance 
-matrix are
-\begin{equation}
-   \widehat{\sigma}_{ij} = \frac{ \widehat{u}_i' \widehat{u}_j }
-   { \sqrt{ \left( T - K_i \right) \cdot \left( T - K_j \right) } }
-   \label{eq:rcov1}
-\end{equation}
-and
-\begin{equation}
-   \widehat{\sigma}_{ij} = \frac{ \widehat{u}_i' \widehat{u}_j }
-   { T - \max \left( K_i , K_j \right) }
-\end{equation}
-where $K_i$ and $K_j$ are the number of regressors in equation
-$i$ and $j$, respectively.
-
-However, this yields an unbiased estimator only if $K_i = K_j$ 
-\citep[p. 469]{judge85}. 
-% Greene (2002, p. 344) says that the second is unbiased if i = j or K_i = K_j,
-% whereas the first is unbiased only if i = j. 
-% However, if K_i = K_j the first and the second are equal.
-% Why is the first biased if K_i = K_j ???????????
-
-A further approach to obtain the estimated residual variance-covariance
-matrix is \citep[p. 309]{zellner62c}
-\begin{eqnarray}
-   \widehat{\sigma}_{ij} & = & 
-   \frac{ \widehat{u}_i' \widehat{u}_j } 
-   { T - K_i - K_j + tr \left[ X_i \left( X_i' X_i \right)^{-1}
-   X_i' X_j \left( X_j' X_j \right)^{-1} X_j' \right] } \\
-   & = &
-   \frac{ \widehat{u}_i' \widehat{u}_j } 
-   { T - K_i - K_j + tr \left[ \left( X_i' X_i \right)^{-1}
-   X_i' X_j \left( X_j' X_j \right)^{-1} X_j' X_i \right] } \\
-   \label{eq:rcov2}
-\end{eqnarray} 
-This yields an unbiased estimator for all elements of $\widehat{\Sigma}$,
-but even if $\widehat{\Sigma}$ is an unbiased estimator of $\Sigma$, 
-its inverse $\widehat{\Sigma}^{-1}$ is not an unbiased estimator 
-of $\Sigma^{-1}$ \citep[p.322]{theil71}.
-Furthermore, the variance-covariance matrix calculated by (\ref{eq:rcov2})
-is not necessarily positive semidefinite \citep[p.322]{theil71}. 
-Hence, "it is doubtful whether [this formula] is really superior to 
-[equation (\ref{eq:rcov0})]" \citep[p.322]{theil71}.
-
-The WLS, SUR, W2SLS and 3SLS parameter estimates are consistent,
-if the estimated residual variance-covariance matrix is calculated
-using the residuals from a first step OLS or 2SLS estimation.
-It is also possible to iterate the estimation procedure and 
-calculate the residual variance-covariance matrix from the
-residuals of the previous iteration. 
-If equation (\ref{eq:rcov0}) is applied to calculate the estimated
-residual variance-covariance matrix, the results are maximum
-likelihood \citep[p. 345]{greene02}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Degrees of freedom}
-
-There exist different approaches to determine the degrees of
-freedom for testing the estimated parameters with a t-test.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Goodness of fit}
-
-The goodness of fit of each equation can be measured by the 
-traditional $R^2$ values:
-\begin{equation}
-   R_i^2 = 1 - \frac{ \widehat{u}_i' \widehat{u}_i }
-   { ( y_i - \overline{y_i} )' ( y_i - \overline{y_i} ) }
-\end{equation}
-where $R_i^2$ is the $R^2$ value of the $i$th equation
-and $\overline{y_i}$ is the mean value of $y_i$.
-
-The goodness of fit of the whole system can be measured by the
-McElroy's $R^2$ value \citep{mcelroy77}: 
-% also: \citep[p. 345]{greene02}
-\begin{equation}
-   R_*^2 = 1 - \frac{ \widehat{u}' \widehat{ \Omega }^{-1} \widehat{u} }
-   { y' \left( \widehat{ \Sigma }^{-1} \otimes
-   \left( I - \frac{i i'}{T} \right) \right) y }
-\end{equation}
-where $T$ is the number of observations in each equation,
-$I$ is an $T \times T$ identity matrix and 
-$i$ is a column vector of $T$ ones.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Using systemfit}\label{sec:UsingSystemfit}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Standard usage}
-
-\pkg{systemfit} is generally called by
-
-\code{
-R> systemfit( method, eqns )
-}
-
-There are two mandatory arguments: \code{method} and \code{eqns}.
-
-The argument \code{method} is a string determining the estimation method.
-It must be one of "OLS", "WLS", "SUR", "2SLS", "W2SLS" or "3SLS".
-
-The other mandatory argument \code{eqns} is a list of the equations 
-to estimate. 
-Each equation is a standard formula in \proglang{R}.
-It starts with a dependent variable on the left hand side.
-After a tilde ($\sim$) the regressors are listed%
-\footnote{For Details see the \proglang{R} help files to \code{formula}}.
-
-This is now demonstrated using an example: \\
-\code{
-R> library( systemfit ) \\
-R> data( kmenta ) \\
-R> attach( kmenta ) \\
-R> fitsur <- systemfit( "SUR", list( q $\sim$ p + d, q $\sim$ p + f + a ) ) \\
-}
-
-The first line loads the \pkg{systemfit} package. 
-The second line loads example data that are included in this package.
-These data come from \cite{kmenta86}.
-They are attached to the \proglang{R} search path in line three.
-In the last line a seemingly unrelated regression is done.
-The first equation represents the demand side of the food market.
-The dependant variable is \code{q} (food consumption per capita). 
-The regressors are \code{p} (ratio of food prices to general consumer prices)
-and \code{d} (disposable income) as well as a constant%
-\footnote{a regression constant is always implied if not explicitly omitted.}.
-The second equation represents the supply side.
-Variable \code{q} (food consumption per capita) is also the dependant 
-variable of this equation. 
-The regressors are again \code{p} (ratio of food prices to general 
-consumer prices) and a constant as well as 
-\code{f} (ratio of preceding year's prices received by farmers) and 
-\code{a} (a time trend in years).
-The regression result is assigned to the variable \code{fitsur}.
-
-Summary results can be printed by\\
-\code{
-R>~summary(~fitsur~)~\\
-~\\
-systemfit~results~\\
-method:~SUR~\\
-\\
-\mbox{}~~~N~DF~~~~~~SSR~~~~~MSE~~~~RMSE~~~~~~~R2~~~Adj~R2 \\
-1~20~17~~65.6829~3.86370~1.96563~0.755019~0.726198~\\
-2~20~16~104.0584~6.50365~2.55023~0.611888~0.539117~\\
-~\\
-The~covariance~matrix~of~the~residuals~used~for~estimation\\
-\mbox{}~~~~~~~~1~~~~~~~2~\\
-1~3.72539~4.13696~\\
-2~4.13696~5.78444~\\
-~\\
-The~covariance~matrix~of~the~residuals\\
-\mbox{}~~~~~~~~1~~~~~~~2~\\
-1~3.86370~4.92431~\\
-2~4.92431~6.50365~\\
-~\\
-The~correlations~of~the~residuals\\
-\mbox{}~~~~~~~~~1~~~~~~~~2~\\
-1~1.000000~0.982348~\\
-2~0.982348~1.000000~\\
-~\\
-The~determinant~of~the~residual~covariance~matrix:~0.879285~\\
-OLS~R-squared~value~of~the~system:~0.683453~\\
-McElroy's~R-squared~value~for~the~system:~0.788722~\\
-~\\
-SUR~estimates~for~1~~(equation~1~)~\\
-Model~Formula:~q~~~p~+~d\\
-\\
-\mbox{}~~~~~~~~~~~~~Estimate~Std.~Error~~~t~value~Pr(>|t|)~\\
-(Intercept)~99.332894~~~7.514452~13.218913~~~~~~~~0~***~\\
-p~~~~~~~~~~~-0.275486~~~0.088509~-3.112513~0.006332~~**~\\
-d~~~~~~~~~~~~~0.29855~~~0.041945~~7.117605~~~~2e-06~***~\\
----~\\
-Signif.~codes:~~0~`***'~0.001~`**'~0.01~`*'~0.05~`.'~0.1~`~'~1~\\
-~\\
-Residual~standard~error:~1.96563~on~17~degrees~of~freedom~\\
-Number~of~observations:~20~Degrees~of~Freedom:~17~\\
-SSR:~65.682902~MSE:~3.8637~Root~MSE:~1.96563~\\
-Multiple~R-Squared:~0.755019~Adjusted~R-Squared:~0.726198~\\
-~\\
-~\\
-SUR~estimates~for~2~~(equation~2~)~\\
-Model~Formula:~q~~~p~+~f~+~a\\
-\\
-\mbox{}~~~~~~~~~~~~~Estimate~Std.~Error~~t~value~Pr(>|t|)~\\
-(Intercept)~61.966166~~~11.08079~5.592215~~~~4e-05~***~\\
-p~~~~~~~~~~~~0.146884~~~0.094435~1.555397~0.139408~\\
-f~~~~~~~~~~~~0.214004~~~0.039868~5.367761~~6.3e-05~***~\\
-a~~~~~~~~~~~~0.339304~~~0.067911~4.996283~0.000132~***~\\
----~\\
-Signif.~codes:~~0~`***'~0.001~`**'~0.01~`*'~0.05~`.'~0.1~`~'~1~\\
-~\\
-Residual~standard~error:~2.550226~on~16~degrees~of~freedom~\\
-Number~of~observations:~20~Degrees~of~Freedom:~16~\\
-SSR:~104.05843~MSE:~6.503652~Root~MSE:~2.550226~\\
-Multiple~R-Squared:~0.611888~Adjusted~R-Squared:~0.539117~\\
-}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{User options}
-
-Following additional options can be set by the user:
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% \subsubsection{Equation labels}
-\paragraph{Equation labels}
-The optional argument \code{eqnlabels} allows the user to label the equations.
-It has to be a vector of strings naming the equations.\\
-\code{
-R>~fitsur~<-~systemfit(~"SUR",~list(~q~~~p~+~d,~q~~~p~+~f~+~a~),\\~
-R+~~~~eqnlabels~=~c(~"demand",~"supply"~)~)\\
-R>~summary(~fitsur~)\\
-systemfit~results\\
-method:~SUR\\
-\\
-\mbox{}~~~~~~~~N~DF~~~~~~SSR~~~~~MSE~~~~RMSE~~~~~~~R2~~~Adj~R2\\
-demand~20~17~~65.6829~3.86370~1.96563~0.755019~0.726198\\
-supply~20~16~104.0584~6.50365~2.55023~0.611888~0.539117\\
-\ldots\\
-}
-If no equation labels are provided, the equations are numbered.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\subsubsection{Instrumental variables}   
-\paragraph{Instrumental variables}   
-\code{inst}one-sided model formula specifying instrumental variables
-   or a list of one-sided model formulas if different instruments should
-   be used for the different equations (only needed for 2SLS, W2SLS and
-   3SLS estimations).
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\subsubsection{Data}   
-\paragraph{Data}   
-\code{data} an optional data frame containing the variables in the model.
-   By default the variables are taken from the environment from which
-   systemfit is called.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\subsubsection{Restrictions}   
-\paragraph{Restrictions}   
-\code{R.restr} an optional j x k matrix to impose linear
-   restrictions on the parameters by \code{R.restr} * $\beta$ = \code{q.restr}
-   (j = number of restrictions, k = number of all parameters,
-   $\beta$ = vector of all parameters).
-
-\code{q.restr} an optional j x 1 matrix to impose linear
-   restrictions (see \code{R.restr}); default is a j x 1 matrix
-   that contains only zeros.
-
-\code{TX} an optional matrix to transform the regressor matrix and,
-   hence, also the coefficient vector (see details).
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\subsubsection{Iteration control}   
-\paragraph{Iteration control}
-\code{maxiter} maximum number of iterations for WLS, SUR, W2SLS and
-   3SLS estimations.
-
-\code{tol} tolerance level indicating when to stop the iteration (only
-   WLS, SUR, W2SLS and 3SLS estimations).
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\subsubsection{Residual covariance matrix}   
-\paragraph{Residual covariance matrix}   
-\code{rcovformula} formula to calculate the estimated residual covariance
-   matrix (see details).
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\subsubsection{3SLS formula}   
-\paragraph{3SLS formula}   
-\code{formula3sls} formula for calculating the 3SLS estimator,
-   one of "GLS", "IV", "GMM", "Schmidt" or "EViews" (see details).
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\subsubsection{Degrees of freedom for t-tests}   
-\paragraph{Degrees of freedom for t-tests}   
-\code{probdfsys} use the degrees of freedom of the whole system
-   (in place of the degrees of freedom of the single equation)
-   to calculate prob values for the t-test of individual parameters.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\subsubsection{Sigma squared}   
-\paragraph{Sigma squared}   
-\code{single.eq.sigma} use different $\sigma^2$s for each
-   single equation to calculate the covariance matrix and the
-   standard errors of the coefficients (only OLS and 2SLS).
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\subsubsection{System options}   
-\paragraph{System options}
-
-\code{solvetol} tolerance level for detecting linear dependencies
-   when inverting a matrix or calculating a determinant (see
-   \code{solve} and \code{det}).
-
-\code{saveMemory} save memory by omitting some calculation that
-   are not crucial for the basic estimation (e.g McElroy's
-   $R^2$).
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Other tools}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Likelihood ratio test}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Hausman test}
-
-\textbf{Man, I really have to clean this up.}
-
-Hausman \citep{hausman1978} developed a test have power against
-alternatives for which $\hat{\theta}$ and $\tilde{\theta}$ diverge
-under misspecification.
-
-The Hausman result if the difference between the covariance of an
-efficient and the covariance of the inefficient estimator is zero, the
-model is correctly specified.
-
-The test is based on the idea that under the hypothesis of no
-correlation, both the limited information estimates and the
-full-information results are consistent, but only the full-information
-estimates are efficient and under the alternative, the limited
-information estimates are consistent, but the full information
-estimates are not.
-
-Since the models that are fit using the \pkg{systemfit} package are
-typically random effects models, the effects may be inconsistent
-between the introduced variables and the regressors.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Summary and outlook}\label{sec:Summmary}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-What should we place in this section?
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Unbalanced Observations}\label{sec:summary_unbalanced_data}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-The \pkg{systemfit} package was originally developed to fit
-simultaneous equations for forestry datasets. Forestry datasets
-typically contain observations for many inexpensive observations
-(diameter at breast height) and very few expensive observations such
-as total tree height. While the package currently does not allow for
-unbalanced datasets, data where the equations contain different
-numbers of observations, future releases of the package will have that
-feature implemented. We are still trying to figure out how develop the
-code for instrumental variable methods (2SLS,3SLS) for the correct
-estimation methods if possible.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Additional Methods}\label{sec:summary_additional_methods}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsubsection{Full Information Maximum Likelihood}
-
-Describe, in a paragraph the process and give some reasons why you
-might want this in the package.
-
-\subsubsection{Generalized Method of Moments}
-
-Describe, in a paragraph the process and give some reasons why you
-might want this in the package.
+\input{systemfit_intro.tex}
+\input{systemfit_estimation.tex}
+\input{systemfit_restrictions.tex}
+\input{systemfit_other.tex}
+\input{systemfit_usage.tex}
+\input{systemfit_summary.tex}
 
@@
 $Log$
+Revision 1.11  2005/09/01 12:27:55  henningsena
+revised structure of the document (sections, subsections, subsubsections) due to removal of nlsystemfit; moved each section into a separate file.
+
 Revision 1.10  2005/09/01 11:49:02  henningsena

### 2005-09-01 paper/trunk/systemfit_estimation.tex
revised structure of the document (sections, subsections, subsubsections) due to removal of nlsystemfit; moved each section into a separate file.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@571 591db2bd-e0a0-4d22-8c19-2192dbb7da31

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Estimation}\label{sec:Estimation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Consider a system of $G$ equations, where the $i$th equation is of
the form 
\begin{equation}
   y_{i} = X_i \beta_i + u_i, \quad i = 1, 2, \ldots, G
\end{equation}
where $y_i$ is a vector of the dependent variable,
$X_i$ is a matrix of the exogenous variables,
$\beta_i$ is the coefficient vector and
$u_i$ is a vector of the disturbance terms of the $i$th equation.

We can write the 'stacked' system as
\begin{equation}
   \left[ \begin{array}{c}
      y_1 \\ y_2\\ \vdots\\ y_G
   \end{array} \right] = 
   \left[ \begin{array}{cccc}
      X_1 & 0 & \cdots & 0\\
      0 & X_2 & \cdots & 0\\
      \vdots & \vdots & \ddots & \vdots\\
      0 & 0 & \cdots & X_G
   \end{array}\right] \cdot 
   \left[ \begin{array}{c}
      \beta_1 \\ \beta_2 \\ \vdots\\ \beta_G
   \end{array} \right] +
   \left[ \begin{array}{c}
      u_1 \\ u_2 \\ \vdots\\ u_G
   \end{array} \right]
\end{equation}

or more simply as
\begin{equation}
   y = X \beta + u 
\end{equation}

We assume that there is no correlation of the disturbance terms 
across observations:
\begin{equation}
   E \left( u_{it} \, u_{jt^*} \right) = 0 
   \; \forall \; t \neq t^*
\end{equation}
where $i$ and $j$ indicate the equation number 
and $t$ and $t^*$ denote the observation number.

However, we explicitly allow for contemporaneous correlation:
\begin{equation}
   E \left( u_{it} \, u_{jt} \right) = \sigma_{ij}
\end{equation}

Thus, the covariance matrix of the total system is
\begin{equation}
   E \left( u \, u' \right) = \Omega = \Sigma \otimes I
\end{equation}
where $\Sigma = \left[ \sigma_{ij} \right]$ and $I$ is an
identity matrix.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Ordinary Least Squares estimation}

The Ordinary Least Squares (OLS) estimator of the system 
is obtained by
\begin{equation}
   \widehat{\beta} = \left( X'X \right)^{-1} X'y
\end{equation}

These estimates are efficient only if the disturbance terms are not 
contemporaneously correlated, which means 
\begin{equation}
   \sigma_{ij} = 0 \; \forall \; i \neq j
\end{equation}

If the whole system is treated as one single equation, 
the variance-covariance matrix of the estimated parameters is
\begin{equation}
   Var \left[ \widehat{\beta} \right] = \sigma^2 \left( X'X \right)^{-1}
\end{equation}
with $\sigma^2 = E \left( u' u \right)$.
This assumes that the disturbances of all equations have the
same variance.

If the disturbance terms of the individual equations 
are allowed to have different variances, 
the variance-covariance matrix of the estimated parameters is
\begin{equation}
   Var \left[ \widehat{\beta} \right] = \left( X' \Omega^{-1} X \right)^{-1}
   \label{eq:OLSvcm2}
\end{equation}
with $\Omega = \Sigma \otimes I$, 
$\sigma_{ij} = 0 \; \forall \; i \neq j$ and
$\sigma_{ii} = E \left( u_i' u_i \right)$.

If no cross-equation restrictions are imposed, the simultaneous 
OLS estimation of the system leads to the same parameter estimates 
as an equation-wise OLS estimation. The variance-covariance matrix 
of the parameters from an equation-wise OLS estimation is equal to 
the variance-covariance matrix obtained by equation 
(\ref{eq:OLSvcm2}).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Weighted least squares}

The Weighted Least Squares (WLS) estimator of the system 
is obtained by
\begin{equation}
   \widehat{\beta} = \left( X' \Omega^{-1} X \right)^{-1} X' \Omega^{-1} y
\end{equation}
with $\Omega = \Sigma \otimes I$, 
$\sigma_{ij} = 0 \; \forall \; i \neq j$ and
$\sigma_{ii} = E \left( u_i' u_i \right)$.

Like the OLS estimates these estimates are only efficient 
if the disturbance terms are not contemporaneously correlated.

The variance-covariance matrix of the estimated parameters is
\begin{equation}
   Var \left[ \widehat{\beta} \right] = \left( X' \Omega^{-1} X \right)^{-1}
\end{equation}

If no cross-equation restrictions are imposed, 
the parameter estimates and the variance-covariance matrix 
of the parameters are equal to the OLS estimates%
\footnote{The variance-covariance matrices of the 
estimated parameters might differ, 
if the WLS estimation is based on a different formula 
to calculate the (diagonal) variance-covariance 
matrix of the residuals (see section \ref{sec:residcov}).}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Seemingly unrelated regression}

When the disturbances are contemporaneously correlated, a Generalized 
Least Squares (GLS) estimation leads to efficient parameter estimates.
In this case the GLS is generally called "Seemingly Unrelated Regression" 
(SUR) (see \citealp{zellner62}). 
Note, while an unbiased OLS or WLS estimation requires only that 
the regressors and the disturbance terms of each single 
equation are uncorrelated,
a consistent SUR estimation requires that all disturbance terms and all 
regressors are uncorrelated:
\begin{equation}
   E \left( u | X \right) = 0
\end{equation}

The SUR estimator can be obtained by:%
\footnote{To calculate $\Omega^{-1}$ it is not necessary to invert 
the (huge) $\Omega$ matrix, 
because it is sufficient to invert the (small) $\Sigma$ matrix and
calculate it by $\Omega^{-1} = \Sigma^{-1} \otimes I$.}

\begin{equation}
   \widehat{\beta} = \left( X' \Omega^{-1} X \right)^{-1} X' \Omega^{-1} y
\end{equation}
with $\Omega = \Sigma \otimes I$ and
$\sigma_{ij} = E \left( u_i' u_j \right)$.

And the variance-covariance matrix of the estimated parameters is
\begin{equation}
   Var \left[ \widehat{\beta} \right] = \left( X' \Omega^{-1} X \right)^{-1}
\end{equation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Two-stage least squares estimation}

If the regressors of one or more equations are correlated 
with the disturbances ($E \left( u_i | X_i \right) \neq 0$), 
the estimated coefficients are biased.
This can be circumvented by an instrumental variable (IV) estimation.
The instrumental variables for each equation $H_i$ 
can be either different or identical for all equations.
The instrumental variables of each equation may not be correlated with 
the disturbance terms of the corresponding equation 
($E \left( u_i | H_i \right) = 0$).

At the first stage new ('fitted') regressors are obtained by
\begin{equation}
   \widehat{X_i} = H_i \left( H_i' H_i \right)^{-1} H_i' X
\end{equation}

At the second stage the unbiased two-stage least squares estimates
of $\beta$ are obtained by:
\begin{equation}
   \widehat{\beta} = \left( \widehat{X}' \widehat{X} \right)^{-1} 
   \widehat{X}' y 
   \label{eq:beta2sls}
\end{equation}

If the whole system is treated as one single equation, 
the variance-covariance matrix of the estimated parameters is
\begin{equation}
   Var \left[ \widehat{\beta} \right] = \sigma^2 \left( \widehat{X}'
   \widehat{X} \right)^{-1}
\end{equation}
with $\sigma^2 = E \left( u' u \right)$.

If the disturbance terms of the individual equations 
are allowed to have different variances, 
the variance-covariance matrix of the estimated parameters is
\begin{equation}
   Var \left[ \widehat{\beta} \right] = \left( \widehat{X}' \Omega^{-1} 
   \widehat{X} \right)^{-1}
\end{equation}
with $\Omega = \Sigma \otimes I$, 
$\sigma_{ij} = 0 \; \forall \; i \neq j$ and
$\sigma_{ii} = E \left( u_i' u_i \right)$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Weighted two-stage least squares estimation}

The Weighted Two-Stage Least Squares (W2SLS) estimator of the system 
is obtained by
\begin{equation}
   \widehat{\beta} = \left( \widehat{X}' \Omega^{-1} \widehat{X} 
   \right)^{-1} \widehat{X}' \Omega^{-1} y
\end{equation}
with $\Omega = \Sigma \otimes I$, 
$\sigma_{ij} = 0 \; \forall \; i \neq j$ and
$\sigma_{ii} = E \left( u_i' u_i \right)$.

The variance-covariance matrix of the estimated parameters is
\begin{equation}
   Var \left[ \widehat{\beta} \right] = \left( \widehat{X}' \Omega^{-1} 
   \widehat{X} \right)^{-1}
\end{equation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Three-stage least squares estimation}

If the regressors are correlated with the disturbances 
($E \left( u | X \right) \neq 0$) and 
the disturbances are contemporaneously correlated, 
an Instrumental Variable (IV) Generalized Least Squares (GLS) 
estimation leads to consistent and efficient estimates. This estimation 
procedure is generally called "Three-stage Least Squares" (3SLS).

The standard 3SLS estimator can be obtained by:
\begin{equation}
   \widehat{\beta} = \left( \widehat{X}' \Omega^{-1} \widehat{X} 
   \right)^{-1} \widehat{X}' \Omega^{-1} y
\end{equation}
with $\Omega = \Sigma \otimes I$ and
$\sigma_{ij} = E \left( u_i' u_j \right)$.

Its variance-covariance matrix is:
\begin{equation}
   Var \left[ \widehat{\beta} \right] = \left( \widehat{X}' \Omega^{-1} 
   \widehat{X} \right)^{-1}
   \label{eq:cov3sls}
\end{equation}

While an unbiased 2SLS or W2SLS estimation requires only that 
the instrumental variables and the disturbance terms of each single 
equation are uncorrelated,
\cite{schmidt90} points out that this estimator is only consistent 
if all disturbance terms and all instrumental variables are uncorrelated:
\begin{equation}
   E \left( u | H \right) = 0
\end{equation}
with
\begin{equation}
   H =
   \left[ \begin{array}{cccc}
      H_1 & 0 & \cdots & 0\\
      0 & H_2 & \cdots & 0\\
      \vdots & \vdots & \ddots & \vdots\\
      0 & 0 & \cdots & H_G
   \end{array}\right] \cdot 
\end{equation}

Since there might be occasions where this cannot be avoided, 
\cite{schmidt90} analyses other approaches to obtain 3SLS estimators:

One of these approaches is based on instrumental variable estimation
(IV-3SLS):
\begin{equation}
   \widehat{\beta} = \left( \widehat{X}' \Omega^{-1} X 
   \right)^{-1} \widehat{X}' \Omega^{-1} y
\end{equation}

The variance-covariance matrix of this IV-3SLS estimator is:
\begin{equation}
   Var \left[ \widehat{\beta} \right] = \left( \widehat{X}' \Omega^{-1} 
   X \right)^{-1}
\end{equation}

Another approach is based on the Generalized Method of Moments (GMM) 
estimator (GMM-3SLS):
\begin{equation}
   \widehat{\beta} = \left( X' H \left( H' \Omega H \right)^{-1}
   H' X \right)^{-1} X' H \left( H' \Omega H \right)^{-1} H' y
   \label{eq:beta3slsGMM}
\end{equation}

The variance-covariance matrix of the GMM-3SLS estimator is:
\begin{equation}
   Var \left[ \widehat{\beta} \right] = 
   \left( X' H \left( H' \Omega H \right)^{-1} H' X \right)^{-1}
\end{equation}

A fourth approach developed by \cite{schmidt90} himself is:
\begin{equation}
   \widehat{\beta} = \left( \widehat{X}' \Omega^{-1} \widehat{X} 
   \right)^{-1} \widehat{X}' \Omega^{-1} 
   H \left( H' H \right)^{-1} H' y 
   \label{eq:beta3slsSchmidt}
\end{equation}

The variance-covariance matrix of this estimator is:
\begin{equation}
   Var \left[ \widehat{\beta} \right] = 
   \left( \widehat{X}' \Omega^{-1}  \widehat{X} \right)^{-1} 
   \widehat{X}' \Omega^{-1} H \left( H' H \right)^{-1} H' \Omega 
   H \left( H' H \right)^{-1} H' \Omega^{-1} \widehat{X}
   \left( \widehat{X}' \Omega^{-1}  \widehat{X} \right)^{-1}
\end{equation}

The econometrics software EViews uses following approach:
\begin{equation}
   \widehat{\beta} = \widehat{\beta}_{2SLS} + 
   \left( \widehat{X}' \Omega^{-1} \widehat{X} \right)^{-1} 
   \widehat{X}' \Omega^{-1} \left( y - X \widehat{\beta}_{2SLS} \right)
\end{equation}
with $\widehat{\beta}_{2SLS}$ is the two-stage least squares estimator
(equation \ref{eq:beta2sls}).
EViews uses the standard 3SLS formula (equation (\ref{eq:cov3sls})) to 
calculate the variance-covariance matrix of the 3SLS estimator.

If the same instrumental variables are used in all equations 
($H_1 = H_2 = \ldots = H_G$), 
all the above mentioned approaches lead to identical parameter estimates.
However, if this is not the case, the results depend on the 
equation used \citep{schmidt90}. 
The only reason to use different instruments in the different equations 
is that the instruments of one equation are correlated with the
disturbance terms in another equation. Otherwise, one would simply use
every instrument in every equation \citep{schmidt90}. 
In this case, only the GMM-3SLS (equation (\ref{eq:beta3slsGMM}))
and the 3SLS estimator developed by \cite{schmidt90} 
(equation (\ref{eq:beta3slsSchmidt})) are consistent.

### 2005-09-01 paper/trunk/systemfit_intro.tex
revised structure of the document (sections, subsections, subsubsections) due to removal of nlsystemfit; moved each section into a separate file.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@571 591db2bd-e0a0-4d22-8c19-2192dbb7da31

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Many theoretical models consist of more than one equation. If these
models are econometrically estimated, the equations generally should
not be estimated individually, but all together, because the
disturbance terms of these equations are likely contemporaneously
correlated. Estimating all equations simultaneously and taking the
covariance structure of the residuals into account leads to efficient
estimates, while ignoring this correlation leads to an inefficient
parameter estimation \citep{zellner62}.

Another reason to estimate an equation system simultaneously are
cross-equation parameter restrictions. These restrictions can be
tested and/or imposed only in a simultaneous estimation
approach. Especially the economic theory suggests many cross-equation
restriction.

The \pkg{systemfit} package provides the capability to estimate
linear and non-linear equation systems in \proglang{R}. Compared to
other statistics and econometrics software the (advanced) user can
control many important details of the estimation procedure. On the
other hand these user options have reasonable defaults that also
beginners can easily use the provided functions.

\textbf{I'll have to update this section to match the recent changes. JDH.}

This paper is organized as follows: In section~\ref{sec:Estimation} we
introduce the mathematics of estimating equation systems.
Section~\ref{sec:Restrictions} shows how linear restrictions can be
imposed.
All other relevant issues are discussed in section~\ref{sec:Other}.
Section~\ref{sec:Usage} demonstrates how to run
\pkg{systemfit}, especially how the features presented in the previous
sections can be used.
Finally, a summary and an outlook are given in
section~\ref{sec:Summmary}.

### 2005-09-01 paper/trunk/systemfit_other.tex
revised structure of the document (sections, subsections, subsubsections) due to removal of nlsystemfit; moved each section into a separate file.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@571 591db2bd-e0a0-4d22-8c19-2192dbb7da31

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Other issues and tools}\label{sec:Other}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Residual variance-covariance matrix}\label{sec:residcov}

Since the true residuals of the estimated equations are generally not known,
the true covariance matrix of the residuals cannot be determined.
Thus, this covariance matrix must be calculated from the
\emph{estimated} residuals.

The estimated covariance matrix of the residuals ($\widehat{\Sigma} = 
\left[ \widehat{\sigma}_{ij} \right]$) 
is often calculated by
\begin{equation}
   \widehat{\sigma}_{ij} = \frac{ \widehat{u}_i' \widehat{u}_j }{ T }
   \label{eq:rcov0}
\end{equation}
where $T$ is the number of observations in each equation.

However, this estimator is biased, because it is not corrected for 
degrees of freedom. 
The usual single-equation procedure to correct for degrees of freedom 
cannot always be applied, because the number of regressors in each equation 
might differ.
Two alternative approaches to calculate the residual variance-covariance 
matrix are
\begin{equation}
   \widehat{\sigma}_{ij} = \frac{ \widehat{u}_i' \widehat{u}_j }
   { \sqrt{ \left( T - K_i \right) \cdot \left( T - K_j \right) } }
   \label{eq:rcov1}
\end{equation}
and
\begin{equation}
   \widehat{\sigma}_{ij} = \frac{ \widehat{u}_i' \widehat{u}_j }
   { T - \max \left( K_i , K_j \right) }
\end{equation}
where $K_i$ and $K_j$ are the number of regressors in equation
$i$ and $j$, respectively.

However, this yields an unbiased estimator only if $K_i = K_j$ 
\citep[p. 469]{judge85}. 
% Greene (2002, p. 344) says that the second is unbiased if i = j or K_i = K_j,
% whereas the first is unbiased only if i = j. 
% However, if K_i = K_j the first and the second are equal.
% Why is the first biased if K_i = K_j ???????????

A further approach to obtain the estimated residual variance-covariance
matrix is \citep[p. 309]{zellner62c}
\begin{eqnarray}
   \widehat{\sigma}_{ij} & = & 
   \frac{ \widehat{u}_i' \widehat{u}_j } 
   { T - K_i - K_j + tr \left[ X_i \left( X_i' X_i \right)^{-1}
   X_i' X_j \left( X_j' X_j \right)^{-1} X_j' \right] } \\
   & = &
   \frac{ \widehat{u}_i' \widehat{u}_j } 
   { T - K_i - K_j + tr \left[ \left( X_i' X_i \right)^{-1}
   X_i' X_j \left( X_j' X_j \right)^{-1} X_j' X_i \right] } \\
   \label{eq:rcov2}
\end{eqnarray} 
This yields an unbiased estimator for all elements of $\widehat{\Sigma}$,
but even if $\widehat{\Sigma}$ is an unbiased estimator of $\Sigma$, 
its inverse $\widehat{\Sigma}^{-1}$ is not an unbiased estimator 
of $\Sigma^{-1}$ \citep[p.322]{theil71}.
Furthermore, the variance-covariance matrix calculated by (\ref{eq:rcov2})
is not necessarily positive semidefinite \citep[p.322]{theil71}. 
Hence, "it is doubtful whether [this formula] is really superior to 
[equation (\ref{eq:rcov0})]" \citep[p.322]{theil71}.

The WLS, SUR, W2SLS and 3SLS parameter estimates are consistent,
if the estimated residual variance-covariance matrix is calculated
using the residuals from a first step OLS or 2SLS estimation.
It is also possible to iterate the estimation procedure and 
calculate the residual variance-covariance matrix from the
residuals of the previous iteration. 
If equation (\ref{eq:rcov0}) is applied to calculate the estimated
residual variance-covariance matrix, the results are maximum
likelihood \citep[p. 345]{greene02}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Degrees of freedom}

There exist different approaches to determine the degrees of
freedom for testing the estimated parameters with a t-test.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Goodness of fit}

The goodness of fit of each equation can be measured by the 
traditional $R^2$ values:
\begin{equation}
   R_i^2 = 1 - \frac{ \widehat{u}_i' \widehat{u}_i }
   { ( y_i - \overline{y_i} )' ( y_i - \overline{y_i} ) }
\end{equation}
where $R_i^2$ is the $R^2$ value of the $i$th equation
and $\overline{y_i}$ is the mean value of $y_i$.

The goodness of fit of the whole system can be measured by the
McElroy's $R^2$ value \citep{mcelroy77}: 
% also: \citep[p. 345]{greene02}
\begin{equation}
   R_*^2 = 1 - \frac{ \widehat{u}' \widehat{ \Omega }^{-1} \widehat{u} }
   { y' \left( \widehat{ \Sigma }^{-1} \otimes
   \left( I - \frac{i i'}{T} \right) \right) y }
\end{equation}
where $T$ is the number of observations in each equation,
$I$ is an $T \times T$ identity matrix and 
$i$ is a column vector of $T$ ones.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Likelihood ratio test}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Hausman test}

\textbf{Man, I really have to clean this up.}

Hausman \citep{hausman1978} developed a test have power against
alternatives for which $\hat{\theta}$ and $\tilde{\theta}$ diverge
under misspecification.

The Hausman result if the difference between the covariance of an
efficient and the covariance of the inefficient estimator is zero, the
model is correctly specified.

The test is based on the idea that under the hypothesis of no
correlation, both the limited information estimates and the
full-information results are consistent, but only the full-information
estimates are efficient and under the alternative, the limited
information estimates are consistent, but the full information
estimates are not.

Since the models that are fit using the \pkg{systemfit} package are
typically random effects models, the effects may be inconsistent
between the introduced variables and the regressors.

### 2005-09-01 paper/trunk/systemfit_restrictions.tex
revised structure of the document (sections, subsections, subsubsections) due to removal of nlsystemfit; moved each section into a separate file.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@571 591db2bd-e0a0-4d22-8c19-2192dbb7da31

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Imposing linear restrictions}\label{sec:Restrictions}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

There are two ways to impose linear parameter restrictions.
First, a matrix $T$ can be provided with
\begin{equation}
   \beta = T \cdot \beta^* \label{eq:T-restr} 
\end{equation}
where $\beta^*$ is a vector of linear independent coefficients.

To impose these restrictions the $X$ matrix is
(post-)multiplied by this $T$ matrix.
\begin{equation}
    X^* = X \cdot T
\end{equation}

Then, a standard estimation is done by substituting $X^*$ for $X$.
This results in the linear independent parameter estimates $\beta^*$ and
the variance-covariance matrix of these parameters. 
The original parameters can be obtained by equation (\ref{eq:T-restr})
and the variance-covariance matrix of the original parameters 
can be obtained by:
\begin{equation}
   Var \left[ \widehat{\beta} \right] = T \cdot Var \left[ \widehat{\beta^*} \right] \cdot T'
\end{equation}

The second way to impose linear parameter restrictions 
can be formulated by
\begin{equation}
   R \beta^0 = q
\end{equation}
where $\beta^0$ is the vector of the restricted coefficients, 
$R$ is a matrix and $q$ is a vector (see \citealp[p. 100]{greene02}). 
Each restriction belongs to one row of $R$ and the corresponding 
element of~$q$.

The first way is less flexible than this latter one%
\footnote{e.g. restrictions like $\beta_1 + \beta_2 = 4$ cannot be imposed
by the first method}, 
but the first way is preferable if equality constraints for coefficients
across many equations of the system are imposed. 
Of course, these restrictions can be also imposed using
the latter method.
However, while the latter method increases the dimension of the 
matrices to be inverted during estimation, the first reduces it. 
Thus, in some cases the latter way leads to estimation problems
(e.g. (near) singularity of the matrices to be inverted),
while the first doesn't.

These two methods can be combined. In this case the restrictions
imposed using the latter method are imposed on the linear independent 
parameters due to the restrictions imposed using the first method:
\begin{equation}
   R \beta^{*0} = q
\end{equation}
where $\beta^{*0}$ is the vector of the restricted $\beta^*$ coefficients.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Restricted OLS estimation}

The OLS estimator restricted by $R \beta^0 = q$ can be obtained by
\begin{equation}
   \left[ \begin{array}{c}
      \widehat{\beta^0} \\ \widehat{\lambda}
   \end{array} \right]
   =
   \left[ \begin{array}{cc}
      X' X & R' \\ 
      R & 0
   \end{array} \right]^{-1}
   \cdot
   \left[ \begin{array}{c}
      X' y \\ q 
   \end{array} \right]
\end{equation}
where $2\lambda$ is a vector of the Lagrangean multipliers of the restrictions.

If the whole system is treated as one single equation, 
the variance-covariance matrix of the estimated parameters is
\begin{equation}
   Var 
   \left[ \begin{array}{c}
      \widehat{\beta} \\ \widehat{\lambda}
   \end{array} \right] 
   = \sigma^2 
   \left[ \begin{array}{cc}
      X' X & R' \\ 
      R & 0
   \end{array} \right]^{-1}
\end{equation}
with $\sigma^2 = E \left( u' u \right)$.

If the disturbance terms of the individual equations 
are allowed to have different variances, 
the variance-covariance matrix of the estimated parameters is
\begin{equation}
   Var 
   \left[ \begin{array}{c}
      \widehat{\beta} \\ \widehat{\lambda}
   \end{array} \right] 
   = 
   \left[ \begin{array}{cc}
      X' \Omega^{-1} X & R' \\ 
      R & 0
   \end{array} \right]^{-1}
\end{equation}
with $\Omega = \Sigma \otimes I$, 
$\sigma_{ij} = 0 \; \forall \; i \neq j$ and
$\sigma_{ii} = E \left( u_i' u_i \right)$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Restricted WLS estimation}

The WLS estimator restricted by $R \beta^0 = q$ can be obtained by
\begin{equation}
   \left[ \begin{array}{c}
      \widehat{\beta^0} \\ \widehat{\lambda}
   \end{array} \right]
   =
   \left[ \begin{array}{cc}
      X' \Omega^{-1} X & R' \\ 
      R & 0
   \end{array} \right]^{-1}
   \cdot
   \left[ \begin{array}{c}
      X' \Omega^{-1} y \\ q 
   \end{array} \right]
\end{equation}
with $\Omega = \Sigma \otimes I$, 
$\sigma_{ij} = 0 \; \forall \; i \neq j$ and
$\sigma_{ii} = E \left( u_i' u_i \right)$.

The variance-covariance matrix of the estimated parameters is
\begin{equation}
   Var 
   \left[ \begin{array}{c}
      \widehat{\beta} \\ \widehat{\lambda}
   \end{array} \right] 
   = 
   \left[ \begin{array}{cc}
      X' \Omega^{-1} X & R' \\ 
      R & 0
   \end{array} \right]^{-1}
\end{equation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Restricted SUR estimation}

The SUR estimator restricted by $R \beta^0 = q$ can be obtained by
\begin{equation}
   \left[ \begin{array}{c}
      \widehat{\beta^0} \\ \widehat{\lambda}
   \end{array} \right]
   =
   \left[ \begin{array}{cc}
      X' \Omega^{-1} X & R' \\ 
      R & 0
   \end{array} \right]^{-1}
   \cdot
   \left[ \begin{array}{c}
      X' \Omega^{-1} y \\ q 
   \end{array} \right]
\end{equation}
with $\Omega = \Sigma \otimes I$ and
$\sigma_{ij} = E \left( u_i' u_j \right)$.

The variance-covariance matrix of the estimated parameters is
\begin{equation}
   Var 
   \left[ \begin{array}{c}
      \widehat{\beta} \\ \widehat{\lambda}
   \end{array} \right] 
   = 
   \left[ \begin{array}{cc}
      X' \Omega^{-1} X & R' \\ 
      R & 0
   \end{array} \right]^{-1}
\end{equation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Restricted 2SLS estimation}

The 2SLS estimator restricted by $R \beta^0 = q$ can be obtained by
\begin{equation}
   \left[ \begin{array}{c}
      \widehat{\beta^0} \\ \widehat{\lambda}
   \end{array} \right]
   =
   \left[ \begin{array}{cc}
      \widehat{X}' \widehat{X} & R' \\ 
      R & 0
   \end{array} \right]^{-1}
   \cdot
   \left[ \begin{array}{c}
      \widehat{X}' y \\ q 
   \end{array} \right]
   \label{eq:beta2SLSr}
\end{equation}

If the whole system is treated as one single equation, 
the variance-covariance matrix of the estimated parameters is
\begin{equation}
   Var 
   \left[ \begin{array}{c}
      \widehat{\beta} \\ \widehat{\lambda}
   \end{array} \right] 
   = \sigma^2 
   \left[ \begin{array}{cc}
      \widehat{X}' \widehat{X} & R' \\ 
      R & 0
   \end{array} \right]^{-1}
\end{equation}
with $\sigma^2 = E \left( u' u \right)$.

If the disturbance terms of the individual equations 
are allowed to have different variances, 
the variance-covariance matrix of the estimated parameters is
\begin{equation}
   Var 
   \left[ \begin{array}{c}
      \widehat{\beta} \\ \widehat{\lambda}
   \end{array} \right] 
   = 
   \left[ \begin{array}{cc}
      \widehat{X}' \Omega^{-1} \widehat{X} & R' \\ 
      R & 0
   \end{array} \right]^{-1}
\end{equation}
with $\Omega = \Sigma \otimes I$, 
$\sigma_{ij} = 0 \; \forall \; i \neq j$ and
$\sigma_{ii} = E \left( u_i' u_i \right)$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Restricted W2SLS estimation}

The W2SLS estimator restricted by $R \beta^0 = q$ can be obtained by
\begin{equation}
   \left[ \begin{array}{c}
      \widehat{\beta^0} \\ \widehat{\lambda}
   \end{array} \right]
   =
   \left[ \begin{array}{cc}
      \widehat{X}' \Omega^{-1} \widehat{X} & R' \\ 
      R & 0
   \end{array} \right]^{-1}
   \cdot
   \left[ \begin{array}{c}
      \widehat{X}' \Omega^{-1} y \\ q 
   \end{array} \right]
\end{equation}
with $\Omega = \Sigma \otimes I$, 
$\sigma_{ij} = 0 \; \forall \; i \neq j$ and
$\sigma_{ii} = E \left( u_i' u_i \right)$.

The variance-covariance matrix of the estimated parameters is
\begin{equation}
   Var 
   \left[ \begin{array}{c}
      \widehat{\beta} \\ \widehat{\lambda}
   \end{array} \right] 
   = 
   \left[ \begin{array}{cc}
      \widehat{X}' \Omega^{-1} \widehat{X} & R' \\ 
      R & 0
   \end{array} \right]^{-1}
\end{equation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Restricted 3SLS estimation}

The standard 3SLS estimator restricted by $R \beta^0 = q$ can be obtained by
\begin{equation}
   \left[ \begin{array}{c}
      \widehat{\beta^0} \\ \widehat{\lambda}
   \end{array} \right]
   =
   \left[ \begin{array}{cc}
      \widehat{X}' \Omega^{-1} \widehat{X} & R' \\ 
      R & 0
   \end{array} \right]^{-1}
   \cdot
   \left[ \begin{array}{c}
      \widehat{X}' \Omega^{-1} y \\ q 
   \end{array} \right]
\end{equation}
with $\Omega = \Sigma \otimes I$ and
$\sigma_{ij} = E \left( u_i' u_j \right)$.

The variance-covariance matrix of this estimator is
\begin{equation}
   Var 
   \left[ \begin{array}{c}
      \widehat{\beta^0} \\ \widehat{\lambda}
   \end{array} \right] 
   = 
   \left[ \begin{array}{cc}
      \widehat{X}' \Omega^{-1} \widehat{X} & R' \\ 
      R & 0
   \end{array} \right]^{-1}
   \label{eq:cov3slsr}
\end{equation}

The IV-3SLS estimator restricted by $R \beta^0 = q$ can be obtained by
\begin{equation}
   \left[ \begin{array}{c}
      \widehat{\beta^0} \\ \widehat{\lambda}
   \end{array} \right]
   =
   \left[ \begin{array}{cc}
      \widehat{X}' \Omega^{-1} X & R' \\ 
      R & 0
   \end{array} \right]^{-1}
   \cdot
   \left[ \begin{array}{c}
      \widehat{X}' \Omega^{-1} y \\ q 
   \end{array} \right]
\end{equation}

with
\begin{equation}
   Var 
   \left[ \begin{array}{c}
      \widehat{\beta^0} \\ \widehat{\lambda}
   \end{array} \right] 
   = 
   \left[ \begin{array}{cc}
      \widehat{X}' \Omega^{-1} \widehat{X} & R' \\ 
      R & 0
   \end{array} \right]^{-1}
\end{equation}

The restricted 3SLS-GMM estimator can be obtained by
\begin{equation}
   \left[ \begin{array}{c}
      \widehat{\beta^0} \\ \widehat{\lambda}
   \end{array} \right]
   =
   \left[ \begin{array}{cc}
      X' H \left( H' \Omega H \right)^{-1} H' X & R' \\ 
      R & 0
   \end{array} \right]^{-1}
   \cdot
   \left[ \begin{array}{c}
      X' H \left( H \Omega H \right)^{-1} H' y \\ q 
   \end{array} \right]
\end{equation}

with
\begin{equation}
   Var 
   \left[ \begin{array}{c}
      \widehat{\beta^0} \\ \widehat{\lambda}
   \end{array} \right] 
   = 
   \left[ \begin{array}{cc}
      X' H \left( H' \Omega H \right)^{-1} H' X & R' \\ 
      R & 0
   \end{array} \right]^{-1}
\end{equation}

The restricted 3SLS estimator based on the suggestion of 
\cite{schmidt90} is:
\begin{equation}
   \left[ \begin{array}{c}
      \widehat{\beta^0} \\ \widehat{\lambda}
   \end{array} \right]
   =
   \left[ \begin{array}{cc}
      \widehat{X}' \Omega^{-1} \widehat{X} & R' \\ 
      R & 0
   \end{array} \right]^{-1}
   \cdot
   \left[ \begin{array}{c}
      \widehat{X}' \Omega^{-1} H \left( H' H \right)^{-1} H' y \\ q 
   \end{array} \right]
\end{equation}

with
\begin{eqnarray}
   Var 
   \left[ \begin{array}{c}
      \widehat{\beta^0} \\ \widehat{\lambda}
   \end{array} \right] 
   & = & 
   \left[ \begin{array}{cc}
      \widehat{X}' \Omega^{-1} \widehat{X} & R' \\ 
      R & 0
   \end{array} \right]^{-1}
   \cdot
   \left[ \begin{array}{cc}
      \widehat{X}' \Omega^{-1} H \left( H' H \right)^{-1} H' \Omega
      H \left( H' H \right)^{-1} H' \Omega^{-1} \widehat{X} & 0' \\ 
      0 & 0
   \end{array} \right]^{-1}
   \nonumber \\
   & & \cdot
   \left[ \begin{array}{cc}
      \widehat{X}' \Omega^{-1} \widehat{X} & R' \\ 
      R & 0
   \end{array} \right]^{-1}
\end{eqnarray}

The econometrics software EViews calculates the restricted 3SLS estimator by:
\begin{equation}
   \left[ \begin{array}{c}
      \widehat{\beta^0} \\ \widehat{\lambda}
   \end{array} \right]
   =
   \left[ \begin{array}{cc}
      \widehat{X}' \Omega^{-1} \widehat{X} & R' \\ 
      R & 0
   \end{array} \right]^{-1}
   \cdot
   \left[ \begin{array}{c}
      \widehat{X}' \Omega^{-1} \left( y - X \widehat{\beta^0}_{2SLS} \right) 
      \\ q 
   \end{array} \right]
\end{equation}
where $\widehat{\beta^0}_{2SLS}$ is the restricted 2SLS estimator calculated
by equation (\ref{eq:beta2SLSr}). 
To calculate the variance-covariance matrix 
EViews uses the standard formula of the restricted 3SLS estimator
(equation (\ref{eq:cov3slsr})).

If the same instrumental variables are used in all equations 
($H_1 = H_2 = \ldots = H_G$), 
all the above mentioned approaches lead to identical parameter estimates
and identical variance-covariance matrices of the estimated parameters.

(***** This has to be checked!!!!! **************)

### 2005-09-01 paper/trunk/systemfit_summary.tex
revised structure of the document (sections, subsections, subsubsections) due to removal of nlsystemfit; moved each section into a separate file.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@571 591db2bd-e0a0-4d22-8c19-2192dbb7da31

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Summary and outlook}\label{sec:Summmary}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

What should we place in this section?

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Unbalanced Observations}\label{sec:summary_unbalanced_data}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The \pkg{systemfit} package was originally developed to fit
simultaneous equations for forestry datasets. Forestry datasets
typically contain observations for many inexpensive observations
(diameter at breast height) and very few expensive observations such
as total tree height. While the package currently does not allow for
unbalanced datasets, data where the equations contain different
numbers of observations, future releases of the package will have that
feature implemented. We are still trying to figure out how develop the
code for instrumental variable methods (2SLS,3SLS) for the correct
estimation methods if possible.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Additional Methods}\label{sec:summary_additional_methods}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsubsection{Full Information Maximum Likelihood}

Describe, in a paragraph the process and give some reasons why you
might want this in the package.

\subsubsection{Generalized Method of Moments}

Describe, in a paragraph the process and give some reasons why you
might want this in the package.

### 2005-09-01 paper/trunk/systemfit_usage.tex
revised structure of the document (sections, subsections, subsubsections) due to removal of nlsystemfit; moved each section into a separate file.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@571 591db2bd-e0a0-4d22-8c19-2192dbb7da31

--- 2005-09-01 paper/trunk/nlsystemfit.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Estimating nonlinear equation systems}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-The \pkg{systemfit} package also contains a method for fitting
-nonlinear systems of equations.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Nonlinear parameter estimation}
-\label{sec:nonlinear_estimation}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-A system of nonlinear equations can be written as:
-
-\begin{equation}
-  \label{eq:non_linear_eq_1}
-  \epsilon_{t} = q( y_t, x_t, \theta )
-\end{equation}
-%%% It is not clear what these variable are. 
-%%% Adding an equation representing only a single equation
-%%% would make things clearer.
-
-\noindent and
-
-\begin{equation}
-  \label{eq:non_linear_eq_2}
-  z_{t} = Z( x_t )
-\end{equation}
-
-where $\epsilon_{t}$ are the residuals from the y observations and
-$z_{t}$ are the function evaluated at the parameter estimates.
-
-\textbf{you really need to check this!!!}
-
-The objective functions for the methods are:
-
-\begin{center}
-\begin{tabular}{|l|c|c|c|} \hline
-  Method & Instruments & Objective Function & Covariance of $\theta$ \\ \hline  
-  OLS & No & $r'r$ & $(X(diag(S)^{-1}\otimes I)X)^{-1}$ \\ \hline
-  SUR & No & $r'(diag(S)_{OLS}^{-1}\otimes I)r$ & $(X(S^{-1}\otimes I)X)^{-1}$ \\ \hline
-  2SLS & Yes & $r'(I \otimes W)r$ & $(X(diag(S)^{-1}\otimes I)X)^{-1}$ \\ \hline 
-  3SLS & Yes & $r'(S_{2SLS}^{-1} \otimes W)r$ & $(X(diag(S)^{-1}\otimes W)X)^{-1}$ \\ \hline
-\end{tabular}
-\end{center}
-
-where, $r$ is a column vector for the residuals for each equation,
-%% what is the difference to \epsilon?
-$X$ is matrix of the partial derivatives of the dependent variable 
-with respect to the parameters $\left( \frac{ \partial y }{ \theta} \right)$,
-%% is this correct?
-%% in the linear section X is used for the regressors, 
-%% thus we should use a different variable name here.
-$W$ is a matrix of the instrument variables, $Z(Z'Z)^{-1}Z$, $Z$ is a
-matrix of the instrument variables, and $I$ is an $n \times n $
-identity matrix and $S$ is the estimated variance-covariance matrix between the
-equations
-
-\begin{equation}
-  \label{eq:non-linear_varcov}
-  \hat{\sigma}_{ij} = (\hat{e}_i' \hat{e}_j) / \sqrt{(T - k_i)*(T- k_j)} 
-\end{equation}
-
-\textbf{The residual variance-covariance matrix can be calculated in 
-different ways (section~\ref{sec:residcov}). 
-It should be relatively easy to implement this also in nlsystemfit(), 
-e.g. the function 'calcRCov' in systemfit() could be moved outside
-systemfit() that it can be used also by nlsystemfit().}
-
-\textbf{You need to clear this up.}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Using systemfit}\label{sec:Usage}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Using nlsystemfit}\label{sec:UsingnlSystemfit}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Standard usage}
 
-\code {nlsystemfit} fits a set of structural nonlinear equations using
-Ordinary Least Squares (OLS), Seemingly Unrelated Regression (SUR),
-Two-Stage Least Squares (2SLS), Three-Stage Least Squares (3SLS) using
-the objective functions described in section
-\ref{sec:nonlinear_estimation}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Standard usage}
-
-Similar to calling the \code{systemfit} function, \code{nlsystemfit}
-is called with a minimum of three arguments,
+\pkg{systemfit} is generally called by
 
 \code{
-R> nlsystemfit( method, eqns, start )
+R> systemfit( method, eqns )
 }
 
-where \code{method} is one of the following estimation methods: "OLS",
-"SUR", "2SLS", or "3SLS", \code{eqns} is a list of equations similar
-to those described for \code{systemfit}, and \code{start} is a list of
-starting values of the parameter estimates.
+There are two mandatory arguments: \code{method} and \code{eqns}.
 
-% The mandatory argument \code{eqns} is a list of the equations  to
-% estimate.  Each equation is a standard formula in \proglang{R}.  It
-% starts with a dependent variable on the left hand side.  After a tilde
-% ($~$) the regressors are listed  \footnote{For Details see the
-%   \proglang{R} help files to \code{formula}}.
+The argument \code{method} is a string determining the estimation method.
+It must be one of "OLS", "WLS", "SUR", "2SLS", "W2SLS" or "3SLS".
+
+The other mandatory argument \code{eqns} is a list of the equations 
+to estimate. 
+Each equation is a standard formula in \proglang{R}.
+It starts with a dependent variable on the left hand side.
+After a tilde ($\sim$) the regressors are listed%
+\footnote{For Details see the \proglang{R} help files to \code{formula}}.
 
 This is now demonstrated using an example: \\
-
-% \code{
-% R> library( systemfit ) \\
-% R> data( ppine ) \\
-% R> hg.formula <- hg $\sim$ exp( h0 + h1*log(tht) + h2*tht$\hat$2 + h3*elev + h4*cr) \\
-% R> dg.formula <- dg $\sim$ exp( d0 + d1*log(dbh) + d2*hg + d3*cr + d4*ba  ) \\
-% R> labels <- list( "height.growth", "diameter.growth" )\\
-% R> inst <- $\sim$ tht + dbh + elev + cr + ba\\
-% R> start.values <- c(h0=-0.5, h1=0.5, h2=-0.001, h3=0.0001, h4=0.08,\\
-% R+   d0=-0.5, d1=0.009, d2=0.25, d3=0.005, d4=-0.02 )\\
-% R> model <- list( hg.formula, dg.formula )\\
-% R> model.3sls <- nlsystemfit( "3SLS", model, start.values, data=ppine,\\
-% R+   eqnlabels=labels, inst=inst ) \\
-% R> summary( model.3sls ) \\
-% }
-
-The nlsystemfit function relies on \code{nlm} to perform the
-minimization of the objective functions and the \code{qr} set of
-functions. If the user does not have a set of estimates for the
-initial parameters, it is suggested using linearized forms in one of
-the linear methods, or simply using \code{nls} to obtain estimates for
-the equations.
-
-The outputs are similar to those for a systemfit object.
-
-\textbf{Describe the code and outputs as Arne did in the previous
-  section.}
+\code{
+R> library( systemfit ) \\
+R> data( kmenta ) \\
+R> attach( kmenta ) \\
+R> fitsur <- systemfit( "SUR", list( q $\sim$ p + d, q $\sim$ p + f + a ) ) \\
+}
 
@@
 The regression result is assigned to the variable \code{fitsur}.
-
-%% Arne -- Did you put the ~'s in by hand?
-%% Yes, I did. I used find and replace to substitute the "~" for " ". 
-%% This is necessary to retain the format of the output.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Other issues}
+\subsection{User options}
 
-\subsubsection{nlsystemfit issue 1}
-% The user should be aware that the function is \bold{VERY} sensative to
-% the starting values and the nlm function may not converge.
+Following additional options can be set by the user:
 
-% did you try optim()? Especially if you provide (analytical) derivatives
-% it is very good in finding the optimum.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% \subsubsection{Equation labels}
+\paragraph{Equation labels}
+The optional argument \code{eqnlabels} allows the user to label the equations.
+It has to be a vector of strings naming the equations.\\
+\code{
+R>~fitsur~<-~systemfit(~"SUR",~list(~q~~~p~+~d,~q~~~p~+~f~+~a~),\\~
+R+~~~~eqnlabels~=~c(~"demand",~"supply"~)~)\\
+R>~summary(~fitsur~)\\
+systemfit~results\\
+method:~SUR\\
+\\
+\mbox{}~~~~~~~~N~DF~~~~~~SSR~~~~~MSE~~~~RMSE~~~~~~~R2~~~Adj~R2\\
+demand~20~17~~65.6829~3.86370~1.96563~0.755019~0.726198\\
+supply~20~16~104.0584~6.50365~2.55023~0.611888~0.539117\\
+\ldots\\
+}
+If no equation labels are provided, the equations are numbered.
 
-% The nlm
-% function will be called with the \code{typsize} argument set the
-% absolute values of the starting values for the OLS and 2SLS
-% methods. For the SUR and 3SLS methods, the \code{typsize} argument is
-% set to the absolute values of the resulting OLS and 2SLS parameter
-% estimates from the nlm result structre. In addition, the starting
-% values for the SUR and 3SLS methods are obtained from the OLS and 2SLS
-% parameter estimates to shorten the number of iterations. The number of
-% iterations reported in the summary are only those used in the last
-% call to nlm, thus the number of iterations in the OLS portion of the
-% SUR fit and the 2SLS portion of the 3SLS fit are not included.  }
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\subsubsection{Instrumental variables}   
+\paragraph{Instrumental variables}   
+\code{inst}one-sided model formula specifying instrumental variables
+   or a list of one-sided model formulas if different instruments should
+   be used for the different equations (only needed for 2SLS, W2SLS and
+   3SLS estimations).
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\subsubsection{Data}   
+\paragraph{Data}   
+\code{data} an optional data frame containing the variables in the model.
+   By default the variables are taken from the environment from which
+   systemfit is called.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\subsubsection{Restrictions}   
+\paragraph{Restrictions}   
+\code{R.restr} an optional j x k matrix to impose linear
+   restrictions on the parameters by \code{R.restr} * $\beta$ = \code{q.restr}
+   (j = number of restrictions, k = number of all parameters,
+   $\beta$ = vector of all parameters).
+
+\code{q.restr} an optional j x 1 matrix to impose linear
+   restrictions (see \code{R.restr}); default is a j x 1 matrix
+   that contains only zeros.
+
+\code{TX} an optional matrix to transform the regressor matrix and,
+   hence, also the coefficient vector (see details).
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\subsubsection{Iteration control}   
+\paragraph{Iteration control}
+\code{maxiter} maximum number of iterations for WLS, SUR, W2SLS and
+   3SLS estimations.
+
+\code{tol} tolerance level indicating when to stop the iteration (only
+   WLS, SUR, W2SLS and 3SLS estimations).
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\subsubsection{Residual covariance matrix}   
+\paragraph{Residual covariance matrix}   
+\code{rcovformula} formula to calculate the estimated residual covariance
+   matrix (see details).
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\subsubsection{3SLS formula}   
+\paragraph{3SLS formula}   
+\code{formula3sls} formula for calculating the 3SLS estimator,
+   one of "GLS", "IV", "GMM", "Schmidt" or "EViews" (see details).
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\subsubsection{Degrees of freedom for t-tests}   
+\paragraph{Degrees of freedom for t-tests}   
+\code{probdfsys} use the degrees of freedom of the whole system
+   (in place of the degrees of freedom of the single equation)
+   to calculate prob values for the t-test of individual parameters.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\subsubsection{Sigma squared}   
+\paragraph{Sigma squared}   
+\code{single.eq.sigma} use different $\sigma^2$s for each
+   single equation to calculate the covariance matrix and the
+   standard errors of the coefficients (only OLS and 2SLS).
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\subsubsection{System options}   
+\paragraph{System options}
+
+\code{solvetol} tolerance level for detecting linear dependencies
+   when inverting a matrix or calculating a determinant (see
+   \code{solve} and \code{det}).
+
+\code{saveMemory} save memory by omitting some calculation that
+   are not crucial for the basic estimation (e.g McElroy's
+   $R^2$).

### 2005-09-07 paper/trunk/systemfit_intro.tex
systemfit_intro.tex: modified/extended introduction
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@572 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 approach. Especially the economic theory suggests many cross-equation
-restriction.
+restrictions.
 
 The \pkg{systemfit} package provides the capability to estimate
-linear and non-linear equation systems in \proglang{R}. Compared to
-other statistics and econometrics software the (advanced) user can
-control many important details of the estimation procedure. On the
-other hand these user options have reasonable defaults that also
+linear and non-linear equation systems in \proglang{R}.
+Although linear and non-linear equation systems can be estimated
+with several other statistical and econometric software packages
+(e.g. \proglang{SAS}, \proglang{EViews}, \proglang{TSP}),
+using \pkg{systemfit} has several advantages.
+First, all estimation procedures are publicly available in the source code.
+Second, the estimation algorithms can be easily modified to meet specific
+requirements.
+Third, the (advanced) user can control many important details of the
+estimation procedure that are generally not available in other software
+packages.
+On the other hand these user options have reasonable defaults that also
 beginners can easily use the provided functions.

### 2005-09-26 paper/trunk/systemfit.tex
some minor changes
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@573 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 $Log$
+Revision 1.12  2005/09/26 11:24:05  henningsena
+some minor changes
+
 Revision 1.11  2005/09/01 12:27:55  henningsena

### 2005-09-26 paper/trunk/systemfit_estimation.tex
some minor changes
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@573 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Ordinary Least Squares estimation}
+\subsection{Ordinary least squares (OLS)}
 
@@
 
-If no cross-equation restrictions are imposed, the simultaneous 
+If no cross-equation parameter restrictions are imposed, the simultaneous 
 OLS estimation of the system leads to the same parameter estimates 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Weighted least squares}
+\subsection{Weighted least squares (WLS)}
 
@@
 
-If no cross-equation restrictions are imposed, 
-the parameter estimates and the variance-covariance matrix 
-of the parameters are equal to the OLS estimates%
-\footnote{The variance-covariance matrices of the 
-estimated parameters might differ, 
-if the WLS estimation is based on a different formula 
-to calculate the (diagonal) variance-covariance 
-matrix of the residuals (see section \ref{sec:residcov}).}.
+If no cross-equation parameter restrictions are imposed, 
+the parameter estimates are equal to the OLS estimates.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Seemingly unrelated regression}
+\subsection{Seemingly unrelated regression (SUR)}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Two-stage least squares estimation}
+\subsection{Two-stage least squares (2SLS)}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Weighted two-stage least squares estimation}
+\subsection{Weighted two-stage least squares (W2SLS)}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Three-stage least squares estimation}
+\subsection{Three-stage least squares (3SLS)}
 
@@
 (equation \ref{eq:beta2sls}).
-EViews uses the standard 3SLS formula (equation (\ref{eq:cov3sls})) to 
+EViews uses the standard 3SLS formula (\ref{eq:cov3sls}) to
 calculate the variance-covariance matrix of the 3SLS estimator.
@@
 every instrument in every equation \citep{schmidt90}. 
-In this case, only the GMM-3SLS (equation (\ref{eq:beta3slsGMM}))
+In this case, only the GMM-3SLS (\ref{eq:beta3slsGMM})
 and the 3SLS estimator developed by \cite{schmidt90} 
-(equation (\ref{eq:beta3slsSchmidt})) are consistent.
+(\ref{eq:beta3slsSchmidt}) are consistent.

### 2005-09-26 paper/trunk/systemfit_restrictions.tex
some minor changes
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@573 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 \end{equation}
-where $\beta^*$ is a vector of linear independent coefficients.
+where $\beta^*$ is a vector of restricted (linear independent) coefficients,
+and $T$ is a matrix with the number of rows equal to the number of
+unrestricted coefficients ($\beta$) and
+the number of columns equal to the number of restricted coefficients
+($\beta^*$).
+$T$ can be used to map each unrestricted coefficient to one or more
+restricted coefficients.
 
@@
 The first way is less flexible than this latter one%
-\footnote{e.g. restrictions like $\beta_1 + \beta_2 = 4$ cannot be imposed
-by the first method}, 
+\footnote{e.g.\ restrictions like $\beta_1 + \beta_2 = 4$ cannot be imposed
+by the first method.}, 
 but the first way is preferable if equality constraints for coefficients
@@
 EViews uses the standard formula of the restricted 3SLS estimator
-(equation (\ref{eq:cov3slsr})).
+(\ref{eq:cov3slsr}).
 

### 2005-09-26 paper/trunk/systemfit_estimation.tex
using now package csquotes to have uniform and correct quotation marks (BibTeX-Style jss.bst has to be modified)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@576 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 
-We can write the 'stacked' system as
+We can write the °stacked° system as
 \begin{equation}
@@
 Least Squares (GLS) estimation leads to efficient parameter estimates.
-In this case the GLS is generally called "Seemingly Unrelated Regression" 
+In this case the GLS is generally called °Seemingly Unrelated Regression°
 (SUR) (see \citealp{zellner62}). 
@@
 estimation leads to consistent and efficient estimates. This estimation 
-procedure is generally called "Three-stage Least Squares" (3SLS).
+procedure is generally called °Three-stage Least Squares° (3SLS).
 

### 2005-10-26 paper/trunk/systemfit.tex
using now package csquotes to have uniform and correct quotation marks (BibTeX-Style jss.bst has to be modified)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@576 591db2bd-e0a0-4d22-8c19-2192dbb7da31
checked out and updated local copy before edits
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@577 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 \documentclass[article]{jss}
+\usepackage{csquotes}
+\MakeOuterQuote{°}
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Standard usage}
+
+\pkg{systemfit} is generally called by
+
+\code{
+R> systemfit( method, eqns )
+}
+
+There are two mandatory arguments: \code{method} and \code{eqns}.
+
+The argument \code{method} is a string determining the estimation method.
+It must be one of "OLS", "WLS", "SUR", "2SLS", "W2SLS" or "3SLS".
+
+The other mandatory argument \code{eqns} is a list of the equations 
+to estimate. 
+Each equation is a standard formula in \proglang{R}.
+It starts with a dependent variable on the left hand side.
+After a tilde ($\sim$) the regressors are listed%
+\footnote{For Details see the \proglang{R} help files to \code{formula}}.
+
+This is now demonstrated using an example: \\
+\code{
+R> library( systemfit ) \\
+R> data( kmenta ) \\
+R> attach( kmenta ) \\
+R> fitsur <- systemfit( "SUR", list( q $\sim$ p + d, q $\sim$ p + f + a ) ) \\
+}
+
+The first line loads the \pkg{systemfit} package. 
+The second line loads example data that are included in this package.
+These data come from \cite{kmenta86}.
+They are attached to the \proglang{R} search path in line three.
+In the last line a seemingly unrelated regression is done.
+The first equation represents the demand side of the food market.
+The dependant variable is \code{q} (food consumption per capita). 
+The regressors are \code{p} (ratio of food prices to general consumer prices)
+and \code{d} (disposable income) as well as a constant%
+\footnote{a regression constant is always implied if not explicitly omitted.}.
+The second equation represents the supply side.
+Variable \code{q} (food consumption per capita) is also the dependant 
+variable of this equation. 
+The regressors are again \code{p} (ratio of food prices to general 
+consumer prices) and a constant as well as 
+\code{f} (ratio of preceding year's prices received by farmers) and 
+\code{a} (a time trend in years).
+The regression result is assigned to the variable \code{fitsur}.
+
+Summary results can be printed by\\
+\code{
+R>~summary(~fitsur~)~\\
+~\\
+systemfit~results~\\
+method:~SUR~\\
+\\
+\mbox{}~~~N~DF~~~~~~SSR~~~~~MSE~~~~RMSE~~~~~~~R2~~~Adj~R2 \\
+1~20~17~~65.6829~3.86370~1.96563~0.755019~0.726198~\\
+2~20~16~104.0584~6.50365~2.55023~0.611888~0.539117~\\
+~\\
+The~covariance~matrix~of~the~residuals~used~for~estimation\\
+\mbox{}~~~~~~~~1~~~~~~~2~\\
+1~3.72539~4.13696~\\
+2~4.13696~5.78444~\\
+~\\
+The~covariance~matrix~of~the~residuals\\
+\mbox{}~~~~~~~~1~~~~~~~2~\\
+1~3.86370~4.92431~\\
+2~4.92431~6.50365~\\
+~\\
+The~correlations~of~the~residuals\\
+\mbox{}~~~~~~~~~1~~~~~~~~2~\\
+1~1.000000~0.982348~\\
+2~0.982348~1.000000~\\
+~\\
+The~determinant~of~the~residual~covariance~matrix:~0.879285~\\
+OLS~R-squared~value~of~the~system:~0.683453~\\
+McElroy's~R-squared~value~for~the~system:~0.788722~\\
+~\\
+SUR~estimates~for~1~~(equation~1~)~\\
+Model~Formula:~q~~~p~+~d\\
+\\
+\mbox{}~~~~~~~~~~~~~Estimate~Std.~Error~~~t~value~Pr(>|t|)~\\
+(Intercept)~99.332894~~~7.514452~13.218913~~~~~~~~0~***~\\
+p~~~~~~~~~~~-0.275486~~~0.088509~-3.112513~0.006332~~**~\\
+d~~~~~~~~~~~~~0.29855~~~0.041945~~7.117605~~~~2e-06~***~\\
+---~\\
+Signif.~codes:~~0~`***'~0.001~`**'~0.01~`*'~0.05~`.'~0.1~`~'~1~\\
+~\\
+Residual~standard~error:~1.96563~on~17~degrees~of~freedom~\\
+Number~of~observations:~20~Degrees~of~Freedom:~17~\\
+SSR:~65.682902~MSE:~3.8637~Root~MSE:~1.96563~\\
+Multiple~R-Squared:~0.755019~Adjusted~R-Squared:~0.726198~\\
+~\\
+~\\
+SUR~estimates~for~2~~(equation~2~)~\\
+Model~Formula:~q~~~p~+~f~+~a\\
+\\
+\mbox{}~~~~~~~~~~~~~Estimate~Std.~Error~~t~value~Pr(>|t|)~\\
+(Intercept)~61.966166~~~11.08079~5.592215~~~~4e-05~***~\\
+p~~~~~~~~~~~~0.146884~~~0.094435~1.555397~0.139408~\\
+f~~~~~~~~~~~~0.214004~~~0.039868~5.367761~~6.3e-05~***~\\
+a~~~~~~~~~~~~0.339304~~~0.067911~4.996283~0.000132~***~\\
+---~\\
+Signif.~codes:~~0~`***'~0.001~`**'~0.01~`*'~0.05~`.'~0.1~`~'~1~\\
+~\\
+Residual~standard~error:~2.550226~on~16~degrees~of~freedom~\\
+Number~of~observations:~20~Degrees~of~Freedom:~16~\\
+SSR:~104.05843~MSE:~6.503652~Root~MSE:~2.550226~\\
+Multiple~R-Squared:~0.611888~Adjusted~R-Squared:~0.539117~\\
+}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{User options}
+
+Following additional options can be set by the user:
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% \subsubsection{Equation labels}
+\paragraph{Equation labels}
+The optional argument \code{eqnlabels} allows the user to label the equations.
+It has to be a vector of strings naming the equations.\\
+\code{
+R>~fitsur~<-~systemfit(~"SUR",~list(~q~~~p~+~d,~q~~~p~+~f~+~a~),\\~
+R+~~~~eqnlabels~=~c(~"demand",~"supply"~)~)\\
+R>~summary(~fitsur~)\\
+systemfit~results\\
+method:~SUR\\
+\\
+\mbox{}~~~~~~~~N~DF~~~~~~SSR~~~~~MSE~~~~RMSE~~~~~~~R2~~~Adj~R2\\
+demand~20~17~~65.6829~3.86370~1.96563~0.755019~0.726198\\
+supply~20~16~104.0584~6.50365~2.55023~0.611888~0.539117\\
+\ldots\\
+}
+If no equation labels are provided, the equations are numbered.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\subsubsection{Instrumental variables}   
+\paragraph{Instrumental variables}   
+\code{inst}one-sided model formula specifying instrumental variables
+   or a list of one-sided model formulas if different instruments should
+   be used for the different equations (only needed for 2SLS, W2SLS and
+   3SLS estimations).
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\subsubsection{Data}   
+\paragraph{Data}   
+\code{data} an optional data frame containing the variables in the model.
+   By default the variables are taken from the environment from which
+   systemfit is called.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\subsubsection{Restrictions}   
+\paragraph{Restrictions}   
+\code{R.restr} an optional j x k matrix to impose linear
+   restrictions on the parameters by \code{R.restr} * $\beta$ = \code{q.restr}
+   (j = number of restrictions, k = number of all parameters,
+   $\beta$ = vector of all parameters).
+
+\code{q.restr} an optional j x 1 matrix to impose linear
+   restrictions (see \code{R.restr}); default is a j x 1 matrix
+   that contains only zeros.
+
+\code{TX} an optional matrix to transform the regressor matrix and,
+   hence, also the coefficient vector (see details).
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\subsubsection{Iteration control}   
+\paragraph{Iteration control}
+\code{maxiter} maximum number of iterations for WLS, SUR, W2SLS and
+   3SLS estimations.
+
+\code{tol} tolerance level indicating when to stop the iteration (only
+   WLS, SUR, W2SLS and 3SLS estimations).
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\subsubsection{Residual covariance matrix}   
+\paragraph{Residual covariance matrix}   
+\code{rcovformula} formula to calculate the estimated residual covariance
+   matrix (see details).
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\subsubsection{3SLS formula}   
+\paragraph{3SLS formula}   
+\code{formula3sls} formula for calculating the 3SLS estimator,
+   one of "GLS", "IV", "GMM", "Schmidt" or "EViews" (see details).
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\subsubsection{Degrees of freedom for t-tests}   
+\paragraph{Degrees of freedom for t-tests}   
+\code{probdfsys} use the degrees of freedom of the whole system
+   (in place of the degrees of freedom of the single equation)
+   to calculate prob values for the t-test of individual parameters.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\subsubsection{Sigma squared}   
+\paragraph{Sigma squared}   
+\code{single.eq.sigma} use different $\sigma^2$s for each
+   single equation to calculate the covariance matrix and the
+   standard errors of the coefficients (only OLS and 2SLS).
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\subsubsection{System options}   
+\paragraph{System options}
+
+\code{solvetol} tolerance level for detecting linear dependencies
+   when inverting a matrix or calculating a determinant (see
+   \code{solve} and \code{det}).
+
+\code{saveMemory} save memory by omitting some calculation that
+   are not crucial for the basic estimation (e.g McElroy's
+   $R^2$).
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Estimating nonlinear equation systems}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+The \pkg{systemfit} package also contains a method for fitting
+nonlinear systems of equations.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Nonlinear parameter estimation}
+\label{sec:nonlinear_estimation}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+A system of nonlinear equations can be written as:
+
+\begin{equation}
+  \label{eq:non_linear_eq_1}
+  \epsilon_{t} = q( y_t, x_t, \beta )
+\end{equation}
+%%% It is not clear what these variable are. 
+%%% Adding an equation representing only a single equation
+%%% would make things clearer.
+
+\noindent and
+
+\begin{equation}
+  \label{eq:non_linear_eq_2}
+  z_{t} = Z( x_t )
+\end{equation}
+
+where $\epsilon_{t}$ are the residuals from the y observations and
+$z_{t}$ are the function evaluated at the parameter estimates.
+
+\textbf{you really need to check this!!!}
+
+The objective functions for the methods are:
+
+\begin{center}
+\begin{tabular}{|l|c|c|c|} \hline
+  Method & Instruments & Objective Function & Covariance of $\theta$ \\ \hline  
+  OLS & No & $r'r$ & $(X(diag(S)^{-1}\otimes I)X)^{-1}$ \\ \hline
+  SUR & No & $r'(diag(S)_{OLS}^{-1}\otimes I)r$ & $(X(S^{-1}\otimes I)X)^{-1}$ \\ \hline
+  2SLS & Yes & $r'(I \otimes W)r$ & $(X(diag(S)^{-1}\otimes I)X)^{-1}$ \\ \hline 
+  3SLS & Yes & $r'(S_{2SLS}^{-1} \otimes W)r$ & $(X(diag(S)^{-1}\otimes W)X)^{-1}$ \\ \hline
+\end{tabular}
+\end{center}
+
+where, $r$ is a column vector for the residuals for each equation,
+%% what is the difference to \epsilon?
+$X$ is matrix of the partial derivatives of the dependent variable 
+with respect to the parameters $\left( \frac{ \partial y }{ \theta} \right)$,
+%% is this correct?
+%% in the linear section X is used for the regressors, 
+%% thus we should use a different variable name here.
+$W$ is a matrix of the instrument variables, $Z(Z'Z)^{-1}Z$, $Z$ is a
+matrix of the instrument variables, and $I$ is an $n \times n $
+identity matrix and $S$ is the estimated variance-covariance matrix between the
+equations
+
+\begin{equation}
+  \label{eq:non-linear_varcov}
+  \hat{\sigma}_{ij} = (\hat{e}_i' \hat{e}_j) / \sqrt{(T - k_i)*(T- k_j)} 
+\end{equation}
+
+\textbf{The residual variance-covariance matrix can be calculated in 
+different ways (section~\ref{sec:residcov}). 
+It should be relatively easy to implement this also in nlsystemfit(), 
+e.g. the function 'calcRCov' in systemfit() could be moved outside
+systemfit() that it can be used also by nlsystemfit().}
+
+\textbf{You need to clear this up.}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Using nlsystemfit}\label{sec:UsingnlSystemfit}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\code {nlsystemfit} fits a set of structural nonlinear equations using
+Ordinary Least Squares (OLS), Seemingly Unrelated Regression (SUR),
+Two-Stage Least Squares (2SLS), Three-Stage Least Squares (3SLS) using
+the objective functions described in section
+\ref{sec:nonlinear_estimation}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Standard usage}
+
+Similar to calling the \code{systemfit} function, \code{nlsystemfit}
+is called with a minimum of three arguments,
+
+\code{
+R> nlsystemfit( method, eqns, start )
+}
+
+where \code{method} is one of the following estimation methods: "OLS",
+"SUR", "2SLS", or "3SLS", \code{eqns} is a list of equations similar
+to those described for \code{systemfit}, and \code{start} is a list of
+starting values of the parameter estimates.
+
+% The mandatory argument \code{eqns} is a list of the equations  to
+% estimate.  Each equation is a standard formula in \proglang{R}.  It
+% starts with a dependent variable on the left hand side.  After a tilde
+% ($~$) the regressors are listed  \footnote{For Details see the
+%   \proglang{R} help files to \code{formula}}.
+
+This is now demonstrated using an example: \\
+
+% \code{
+% R> library( systemfit ) \\
+% R> data( ppine ) \\
+% R> hg.formula <- hg $\sim$ exp( h0 + h1*log(tht) + h2*tht$\hat$2 + h3*elev + h4*cr) \\
+% R> dg.formula <- dg $\sim$ exp( d0 + d1*log(dbh) + d2*hg + d3*cr + d4*ba  ) \\
+% R> labels <- list( "height.growth", "diameter.growth" )\\
+% R> inst <- $\sim$ tht + dbh + elev + cr + ba\\
+% R> start.values <- c(h0=-0.5, h1=0.5, h2=-0.001, h3=0.0001, h4=0.08,\\
+% R+   d0=-0.5, d1=0.009, d2=0.25, d3=0.005, d4=-0.02 )\\
+% R> model <- list( hg.formula, dg.formula )\\
+% R> model.3sls <- nlsystemfit( "3SLS", model, start.values, data=ppine,\\
+% R+   eqnlabels=labels, inst=inst ) \\
+% R> summary( model.3sls ) \\
+% }
+
+The nlsystemfit function relies on \code{nlm} to perform the
+minimization of the objective functions and the \code{qr} set of
+functions. If the user does not have a set of estimates for the
+initial parameters, it is suggested using linearized forms in one of
+the linear methods, or simply using \code{nls} to obtain estimates for
+the equations.
+
+The outputs are similar to those for a systemfit object.
+
+\textbf{Describe the code and outputs as Arne did in the previous
+  section.}
+
+The first line loads the \pkg{systemfit} package. 
+The second line loads example data that are included in this package.
+These data come from \cite{kmenta86}.
+They are attached to the \proglang{R} search path in line three.
+In the last line a seemingly unrelated regression is done.
+The first equation represents the demand side of the food market.
+The dependant variable is \code{q} (food consumption per capita). 
+The regressors are \code{p} (ratio of food prices to general consumer prices)
+and \code{disposable income} as well as a constant%
+\footnote{a regression constant is always implied if not explicitly omitted.}.
+The second equation represents the supply side.
+Variable \code{q} (food consumption per capita) is also the dependant 
+variable of this equation. 
+The regressors are again \code{p} (ratio of food prices to general 
+consumer prices) and a constant as well as 
+\code{f} (ratio of preceding year's prices received by farmers) and 
+\code{a} (a time trend in years).
+The regression result is assigned to the variable \code{fitsur}.
+
+%% Arne -- Did you put the ~'s in by hand?
+%% Yes, I did. I used find and replace to substitute the "~" for " ". 
+%% This is necessary to retain the format of the output.
+
+Summary results can be printed by\\
+\code{
+R>~summary(~fitsur~)~\\
+~\\
+systemfit~results~\\
+method:~SUR~\\
+\\
+\mbox{}~~~N~DF~~~~~~SSR~~~~~MSE~~~~RMSE~~~~~~~R2~~~Adj~R2 \\
+1~20~17~~65.6829~3.86370~1.96563~0.755019~0.726198~\\
+2~20~16~104.0584~6.50365~2.55023~0.611888~0.539117~\\
+~\\
+The~covariance~matrix~of~the~residuals~used~for~estimation\\
+\mbox{}~~~~~~~~1~~~~~~~2~\\
+1~3.72539~4.13696~\\
+2~4.13696~5.78444~\\
+~\\
+The~covariance~matrix~of~the~residuals\\
+\mbox{}~~~~~~~~1~~~~~~~2~\\
+1~3.86370~4.92431~\\
+2~4.92431~6.50365~\\
+~\\
+The~correlations~of~the~residuals\\
+\mbox{}~~~~~~~~~1~~~~~~~~2~\\
+1~1.000000~0.982348~\\
+2~0.982348~1.000000~\\
+~\\
+The~determinant~of~the~residual~covariance~matrix:~0.879285~\\
+OLS~R-squared~value~of~the~system:~0.683453~\\
+McElroy's~R-squared~value~for~the~system:~0.788722~\\
+~\\
+SUR~estimates~for~1~~(equation~1~)~\\
+Model~Formula:~q~~~p~+~d\\
+\\
+\mbox{}~~~~~~~~~~~~~Estimate~Std.~Error~~~t~value~Pr(>|t|)~\\
+(Intercept)~99.332894~~~7.514452~13.218913~~~~~~~~0~***~\\
+p~~~~~~~~~~~-0.275486~~~0.088509~-3.112513~0.006332~~**~\\
+d~~~~~~~~~~~~~0.29855~~~0.041945~~7.117605~~~~2e-06~***~\\
+---~\\
+Signif.~codes:~~0~`***'~0.001~`**'~0.01~`*'~0.05~`.'~0.1~`~'~1~\\
+~\\
+Residual~standard~error:~1.96563~on~17~degrees~of~freedom~\\
+Number~of~observations:~20~Degrees~of~Freedom:~17~\\
+SSR:~65.682902~MSE:~3.8637~Root~MSE:~1.96563~\\
+Multiple~R-Squared:~0.755019~Adjusted~R-Squared:~0.726198~\\
+~\\
+~\\
+SUR~estimates~for~2~~(equation~2~)~\\
+Model~Formula:~q~~~p~+~f~+~a\\
+\\
+\mbox{}~~~~~~~~~~~~~Estimate~Std.~Error~~t~value~Pr(>|t|)~\\
+(Intercept)~61.966166~~~11.08079~5.592215~~~~4e-05~***~\\
+p~~~~~~~~~~~~0.146884~~~0.094435~1.555397~0.139408~\\
+f~~~~~~~~~~~~0.214004~~~0.039868~5.367761~~6.3e-05~***~\\
+a~~~~~~~~~~~~0.339304~~~0.067911~4.996283~0.000132~***~\\
+---~\\
+Signif.~codes:~~0~`***'~0.001~`**'~0.01~`*'~0.05~`.'~0.1~`~'~1~\\
+~\\
+Residual~standard~error:~2.550226~on~16~degrees~of~freedom~\\
+Number~of~observations:~20~Degrees~of~Freedom:~16~\\
+SSR:~104.05843~MSE:~6.503652~Root~MSE:~2.550226~\\
+Multiple~R-Squared:~0.611888~Adjusted~R-Squared:~0.539117~\\
+}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Other issues}
+
+\subsubsection{nlsystemfit issue 1}
+% The user should be aware that the function is \bold{VERY} sensative to
+% the starting values and the nlm function may not converge.
+
+% did you try optim()? Especially if you provide (analytical) derivatives
+% it is very good in finding the optimum.
+
+% The nlm
+% function will be called with the \code{typsize} argument set the
+% absolute values of the starting values for the OLS and 2SLS
+% methods. For the SUR and 3SLS methods, the \code{typsize} argument is
+% set to the absolute values of the resulting OLS and 2SLS parameter
+% estimates from the nlm result structre. In addition, the starting
+% values for the SUR and 3SLS methods are obtained from the OLS and 2SLS
+% parameter estimates to shorten the number of iterations. The number of
+% iterations reported in the summary are only those used in the last
+% call to nlm, thus the number of iterations in the OLS portion of the
+% SUR fit and the 2SLS portion of the 3SLS fit are not included.  }
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Other tools}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Likelihood ratio test}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Hausman test}
+
+\textbf{Man, I really have to clean this up.}
+
+Hausman \citep{hausman1978} developed a test have power against
+alternatives for which $\hat{\theta}$ and $\tilde{\theta}$ diverge
+under misspecification.
+
+The Hausman result if the difference between the covariance of an
+efficient and the covariance of the inefficient estimator is zero, the
+model is correctly specified.
+
+The test is based on the idea that under the hypothesis of no
+correlation, both the limited information estimates and the
+full-information results are consistent, but only the full-information
+estimates are efficient and under the alternative, the limited
+information estimates are consistent, but the full information
+estimates are not.
+
+Since the models that are fit using the \pkg{systemfit} package are
+typically random effects models, the effects may be inconsistent
+between the introduced variables and the regressors.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Summary and outlook}\label{sec:Summmary}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+What should we place in this section?
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Unbalanced Observations}\label{sec:summary_unbalanced_data}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+The \pkg{systemfit} package was originally developed to fit
+simultaneous equations for forestry datasets. Forestry datasets
+typically contain observations for many inexpensive observations
+(diameter at breast height) and very few expensive observations such
+as total tree height. While the package currently does not allow for
+unbalanced datasets, data where the equations contain different
+numbers of observations, future releases of the package will have that
+feature implemented. We are still trying to figure out how develop the
+code for instrumental variable methods (2SLS,3SLS) for the correct
+estimation methods if possible.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Additional Methods}\label{sec:summary_additional_methods}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsubsection{Full Information Maximum Likelihood}
+
+Describe, in a paragraph the process and give some reasons why you
+might want this in the package.
+
+\subsubsection{Generalized Method of Moments}
+
+Describe, in a paragraph the process and give some reasons why you
+might want this in the package.
+
 \section{CVS Revision Log}
@@
 $Log$
+Revision 1.14  2005/10/26 21:16:53  hamannj
+checked out and updated local copy before edits
+
+Revision 1.13  2005/09/26 20:34:12  henningsena
+using now package csquotes to have uniform and correct quotation marks (BibTeX-Style jss.bst has to be modified)
+
 Revision 1.12  2005/09/26 11:24:05  henningsena

### 2005-11-02 paper/trunk/systemfit.tex
small changes
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@578 591db2bd-e0a0-4d22-8c19-2192dbb7da31

--- 2005-09-26 paper/trunk/systemfit.tex
@@
 %% an abstract and keywords
-\Abstract{ Many statistical analysis are based on models that consist
-  of  more than one equation.  These equations should be fitted
-  simultaneously rather than equation  by equation.  The package
-  \pkg{systemfit} provides the capability to estimate systems  of
-  linear and non-linear equations in \proglang{R}.
-}
+\Abstract{Many statistical analysis are based on sets of models that
+  consist of  more than one equation which contain relationships among
+  the various equations.  Ignoring these relationships can result in
+  biased and possibly inconsistent parameter estimates. These sets of
+  equations should be fitted simultaneously rather than equation  by
+  equation.  The package \pkg{systemfit} provides the capability to
+  estimate systems of linear and non-linear equations, using limited
+  and full-information methods within the \proglang{R} programming
+  environment.  }
 
-\Keywords{R, simultaneous equation system, seemingly unrelated regression}
+\Keywords{R, simultaneous equation system, seemingly unrelated
+  regression, two-stage least squares, three-stage least squares}
 
@@
 \input{systemfit_restrictions.tex}
+\input{systemfit_usage.tex}
 \input{systemfit_other.tex}
-\input{systemfit_usage.tex}
 \input{systemfit_summary.tex}
+
+% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% \section{Estimating nonlinear equation systems}
+% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% The \pkg{systemfit} package also contains a method for fitting
+% nonlinear systems of equations.
+
+% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% \subsection{Nonlinear parameter estimation}
+% \label{sec:nonlinear_estimation}
+% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% A system of nonlinear equations can be written as:
+
+% \begin{equation}
+%   \label{eq:non_linear_eq_1}
+%   \epsilon_{t} = q( y_t, x_t, \beta )
+% \end{equation}
+% %%% It is not clear what these variable are. 
+% %%% Adding an equation representing only a single equation
+% %%% would make things clearer.
+
+% \noindent and
+
+% \begin{equation}
+%   \label{eq:non_linear_eq_2}
+%   z_{t} = Z( x_t )
+% \end{equation}
+
+% where $\epsilon_{t}$ are the residuals from the y observations and
+% $z_{t}$ are the function evaluated at the parameter estimates.
+
+% \textbf{you really need to check this!!!}
+
+% The objective functions for the methods are:
+
+% \begin{center}
+% \begin{tabular}{|l|c|c|c|} \hline
+%   Method & Instruments & Objective Function & Covariance of $\theta$ \\ \hline  
+%   OLS & No & $r'r$ & $(X(diag(S)^{-1}\otimes I)X)^{-1}$ \\ \hline
+%   SUR & No & $r'(diag(S)_{OLS}^{-1}\otimes I)r$ & $(X(S^{-1}\otimes I)X)^{-1}$ \\ \hline
+%   2SLS & Yes & $r'(I \otimes W)r$ & $(X(diag(S)^{-1}\otimes I)X)^{-1}$ \\ \hline 
+%   3SLS & Yes & $r'(S_{2SLS}^{-1} \otimes W)r$ & $(X(diag(S)^{-1}\otimes W)X)^{-1}$ \\ \hline
+% \end{tabular}
+% \end{center}
+
+% where, $r$ is a column vector for the residuals for each equation,
+% %% what is the difference to \epsilon?
+% $X$ is matrix of the partial derivatives of the dependent variable 
+% with respect to the parameters $\left( \frac{ \partial y }{ \theta} \right)$,
+% %% is this correct?
+% %% in the linear section X is used for the regressors, 
+% %% thus we should use a different variable name here.
+% $W$ is a matrix of the instrument variables, $Z(Z'Z)^{-1}Z$, $Z$ is a
+% matrix of the instrument variables, and $I$ is an $n \times n $
+% identity matrix and $S$ is the estimated variance-covariance matrix between the
+% equations
+
+% \begin{equation}
+%   \label{eq:non-linear_varcov}
+%   \hat{\sigma}_{ij} = (\hat{e}_i' \hat{e}_j) / \sqrt{(T - k_i)*(T- k_j)} 
+% \end{equation}
+
+% \textbf{The residual variance-covariance matrix can be calculated in 
+% different ways (section~\ref{sec:residcov}). 
+% It should be relatively easy to implement this also in nlsystemfit(), 
+% e.g. the function 'calcRCov' in systemfit() could be moved outside
+% systemfit() that it can be used also by nlsystemfit().}
+
+% \textbf{You need to clear this up.}
+
+% % %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% % \subsection{Using nlsystemfit}\label{sec:UsingnlSystemfit}
+% % %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% % \code {nlsystemfit} fits a set of structural nonlinear equations using
+% % Ordinary Least Squares (OLS), Seemingly Unrelated Regression (SUR),
+% % Two-Stage Least Squares (2SLS), Three-Stage Least Squares (3SLS) using
+% % the objective functions described in section
+% % \ref{sec:nonlinear_estimation}.
+
+% % %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% % \subsubsection{Standard usage}
+
+% % Similar to calling the \code{systemfit} function, \code{nlsystemfit}
+% % is called with a minimum of three arguments,
+
+% % \code{
+% % R> nlsystemfit( method, eqns, start )
+% % }
+
+% % where \code{method} is one of the following estimation methods: "OLS",
+% % "SUR", "2SLS", or "3SLS", \code{eqns} is a list of equations similar
+% % to those described for \code{systemfit}, and \code{start} is a list of
+% % starting values of the parameter estimates.
+
+% % % The mandatory argument \code{eqns} is a list of the equations  to
+% % % estimate.  Each equation is a standard formula in \proglang{R}.  It
+% % % starts with a dependent variable on the left hand side.  After a tilde
+% % % ($~$) the regressors are listed  \footnote{For Details see the
+% % %   \proglang{R} help files to \code{formula}}.
+
+% % This is now demonstrated using an example: \\
+
+% % % \code{
+% % % R> library( systemfit ) \\
+% % % R> data( ppine ) \\
+% % % R> hg.formula <- hg $\sim$ exp( h0 + h1*log(tht) + h2*tht$\hat$2 + h3*elev + h4*cr) \\
+% % % R> dg.formula <- dg $\sim$ exp( d0 + d1*log(dbh) + d2*hg + d3*cr + d4*ba  ) \\
+% % % R> labels <- list( "height.growth", "diameter.growth" )\\
+% % % R> inst <- $\sim$ tht + dbh + elev + cr + ba\\
+% % % R> start.values <- c(h0=-0.5, h1=0.5, h2=-0.001, h3=0.0001, h4=0.08,\\
+% % % R+   d0=-0.5, d1=0.009, d2=0.25, d3=0.005, d4=-0.02 )\\
+% % % R> model <- list( hg.formula, dg.formula )\\
+% % % R> model.3sls <- nlsystemfit( "3SLS", model, start.values, data=ppine,\\
+% % % R+   eqnlabels=labels, inst=inst ) \\
+% % % R> summary( model.3sls ) \\
+% % % }
+
+% % The nlsystemfit function relies on \code{nlm} to perform the
+% % minimization of the objective functions and the \code{qr} set of
+% % functions. If the user does not have a set of estimates for the
+% % initial parameters, it is suggested using linearized forms in one of
+% % the linear methods, or simply using \code{nls} to obtain estimates for
+% % the equations.
+
+% % The outputs are similar to those for a systemfit object.
 
@@
 $Log$
+Revision 1.15  2005/11/02 18:15:40  hamannj
+small changes
+
+Revision 1.14  2005/10/26 21:16:53  hamannj
+checked out and updated local copy before edits
+
+Revision 1.13  2005/09/26 20:34:12  henningsena
+using now package csquotes to have uniform and correct quotation marks (BibTeX-Style jss.bst has to be modified)
+
 Revision 1.12  2005/09/26 11:24:05  henningsena

### 2005-11-02 paper/trunk/systemfit_estimation.tex
small changes
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@578 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 \begin{equation}
-   y_{i} = X_i \beta_i + u_i, \quad i = 1, 2, \ldots, G
+   y_{i} = \mathbf{X}_i \beta_i + u_i, \quad i = 1, 2, \ldots, G
 \end{equation}
-where $y_i$ is a vector of the dependent variable,
-$X_i$ is a matrix of the exogenous variables,
-$\beta_i$ is the coefficient vector and
-$u_i$ is a vector of the disturbance terms of the $i$th equation.
+where $y_i$ is a vector of the dependent variable, $\mathbf{X}_i$ is a
+matrix of the exogenous variables, $\beta_i$ is the coefficient vector
+and $u_i$ is a vector of the disturbance terms of the $i$th equation.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Ordinary least squares (OLS)}
+\subsection{Ordinary Least Squares (OLS)}
 
@@
 is obtained by
+
 \begin{equation}
@@
 contemporaneously correlated, which means
+
 \begin{equation}
@@
 the variance-covariance matrix of the estimated parameters is
+
 \begin{equation}
@@
 \end{equation}
-with $\sigma^2 = E \left( u' u \right)$.
-This assumes that the disturbances of all equations have the
-same variance.
 
-If the disturbance terms of the individual equations 
-are allowed to have different variances, 
-the variance-covariance matrix of the estimated parameters is
+with $\sigma^2 = E \left( u' u \right)$.  This assumes that the
+disturbances of all equations have the same variance.
+
+If the disturbance terms of the individual equations  are allowed to
+have different variances,  the variance-covariance matrix of the
+estimated parameters is
+
 \begin{equation}
@@
 \end{equation}
-with $\Omega = \Sigma \otimes I$, 
-$\sigma_{ij} = 0 \; \forall \; i \neq j$ and
-$\sigma_{ii} = E \left( u_i' u_i \right)$.
+
+\noindent with $\Omega = \Sigma \otimes I$, 
+$\sigma_{ij} = 0 \; \forall \; i \neq j$ and $\sigma_{ii} = E \left(
+  u_i' u_i \right)$.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Weighted least squares (WLS)}
+\subsection{Weighted Least Squares (WLS)}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Seemingly unrelated regression (SUR)}
+\subsection{Seemingly Unrelated Regression (SUR)}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Two-stage least squares (2SLS)}
+\subsection{Two-Stage Least Squares (2SLS)}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Weighted two-stage least squares (W2SLS)}
+\subsection{Weighted Two-Stage Least Squares (W2SLS)}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Three-stage least squares (3SLS)}
+\subsection{Three-Stage Least Squares (3SLS)}
 
@@
 (\ref{eq:beta3slsSchmidt}) are consistent.
+
+%%% Local Variables: 
+%%% mode: latex
+%%% TeX-master: "systemfit"
+%%% End:

### 2005-11-02 paper/trunk/systemfit_intro.tex
some minor changes
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@573 591db2bd-e0a0-4d22-8c19-2192dbb7da31
small changes
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@578 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
+%       $Id$
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Many theoretical models consist of more than one equation. If these
-models are econometrically estimated, the equations generally should
-not be estimated individually, but all together, because the
-disturbance terms of these equations are likely contemporaneously
-correlated. Estimating all equations simultaneously and taking the
-covariance structure of the residuals into account leads to efficient
-estimates, while ignoring this correlation leads to an inefficient
-parameter estimation \citep{zellner62}.
+models are econometrically estimated, it may not be appropriate to
+estimate the equations individually, but all together, because the
+disturbance terms of these equations are likely to be
+contemporaneously correlated \citep{theil71} (make sure this is the
+correct paper). Estimating all equations simultaneously, taking the
+covariance structure of the residuals into account, leads to efficient
+estimates. Ignoring this correlation leads to an inefficient parameter
+estimation \citep{zellner62}.
 
@@
 cross-equation parameter restrictions. These restrictions can be
-tested and/or imposed only in a simultaneous estimation
-approach. Especially the economic theory suggests many cross-equation
-restrictions.
+tested and/or imposed only in a simultaneous estimation approach
+especially the economic theory suggests many cross-equation
+restrictions \textbf{(is this statement really true?)}.
 
-The \pkg{systemfit} package provides the capability to estimate
-linear and non-linear equation systems in \proglang{R}.
-Although linear and non-linear equation systems can be estimated
-with several other statistical and econometric software packages
-(e.g. \proglang{SAS}, \proglang{EViews}, \proglang{TSP}),
-using \pkg{systemfit} has several advantages.
-First, all estimation procedures are publicly available in the source code.
-Second, the estimation algorithms can be easily modified to meet specific
-requirements.
-Third, the (advanced) user can control many important details of the
-estimation procedure that are generally not available in other software
-packages.
-On the other hand these user options have reasonable defaults that also
-beginners can easily use the provided functions.
-
-\textbf{I'll have to update this section to match the recent changes. JDH.}
+The \pkg{systemfit} package provides the capability to estimate linear
+and non-linear equation systems in \proglang{R}.  Although linear and
+non-linear equation systems can be estimated with several other
+statistical and econometric software packages (e.g. \proglang{SAS},
+\proglang{EViews}, \proglang{TSP}), \pkg{systemfit} has several
+advantages.  First, all estimation procedures are publicly available
+in the source code.  Second, the estimation algorithms can be easily
+modified to meet specific requirements.  Third, the (advanced) user
+can control many important details of the estimation procedure that
+are generally not available in other software packages.  On the other
+hand, these user options have reasonable defaults that also beginners
+can easily use the provided functions.
 
@@
 Section~\ref{sec:Restrictions} shows how linear restrictions can be
-imposed.
-All other relevant issues are discussed in section~\ref{sec:Other}.
-Section~\ref{sec:Usage} demonstrates how to run
+imposed.  Section~\ref{sec:Usage} demonstrates how to run
 \pkg{systemfit}, especially how the features presented in the previous
-sections can be used.
-Finally, a summary and an outlook are given in
-section~\ref{sec:Summmary}.
+sections can be used. All other relevant issues are discussed in
+section~\ref{sec:Other}. Finally, a summary and outlook are presented
+in section~\ref{sec:Summmary}.
+
+%%% Local Variables: 
+%%% mode: latex
+%%% TeX-master: "systemfit"
+%%% End:

### 2005-11-02 paper/trunk/systemfit_other.tex
some minor changes
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@573 591db2bd-e0a0-4d22-8c19-2192dbb7da31
using now package csquotes to have uniform and correct quotation marks (BibTeX-Style jss.bst has to be modified)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@576 591db2bd-e0a0-4d22-8c19-2192dbb7da31
small changes
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@578 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 However, this estimator is biased, because it is not corrected for
-degrees of freedom. 
-The usual single-equation procedure to correct for degrees of freedom 
-cannot always be applied, because the number of regressors in each equation 
-might differ.
-Two alternative approaches to calculate the residual variance-covariance 
-matrix are
+degrees of freedom.  The usual single-equation procedure to correct
+for degrees of freedom  cannot always be applied, because the number
+of regressors in each equation  might differ.  Two alternative
+approaches to calculate the residual variance-covariance  matrix are
 \begin{equation}
@@
 is not necessarily positive semidefinite \citep[p.322]{theil71}. 
-Hence, "it is doubtful whether [this formula] is really superior to 
-[equation (\ref{eq:rcov0})]" \citep[p.322]{theil71}.
+Hence, °it is doubtful whether [this formula] is really superior to 
+[(\ref{eq:rcov0})]° \citep[p.322]{theil71}.
 
@@
 
-There exist different approaches to determine the degrees of
-freedom for testing the estimated parameters with a t-test.
+Different approaches to determine the degrees of freedom for testing
+the estimated parameters with a t-test \textbf{(cite this one)}.
+
+\textbf{Arne, do you have a citation for this?}
+
+The \code{probdfsys} can be used to set the degrees of freedom when
+calculating the probability values for the t-test for individual
+parameters. This argument can set the degrees of freedom for the
+entire system instead of using the degrees of freedom for each of the
+individual equations.
+
+%    \item{probdfsys}{use the degrees of freedom of the whole system
+%       (in place of the degrees of freedom of the single equation)
+%       to calculate prob values for the t-test of individual parameters.}
 
@@
 
+Currently, the \code{systemfit} package can perform a Likelihood Ratio
+test to test linear restrictions when estimating SUR equation systems.
+
+This test calculates the likelihood ratio value by calculating the
+estimated residual covariance matrix $hat{S}$ by the maximum likelihood
+formula
+
+\begin{equation}
+hat{s}_{ij} = \frac{\hat{e}_i^{'} \hat{e}_j}{T} 
+\end{equation}
+where $T$ is vector of ones.
+
+% (no matter which formula is used in the estimation).
+
+You should really get some of the text from your stats book.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\textbf{Man, I really have to clean this up.}
-
 Hausman \citep{hausman1978} developed a test have power against
 alternatives for which $\hat{\theta}$ and $\tilde{\theta}$ diverge
-under misspecification.
-
-The Hausman result if the difference between the covariance of an
-efficient and the covariance of the inefficient estimator is zero, the
-model is correctly specified.
+under misspecification. The resulting test statistic, distributed as a
+$\chi^2$, is the difference between the covariance of an efficient and
+the covariance of the inefficient estimator is zero, if the model is
+correctly specified.
 
@@
 
-Since the models that are fit using the \pkg{systemfit} package are
-typically random effects models, the effects may be inconsistent
-between the introduced variables and the regressors.
+% Since the models that are fit using the \pkg{systemfit} package are
+% typically random effects models, the effects may be inconsistent
+% between the introduced variables and the regressors.
+
+% results2sls  result of a 2SLS (limited information) estimation returned by systemfit. 
+% results3sls  result of a 3SLS (full information) estimation returned by systemfit.
+
+The null hypotheses of the test is that all exogenous variables are
+uncorrelated with all disturbance terms. Under this hypothesis both
+the 2SLS and the 3SLS estimator are consistent but only the 3SLS
+estimator is (asymptotically) efficient. Under the alternative
+hypothesis the 2SLS estimator is consistent but the 3SLS estimator is
+inconsistent.
+
+The Hausman test statistic is,
+
+\begin{equation}
+m = ( b_2 - b_3 )^{'} ( V_2 - V_3 ) ( b_2 - b_3 ) 
+\label{eq:hausman}
+\end{equation}
+where $b_2$ and $V_2$ are the estimated coefficients and the
+variance-covariance matrix from a 2SLS estimation and $b_3$ and $V_3$
+are the estimated coefficients and their variance covariance-matrix
+from a 3SLS estimation.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Nonlinear estimation}
+
+As linear systems of equations can be estimated using
+\code{systemfit}, nonlinear systems of equations can be estimated
+using a similar function \code{nlsystemfit}. The usage for the
+nonlinear estimation function is similar to \code{systemfit} and can
+estimate system of equations using OLS, SUR, 2SLS and 3SLS methods.
+
+\begin{equation}
+  \label{eq:non_linear_eq_1}
+  \epsilon_{t} = q( y_t, x_t, \beta )
+\end{equation}
+
+\noindent and
+
+\begin{equation}
+  \label{eq:non_linear_eq_2}
+  z_{t} = Z( x_t )
+\end{equation}
+where $\epsilon_{t}$ are the residuals from the $y_t$ observations,
+$x_t$ is a matrix of independent variables, $\beta$ is a vector is
+parameter estimates and $z_{t}$ are the functions evaluated at the
+parameter estimates.
+
+Similar to calling the \code{systemfit} function, \code{nlsystemfit}
+is called with a minimum of three arguments,
+
+\code{
+R> nlsystemfit( method, eqns, start )
+}
+
+where \code{method} is one of the following estimation methods: "OLS",
+"SUR", "2SLS", or "3SLS", \code{eqns} is a list of equations similar
+to those described for \code{systemfit}, and \code{start} is a list of
+starting values of the parameter estimates. The nlsystemfit function
+relies on \code{nlm} to perform the minimization of the objective
+functions and the \code{qr} set of functions. If the user does not
+have a set of estimates for the initial parameters, it is suggested
+using linearized forms in one of the linear methods, or simply using
+\code{nls} to obtain estimates for the equations.
+
+% The mandatory argument
+% \code{eqns} is a list of the equations to estimate.  Each equation is
+% a standard formula in \proglang{R}.  It starts with a dependent
+% variable on the left hand side.  After a tilde ($~$) the regressors
+% are listed  \footnote{For Details see the \proglang{R} help files to
+%   \code{formula}}.
+
+% The objective functions for the methods are:
+
+% \begin{center}
+% \begin{tabular}{|l|c|c|c|} \hline
+%   Method & Instruments & Objective Function & Covariance of $\theta$ \\ \hline  
+%   OLS & No & $r'r$ & $(X(diag(S)^{-1}\otimes I)X)^{-1}$ \\ \hline
+%   SUR & No & $r'(diag(S)_{OLS}^{-1}\otimes I)r$ & $(X(S^{-1}\otimes I)X)^{-1}$ \\ \hline
+%   2SLS & Yes & $r'(I \otimes W)r$ & $(X(diag(S)^{-1}\otimes I)X)^{-1}$ \\ \hline 
+%   3SLS & Yes & $r'(S_{2SLS}^{-1} \otimes W)r$ & $(X(diag(S)^{-1}\otimes W)X)^{-1}$ \\ \hline
+% \end{tabular}
+% \end{center}
+
+% where, $r$ is a column vector for the residuals for each equation,
+% what is the difference to \epsilon?
+% $X$ is matrix of the partial derivatives of the dependent variable 
+% with respect to the parameters $\left( \frac{ \partial y }{ \theta} \right)$,
+% is this correct?
+% in the linear section X is used for the regressors, 
+% thus we should use a different variable name here.
+% $W$ is a matrix of the instrument variables, $Z(Z'Z)^{-1}Z$, $Z$ is a
+% matrix of the instrument variables, and $I$ is an $n \times n $
+% identity matrix and $S$ is the estimated variance-covariance matrix between the
+% equations
+
+% \begin{equation}
+%   \label{eq:non-linear_varcov}
+%   \hat{\sigma}_{ij} = (\hat{e}_i' \hat{e}_j) / \sqrt{(T - k_i)*(T- k_j)} 
+% \end{equation}
+
+% Do we actually have the variance-covariance function in place.
+
+% The argument \code{method} is a string determining the estimation method.
+% It must be one of "OLS", "WLS", "SUR", "2SLS", "W2SLS" or "3SLS".
+
+% \textbf{The residual variance-covariance matrix can be calculated in
+%   different ways (section~\ref{sec:residcov}).  It should be
+%   relatively easy to implement this also in nlsystemfit(),  e.g. the
+%   function 'calcRCov' in systemfit() could be moved outside
+%   systemfit() that it can be used also by nlsystemfit().}
+
+% \textbf{You need to clear this up.}
+
+%%% Local Variables: 
+%%% mode: latex
+%%% TeX-master: "systemfit"
+%%% End:

### 2005-11-02 paper/trunk/systemfit_summary.tex
some minor changes
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@574 591db2bd-e0a0-4d22-8c19-2192dbb7da31
small changes
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@578 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 
-What should we place in this section?
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Unbalanced Observations}\label{sec:summary_unbalanced_data}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 The \pkg{systemfit} package was originally developed to fit
 simultaneous equations for forestry datasets. Forestry datasets
-typically contain observations for many inexpensive observations
-(diameter at breast height) and very few expensive observations such
-as total tree height. While the package currently does not allow for
-unbalanced datasets, data where the equations contain different
-numbers of observations, future releases of the package will have that
-feature implemented. We are still trying to figure out how develop the
-code for instrumental variable methods (2SLS,3SLS) for the correct
-estimation methods if possible.
+typically contain observations of many inexpensive observations (stem
+diameter) and few expensive observations such as stem height and being
+able the utilize as many observations as possible is advantageous. As
+most software packages simply drop data points that do not contain
+both observations, the \pkg{systemfit} package currently does not
+allow for unbalanced datasets, but it is our intention to include this
+capability for certain estimation methods in future releases.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Additional Methods}\label{sec:summary_additional_methods}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+In this article, we have described some of the basic features of the
+\pkg{systemfit} package for estimation of linear systems of
+equations. It has been tested on a variety of datasets and has
+produced satisfactory for a few years. While more sophisticated tools
+exist, the \pkg{systemfit} package will perform the basic fitting
+methods. We wish to include more sophisticated estimation methods such
+as full-information maximum likelihood and generalized methods of
+moments.
 
-\subsubsection{Full Information Maximum Likelihood}
-
-Describe, in a paragraph the process and give some reasons why you
-might want this in the package.
-
-\subsubsection{Generalized Method of Moments}
-
-Describe, in a paragraph the process and give some reasons why you
-might want this in the package.
+%%% Local Variables: 
+%%% mode: latex
+%%% TeX-master: "systemfit"
+%%% End:

### 2005-11-02 paper/trunk/systemfit.tex
small last changes
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@579 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 %% an abstract and keywords
-\Abstract{Many statistical analysis are based on sets of models that
-  consist of  more than one equation which contain relationships among
-  the various equations.  Ignoring these relationships can result in
-  biased and possibly inconsistent parameter estimates. These sets of
-  equations should be fitted simultaneously rather than equation  by
-  equation.  The package \pkg{systemfit} provides the capability to
-  estimate systems of linear and non-linear equations, using limited
-  and full-information methods within the \proglang{R} programming
+\Abstract{Many statistical analysis are based on models that consist
+  of  more than one equation. If these models contain relationships
+  among the various equations, ignoring these relationships can result
+  in biased and possibly inconsistent parameter estimates. If
+  relationships exist among the equations, the sets of equations
+  should be fitted simultaneously rather than equation by equation.
+  The package \pkg{systemfit} provides the capability to estimate
+  systems of linear and non-linear equations, using limited and
+  full-information methods within the \proglang{R} programming
   environment.  }
@@
 
-\section{CVS Revision Log}
-\begin{verbatim}
-$Log$
-Revision 1.15  2005/11/02 18:15:40  hamannj
-small changes
+% \section{CVS Revision Log}
+% \begin{verbatim}
+% $Log$
+% Revision 1.16  2005/11/02 18:31:45  hamannj
+% small last changes
+%
+% Revision 1.15  2005/11/02 18:15:40  hamannj
+% small changes
 
-Revision 1.14  2005/10/26 21:16:53  hamannj
-checked out and updated local copy before edits
+% Revision 1.14  2005/10/26 21:16:53  hamannj
+% checked out and updated local copy before edits
 
-Revision 1.13  2005/09/26 20:34:12  henningsena
-using now package csquotes to have uniform and correct quotation marks (BibTeX-Style jss.bst has to be modified)
+% Revision 1.13  2005/09/26 20:34:12  henningsena
+% using now package csquotes to have uniform and correct quotation marks (BibTeX-Style jss.bst has to be modified)
 
-Revision 1.12  2005/09/26 11:24:05  henningsena
-some minor changes
+% Revision 1.12  2005/09/26 11:24:05  henningsena
+% some minor changes
 
-Revision 1.11  2005/09/01 12:27:55  henningsena
-revised structure of the document (sections, subsections, subsubsections) due to removal of nlsystemfit; moved each section into a separate file.
+% Revision 1.11  2005/09/01 12:27:55  henningsena
+% revised structure of the document (sections, subsections, subsubsections) due to removal of nlsystemfit; moved each section into a separate file.
 
-Revision 1.10  2005/09/01 11:49:02  henningsena
-moved section about non-linear estimation from systemfit.tex into new file nlsystemfit.tex, because this section should not be included in (this version of) the article
+% Revision 1.10  2005/09/01 11:49:02  henningsena
+% moved section about non-linear estimation from systemfit.tex into new file nlsystemfit.tex, because this section should not be included in (this version of) the article
 
-Revision 1.9  2004/11/16 10:59:09  henningsena
-some minor changes and some comments on nlsystemfit
+% Revision 1.9  2004/11/16 10:59:09  henningsena
+% some minor changes and some comments on nlsystemfit
 
-Revision 1.8  2004/11/16 08:55:53  henningsena
-merged file and resolved conflict due to some minor changes I did while Jeff was adding nlsystemfit.
+% Revision 1.8  2004/11/16 08:55:53  henningsena
+% merged file and resolved conflict due to some minor changes I did while Jeff was adding nlsystemfit.
 
-Revision 1.7  2004/11/16 05:34:47  hamannj
-question about nlsystemfit code section (example)
+% Revision 1.7  2004/11/16 05:34:47  hamannj
+% question about nlsystemfit code section (example)
 
-Revision 1.6  2004/11/16 05:32:46  hamannj
-made small changes because my code section wouldn't compile. argh.
+% Revision 1.6  2004/11/16 05:32:46  hamannj
+% made small changes because my code section wouldn't compile. argh.
 
-Revision 1.5  2004/11/16 05:31:19  hamannj
-added cvs header and log section. to be removed when submitted.
+% Revision 1.5  2004/11/16 05:31:19  hamannj
+% added cvs header and log section. to be removed when submitted.
 
-\end{verbatim}
+% \end{verbatim}
 

### 2005-11-02 paper/trunk/systemfit_other.tex
small last changes
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@579 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 test to test linear restrictions when estimating SUR equation systems.
+The function returns the empirical likelihood ratio, the $p$-value and
+the degrees of freedom of the test which is equivalent to the number
+of restrictions.
 
-This test calculates the likelihood ratio value by calculating the
-estimated residual covariance matrix $hat{S}$ by the maximum likelihood
-formula
+% This test calculates the likelihood ratio value by calculating the
+% estimated residual covariance matrix $\hat{S}$ by the maximum
+% likelihood formula,
 
-\begin{equation}
-hat{s}_{ij} = \frac{\hat{e}_i^{'} \hat{e}_j}{T} 
-\end{equation}
-where $T$ is vector of ones.
-
-% (no matter which formula is used in the estimation).
-
-You should really get some of the text from your stats book.
+% \begin{equation}
+% \hat{s}_{ij} = \frac{\hat{e}_i^{'} \hat{e}_j}{T} 
+% \end{equation}
+% where $T$ is vector of ones.
 

### 2005-11-02 paper/trunk/systemfit_other.tex
small last changes
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@580 591db2bd-e0a0-4d22-8c19-2192dbb7da31
added hausman citation
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@581 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 
-%    \item{probdfsys}{use the degrees of freedom of the whole system
-%       (in place of the degrees of freedom of the single equation)
-%       to calculate prob values for the t-test of individual parameters.}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-% Since the models that are fit using the \pkg{systemfit} package are
-% typically random effects models, the effects may be inconsistent
-% between the introduced variables and the regressors.
-
-% results2sls  result of a 2SLS (limited information) estimation returned by systemfit. 
-% results3sls  result of a 3SLS (full information) estimation returned by systemfit.
-
 The null hypotheses of the test is that all exogenous variables are
@@
 As linear systems of equations can be estimated using
-\code{systemfit}, nonlinear systems of equations can be estimated
-using a similar function \code{nlsystemfit}. The usage for the
-nonlinear estimation function is similar to \code{systemfit} and can
-estimate system of equations using OLS, SUR, 2SLS and 3SLS methods.
+\code{systemfit}, nonlinear systems of equations \citep{hausman-1975}
+can be estimated using a similar function \code{nlsystemfit}. The
+usage for the nonlinear estimation function is similar to
+\code{systemfit} and can estimate system of equations using OLS, SUR,
+2SLS and 3SLS methods.
 
@@
 \code{nls} to obtain estimates for the equations.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% left over from the older docs
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2005-11-07 paper/trunk/systemfit_other.tex
just changed some line breaks (deleted new ones and restored deleted ones) to make it easier to find the changes that Jeff committed in the last revision(s)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@582 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 However, this estimator is biased, because it is not corrected for
-degrees of freedom.  The usual single-equation procedure to correct
-for degrees of freedom  cannot always be applied, because the number
-of regressors in each equation  might differ.  Two alternative
-approaches to calculate the residual variance-covariance  matrix are
+degrees of freedom.
+The usual single-equation procedure to correct for degrees of freedom
+cannot always be applied, because the number of regressors in each equation
+might differ.
+Two alternative approaches to calculate the residual variance-covariance
+matrix are
 \begin{equation}
@@
 
-Different approaches to determine the degrees of freedom for testing
-the estimated parameters with a t-test \textbf{(cite this one)}.
+Different approaches to determine the degrees of
+freedom for testing the estimated parameters with a t-test
+\textbf{(cite this one)}.
 
@@
 calculating the probability values for the t-test for individual
-parameters. This argument can set the degrees of freedom for the
-entire system instead of using the degrees of freedom for each of the
+parameters.
+This argument can set the degrees of freedom for the entire system
+instead of using the degrees of freedom for each of the
 individual equations.
@@
 alternatives for which $\hat{\theta}$ and $\tilde{\theta}$ diverge
-under misspecification. The resulting test statistic, distributed as a
-$\chi^2$, is the difference between the covariance of an efficient and
-the covariance of the inefficient estimator is zero, if the model is
-correctly specified.
+under misspecification.
+The resulting test statistic, distributed as a
+$\chi^2$, is the difference between the covariance of an
+efficient and the covariance of the inefficient estimator is zero, if the
+model is correctly specified.
 
@@
 The null hypotheses of the test is that all exogenous variables are
-uncorrelated with all disturbance terms. Under this hypothesis both
-the 2SLS and the 3SLS estimator are consistent but only the 3SLS
-estimator is (asymptotically) efficient. Under the alternative
-hypothesis the 2SLS estimator is consistent but the 3SLS estimator is
-inconsistent.
+uncorrelated with all disturbance terms.
+Under this hypothesis both the 2SLS and the 3SLS estimator are consistent
+but only the 3SLS estimator is (asymptotically) efficient.
+Under the alternative hypothesis the 2SLS estimator is consistent but the
+3SLS estimator is inconsistent.
 
@@
 \code{systemfit}, nonlinear systems of equations \citep{hausman-1975}
-can be estimated using a similar function \code{nlsystemfit}. The
-usage for the nonlinear estimation function is similar to
+can be estimated using a similar function \code{nlsystemfit}.
+The usage for the nonlinear estimation function is similar to
 \code{systemfit} and can estimate system of equations using OLS, SUR,
@@
 to those described for \code{systemfit}, and \code{start} is a list of
-starting values of the parameter estimates. The nlsystemfit function
-relies on \code{nlm} to perform the minimization of the objective
-functions and the \code{qr} set of functions. If the user does not
-have a set of estimates for the initial parameters, it is suggested
-using linearized forms in one of the linear methods, or simply using
-\code{nls} to obtain estimates for the equations.
+starting values of the parameter estimates.
+The nlsystemfit function relies on \code{nlm} to perform the minimization
+of the objective functions and the \code{qr} set of functions.
+If the user does not have a set of estimates for the initial parameters,
+it is suggested using linearized forms in one of the linear methods,
+or simply using \code{nls} to obtain estimates for the equations.
 

### 2005-11-07 paper/trunk/systemfit_other.tex
a consistent BibTeX key pattern is now used throughout the whole article: [lastname of first author][last two digits of year]
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@583 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 
-Hausman \citep{hausman1978} developed a test have power against
+Hausman \citep{hausman78} developed a test have power against
 alternatives for which $\hat{\theta}$ and $\tilde{\theta}$ diverge
@@
 As linear systems of equations can be estimated using
-\code{systemfit}, nonlinear systems of equations \citep{hausman-1975}
+\code{systemfit}, nonlinear systems of equations \citep{hausman75}
 can be estimated using a similar function \code{nlsystemfit}.

### 2005-11-07 paper/trunk/systemfit.tex
just changed some line breaks (deleted new ones and restored deleted ones) to make it easier to find the changes that Jeff committed in the last revision(s)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@582 591db2bd-e0a0-4d22-8c19-2192dbb7da31
a few minor changes (predominantly adding a "\ " after a period (.) that does not indicate an end of a sentence)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@584 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 %% for pretty printing and a nice hypersummary also set:
-\Plainauthor{Arne Henningsen, Jeff D. Hamann} %% comma-separated
+\Plainauthor{Arne Henningsen, Jeff D.\ Hamann} %% comma-separated
 \Plaintitle{systemfit: A Package to Estimate Simultaneous Equation Systems 
@@
 
-\Keywords{R, simultaneous equation system, seemingly unrelated
-  regression, two-stage least squares, three-stage least squares}
+\Keywords{R, simultaneous equation system, seemingly unrelated regression,
+   two-stage least squares, three-stage least squares}
 
@@
    \\
-   Jeff D. Hamann\\
+   Jeff D.\ Hamann\\
    Forest Informatics, Inc.\\
@@
 % $Log$
+% Revision 1.18  2005/11/07 14:08:56  henningsena
+% a few minor changes (predominantly adding a "\ " after a period (.) that does not indicate an end of a sentence)
+%
+% Revision 1.17  2005/11/07 10:21:59  henningsena
+% just changed some line breaks (deleted new ones and restored deleted ones) to make it easier to find the changes that Jeff committed in the last revision(s)
+%
 % Revision 1.16  2005/11/02 18:31:45  hamannj

### 2005-11-07 paper/trunk/systemfit_intro.tex
just changed some line breaks (deleted new ones and restored deleted ones) to make it easier to find the changes that Jeff committed in the last revision(s)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@582 591db2bd-e0a0-4d22-8c19-2192dbb7da31
a few minor changes (predominantly adding a "\ " after a period (.) that does not indicate an end of a sentence)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@584 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 estimate the equations individually, but all together, because the
-disturbance terms of these equations are likely to be
-contemporaneously correlated \citep{theil71} (make sure this is the
-correct paper). Estimating all equations simultaneously, taking the
+disturbance terms of these equations are likely to be contemporaneously
+correlated. %\citep{theil71} (make sure this is the correct paper).
+Estimating all equations simultaneously, taking the
 covariance structure of the residuals into account, leads to efficient
-estimates. Ignoring this correlation leads to an inefficient parameter
-estimation \citep{zellner62}.
+estimates.
+Ignoring this correlation leads to an inefficient
+parameter estimation \citep{zellner62}.
 
@@
 cross-equation parameter restrictions. These restrictions can be
-tested and/or imposed only in a simultaneous estimation approach
-especially the economic theory suggests many cross-equation
-restrictions \textbf{(is this statement really true?)}.
+tested and/or imposed only in a simultaneous estimation
+approach especially the economic theory suggests many cross-equation
+restrictions.
 
-The \pkg{systemfit} package provides the capability to estimate linear
-and non-linear equation systems in \proglang{R}.  Although linear and
-non-linear equation systems can be estimated with several other
-statistical and econometric software packages (e.g. \proglang{SAS},
-\proglang{EViews}, \proglang{TSP}), \pkg{systemfit} has several
-advantages.  First, all estimation procedures are publicly available
-in the source code.  Second, the estimation algorithms can be easily
-modified to meet specific requirements.  Third, the (advanced) user
-can control many important details of the estimation procedure that
-are generally not available in other software packages.  On the other
-hand, these user options have reasonable defaults that also beginners
-can easily use the provided functions.
+The \pkg{systemfit} package provides the capability to estimate
+linear and non-linear equation systems in \proglang{R}.
+Although linear and non-linear equation systems can be estimated
+with several other statistical and econometric software packages
+(e.g.\ \proglang{SAS}, \proglang{EViews}, \proglang{TSP}),
+\pkg{systemfit} has several advantages.
+First, all estimation procedures are publicly available in the source code.
+Second, the estimation algorithms can be easily modified to meet specific
+requirements.
+Third, the (advanced) user can control many important details of the
+estimation procedure that are generally not available in other software
+packages.
+On the other hand, these user options have reasonable defaults that also
+beginners can easily use the provided functions.
 
@@
 Section~\ref{sec:Restrictions} shows how linear restrictions can be
-imposed.  Section~\ref{sec:Usage} demonstrates how to run
+imposed.
+Section~\ref{sec:Usage} demonstrates how to run
 \pkg{systemfit}, especially how the features presented in the previous
-sections can be used. All other relevant issues are discussed in
-section~\ref{sec:Other}. Finally, a summary and outlook are presented
-in section~\ref{sec:Summmary}.
+sections can be used.
+All other relevant issues are discussed in section~\ref{sec:Other}.
+Finally, a summary and outlook are presented in
+section~\ref{sec:Summmary}.
 

### 2005-11-07 paper/trunk/systemfit_other.tex
a few minor changes (predominantly adding a "\ " after a period (.) that does not indicate an end of a sentence)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@584 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 However, this yields an unbiased estimator only if $K_i = K_j$ 
-\citep[p. 469]{judge85}. 
+\citep[p.\ 469]{judge85}. 
 % Greene (2002, p. 344) says that the second is unbiased if i = j or K_i = K_j,
@@
 A further approach to obtain the estimated residual variance-covariance
-matrix is \citep[p. 309]{zellner62c}
+matrix is \citep[p.\ 309]{zellner62c}
 \begin{eqnarray}
@@
 its inverse $\widehat{\Sigma}^{-1}$ is not an unbiased estimator 
-of $\Sigma^{-1}$ \citep[p.322]{theil71}.
+of $\Sigma^{-1}$ \citep[p.\ 322]{theil71}.
 Furthermore, the variance-covariance matrix calculated by (\ref{eq:rcov2})
-is not necessarily positive semidefinite \citep[p.322]{theil71}. 
+is not necessarily positive semidefinite \citep[p.\ 322]{theil71}. 
 Hence, °it is doubtful whether [this formula] is really superior to 
-[(\ref{eq:rcov0})]° \citep[p.322]{theil71}.
+[(\ref{eq:rcov0})]° \citep[p.\ 322]{theil71}.
 
@@
 residual variance-covariance matrix, the results are maximum
-likelihood \citep[p. 345]{greene02}.
+likelihood \citep[p.\ 345]{greene02}.
 
@@
 McElroy's $R^2$ value \citep{mcelroy77}: 
-% also: \citep[p. 345]{greene02}
+% also: \citep[p.\ 345]{greene02}
 \begin{equation}

### 2005-11-07 paper/trunk/systemfit_restrictions.tex
small changes
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@578 591db2bd-e0a0-4d22-8c19-2192dbb7da31
small last changes
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@579 591db2bd-e0a0-4d22-8c19-2192dbb7da31
just changed some line breaks (deleted new ones and restored deleted ones) to make it easier to find the changes that Jeff committed in the last revision(s)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@582 591db2bd-e0a0-4d22-8c19-2192dbb7da31
a few minor changes (predominantly adding a "\ " after a period (.) that does not indicate an end of

@@
 
-To impose these restrictions the $X$ matrix is
+To impose these restrictions, the $X$ matrix is
 (post-)multiplied by this $T$ matrix.
@@
 where $\beta^0$ is the vector of the restricted coefficients, 
-$R$ is a matrix and $q$ is a vector (see \citealp[p. 100]{greene02}). 
+$R$ is a matrix and $q$ is a vector (see \citealp[p.\ 100]{greene02}). 
 Each restriction belongs to one row of $R$ and the corresponding 
@@
 Thus, in some cases the latter way leads to estimation problems
-(e.g. (near) singularity of the matrices to be inverted),
+(e.g.\ (near) singularity of the matrices to be inverted),
 while the first doesn't.
@@
 
-(***** This has to be checked!!!!! **************)
+%%% Local Variables: 
+%%% mode: latex
+%%% TeX-master: "systemfit"
+%%% End:

### 2005-11-08 paper/trunk/systemfit_summary.tex
just changed some line breaks (deleted new ones and restored deleted ones) to make it easier to find the changes that Jeff committed in the last revision(s)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@582 591db2bd-e0a0-4d22-8c19-2192dbb7da31
added serial correlation and heteroscedastic paragraph
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@585 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 simultaneous equations for forestry datasets. Forestry datasets
-typically contain observations of many inexpensive observations (stem
-diameter) and few expensive observations such as stem height and being
-able the utilize as many observations as possible is advantageous. As
-most software packages simply drop data points that do not contain
+typically contain observations of many inexpensive observations
+(stem diameter) and few expensive observations such
+as stem height and being able the utilize as many observations as
+possible is advantageous.
+As most software packages simply drop data points that do not contain
 both observations, the \pkg{systemfit} package currently does not
@@
 capability for certain estimation methods in future releases.
+
+For all of the methods developed in the package, the disturbances of
+the individual equations are assumed to be independent and identically
+distributed (iid) and the models are fit using least squares.  
+The package could be enhanced by the inclusion of methods to fit
+equations with serially correlated and heteroscedastic disturbances.
 

### 2005-11-08 paper/trunk/systemfit.tex
made seemingly massive changes... but hopefully not for the worse...
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@586 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 \MakeOuterQuote{°}
+
+\usepackage{lineno}
+
+\linenumbers
 
@@
 %% an abstract and keywords
-\Abstract{Many statistical analysis are based on models that consist
-  of  more than one equation. If these models contain relationships
-  among the various equations, ignoring these relationships can result
-  in biased and possibly inconsistent parameter estimates. If
-  relationships exist among the equations, the sets of equations
-  should be fitted simultaneously rather than equation by equation.
-  The package \pkg{systemfit} provides the capability to estimate
-  systems of linear and non-linear equations, using limited and
-  full-information methods within the \proglang{R} programming
-  environment.  }
+\Abstract{
 
-\Keywords{R, simultaneous equation system, seemingly unrelated regression,
-   two-stage least squares, three-stage least squares}
+  Many statistical analyses are based on models containing systems of
+  structurally related equations. If the cross-equation disturbances
+  are correlated, full information methods are required
+  \citep{zellner62}. If exogenous variables are stochastically
+  dependent on the disturbances in the system
+  \citep{zellner62b,theil71}, limited information methods should be
+  used for parameter estimation. The package \pkg{systemfit} provides
+  the capability to estimate systems of linear and non-linear
+  structural equations for limited (ordinary least squares (OLS), two
+  stage-least squares (2SLS)) and full-information
+  (seemingly-unrelated regression (SUR) and three-stage least squares
+  (3SLS)) methods within the \proglang{R} programming environment.}
+
+\Keywords{R, simultaneous equations estimation, ordinary least
+  squares, seemingly unrelated regression, two-stage least squares,
+  three-stage least squares}
 
@@
 
-% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% \section{Estimating nonlinear equation systems}
-% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% The \pkg{systemfit} package also contains a method for fitting
-% nonlinear systems of equations.
-
-% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% \subsection{Nonlinear parameter estimation}
-% \label{sec:nonlinear_estimation}
-% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% A system of nonlinear equations can be written as:
-
-% \begin{equation}
-%   \label{eq:non_linear_eq_1}
-%   \epsilon_{t} = q( y_t, x_t, \beta )
-% \end{equation}
-% %%% It is not clear what these variable are. 
-% %%% Adding an equation representing only a single equation
-% %%% would make things clearer.
-
-% \noindent and
-
-% \begin{equation}
-%   \label{eq:non_linear_eq_2}
-%   z_{t} = Z( x_t )
-% \end{equation}
-
-% where $\epsilon_{t}$ are the residuals from the y observations and
-% $z_{t}$ are the function evaluated at the parameter estimates.
-
-% \textbf{you really need to check this!!!}
-
-% The objective functions for the methods are:
-
-% \begin{center}
-% \begin{tabular}{|l|c|c|c|} \hline
-%   Method & Instruments & Objective Function & Covariance of $\theta$ \\ \hline  
-%   OLS & No & $r'r$ & $(X(diag(S)^{-1}\otimes I)X)^{-1}$ \\ \hline
-%   SUR & No & $r'(diag(S)_{OLS}^{-1}\otimes I)r$ & $(X(S^{-1}\otimes I)X)^{-1}$ \\ \hline
-%   2SLS & Yes & $r'(I \otimes W)r$ & $(X(diag(S)^{-1}\otimes I)X)^{-1}$ \\ \hline 
-%   3SLS & Yes & $r'(S_{2SLS}^{-1} \otimes W)r$ & $(X(diag(S)^{-1}\otimes W)X)^{-1}$ \\ \hline
-% \end{tabular}
-% \end{center}
-
-% where, $r$ is a column vector for the residuals for each equation,
-% %% what is the difference to \epsilon?
-% $X$ is matrix of the partial derivatives of the dependent variable 
-% with respect to the parameters $\left( \frac{ \partial y }{ \theta} \right)$,
-% %% is this correct?
-% %% in the linear section X is used for the regressors, 
-% %% thus we should use a different variable name here.
-% $W$ is a matrix of the instrument variables, $Z(Z'Z)^{-1}Z$, $Z$ is a
-% matrix of the instrument variables, and $I$ is an $n \times n $
-% identity matrix and $S$ is the estimated variance-covariance matrix between the
-% equations
-
-% \begin{equation}
-%   \label{eq:non-linear_varcov}
-%   \hat{\sigma}_{ij} = (\hat{e}_i' \hat{e}_j) / \sqrt{(T - k_i)*(T- k_j)} 
-% \end{equation}
-
-% \textbf{The residual variance-covariance matrix can be calculated in 
-% different ways (section~\ref{sec:residcov}). 
-% It should be relatively easy to implement this also in nlsystemfit(), 
-% e.g. the function 'calcRCov' in systemfit() could be moved outside
-% systemfit() that it can be used also by nlsystemfit().}
-
-% \textbf{You need to clear this up.}
-
-% % %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% % \subsection{Using nlsystemfit}\label{sec:UsingnlSystemfit}
-% % %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% % \code {nlsystemfit} fits a set of structural nonlinear equations using
-% % Ordinary Least Squares (OLS), Seemingly Unrelated Regression (SUR),
-% % Two-Stage Least Squares (2SLS), Three-Stage Least Squares (3SLS) using
-% % the objective functions described in section
-% % \ref{sec:nonlinear_estimation}.
-
-% % %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% % \subsubsection{Standard usage}
-
-% % Similar to calling the \code{systemfit} function, \code{nlsystemfit}
-% % is called with a minimum of three arguments,
-
-% % \code{
-% % R> nlsystemfit( method, eqns, start )
-% % }
-
-% % where \code{method} is one of the following estimation methods: "OLS",
-% % "SUR", "2SLS", or "3SLS", \code{eqns} is a list of equations similar
-% % to those described for \code{systemfit}, and \code{start} is a list of
-% % starting values of the parameter estimates.
-
-% % % The mandatory argument \code{eqns} is a list of the equations  to
-% % % estimate.  Each equation is a standard formula in \proglang{R}.  It
-% % % starts with a dependent variable on the left hand side.  After a tilde
-% % % ($~$) the regressors are listed  \footnote{For Details see the
-% % %   \proglang{R} help files to \code{formula}}.
-
-% % This is now demonstrated using an example: \\
-
-% % % \code{
-% % % R> library( systemfit ) \\
-% % % R> data( ppine ) \\
-% % % R> hg.formula <- hg $\sim$ exp( h0 + h1*log(tht) + h2*tht$\hat$2 + h3*elev + h4*cr) \\
-% % % R> dg.formula <- dg $\sim$ exp( d0 + d1*log(dbh) + d2*hg + d3*cr + d4*ba  ) \\
-% % % R> labels <- list( "height.growth", "diameter.growth" )\\
-% % % R> inst <- $\sim$ tht + dbh + elev + cr + ba\\
-% % % R> start.values <- c(h0=-0.5, h1=0.5, h2=-0.001, h3=0.0001, h4=0.08,\\
-% % % R+   d0=-0.5, d1=0.009, d2=0.25, d3=0.005, d4=-0.02 )\\
-% % % R> model <- list( hg.formula, dg.formula )\\
-% % % R> model.3sls <- nlsystemfit( "3SLS", model, start.values, data=ppine,\\
-% % % R+   eqnlabels=labels, inst=inst ) \\
-% % % R> summary( model.3sls ) \\
-% % % }
-
-% % The nlsystemfit function relies on \code{nlm} to perform the
-% % minimization of the objective functions and the \code{qr} set of
-% % functions. If the user does not have a set of estimates for the
-% % initial parameters, it is suggested using linearized forms in one of
-% % the linear methods, or simply using \code{nls} to obtain estimates for
-% % the equations.
-
-% % The outputs are similar to those for a systemfit object.
-
 % \section{CVS Revision Log}
@@
 % $Log$
+% Revision 1.19  2005/11/08 06:31:48  hamannj
+% made seemingly massive changes... but hopefully not for the worse...
+%
 % Revision 1.18  2005/11/07 14:08:56  henningsena

### 2005-11-08 paper/trunk/systemfit_intro.tex
made seemingly massive changes... but hopefully not for the worse...
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@586 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 
-Many theoretical models consist of more than one equation. If these
-models are econometrically estimated, it may not be appropriate to
-estimate the equations individually, but all together, because the
-disturbance terms of these equations are likely to be contemporaneously
-correlated. %\citep{theil71} (make sure this is the correct paper).
-Estimating all equations simultaneously, taking the
-covariance structure of the residuals into account, leads to efficient
+Many theoretical models consist of more than one equation. These
+models can contain variables that appear on the left-hand side in one
+equation and on the right-hand side of another equation
+(simultaneity-bias). The system can have non-zero off-diagonal
+covariance elements resulting from related disturbnaces
+(contemporaneous correlation). Ignoring contemporaneous correlation
+leads to inefficient parameter estimates \citep{zellner62} and
+ignoring simultaneity-bias can lead to inconsistent parameter
 estimates.
-Ignoring this correlation leads to an inefficient
-parameter estimation \citep{zellner62}.
+
+% If these models are econometrically estimated, it may not be
+% appropriate to estimate the equations individually, but all together,
+% because the disturbance terms of these equations are likely to be
+% contemporaneously correlated.  Estimating all equations
+% simultaneously, taking the covariance structure of the residuals into
+% account, leads to efficient estimates.  Ignoring this correlation
+% leads to an inefficient parameter estimation \citep{zellner62}.
 
@@
 cross-equation parameter restrictions. These restrictions can be
-tested and/or imposed only in a simultaneous estimation
-approach especially the economic theory suggests many cross-equation
-restrictions.
+tested and/or imposed only in a simultaneous estimation approach
+especially the economic theory suggests many cross-equation
+restrictions. (But we don't say why this is important...) Are we
+imposing restrictions in the parameters, or the variance-covariance
+matrix or both and for what pupose?
+
+For all of the methods developed in the package, the disturbnaces of
+the individual equations are assumed to be independent and identially
+distributed (iid).  
+In the future, we would like to add the ability to fit equations were
+the disturbances are serially correlated (wikins 1969).
 
@@
 requirements.
-Third, the (advanced) user can control many important details of the
-estimation procedure that are generally not available in other software
-packages.
-On the other hand, these user options have reasonable defaults that also
-beginners can easily use the provided functions.
+Third, the (advanced) user can control estimation details generally
+not available in other software packages by overriding reasonable defaults.
 

### 2006-02-14 paper/trunk/systemfit_intro.tex
really small change after read. it's ready for submission!
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@589 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 especially the economic theory suggests many cross-equation
-restrictions. (But we don't say why this is important...) Are we
-imposing restrictions in the parameters, or the variance-covariance
-matrix or both and for what pupose?
+restrictions.
 
-For all of the methods developed in the package, the disturbnaces of
-the individual equations are assumed to be independent and identially
-distributed (iid).  
-In the future, we would like to add the ability to fit equations were
-the disturbances are serially correlated (wikins 1969).
+% (But we don't say why this is important...) Are we
+% imposing restrictions in the parameters, or the variance-covariance
+% matrix or both and for what pupose?
 
-The \pkg{systemfit} package provides the capability to estimate
-linear and non-linear equation systems in \proglang{R}.
-Although linear and non-linear equation systems can be estimated
-with several other statistical and econometric software packages
-(e.g.\ \proglang{SAS}, \proglang{EViews}, \proglang{TSP}),
-\pkg{systemfit} has several advantages.
-First, all estimation procedures are publicly available in the source code.
-Second, the estimation algorithms can be easily modified to meet specific
-requirements.
-Third, the (advanced) user can control estimation details generally
-not available in other software packages by overriding reasonable defaults.
+% For all of the methods developed in the package, the disturbnaces of
+% the individual equations are assumed to be independent and identially
+% distributed (iid).  
+% In the future, we would like to add the ability to fit equations were
+% the disturbances are serially correlated (wikins 1969).
 
-This paper is organized as follows: In section~\ref{sec:Estimation} we
+The \pkg{systemfit} package provides the capability to estimate linear
+and non-linear equation systems in \proglang{R}.  Although linear and
+non-linear equation systems can be estimated with several other
+statistical and econometric software packages (e.g.\ \proglang{SAS},
+\proglang{EViews}, \proglang{TSP}), \pkg{systemfit} has several
+advantages.  First, all estimation procedures are publicly available
+in the source code.  Second, the estimation algorithms can be easily
+modified to meet specific requirements.  Third, the (advanced) user
+can control estimation details generally not available in other
+software packages by overriding reasonable defaults.
+
+This paper is organized as follows: In Section~\ref{sec:Estimation} we
 introduce the mathematics of estimating equation systems.
 Section~\ref{sec:Restrictions} shows how linear restrictions can be
-imposed.
-Section~\ref{sec:Usage} demonstrates how to run
+imposed.  Section~\ref{sec:Usage} demonstrates how to run
 \pkg{systemfit}, especially how the features presented in the previous
-sections can be used.
-All other relevant issues are discussed in section~\ref{sec:Other}.
-Finally, a summary and outlook are presented in
-section~\ref{sec:Summmary}.
+sections can be used.  All other relevant issues are discussed in
+Section~\ref{sec:Other}.  Finally, a summary and outlook are presented
+in Section~\ref{sec:Summmary}.
 

### 2006-02-14 paper/trunk/systemfit_other.tex
made seemingly massive changes... but hopefully not for the worse...
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@586 591db2bd-e0a0-4d22-8c19-2192dbb7da31
really small change after read. it's ready for submission!
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@589 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 
+Don't we already mention this in the usage section.
+
 Different approaches to determine the degrees of
@@
 \textbf{Arne, do you have a citation for this?}
+
+\paragraph{Degrees of freedom for t-tests}   
+\code{probdfsys} use the degrees of freedom of the whole system
+   (in place of the degrees of freedom of the single equation)
+   to calculate prob values for the t-test of individual parameters.
 
@@
 alternatives for which $\hat{\theta}$ and $\tilde{\theta}$ diverge
-under misspecification.
-The resulting test statistic, distributed as a
-$\chi^2$, is the difference between the covariance of an
-efficient and the covariance of the inefficient estimator is zero, if the
-model is correctly specified.
+under misspecification.  The resulting test statistic, distributed as
+a $\chi^2$, is the difference between the covariance of an efficient
+estimator and the covariance of the inefficient estimator. If the
+difference is zero, the model is correctly specified.
 
-The test is based on the idea that under the hypothesis of no
+The test is based on the idea that under the null hypothesis of no
 correlation, both the limited information estimates and the
@@
 The null hypotheses of the test is that all exogenous variables are
-uncorrelated with all disturbance terms.
-Under this hypothesis both the 2SLS and the 3SLS estimator are consistent
-but only the 3SLS estimator is (asymptotically) efficient.
-Under the alternative hypothesis the 2SLS estimator is consistent but the
-3SLS estimator is inconsistent.
+uncorrelated with all disturbance terms.  Under this hypothesis both
+the 2SLS and the 3SLS estimator are consistent but only the 3SLS
+estimator is (asymptotically) efficient.  Under the alternative
+hypothesis the 2SLS estimator is consistent but the 3SLS estimator is
+inconsistent.
 
@@
 \begin{equation}
-m = ( b_2 - b_3 )^{'} ( V_2 - V_3 ) ( b_2 - b_3 ) 
+m = ( \beta_2 - \beta_3 )^{'} ( \Omega_2 - \Omega_3 ) ( \beta_2 - \beta_3 ) 
 \label{eq:hausman}
 \end{equation}
-where $b_2$ and $V_2$ are the estimated coefficients and the
-variance-covariance matrix from a 2SLS estimation and $b_3$ and $V_3$
-are the estimated coefficients and their variance covariance-matrix
-from a 3SLS estimation.
+where $\beta_2$ and $\Omega_2$ are the estimated coefficients and the
+variance-covariance matrix from a 2SLS estimation and $\beta_3$ and
+$\Omega_3$ are the estimated coefficients and their variance
+covariance-matrix from a 3SLS estimation.
 
@@
 As linear systems of equations can be estimated using
-\code{systemfit}, nonlinear systems of equations \citep{hausman75}
-can be estimated using a similar function \code{nlsystemfit}.
-The usage for the nonlinear estimation function is similar to
-\code{systemfit} and can estimate system of equations using OLS, SUR,
-2SLS and 3SLS methods.
+\code{systemfit}, nonlinear systems of equations \citep{hausman75} can
+be estimated using a similar function; \code{nlsystemfit}.  The usage
+for the nonlinear estimation function is similar to \code{systemfit}
+and can estimate system of equations using OLS, SUR, 2SLS and 3SLS
+methods. The equations are defined as,
 
@@
 to those described for \code{systemfit}, and \code{start} is a list of
-starting values of the parameter estimates.
-The nlsystemfit function relies on \code{nlm} to perform the minimization
-of the objective functions and the \code{qr} set of functions.
-If the user does not have a set of estimates for the initial parameters,
-it is suggested using linearized forms in one of the linear methods,
-or simply using \code{nls} to obtain estimates for the equations.
+starting values of the parameter estimates.  The nlsystemfit function
+relies on \code{nlm} to perform the minimization of the objective
+functions and the \code{qr} set of functions.  If the user does not
+have a set of estimates for the initial parameters, it is suggested
+using linearized forms in one of the linear methods, or simply using
+\code{nls} to obtain estimates for the equations. The outputs are
+similar to those obtained by \code{systemfit} objects.
 

### 2006-02-15 paper/trunk/systemfit_intro.tex
restored line breaks of revision 35 to improve output of "svn diff"
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@590 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 
-The \pkg{systemfit} package provides the capability to estimate linear
-and non-linear equation systems in \proglang{R}.  Although linear and
-non-linear equation systems can be estimated with several other
-statistical and econometric software packages (e.g.\ \proglang{SAS},
-\proglang{EViews}, \proglang{TSP}), \pkg{systemfit} has several
-advantages.  First, all estimation procedures are publicly available
-in the source code.  Second, the estimation algorithms can be easily
-modified to meet specific requirements.  Third, the (advanced) user
-can control estimation details generally not available in other
-software packages by overriding reasonable defaults.
+The \pkg{systemfit} package provides the capability to estimate
+linear and non-linear equation systems in \proglang{R}.
+Although linear and non-linear equation systems can be estimated
+with several other statistical and econometric software packages
+(e.g.\ \proglang{SAS}, \proglang{EViews}, \proglang{TSP}),
+\pkg{systemfit} has several advantages.
+First, all estimation procedures are publicly available in the source code.
+Second, the estimation algorithms can be easily modified to meet specific
+requirements.
+Third, the (advanced) user can control estimation details generally
+not available in other software packages by overriding reasonable defaults.
 
@@
 Section~\ref{sec:Restrictions} shows how linear restrictions can be
-imposed.  Section~\ref{sec:Usage} demonstrates how to run
+imposed.
+Section~\ref{sec:Usage} demonstrates how to run
 \pkg{systemfit}, especially how the features presented in the previous
-sections can be used.  All other relevant issues are discussed in
-Section~\ref{sec:Other}.  Finally, a summary and outlook are presented
-in Section~\ref{sec:Summmary}.
+sections can be used.
+All other relevant issues are discussed in Section~\ref{sec:Other}.
+Finally, a summary and outlook are presented in
+Section~\ref{sec:Summmary}.
 

### 2006-02-15 paper/trunk/systemfit_other.tex
restored line breaks of revision 35 to improve output of "svn diff"
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@590 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 alternatives for which $\hat{\theta}$ and $\tilde{\theta}$ diverge
-under misspecification.  The resulting test statistic, distributed as
-a $\chi^2$, is the difference between the covariance of an efficient
-estimator and the covariance of the inefficient estimator. If the
-difference is zero, the model is correctly specified.
+under misspecification.
+The resulting test statistic, distributed as a
+$\chi^2$, is the difference between the covariance of an
+efficient estimator and the covariance of the inefficient estimator.
+If the difference is zero, the model is correctly specified.
 
@@
 The null hypotheses of the test is that all exogenous variables are
-uncorrelated with all disturbance terms.  Under this hypothesis both
-the 2SLS and the 3SLS estimator are consistent but only the 3SLS
-estimator is (asymptotically) efficient.  Under the alternative
-hypothesis the 2SLS estimator is consistent but the 3SLS estimator is
-inconsistent.
+uncorrelated with all disturbance terms.
+Under this hypothesis both the 2SLS and the 3SLS estimator are consistent
+but only the 3SLS estimator is (asymptotically) efficient.
+Under the alternative hypothesis the 2SLS estimator is consistent but the
+3SLS estimator is inconsistent.
 
@@
 As linear systems of equations can be estimated using
-\code{systemfit}, nonlinear systems of equations \citep{hausman75} can
-be estimated using a similar function; \code{nlsystemfit}.  The usage
-for the nonlinear estimation function is similar to \code{systemfit}
-and can estimate system of equations using OLS, SUR, 2SLS and 3SLS
-methods. The equations are defined as,
+\code{systemfit}, nonlinear systems of equations \citep{hausman75}
+can be estimated using a similar function; \code{nlsystemfit}.
+The usage for the nonlinear estimation function is similar to
+\code{systemfit} and can estimate system of equations using OLS, SUR,
+2SLS and 3SLS methods.
+The equations are defined as,
 
@@
 to those described for \code{systemfit}, and \code{start} is a list of
-starting values of the parameter estimates.  The nlsystemfit function
-relies on \code{nlm} to perform the minimization of the objective
-functions and the \code{qr} set of functions.  If the user does not
-have a set of estimates for the initial parameters, it is suggested
-using linearized forms in one of the linear methods, or simply using
-\code{nls} to obtain estimates for the equations. The outputs are
-similar to those obtained by \code{systemfit} objects.
+starting values of the parameter estimates.
+The nlsystemfit function relies on \code{nlm} to perform the minimization
+of the objective functions and the \code{qr} set of functions.
+If the user does not have a set of estimates for the initial parameters,
+it is suggested using linearized forms in one of the linear methods,
+or simply using \code{nls} to obtain estimates for the equations.
+The outputs are similar to those obtained by \code{systemfit} objects.
 

### 2006-02-15 paper/trunk/systemfit_other.tex
updated the hausman section.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@592 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 Hausman \citep{hausman78} developed a test have power against
-alternatives for which $\hat{\theta}$ and $\tilde{\theta}$ diverge
+alternatives for which $\hat{\beta}_2$ and $\hat{\beta}_3$ diverge
 under misspecification.
@@
 \begin{equation}
-m = ( \beta_2 - \beta_3 )^{'} ( \Omega_2 - \Omega_3 ) ( \beta_2 - \beta_3 ) 
+  m = ( \hat{\beta}_2 - \hat{\beta}_3 )^{'} ( \hat{\Omega}_2 - \hat{\Omega}_3 ) ( \hat{\beta}_2 - \hat{\beta}_3 ) 
 \label{eq:hausman}
 \end{equation}
-where $\beta_2$ and $\Omega_2$ are the estimated coefficients and the
-variance-covariance matrix from a 2SLS estimation and $\beta_3$ and
-$\Omega_3$ are the estimated coefficients and their variance
-covariance-matrix from a 3SLS estimation.
+where $\hat{\beta}_2$ and $\hat{\Omega}_2$ are the estimated
+coefficient and variance-covariance matrix from 2SLS estimation and
+$\hat{\beta}_3$ and $\hat{\Omega}_3$ are the estimated coefficients
+and variance covariance-matrix from 3SLS estimation.
 

### 2006-02-15 paper/trunk/systemfit_estimation.tex
small last changes
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@579 591db2bd-e0a0-4d22-8c19-2192dbb7da31
just changed some line breaks (deleted new ones and restored deleted ones) to make it easier to find the changes that Jeff committed in the last revision(s)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@582 591db2bd-e0a0-4d22-8c19-2192dbb7da31
changed a few line breaks that each sentence starts in a new line now
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@591 591db2bd-e0a0-4d22-8c19-2192dbb7da31
several minor revisions in sections Introduct

--- 2005-09-26 paper/trunk/systemfit_estimation.tex
@@
 \end{equation}
-
 These estimates are efficient only if the disturbance terms are not
 contemporaneously correlated, which means
-\begin{equation}
-   \sigma_{ij} = 0 \; \forall \; i \neq j
-\end{equation}
-
+$\sigma_{ij} = 0 \; \forall \; i \neq j$.
 If the whole system is treated as one single equation,
@@
 OLS estimation of the system leads to the same parameter estimates 
-as an equation-wise OLS estimation. The variance-covariance matrix 
-of the parameters from an equation-wise OLS estimation is equal to 
-the variance-covariance matrix obtained by equation 
-(\ref{eq:OLSvcm2}).
+as an equation-wise OLS estimation.
+The variance-covariance matrix of the parameters from an equation-wise
+OLS estimation is equal to the variance-covariance matrix obtained
+by equation (\ref{eq:OLSvcm2}).
 
@@
 $\sigma_{ii} = E \left( u_i' u_i \right)$.
-
 Like the OLS estimates these estimates are only efficient
 if the disturbance terms are not contemporaneously correlated.
-
 The variance-covariance matrix of the estimated parameters is
@@
 \end{equation}
-
 If no cross-equation parameter restrictions are imposed,
@@
 the regressors and the disturbance terms of each single 
-equation are uncorrelated,
+equation are uncorrelated
+$( E \left[ u_i | X_i \right] = 0 \; \forall \; i )$,
 a consistent SUR estimation requires that all disturbance terms and all 
-regressors are uncorrelated:
-\begin{equation}
-   E \left( u | X \right) = 0
-\end{equation}
+regressors are uncorrelated
+$( E \left[ u | X \right] = 0 )$.
 
 The SUR estimator can be obtained by:%
-\footnote{To calculate $\Omega^{-1}$ it is not necessary to invert 
+\footnote{
+To calculate $\Omega^{-1}$ it is not necessary to invert 
 the (huge) $\Omega$ matrix, 
 because it is sufficient to invert the (small) $\Sigma$ matrix and
-calculate it by $\Omega^{-1} = \Sigma^{-1} \otimes I$.}
-
+calculate it by $\Omega^{-1} = \Sigma^{-1} \otimes I$
+\citep[p.\ 342]{greene03}.
+}
 \begin{equation}
@@
 $\sigma_{ij} = E \left( u_i' u_j \right)$.
-
 And the variance-covariance matrix of the estimated parameters is
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Two-stage least squares (2SLS)}
+\subsection{Two-Stage least squares (2SLS)}
 
@@
 the estimated coefficients are biased.
-This can be circumvented by an instrumental variable (IV) estimation.
+This can be circumvented by a two-stage least squares (2SLS) estimation.
 The instrumental variables for each equation $H_i$ 
@@
 \end{equation}
-
 At the second stage the unbiased two-stage least squares estimates
@@
 \end{equation}
-
 If the whole system is treated as one single equation,
@@
 with $\sigma^2 = E \left( u' u \right)$.
-
 If the disturbance terms of the individual equations
@@
 $\sigma_{ii} = E \left( u_i' u_i \right)$.
-
 The variance-covariance matrix of the estimated parameters is
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Three-stage least squares (3SLS)}
+\subsection{Three-Stage least squares (3SLS)}
 
@@
 the disturbances are contemporaneously correlated, 
-an Instrumental Variable (IV) Generalized Least Squares (GLS) 
-estimation leads to consistent and efficient estimates. This estimation 
-procedure is generally called °Three-stage Least Squares° (3SLS).
+a Generalized Least Squares (GLS) version of the two-stage least squares
+estimation leads to consistent and efficient estimates.
+This estimation procedure is generally called °Three-stage Least
+Squares° (3SLS).
 
@@
 $\sigma_{ij} = E \left( u_i' u_j \right)$.
-
 Its variance-covariance matrix is:
@@
 \end{equation}
-
 While an unbiased 2SLS or W2SLS estimation requires only that
 the instrumental variables and the disturbance terms of each single 
-equation are uncorrelated,
+equation are uncorrelated
+$( E \left[ u_i | H_i \right]) = 0 \; \forall \; i )$,
 \cite{schmidt90} points out that this estimator is only consistent 
-if all disturbance terms and all instrumental variables are uncorrelated:
-\begin{equation}
-   E \left( u | H \right) = 0
-\end{equation}
+if all disturbance terms and all instrumental variables are uncorrelated
+$( E \left[ u | H \right]) = 0 )$
 with
@@
       0 & 0 & \cdots & H_G
-   \end{array}\right] \cdot 
+   \end{array}\right]
 \end{equation}
-
 Since there might be occasions where this cannot be avoided,
@@
 \end{equation}
-
 The variance-covariance matrix of this IV-3SLS estimator is:
@@
 \end{equation}
-
 Another approach is based on the Generalized Method of Moments (GMM)
@@
 \end{equation}
-
 The variance-covariance matrix of the GMM-3SLS estimator is:
@@
 \end{equation}
-
 A fourth approach developed by \cite{schmidt90} himself is:
@@
 \end{equation}
-
 The variance-covariance matrix of this estimator is:
@@
 \end{equation}
-
 The econometrics software EViews uses following approach:
@@
 is that the instruments of one equation are correlated with the
-disturbance terms in another equation. Otherwise, one would simply use
-every instrument in every equation \citep{schmidt90}. 
+disturbance terms in another equation.
+Otherwise, one would simply use every instrument in every equation
+\citep{schmidt90}.
 In this case, only the GMM-3SLS (\ref{eq:beta3slsGMM})

### 2006-02-15 paper/trunk/systemfit_intro.tex
changed a few line breaks that each sentence starts in a new line now
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@591 591db2bd-e0a0-4d22-8c19-2192dbb7da31
several minor revisions in sections Introduction, Estimation, Restrictions, and Other Issues
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@593 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 
-Many theoretical models consist of more than one equation. These
-models can contain variables that appear on the left-hand side in one
-equation and on the right-hand side of another equation
-(simultaneity-bias). The system can have non-zero off-diagonal
-covariance elements resulting from related disturbnaces
-(contemporaneous correlation). Ignoring contemporaneous correlation
-leads to inefficient parameter estimates \citep{zellner62} and
-ignoring simultaneity-bias can lead to inconsistent parameter
-estimates.
+Many theoretical models consist of more than one equation.
+These models can contain variables that appear on the left-hand side in one
+equation and on the right-hand side of another equation (simultaneity-bias).
+The system can have non-zero off-diagonal covariance elements resulting from
+related disturbances (contemporaneous correlation).
+Ignoring contemporaneous correlation leads to inefficient parameter
+estimates \citep{zellner62} and ignoring simultaneity-bias can lead to
+inconsistent parameter estimates.
 
@@
 Another reason to estimate an equation system simultaneously are
-cross-equation parameter restrictions. These restrictions can be
-tested and/or imposed only in a simultaneous estimation approach
-especially the economic theory suggests many cross-equation
-restrictions.
+cross-equation parameter restrictions.
+These restrictions can be tested and/or imposed only in a simultaneous
+estimation approach.
+Especially the economic theory suggests many cross-equation parameter
+restrictions (e.g.\ the symmetry restriction in demand models).
 
@@
 
-% For all of the methods developed in the package, the disturbnaces of
+% For all of the methods developed in the package, the disturbances of
 % the individual equations are assumed to be independent and identially
@@
 The \pkg{systemfit} package provides the capability to estimate
-linear and non-linear equation systems in \proglang{R}.
+linear and non-linear equation systems in \proglang{R}
+\citep{r-project}.
 Although linear and non-linear equation systems can be estimated

### 2006-02-15 paper/trunk/systemfit_other.tex
several minor revisions in sections Introduction, Estimation, Restrictions, and Other Issues
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@593 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 \emph{estimated} residuals. 
-
-The estimated covariance matrix of the residuals ($\widehat{\Sigma} = 
-\left[ \widehat{\sigma}_{ij} \right]$) 
-is often calculated by
+Generally, the estimated covariance matrix of the residuals
+($\widehat{\Sigma} = \left[ \widehat{\sigma}_{ij} \right]$)
+can be calculated from the residuals of a first-step OLS or 2SLS estimation.
+The following formula is often applied:
 \begin{equation}
@@
    { T - \max \left( K_i , K_j \right) }
+   \label{eq:rcov3}
 \end{equation}
@@
 \citep[p.\ 469]{judge85}. 
-% Greene (2002, p. 344) says that the second is unbiased if i = j or K_i = K_j,
+% Greene (2003, p. 344) says that the second is unbiased if i = j or K_i = K_j,
 % whereas the first is unbiased only if i = j. 
@@
 if the estimated residual variance-covariance matrix is calculated
-using the residuals from a first step OLS or 2SLS estimation.
+using the residuals from a first-step OLS or 2SLS estimation.
 It is also possible to iterate the estimation procedure and 
@@
 residual variance-covariance matrix, the results are maximum
-likelihood \citep[p.\ 345]{greene02}.
+likelihood \citep[p.\ 345]{greene03}.
 
@@
 McElroy's $R^2$ value \citep{mcelroy77}: 
-% also: \citep[p.\ 345]{greene02}
+% also: \citep[p.\ 345]{greene03}
 \begin{equation}

### 2006-02-15 paper/trunk/systemfit_restrictions.tex
changed a few line breaks that each sentence starts in a new line now
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@591 591db2bd-e0a0-4d22-8c19-2192dbb7da31
several minor revisions in sections Introduction, Estimation, Restrictions, and Other Issues
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@593 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 \end{equation}
-
 Then, a standard estimation is done by substituting $X^*$ for $X$.
@@
 \end{equation}
-
 The second way to impose linear parameter restrictions
@@
 where $\beta^0$ is the vector of the restricted coefficients, 
-$R$ is a matrix and $q$ is a vector (see \citealp[p.\ 100]{greene02}). 
+$R$ is a matrix and $q$ is a vector (see \citealp[p.\ 100]{greene03}).
 Each restriction belongs to one row of $R$ and the corresponding 
@@
 The first way is less flexible than this latter one%
-\footnote{e.g.\ restrictions like $\beta_1 + \beta_2 = 4$ cannot be imposed
-by the first method.}, 
+\footnote{
+While restrictions like $\beta_1 = 2 \beta_2$ can be imposed by
+both methods,
+restrictions like $\beta_1 + \beta_2 = 4$ can be imposed only
+by the second method.
+}, 
 but the first way is preferable if equality constraints for coefficients
@@
 
-These two methods can be combined. In this case the restrictions
-imposed using the latter method are imposed on the linear independent 
-parameters due to the restrictions imposed using the first method:
+These two methods can be combined.
+In this case the restrictions imposed using the latter method are
+imposed on the linear independent parameters due to the restrictions
+imposed using the first method:
 \begin{equation}
@@
 \end{equation}
-where $2\lambda$ is a vector of the Lagrangean multipliers of the restrictions.
-
+where $\lambda$ is a vector of the Lagrangean multipliers of the restrictions.
 If the whole system is treated as one single equation,
@@
 $\sigma_{ij} = E \left( u_i' u_j \right)$.
-
 The variance-covariance matrix of the estimated parameters is
@@
 $\sigma_{ij} = E \left( u_i' u_j \right)$.
-
 The variance-covariance matrix of this estimator is
@@
 \end{equation}
-
 The IV-3SLS estimator restricted by $R \beta^0 = q$ can be obtained by
@@
 \end{equation}
-
 with
@@
 \end{equation}
-
 The restricted 3SLS-GMM estimator can be obtained by
@@
 \end{equation}
-
 The restricted 3SLS estimator based on the suggestion of
@@
 \end{eqnarray}
-
 The econometrics software EViews calculates the restricted 3SLS estimator by:
@@
 To calculate the variance-covariance matrix 
-EViews uses the standard formula of the restricted 3SLS estimator
-(\ref{eq:cov3slsr}).
+EViews uses the standard formula of the restricted 3SLS
+estimator~(\ref{eq:cov3slsr}).
 

### 2006-02-15 paper/trunk/systemfit_estimation.tex
considerably revised and extended "Usage" section
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@594 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
    Var \left[ \widehat{\beta} \right] = \sigma^2 \left( X'X \right)^{-1}
+   \label{eq:olsVarSameSigma}
 \end{equation}
@@
    Var \left[ \widehat{\beta} \right] = \left( X' \Omega^{-1} X \right)^{-1}
-   \label{eq:OLSvcm2}
+   \label{eq:olsVarSingleSigma}
 \end{equation}
@@
 OLS estimation is equal to the variance-covariance matrix obtained
-by equation (\ref{eq:OLSvcm2}).
+by equation (\ref{eq:olsVarSingleSigma}).
 
@@
    \widehat{X} \right)^{-1}
+   \label{eq:2slsVarSameSigma}
 \end{equation}
@@
    \widehat{X} \right)^{-1}
+   \label{eq:2slsVarSingleSigma}
 \end{equation}
@@
 \subsection{Three-Stage least squares (3SLS)}
+\label{sec:3sls}
 
@@
    \right)^{-1} \widehat{X}' \Omega^{-1} y
+   \label{eq:3slsGls}
 \end{equation}
@@
    \right)^{-1} \widehat{X}' \Omega^{-1} y
+   \label{eq:3slsIv}
 \end{equation}
@@
    H' X \right)^{-1} X' H \left( H' \Omega H \right)^{-1} H' y
-   \label{eq:beta3slsGMM}
+   \label{eq:3slsGmm}
 \end{equation}
@@
    H \left( H' H \right)^{-1} H' y 
-   \label{eq:beta3slsSchmidt}
+   \label{eq:3slsSchmidt}
 \end{equation}
@@
    \widehat{X}' \Omega^{-1} \left( y - X \widehat{\beta}_{2SLS} \right)
+   \label{eq:3slsEViews}
 \end{equation}
@@
 \citep{schmidt90}.
-In this case, only the GMM-3SLS (\ref{eq:beta3slsGMM})
+In this case, only the GMM-3SLS (\ref{eq:3slsGmm})
 and the 3SLS estimator developed by \cite{schmidt90} 
-(\ref{eq:beta3slsSchmidt}) are consistent.
+(\ref{eq:3slsSchmidt}) are consistent.
 

### 2006-02-16 paper/trunk/systemfit_other.tex
revised section "other issues" (mainly subsections "degrees of freedom" and "hausman test"); argument "probdfsys" in section "using systemfit" now refers to the subsection "degrees of freedom" in section  "other issues"
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@598 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 where $T$ is the number of observations in each equation.
-
 However, this estimator is biased, because it is not corrected for
@@
 $i$ and $j$, respectively.
-
-However, this yields an unbiased estimator only if $K_i = K_j$ 
+However, these formulas yield unbiased estimators only if $K_i = K_j$
 \citep[p.\ 469]{judge85}. 
@@
    { T - K_i - K_j + tr \left[ X_i \left( X_i' X_i \right)^{-1}
-   X_i' X_j \left( X_j' X_j \right)^{-1} X_j' \right] } \\
+   X_i' X_j \left( X_j' X_j \right)^{-1} X_j' \right] }
+   \label{eq:rcov2} \\
    & = &
@@
    { T - K_i - K_j + tr \left[ \left( X_i' X_i \right)^{-1}
-   X_i' X_j \left( X_j' X_j \right)^{-1} X_j' X_i \right] } \\
-   \label{eq:rcov2}
+   X_i' X_j \left( X_j' X_j \right)^{-1} X_j' X_i \right] }
 \end{eqnarray} 
@@
 If equation (\ref{eq:rcov0}) is applied to calculate the estimated
-residual variance-covariance matrix, the results are maximum
+residual variance-covariance matrix, the results converge to maximum
 likelihood \citep[p.\ 345]{greene03}.
@@
 \subsection{Degrees of freedom}
+\label{sec:degreesOfFreedom}
 
-Don't we already mention this in the usage section.
+To our knowledge the question about how to determine the degrees
+of freedom for single-parameter t-tests is not comprehensively
+discussed in the literature.
+While sometimes the degrees of freedom of the whole system
+(total number of observations in all equations minus
+total number of estimated parameters)
+are applied,
+in other cases the degrees of freedom of each single equation
+(number of observations in the equations minus
+number of estimated parameters in the equation)
+are used.
+Asymptotically, this distinction doesn't make a difference.
+However, in many empirical applications, the number of observations
+of each equation is rather small, and
+therefore, this makes a difference.
 
-Different approaches to determine the degrees of
-freedom for testing the estimated parameters with a t-test
-\textbf{(cite this one)}.
+If a system of equations is estimated by unrestricted OLS and
+the variance covariance matrix of the parameters is calculated
+by~(\ref{eq:olsVarSingleSigma}),
+the estimated parameters and their standard errors are identical
+to an equation-wise OLS estimation.
+In this case, it is reasonable to use the degrees of freedom of
+each single equation,
+because this yields also the same p-values as the equation-wise
+OLS estimation.
 
-\textbf{Arne, do you have a citation for this?}
-
-\paragraph{Degrees of freedom for t-tests}   
-\code{probdfsys} use the degrees of freedom of the whole system
-   (in place of the degrees of freedom of the single equation)
-   to calculate prob values for the t-test of individual parameters.
-
-The \code{probdfsys} can be used to set the degrees of freedom when
-calculating the probability values for the t-test for individual
-parameters.
-This argument can set the degrees of freedom for the entire system
-instead of using the degrees of freedom for each of the
-individual equations.
+In contrast, if a system of equations is estimated with many
+cross-equation restrictions and
+the variance covariance matrix of an OLS estimation is calculated
+by~(\ref{eq:olsVarSameSigma}),
+the system estimation is similar to a single equation estimation.
+Therefore, in this case, it seems to be reasonable to use the degrees
+of freedom of whole system.
 
@@
 
-The goodness of fit of each equation can be measured by the 
+The goodness of fit of each single equation can be measured by the
 traditional $R^2$ values:
@@
 
-Hausman \citep{hausman78} developed a test have power against
-alternatives for which $\hat{\beta}_2$ and $\hat{\beta}_3$ diverge
-under misspecification.
-The resulting test statistic, distributed as a
-$\chi^2$, is the difference between the covariance of an
-efficient estimator and the covariance of the inefficient estimator.
-If the difference is zero, the model is correctly specified.
-
-The test is based on the idea that under the null hypothesis of no
-correlation, both the limited information estimates and the
-full-information results are consistent, but only the full-information
-estimates are efficient and under the alternative, the limited
-information estimates are consistent, but the full information
-estimates are not.
-
+Hausman \citep{hausman78} developed a test for misspecification.
 The null hypotheses of the test is that all exogenous variables are
@@
 but only the 3SLS estimator is (asymptotically) efficient.
-Under the alternative hypothesis the 2SLS estimator is consistent but the
-3SLS estimator is inconsistent.
-
+Under the alternative hypothesis the 2SLS estimator is consistent
+but the 3SLS estimator is inconsistent.
+The resulting test statistic, distributed as a
+$\chi^2$, is the difference between the covariance of an
+3SLS estimator and the covariance of an 2SLS estimator.
+If the difference is (close to) zero, the model is correctly specified.
 The Hausman test statistic is,
-
 \begin{equation}
-  m = ( \hat{\beta}_2 - \hat{\beta}_3 )^{'} ( \hat{\Omega}_2 - \hat{\Omega}_3 ) ( \hat{\beta}_2 - \hat{\beta}_3 ) 
+  m = \left( \hat{\beta}_2 - \hat{\beta}_3 \right)^{'}
+      \left( Var \left[ \hat{\beta}_2 \right] -
+             Var \left[ \hat{\beta}_3 \right] \right)
+      \left( \hat{\beta}_2 - \hat{\beta}_3 \right)
 \label{eq:hausman}
 \end{equation}
-where $\hat{\beta}_2$ and $\hat{\Omega}_2$ are the estimated
-coefficient and variance-covariance matrix from 2SLS estimation and
-$\hat{\beta}_3$ and $\hat{\Omega}_3$ are the estimated coefficients
-and variance covariance-matrix from 3SLS estimation.
+where $\hat{\beta}_2$ and $Var \left[ \hat{\beta}_2 \right]$ are the estimated
+coefficient and variance-covariance matrix from 2SLS estimation, and
+$\hat{\beta}_3$ and $Var \left[ \hat{\beta}_3 \right]$ are the estimated
+coefficients and variance covariance-matrix from 3SLS estimation.
 

### 2006-02-16 paper/trunk/systemfit_other.tex
added description of methods "WSUR" and "W3SLS"
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@599 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 using the residuals from a first-step OLS or 2SLS estimation.
+The software LIMDEP calculates SUR or 3SLS estimates by a
+slightly different approach.
+It uses the residuals of a first-step OLS or 2SLS estimation
+for a second-step WLS or W2SLS estimation,
+and then calculates the residual variance covariance matrix
+from the residuals of the WLS or W2SLS estimation
+to estimates the model by SUR or 3SLS in a third step
+\citep{greene06}.
 It is also possible to iterate the estimation procedure and

### 2006-02-16 paper/trunk/systemfit_other.tex
added a subsection about how to use "hausman.systemfit" in section "using systemfit"; corrected description of hausman test in section "other issues"
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@600 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 \subsection{Hausman test}
+\label{sec:hausman}
 
@@
 but the 3SLS estimator is inconsistent.
-The resulting test statistic, distributed as a
-$\chi^2$, is the difference between the covariance of an
-3SLS estimator and the covariance of an 2SLS estimator.
-If the difference is (close to) zero, the model is correctly specified.
 The Hausman test statistic is,
@@
 coefficients and variance covariance-matrix from 3SLS estimation.
+Under the null hypotheses this test statistic has a
+$\chi^2$ distribution with degrees of freedom equal to the number of
+estimated parameters.
 

### 2006-02-16 paper/trunk/systemfit_other.tex
revised paragraph about the WSUR and W3SLS method and paragraph about iterated WLS/SUR/W2SLS/3SLS
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@601 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 using the residuals from a first-step OLS or 2SLS estimation.
-The software LIMDEP calculates SUR or 3SLS estimates by a
-slightly different approach.
+There exists also an alternative slightly different approach.%
+\footnote{
+For instance, this approach is applied by
+the command °TSCS° of the software LIMDEP that carries out SUR estimations
+in which all coefficient vectors are constrained to be equal
+\citep{greene06}.
+}
 It uses the residuals of a first-step OLS or 2SLS estimation
-for a second-step WLS or W2SLS estimation,
-and then calculates the residual variance covariance matrix
-from the residuals of the WLS or W2SLS estimation
-to estimates the model by SUR or 3SLS in a third step
-\citep{greene06}.
-It is also possible to iterate the estimation procedure and
-calculate the residual variance-covariance matrix from the
-residuals of the previous iteration. 
+to apply a WLS or W2SLS estimation on a second step.
+Then, it calculates the residual variance covariance matrix
+from the residuals of the second-step estimation
+to estimates the model by SUR or 3SLS in a third step.
+If no cross-equation restrictions are imposed,
+the parameter estimates of OLS and WLS as well as 2SLS and W2SLS are identical.
+Hence, in this case the both approach generate the same results.
+
+It is also possible to iterate WLS, SUR, W2SLS and 3SLS estimations.
+At each iteration the residual variance-covariance matrix is calculated
+from the residuals of the previous iteration.
 If equation (\ref{eq:rcov0}) is applied to calculate the estimated
-residual variance-covariance matrix, the results converge to maximum
+residual variance-covariance matrix,
+an iterated SUR estimation converges to maximum
 likelihood \citep[p.\ 345]{greene03}.

### 2006-02-17 paper/trunk/nlsystemfit.tex
substituted "covariance matrix" for "variance-covariance matrix" and "\Cov [ . ]" for "Var [ . ]", where "\Cov" is a command for "\text{Cov}"
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@605 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 matrix of the instrument variables, and $I$ is an $n \times n $
-identity matrix and $S$ is the estimated variance-covariance matrix between the
+identity matrix and $S$ is the estimated covariance matrix between the
 equations
@@
 
-\textbf{The residual variance-covariance matrix can be calculated in 
+\textbf{The residual covariance matrix can be calculated in
 different ways (section~\ref{sec:residcov}). 

### 2006-02-17 paper/trunk/systemfit_estimation.tex
added description of methods "WSUR" and "W3SLS"
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@599 591db2bd-e0a0-4d22-8c19-2192dbb7da31
substituted "covariance matrix" for "variance-covariance matrix" and "\Cov [ . ]" for "Var [ . ]", where "\Cov" is a command for "\text{Cov}"
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@605 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 \subsection{Ordinary least squares (OLS)}
+\label{sec:ols}
 
@@
 If the whole system is treated as one single equation,
-the variance-covariance matrix of the estimated parameters is
+the covariance matrix of the estimated parameters is
 \begin{equation}
-   Var \left[ \widehat{\beta} \right] = \sigma^2 \left( X'X \right)^{-1}
-   \label{eq:olsVarSameSigma}
+   \Cov \left[ \widehat{\beta} \right] = \sigma^2 \left( X'X \right)^{-1}
+   \label{eq:olsCovSameSigma}
 \end{equation}
@@
 are allowed to have different variances,
-the variance-covariance matrix of the estimated parameters is
+the covariance matrix of the estimated parameters is
 \begin{equation}
-   Var \left[ \widehat{\beta} \right] = \left( X' \Omega^{-1} X \right)^{-1}
-   \label{eq:olsVarSingleSigma}
+   \Cov \left[ \widehat{\beta} \right] = \left( X' \Omega^{-1} X \right)^{-1}
+   \label{eq:olsCovSingleSigma}
 \end{equation}
@@
 as an equation-wise OLS estimation.
-The variance-covariance matrix of the parameters from an equation-wise
-OLS estimation is equal to the variance-covariance matrix obtained
-by equation (\ref{eq:olsVarSingleSigma}).
+The covariance matrix of the parameters from an equation-wise
+OLS estimation is equal to the covariance matrix obtained
+by equation (\ref{eq:olsCovSingleSigma}).
 
@@
 if the disturbance terms are not contemporaneously correlated.
-The variance-covariance matrix of the estimated parameters is
+The covariance matrix of the estimated parameters is
 \begin{equation}
-   Var \left[ \widehat{\beta} \right] = \left( X' \Omega^{-1} X \right)^{-1}
+   \Cov \left[ \widehat{\beta} \right] = \left( X' \Omega^{-1} X \right)^{-1}
 \end{equation}
@@
 $\sigma_{ij} = E \left( u_i' u_j \right)$.
-And the variance-covariance matrix of the estimated parameters is
+And the covariance matrix of the estimated parameters is
 \begin{equation}
-   Var \left[ \widehat{\beta} \right] = \left( X' \Omega^{-1} X \right)^{-1}
+   \Cov \left[ \widehat{\beta} \right] = \left( X' \Omega^{-1} X \right)^{-1}
 \end{equation}
@@
 If the whole system is treated as one single equation,
-the variance-covariance matrix of the estimated parameters is
+the covariance matrix of the estimated parameters is
 \begin{equation}
-   Var \left[ \widehat{\beta} \right] = \sigma^2 \left( \widehat{X}'
+   \Cov \left[ \widehat{\beta} \right] = \sigma^2 \left( \widehat{X}'
    \widehat{X} \right)^{-1}
-   \label{eq:2slsVarSameSigma}
+   \label{eq:2slsCovSameSigma}
 \end{equation}
@@
 are allowed to have different variances, 
-the variance-covariance matrix of the estimated parameters is
+the covariance matrix of the estimated parameters is
 \begin{equation}
-   Var \left[ \widehat{\beta} \right] = \left( \widehat{X}' \Omega^{-1} 
+   \Cov \left[ \widehat{\beta} \right] = \left( \widehat{X}' \Omega^{-1}
    \widehat{X} \right)^{-1}
-   \label{eq:2slsVarSingleSigma}
+   \label{eq:2slsCovSingleSigma}
 \end{equation}
@@
 $\sigma_{ii} = E \left( u_i' u_i \right)$.
-The variance-covariance matrix of the estimated parameters is
+The covariance matrix of the estimated parameters is
 \begin{equation}
-   Var \left[ \widehat{\beta} \right] = \left( \widehat{X}' \Omega^{-1} 
+   \Cov \left[ \widehat{\beta} \right] = \left( \widehat{X}' \Omega^{-1}
    \widehat{X} \right)^{-1}
@@
 $\sigma_{ij} = E \left( u_i' u_j \right)$.
-Its variance-covariance matrix is:
+Its covariance matrix is:
 \begin{equation}
-   Var \left[ \widehat{\beta} \right] = \left( \widehat{X}' \Omega^{-1} 
+   \Cov \left[ \widehat{\beta} \right] = \left( \widehat{X}' \Omega^{-1}
    \widehat{X} \right)^{-1}
@@
 \end{equation}
-The variance-covariance matrix of this IV-3SLS estimator is:
+The covariance matrix of this IV-3SLS estimator is:
 \begin{equation}
-   Var \left[ \widehat{\beta} \right] = \left( \widehat{X}' \Omega^{-1} 
+   \Cov \left[ \widehat{\beta} \right] = \left( \widehat{X}' \Omega^{-1}
    X \right)^{-1}
@@
 \end{equation}
-The variance-covariance matrix of the GMM-3SLS estimator is:
+The covariance matrix of the GMM-3SLS estimator is:
 \begin{equation}
-   Var \left[ \widehat{\beta} \right] = 
+   \Cov \left[ \widehat{\beta} \right] =
    \left( X' H \left( H' \Omega H \right)^{-1} H' X \right)^{-1}
@@
 \end{equation}
-The variance-covariance matrix of this estimator is:
+The covariance matrix of this estimator is:
 \begin{equation}
-   Var \left[ \widehat{\beta} \right] = 
+   \Cov \left[ \widehat{\beta} \right] =
    \left( \widehat{X}' \Omega^{-1}  \widehat{X} \right)^{-1} 
@@
 EViews uses the standard 3SLS formula (\ref{eq:cov3sls}) to
-calculate the variance-covariance matrix of the 3SLS estimator.
+calculate the covariance matrix of the 3SLS estimator.
 

### 2006-02-17 paper/trunk/systemfit_other.tex
substituted "covariance matrix" for "variance-covariance matrix" and "\Cov [ . ]" for "Var [ . ]", where "\Cov" is a command for "\text{Cov}"
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@605 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Residual variance-covariance matrix}\label{sec:residcov}
+\subsection{Residual covariance matrix}\label{sec:residcov}
 
@@
 might differ.
-Two alternative approaches to calculate the residual variance-covariance
+Two alternative approaches to calculate the residual covariance
 matrix are
@@
 
-A further approach to obtain the estimated residual variance-covariance
+A further approach to obtain the estimated residual covariance
 matrix is \citep[p.\ 309]{zellner62c}
@@
 of $\Sigma^{-1}$ \citep[p.\ 322]{theil71}.
-Furthermore, the variance-covariance matrix calculated by (\ref{eq:rcov2})
+Furthermore, the covariance matrix calculated by (\ref{eq:rcov2})
 is not necessarily positive semidefinite \citep[p.\ 322]{theil71}. 
@@
 The WLS, SUR, W2SLS and 3SLS parameter estimates are consistent,
-if the estimated residual variance-covariance matrix is calculated
+if the estimated residual covariance matrix is calculated
 using the residuals from a first-step OLS or 2SLS estimation.
@@
 to apply a WLS or W2SLS estimation on a second step.
-Then, it calculates the residual variance covariance matrix
+Then, it calculates the residual covariance matrix
 from the residuals of the second-step estimation
@@
 It is also possible to iterate WLS, SUR, W2SLS and 3SLS estimations.
-At each iteration the residual variance-covariance matrix is calculated
+At each iteration the residual covariance matrix is calculated
 from the residuals of the previous iteration.
 If equation (\ref{eq:rcov0}) is applied to calculate the estimated
-residual variance-covariance matrix,
+residual covariance matrix,
 an iterated SUR estimation converges to maximum
@@
 If a system of equations is estimated by unrestricted OLS and
-the variance covariance matrix of the parameters is calculated
-by~(\ref{eq:olsVarSingleSigma}),
+the covariance matrix of the parameters is calculated
+by~(\ref{eq:olsCovSingleSigma}),
 the estimated parameters and their standard errors are identical
@@
 cross-equation restrictions and
-the variance covariance matrix of an OLS estimation is calculated
-by~(\ref{eq:olsVarSameSigma}),
+the covariance matrix of an OLS estimation is calculated
+by~(\ref{eq:olsCovSameSigma}),
 the system estimation is similar to a single equation estimation.
@@
   m = \left( \hat{\beta}_2 - \hat{\beta}_3 \right)^{'}
-      \left( Var \left[ \hat{\beta}_2 \right] -
-             Var \left[ \hat{\beta}_3 \right] \right)
+      \left( \Cov \left[ \hat{\beta}_2 \right] -
+             \Cov \left[ \hat{\beta}_3 \right] \right)
       \left( \hat{\beta}_2 - \hat{\beta}_3 \right)
@@
 \end{equation}
-where $\hat{\beta}_2$ and $Var \left[ \hat{\beta}_2 \right]$ are the estimated
-coefficient and variance-covariance matrix from 2SLS estimation, and
-$\hat{\beta}_3$ and $Var \left[ \hat{\beta}_3 \right]$ are the estimated
-coefficients and variance covariance-matrix from 3SLS estimation.
+where $\hat{\beta}_2$ and $\Cov \left[ \hat{\beta}_2 \right]$ are the estimated
+coefficient and covariance matrix from 2SLS estimation, and
+$\hat{\beta}_3$ and $\Cov \left[ \hat{\beta}_3 \right]$ are the estimated
+coefficients and covariance matrix from 3SLS estimation.
 Under the null hypotheses this test statistic has a
@@
 % matrix of the instrument variables, and $I$ is an $n \times n $
-% identity matrix and $S$ is the estimated variance-covariance matrix between the
+% identity matrix and $S$ is the estimated covariance matrix between the
 % equations
@@
 
-% Do we actually have the variance-covariance function in place.
+% Do we actually have the covariance function in place.
 
@@
 
-% \textbf{The residual variance-covariance matrix can be calculated in
+% \textbf{The residual covariance matrix can be calculated in
 %   different ways (section~\ref{sec:residcov}).  It should be

### 2006-02-17 paper/trunk/systemfit_restrictions.tex
considerably revised and extended "Usage" section
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@594 591db2bd-e0a0-4d22-8c19-2192dbb7da31
added description of methods "WSUR" and "W3SLS"
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@599 591db2bd-e0a0-4d22-8c19-2192dbb7da31
substituted "covariance matrix" for "variance-covariance matrix" and "\Cov [ . ]" for "Var [ . ]", where "\Cov" is a command for "\text{Cov}"
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@605 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 This results in the linear independent parameter estimates $\beta^*$ and
-the variance-covariance matrix of these parameters. 
+the covariance matrix of these parameters.
 The original parameters can be obtained by equation (\ref{eq:T-restr})
-and the variance-covariance matrix of the original parameters 
+and the covariance matrix of the original parameters
 can be obtained by:
 \begin{equation}
-   Var \left[ \widehat{\beta} \right] = T \cdot Var \left[ \widehat{\beta^*} \right] \cdot T'
+   \Cov \left[ \widehat{\beta} \right] = T \cdot \Cov \left[ \widehat{\beta^*} \right] \cdot T'
 \end{equation}
@@
 \subsection{Restricted OLS estimation}
+\label{sec:olsR}
 
@@
 If the whole system is treated as one single equation,
-the variance-covariance matrix of the estimated parameters is
+the covariance matrix of the estimated parameters is
 \begin{equation}
-   Var 
+   \Cov 
    \left[ \begin{array}{c}
@@
 are allowed to have different variances, 
-the variance-covariance matrix of the estimated parameters is
+the covariance matrix of the estimated parameters is
 \begin{equation}
-   Var 
+   \Cov 
    \left[ \begin{array}{c}
@@
 $\sigma_{ii} = E \left( u_i' u_i \right)$.
-The variance-covariance matrix of the estimated parameters is
+The covariance matrix of the estimated parameters is
 \begin{equation}
-   Var 
+   \Cov 
    \left[ \begin{array}{c}
@@
 $\sigma_{ij} = E \left( u_i' u_j \right)$.
-The variance-covariance matrix of the estimated parameters is
+The covariance matrix of the estimated parameters is
 \begin{equation}
-   Var 
+   \Cov 
    \left[ \begin{array}{c}
@@
 \subsection{Restricted 3SLS estimation}
+\label{sec:3slsR}
 
@@
    \end{array} \right]
+   \label{eq:3slsGlsR}
 \end{equation}
@@
 $\sigma_{ij} = E \left( u_i' u_j \right)$.
-The variance-covariance matrix of this estimator is
+The covariance matrix of this estimator is
 \begin{equation}
-   Var 
+   \Cov 
    \left[ \begin{array}{c}
@@
    \end{array} \right]
+   \label{eq:3slsIvR}
 \end{equation}
@@
 \begin{equation}
-   Var 
+   \Cov 
    \left[ \begin{array}{c}
@@
    \end{array} \right]
+   \label{eq:3slsGmmR}
 \end{equation}
@@
    \end{array} \right]
+   \label{eq:3slsSchmidtR}
 \end{equation}
@@
 \begin{eqnarray}
-   Var 
+   \Cov 
    \left[ \begin{array}{c}
@@
    \end{array} \right]
+   \label{eq:3slsEViewsR}
 \end{equation}
@@
 by equation (\ref{eq:beta2SLSr}). 
-To calculate the variance-covariance matrix 
+To calculate the covariance matrix
 EViews uses the standard formula of the restricted 3SLS
@@
 all the above mentioned approaches lead to identical parameter estimates
-and identical variance-covariance matrices of the estimated parameters.
+and identical covariance matrices of the estimated parameters.
 

### 2006-02-17 trunk/R/lrtest.R
moved code for Hausman test and LR test into separate files
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@163 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@ -1,0 +1,23 @@
+## Likelihood Ratio Test
+lrtest.systemfit <- function( resultc, resultu ) {
+  lrtest <- list()
+  if( resultc$method %in% c( "SUR", "WSUR" ) &
+      resultu$method %in% c( "SUR", "WSUR" ) ) {
+    n   <- resultu$eq[[1]]$n
+    lrtest$df  <- resultu$ki - resultc$ki
+    if(resultc$rcovformula != resultu$rcovformula) {
+      stop( paste( "both estimations must use the same formula to calculate",
+                   "the residual covariance matrix!" ) )
+    }
+    if(resultc$rcovformula == 0) {
+      lrtest$lr  <- n * ( log( resultc$drcov ) - log( resultu$drcov ) )
+    } else {
+      residc <- array(resultc$resids,c(n,resultc$g))
+      residu <- array(resultu$resids,c(n,resultu$g))
+      lrtest$lr <- n * ( log( det( (t(residc) %*% residc)) ) -
+                         log( det( (t(residu) %*% residu))))
+    }
+    lrtest$p <- 1-pchisq( lrtest$lr, lrtest$df )
+  }
+  lrtest
+}

### 2006-02-17 trunk/R/systemfit.R

@@ -1137,61 +1137,6 @@ se.ratio.systemfit <- function( resultsi, resultsj, eqni ) {
 }
 
 
-## this function returns test statistic for
-## the hausman test
-# The m-statistic is then distributed with k degrees of freedom, where k
-# is the rank of the matrix .A generalized inverse is used, as
-# recommended by Hausman (1982).
-
-hausman.systemfit <- function( results2sls, results3sls ) {
-
-   result <- list()
-
-   result$q <- results2sls$b - results3sls$b
-   result$qVar <- results2sls$bcov - results3sls$bcov
-
-#    if( min( eigen( hausman$qVar )$values ) < 0 ) {
-#       warning( "the matrix V is not 'positive definite'" )
-#    }
-
-   result$statistic <- t( result$q ) %*% solve( result$qVar, result$q )
-   names( result$statistic ) <- "Hausman"
-   result$parameter <- nrow( result$qVar )
-   names( result$parameter ) <- "df"
-   result$p.value <- 1 - pchisq( result$statistic, result$parameter )
-   result$method = paste( "Hausman specification test for consistency of",
-      "the 3SLS estimation" )
-   result$data.name = results2sls$data.name
-   class( result ) <- "htest"
-   return( result )
-}
-
-
-## Likelihood Ratio Test
-lrtest.systemfit <- function( resultc, resultu ) {
-  lrtest <- list()
-  if( resultc$method %in% c( "SUR", "WSUR" ) &
-      resultu$method %in% c( "SUR", "WSUR" ) ) {
-    n   <- resultu$eq[[1]]$n
-    lrtest$df  <- resultu$ki - resultc$ki
-    if(resultc$rcovformula != resultu$rcovformula) {
-      stop( paste( "both estimations must use the same formula to calculate",
-                   "the residual covariance matrix!" ) )
-    }
-    if(resultc$rcovformula == 0) {
-      lrtest$lr  <- n * ( log( resultc$drcov ) - log( resultu$drcov ) )
-    } else {
-      residc <- array(resultc$resids,c(n,resultc$g))
-      residu <- array(resultu$resids,c(n,resultu$g))
-      lrtest$lr <- n * ( log( det( (t(residc) %*% residc)) ) -
-                         log( det( (t(residu) %*% residu))))
-    }
-    lrtest$p <- 1-pchisq( lrtest$lr, lrtest$df )
-  }
-  lrtest
-}
-
-
 ## return all coefficients
 coef.systemfit <- function( object, ... ) {
    object$b

### 2006-02-17 trunk/R/waldtest.R
revised Wald-test documentation and print method
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@166 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@ -22,6 +22,7 @@ print.waldtest.systemfit <- function( x, digits = 4, ... ){
    cat( "\n", "Wald-test for linear parameter restrictions",
       " in equation systems\n", sep = "" )
    cat( "Wald-statistic:", formatC( x$statistic, digits = digits ), "\n" )
-   cat( "p-value:", formatC( x$p.value, digits = digits ), "\n" )
+   cat( "degrees of freedom:", x$nRestr, "\n" )
+   cat( "p-value:", formatC( x$p.value, digits = digits ), "\n\n" )
    invisible( x )
 }

### 2006-02-17 trunk/R/ftest.R
revised F-test calculation, documentation and print method (it uses now the residual covariance matrix that was used for the estimation)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@167 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@ -4,9 +4,13 @@ ftest.systemfit <- function( object, R.restr,
    coef <- coef( object )
    vcov <- vcov( object )
    resid <- unlist( residuals( object ) )
-   rcov  <- object$rcov
    nEq   <- object$g
    nObsEq <- object$n / nEq
+   if( is.null( object$rcovest ) ) {
+      rcov <- diag( nEq )
+   } else {
+      rcov <- object$rcovest
+   }
 
    result <- list()
 
@@ -20,6 +24,7 @@ ftest.systemfit <- function( object, R.restr,
    denominator <- t( resid ) %*%
       ( solve( rcov ) %x% diag( nObsEq ) ) %*%
       resid
+   #print( denominator )
 
    result$statistic <- ( numerator / result$nRestr ) /
       ( denominator / result$dfSys )
@@ -34,6 +39,8 @@ print.ftest.systemfit <- function( x, digits = 4, ... ){
    cat( "\n", "F-test for linear parameter restrictions",
       " in equation systems\n", sep = "" )
    cat( "F-statistic:", formatC( x$statistic, digits = digits ), "\n" )
-   cat( "p-value:", formatC( x$p.value, digits = digits ), "\n" )
+   cat( "degrees of freedom of the numerator:", x$nRestr, "\n" )
+   cat( "degrees of freedom of the denominator:", x$dfSys, "\n" )
+   cat( "p-value:", formatC( x$p.value, digits = digits ), "\n\n" )
    invisible( x )
 }

### 2006-02-17 trunk/R/lrtest.R
considerably revised LR-test documentation, added print method, renamed objects returned by lrtest.systemfit
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@168 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@ -4,20 +4,31 @@ lrtest.systemfit <- function( resultc, resultu ) {
   if( resultc$method %in% c( "SUR", "WSUR" ) &
       resultu$method %in% c( "SUR", "WSUR" ) ) {
     n   <- resultu$eq[[1]]$n
-    lrtest$df  <- resultu$ki - resultc$ki
+    lrtest$nRestr  <- resultu$ki - resultc$ki
     if(resultc$rcovformula != resultu$rcovformula) {
       stop( paste( "both estimations must use the same formula to calculate",
                    "the residual covariance matrix!" ) )
     }
     if(resultc$rcovformula == 0) {
-      lrtest$lr  <- n * ( log( resultc$drcov ) - log( resultu$drcov ) )
+      lrtest$statistic  <- n * ( log( resultc$drcov ) - log( resultu$drcov ) )
     } else {
       residc <- array(resultc$resids,c(n,resultc$g))
       residu <- array(resultu$resids,c(n,resultu$g))
-      lrtest$lr <- n * ( log( det( (t(residc) %*% residc)) ) -
+      lrtest$statistic <- n * ( log( det( (t(residc) %*% residc)) ) -
                          log( det( (t(residu) %*% residu))))
     }
-    lrtest$p <- 1-pchisq( lrtest$lr, lrtest$df )
+    lrtest$p.value <- 1 - pchisq( lrtest$statistic, lrtest$nRestr )
   }
-  lrtest
+  class( lrtest ) <- "lrtest.systemfit"
+  return( lrtest )
 }
+
+print.lrtest.systemfit <- function( x, digits = 4, ... ){
+   cat( "\n", "Likelihood-Ratio-test for parameter restrictions",
+      " in equation systems\n", sep = "" )
+   cat( "LR-statistic:", formatC( x$statistic, digits = digits ), "\n" )
+   cat( "degrees of freedom:", x$nRestr, "\n" )
+   cat( "p-value:", formatC( x$p.value, digits = digits ), "\n\n" )
+   invisible( x )
+}
+

### 2006-02-17 paper/trunk/systemfit_other.tex
Revised/extended subsection about testing linear parameter restrictions (F-test, Wald-test, LR-test)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@606 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Likelihood ratio test}
+\subsection{Testing Linear Restrictions}
 
-Currently, the \code{systemfit} package can perform a Likelihood Ratio
-test to test linear restrictions when estimating SUR equation systems.
-The function returns the empirical likelihood ratio, the $p$-value and
-the degrees of freedom of the test which is equivalent to the number
-of restrictions.
+Linear restrictions can be tested by an F test, Wald test or
+likelihood-ratio (LR) test.
 
-% This test calculates the likelihood ratio value by calculating the
-% estimated residual covariance matrix $\hat{S}$ by the maximum
-% likelihood formula,
+The F-statistic for sytems of equations is
+\begin{equation}
+F = \frac{
+   ( R \hat{b} - q )'
+   ( R ( X' ( \hat{\Sigma} \otimes I )^{-1} X )^{-1} R' )^{-1}
+   ( R \hat{b} - q ) /
+   j
+}{
+   \hat{e}' ( \Sigma \otimes I )^{-1} \hat{e} /
+   ( M \cdot T - K )
+}
+\end{equation}
+where $j$ is the number of restrictions,
+$M$ is the number of equations,
+$T$ is the number of observations per equation,
+$K$ is the total number of estimated coefficients, and
+$\hat{\Sigma}$ is the estimated residual covariance matrix
+used in the estimation.
+Asymptotically, $F$ has an F-distribution
+with $j$ and $M \cdot T - K$ degrees of freedom
+under the null hypothesis
+\citep[p.\ 346]{greene03}.
 
-% \begin{equation}
-% \hat{s}_{ij} = \frac{\hat{e}_i^{'} \hat{e}_j}{T} 
-% \end{equation}
-% where $T$ is vector of ones.
+The Wald-statistic for sytems of equations is
+\begin{equation}
+W =
+   ( R \hat{b} - q )'
+   ( R \widehat{Cov} [ \hat{b} ] R' )^{-1}
+   ( R \hat{b} - q )
+\end{equation}
+Asymptotically, $W$ has a $\chi^2$
+distribution with $j$ degrees of freedom
+under the null hypothesis
+\citep[p.\ 347]{greene03}.
+
+The LR-statistic for sytems of equations is
+\begin{equation}
+LR = T \cdot \left(
+   log \left| \hat{ \hat{ \Sigma } }_r \right|
+   - log \left| \hat{ \hat{ \Sigma } }_u \right|
+   \right)
+\end{equation}
+where $T$ is the number of observations per equation, and 
+$\hat{\hat{\Sigma}}_r$ and $\hat{\hat{\Sigma}}_u$ are
+the residual covariance matrices calculated by formula (\ref{eq:rcov0})
+of the restricted and unrestricted estimation, respectively.
+Asymptotically, $LR$ has a $\chi^2$
+distribution with $j$ degrees of freedom
+under the null hypothesis
+\citep[p.\ 349]{greene03}.
 

### 2006-02-18 paper/trunk/systemfit_intro.tex
substituted "covariance matrix" for "variance-covariance matrix" and "\Cov [ . ]" for "Var [ . ]", where "\Cov" is a command for "\text{Cov}"
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@605 591db2bd-e0a0-4d22-8c19-2192dbb7da31
considerably revised introduction
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@609 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 
-Many theoretical models consist of more than one equation.
-These models can contain variables that appear on the left-hand side in one
-equation and on the right-hand side of another equation (simultaneity-bias).
-The system can have non-zero off-diagonal covariance elements resulting from
-related disturbances (contemporaneous correlation).
-Ignoring contemporaneous correlation leads to inefficient parameter
-estimates \citep{zellner62} and ignoring simultaneity-bias can lead to
-inconsistent parameter estimates.
-
-% If these models are econometrically estimated, it may not be
-% appropriate to estimate the equations individually, but all together,
-% because the disturbance terms of these equations are likely to be
-% contemporaneously correlated.  Estimating all equations
-% simultaneously, taking the covariance structure of the residuals into
-% account, leads to efficient estimates.  Ignoring this correlation
-% leads to an inefficient parameter estimation \citep{zellner62}.
-
+Many theoretical models that are econometrically estimated
+consist of more than one equation.
+In this case, the disturbance terms of these equations are likely
+to be contemporaneously correlated,
+because some unconsidered factors
+that influence the disturbance term in one equation
+probably influence the disturbance terms in other equations
+of these model, too.
+Ignoring this contemporaneous correlation
+and estimating these equations separately
+leads to inefficient parameter estimates.
+However, estimating all equations simultaneously,
+taking the covariance structure of the residuals into account,
+leads to efficient estimates.
+This estimation procedure is generally called
+°Seemingly Unrelated Regression° (SUR) \citep{zellner62}.
 Another reason to estimate an equation system simultaneously are
-cross-equation parameter restrictions.
+cross-equation parameter restrictions.%
+\footnote{
+Especially the economic theory suggests many cross-equation parameter
+restrictions (e.g.\ the symmetry restriction in demand models).
+}
 These restrictions can be tested and/or imposed only in a simultaneous
 estimation approach.
-Especially the economic theory suggests many cross-equation parameter
-restrictions (e.g.\ the symmetry restriction in demand models).
 
-% (But we don't say why this is important...) Are we
-% imposing restrictions in the parameters, or the variance-covariance
-% matrix or both and for what pupose?
+Furthermore, these models can contain variables
+that appear on the left-hand side in one equation
+and on the right-hand side of another equation.
+Ignoring the endogeneity of these variables can lead to inconsistent
+parameter estimates.
+This simultaneity bias can be circumvented by applying
+a °Two-Stage Least Squares° (2SLS) or °Three-Stage Least Squares° (3SLS)
+estimation of the equation system.
 
 % For all of the methods developed in the package, the disturbances of
-% the individual equations are assumed to be independent and identially
+% the individual equations are assumed to be independent and identically
 % distributed (iid).  
@@
 
-This paper is organized as follows: In Section~\ref{sec:Estimation} we
+This paper is organized as follows:
+In the following Section~\ref{sec:Estimation} we
 introduce the mathematics of estimating equation systems.
@@
 imposed.
+Some other relevant issues about estimation of equation systems
+are discussed in Section~\ref{sec:Other}.
 Section~\ref{sec:Usage} demonstrates how to run
@@
 sections can be used.
-All other relevant issues are discussed in Section~\ref{sec:Other}.
 Finally, a summary and outlook are presented in

### 2006-02-18 paper/trunk/systemfit_summary.tex
made small comment to test the repos on the new server
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@587 591db2bd-e0a0-4d22-8c19-2192dbb7da31
made another small comment to test the repos
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@588 591db2bd-e0a0-4d22-8c19-2192dbb7da31
really small change after read. it's ready for submission!
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@589 591db2bd-e0a0-4d22-8c19-2192dbb7da31
changed a few line breaks that each sentence starts in a new line now
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@

--- 2005-11-02 paper/trunk/systemfit_summary.tex
@@
 
-The \pkg{systemfit} package was originally developed to fit
-simultaneous equations for forestry datasets. Forestry datasets
-typically contain observations of many inexpensive observations (stem
-diameter) and few expensive observations such as stem height and being
-able the utilize as many observations as possible is advantageous. As
-most software packages simply drop data points that do not contain
-both observations, the \pkg{systemfit} package currently does not
-allow for unbalanced datasets, but it is our intention to include this
-capability for certain estimation methods in future releases.
-
 In this article, we have described some of the basic features of the
 \pkg{systemfit} package for estimation of linear systems of
-equations. It has been tested on a variety of datasets and has
-produced satisfactory for a few years. While more sophisticated tools
-exist, the \pkg{systemfit} package will perform the basic fitting
-methods. We wish to include more sophisticated estimation methods such
-as full-information maximum likelihood and generalized methods of
-moments.
+equations with various estimation methods.
+Many details of the estimation can be controlled by the user.
+Furthermore, the package provides some statistical tests
+for parameter restrictions and consistency of 3SLS estimation.
+It has been tested on a variety of datasets and has produced satisfactory
+for a few years.%
+\footnote{
+The code to replicate results published in the literature is given
+in the documentation of the cooresponding data sets
+included in the \pkg{systemfit} package
+(see the documentation of \code{GrunfeldGreen}, \code{GrunfeldTheil},
+\code{KleinI} and \code{kmenta}).
+}
+While the \pkg{systemfit} package performs the basic fitting methods,
+more sophisticated tools exist.
+We hope to implement missing functionalities
+in the near future.
+Some of these are discussed in the following.
+
+\paragraph{Unbalanced datasets}
+Currently, the \pkg{systemfit} package requires
+that all equations have the same number of observations.
+However, many data sets have unbalanced observations.%
+\footnote{
+For instance,
+forestry datasets typically contain observations of many inexpensive
+observations (stem diameter) and few expensive observations such
+as stem height.
+}
+Simply dropping data points
+that do not contain observations for all equations
+may considerably reduce the number of observations,
+and thus, the information utilized in the estimation.
+Hence, it is our intention to include the capability for estimations
+with unbalanced data sets as described in \citet{schmidt77}
+in future releases of \pkg{systemfit}.
+
+\paragraph{Serial correlation and heteroscedasticity}
+For all of the methods developed in the package, the disturbances of
+the individual equations are assumed to be independent and identically
+distributed (iid).
+The package could be enhanced by the inclusion of methods to fit
+equations with serially correlated and heteroscedastic disturbances.
+
+\paragraph{Estimation methods}
+We wish to include more sophisticated estimation methods such
+as limited-information maximum likelihood (LIML),
+full-information maximum likelihood (FIML),
+and generalized methods of moments (GMM), in future releases.
+
+%% testing for the new svn repos
+%% It works great!!!
 

### 2006-02-18 paper/trunk/nlsystemfit.tex
moved subsection about non-linear estimation from section "other issues" to the file "nlsystemfit.tex" that is (currently) not included in the article; added paragraph about nlsystemfit in section "summary and outlook"; removed "non-linear" from section "introduction"
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@612 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 % SUR fit and the 2SLS portion of the 3SLS fit are not included.  }
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Nonlinear estimation}
+
+As linear systems of equations can be estimated using
+\code{systemfit}, nonlinear systems of equations \citep{hausman75}
+can be estimated using a similar function; \code{nlsystemfit}.
+The usage for the nonlinear estimation function is similar to
+\code{systemfit} and can estimate system of equations using OLS, SUR,
+2SLS and 3SLS methods.
+The equations are defined as,
+
+\begin{equation}
+  \label{eq:non_linear_eq_1}
+  \epsilon_{t} = q( y_t, x_t, \beta )
+\end{equation}
+
+\noindent and
+
+\begin{equation}
+  \label{eq:non_linear_eq_2}
+  z_{t} = Z( x_t )
+\end{equation}
+where $\epsilon_{t}$ are the residuals from the $y_t$ observations,
+$x_t$ is a matrix of independent variables, $\beta$ is a vector is
+parameter estimates and $z_{t}$ are the functions evaluated at the
+parameter estimates.
+
+Similar to calling the \code{systemfit} function, \code{nlsystemfit}
+is called with a minimum of three arguments,
+
+\code{
+R> nlsystemfit( method, eqns, start )
+}
+
+where \code{method} is one of the following estimation methods: "OLS",
+"SUR", "2SLS", or "3SLS", \code{eqns} is a list of equations similar
+to those described for \code{systemfit}, and \code{start} is a list of
+starting values of the parameter estimates.
+The nlsystemfit function relies on \code{nlm} to perform the minimization
+of the objective functions and the \code{qr} set of functions.
+If the user does not have a set of estimates for the initial parameters,
+it is suggested using linearized forms in one of the linear methods,
+or simply using \code{nls} to obtain estimates for the equations.
+The outputs are similar to those obtained by \code{systemfit} objects.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% left over from the older docs
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% The mandatory argument
+% \code{eqns} is a list of the equations to estimate.  Each equation is
+% a standard formula in \proglang{R}.  It starts with a dependent
+% variable on the left hand side.  After a tilde ($~$) the regressors
+% are listed  \footnote{For Details see the \proglang{R} help files to
+%   \code{formula}}.
+
+% The objective functions for the methods are:
+
+% \begin{center}
+% \begin{tabular}{|l|c|c|c|} \hline
+%   Method & Instruments & Objective Function & Covariance of $\theta$ \\ \hline  
+%   OLS & No & $r'r$ & $(X(diag(S)^{-1}\otimes I)X)^{-1}$ \\ \hline
+%   SUR & No & $r'(diag(S)_{OLS}^{-1}\otimes I)r$ & $(X(S^{-1}\otimes I)X)^{-1}$ \\ \hline
+%   2SLS & Yes & $r'(I \otimes W)r$ & $(X(diag(S)^{-1}\otimes I)X)^{-1}$ \\ \hline 
+%   3SLS & Yes & $r'(S_{2SLS}^{-1} \otimes W)r$ & $(X(diag(S)^{-1}\otimes W)X)^{-1}$ \\ \hline
+% \end{tabular}
+% \end{center}
+
+% where, $r$ is a column vector for the residuals for each equation,
+% what is the difference to \epsilon?
+% $X$ is matrix of the partial derivatives of the dependent variable 
+% with respect to the parameters $\left( \frac{ \partial y }{ \theta} \right)$,
+% is this correct?
+% in the linear section X is used for the regressors, 
+% thus we should use a different variable name here.
+% $W$ is a matrix of the instrument variables, $Z(Z'Z)^{-1}Z$, $Z$ is a
+% matrix of the instrument variables, and $I$ is an $n \times n $
+% identity matrix and $S$ is the estimated covariance matrix between the
+% equations
+
+% \begin{equation}
+%   \label{eq:non-linear_varcov}
+%   \hat{\sigma}_{ij} = (\hat{e}_i' \hat{e}_j) / \sqrt{(T - k_i)*(T- k_j)} 
+% \end{equation}
+
+% Do we actually have the covariance function in place.
+
+% The argument \code{method} is a string determining the estimation method.
+% It must be one of "OLS", "WLS", "SUR", "2SLS", "W2SLS" or "3SLS".
+
+% \textbf{The residual covariance matrix can be calculated in
+%   different ways (section~\ref{sec:residcov}).  It should be
+%   relatively easy to implement this also in nlsystemfit(),  e.g. the
+%   function 'calcRCov' in systemfit() could be moved outside
+%   systemfit() that it can be used also by nlsystemfit().}
+
+% \textbf{You need to clear this up.}

### 2006-02-18 paper/trunk/systemfit_intro.tex
moved subsection about non-linear estimation from section "other issues" to the file "nlsystemfit.tex" that is (currently) not included in the article; added paragraph about nlsystemfit in section "summary and outlook"; removed "non-linear" from section "introduction"
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@612 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 The \pkg{systemfit} package provides the capability to estimate
-linear and non-linear equation systems in \proglang{R}
+linear equation systems in \proglang{R}
 \citep{r-project}.
-Although linear and non-linear equation systems can be estimated
+Although linear equation systems can be estimated
 with several other statistical and econometric software packages

### 2006-02-18 paper/trunk/systemfit_other.tex
added section about how to use systemfit to test linear restrictions; changed restriction: beta_2 = -beta_5 -> beta_2 = -beta_6; a very few other corrections
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@607 591db2bd-e0a0-4d22-8c19-2192dbb7da31
moved subsection about non-linear estimation from section "other issues" to the file "nlsystemfit.tex" that is (currently) not included in the article; added paragraph about nlsystemfit in section "summary and outlook"; removed "non-linear" from section "introduction"
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@612 591d

@@
 \subsection{Testing Linear Restrictions}
+\label{sec:testingRestrictions}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Nonlinear estimation}
-
-As linear systems of equations can be estimated using
-\code{systemfit}, nonlinear systems of equations \citep{hausman75}
-can be estimated using a similar function; \code{nlsystemfit}.
-The usage for the nonlinear estimation function is similar to
-\code{systemfit} and can estimate system of equations using OLS, SUR,
-2SLS and 3SLS methods.
-The equations are defined as,
-
-\begin{equation}
-  \label{eq:non_linear_eq_1}
-  \epsilon_{t} = q( y_t, x_t, \beta )
-\end{equation}
-
-\noindent and
-
-\begin{equation}
-  \label{eq:non_linear_eq_2}
-  z_{t} = Z( x_t )
-\end{equation}
-where $\epsilon_{t}$ are the residuals from the $y_t$ observations,
-$x_t$ is a matrix of independent variables, $\beta$ is a vector is
-parameter estimates and $z_{t}$ are the functions evaluated at the
-parameter estimates.
-
-Similar to calling the \code{systemfit} function, \code{nlsystemfit}
-is called with a minimum of three arguments,
-
-\code{
-R> nlsystemfit( method, eqns, start )
-}
-
-where \code{method} is one of the following estimation methods: "OLS",
-"SUR", "2SLS", or "3SLS", \code{eqns} is a list of equations similar
-to those described for \code{systemfit}, and \code{start} is a list of
-starting values of the parameter estimates.
-The nlsystemfit function relies on \code{nlm} to perform the minimization
-of the objective functions and the \code{qr} set of functions.
-If the user does not have a set of estimates for the initial parameters,
-it is suggested using linearized forms in one of the linear methods,
-or simply using \code{nls} to obtain estimates for the equations.
-The outputs are similar to those obtained by \code{systemfit} objects.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% left over from the older docs
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% The mandatory argument
-% \code{eqns} is a list of the equations to estimate.  Each equation is
-% a standard formula in \proglang{R}.  It starts with a dependent
-% variable on the left hand side.  After a tilde ($~$) the regressors
-% are listed  \footnote{For Details see the \proglang{R} help files to
-%   \code{formula}}.
-
-% The objective functions for the methods are:
-
-% \begin{center}
-% \begin{tabular}{|l|c|c|c|} \hline
-%   Method & Instruments & Objective Function & Covariance of $\theta$ \\ \hline  
-%   OLS & No & $r'r$ & $(X(diag(S)^{-1}\otimes I)X)^{-1}$ \\ \hline
-%   SUR & No & $r'(diag(S)_{OLS}^{-1}\otimes I)r$ & $(X(S^{-1}\otimes I)X)^{-1}$ \\ \hline
-%   2SLS & Yes & $r'(I \otimes W)r$ & $(X(diag(S)^{-1}\otimes I)X)^{-1}$ \\ \hline 
-%   3SLS & Yes & $r'(S_{2SLS}^{-1} \otimes W)r$ & $(X(diag(S)^{-1}\otimes W)X)^{-1}$ \\ \hline
-% \end{tabular}
-% \end{center}
-
-% where, $r$ is a column vector for the residuals for each equation,
-% what is the difference to \epsilon?
-% $X$ is matrix of the partial derivatives of the dependent variable 
-% with respect to the parameters $\left( \frac{ \partial y }{ \theta} \right)$,
-% is this correct?
-% in the linear section X is used for the regressors, 
-% thus we should use a different variable name here.
-% $W$ is a matrix of the instrument variables, $Z(Z'Z)^{-1}Z$, $Z$ is a
-% matrix of the instrument variables, and $I$ is an $n \times n $
-% identity matrix and $S$ is the estimated covariance matrix between the
-% equations
-
-% \begin{equation}
-%   \label{eq:non-linear_varcov}
-%   \hat{\sigma}_{ij} = (\hat{e}_i' \hat{e}_j) / \sqrt{(T - k_i)*(T- k_j)} 
-% \end{equation}
-
-% Do we actually have the covariance function in place.
-
-% The argument \code{method} is a string determining the estimation method.
-% It must be one of "OLS", "WLS", "SUR", "2SLS", "W2SLS" or "3SLS".
-
-% \textbf{The residual covariance matrix can be calculated in
-%   different ways (section~\ref{sec:residcov}).  It should be
-%   relatively easy to implement this also in nlsystemfit(),  e.g. the
-%   function 'calcRCov' in systemfit() could be moved outside
-%   systemfit() that it can be used also by nlsystemfit().}
-
-% \textbf{You need to clear this up.}
-
 %%% Local Variables: 

### 2006-02-18 paper/trunk/systemfit_summary.tex
moved subsection about non-linear estimation from section "other issues" to the file "nlsystemfit.tex" that is (currently) not included in the article; added paragraph about nlsystemfit in section "summary and outlook"; removed "non-linear" from section "introduction"
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@612 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 
+\paragraph{Non-linear estimation}
+The \pkg{systemfit} package provides a function to estimate
+systems of non-linear estimations.
+However, the function \code{nlsystemfit} is currently under development
+and the results are not yet always reliable
+due to convergence problems.
+We hope to improve this behavior soon.
+
 %% testing for the new svn repos

### 2006-02-19 paper/trunk/systemfit_estimation.tex
several minor revisions (mainly suggested by Geraldine)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@613 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 \end{equation}
-where $\Sigma = \left[ \sigma_{ij} \right]$ and $I$ is an
-identity matrix.
+where $\Sigma = \left[ \sigma_{ij} \right]$ is the residual covariance matrix
+and $I$ is an identity matrix.
 
@@
 \begin{equation}
-   \widehat{\beta} = \left( X'X \right)^{-1} X'y
+   \widehat{\beta}_{OLS} = \left( X'X \right)^{-1} X'y
+   \label{eq:ols}
 \end{equation}
@@
 \begin{equation}
-   \Cov \left[ \widehat{\beta} \right] = \sigma^2 \left( X'X \right)^{-1}
+   \Cov \left[ \widehat{\beta}_{OLS} \right] = \sigma^2 \left( X'X \right)^{-1}
    \label{eq:olsCovSameSigma}
@@
 \begin{equation}
-   \Cov \left[ \widehat{\beta} \right] = \left( X' \Omega^{-1} X \right)^{-1}
+   \Cov \left[ \widehat{\beta}_{OLS} \right] = \left( X' \Omega^{-1} X \right)^{-1}
    \label{eq:olsCovSingleSigma}
@@
 \begin{equation}
-   \widehat{\beta} = \left( X' \Omega^{-1} X \right)^{-1} X' \Omega^{-1} y
+   \widehat{\beta}_{WLS} = \left( X' \Omega^{-1} X \right)^{-1} X' \Omega^{-1} y
 \end{equation}
@@
 \begin{equation}
-   \Cov \left[ \widehat{\beta} \right] = \left( X' \Omega^{-1} X \right)^{-1}
+   \Cov \left[ \widehat{\beta}_{WLS} \right] = \left( X' \Omega^{-1} X \right)^{-1}
 \end{equation}
@@
 \begin{equation}
-   \widehat{\beta} = \left( X' \Omega^{-1} X \right)^{-1} X' \Omega^{-1} y
+   \widehat{\beta}_{SUR} = \left( X' \Omega^{-1} X \right)^{-1} X' \Omega^{-1} y
 \end{equation}
@@
 \begin{equation}
-   \Cov \left[ \widehat{\beta} \right] = \left( X' \Omega^{-1} X \right)^{-1}
+   \Cov \left[ \widehat{\beta}_{SUR} \right] = \left( X' \Omega^{-1} X \right)^{-1}
 \end{equation}
@@
 the estimated coefficients are biased.
-This can be circumvented by a two-stage least squares (2SLS) estimation.
+This can be circumvented by an instrumental variable (IV)
+two-stage least squares (2SLS) estimation.
 The instrumental variables for each equation $H_i$ 
@@
 \begin{equation}
-   \widehat{\beta} = \left( \widehat{X}' \widehat{X} \right)^{-1} 
+   \widehat{\beta}_{2SLS} = \left( \widehat{X}' \widehat{X} \right)^{-1} 
    \widehat{X}' y 
@@
 \begin{equation}
-   \Cov \left[ \widehat{\beta} \right] = \sigma^2 \left( \widehat{X}'
+   \Cov \left[ \widehat{\beta}_{2SLS} \right] = \sigma^2 \left( \widehat{X}'
    \widehat{X} \right)^{-1}
@@
 \begin{equation}
-   \Cov \left[ \widehat{\beta} \right] = \left( \widehat{X}' \Omega^{-1}
+   \Cov \left[ \widehat{\beta}_{2SLS} \right] = \left( \widehat{X}' \Omega^{-1}
    \widehat{X} \right)^{-1}
@@
 \begin{equation}
-   \widehat{\beta} = \left( \widehat{X}' \Omega^{-1} \widehat{X} 
+   \widehat{\beta}_{W2SLS} = \left( \widehat{X}' \Omega^{-1} \widehat{X} 
    \right)^{-1} \widehat{X}' \Omega^{-1} y
@@
 \begin{equation}
-   \Cov \left[ \widehat{\beta} \right] = \left( \widehat{X}' \Omega^{-1}
+   \Cov \left[ \widehat{\beta}_{W2SLS} \right] = \left( \widehat{X}' \Omega^{-1}
    \widehat{X} \right)^{-1}
@@
 \begin{equation}
-   \widehat{\beta} = \left( \widehat{X}' \Omega^{-1} \widehat{X} 
+   \widehat{\beta}_{3SLS} = \left( \widehat{X}' \Omega^{-1} \widehat{X} 
    \right)^{-1} \widehat{X}' \Omega^{-1} y
@@
 \begin{equation}
-   \Cov \left[ \widehat{\beta} \right] = \left( \widehat{X}' \Omega^{-1}
+   \Cov \left[ \widehat{\beta}_{3SLS} \right] = \left( \widehat{X}' \Omega^{-1}
    \widehat{X} \right)^{-1}
@@
 One of these approaches is based on instrumental variable estimation
-(IV-3SLS):
+(3SLS-IV):
 \begin{equation}
-   \widehat{\beta} = \left( \widehat{X}' \Omega^{-1} X 
+   \widehat{\beta}_{3SLS-IV} = \left( \widehat{X}' \Omega^{-1} X
    \right)^{-1} \widehat{X}' \Omega^{-1} y
@@
 \end{equation}
-The covariance matrix of this IV-3SLS estimator is:
+The covariance matrix of this 3SLS-IV estimator is:
 \begin{equation}
-   \Cov \left[ \widehat{\beta} \right] = \left( \widehat{X}' \Omega^{-1}
+   \Cov \left[ \widehat{\beta}_{3SLS-IV} \right] = \left( \widehat{X}' \Omega^{-1}
    X \right)^{-1}
@@
 Another approach is based on the Generalized Method of Moments (GMM)
-estimator (GMM-3SLS):
+estimator (3SLS-GMM):
 \begin{equation}
-   \widehat{\beta} = \left( X' H \left( H' \Omega H \right)^{-1}
+   \widehat{\beta}_{3SLS-GMM} = \left( X' H \left( H' \Omega H \right)^{-1}
    H' X \right)^{-1} X' H \left( H' \Omega H \right)^{-1} H' y
@@
 \end{equation}
-The covariance matrix of the GMM-3SLS estimator is:
+The covariance matrix of the 3SLS-GMM estimator is:
 \begin{equation}
-   \Cov \left[ \widehat{\beta} \right] =
+   \Cov \left[ \widehat{\beta}_{3SLS-GMM} \right] =
    \left( X' H \left( H' \Omega H \right)^{-1} H' X \right)^{-1}
@@
 \begin{equation}
-   \widehat{\beta} = \left( \widehat{X}' \Omega^{-1} \widehat{X} 
+   \widehat{\beta}_{3SLS-Schmidt} = \left( \widehat{X}' \Omega^{-1} \widehat{X}
    \right)^{-1} \widehat{X}' \Omega^{-1} 
@@
 \begin{equation}
-   \Cov \left[ \widehat{\beta} \right] =
+   \Cov \left[ \widehat{\beta}_{3SLS-Schmidt} \right] =
    \left( \widehat{X}' \Omega^{-1}  \widehat{X} \right)^{-1} 
@@
 \begin{equation}
-   \widehat{\beta} = \widehat{\beta}_{2SLS} + 
+   \widehat{\beta}_{3SLS-EViews} = \widehat{\beta}_{2SLS} + 
    \left( \widehat{X}' \Omega^{-1} \widehat{X} \right)^{-1} 
@@
 \end{equation}
-with $\widehat{\beta}_{2SLS}$ is the two-stage least squares estimator
-(equation \ref{eq:beta2sls}).
+where $\widehat{\beta}_{2SLS}$ is the two-stage least squares estimator
+as defined by (\ref{eq:beta2sls}).
 EViews uses the standard 3SLS formula (\ref{eq:cov3sls}) to
@@
 However, if this is not the case, the results depend on the 
-equation used \citep{schmidt90}. 
-The only reason to use different instruments in the different equations 
-is that the instruments of one equation are correlated with the
-disturbance terms in another equation.
-Otherwise, one would simply use every instrument in every equation
+method used \citep{schmidt90}.
+The only reason to use different instruments for different equations
+is a correlation of the instruments of one equation with the
+disturbance terms of another equation.
+Otherwise, one could simply use all instruments in every equation
 \citep{schmidt90}.
-In this case, only the GMM-3SLS (\ref{eq:3slsGmm})
+In this case, only the 3SLS-GMM (\ref{eq:3slsGmm})
 and the 3SLS estimator developed by \cite{schmidt90} 

### 2006-02-19 paper/trunk/systemfit_other.tex
several minor revisions (mainly suggested by Geraldine)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@613 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 where $T$ is the number of observations in each equation.
-However, this estimator is biased, because it is not corrected for
-degrees of freedom.
+However, in finite samples this estimator is biased,
+because it is not corrected for degrees of freedom.
 The usual single-equation procedure to correct for degrees of freedom
@@
 }
-It uses the residuals of a first-step OLS or 2SLS estimation
+This alternative approach uses the residuals of a first-step OLS or 2SLS estimation
 to apply a WLS or W2SLS estimation on a second step.
@@
 the parameter estimates of OLS and WLS as well as 2SLS and W2SLS are identical.
-Hence, in this case the both approach generate the same results.
+Hence, in this case both approaches generate the same results.
 
@@
 discussed in the literature.
-While sometimes the degrees of freedom of the whole system
+While sometimes the degrees of freedom of the entire system
 (total number of observations in all equations minus
@@
 of each equation is rather small, and
-therefore, this makes a difference.
+therefore, it matters.
 
-If a system of equations is estimated by unrestricted OLS and
+If a system of equations is estimated by an unrestricted OLS and
 the covariance matrix of the parameters is calculated
@@
 each single equation,
-because this yields also the same p-values as the equation-wise
+because this yields the same p-values as the equation-wise
 OLS estimation.
@@
 
-The F-statistic for sytems of equations is
+The F-statistic for systems of equations is
 \begin{equation}
 F = \frac{
-   ( R \hat{b} - q )'
+   ( R \hat{\beta} - q )'
    ( R ( X' ( \hat{\Sigma} \otimes I )^{-1} X )^{-1} R' )^{-1}
-   ( R \hat{b} - q ) /
+   ( R \hat{\beta} - q ) /
    j
 }{
-   \hat{e}' ( \Sigma \otimes I )^{-1} \hat{e} /
+   \hat{u}' ( \Sigma \otimes I )^{-1} \hat{u} /
    ( M \cdot T - K )
@@
 used in the estimation.
-Asymptotically, $F$ has an F-distribution
+Under the null hypothesis, $F$ has an F-distribution
 with $j$ and $M \cdot T - K$ degrees of freedom
-under the null hypothesis
 \citep[p.\ 346]{greene03}.
 
-The Wald-statistic for sytems of equations is
+The Wald-statistic for systems of equations is
 \begin{equation}
 W =
-   ( R \hat{b} - q )'
-   ( R \widehat{Cov} [ \hat{b} ] R' )^{-1}
-   ( R \hat{b} - q )
+   ( R \hat{\beta} - q )'
+   ( R \widehat{Cov} [ \hat{\beta} ] R' )^{-1}
+   ( R \hat{\beta} - q )
 \end{equation}
@@
 
-The LR-statistic for sytems of equations is
+The LR-statistic for systems of equations is
 \begin{equation}

### 2006-02-19 paper/trunk/systemfit_restrictions.tex
several minor revisions (mainly suggested by Geraldine)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@613 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 There are two ways to impose linear parameter restrictions.
-First, a matrix $T$ can be provided with
+First, a matrix $T$ can be specified that
 \begin{equation}
@@
 \end{equation}
-Then, a standard estimation is done by substituting $X^*$ for $X$.
+Then, $X^*$ is substituted for $X$ and a standard estimation as described
+in the previous section is done
+(equations~\ref{eq:ols}--\ref{eq:3slsEViews}).
 This results in the linear independent parameter estimates $\beta^*$ and
-the covariance matrix of these parameters.
+their covariance matrix.
 The original parameters can be obtained by equation (\ref{eq:T-restr})
@@
 \begin{equation}
-   \Cov \left[ \widehat{\beta} \right] = T \cdot \Cov \left[ \widehat{\beta^*} \right] \cdot T'
+   \Cov \left[ \widehat{\beta} \right] = T \cdot \Cov \left[ \widehat{\beta}^* \right] \cdot T'
 \end{equation}
+
 The second way to impose linear parameter restrictions
@@
 where $\beta^0$ is the vector of the restricted coefficients, 
-$R$ is a matrix and $q$ is a vector (see \citealp[p.\ 100]{greene03}).
-Each restriction belongs to one row of $R$ and the corresponding 
-element of~$q$.
+and $R$ and $q$ are a matrix and vector, respectively,
+to impose the restrictions \citep[see][p.\ 100]{greene03}.
+Each linear independent restriction is represented by one row of $R$
+and the corresponding element of~$q$.
 
@@
    \left[ \begin{array}{c}
-      \widehat{\beta^0} \\ \widehat{\lambda}
+      \widehat{\beta}^0_{OLS} \\ \widehat{\lambda}
    \end{array} \right]
@@
    \left[ \begin{array}{c}
-      \widehat{\beta} \\ \widehat{\lambda}
+      \widehat{\beta}^0_{OLS} \\ \widehat{\lambda}
    \end{array} \right] 
@@
    \left[ \begin{array}{c}
-      \widehat{\beta^0} \\ \widehat{\lambda}
+      \widehat{\beta}^0_{WLS} \\ \widehat{\lambda}
    \end{array} \right]
@@
    \left[ \begin{array}{c}
-      \widehat{\beta} \\ \widehat{\lambda}
+      \widehat{\beta}^0_{WLS} \\ \widehat{\lambda}
    \end{array} \right] 
@@
    \left[ \begin{array}{c}
-      \widehat{\beta^0} \\ \widehat{\lambda}
+      \widehat{\beta}^0_{SUR} \\ \widehat{\lambda}
    \end{array} \right]
@@
    \left[ \begin{array}{c}
-      \widehat{\beta} \\ \widehat{\lambda}
+      \widehat{\beta}^0_{SUR} \\ \widehat{\lambda}
    \end{array} \right] 
@@
    \left[ \begin{array}{c}
-      \widehat{\beta^0} \\ \widehat{\lambda}
+      \widehat{\beta}^0_{2SLS} \\ \widehat{\lambda}
    \end{array} \right]
@@
    \left[ \begin{array}{c}
-      \widehat{\beta} \\ \widehat{\lambda}
+      \widehat{\beta}^0_{2SLS} \\ \widehat{\lambda}
    \end{array} \right] 
@@
    \left[ \begin{array}{c}
-      \widehat{\beta^0} \\ \widehat{\lambda}
+      \widehat{\beta}^0_{W2SLS} \\ \widehat{\lambda}
    \end{array} \right]
@@
    \left[ \begin{array}{c}
-      \widehat{\beta} \\ \widehat{\lambda}
+      \widehat{\beta}^0_{W2SLS} \\ \widehat{\lambda}
    \end{array} \right] 
@@
    \left[ \begin{array}{c}
-      \widehat{\beta^0} \\ \widehat{\lambda}
+      \widehat{\beta}^0_{3SLS} \\ \widehat{\lambda}
    \end{array} \right]
@@
    \left[ \begin{array}{c}
-      \widehat{\beta^0} \\ \widehat{\lambda}
+      \widehat{\beta}^0_{3SLS} \\ \widehat{\lambda}
    \end{array} \right] 
@@
 \end{equation}
-The IV-3SLS estimator restricted by $R \beta^0 = q$ can be obtained by
+The 3SLS-IV estimator restricted by $R \beta^0 = q$ can be obtained by
 \begin{equation}
    \left[ \begin{array}{c}
-      \widehat{\beta^0} \\ \widehat{\lambda}
+      \widehat{\beta}^0_{3SLS-IV} \\ \widehat{\lambda}
    \end{array} \right]
@@
    \left[ \begin{array}{c}
-      \widehat{\beta^0} \\ \widehat{\lambda}
+      \widehat{\beta}^0_{3SLS-IV} \\ \widehat{\lambda}
    \end{array} \right] 
@@
    \left[ \begin{array}{c}
-      \widehat{\beta^0} \\ \widehat{\lambda}
+      \widehat{\beta}^0_{3SLS-GMM} \\ \widehat{\lambda}
    \end{array} \right]
@@
    \left[ \begin{array}{c}
-      \widehat{\beta^0} \\ \widehat{\lambda}
+      \widehat{\beta}^0_{3SLS-GMM} \\ \widehat{\lambda}
    \end{array} \right] 
@@
    \left[ \begin{array}{c}
-      \widehat{\beta^0} \\ \widehat{\lambda}
+      \widehat{\beta}^0_{3SLS-Schmidt} \\ \widehat{\lambda}
    \end{array} \right]
@@
    \left[ \begin{array}{c}
-      \widehat{\beta^0} \\ \widehat{\lambda}
+      \widehat{\beta}^0_{3SLS-Schmidt} \\ \widehat{\lambda}
    \end{array} \right] 
@@
    \end{array} \right]^{-1}
-   \cdot
+   \\
+   & & \cdot
    \left[ \begin{array}{cc}
@@
    \end{array} \right]^{-1}
+   \nonumber
 \end{eqnarray}
@@
    \left[ \begin{array}{c}
-      \widehat{\beta^0} \\ \widehat{\lambda}
+      \widehat{\beta}^0_{3SLS-EViews} \\ \widehat{\lambda}
    \end{array} \right]
@@
    \left[ \begin{array}{c}
-      \widehat{X}' \Omega^{-1} \left( y - X \widehat{\beta^0}_{2SLS} \right) 
+      \widehat{X}' \Omega^{-1} \left( y - X \widehat{\beta}^0_{2SLS} \right)
       \\ q 
@@
 \end{equation}
-where $\widehat{\beta^0}_{2SLS}$ is the restricted 2SLS estimator calculated
+where $\widehat{\beta}^0_{2SLS}$ is the restricted 2SLS estimator calculated
 by equation (\ref{eq:beta2SLSr}). 

### 2006-02-19 paper/trunk/systemfit_estimation.tex
rorganized the sections and subsections: there now a section called "Statistical Background" that includes the subsections (former sections) "Estimation", "Restrictions" and "Other issues" (a section of the "statistical background" is suggested by the JSS
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@614 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Estimation}\label{sec:Estimation}
+\subsection{Estimation}\label{sec:Estimation}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-Consider a system of $G$ equations, where the $i$th equation is of
-the form 
-\begin{equation}
-   y_{i} = X_i \beta_i + u_i, \quad i = 1, 2, \ldots, G
-\end{equation}
-where $y_i$ is a vector of the dependent variable,
-$X_i$ is a matrix of the exogenous variables,
-$\beta_i$ is the coefficient vector and
-$u_i$ is a vector of the disturbance terms of the $i$th equation.
-
-We can write the °stacked° system as
-\begin{equation}
-   \left[ \begin{array}{c}
-      y_1 \\ y_2\\ \vdots\\ y_G
-   \end{array} \right] = 
-   \left[ \begin{array}{cccc}
-      X_1 & 0 & \cdots & 0\\
-      0 & X_2 & \cdots & 0\\
-      \vdots & \vdots & \ddots & \vdots\\
-      0 & 0 & \cdots & X_G
-   \end{array}\right] \cdot 
-   \left[ \begin{array}{c}
-      \beta_1 \\ \beta_2 \\ \vdots\\ \beta_G
-   \end{array} \right] +
-   \left[ \begin{array}{c}
-      u_1 \\ u_2 \\ \vdots\\ u_G
-   \end{array} \right]
-\end{equation}
-
-or more simply as
-\begin{equation}
-   y = X \beta + u 
-\end{equation}
-
-We assume that there is no correlation of the disturbance terms 
-across observations:
-\begin{equation}
-   E \left( u_{it} \, u_{jt^*} \right) = 0 
-   \; \forall \; t \neq t^*
-\end{equation}
-where $i$ and $j$ indicate the equation number 
-and $t$ and $t^*$ denote the observation number.
-
-However, we explicitly allow for contemporaneous correlation:
-\begin{equation}
-   E \left( u_{it} \, u_{jt} \right) = \sigma_{ij}
-\end{equation}
-
-Thus, the covariance matrix of the total system is
-\begin{equation}
-   E \left( u \, u' \right) = \Omega = \Sigma \otimes I
-\end{equation}
-where $\Sigma = \left[ \sigma_{ij} \right]$ is the residual covariance matrix
-and $I$ is an identity matrix.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Ordinary least squares (OLS)}
-\label{sec:ols}
+\subsubsection{Ordinary least squares (OLS)}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Weighted least squares (WLS)}
+\subsubsection{Weighted least squares (WLS)}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Seemingly unrelated regression (SUR)}
+\subsubsection{Seemingly unrelated regression (SUR)}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Two-Stage least squares (2SLS)}
+\subsubsection{Two-Stage least squares (2SLS)}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Weighted two-stage least squares (W2SLS)}
+\subsubsection{Weighted two-stage least squares (W2SLS)}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Three-Stage least squares (3SLS)}
-\label{sec:3sls}
+\subsubsection{Three-Stage least squares (3SLS)}
 

### 2006-02-19 paper/trunk/systemfit_intro.tex
rorganized the sections and subsections: there now a section called "Statistical Background" that includes the subsections (former sections) "Estimation", "Restrictions" and "Other issues" (a section of the "statistical background" is suggested by the JSS
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@614 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 This paper is organized as follows:
-In the following Section~\ref{sec:Estimation} we
-introduce the mathematics of estimating equation systems.
-Section~\ref{sec:Restrictions} shows how linear restrictions can be
-imposed.
-Some other relevant issues about estimation of equation systems
-are discussed in Section~\ref{sec:Other}.
+In the following Section~\ref{sec:statistics} we
+introduce the statistical background of estimating equation systems.
 Section~\ref{sec:Usage} demonstrates how to run
-\pkg{systemfit}, especially how the features presented in the previous
-sections can be used.
+\pkg{systemfit} and especially how the features presented in the previous
+section can be used.
 Finally, a summary and outlook are presented in

### 2006-02-19 paper/trunk/systemfit_statistics.tex
rorganized the sections and subsections: there now a section called "Statistical Background" that includes the subsections (former sections) "Estimation", "Restrictions" and "Other issues" (a section of the "statistical background" is suggested by the JSS
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@614 591db2bd-e0a0-4d22-8c19-2192dbb7da31

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Statistical Background}\label{sec:statistics}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In this section we provide the statistical background
of the functionality provided in the \pkg{systemfit} package.
After introducing notations and assumptions,
we provide the formulas to estimate linear equation systems.
Then, we shows how linear restrictions can be imposed.
Finally, some other relevant issues about estimation of equation systems
are discussed.

Consider a system of $G$ equations, where the $i$th equation is of
the form 
\begin{equation}
   y_{i} = X_i \beta_i + u_i, \quad i = 1, 2, \ldots, G
\end{equation}
where $y_i$ is a vector of the dependent variable,
$X_i$ is a matrix of the exogenous variables,
$\beta_i$ is the coefficient vector and
$u_i$ is a vector of the disturbance terms of the $i$th equation.

We can write the °stacked° system as
\begin{equation}
   \left[ \begin{array}{c}
      y_1 \\ y_2\\ \vdots\\ y_G
   \end{array} \right] = 
   \left[ \begin{array}{cccc}
      X_1 & 0 & \cdots & 0\\
      0 & X_2 & \cdots & 0\\
      \vdots & \vdots & \ddots & \vdots\\
      0 & 0 & \cdots & X_G
   \end{array}\right] \cdot 
   \left[ \begin{array}{c}
      \beta_1 \\ \beta_2 \\ \vdots\\ \beta_G
   \end{array} \right] +
   \left[ \begin{array}{c}
      u_1 \\ u_2 \\ \vdots\\ u_G
   \end{array} \right]
\end{equation}

or more simply as
\begin{equation}
   y = X \beta + u 
\end{equation}

We assume that there is no correlation of the disturbance terms 
across observations:
\begin{equation}
   E \left( u_{it} \, u_{jt^*} \right) = 0 
   \; \forall \; t \neq t^*
\end{equation}
where $i$ and $j$ indicate the equation number 
and $t$ and $t^*$ denote the observation number.

However, we explicitly allow for contemporaneous correlation:
\begin{equation}
   E \left( u_{it} \, u_{jt} \right) = \sigma_{ij}
\end{equation}

Thus, the covariance matrix of the total system is
\begin{equation}
   E \left( u \, u' \right) = \Omega = \Sigma \otimes I
\end{equation}
where $\Sigma = \left[ \sigma_{ij} \right]$ is the residual covariance matrix
and $I$ is an identity matrix.

\input{systemfit_estimation.tex}
\input{systemfit_restrictions.tex}
\input{systemfit_other.tex}

### 2006-02-19 paper/trunk/systemfit_summary.tex
rorganized the sections and subsections: there now a section called "Statistical Background" that includes the subsections (former sections) "Estimation", "Restrictions" and "Other issues" (a section of the "statistical background" is suggested by the JSS
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@614 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 
-\paragraph{Unbalanced datasets}
+\subsubsection*{Unbalanced datasets}
 Currently, the \pkg{systemfit} package requires
@@
 
-\paragraph{Serial correlation and heteroscedasticity}
+\subsubsection*{Serial correlation and heteroscedasticity}
 For all of the methods developed in the package, the disturbances of
@@
 
-\paragraph{Estimation methods}
+\subsubsection*{Estimation methods}
 We wish to include more sophisticated estimation methods such
@@
 
-\paragraph{Non-linear estimation}
+\subsubsection*{Non-linear estimation}
 The \pkg{systemfit} package provides a function to estimate

### 2006-02-19 paper/trunk/systemfit.tex
considerably revised and extended "Usage" section
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@594 591db2bd-e0a0-4d22-8c19-2192dbb7da31
changed position of sections "Usaging Systemfit" and "Other Issues"
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@596 591db2bd-e0a0-4d22-8c19-2192dbb7da31
substituted "covariance matrix" for "variance-covariance matrix" and "\Cov [ . ]" for "Var [ . ]", where "\Cov" is a command for "\text{Cov}"
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@605 591db2bd-e0a0-4d22-8c19-2192dbb7da31
added package "thumbpdf" to cre

@@
 \documentclass[article]{jss}
+\usepackage[pdftex]{thumbpdf}
+%   run: pdflatex *.tex -> thumbpdf *.tex -> pdflatex *.tex
 \usepackage{csquotes}
 \MakeOuterQuote{°}
+\usepackage{amsmath}
+\usepackage{float}
 
@@
 \linenumbers
+
+\newcommand{\Cov}{\text{Cov}}
+\clubpenalty=10000
+\widowpenalty=10000
 
@@
 \Abstract{ 
+Many statistical analyses are based on models containing systems of
+structurally related equations.
+Since in most cases cross-equation disturbances are correlated,
+full information methods are required \citep{zellner62}.
+If exogenous variables are stochastically dependent
+on the disturbances in the system,
+instrumental variable estimations should be used for parameter estimation
+\citep{zellner62b}
+The package \pkg{systemfit} provides the capability
+to estimate systems of linear equations
+within the \proglang{R} programming environment.
+}
 
-  Many statistical analyses are based on models containing systems of
-  structurally related equations. If the cross-equation disturbances
-  are correlated, full information methods are required
-  \citep{zellner62}. If exogenous variables are stochastically
-  dependent on the disturbances in the system
-  \citep{zellner62b,theil71}, limited information methods should be
-  used for parameter estimation. The package \pkg{systemfit} provides
-  the capability to estimate systems of linear and non-linear
-  structural equations for limited (ordinary least squares (OLS), two
-  stage-least squares (2SLS)) and full-information
-  (seemingly-unrelated regression (SUR) and three-stage least squares
-  (3SLS)) methods within the \proglang{R} programming environment.}
-
-\Keywords{R, simultaneous equations estimation, ordinary least
-  squares, seemingly unrelated regression, two-stage least squares,
+\Keywords{R, simultaneous equations systems,
+  seemingly unrelated regression, two-stage least squares,
   three-stage least squares}
@@
 \input{systemfit_intro.tex}
-\input{systemfit_estimation.tex}
-\input{systemfit_restrictions.tex}
+\input{systemfit_statistics.tex}
 \input{systemfit_usage.tex}
-\input{systemfit_other.tex}
 \input{systemfit_summary.tex}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\bibliography{systemfit} % a subset of my big bibtex file
-%\bibliography{/home/suapm095/Documents/Literatur/arne}
+%\bibliography{systemfit} % a subset of my big bibtex file
+\bibliography{/home/suapm095/Documents/Literatur/arne}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2006-02-19 paper/trunk/systemfit_estimation.tex
added reference to Zellner & Theil (1962) in the paragraph about 3SLS estimation
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@616 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 In this case the GLS is generally called °Seemingly Unrelated Regression°
-(SUR) (see \citealp{zellner62}). 
+(SUR) \citep{zellner62}.
 Note, while an unbiased OLS or WLS estimation requires only that 
@@
 This estimation procedure is generally called °Three-stage Least
-Squares° (3SLS).
+Squares° (3SLS) \citep{zellner62b}.
 

### 2006-02-19 paper/trunk/systemfit_code.tex
added section about the code inside the systemfit package (this is suggested by the JSS)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@617 591db2bd-e0a0-4d22-8c19-2192dbb7da31

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Source code}\label{sec:code}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The \pkg{systemfit} package includes functions
to estimate systems of equations
(\code{systemfit}, \code{systemfitClassic})
and to test hypotheses in these systems
(\code{ftest.systemfit}, \code{wald\-test.sys\-tem\-fit},
\code{lrtest.systemfit}, \code{hausman.systemfit}).
Furthermore, this package provides some helper functions
e.g. to extract the estimated coefficients (\code{coef.systemfit})
or to calculate predicted values (\code{predict.systemfit}).

The source code of the \pkg{systemfit} is publicly available
for download from °CRAN° (The Comprehensive R Archive Network,
\url{http://cran.r-project.org/src/contrib/Descriptions/systemfit.html}).
Since the whole package has more than 2,100 lines of code,
it is not presented in this article.
Furthermore, the code corresponds exactly to the procedures and formulas
described in the previous section.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection[Basic function systemfit]{The basic function \code{systemfit}}
The basic functionality of this package is provided by the
function \code{systemfit}.
This function estimates the equation system as described
in sections~\ref{sec:Estimation}.
If parameter restrictions are provided, the formulas in
section~\ref{sec:Restrictions} are applied.
Furthermore, the user can control several details of the estimation.
For instance, the
formula to calculate the residual covariance matrix
(see section~\ref{sec:residcov}),
the degrees of freedom for the t-tests
(see section~\ref{sec:degreesOfFreedom}), or
the formula for the 3SLS estimation
(see sections~\ref{sec:Estimation} and~\ref{sec:Restrictions})
can be specified by the user.
The \code{systemfit} function returns many objects
that users might be interest in.
A complete list is available in the documentation of this function
that is included in the package.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection[Wrapper function systemfitClassic]{The wrapper function
   \code{systemfitClassic}}
Furthermore, the \pkg{systemfit} package includes the function
\code{systemfitClassic}.
This is a wrapper function for \code{systemfit}
that can be applied to (classical) panel-like data in long format
if the regressors are the same for all equations.
The data are reshaped and the formulas are modified to enable
an estimation using the standard \code{systemfit} function.
The user can specify whether the coefficients should be restricted
to be equal in all equations.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Statistical tests}
The statistical tests described in
sections~\ref{sec:testingRestrictions} and~\ref{sec:hausman}
are implemented as specified in these sections.
The functions \code{ftest.systemfit}, \code{waldtest.systemfit} and
\code{lrtest.systemfit} test linear restrictions on the
estimated parameters.
On the other hand, the function \code{hausman.systemfit}
tests the consistency of the 3SLS estimator.
All functions return the empirical test statistic,
the degree(s) of freedom, and the p-value.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Efficiency of the code}
We have followed \cite{bates04} to make the code faster and more stable.
First, if a formula contains an inverse of a matrix
that is post-multiplied by a vector,
we use \code{solve(A, b)} instead of \code{solve(A) \%*\% b}.
Second, we calculate crossproducts
by \code{crossprod(X)} or \code{crossprod(X, y)}
instead of \code{t(X) \%*\% X} or \code{t(X) \%*\% y},
respectively.

To calculate $\Omega^{-1}$ in sections~\ref{sec:Estimation}
and~\ref{sec:Restrictions} it is not necessary to invert
the (huge) $\Omega$ matrix,
because it is sufficient to invert the (small) $\Sigma$ matrix and
calculate it by $\Omega^{-1} = \Sigma^{-1} \otimes I$
\citep[p.\ 342]{greene03}.

### 2006-02-20 paper/trunk/runSweave.sh
added section to test the realiability of the results returned by systemfit
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@621 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@ -1,2 +1,4 @@
 #!/bin/sh
 echo "Sweave(\"systemfit_usage.Snw\")" | R --no-save --no-restore
+
+echo "Sweave(\"systemfit_reliability.Snw\")" | R --no-save --no-restore

### 2006-02-20 paper/trunk/systemfit_summary.tex
two modifications in the introduction and in the summary due to the inclusion of section "testing reliability" (in the previous revision)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@622 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 It has been tested on a variety of datasets and has produced satisfactory
-for a few years.%
-\footnote{
-The code to replicate results published in the literature is given
-in the documentation of the cooresponding data sets
-included in the \pkg{systemfit} package
-(see the documentation of \code{GrunfeldGreen}, \code{GrunfeldTheil},
-\code{KleinI} and \code{kmenta}).
-}
+for a few years.
 While the \pkg{systemfit} package performs the basic fitting methods,

### 2006-02-23 paper/trunk/systemfit_intro.tex
added section about the code inside the systemfit package (this is suggested by the JSS)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@617 591db2bd-e0a0-4d22-8c19-2192dbb7da31
two modifications in the introduction and in the summary due to the inclusion of section "testing reliability" (in the previous revision)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@622 591db2bd-e0a0-4d22-8c19-2192dbb7da31
corrected a grammar mistake in the introduction
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@625 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 probably influence the disturbance terms in other equations
-of these model, too.
+of this model, too.
 Ignoring this contemporaneous correlation
@@
 introduce the statistical background of estimating equation systems.
+The implementation of the statistical procedures in \proglang{R}
+is shortly explained in Section~\ref{sec:code}.
 Section~\ref{sec:Usage} demonstrates how to run
@@
 section can be used.
+In Section~\ref{sec:reliability} the reliability of the results from
+\pkg{systemfit} are tested.
 Finally, a summary and outlook are presented in

### 2006-02-23 paper/branches/useR2006/systemfit.tex
branches/useR2006: added abstract for useR 2006
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@626 591db2bd-e0a0-4d22-8c19-2192dbb7da31

--- 2006-02-23 paper/trunk/systemfit_intro.tex
@@
-%       $Id$
+\documentclass[12pt]{scrartcl}
+\input{systemfit_commands}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Introduction}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{document}
+
+\begin{center}
+
+\title{
+systemfit: A Package to Estimate\\
+Simultaneous Equation Systems in \proglang{R}
+}
+
+\vspace{6mm}
+
+\author{
+Arne Henningsen and Jeff D.\ Hamann
+}
+
+\end{center}
 
@@
 
-% For all of the methods developed in the package, the disturbances of
-% the individual equations are assumed to be independent and identically
-% distributed (iid).  
-% In the future, we would like to add the ability to fit equations were
-% the disturbances are serially correlated (wikins 1969).
-
 The \pkg{systemfit} package provides the capability to estimate
@@
 Although linear equation systems can be estimated
-with several other statistical and econometric software packages
-(e.g.\ \proglang{SAS}, \proglang{EViews}, \proglang{TSP}),
+with several other statistical and econometric software packages,
 \pkg{systemfit} has several advantages.
@@
 
-This paper is organized as follows:
-In the following Section~\ref{sec:statistics} we
-introduce the statistical background of estimating equation systems.
-The implementation of the statistical procedures in \proglang{R}
-is shortly explained in Section~\ref{sec:code}.
-Section~\ref{sec:Usage} demonstrates how to run
-\pkg{systemfit} and especially how the features presented in the previous
-section can be used.
-In Section~\ref{sec:reliability} the reliability of the results from
-\pkg{systemfit} are tested.
-Finally, a summary and outlook are presented in
-Section~\ref{sec:Summmary}.
+The \pkg{systemfit} package has been tested on a variety of datasets and
+has produced satisfactory for a few years.
+On the useR! conference,
+we would like to present some of the basic features of the
+\pkg{systemfit} package,
+some of the many details that can be controlled by the user,
+and the statistical tests for parameter restrictions and
+consistency of 3SLS estimation that are included in the package.
+While the \pkg{systemfit} package performs the basic fitting methods,
+more sophisticated methods are still missing.
+We hope to implement missing functionalities
+in the near future
+--- maybe with the help of other useRs.
 
-%%% Local Variables: 
-%%% mode: latex
-%%% TeX-master: "systemfit"
-%%% End:
+\bibliographystyle{jss}
+\bibliography{systemfit} % a subset of my big bibtex file
+%\bibliography{/home/suapm095/Documents/Literatur/arne}
+
+\end{document}

### 2006-02-25 paper/trunk/systemfit.tex
added section about the code inside the systemfit package (this is suggested by the JSS)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@617 591db2bd-e0a0-4d22-8c19-2192dbb7da31
added section to test the realiability of the results returned by systemfit
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@621 591db2bd-e0a0-4d22-8c19-2192dbb7da31
small edits to abstract
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@628 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 
+\hyphenation{sys-tem-fit}
+
 \author{Arne Henningsen\\University of Kiel\And
@@
 structurally related equations.
-Since in most cases cross-equation disturbances are correlated,
+In cases where cross-equation disturbances are correlated,
 full information methods are required \citep{zellner62}.
@@
 on the disturbances in the system,
-instrumental variable estimations should be used for parameter estimation
+then instrumental variable estimation methods should be used
 \citep{zellner62b}
@@
   seemingly unrelated regression, two-stage least squares,
-  three-stage least squares}
+  three-stage least squares, parameter restriction}
 
@@
 \input{systemfit_statistics.tex}
+\input{systemfit_code.tex}
 \input{systemfit_usage.tex}
+\input{systemfit_reliability.tex}
 \input{systemfit_summary.tex}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\bibliography{systemfit} % a subset of my big bibtex file
-\bibliography{/home/suapm095/Documents/Literatur/arne}
+\bibliography{systemfit} % a subset of my big bibtex file
+%\bibliography{/home/suapm095/Documents/Literatur/arne}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2006-02-25 paper/trunk/systemfit_intro.tex
small edits by jeff so that it reads better
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@629 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 consist of more than one equation.
-In this case, the disturbance terms of these equations are likely
+The disturbance terms of these equations are likely
 to be contemporaneously correlated,
-because some unconsidered factors
+because unconsidered factors
 that influence the disturbance term in one equation
-probably influence the disturbance terms in other equations
-of this model, too.
+probably influence the disturbance terms in other equations.
 Ignoring this contemporaneous correlation
@@
 This estimation procedure is generally called
-°Seemingly Unrelated Regression° (SUR) \citep{zellner62}.
+"Seemingly Unrelated Regression" (SUR) \citep{zellner62}.
 Another reason to estimate an equation system simultaneously are
@@
 parameter estimates.
-This simultaneity bias can be circumvented by applying
-a °Two-Stage Least Squares° (2SLS) or °Three-Stage Least Squares° (3SLS)
-estimation of the equation system.
+This simultaneity bias can be corrected for in each equation by
+applying a "Two-Stage Least Squares" (2SLS) method or for all
+equations simultaneously when combined with SUR resulting in a
+"Three-Stage Least Squares" (3SLS) estimation of the system of
+equations.
 
@@
 
-This paper is organized as follows:
-In the following Section~\ref{sec:statistics} we
-introduce the statistical background of estimating equation systems.
-The implementation of the statistical procedures in \proglang{R}
-is shortly explained in Section~\ref{sec:code}.
+% This paper is organized as follows:
+In Section~\ref{sec:statistics} we introduce the statistical
+background of estimating equation systems.
+The implementation of the statistical procedures in \proglang{R} is
+shortly explained in Section~\ref{sec:code}.
 Section~\ref{sec:Usage} demonstrates how to run
-\pkg{systemfit} and especially how the features presented in the previous
-section can be used.
+\pkg{systemfit} and how some of the features presented in the previous
+section can be utilized.
 In Section~\ref{sec:reliability} the reliability of the results from
-\pkg{systemfit} are tested.
+\pkg{systemfit} are presented.
 Finally, a summary and outlook are presented in

### 2006-02-25 paper/trunk/systemfit_estimation.tex
added section about the code inside the systemfit package (this is suggested by the JSS)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@617 591db2bd-e0a0-4d22-8c19-2192dbb7da31
the capitalization of all (sub)(sub)section headers is now in sentence-style (according to the guidelines of the JSS)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@618 591db2bd-e0a0-4d22-8c19-2192dbb7da31
very small edits to improve readibility
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@630 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 Least Squares (GLS) estimation leads to efficient parameter estimates.
-In this case the GLS is generally called °Seemingly Unrelated Regression°
+In this case, the GLS is generally called "Seemingly Unrelated Regression"
 (SUR) \citep{zellner62}.
-Note, while an unbiased OLS or WLS estimation requires only that 
+It should be noted that while an unbiased OLS or WLS estimation requires only that 
 the regressors and the disturbance terms of each single 
@@
 
-The SUR estimator can be obtained by:%
-\footnote{
-To calculate $\Omega^{-1}$ it is not necessary to invert 
-the (huge) $\Omega$ matrix, 
-because it is sufficient to invert the (small) $\Sigma$ matrix and
-calculate it by $\Omega^{-1} = \Sigma^{-1} \otimes I$
-\citep[p.\ 342]{greene03}.
-}
+The SUR estimator can be obtained by:
 \begin{equation}
    \widehat{\beta}_{SUR} = \left( X' \Omega^{-1} X \right)^{-1} X' \Omega^{-1} y
+   \label{eq:sur}
 \end{equation}
@@
    \Cov \left[ \widehat{\beta}_{SUR} \right] = \left( X' \Omega^{-1} X \right)^{-1}
+   \label{eq:surCov}
 \end{equation}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Two-Stage least squares (2SLS)}
+\subsubsection{Two-stage least squares (2SLS)}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Three-Stage least squares (3SLS)}
+\subsubsection{Three-stage least squares (3SLS)}
 

### 2006-02-25 paper/trunk/systemfit_statistics.tex
the capitalization of all (sub)(sub)section headers is now in sentence-style (according to the guidelines of the JSS)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@618 591db2bd-e0a0-4d22-8c19-2192dbb7da31
added footer for nested docs
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@633 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Statistical Background}\label{sec:statistics}
+\section{Statistical background}\label{sec:statistics}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 In this section we provide the statistical background
-of the functionality provided in the \pkg{systemfit} package.
+of the functionality provided by the \pkg{systemfit} package.
 After introducing notations and assumptions,
-we provide the formulas to estimate linear equation systems.
-Then, we shows how linear restrictions can be imposed.
-Finally, some other relevant issues about estimation of equation systems
-are discussed.
+we provide the formulas to estimate systems of linear equations.
+We then demonstrate how to impose linear restrictions on parameters.
+Finally, we present additional relevant issues about estimation of
+equation systems.
 
@@
 \input{systemfit_other.tex}
+
+%%% Local Variables: 
+%%% mode: latex
+%%% TeX-master: "systemfit"
+%%% End:

### 2006-02-25 paper/trunk/systemfit_usage.tex
small change for readibility (of->on)... very small indeed
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@634 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 
+In this section we demonstrate how to use the \pkg{systemfit} package.
+First, we show the standard usage of \code{systemfit} by a simple example.
+Second, several options that can be specified by the user are presented.
+Then, the wrapper function \code{systemfitClassic} is described.
+Finally, we demonstrate how to apply some statistical tests.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Standard usage}
+\subsection[Standard usage of systemfit]{Standard usage of \code{systemfit}}
+\label{sec:standard-usage}
 
-\pkg{systemfit} is generally called by
-
-\code{
-R> systemfit( method, eqns )
-}
+As described in the previous section,
+equation systems can be econometrically estimated
+with the function \code{systemfit}.
+It is generally called by
+\begin{Schunk}
+\begin{Sinput}
+> systemfit(method, eqns)
+\end{Sinput}
+\end{Schunk}
 
 There are two mandatory arguments: \code{method} and \code{eqns}.
+The argument \code{method} is a string determining the estimation method.
+It must be either °OLS°, °WLS°, °SUR°, °WSUR°, °2SLS°, °W2SLS°, °3SLS°,
+or °W3SLS°.
+While six of these methods correspond to the estimation methods
+described in sections~\ref{sec:Estimation} and~\ref{sec:Restrictions},
+the methods °WSUR° and °W3SLS° are °SUR° and °3SLS° estimations
+using the residual covariance matrices from
+°WLS° and °W2SLS° estimations, respectively
+(see section~\ref{sec:residcov}).
+The other mandatory argument, \code{eqns}, is a list of the equations
+to be estimated.
+Each equation is a standard formula in \proglang{R} and
+starts with a dependent variable on the left hand side.
+After a tilde ($\sim$) the regressors are listed,
+separated by plus signs%
+\footnote{
+For Details see the \proglang{R} help files to \code{formula}.
+}.
 
-The argument \code{method} is a string determining the estimation method.
-It must be one of "OLS", "WLS", "SUR", "2SLS", "W2SLS" or "3SLS".
-
-The other mandatory argument \code{eqns} is a list of the equations 
-to estimate. 
-Each equation is a standard formula in \proglang{R}.
-It starts with a dependent variable on the left hand side.
-After a tilde ($\sim$) the regressors are listed%
-\footnote{For Details see the \proglang{R} help files to \code{formula}}.
-
-This is now demonstrated using an example: \\
-\code{
-R> library( systemfit ) \\
-R> data( kmenta ) \\
-R> attach( kmenta ) \\
-R> fitsur <- systemfit( "SUR", list( q $\sim$ p + d, q $\sim$ p + f + a ) ) \\
-}
+The following demonstration uses an example taken from
+\citet[p.\ 685]{kmenta86}.
+We want to estimate a small model of US the food market:
+\begin{align}
+\texttt{consump} & = \beta_1 + \beta_2 * \texttt{price}
+   + \beta_3 * \texttt{income} \\
+\texttt{consump} & = \beta_4 + \beta_5 * \texttt{price}
+   + \beta_6 * \texttt{farmPrice} + \beta_7 * \texttt{trend}
+\end{align}
+The first equation represents the demand side of the food market.
+Variable \code{consump} (food consumption per capita) is the dependant variable.
+The regressors are \code{price} (ratio of food prices to general consumer prices)
+and \code{income} (disposable income) as well as a constant.
+The second equation specifies the supply side of the food market.
+Variable \code{consump} is the dependant variable of this equation as well.
+The regressors are again \code{price} (ratio of food prices to general 
+consumer prices) and a constant as well as 
+\code{farmPrice} (ratio of preceding year's prices received by farmers) and 
+\code{trend} (a time trend in years).
+These equations can be estimated as SUR in \proglang{R} by
+\begin{Schunk}
+\begin{Sinput}
+> library(systemfit)
+> data(Kmenta)
+> attach(Kmenta)
+> eqDemand <- consump ~ price + income
+> eqSupply <- consump ~ price + farmPrice + trend
+> fitsur <- systemfit("SUR", list(demand = eqDemand, supply = eqSupply))
+\end{Sinput}
+\end{Schunk}
 
 The first line loads the \pkg{systemfit} package.
-The second line loads example data that are included in this package.
-These data come from \cite{kmenta86}.
+The second line loads example data that are included with the package.
 They are attached to the \proglang{R} search path in line three.
-In the last line a seemingly unrelated regression is done.
-The first equation represents the demand side of the food market.
-The dependant variable is \code{q} (food consumption per capita). 
-The regressors are \code{p} (ratio of food prices to general consumer prices)
-and \code{d} (disposable income) as well as a constant%
-\footnote{a regression constant is always implied if not explicitly omitted.}.
-The second equation represents the supply side.
-Variable \code{q} (food consumption per capita) is also the dependant 
-variable of this equation. 
-The regressors are again \code{p} (ratio of food prices to general 
-consumer prices) and a constant as well as 
-\code{f} (ratio of preceding year's prices received by farmers) and 
-\code{a} (a time trend in years).
-The regression result is assigned to the variable \code{fitsur}.
+In the fourth and fifth line, the demand and supply equations are specified,
+respectively%
+\footnote{
+A regression constant is always implied if not explicitly omitted.
+}.
+Finally, in the last line, a seemingly unrelated regression is performed
+and the regression result is assigned to the variable \code{fitsur}.
 
-Summary results can be printed by\\
-\code{
-R>~summary(~fitsur~)~\\
-~\\
-systemfit~results~\\
-method:~SUR~\\
-\\
-\mbox{}~~~N~DF~~~~~~SSR~~~~~MSE~~~~RMSE~~~~~~~R2~~~Adj~R2 \\
-1~20~17~~65.6829~3.86370~1.96563~0.755019~0.726198~\\
-2~20~16~104.0584~6.50365~2.55023~0.611888~0.539117~\\
-~\\
-The~covariance~matrix~of~the~residuals~used~for~estimation\\
-\mbox{}~~~~~~~~1~~~~~~~2~\\
-1~3.72539~4.13696~\\
-2~4.13696~5.78444~\\
-~\\
-The~covariance~matrix~of~the~residuals\\
-\mbox{}~~~~~~~~1~~~~~~~2~\\
-1~3.86370~4.92431~\\
-2~4.92431~6.50365~\\
-~\\
-The~correlations~of~the~residuals\\
-\mbox{}~~~~~~~~~1~~~~~~~~2~\\
-1~1.000000~0.982348~\\
-2~0.982348~1.000000~\\
-~\\
-The~determinant~of~the~residual~covariance~matrix:~0.879285~\\
-OLS~R-squared~value~of~the~system:~0.683453~\\
-McElroy's~R-squared~value~for~the~system:~0.788722~\\
-~\\
-SUR~estimates~for~1~~(equation~1~)~\\
-Model~Formula:~q~~~p~+~d\\
-\\
-\mbox{}~~~~~~~~~~~~~Estimate~Std.~Error~~~t~value~Pr(>|t|)~\\
-(Intercept)~99.332894~~~7.514452~13.218913~~~~~~~~0~***~\\
-p~~~~~~~~~~~-0.275486~~~0.088509~-3.112513~0.006332~~**~\\
-d~~~~~~~~~~~~~0.29855~~~0.041945~~7.117605~~~~2e-06~***~\\
----~\\
-Signif.~codes:~~0~`***'~0.001~`**'~0.01~`*'~0.05~`.'~0.1~`~'~1~\\
-~\\
-Residual~standard~error:~1.96563~on~17~degrees~of~freedom~\\
-Number~of~observations:~20~Degrees~of~Freedom:~17~\\
-SSR:~65.682902~MSE:~3.8637~Root~MSE:~1.96563~\\
-Multiple~R-Squared:~0.755019~Adjusted~R-Squared:~0.726198~\\
-~\\
-~\\
-SUR~estimates~for~2~~(equation~2~)~\\
-Model~Formula:~q~~~p~+~f~+~a\\
-\\
-\mbox{}~~~~~~~~~~~~~Estimate~Std.~Error~~t~value~Pr(>|t|)~\\
-(Intercept)~61.966166~~~11.08079~5.592215~~~~4e-05~***~\\
-p~~~~~~~~~~~~0.146884~~~0.094435~1.555397~0.139408~\\
-f~~~~~~~~~~~~0.214004~~~0.039868~5.367761~~6.3e-05~***~\\
-a~~~~~~~~~~~~0.339304~~~0.067911~4.996283~0.000132~***~\\
----~\\
-Signif.~codes:~~0~`***'~0.001~`**'~0.01~`*'~0.05~`.'~0.1~`~'~1~\\
-~\\
-Residual~standard~error:~2.550226~on~16~degrees~of~freedom~\\
-Number~of~observations:~20~Degrees~of~Freedom:~16~\\
-SSR:~104.05843~MSE:~6.503652~Root~MSE:~2.550226~\\
-Multiple~R-Squared:~0.611888~Adjusted~R-Squared:~0.539117~\\
-}
+Summary results can be printed by
+\begin{Schunk}
+\begin{Sinput}
+> summary(fitsur)
+\end{Sinput}
+\begin{Soutput}
+systemfit results 
+method: SUR
+
+        N DF      SSR     MSE    RMSE       R2   Adj R2
+demand 20 17  65.6829 3.86370 1.96563 0.755019 0.726198
+supply 20 16 104.0584 6.50365 2.55023 0.611888 0.539117
+
+The covariance matrix of the residuals used for estimation
+        demand  supply
+demand 3.72539 4.13696
+supply 4.13696 5.78444
+
+The covariance matrix of the residuals
+        demand  supply
+demand 3.86370 4.92431
+supply 4.92431 6.50365
+
+The correlations of the residuals
+         demand   supply
+demand 1.000000 0.982348
+supply 0.982348 1.000000
+
+The determinant of the residual covariance matrix: 0.879285
+OLS R-squared value of the system: 0.683453
+McElroy's R-squared value for the system: 0.788722
+
+SUR estimates for 'demand' (equation 1)
+Model Formula: consump ~ price + income
+
+             Estimate Std. Error   t value Pr(>|t|)    
+(Intercept) 99.332894   7.514452 13.218913        0 ***
+price       -0.275486   0.088509 -3.112513 0.006332  **
+income        0.29855   0.041945  7.117605    2e-06 ***
+---
+Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
+
+Residual standard error: 1.96563 on 17 degrees of freedom
+Number of observations: 20 Degrees of Freedom: 17 
+SSR: 65.682902 MSE: 3.8637 Root MSE: 1.96563 
+Multiple R-Squared: 0.755019 Adjusted R-Squared: 0.726198
+
+SUR estimates for 'supply' (equation 2)
+Model Formula: consump ~ price + farmPrice + trend
+
+             Estimate Std. Error  t value Pr(>|t|)    
+(Intercept) 61.966166   11.08079 5.592215    4e-05 ***
+price        0.146884   0.094435 1.555397 0.139408    
+farmPrice    0.214004   0.039868 5.367761  6.3e-05 ***
+trend        0.339304   0.067911 4.996283 0.000132 ***
+---
+Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
+
+Residual standard error: 2.550226 on 16 degrees of freedom
+Number of observations: 20 Degrees of Freedom: 16 
+SSR: 104.05843 MSE: 6.503652 Root MSE: 2.550226 
+Multiple R-Squared: 0.611888 Adjusted R-Squared: 0.539117 
+\end{Soutput}
+\end{Schunk}
+
+First, the estimation method is reported
+and a few summary statistics for each equation are given.
+Then, some results regarding the whole equation system are printed:
+covariance matrix and correlations of the residuals,
+log of the determinant of the residual covariance matrix,
+$R^2$ value of the whole system, and
+McElroy's $R^2$ value.
+If the model is estimated by (W)SUR or (W)3SLS,
+the covariance matrix used for estimation is printed additionally.
+Finally, the estimation results of each equation are reported:
+formula of the estimated equation,
+estimated parameters, their standard errors, t-values, p-values
+and codes indicating their statistical significance,
+and some other statistics like
+standard error of the residuals or
+$R^2$ value of the equation.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{User options}
+\subsection[User options of systemfit]{User options of \code{systemfit}}
+\label{sec:user-options}
 
-Following additional options can be set by the user:
+The user can modify the default estimation method by providing
+additional optional arguments,
+e.g.\ to specify instrumental variables
+or to impose parameter restrictions.
+All optional arguments are described in the following:
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% \subsubsection{Equation labels}
-\paragraph{Equation labels}
+\subsubsection{Equation labels}
 The optional argument \code{eqnlabels} allows the user to label the equations.
-It has to be a vector of strings naming the equations.\\
-\code{
-R>~fitsur~<-~systemfit(~"SUR",~list(~q~~~p~+~d,~q~~~p~+~f~+~a~),\\~
-R+~~~~eqnlabels~=~c(~"demand",~"supply"~)~)\\
-R>~summary(~fitsur~)\\
-systemfit~results\\
-method:~SUR\\
-\\
-\mbox{}~~~~~~~~N~DF~~~~~~SSR~~~~~MSE~~~~RMSE~~~~~~~R2~~~Adj~R2\\
-demand~20~17~~65.6829~3.86370~1.96563~0.755019~0.726198\\
-supply~20~16~104.0584~6.50365~2.55023~0.611888~0.539117\\
-\ldots\\
-}
-If no equation labels are provided, the equations are numbered.
+It has to be a vector of strings.
+If this argument is not provided, the labels are taken from the names of
+the equations in argument \code{eqns}.
+And if the equations have no names, they are numbered consecutively.
+Hence, the following command has the same effect as the command above.
+\begin{Schunk}
+\begin{Sinput}
+> fitsur <- systemfit("SUR", list(eqDemand, eqSupply), eqnlabels = c("demand", 
++     "supply"))
+\end{Sinput}
+\end{Schunk}
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\subsubsection{Instrumental variables}   
-\paragraph{Instrumental variables}   
-\code{inst}one-sided model formula specifying instrumental variables
-   or a list of one-sided model formulas if different instruments should
-   be used for the different equations (only needed for 2SLS, W2SLS and
-   3SLS estimations).
+\subsubsection{Instrumental variables}
+The instruments for a 2SLS, W2SLS or 3SLS estimation can be specified
+by the argument \code{inst}.
+If the same instruments should be used for all equations,
+\code{inst} must be a one-sided formula.
+If different instruments should be used for the equations,
+\code{inst} must be a list that contains a one-sided formula
+for each equation.
+The first example uses the same instruments for all equations,
+and the second uses different instruments:
+\begin{Schunk}
+\begin{Sinput}
+> fit3sls <- systemfit("3SLS", list(demand = eqDemand, supply = eqSupply), 
++     inst = ~income + farmPrice + trend)
+> fit3sls2 <- systemfit("3SLS", list(demand = eqDemand, supply = eqSupply), 
++     inst = list(~farmPrice + trend, ~income + farmPrice + trend))
+\end{Sinput}
+\end{Schunk}
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\subsubsection{Data}   
-\paragraph{Data}   
-\code{data} an optional data frame containing the variables in the model.
-   By default the variables are taken from the environment from which
-   systemfit is called.
+\subsubsection{Data}
+Having all data in the global environment or attached to the search path
+is often inconvenient.
+Therefore, a data frame \code{data} can be provided
+that contains the variables of the model.
+In the following example we do not need to attach the data frame
+\code{Kmenta} before calling systemfit:
+\begin{Schunk}
+\begin{Sinput}
+> fitsur <- systemfit("SUR", list(eqDemand, eqSupply), data = Kmenta)
+\end{Sinput}
+\end{Schunk}
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\subsubsection{Restrictions}   
-\paragraph{Restrictions}   
-\code{R.restr} an optional j x k matrix to impose linear
-   restrictions on the parameters by \code{R.restr} * $\beta$ = \code{q.restr}
-   (j = number of restrictions, k = number of all parameters,
-   $\beta$ = vector of all parameters).
+\subsubsection{Parameter restrictions}
+As outlined in section~\ref{sec:Restrictions}, parameter restrictions
+can be imposed in two ways.
+The first way is to use the transformation matrix $T$
+that can be specified by the argument \code{TX}.
+If we want to impose the restriction, say $\beta_2 = - \beta_6$,
+we can do this as follows
+\begin{Schunk}
+\begin{Sinput}
+> tx <- matrix(0, nrow = 7, ncol = 6)
+> tx[1, 1] <- 1
+> tx[2, 2] <- 1
+> tx[3, 3] <- 1
+> tx[4, 4] <- 1
+> tx[5, 5] <- 1
+> tx[6, 2] <- -1
+> tx[7, 6] <- 1
+> fitsurTx <- systemfit("SUR", list(eqDemand, eqSupply), TX = tx)
+\end{Sinput}
+\end{Schunk}
+The first line creates a $7 \times 6$ matrix of zeros,
+where 7 is the number of unrestricted coefficients and
+6 is the number of restricted coefficients.
+The following seven lines specify this matrix in a way
+that the unrestricted coefficients $( \beta )$ are assigned
+to the restricted coefficients $( \beta^* )$ with
+$\beta_1 = \beta^*_1$,
+$\beta_2 = \beta^*_2$,
+$\beta_3 = \beta^*_3$,
+$\beta_4 = \beta^*_4$,
+$\beta_5 = \beta^*_5$,
+$\beta_6 = -\beta^*_2$, and
+$\beta_7 = \beta^*_6$.
+Finally the model is estimated with restriction
+$\beta^*_2 = \beta_2 = - \beta_6$ imposed.
 
-\code{q.restr} an optional j x 1 matrix to impose linear
-   restrictions (see \code{R.restr}); default is a j x 1 matrix
-   that contains only zeros.
-
-\code{TX} an optional matrix to transform the regressor matrix and,
-   hence, also the coefficient vector (see details).
+The second way to impose parameter restrictions is to use
+the matrix $R$ and the vector $q$
+(see section~\ref{sec:Restrictions}).
+Matrix $R$ can be specified with the argument \code{R.restr}
+and vector $q$ with argument \code{q.restr}.
+We convert the restriction specified above to $\beta_2 + \beta_6 = 0$
+and impose it in the second way:
+\begin{Schunk}
+\begin{Sinput}
+> Rmat <- matrix(0, nrow = 1, ncol = 7)
+> Rmat[1, 2] <- 1
+> Rmat[1, 6] <- 1
+> qvec <- c(0)
+> fitsurRmat <- systemfit("SUR", list(eqDemand, eqSupply), R.restr = Rmat, 
++     q.restr = qvec)
+\end{Sinput}
+\end{Schunk}
+\label{code:Rmat}
+The first line creates a $1 \times 7$ matrix of zeros,
+where 1 is the number of restrictions and
+7 is the number of unrestricted coefficients.
+The following two lines specify this matrix in a way
+that the multiplication with the parameter vector
+results in $ \beta_2 + \beta_6 $.
+The fourth line creates a vector with a single element
+that contains the left hand side of the restriction, i.e.\ zero.
+Finally the model is estimated with restriction
+$\beta_2 + \beta_6 = 0$ imposed.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\subsubsection{Iteration control}   
-\paragraph{Iteration control}
-\code{maxiter} maximum number of iterations for WLS, SUR, W2SLS and
-   3SLS estimations.
-
-\code{tol} tolerance level indicating when to stop the iteration (only
-   WLS, SUR, W2SLS and 3SLS estimations).
+\subsubsection{Iteration control}
+The estimation methods WLS, SUR, W2SLS and 3SLS need a
+covariance matrix of the residuals
+that can be calculated from a first-step OLS or 2SLS estimation
+(see section~\ref{sec:residcov}).
+If the argument \code{maxiter} is set to a number large than one,
+this procedure is iterated and at each iteration the covariance
+matrix is calculated from the previous step estimation.
+This iteration is repeated until the maximum number of iterations
+is reached or the parameter estimates have converged.
+The maximum number of iterations is specified by the argument \code{maxiter}.
+Its default value is one, which means no iteration.
+The convergence criterion is
+\begin{equation}
+   \sqrt{ \frac{ \sum_i (b_{i,g} - b_{i,g-1})^2 }{ \sum_i b_{i,g-1}^2 }}
+      < \texttt{tol}
+\end{equation}
+where $b_{i,g}$ is the $i$th coefficient of the $g$th iteration.
+The default value of \code{tol} is $10^{-5}$.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\subsubsection{Residual covariance matrix}   
-\paragraph{Residual covariance matrix}   
-\code{rcovformula} formula to calculate the estimated residual covariance
-   matrix (see details).
+\subsubsection{Residual covariance matrix}
+It was explained in section~\ref{sec:residcov} that several different
+formulas have been proposed to calculate the residual
+covariance matrix.
+The user can specify, which formula \pkg{systemfit} should use.
+Possible values of the argument \code{rcovformula} are presented in
+table~\ref{tab:rcovformula}.
+By default, \pkg{systemfit} uses equation (\ref{eq:rcov1}).
+
+\begin{table}[H]
+\caption{Possible values of argument \code{rcovformula}}
+\label{tab:rcovformula}
+\centering
+\begin{tabular}{|c|c|}
+\hline
+argument & equation \\
+\code{rcovformula} & \\
+\hline
+0              & \ref{eq:rcov0} \\
+\hline
+1 or 'geomean' & \ref{eq:rcov1} \\
+\hline
+2 or 'Theil'   & \ref{eq:rcov2} \\
+\hline
+3 or 'max'     & \ref{eq:rcov3} \\
+\hline
+\end{tabular}
+\end{table}
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\subsubsection{3SLS formula}   
-\paragraph{3SLS formula}   
-\code{formula3sls} formula for calculating the 3SLS estimator,
-   one of "GLS", "IV", "GMM", "Schmidt" or "EViews" (see details).
+\subsubsection{3SLS formula}
+As discussed in sections~\ref{sec:Estimation} and~\ref{sec:Restrictions},
+there exist several different
+formulas to perform a 3SLS estimation.
+The user can specify the applied formula by the argument \code{formula3sls}.
+Possible values are presented in table~\ref{tab:formula3sls}.
+The default value is 'GLS'.
+
+\begin{table}[H]
+\caption{Possible values of argument \code{formula3sls}}
+\label{tab:formula3sls}
+\centering
+\begin{tabular}{|c|c|c|}
+\hline
+argument           & equation       & equation \\
+\code{formula3sls} & (unrestricted) & (restricted) \\
+\hline
+'GLS'     & \ref{eq:3slsGls}     & \ref{eq:3slsGlsR} \\
+\hline
+'IV'      & \ref{eq:3slsIv}      & \ref{eq:3slsIvR} \\
+\hline
+'GMM'     & \ref{eq:3slsGmm}     & \ref{eq:3slsGmmR} \\
+\hline
+'Schmidt' & \ref{eq:3slsSchmidt} & \ref{eq:3slsSchmidtR} \\
+\hline
+'EViews'  & \ref{eq:3slsEViews}  & \ref{eq:3slsEViewsR} \\
+\hline
+\end{tabular}
+\end{table}
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\subsubsection{Degrees of freedom for t-tests}   
-\paragraph{Degrees of freedom for t-tests}   
-\code{probdfsys} use the degrees of freedom of the whole system
-   (in place of the degrees of freedom of the single equation)
-   to calculate prob values for the t-test of individual parameters.
+\subsubsection{Degrees of freedom for t-tests}
+There exist two different approaches to determine the degrees of freedom
+to perform t-tests on the estimated parameters
+(section~\ref{sec:degreesOfFreedom}).
+This can be specified with argument \code{probdfsys}.
+If it is \code{TRUE},
+the degrees of freedom of the whole system are taken.
+In contrast, if \code{probdfsys} is \code{FALSE},
+the degrees of freedom of the single equation are taken.
+By default, \code{probdfsys} is \code{TRUE}, if any restrictions
+are specified using either the argument \code{R.restr} or the
+argument \code{TX}, and it is \code{FALSE} otherwise.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\subsubsection{Sigma squared}   
-\paragraph{Sigma squared}   
-\code{single.eq.sigma} use different $\sigma^2$s for each
-   single equation to calculate the covariance matrix and the
-   standard errors of the coefficients (only OLS and 2SLS).
+\subsubsection{Sigma squared}
+In case of OLS or 2SLS estimations, argument \code{single.eq.sigma}
+can be used to specify,
+whether different $\sigma^2$s for each single equation
+or the same $\sigma^2$ for all equations should be used.
+If argument \code{single.eq.sigma} is \code{TRUE},
+equations (\ref{eq:olsCovSingleSigma})
+and (\ref{eq:2slsCovSingleSigma}) are applied.
+In contrast, if argument \code{single.eq.sigma} is \code{FALSE},
+equations (\ref{eq:olsCovSameSigma})
+and (\ref{eq:2slsCovSameSigma}) are applied.
+By default, \code{single.eq.sigma} is \code{FALSE}, if any restrictions
+are specified using either the argument \code{R.restr} or the
+argument \code{TX}, and it is \code{TRUE} otherwise.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\subsubsection{System options}   
-\paragraph{System options}
+\subsubsection{System options}
+Finally, two options regarding some internal calculations are available.
+First, argument \code{solvetol} specifies the tolerance level
+for detecting linear dependencies when inverting a matrix or
+calculating a determinant (using functions \code{solve} and \code{det}).
+The default value depends on the used computer system and is equal to the
+default tolerance level of \code{solve} and \code{det}.
+Second, argument \code{saveMemory} can be used in case of large data sets
+to accelerate the estimation by omitting some calculation
+that are not crucial for the basic estimation.
+Currently, only the calculation of McElroy's $R^2$ is omitted.
+The default value of argument \code{saveMemory} is \code{TRUE},
+if the argument \code{data} is specified and the number of observations
+times the number of equations is larger than 1000,
+and it is \code{FALSE} otherwise.
 
-\code{solvetol} tolerance level for detecting linear dependencies
-   when inverting a matrix or calculating a determinant (see
-   \code{solve} and \code{det}).
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection[Wrapper function systemfitClassic]{The wrapper function
+   \code{systemfitClassic}}
 
-\code{saveMemory} save memory by omitting some calculation that
-   are not crucial for the basic estimation (e.g McElroy's
-   $R^2$).
+The wrapper function \code{systemfitClassic}
+can be applied to (classical) panel-like data in long format
+if the regressors are the same for all equations.
+This function is called by
+\begin{Schunk}
+\begin{Sinput}
+> systemfitClassic(method, formula, eqnVar, timeVar, data)
+\end{Sinput}
+\end{Schunk}
+
+The mandatory arguments are \code{method}, \code{formula}, \code{eqnVar},
+and \code{timeVar}.
+Argument \code{method} is the same as in \code{systemfit}
+(see section~\ref{sec:standard-usage}).
+The second argument \code{formula} is a standard formula in \proglang{R}
+that will be applied to all equations.
+Argument \code{eqnVar} specifies the variable name indicating
+the equation to which the observation belongs,
+and argument \code{timeVar} specifies the variable name indicating
+the time.
+Finally, \code{data} is a \code{data.frame} that contains all required data.
+
+We demonstrate the usage of \code{systemfitClassic}
+using an example taken from \citet[pp.\ 295, 300]{theil71}
+that is based on \citet{grunfeld58}.
+We want to estimate a model for gross investment of 2 US firms
+in the years 1935--1954:
+\begin{equation}
+\texttt{invest}_{it} = \beta_1 + \beta_2 * \texttt{value}_{it} +
+   \beta_3 * \texttt{capital}_{it}
+\end{equation}
+where \code{invest} is the gross investment of firm $i$ in year $t$,
+\code{value} is the market value of the firm at the end of the
+previous year, and
+\code{capital} is the capital stock of the firm at the end of the
+previous year.
+
+This model can be estimated by
+\begin{Schunk}
+\begin{Sinput}
+> data("GrunfeldTheil")
+> theilSur <- systemfitClassic("SUR", invest ~ value + capital, 
++     "firm", "year", data = GrunfeldTheil)
+\end{Sinput}
+\end{Schunk}
+The first line loads example data that are included with the package.
+And then, a seemingly unrelated regression is performed,
+where the variable °firm° indicates the firm
+and the variable °year° indicates the time.
+
+The function \code{systemfitClassic} has also an optional argument
+\code{pooled} that is a logical variable indicating
+whether the coefficients should be restricted to be equal in all equations.
+By default, this argument is set to °\code{FALSE}°.
+In addition all optional arguments of \code{systemfit}
+(see section~\ref{sec:user-options})
+except for \code{eqnLabels} and \code{TX}
+can be used with \code{systemfitClassic}, too.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Testing linear restrictions}
+
+As described in section~\ref{sec:testingRestrictions},
+linear restrictions can be tested by an F test, Wald test or
+LR test.
+The corresponding functions in package \pkg{systemfit} are
+\code{ftest.systemfit}, \code{waldtest.systemfit}, and
+\code{lrtest.systemfit}, respectively.
+
+We will now test the restriction $\beta_2 = -\beta_6$
+that was specified by the matrix \code{Rmat} and the vector \code{qvec}
+in the example above (p.~\pageref{code:Rmat}).
+\begin{Schunk}
+\begin{Sinput}
+> ftest.systemfit(fitsur, Rmat, qvec)
+\end{Sinput}
+\begin{Soutput}
+F-test for linear parameter restrictions in equation systems
+F-statistic: 0.9322 
+degrees of freedom of the numerator: 1 
+degrees of freedom of the denominator: 33 
+p-value: 0.3413 
+\end{Soutput}
+\begin{Sinput}
+> waldtest.systemfit(fitsur, Rmat, qvec)
+\end{Sinput}
+\begin{Soutput}
+Wald-test for linear parameter restrictions in equation systems
+Wald-statistic: 0.6092 
+degrees of freedom: 1 
+p-value: 0.4351 
+\end{Soutput}
+\begin{Sinput}
+> lrtest.systemfit(fitsurRmat, fitsur)
+\end{Sinput}
+\begin{Soutput}
+Likelihood-Ratio-test for parameter restrictions in equation systems
+LR-statistic: 1.004 
+degrees of freedom: 1 
+p-value: 0.3163 
+\end{Soutput}
+\end{Schunk}
+The linear restrictions are tested by an F test first,
+then by a Wald test, and finally by an LR test.
+The functions \code{ftest.systemfit} and \code{waldtest.systemfit}
+have three arguments.
+The first argument must be an unrestricted regression
+returned by \code{systemfit}.
+The second and third argument are the restriction matrix $R$ and
+the vector $q$ as described in section~\ref{sec:Restrictions}.
+In contrast, the function \code{lrtest.systemfit} needs two arguments.
+The first argument must be a restricted and the second an unrestricted
+regression returned by \code{systemfit}.
+
+All tests print a short explanation first.
+Then the empirical test statistic and the degree(s) of freedom are reported.
+Finally the p-value is printed.
+While there is some variation of the p-values across
+the three different tests,
+all tests suggest the same decision:
+The null hypothesis $\beta_2 = -\beta_6$ cannot be rejected at any
+reasonable level of significance.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Hausman test}
+A Hausman test, which is described in section~\ref{sec:hausman},
+can be carried out with following commands:
+\begin{Schunk}
+\begin{Sinput}
+> fit2sls <- systemfit("2SLS", list(demand = eqDemand, supply = eqSupply), 
++     inst = ~income + farmPrice + trend, data = Kmenta)
+> fit3sls <- systemfit("3SLS", list(demand = eqDemand, supply = eqSupply), 
++     inst = ~income + farmPrice + trend, data = Kmenta)
+> hausman.systemfit(fit2sls, fit3sls)
+\end{Sinput}
+\begin{Soutput}
+	Hausman specification test for consistency of the 3SLS estimation
+
+data:  Kmenta 
+Hausman = 2.5357, df = 7, p-value = 0.9244
+\end{Soutput}
+\end{Schunk}
+First of all, the model is estimated by 2SLS and then by 3SLS.
+Finally, in the last line
+the test is carried out by the command \code{hausman.systemfit}.
+This function requires two arguments:
+the result of a 2SLS estimation and
+the result of a 3SLS estimation.
+The Hausman test statistic is
+2.536,
+which has a $\chi^2$ distribution
+with 7 degrees of freedom
+under the null hypothesis.
+The corresponding p-value is
+0.924.
+This shows that the null hypothesis is not rejected
+at any reasonable level of significance.
+Hence, we can assume that the 3SLS estimator is consistent.

### 2006-02-25 paper/trunk/systemfit_other.tex
rorganized the sections and subsections: there now a section called "Statistical Background" that includes the subsections (former sections) "Estimation", "Restrictions" and "Other issues" (a section of the "statistical background" is suggested by the JSS
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@614 591db2bd-e0a0-4d22-8c19-2192dbb7da31
the capitalization of all (sub)(sub)section headers is now in sentence-style (according to the guidelines of the JSS)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@618 591db2bd-e0a0-4d22-8c19-2192dbb7da31
changed reference fo

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Other issues and tools}\label{sec:Other}
+%\section{Other issues and tools}\label{sec:Other}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+In some uncommon cases,
+for instance in pooled estimations,
+where the coefficients are restricted to be equal in all equations,
+the means of the residuals of each equation are not equal to zero
+$( \overline{ \widehat{u} }_i \neq 0 )$.
+Therefore, we subtract the means from the residuals
+and substitute $\widehat{u}_i - \overline{ \widehat{u} }_i$
+for $\widehat{u}_i$ in (\ref{eq:rcov0}--\ref{eq:rcov2}).
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Testing Linear Restrictions}
+\subsection{Testing linear restrictions}
 \label{sec:testingRestrictions}
@@
 with $j$ and $M \cdot T - K$ degrees of freedom
-\citep[p.\ 346]{greene03}.
+\citep[p.\ 314]{theil71}.
 

### 2006-02-25 paper/trunk/systemfit_other.tex
I'm not sure what I changed here.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@644 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 
-Hausman \citep{hausman78} developed a test for misspecification.
+\citet{hausman78} developed a test for misspecification.
 The null hypotheses of the test is that all exogenous variables are
@@
 \begin{equation}
-  m = \left( \hat{\beta}_2 - \hat{\beta}_3 \right)^{'}
-      \left( \Cov \left[ \hat{\beta}_2 \right] -
-             \Cov \left[ \hat{\beta}_3 \right] \right)
-      \left( \hat{\beta}_2 - \hat{\beta}_3 \right)
+  m = \left( \hat{\beta}_{2SLS} - \hat{\beta}_{3SLS} \right)^{'}
+      \left( \Cov \left[ \hat{\beta}_{2SLS} \right] -
+             \Cov \left[ \hat{\beta}_{3SLS} \right] \right)
+      \left( \hat{\beta}_{2SLS} - \hat{\beta}_{3SLS} \right)
 \label{eq:hausman}
 \end{equation}
-where $\hat{\beta}_2$ and $\Cov \left[ \hat{\beta}_2 \right]$ are the estimated
+where $\hat{\beta}_{2SLS}$ and $\Cov \left[ \hat{\beta}_{2SLS} \right]$ are the estimated
 coefficient and covariance matrix from 2SLS estimation, and
-$\hat{\beta}_3$ and $\Cov \left[ \hat{\beta}_3 \right]$ are the estimated
+$\hat{\beta}_{3SLS}$ and $\Cov \left[ \hat{\beta}_{3SLS} \right]$ are the estimated
 coefficients and covariance matrix from 3SLS estimation.

### 2006-02-25 paper/trunk/systemfit_summary.tex
added "\nopagebreak" to avoid a page break directly after the section header
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@637 591db2bd-e0a0-4d22-8c19-2192dbb7da31
gotta get my changes in...
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@645 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
+\nopagebreak
 In this article, we have described some of the basic features of the
 \pkg{systemfit} package for estimation of linear systems of
-equations with various estimation methods.
+equations.
 Many details of the estimation can be controlled by the user.
@@
 in the near future.
-Some of these are discussed in the following.
+% Some of these are discussed in the following.
 
@@
 The package could be enhanced by the inclusion of methods to fit
-equations with serially correlated and heteroscedastic disturbances.
+equations with serially correlated and heteroscedastic disturbances
+(parks, 1967).
 
 \subsubsection*{Estimation methods}
-We wish to include more sophisticated estimation methods such
-as limited-information maximum likelihood (LIML),
-full-information maximum likelihood (FIML),
-and generalized methods of moments (GMM), in future releases.
+In the future, we wish to include more sophisticated estimation
+methods such as limited-information maximum likelihood (LIML),
+full-information maximum likelihood (FIML), and generalized methods of
+moments (GMM).
 
 \subsubsection*{Non-linear estimation}
-The \pkg{systemfit} package provides a function to estimate
+Finally, the \pkg{systemfit} package provides a function to estimate
 systems of non-linear estimations.
-However, the function \code{nlsystemfit} is currently under development
-and the results are not yet always reliable
-due to convergence problems.
-We hope to improve this behavior soon.
+However, the function \code{nlsystemfit} is currently under
+development and the results are not yet always reliable due to
+convergence difficulties.
+% We hope to improve this behavior soon.
 

### 2006-02-25 paper/trunk/systemfit_summary.tex
added parks citation to the file
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@648 591db2bd-e0a0-4d22-8c19-2192dbb7da31
updated the unbalanced section.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@649 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 For instance,
-forestry datasets typically contain observations of many inexpensive
-observations (stem diameter) and few expensive observations such
-as stem height.
+forestry datasets typically contain many observations of inexpensive
+variables (stem diameter, tree count) and few expensive variables such
+as stem height or volume.
 }
-Simply dropping data points
-that do not contain observations for all equations
-may considerably reduce the number of observations,
-and thus, the information utilized in the estimation.
+Simply dropping data points that do not contain observations for all
+equations may reduce the number of observations considerably, and
+thus, the information utilized in the estimation.
 Hence, it is our intention to include the capability for estimations
-with unbalanced data sets as described in \citet{schmidt77}
-in future releases of \pkg{systemfit}.
+with unbalanced data sets as described in \citet{schmidt77} in future
+releases of \pkg{systemfit}.
 
@@
 equations with serially correlated and heteroscedastic disturbances
-(parks, 1967).
+\citep{parks67}.
 
@@
 convergence difficulties.
-% We hope to improve this behavior soon.
-
-%% testing for the new svn repos
-%% It works great!!!
 

### 2006-02-28 trunk/R/systemfit.R
corrected preparation of residuals if no labels are provided
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@185 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@ -1175,7 +1175,11 @@ coef.summary.systemfit.equation <- function( object, ... ) {
 
 ## return all residuals
 residuals.systemfit <- function( object, ... ) {
-   eqNames <- colnames( object$rcov )
+   if( is.null( colnames( object$rcov ) ) ) {
+      eqNames <- paste( "eq", c( 1:object$g ) )
+   } else {
+      eqNames <- colnames( object$rcov )
+   }
    residuals <- data.frame( obsNo = c( 1:length( object$eq[[1]]$residuals ) ) )
    for( i in 1:object$g ) {
       residuals[[ eqNames[ i ] ]] <- object$eq[[i]]$residuals

### 2006-02-28 paper/trunk/systemfit_other.tex
modified discussion about centering the residuals before calculating its covariance matrix; added description of new argument "CenterResiduals"; modified replication of residual covariance matrix in Greene (2003), Table 14.1 (p. 351) and errata.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@653 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 $( \overline{ \widehat{u} }_i \neq 0 )$.
-Therefore, we subtract the means from the residuals
-and substitute $\widehat{u}_i - \overline{ \widehat{u} }_i$
+Therefore, it might be argued
+that the residual covariance matrix should be calculated
+by subtracting the means from the residuals
+and substituting $\widehat{u}_i - \overline{ \widehat{u} }_i$
 for $\widehat{u}_i$ in (\ref{eq:rcov0}--\ref{eq:rcov2}).

### 2006-03-03 paper/trunk/systemfit_code.tex
fixed quote characters
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@631 591db2bd-e0a0-4d22-8c19-2192dbb7da31
added paragraph in subsection "efficiency of the code" that explains how we compute "X' Omega^{-1}" without computing the (potentially huge) Omega^{-1} matrix
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@654 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 
-To calculate $\Omega^{-1}$ in sections~\ref{sec:Estimation}
-and~\ref{sec:Restrictions} it is not necessary to invert
-the (huge) $\Omega$ matrix,
-because it is sufficient to invert the (small) $\Sigma$ matrix and
-calculate it by $\Omega^{-1} = \Sigma^{-1} \otimes I$
-\citep[p.\ 342]{greene03}.
+The matrix $\Omega^{-1}$ that is used to compute the estimated
+coefficients and their covariance matrix is of size
+$( G \cdot T ) \times ( G \cdot T )$
+(see sections~\ref{sec:Estimation} and~\ref{sec:Restrictions}).
+In case of large data sets, this matrix $\Omega^{-1}$ gets really huge
+and needs a lot of memory.
+Therefore, we use the following transformation and compute $X' \Omega^{-1}$
+by deviding the $X$ matrix into submatrices,
+doing some calculations with these submatrices,
+adding up some of these submatrices, and
+finally putting the submatrices together:
+\begin{equation}
+X' \Omega^{-1}
+%= X' \left( \Sigma^{-1} \otimes I \right)
+= \sum_{i=1} \left[
+   {X^1}' \sigma^{1i} \quad {X^2}' \sigma^{2i}
+   \quad \cdots \quad  \quad {X^G}' \sigma^{Gi}
+\right]
+\end{equation}
+where $\sigma^{ij}$ are the elements of the matrix $\Sigma^{-1}$,
+and $X^i$ is a submatrix of $X$ that contains the rows
+that belong to the $i$'s equation.

### 2006-03-13 paper/trunk/systemfit_abstract.tex
moved abstract and keywords into a separate file; this makes it possible to include the abstract in the package vignette (which will be created soon) without having a second copy of the abstract
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@656 591db2bd-e0a0-4d22-8c19-2192dbb7da31

--- 2006-02-25 paper/trunk/systemfit.tex
@@
-%       $Id$
-
-\documentclass[article]{jss}
-\usepackage[pdftex]{thumbpdf}
-%   run: pdflatex *.tex -> thumbpdf *.tex -> pdflatex *.tex
-\usepackage{csquotes}
-\MakeOuterQuote{°}
-\usepackage{amsmath}
-\usepackage{float}
-
-\usepackage{lineno}
-
-\linenumbers
-
-\newcommand{\Cov}{\text{Cov}}
-\clubpenalty=10000
-\widowpenalty=10000
-
-\hyphenation{sys-tem-fit}
-
-\author{Arne Henningsen\\University of Kiel\And
-        Jeff D. Hamann\\ Forest Informatics, Inc. }
-
-\title{systemfit: A Package to Estimate\\
-       Simultaneous Equation Systems in \proglang{R}}
-
-%% for pretty printing and a nice hypersummary also set:
-\Plainauthor{Arne Henningsen, Jeff D.\ Hamann} %% comma-separated
-\Plaintitle{systemfit: A Package to Estimate Simultaneous Equation Systems 
-            in R} %% without formatting
-\Shorttitle{systemfit} %% a short title (if necessary)
-
-%% an abstract and keywords
 \Abstract{
@@
   seemingly unrelated regression, two-stage least squares,
-  three-stage least squares, parameter restriction}
+  three-stage least squares}
 
 %% at least one keyword must be supplied
-
-%% publication information
-%% NOTE: This needs to filled out ONLY IF THE PAPER WAS ACCEPTED.
-%% If it was not (yet) accepted, leave them commented.
-%% \Volume{13}
-%% \Issue{9}
-%% \Month{September}
-%% \Year{2004}
-%% \Submitdate{2004-09-29}
-%% \Acceptdate{2004-09-29}
-
-%% The address of (at least) one author should be given
-%% in the following format:
-\Address{
-   Arne Henningsen\\
-   Department of Agricultural Economics\\
-   University of Kiel\\
-   D-24098 Kiel, Germany\\
-   E-mail: \email{ahenningsen@agric-econ.uni-kiel.de}\\
-   URL: \url{http://www.uni-kiel.de/agrarpol/ahenningsen/}\\
-   \\
-   Jeff D.\ Hamann\\
-   Forest Informatics, Inc.\\
-   PO Box 1421\\
-   Corvallis, Oregon 97339-1421\\
-   e-mail: \email{jeff.hamann@forestinformatics.com}\\
-   URL: \url{http://www.forestinformatics.com}\\
-}
-%% It is also possible to add a telephone and fax number
-%% before the e-mail in the following format:
-%% Telephone: +43/1/31336-5053
-%% Fax: +43/1/31336-734
-
-%% for those who use Sweave please include the following line (with % symbols):
-%% need no \usepackage{Sweave.sty}
-
-%% end of declarations %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{document}
-
-\input{systemfit_intro.tex}
-\input{systemfit_statistics.tex}
-\input{systemfit_code.tex}
-\input{systemfit_usage.tex}
-\input{systemfit_reliability.tex}
-\input{systemfit_summary.tex}
-
-% \section{CVS Revision Log}
-% \begin{verbatim}
-% $Log$
-% Revision 1.19  2005/11/08 06:31:48  hamannj
-% made seemingly massive changes... but hopefully not for the worse...
-%
-% Revision 1.18  2005/11/07 14:08:56  henningsena
-% a few minor changes (predominantly adding a "\ " after a period (.) that does not indicate an end of a sentence)
-%
-% Revision 1.17  2005/11/07 10:21:59  henningsena
-% just changed some line breaks (deleted new ones and restored deleted ones) to make it easier to find the changes that Jeff committed in the last revision(s)
-%
-% Revision 1.16  2005/11/02 18:31:45  hamannj
-% small last changes
-%
-% Revision 1.15  2005/11/02 18:15:40  hamannj
-% small changes
-
-% Revision 1.14  2005/10/26 21:16:53  hamannj
-% checked out and updated local copy before edits
-
-% Revision 1.13  2005/09/26 20:34:12  henningsena
-% using now package csquotes to have uniform and correct quotation marks (BibTeX-Style jss.bst has to be modified)
-
-% Revision 1.12  2005/09/26 11:24:05  henningsena
-% some minor changes
-
-% Revision 1.11  2005/09/01 12:27:55  henningsena
-% revised structure of the document (sections, subsections, subsubsections) due to removal of nlsystemfit; moved each section into a separate file.
-
-% Revision 1.10  2005/09/01 11:49:02  henningsena
-% moved section about non-linear estimation from systemfit.tex into new file nlsystemfit.tex, because this section should not be included in (this version of) the article
-
-% Revision 1.9  2004/11/16 10:59:09  henningsena
-% some minor changes and some comments on nlsystemfit
-
-% Revision 1.8  2004/11/16 08:55:53  henningsena
-% merged file and resolved conflict due to some minor changes I did while Jeff was adding nlsystemfit.
-
-% Revision 1.7  2004/11/16 05:34:47  hamannj
-% question about nlsystemfit code section (example)
-
-% Revision 1.6  2004/11/16 05:32:46  hamannj
-% made small changes because my code section wouldn't compile. argh.
-
-% Revision 1.5  2004/11/16 05:31:19  hamannj
-% added cvs header and log section. to be removed when submitted.
-
-% \end{verbatim}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\bibliography{systemfit} % a subset of my big bibtex file
-%\bibliography{/home/suapm095/Documents/Literatur/arne}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\end{document}

### 2006-03-15 paper/tags/JSS_20060315/systemfit_summary.tex
created tag for version submitted to the JSS on March 15, 2006
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@659 591db2bd-e0a0-4d22-8c19-2192dbb7da31

--- 2006-02-25 paper/trunk/systemfit_summary.tex
@@
 methods such as limited-information maximum likelihood (LIML),
-full-information maximum likelihood (FIML), and generalized methods of
-moments (GMM).
+full-information maximum likelihood (FIML), generalized methods of
+moments (GMM) and spatial econometric methods.
 

### 2006-03-15 tags/JSS_20060315/tests/test_2sls.R
created tag for version submitted to the JSS on March 15, 2006
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@200 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@ -1,0 +1,250 @@
+
+library( systemfit )
+data( "Kmenta" )
+
+demand <- consump ~ price + income
+supply <- consump ~ price + farmPrice + trend
+inst <- ~ income + farmPrice + trend
+inst1  <- ~ income + farmPrice
+instlist <- list( inst1, inst )
+labels <- list( "demand", "supply" )
+system <- list( demand, supply )
+restrm <- matrix(0,1,7)  # restriction matrix "R"
+restrm[1,3] <-  1
+restrm[1,7] <- -1
+restr2m <- matrix(0,2,7)  # restriction matrix "R" 2
+restr2q <- matrix(0,2,1)  # restriction vector "q" 2
+restr2m[1,3] <-  1
+restr2m[1,7] <- -1
+restr2m[2,2] <- -1
+restr2m[2,5] <-  1
+restr2q[2,1] <-  0.5
+tc <- matrix(0,7,6)
+tc[1,1] <- 1
+tc[2,2] <- 1
+tc[3,3] <- 1
+tc[4,4] <- 1
+tc[5,5] <- 1
+tc[6,6] <- 1
+tc[7,3] <- 1
+restr3m <- matrix(0,1,6)  # restriction matrix "R" 2
+restr3q <- matrix(0,1,1)  # restriction vector "q" 2
+restr3m[1,2] <- -1
+restr3m[1,5] <-  1
+restr3q[1,1] <-  0.5
+
+# It is not possible to estimate 2SLS with systemfit exactly
+# as EViews does, because EViews uses
+# rcovformula == 1 for the coefficient covariance matrix and
+# rcovformula == 0 for the residual covariance matrix.
+# systemfit uses always the same formulas for both calculations.
+
+## *************** 2SLS estimation ************************
+## ************ 2SLS estimation (default)*********************
+fit2sls1 <- systemfit( "2SLS", system, labels, data = Kmenta, inst = inst )
+print( summary( fit2sls1 ) )
+print( round( fit2sls1$bcov, digits = 6 ) )
+
+## *************** 2SLS estimation (single.eq.sigma=F)*******************
+fit2sls1s <- systemfit( "2SLS", system, labels, data = Kmenta, inst = inst,
+   single.eq.sigma = FALSE )
+print( summary( fit2sls1s ) )
+print( round( fit2sls1s$bcov, digits = 6 ) )
+
+## ********************* 2SLS (probdfsys = TRUE) *****************
[…]

### 2006-03-15 tags/JSS_20060315/tests/test_ols.R

@@ -1,0 +1,209 @@
+
+library( systemfit )
+data( "Kmenta" )
+
+demand <- consump ~ price + income
+supply <- consump ~ price + farmPrice + trend
+labels <- list( "demand", "supply" )
+system <- list( demand, supply )
+restrm <- matrix(0,1,7)  # restriction matrix "R"
+restrm[1,3] <-  1
+restrm[1,7] <- -1
+restr2m <- matrix(0,2,7)  # restriction matrix "R" 2
+restr2q <- matrix(0,2,1)  # restriction vector "q" 2
+restr2m[1,3] <-  1
+restr2m[1,7] <- -1
+restr2m[2,2] <- -1
+restr2m[2,5] <-  1
+restr2q[2,1] <-  0.5
+tc <- matrix(0,7,6)
+tc[1,1] <- 1
+tc[2,2] <- 1
+tc[3,3] <- 1
+tc[4,4] <- 1
+tc[5,5] <- 1
+tc[6,6] <- 1
+tc[7,3] <- 1
+restr3m <- matrix(0,1,6)  # restriction matrix "R" 2
+restr3q <- matrix(0,1,1)  # restriction vector "q" 2
+restr3m[1,2] <- -1
+restr3m[1,5] <-  1
+restr3q[1,1] <-  0.5
+
+# It is not possible to estimate OLS with systemfit
+# exactly as EViews does, because EViews uses
+# rcovformula == 1 for the coefficient covariance matrix and
+# rcovformula == 0 for the residual covariance matrix, while
+# systemfit uses always the same formulas for both calculations.
+
+## *************** OLS estimation ************************
+## ********** OLS estimation (default) ********************
+fitols1 <- systemfit( "OLS", system, labels, data = Kmenta )
+print( summary( fitols1 ) )
+print( round( fitols1$bcov, digits = 6 ) )
+
+## ********** OLS estimation (no single.eq sigma=F) ******************
+fitols1s <- systemfit( "OLS", system, labels, data = Kmenta,
+   single.eq.sigma = FALSE )
+print( summary( fitols1s ) )
+print( round( fitols1s$bcov, digits = 6 ) )
+
+## ****************  OLS (probdfsys=T) ***********************
+fitols1p <- systemfit( "OLS", system, labels, data = Kmenta,
+   probdfsys = TRUE )
+print( summary( fitols1p ) )
+print( round( fitols1p$bcov, digits = 6 ) )
+
+## ****************  OLS (rcovformula=0) ***********************
[…]

### 2006-03-15 tags/JSS_20060315/tests/test_wls.R

@@ -1,0 +1,150 @@
+
+library( systemfit )
+data( "Kmenta" )
+
+demand <- consump ~ price + income
+supply <- consump ~ price + farmPrice + trend
+labels <- list( "demand", "supply" )
+system <- list( demand, supply )
+restrm <- matrix(0,1,7)  # restriction matrix "R"
+restrm[1,3] <-  1
+restrm[1,7] <- -1
+restr2m <- matrix(0,2,7)  # restriction matrix "R" 2
+restr2q <- matrix(0,2,1)  # restriction vector "q" 2
+restr2m[1,3] <-  1
+restr2m[1,7] <- -1
+restr2m[2,2] <- -1
+restr2m[2,5] <-  1
+restr2q[2,1] <-  0.5
+tc <- matrix(0,7,6)
+tc[1,1] <- 1
+tc[2,2] <- 1
+tc[3,3] <- 1
+tc[4,4] <- 1
+tc[5,5] <- 1
+tc[6,6] <- 1
+tc[7,3] <- 1
+restr3m <- matrix(0,1,6)  # restriction matrix "R" 2
+restr3q <- matrix(0,1,1)  # restriction vector "q" 2
+restr3m[1,2] <- -1
+restr3m[1,5] <-  1
+restr3q[1,1] <-  0.5
+
+
+## *************** WLS estimation ************************
+fitwls1 <- systemfit( "WLS", system, labels, data = Kmenta )
+print( summary( fitwls1 ) )
+print( round( fitwls1$bcov, digits = 6 ) )
+
+## *************** WLS estimation (EViews-like) ************************
+fitwls1e <- systemfit( "WLS", system, labels, data = Kmenta, rcovformula = 0,
+   probdfsys = TRUE )
+print( summary( fitwls1e ) )
+print( round( fitwls1e$bcov, digits = 6 ) )
+
+## ************** WLS with cross-equation restriction ***************
+fitwls2 <- systemfit( "WLS", system, labels, data = Kmenta, R.restr = restrm )
+print( summary( fitwls2 ) )
+print( round( fitwls2$bcov, digits = 6 ) )
+
+## ************** WLS with cross-equation restriction (EViews-like) *******
+fitwls2e <- systemfit( "WLS", system, labels, data = Kmenta, R.restr = restrm,
+   rcovformula = 0 )
+print( summary( fitwls2e ) )
+print( round( fitwls2e$bcov, digits = 6 ) )
+
+## ******* WLS with cross-equation restriction via TX **********
[…]

### 2006-04-28 paper/trunk/systemfit_summary.tex
several small mainly cosmetic changes (e.g. dashes)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@665 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 In the future, we wish to include more sophisticated estimation
-methods such as limited-information maximum likelihood (LIML),
-full-information maximum likelihood (FIML), and generalized methods of
-moments (GMM).
+methods such as limited information maximum likelihood (LIML),
+full information maximum likelihood (FIML), generalized methods of
+moments (GMM) and spatial econometric methods.
 

### 2006-05-11 paper/tags/arne_diss_submit/systemfit_code.tex
created tag for the revision that is included in Arnes dissertation
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@667 591db2bd-e0a0-4d22-8c19-2192dbb7da31

--- 2006-03-03 paper/trunk/systemfit_code.tex
@@
 (see section~\ref{sec:residcov}),
-the degrees of freedom for the t-tests
+the degrees of freedom for the $t$ tests
 (see section~\ref{sec:degreesOfFreedom}), or
@@
 All functions return the empirical test statistic,
-the degree(s) of freedom, and the p-value.
+the degree(s) of freedom, and the $p$ value.
 
@@
 that is post-multiplied by a vector,
-we use \code{solve(A, b)} instead of \code{solve(A) \%*\% b}.
+we use \code{solve(A,b)} instead of \code{solve(A) \%*\% b}.
 Second, we calculate crossproducts
-by \code{crossprod(X)} or \code{crossprod(X, y)}
+by \code{crossprod(X)} or \code{crossprod(X,y)}
 instead of \code{t(X) \%*\% X} or \code{t(X) \%*\% y},
@@
 Therefore, we use the following transformation and compute $X' \Omega^{-1}$
-by deviding the $X$ matrix into submatrices,
+by dividing the $X$ matrix into submatrices,
 doing some calculations with these submatrices,
@@
 %= X' \left( \Sigma^{-1} \otimes I \right)
-= \sum_{i=1} \left[
-   {X^1}' \sigma^{1i} \quad {X^2}' \sigma^{2i}
-   \quad \cdots \quad  \quad {X^G}' \sigma^{Gi}
-\right]
+= \sum_{i=1} \left[ \begin{array}{c}
+   \sigma^{1i} {X^1} \\
+   \sigma^{2i} {X^2} \\
+   \vdots \\
+   \sigma^{Gi} {X^G} \\
+\end{array} \right]'
 \end{equation}

### 2006-05-11 paper/tags/arne_diss_submit/systemfit_intro.tex
created tag for the revision that is included in Arnes dissertation
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@667 591db2bd-e0a0-4d22-8c19-2192dbb7da31

--- 2006-02-25 paper/trunk/systemfit_intro.tex
@@
 This estimation procedure is generally called
-"Seemingly Unrelated Regression" (SUR) \citep{zellner62}.
+°Seemingly Unrelated Regression° (SUR) \citep{zellner62}.
 Another reason to estimate an equation system simultaneously are
@@
 This simultaneity bias can be corrected for in each equation by
-applying a "Two-Stage Least Squares" (2SLS) method or for all
+applying a °Two-Stage Least Squares° (2SLS) method or for all
 equations simultaneously when combined with SUR resulting in a
-"Three-Stage Least Squares" (3SLS) estimation of the system of
+°Three-Stage Least Squares° (3SLS) estimation of the system of
 equations.

### 2006-05-11 paper/tags/arne_diss_submit/systemfit_other.tex
created tag for the revision that is included in Arnes dissertation
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@667 591db2bd-e0a0-4d22-8c19-2192dbb7da31

--- 2006-02-28 paper/trunk/systemfit_other.tex
@@
 
-The WLS, SUR, W2SLS and 3SLS parameter estimates are consistent,
+The WLS, SUR, W2SLS and 3SLS parameter estimates are consistent
 if the estimated residual covariance matrix is calculated
@@
 To our knowledge the question about how to determine the degrees
-of freedom for single-parameter t-tests is not comprehensively
+of freedom for single-parameter $t$ tests is not comprehensively
 discussed in the literature.
@@
 are used.
-Asymptotically, this distinction doesn't make a difference.
+Asymptotically, this distinction does not make a difference.
 However, in many empirical applications, the number of observations
@@
 each single equation,
-because this yields the same p-values as the equation-wise
+because this yields the same $p$ values as the equation-wise
 OLS estimation.
@@
 Therefore, in this case, it seems to be reasonable to use the degrees
-of freedom of whole system.
+of freedom of the entire system.
 
@@
 
-Linear restrictions can be tested by an F test, Wald test or
-likelihood-ratio (LR) test.
+Linear restrictions can be tested by an $F$ test, Wald test or
+likelihood ratio (LR) test.
 
-The F-statistic for systems of equations is
+The $F$ statistic for systems of equations is
 \begin{equation}
@@
 used in the estimation.
-Under the null hypothesis, $F$ has an F-distribution
+Under the null hypothesis, $F$ has an $F$ distribution
 with $j$ and $M \cdot T - K$ degrees of freedom
@@
 
-The Wald-statistic for systems of equations is
+The Wald statistic for systems of equations is
 \begin{equation}
@@
 
-The LR-statistic for systems of equations is
+The LR statistic for systems of equations is
 \begin{equation}
@@
 \citet{hausman78} developed a test for misspecification.
-The null hypotheses of the test is that all exogenous variables are
+The null hypothesis of the test is that all exogenous variables are
 uncorrelated with all disturbance terms.
@@
 coefficients and covariance matrix from 3SLS estimation.
-Under the null hypotheses this test statistic has a
+Under the null hypothesis this test statistic has a
 $\chi^2$ distribution with degrees of freedom equal to the number of

### 2006-05-11 paper/tags/arne_diss_submit/systemfit_restrictions.tex
created tag for the revision that is included in Arnes dissertation
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@667 591db2bd-e0a0-4d22-8c19-2192dbb7da31

--- 2006-02-19 paper/trunk/systemfit_restrictions.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Imposing linear restrictions}\label{sec:Restrictions}
+\subsection{Imposing linear restrictions}\label{sec:Restrictions}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
+It is common to perform hypothesis tests by imposing restrictions on
+the parameter estimates.
 There are two ways to impose linear parameter restrictions.
@@
 To impose these restrictions, the $X$ matrix is
-(post-)multiplied by this $T$ matrix.
+(post-)\hspace{0pt}multiplied by this $T$ matrix.
 \begin{equation}
@@
 (e.g.\ (near) singularity of the matrices to be inverted),
-while the first doesn't.
+while the first does not.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Restricted OLS estimation}
-\label{sec:olsR}
+\subsubsection{Restricted OLS estimation}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Restricted WLS estimation}
+\subsubsection{Restricted WLS estimation}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Restricted SUR estimation}
+\subsubsection{Restricted SUR estimation}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Restricted 2SLS estimation}
+\subsubsection{Restricted 2SLS estimation}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Restricted W2SLS estimation}
+\subsubsection{Restricted W2SLS estimation}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Restricted 3SLS estimation}
-\label{sec:3slsR}
+\subsubsection{Restricted 3SLS estimation}
 

### 2006-05-11 paper/tags/arne_diss_submit/runSweave.sh

@@ -1,0 +1,4 @@
+#!/bin/sh
+echo "Sweave(\"systemfit_usage.Snw\")" | R --no-save --no-restore
+
+echo "Sweave(\"systemfit_reliability.Snw\")" | R --no-save --no-restore

### 2006-05-22 paper/branches/useR2006/Poster/systemfit_poster.tex
added (draft of) poster for useR! 2006
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@668 591db2bd-e0a0-4d22-8c19-2192dbb7da31

\documentclass[portrait,a0b,final]{a0poster}
\input{systemfit_poster_commands}

\begin{document}
% set background color (gradient)
\psframe[fillstyle=gradient, gradbegin=white, gradend=whiteblue,
   gradmidpoint=1.0](-3,3)(1.1\textwidth,-1.1\textheight)

\begin{center}
% \fbox{
\begin{postercolumn}{1}
% main header
\posterbox{0.96\textwidth}{}{linewidth=2mm, framearc=0.3,
   linecolor=lightblue, fillstyle=gradient, gradangle=0,
   gradbegin=white, gradend=whiteblue, gradmidpoint=1.0,
   framesep=1em}{
% left useR logo
\begin{minipage}[c][9cm][c]{0.2\textwidth}
  \begin{center}
    \includegraphics[width=12cm]{useR-large}
  \end{center}
\end{minipage}
% title etc
\begin{minipage}[c][13cm][c]{0.59\textwidth}
  \begin{center}
    {\VeryHuge \textbf{systemfit}}\\[12mm]
    {\Huge Simultaneous Equation Systems in R}\\[8mm]
    {\Huge Arne Henningsen and Jeff D.\ Hamann}\\[8mm]
    {\LARGE University of Kiel (Germany) and Forest Informatics, Inc. (Corvallis, OR, USA)}
  \end{center}
\end{minipage}
% right useR logo
\begin{minipage}[c][9cm][c]{0.2\textwidth}
  \begin{center}
    \includegraphics[width=12cm]{useR-large}
  \end{center}
\end{minipage}
}
\end{postercolumn}

\vspace{15mm}

% \fbox{
\begin{postercolumn}{0.49}
% Features
\posterbox{0.92\textwidth}{}{linewidth=2mm, framearc=0.1,
   linecolor=lightblue, fillstyle=gradient, gradangle=0,
   gradbegin=white, gradend=whiteblue, gradmidpoint=1.0,
   framesep=1em}{
{\boxheaderfontsize \textbf{Features}}
{\standardfontsize
\begin{itemize}
\item several estimation methods, e.g.:
\begin{itemize}
\item Ordinary Least Squares (OLS)
\item Two-Stage Least Squares (2SLS)
\item Seemingly Unrelated Regression (SUR)
\item Three-Stage Least Squares (3SLS)
\end{itemize}
\item imposition of linear restrictions by two different methods
\item instrumental variables: the same for all equations or different for each equation
\item several different formulas to calculate the residual covariance matrix
\item several different formulas for the 3SLS estimation
\item $t$ tests: degrees of freedom of the whole system
or degrees of freedom of the single equation
\item OLS/2SLS estimation: different $\sigma^2$s for each single equation
or the same $\sigma^2$ for all equations
\item wrapper function for (classical) panel-like data in long format
(\texttt{systemfitClassic})
\item testing linear hypothesis using the F-, Wald-, and LR-statistic
\item Hausman test for the consistency of the 3SLS estimator
\end{itemize}
}}

\end{postercolumn}
% }
\hfill
% \fbox{
\begin{postercolumn}{0.49}
% Advantages
\posterbox{0.92\textwidth}{}{linewidth=2mm, framearc=0.1,
   linecolor=lightblue, fillstyle=gradient, gradangle=0,
   gradbegin=white, gradend=whiteblue, gradmidpoint=1.0,
   framesep=1em}{
{\boxheaderfontsize \textbf{Advantages of \texttt{systemfit}}}
{\standardfontsize
\begin{itemize}
\item all estimation procedures are publicly available in the source code
\item estimation algorithms can be easily modified to meet specific
requirements
\item the (advanced) user can control many estimation details generally
not available in other software packages
\item novices can safely ignore these estimation details
because they have reasonable defaults
\item reliability of \texttt{systemfit} has been demonstrated by comparing its estimation
results with results published in the literature
\end{itemize}
}}

\vspace{15mm}

% Future
\posterbox{0.92\textwidth}{}{linewidth=2mm, framearc=0.1,
   linecolor=lightblue, fillstyle=gradient, gradangle=0,
   gradbegin=white, gradend=whiteblue, gradmidpoint=1.0,
   framesep=1em}{
{\boxheaderfontsize \textbf{Plans for the Future}}
{\standardfontsize
\begin{itemize}
\item estimation with unbalanced data sets
\item estimation methods: LIML, FIML,  and GMM
\item fitting equation systems with serially correlated and
heteroscedastic disturbances
\item spatial econometric methods
\item improving the function \texttt{nlsystemfit}
to estimate systems of non-linear estimations
\end{itemize}
}}

\end{postercolumn}
% }

\vspace{15mm}

\begin{postercolumn}{1}
\posterbox{0.96\textwidth}{}{linewidth=2mm, framearc=0.3,
   linecolor=lightblue, fillstyle=gradient, gradangle=0,
   gradbegin=white, gradend=whiteblue, gradmidpoint=1.0,
   framesep=1em}{
\begin{spacing}{1.5}
{\Large
\textbf{Arne Henningsen} $\cdot$
Department of Agricultural Economics $\cdot$
University of Kiel $\cdot$
Olshausenstr.~40 $\cdot$
D-24098 Kiel (Germany) $\cdot$
Phone~+49-431-880-4445 $\cdot$
Fax~+49-431-880-1397 $\cdot$
\url{ahenningsen@agric-econ.uni-kiel.de} $\cdot$
\url{http://www.uni-kiel.de/agrarpol/ahenningsen/}\\[5mm]
\textbf{Jeff D. Hamann} $\cdot$
Forest Informatics, Inc. $\cdot$
PO Box 1421 $\cdot$
Corvallis, Oregon USA 97339-1421 $\cdot$
Phone~+1-541-754-1428 $\cdot$
Fax~+1-541-752-0288 $\cdot$
\url{jeff.hamann@forestinformatics.com} $\cdot$
\url{http://www.forestinformatics.com}
}
\end{spacing}
}
\end{postercolumn}
% }
\end{center}
\end{document}

### 2006-06-05 paper/branches/useR2006/Poster/systemfit_poster.tex
Poster for useR! 2006: added box "Motivation"
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@669 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 \begin{postercolumn}{0.49}
+% Motivation
+\posterbox{0.92\textwidth}{}{linewidth=2mm, framearc=0.1,
+   linecolor=lightblue, fillstyle=gradient, gradangle=0,
+   gradbegin=white, gradend=whiteblue, gradmidpoint=1.0,
+   framesep=1em}{
+{\boxheaderfontsize \textbf{Motivation}}
+{\standardfontsize
+\begin{itemize}
+\item many theoretical models consist of more than one equation\\
+$\Rightarrow$ contemporaneously correlation of disturbance terms (likely)\\
+$\Rightarrow$ \parbox[t]{0.95\textwidth}{\raggedright
+   simultaneously estimation of all equations as 
+   °Seemingly Unrelated Regression° (SUR) leads to efficient results}
+\mbox{}
+\item theoretically derived cross-equation parameter restrictions\\
+$\Rightarrow$ simultaneous estimation of all equations required
+\item endogeneity of some variables\\
+$\Rightarrow$ \parbox[t]{0.95\textwidth}{\raggedright
+   estimation using °Two-Stage Least Squares° (2SLS)
+   or °Three-Stage Least Squares° (3SLS) required}
+\end{itemize}
+}}
+
+\vspace{15mm}
+
 % Features
@@
    framesep=1em}{
-{\boxheaderfontsize \textbf{Features}}
+{\boxheaderfontsize \textbf{Features of \texttt{systemfit}}}
 {\standardfontsize

### 2006-06-05 paper/branches/useR2006/Poster/systemfit_poster.tex
Poster for useR! 2006: added box "Example"
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@670 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 \begin{postercolumn}{0.49}
+
+\begin{SaveVerbatim}{specSystem}
+eqDemand <- consump ~ price + income
+eqSupply <- consump ~ price + farmPrice + trend
+system <- list(demand=eqDemand, supply=eqSupply)
+\end{SaveVerbatim}
+% \end{Verbatim}
+\begin{SaveVerbatim}{estSystem}
+fitsur <- systemfit( "SUR", system )
+\end{SaveVerbatim}
+% \end{Verbatim}
+\begin{SaveVerbatim}{printResults}
+summary( fitsur )
+\end{SaveVerbatim}
+% \end{Verbatim}
+
+% Example
+\posterbox{0.92\textwidth}{}{linewidth=2mm, framearc=0.1,
+   linecolor=lightblue, fillstyle=gradient, gradangle=0,
+   gradbegin=white, gradend=whiteblue, gradmidpoint=1.0,
+   framesep=1em}{
+{\boxheaderfontsize \textbf{Example}}
+{\standardfontsize
+\begin{itemize}
+\item from Kmenta (1986): Elements of Econometrics, p.~685
+\item specification of the equation system:
+   \bigskip\\
+   \BUseVerbatim{specSystem}
+\item estimation using method °SUR°:\\
+   \BUseVerbatim{estSystem}
+\item printing summary results:\\
+   \BUseVerbatim{printResults}
+\end{itemize}
+}}
+
+\vspace{15mm}
+
 % Advantages

### 2006-06-06 paper/branches/useR2006/Poster/systemfit_poster.tex
Poster for useR! 2006: improved placement and sizes of boxes
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@671 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
-\documentclass[portrait,a0b,final]{a0poster}
+\documentclass[portrait,a0b]{a0poster}
 \input{systemfit_poster_commands}
@@
 % Motivation
-\posterbox{0.92\textwidth}{}{linewidth=2mm, framearc=0.1,
+\posterbox{0.92\textwidth}{245mm}{linewidth=2mm, framearc=0.1,
    linecolor=lightblue, fillstyle=gradient, gradangle=0,
@@
 % Example
-\posterbox{0.92\textwidth}{}{linewidth=2mm, framearc=0.1,
+\posterbox{0.92\textwidth}{245mm}{linewidth=2mm, framearc=0.1,
    linecolor=lightblue, fillstyle=gradient, gradangle=0,
@@
 \end{itemize}
+\vspace{4mm}
 }}

### 2006-06-08 paper/branches/useR2006/Poster/systemfit_poster.tex
Poster for useR! 2006: a few minor corrections (this versions will be presented at useR! 2006)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@672 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 \item many theoretical models consist of more than one equation\\
-$\Rightarrow$ contemporaneously correlation of disturbance terms (likely)\\
+$\Rightarrow$ contemporaneous correlation of disturbance terms (likely)\\
 $\Rightarrow$ \parbox[t]{0.95\textwidth}{\raggedright
-   simultaneously estimation of all equations as 
+   simultaneous estimation of all equations as
    °Seemingly Unrelated Regression° (SUR) leads to efficient results}
@@
 (\texttt{systemfitClassic})
-\item testing linear hypothesis using the F-, Wald-, and LR-statistic
+\item testing linear hypotheses using the F-, Wald-, and LR-statistic
 \item Hausman test for the consistency of the 3SLS estimator

### 2006-06-09 paper/branches/useR2006/Slides/systemfit_slides.tex
useR! 2006: added first draft of slides
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@673 591db2bd-e0a0-4d22-8c19-2192dbb7da31

\documentclass{beamer}
%\documentclass[notes=show,handout]{beamer}
%\documentclass[handout]{beamer}

\input{systemfit_slides_commands.tex}

\title[systemfit]{
   \textbf{systemfit}\\ Simultaneous Equation Systems in R}
\author{Arne Henningsen and Jeff D.\ Hamann}
\date{University of Kiel (Germany) and\\Forest Informatics, Inc. (Corvallis, OR, USA)}

\begin{document}
\frame{ \titlepage }
\note{
\begin{itemize}
\item Welcome!
\item authors
\item package systemfit: Estimating simultaneous equation systems in R
\end{itemize}
}

% ============== Introduction ==============================
\section{Introduction}

% ----------------- Motivation ----------------------------
\subsection{Motivation}
\frame{
   \frametitle{Motivation}
\begin{itemize}
\item many theoretical models consist of more than one equation
   \begin{itemize}
   \item contemporaneous correlation of disturbance terms (likely)
   \item simultaneous estimation of all equations as
      °Seemingly Unrelated Regression° (SUR) leads to efficient results
   \end{itemize}
\item theoretically derived cross-equation parameter restrictions\\
   \begin{itemize}
   \item simultaneous estimation of all equations required
   \end{itemize}
\item endogeneity of some variables\\
   \begin{itemize}
   \item estimation using °Two-Stage Least Squares° (2SLS)
      or °Three-Stage Least Squares° (3SLS) required
   \end{itemize}
\end{itemize}
}
\note{
\begin{itemize}
\item ?
\end{itemize}
}

% ----------------- Outline ----------------------------
\subsection{Outline}
\frame{
   \frametitle{Outline}
\begin{itemize}
\item Introduction
\item Features of systemfit
\item Examples
\item Advantages
\item Plans for the Future
\end{itemize}
}
\note{
\begin{itemize}
\item ?
\end{itemize}
}

% ============== Features ==============================
\section{Features}

% ----------------- Estimation Methods ----------------------------
\subsection{Methods}
\frame{
   \frametitle{Estimation Methods}
\begin{itemize}
\item Ordinary Least Squares (OLS)
\item Two-Stage Least Squares (2SLS)
\item Seemingly Unrelated Regression (SUR)
\item Three-Stage Least Squares (3SLS)
\item \ldots
\end{itemize}
}
\note{
\begin{itemize}
\item 2SLS often used for single-equations
\end{itemize}
}

% ----------------- Estimation Control ----------------------------
\subsection{Estimation Control}
\frame{
   \frametitle{Estimation Control}
\begin{itemize}
\item imposition of linear restrictions
\item instrumental variables
\item formulas for the residual covariance matrix
\item formulas for 3SLS estimation
\item degrees of freedom for $t$ tests
\item homogenous residual variance in OLS/2SLS estimations
\end{itemize}
}
\note{
\begin{itemize}
\item imposition of linear restrictions by two different methods
\item instrumental variables: the same for all equations or different for each equation
\item several different formulas to calculate the residual covariance matrix
\item several different formulas for the 3SLS estimation
\item $t$ tests: degrees of freedom of the whole system
   or degrees of freedom of the single equation
\item OLS/2SLS estimation: different $\sigma^2$s for each single equation
   or the same $\sigma^2$ for all equations
\end{itemize}
}

% ----------------- Other Tools ----------------------------
\subsection{Other Tools}
\frame{
   \frametitle{Other Tools}
\begin{itemize}
\item wrapper function for (classical) panel-like data in long format:
   \texttt{systemfitClassic}
\item testing linear hypotheses using the F-, Wald-, and LR-statistic
\item Hausman test for the consistency of the 3SLS estimator
\end{itemize}
}
\note{
\begin{itemize}
\item ?
\end{itemize}
}

% ======================= Example ======================================
\section{Example}

% ----------------- Commands ----------------------------
\subsection{Commands}
\begin{SaveVerbatim}{specSystem}
eqDemand <- consump ~ price + income
eqSupply <- consump ~ price + farmPrice + trend
system <- list(demand=eqDemand, supply=eqSupply)
\end{SaveVerbatim}
% \end{Verbatim}
\begin{SaveVerbatim}{estSystem}
fitsur <- systemfit( "SUR", system )
\end{SaveVerbatim}
% \end{Verbatim}
\begin{SaveVerbatim}{printResults}
summary( fitsur )
\end{SaveVerbatim}
% \end{Verbatim}
\frame{
   \frametitle{Example: Commands}
\begin{itemize}
\item from Kmenta (1986): Elements of Econometrics, p.~685
\item specification of the equation system:\\[2mm]
   \BUseVerbatim{specSystem}
\item estimation using method °SUR°:
   \BUseVerbatim{estSystem}
\item printing summary results:\\
   \BUseVerbatim{printResults}
\end{itemize}
}
\note{
\begin{itemize}
\item ?
\end{itemize}
}

% ----------------- Output ----------------------------
\subsection{Output}
\begin{SaveVerbatim}{output}
???
\end{SaveVerbatim}
% \end{Verbatim}
\frame{
   \frametitle{Example: Output}
\BUseVerbatim{output}
}
\note{
\begin{itemize}
\item ?
\end{itemize}
}

% ============== Advantages ==============================
\section{Advantages}
\frame{
   \frametitle{Advantages}
\begin{itemize}
\item all estimation procedures are publicly available
\item estimation algorithms can be easily modified
\item many estimation details can be controlled by the user
\item novices can safely ignore these estimation details
\item reliability of \texttt{systemfit} has been demonstrated
\end{itemize}
}
\note{
\begin{itemize}
\item all estimation procedures are publicly available in the source code
\item estimation algorithms can be easily modified to meet specific
requirements
\item the (advanced) user can control many estimation details generally
not available in other software packages
\item novices can safely ignore these estimation details
because they have reasonable defaults
\item reliability of \texttt{systemfit} has been demonstrated by comparing its estimation
results with results published in the literature
\end{itemize}
}

% ============== Future ==============================
\section{Future}

% ----------------- General Plans for the Future ----------------------------
\subsection{General}
\frame{
   \frametitle{Plans for the Future}
\begin{itemize}
\item estimation with unbalanced data sets
\item estimation methods: LIML, FIML,  and GMM
\item fitting equation systems with serially correlated and
heteroscedastic disturbances
\item spatial econometric methods
\item improving the function \texttt{nlsystemfit}
to estimate systems of non-linear estimations
\end{itemize}
}
\note{
\begin{itemize}
\item ?
\end{itemize}
}

% ----------------- Arguments ----------------------------
\subsection{Arguments}
\frame{
   \frametitle{User Interface: Arguments}
Arguments of \texttt{systemfit}:\\[3mm]
\begin{minipage}{0.22\textwidth}
\begin{itemize}
\item method
\item eqns
\item eqnlabels
\item inst
\item data
\item R.restr
\end{itemize}
\end{minipage}
\begin{minipage}{0.32\textwidth}
\begin{itemize}
\item q.restr
\item TX
\item maxiter
\item tol
\item rcovformula
\item centerResiduals
\end{itemize}
\end{minipage}
\begin{minipage}{0.40\textwidth}
\begin{itemize}
\item formula3sls
\item probdfsys
\item single.eq.sigma
\item solvetol
\item saveMemory
\item (more in the future)
\end{itemize}
\end{minipage}\\[5mm]
\textbf{Too many?}
}
\note{
\begin{itemize}
\item ?
\end{itemize}
}

\frame{
   \frametitle{Arguments}
Reducing arguments:\\[3mm]
\begin{minipage}{0.25\textwidth}
\begin{itemize}
\item method
\item eqns
\item inst
\item data
\end{itemize}
\end{minipage}
\begin{minipage}{0.55\textwidth}
\begin{itemize}
\item R.restr
\item q.restr
\item TX
\item \textbf{control} (like in \texttt{optim})
\end{itemize}
\end{minipage}\\[5mm]
However: This would break existing code!
}
\note{
\begin{itemize}
\item ?
\end{itemize}
}

% ==================== The End ===========================
\section{ }
\frame{
   \frametitle{The End . . .}

\begin{center}
{\LARGE \emph{Thank you for your attention!}}
\end{center}
}
\note{
   Thank you for your attention!
}

% ------------ End -------------

\end{document}

### 2006-06-10 paper/branches/useR2006/Slides/systemfit_slides.tex
useR! 2006: added slides showing output of summary.systemfit(); a few minor other modifications
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@674 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 \begin{SaveVerbatim}{estSystem}
-fitsur <- systemfit( "SUR", system )
+fitsur <- systemfit("SUR", system, data=Kmenta)
 \end{SaveVerbatim}
@@
 \item specification of the equation system:\\[2mm]
-   \BUseVerbatim{specSystem}
+   {\small \BUseVerbatim{specSystem}}
 \item estimation using method °SUR°:
-   \BUseVerbatim{estSystem}
+   {\small \BUseVerbatim{estSystem}}
 \item printing summary results:\\
-   \BUseVerbatim{printResults}
+   {\small \BUseVerbatim{printResults}}
 \end{itemize}
@@
 \subsection{Output}
-\begin{SaveVerbatim}{output}
-???
+\begin{SaveVerbatim}{outputSystem}
+systemfit results
+method: SUR
+
+        N DF      SSR     MSE    RMSE       R2   Adj R2
+demand 20 17  65.6829 3.86370 1.96563 0.755019 0.726198
+supply 20 16 104.0584 6.50365 2.55023 0.611888 0.539117
+
+[...]
+
+The correlations of the residuals
+         demand   supply
+demand 1.000000 0.982348
+supply 0.982348 1.000000
+
+The determinant of the residual covariance matrix: 0.879285
+OLS R-squared value of the system: 0.683453
+McElroy's R-squared value for the system: 0.788722
 \end{SaveVerbatim}
@@
 \frame{
-   \frametitle{Example: Output}
-\BUseVerbatim{output}
+   \frametitle{Results of the Entire System}
+{\small \BUseVerbatim{outputSystem}}
+}
+\note{
+\begin{itemize}
+\item ?
+\end{itemize}
+}
+
+\begin{SaveVerbatim}{outputEq}
+SUR estimates for 'demand' (equation 1)
+Model Formula: consump ~ price + income
+
+             Estimate Std. Error   t value Pr(>|t|)
+(Intercept) 99.332894   7.514452 13.218913        0 ***
+price       -0.275486   0.088509 -3.112513 0.006332  **
+income        0.29855   0.041945  7.117605    2e-06 ***
+---
+Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
+
+Residual standard error: 1.96563 on 17 degrees of freedom
+Number of observations: 20 Degrees of Freedom: 17
+SSR: 65.682902 MSE: 3.8637 Root MSE: 1.96563
+Multiple R-Squared: 0.755019 Adjusted R-Squared: 0.726198
+\end{SaveVerbatim}
+% \end{Verbatim}
+\frame{
+   \frametitle{Results of a Single Equation}
+{\small \BUseVerbatim{outputEq}}
 }
@@
    \frametitle{Arguments}
-Reducing arguments:\\[3mm]
+Reducing arguments?\\[3mm]
 \begin{minipage}{0.25\textwidth}

### 2006-06-12 paper/branches/useR2006/Slides/systemfit_slides.tex
useR! 2006 Slides: added notes; some other improvements; removed slide about advantages
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@675 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 \begin{itemize}
-\item Welcome!
-\item authors
-\item package systemfit: Estimating simultaneous equation systems in R
+\item Welcome to my presentation about the R package systemfit.
+\item This package is written by me, Arne Henningsen,
+   and by my co-author Jeff Hamann.
+\item The package can used to estimate simultaneous equation systems in R.
 \end{itemize}
@@
    \frametitle{Motivation}
-\begin{itemize}
-\item many theoretical models consist of more than one equation
+Many theoretical models consist of more than one equation
    \begin{itemize}
@@
    \end{itemize}
-\item theoretically derived cross-equation parameter restrictions\\
+\medskip
+Theoretically derived cross-equation parameter restrictions\\
    \begin{itemize}
@@
    \end{itemize}
-\item endogeneity of some variables\\
+\medskip
+Endogeneity of some variables
    \begin{itemize}
@@
    \end{itemize}
-\end{itemize}
+\bigskip
+$\Rightarrow$ All these models can be estimated by \texttt{systemfit}.
 }
@@
 \begin{itemize}
-\item ?
+\item Many theoretical models that are econometrically estimated
+   consist of more than one equation.
+\item For instance, economists like me model the supply
+   and the demand of a certain good.
+\item Or forest engineers like my co-author Jeff model the growth of trees
+   and take the height as well as the diameter of the trees into account.
+\item The disturbance terms of the single equations of such a equation system
+   are likely contemporaneously correlated.
+\item For instance, if you have a specific tree with a below average
+   diameter growth,
+   the growth of the height will probably be also below average.
+\item A standard OLS regression of each equation separately
+   ignores this correlation and leads to inefficient estimates.
+\item However, efficients estimates can be obtained,
+   if all equations are estimated simultaneously,
+   using Zellner's °Seemingly Unrelated Regression° method.
+\myhrule
+\item Furthermore, when the underlying theory suggests cross-equation
+   parameter restrictions,
+   a simultaneous estimation of all equations required, too.
+\myhrule
+\item Finally, it might be that some variables are the endogenous variable
+    in some equations and a regressor in other equations.
+\item In this case, °Two-Stage Least Squares° or °Three-Stage Least Squares°
+   regressions of the system have to be applied to obtain unbiased estimates.
+\myhrule
+\item All these models can be estimated by \texttt{systemfit}.
 \end{itemize}
@@
 \begin{itemize}
+\setlength{\itemsep}{0.75em}
 \item Introduction
@@
 \item Examples
-\item Advantages
+% \item Advantages
 \item Plans for the Future
@@
 \begin{itemize}
-\item ?
+\item After this introduction, I will present the features of systemfit.
+\item Then, I will demonstrate the usage of systemfit by a simple example.
+% \item Thereafter, I will highlight some advantages of systemfit compared to
+%    other software packages.
+\item Finally, I will present our future plans to extend and improve systemfit.
 \end{itemize}
@@
 \begin{itemize}
+\setlength{\itemsep}{0.75em}
 \item Ordinary Least Squares (OLS)
@@
 \begin{itemize}
-\item 2SLS often used for single-equations
+\item The most important features of \texttt{systemfit}
+   are -- of course -- the estimation methods.
+\item It provides
+   \begin{itemize}
+   \item Ordinary Least Squares
+   \item Two-Stage Least Squares
+   \item Seemingly Unrelated Regression
+   \item Three-Stage Least Squares
+   \item and some other less important methods
+   \end{itemize}
+\item Although the °Two-Stage Least Squares° estimator is designed
+   for estimating equation systems,
+   it can be applied to single equations, too.
+\item I have heard from several people -- among them Jeff Racine --
+   that they generally use systemfit to fit single-equation models
+   by °Two-Stage Least Squares°.
 \end{itemize}
@@
 \begin{itemize}
+\setlength{\itemsep}{0.75em}
 \item imposition of linear restrictions
 \item instrumental variables
+\item iterated estimation
 \item formulas for the residual covariance matrix
 \item formulas for 3SLS estimation
-\item degrees of freedom for $t$ tests
-\item homogenous residual variance in OLS/2SLS estimations
+% \item degrees of freedom for $t$ tests
+% \item homogenous residual variance in OLS/2SLS estimations
+\item \ldots
 \end{itemize}
@@
 \begin{itemize}
-\item imposition of linear restrictions by two different methods
-\item instrumental variables: the same for all equations or different for each equation
-\item several different formulas to calculate the residual covariance matrix
-\item several different formulas for the 3SLS estimation
-\item $t$ tests: degrees of freedom of the whole system
-   or degrees of freedom of the single equation
-\item OLS/2SLS estimation: different $\sigma^2$s for each single equation
-   or the same $\sigma^2$ for all equations
+\item Besides the estimation method,
+   the user can control many details of the estimation.
+\item For instance, she or he can impose linear restrictions
+   by two different methods,
+\item and can specify instrumental variables:
+   either the same for all equations or different for each equation.
+\item Furthermore, it can be indicated whether the feasable least squares
+   estimations should be iterated.
+\item The literature suggests several different formulas to calculate
+   the residual covariance matrix and the Three-Stage Least Squares estimator.
+\item To my knowledge, systemfit is the only popular software
+   where the user can decide which formula is applied.
+\item There exist some further details that can be controlled by the user.
+\item If you are interested in them please don't hesitate to ask me or
+   take a look into the documentation.
 \end{itemize}
@@
 \begin{itemize}
+\setlength{\itemsep}{0.75em}
 \item wrapper function for (classical) panel-like data in long format:
@@
 \begin{itemize}
-\item ?
+\item The package systemfit does not just provide the function systemfit
+   to estimate equation systems,
+   but also a few other functions.
+\item For instance, the package includes the wrapper function systemfitClassic
+   that can be applied to panel-like data in long format.
+\item Then, there are functions to test linear hypotheses using the
+   F-, Wald-, and Likelihood-ratio-statistic
+\item Finally, the consistency of the 3SLS estimator can be tested
+   by a hausman test.
 \end{itemize}
@@
 \begin{itemize}
+\setlength{\itemsep}{0.75em}
 \item from Kmenta (1986): Elements of Econometrics, p.~685
@@
 % ============== Advantages ==============================
-\section{Advantages}
-\frame{
-   \frametitle{Advantages}
-\begin{itemize}
-\item all estimation procedures are publicly available
-\item estimation algorithms can be easily modified
-\item many estimation details can be controlled by the user
-\item novices can safely ignore these estimation details
-\item reliability of \texttt{systemfit} has been demonstrated
-\end{itemize}
-}
-\note{
-\begin{itemize}
-\item all estimation procedures are publicly available in the source code
-\item estimation algorithms can be easily modified to meet specific
-requirements
-\item the (advanced) user can control many estimation details generally
-not available in other software packages
-\item novices can safely ignore these estimation details
-because they have reasonable defaults
-\item reliability of \texttt{systemfit} has been demonstrated by comparing its estimation
-results with results published in the literature
-\end{itemize}
-}
+% \section{Advantages}
+% \frame{
+%    \frametitle{Advantages}
+% \begin{itemize}
+% \item all estimation procedures are publicly available
+% \item estimation algorithms can be easily modified
+% \item many estimation details can be controlled by the user
+% \item novices can safely ignore these estimation details
+% \item reliability of \texttt{systemfit} has been demonstrated
+% \end{itemize}
+% }
+% \note{
+% \begin{itemize}
+% \item all estimation procedures are publicly available in the source code
+% \item estimation algorithms can be easily modified to meet specific
+% requirements
+% \item the (advanced) user can control many estimation details generally
+% not available in other software packages
+% \item novices can safely ignore these estimation details
+% because they have reasonable defaults
+% \item reliability of \texttt{systemfit} has been demonstrated by comparing its estimation
+% results with results published in the literature
+% \end{itemize}
+% }
 
@@
 \begin{itemize}
+\setlength{\itemsep}{0.75em}
 \item estimation with unbalanced data sets

### 2006-06-13 paper/branches/useR2006/Slides/systemfit_slides.tex
useR! 2006 Slides: added some further notes; some minor corrections
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@676 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
-\documentclass{beamer}
-%\documentclass[notes=show,handout]{beamer}
+%\documentclass{beamer}
+\documentclass[notes=show,handout]{beamer}
 %\documentclass[handout]{beamer}
@@
    and by my co-author Jeff Hamann.
-\item The package can used to estimate simultaneous equation systems in R.
+\item The package can be used to estimate simultaneous equation systems in R.
 \end{itemize}
@@
    the growth of the height will probably be also below average.
-\item A standard OLS regression of each equation separately
+\item A standard Ordinary Least Squares regression of each equation separately
    ignores this correlation and leads to inefficient estimates.
 \item However, efficients estimates can be obtained,
-   if all equations are estimated simultaneously,
+   if all equations are estimated simultaneously
    using Zellner's °Seemingly Unrelated Regression° method.
-\myhrule
-\item Furthermore, when the underlying theory suggests cross-equation
+\item Furthermore, when the theory suggests cross-equation
    parameter restrictions,
-   a simultaneous estimation of all equations required, too.
-\myhrule
+   a simultaneous estimation of all equations is required, too.
 \item Finally, it might be that some variables are the endogenous variable
@@
 \item In this case, °Two-Stage Least Squares° or °Three-Stage Least Squares°
-   regressions of the system have to be applied to obtain unbiased estimates.
-\myhrule
+   regressions have to be applied to obtain unbiased estimates.
 \item All these models can be estimated by \texttt{systemfit}.
@@
 %    other software packages.
-\item Finally, I will present our future plans to extend and improve systemfit.
+\item And finally, I will present our future plans to extend and improve systemfit.
 \end{itemize}
@@
 \item Besides the estimation method,
-   the user can control many details of the estimation.
+   the user can control many further details of the estimation.
 \item For instance, she or he can impose linear restrictions
-   by two different methods,
-\item and can specify instrumental variables:
-   either the same for all equations or different for each equation.
-\item Furthermore, it can be indicated whether the feasable least squares
-   estimations should be iterated.
+   by two different methods.
+\item Instrumental variables can be provided
+   either by specifying the same variables for all equations or
+   by specifying different variables for each equation.
+\item Furthermore, it can be indicated whether feasible least squares
+   estimations should be iterated or not.
 \item The literature suggests several different formulas to calculate
    the residual covariance matrix and the Three-Stage Least Squares estimator.
-\item To my knowledge, systemfit is the only popular software
-   where the user can decide which formula is applied.
+\item To my knowledge, systemfit is the only statistical software
+   where the user can decide which formula should be applied.
 \item There exist some further details that can be controlled by the user.
-\item If you are interested in them please don't hesitate to ask me or
+\item If you are interested in these, please don't hesitate to ask me or
    take a look into the documentation.
@@
 \item Then, there are functions to test linear hypotheses using the
-   F-, Wald-, and Likelihood-ratio-statistic
-\item Finally, the consistency of the 3SLS estimator can be tested
-   by a hausman test.
+   F-, Wald-, and Likelihood-ratio-statistic.
+\item Finally, the consistency of the Three-Stage Least Squares estimator
+   can be tested by a Hausman test.
 \end{itemize}
@@
 eqSupply <- consump ~ price + farmPrice + trend
-system <- list(demand=eqDemand, supply=eqSupply)
+eqSystem <- list(demand=eqDemand, supply=eqSupply)
 \end{SaveVerbatim}
@@
 \begin{SaveVerbatim}{estSystem}
-fitsur <- systemfit("SUR", system, data=Kmenta)
+fitsur <- systemfit("SUR", eqSystem, data=Kmenta)
 \end{SaveVerbatim}
@@
 \begin{itemize}
-\item ?
+\item Now, I will shortly demonstrate how to use systemfit.
+\item I have selected an example from Kmenta's textbook °Elements of
+   Econometrics°.
+\item In this example, the demand and the supply of food in the U.S.\ is
+   analyzed.
+\item The demand and the supply equation are specified first.
+\item The consumed food quantity is a function of the food price and
+   the income of the consumers.
+\item The produced quantity equals the consumed quantity and depends
+   on the food price, the average price of farm products
+   and a time trend to account for technical progress.
+\item In the third line, these two equations are combined into a list
+   named °eqSystem°.
+\item Then the equation system is estimated by °Seemingly Unrelated
+   Regression°
+\item Of course, the food price is endogenous and therefore,
+   the system should be estimated by Three-Stage Least Squares
+   using instrumental variables.
+   However, I want to keep this example as simple as possible.
+\item Finally, summary results of the regression are printed.
 \end{itemize}
@@
 \begin{itemize}
-\item ?
+\item First, the summary results of the entire system are presented.
+\item For instance, the number of observations, \texttt{N},
+   the degrees of freedom, \texttt{DF},
+   as well as R-squared and adjusted R-squared values
+   are printed for each equation.
+\item Besides some other results, the correlation matrix of the residuals
+   is presented.
+\item A coefficient of correlation of 0.98 shows that the residuals
+   of the two equations are highly correlated.
+\item This indicates that OLS results that do not take this correlation
+   into account are rather inefficient.
+\item Finally, the determinant of the residual covariance matrix,
+   the OLS R-squared value and McElroy's R-squared value of
+   the equation system are printed.
 \end{itemize}
@@
 \begin{itemize}
-\item ?
+\item Summary results of each equation are printed then.
+\item The results of the demand equation are presented on this slide.
+\item The output is pretty similar to the output of °summary.lm°.
+\item For instance, the estimated coefficients, their standard errors,
+   the t-values, and the marginal levels of significance are shown.
 \end{itemize}
@@
 \item improving the function \texttt{nlsystemfit}
-to estimate systems of non-linear estimations
+to estimate systems of non-linear equations
 \end{itemize}
@@
 \begin{itemize}
-\item ?
+\item Although systemfit has many features and produces
+   reliable results,
+   there are still many things that can be added to the package.
+\item For instance, we plan to allow for an estimation with
+   unbalanced data sets -- that are datasets with unequal numbers
+   of observations for each equation.
+\item Then we would like to add the estimation methods
+   °Limited Information Maximum Likelihood°,
+   °Full Information Maximum Likelihood°, and
+   °Generalized Method of Moments°.
+\item Furthermore, we plan to add estimation procedures
+   that account for serially correlated and heteroscedastic disturbances
+   as well as for spatial data.
+\item Finally, we would like to improve the function \texttt{nlsystemfit}
+   to estimate systems of non-linear equations.
 \end{itemize}
@@
 \begin{itemize}
-\item ?
+\item As I have indicated before,
+   the user can control many estimation details of systemfit.
+\item However, this has also an important drawback:
+\item the systemfit function has a lot of arguments.
+\item And the number of arguments will increase when we add further features
+   in the future.
+\item Although most arguments have reasonable default values,
+   the long list of arguments might confuse many newbies.
 \end{itemize}
@@
 \begin{itemize}
-\item ?
+\item Therefore, we are consider whether we should reduce the number
+   of arguments.
+\item All arguments that are generally used by advanced users only,
+   could be subsumed into a list of control parameters
+   like in the function °optim°.
+\item However: This would break existing code!
+\item I would like to know what you think about this issue.
+\item Do you think we should reduce the arguments
+   or keep all arguments as they are?
+\item Or do you have a better solution to make the function more concise?
 \end{itemize}

### 2006-06-15 paper/branches/useR2006/Slides/systemfit_slides.tex
useR 2006 Slides: some modifications (mainly of the notes)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@678 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
-%\documentclass{beamer}
-\documentclass[notes=show,handout]{beamer}
+\documentclass{beamer}
+%\documentclass[notes=show,handout]{beamer}
 %\documentclass[handout]{beamer}
@@
    \textbf{systemfit}\\ Simultaneous Equation Systems in R}
-\author{Arne Henningsen and Jeff D.\ Hamann}
-\date{University of Kiel (Germany) and\\Forest Informatics, Inc. (Corvallis, OR, USA)}
+\author{Arne Henningsen (University of Kiel, Germany)\\[0.4em]
+   Jeff D.\ Hamann (Forest Informatics, Inc., Corvallis, USA)}
+\date{Vienna, June 16, 2006}
 
@@
 \item Welcome to my presentation about the R package systemfit.
-\item This package is written by me, Arne Henningsen,
+\item The package can be used to estimate simultaneous equation systems in R.
+\item It has been written by me, Arne Henningsen,
    and by my co-author Jeff Hamann.
-\item The package can be used to estimate simultaneous equation systems in R.
 \end{itemize}
@@
 \bigskip
-$\Rightarrow$ All these models can be estimated by \texttt{systemfit}.
+$\Rightarrow$ All these models can be estimated by \texttt{systemfit}
 }
@@
    and take the height as well as the diameter of the trees into account.
-\item The disturbance terms of the single equations of such a equation system
+\item The disturbance terms of the equations of an equation system
    are likely contemporaneously correlated.
@@
    ignores this correlation and leads to inefficient estimates.
-\item However, efficients estimates can be obtained,
+\item However, efficient estimates can be obtained,
    if all equations are estimated simultaneously
    using Zellner's °Seemingly Unrelated Regression° method.
-\item Furthermore, when the theory suggests cross-equation
+\item Moreover, if the underlying theory suggests cross-equation
    parameter restrictions,
@@
 \item Features of systemfit
-\item Examples
+\item Example
 % \item Advantages
@@
 \item instrumental variables
-\item iterated estimation
+\item iteration of FGLS estimation
 \item formulas for the residual covariance matrix
@@
    by specifying different variables for each equation.
-\item Furthermore, it can be indicated whether feasible least squares
+\item Furthermore, it can be indicated whether feasible generalized least squares
    estimations should be iterated or not.
@@
 \setlength{\itemsep}{0.75em}
-\item wrapper function for (classical) panel-like data in long format:
-   \texttt{systemfitClassic}
+\item \texttt{systemfitClassic}: wrapper function for (classical)
+   panel-like data in long format
 \item testing linear hypotheses using the F-, Wald-, and LR-statistic
@@
 \item Of course, the food price is endogenous and therefore,
-   the system should be estimated by Three-Stage Least Squares
+   the system should be estimated by °Three-Stage Least Squares°
    using instrumental variables.
@@
 \begin{itemize}
-\item First, the summary results of the entire system are presented.
+\item First, the general results of the entire system are presented.
 \item For instance, the number of observations, \texttt{N},
@@
    is presented.
-\item A coefficient of correlation of 0.98 shows that the residuals
+\item A coefficient of correlation of about 0.98 shows that the residuals
    of the two equations are highly correlated.
@@
 to estimate systems of non-linear equations
+\item \ldots
 \end{itemize}
@@
 \item For instance, we plan to allow for an estimation with
-   unbalanced data sets -- that are datasets with unequal numbers
+   unbalanced data sets -- these are datasets with unequal numbers
    of observations for each equation.
@@
    to estimate systems of non-linear equations.
+\item And there are of other things that can be improved.
+\item If you are interested,
+   you are invited to help us improving systemfit.
 \end{itemize}
@@
 \begin{itemize}
+\item At the end of my presentation,
+   I would like to ask a question to the users of systemfit.
 \item As I have indicated before,
@@
    in the future.
-\item Although most arguments have reasonable default values,
+\item Although most arguments have reasonable default values
+   and can be ignored by unexperienced users,
    the long list of arguments might confuse many newbies.
@@
    of arguments.
-\item All arguments that are generally used by advanced users only,
+\item All arguments that are generally not used by unadvanced,
    could be subsumed into a list of control parameters
@@
 \note{
-   Thank you for your attention!
+   Thank you very much for your attention!
 }

### 2006-08-28 paper/trunk/JSS/response_A.tex
JSS/reviewer_*.*: added reviews
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@682 591db2bd-e0a0-4d22-8c19-2192dbb7da31

\documentclass[12pt,english]{scrartcl}
\input{response_commands}

\begin{document}
\begin{center}
{\Large\textbf{
systemfit: A Package to Estimate
Simultaneous Equation Systems in R (JSS~224)}}

\bigskip

{\Large\textbf{Responses to Reviewer A}}
\end{center}

We thank this reviewer for a very helpful set of comments and suggestions
for improving the paper.
The reviewer has made one overall comment, 9 major comments and
several minor comments.
We have structured our reply accordingly.

\section{Overall Comment}
% ########################################################
% ## Review for                                         ##
% ##   systemfit: A Package to Estimate Simultaneous    ##
% ##   Equation Systems in R                            ##
% ##   Arne Henningsen, Jeff D. Hamann                  ##
% ## submitted for publication in                       ##
% ##   Journal of Statistical Software                  ##
% ########################################################
% 
% The paper describes the CRAN package "systemfit" for fitting simultaneous
% equation models in R. Estimation procedures include standard techniques in
% econometrics such as SUR, 2SLS and 3SLS.
% The paper can be divided roughly into three parts: Sections 1/2 explain
% the underlying theory (~12 pages), Sections 3/4 describe the code in systemfit
% and how to use it (~10 pages) and Section 5 reproduces results from various
% textbooks (~8 pages).
% The paper and software are both interesting and mostly well-written: they
% deserve publication in JSS. However, some major and minor comments have to
% be addressed before it can be accepted. In particular, the revision has to
% streamline exposition and notation of standard results, and to combine the
% (established) theory with the implementation more closely (otherwise the reader
% has to wait more than 10 pages before s/he learns something about the package).
% I suggest to eliminate Section 2 in its current form and to re-write the
% description of the theory in a more general and unified way while weaving
% it with the translation into computational tools. As for the software itself,
% several modifications are necessary to make it more intuitive for R users
% and to improve how the underlying concepts are turned into computational 
% tools. More details are described below.
% 
% 
\section{Major Points}
% MAJOR POINTS:
\subsection{Title}
%   o title
% 
%     The title sounds a bit awkward in English: first, "a package *for*
%     estimat*ing*" would be more appropriate, secondly "systems of
%     equations" seems to be preferred over "equation systems" in English.
% 
\subsection{Abstract/Introduction}
%   o abstract/introduction
% 
%     Both, abstract and introduction, talk rather generally about
%     systems of equations reviewing rather well-known theoretical
%     results. I would suggest to revise both, focusing on what
%     *can* be done *in R* rather than giving a tutorial what *should*
%     be done *in general*. Of course, it is ok to review some theory,
%     however, practitioners should know what they are doing and the
%     paper should mainly teach them how to do it in R.
%
\subsection{Estimation Methods (2.1)}
%   o estimation methods (2.1)
% 
%     The description of the estimation methods can be shortened considerably
%     and should focus on the unifying properties of the different estimators.
%     All methods from OLS through to standard 3SLS comply with formulas
%     (12) and (13), only with differing choices for Omega (s*I, diagonal or
%     full Omega) and X (original X vs \hat X from instruments). The exposition
%     could be streamlined considerably if this structure were exploited.
% 
%     Terminology for the different estimators also varies between various
%     communities using systems of equations models. The paper should address
%     this. For example, in econometrics the GLS estimator would be an estimator
%     with known Omega while FGLS (feasible GLS) is the estimator based on
%     an estimated Omega (see Greene, 2003). The current discussion makes this
%     even more problematic because Omega is treated as known throughout
%     Section 2. The discussion of the estimators needs to be combined with
%     discussion of the estimators for Omega.
% 
%     For the varying 3SLS estimators, it would help to point out why/when
%     these are needed at the beginning rather than at the end of the 3SLS
%     section. 
%
\subsection{Imposing Linear Restrictions (2.2)}
%   o imposing linear restrictions (2.2)
% 
%     I found this section somewhat confusing. At first, it should be made
%     clear that this is about estimating the coefficients under linear
%     restrictions (and not inference or something else). Then, it needs
%     to be explained more concisely that all linear restrictions are of
%     type (33) and that the rest of the discussion is about different
%     computational algorithms for obtaining the estimates under restriction.
%     Again, the exposition on pages 7-9 can be improved (and shortened)
%     if the unifying structure is exploited. Also the notation can be
%     streamlined concerning beta, \hat beta, beta^0, beta^* etc.
%
\subsection{systemfitClassic (4.3)}
%   o systemfitClassic (4.3)
% 
%     First, this function deserves a better name: it is long, unintuitive
%     and does not reflect the underlying conceptual tools. If the requirement
%     of using the same formula for all equations would be relaxed (but
%     still supported) a conceivable name would be sur() for SUR type 
%     models. But it would also be conceivable to stick to the current
%     requirement of a single formula (then, still a better name would
%     be needed) and to set up a second wrapper sur() etc. In any case,
%     the function (its name, arguments and output) should reflect what
%     is conceptually intuitive for this kinds of models (for the communities
%     using it).
%     Second, the interface should also be re-designed: the standard in R is
%     that the formula/s is/are the first argument (which would also be more
%     convenient for systemfit()), e.g.,
%       function(formula, method = "FGLS", data = ..., ...)
%     Currently, the data argument seems to be mandatory (opposed to the
%     paper documentation). The specification of eqnVar and timeVar via
%     strings seems to be uncommon and suboptimal. Probably some kind of
%     formula would be preferable. Maybe a formula similar to those used in
%     nlme could be employed? If more than one formula is specified, one has
%     to pay attention that no problems are used by different patterns of
%     missingness in differing equations.
%
\subsection{Testing Linear Restrictions (4.4)}
%   o testing linear restrictions (4.4)
% 
%     The package introduces several functions that are somewhat orthogonal
%     to exisiting methodology. A single method (with "test" argument) for
%     the linear.hypothesis() function from "car" might seem more appropriate
%     here. The Wald test can already now be reproduced using the default
%     linear.hypothesis() method (except for formula formatting)
%       linear.hypothesis(fitsur, Rmat, qvec)    
%     If the coefficient naming would be modified (see below), the convenience
%     interface for specifying the restrictions in linear.hypothesis() could
%     also be re-used.    
%     Also two of the functions look like methods to the generics lrtest() and
%     waldtest() in package "lmtest", although this is not the case. Furthermore,
%     some more general method of comparing various fitted models would be nice,
%     e.g., an anova() method or methods to waldtest()/lrtest() although the 
%     latter two seem to be under development.
%     The output formatting of the current function is also not very elegant,
%     a new class along with its print method is introduced for each test. It
%     would be much easier to re-use "htest" or (probably better) "anova" as
%     classes for this.
%
\subsection{Organize Internals of systemfit}
%   o organize internals of systemfit
% 
%     The internals of a fitted "systemfit" object could be re-organized. Currently,
%     there are a lot of single letter elements - more standard naming as in "lm"
%     objects etc. would be more desirable. Also, computing the partial t tests
%     should/could be done in the summary() method, avoiding that certain parameters
%     (not affecting the fit itself, such as degrees of freedom) necessarily have to
%     be specified for model fitting.
% 
%     The technical control parameters could also be collected in a list of arguments, e.g.,
%       control = systemfit.control()
%     that sets suitable defaults but can also easily be changed. If possible `...'
%     arguments could be passed to systemfit.control() assuring backward compatibility
%     and facilitating interactive usage. The names of various
%     arguments should be re-thought, e.g. "tx", "R.restr" and "q.rests" seem rather
%     arbitrary. Similar comments hold for other arguments, see also below for some
%     more comments. As noted above, the standard in R seems to supply the formula(s)
%     as the first argument and to have then some `method' argument.
% 
%     Some more standard elements of a model fit should also be kept, in particular
%     the call so that update() can work with "systemfit" objects, e.g.,
%       fitsur <- systemfit("SUR", list(demand = eqDemand, supply = eqSupply))
%       fitsur$call <- call("systemfit", "SUR", list(demand = eqDemand, supply = eqSupply))
%       fitsurRmat <- update(fm, R.restr = Rmat, q.restr = qvec)
%     The call can easily be captured via match.call(). Similarly, one could think 
%     about how the terms and formula can be preserved and possibly made available
%     in terms() and formula() calls. Other standard methods would also be nice to have,
%     e.g., df.residual() or logLik(). (By the way: The row/column naming in fitted()
%     and residuals() should match, probably re-using equation and observation names.)
% 
%     The coefficient naming should be improved, e.g., "<eqname>_<varname>"
%     where <eqname> can be set by default to "eq1", "eq2" etc. but should use the
%     equation names (if supplied). Spaces in names should be avoided.
%     The coefficient naming in the systemfitClassic() interface could also be 
%     improved, see e.g. the output on p.27, l.546: "eq 2 capital.Chrysler" redundantly
%     codes the firm (both by eq and name) and "Chrysler_capital" or "Chryler.capital"
%     would be preferable.
% 
%     It would also be nice to have some more control about the amount of output of
%     the summary() method. Currently, the output can become very long easily when the
%     number of equations increases. One could want to control the equations and
%     correlations used, e.g., by something like
%       summary(sfobject, equation = ..., cor = ...)
% 
%
\subsection{Computations in systemfit}
%   o computations in systemfit()
% 
%     A few computations within the systemfit() functions can be further streamlined.
%     A simple modification is to use
%       method %in% c("OLS", "WLS", ...)
%     instead of
%       method == "OLS" | method == "WLS" | method == ...
%     Furthermore, various marix computations can be made even more stable. Currently,
%     the code still has    
%       solve(crossprod(X), crossprod(X,Y))    
%       solve(crossprod(X))
%     which looks very similar to the code presented in Bates (2004). However, Bates
%     (2004) uses the "Matrix" package, which has specialized solve() for certain
%     matrix types. To perform this `by hand' without Matrix, one way would be re-using
%     lm.fit() for the first command. Alternatively, the QR or Cholesky decomposition
%     could be carried out using qr() or chol(). The latter command could be replaced
%     by
%      chol2inv(qr.R(qr(X)))
%      chol2inv(chol(crossprod(X)))
%     where chol() should be faster but qr() might be more accurate.
%
\subsection{Replication of Textbook Results (5)}
%   o Replication of textbook results (5)
% 
%     Having a section like this is excellent and should help teachers and students
%     for using systemfit in their lectures. Furthermore, it encourages reproducible
%     research which always deserves support. However, I would suggest to streamline
%     the section to make it more readable. The section should discuss its aims (where
%     I would add replication to reliability) and summarize the most important results
%     and experiences from the replication. Then, I would pick a `typical' example and
%     discuss it in some more detail with full R output. For the remaining examples,
%     the input suffices as long as there are no deviations of the systemfit results
%     from the published numbers.
% 
%     In addition to Table 14.1 and 14.2, I tried to replicate Table 14.3 with
%     ML estimates (using iterated FGLS) in Greene (2003) via 
%       systemfitClassic("WSUR", invest ~ value + capital, "firm", "year",
%                        data = GrunfeldGreene, rcovformula = 0, pooled = TRUE, maxiter = 1000)    
%     for the pooled model which yielded rather similar results; and via
%       systemfitClassic("SUR", invest ~ value + capital, "firm", "year",
%                        data = GrunfeldGreene, rcovformula = 0, maxiter = 1000)    
%     for the full model which is not unsimilar to Greene's results but also not very
%     close either. Is this a problem of systemfit or of Greene or of me fitting the
%     wrong model?
% 
%     Having two Grunfeld data sets is confusing as the data used by Theil is simply
%     a subset of the data used by Greene:
%       subset(GrunfeldGreene, firm %in% c("General Electric", "Westinghouse"))
%     Of course, there are different versions of this data set, but neither the paper
%     nor the manual pages of GrunfeldGreene and GrunfeldTheil appropriately discuss
%     the problem. From the manual and the paper, the reader is only able to feel
%     vaguely insecure about the appropriateness of the selected subsets. Also note
%     that a more complete (and slightly different) version of the data set is
%     available in package "Ecdat".
% 
%
\section{Minor Points and Typos}
% MINOR POINTS & TYPOS
% 
%   o p.1, l.3. Is a "theoretical model that is econometrically estimated"
%     a well-defined term?
% 
%   o p.2, Eq.2. The \cdot between the regressor matrix and its coefficients
%     should be omitted (since this is not a scalar multiplication).
% 
%   o p.2, l.42. For symmetry between the two index types, it might be
%     more convenient to denote the equation by i/j and the observation
%     by s/t rather than t/t*, in particular as * is also used later for 
%     other purposes.
% 
%   o p.3, l.45. While it is possible to call u "residual" and \hat u an 
%     "estimated residual", the more standard (and less confusing) terminology
%     is to call u "disturbance" or "error term" and the estimated quantity
%     \hat u "residual".
% 
%   o p.3, l.52. The equation $\sigma^2 = E(u'u)$ is wrong because the 
%     right-hand side is a sum over squared errors. This mistake is repeated
%     throughout the paper. Also (as noted above), no distinction between
%     population and sample covariances is made which leads to confusion.
% 
%   o p.3, l.69. "the GLS is" -> "the GLS estimator is"
%                "called ... SUR" -> "called ... SUR estimator"
% 
%   o p.4, l.72/3. The "E[u|X] = 0" type equations imply (but are not
%     equivalent to) the assumptions of uncorrelated disturbances.
% 
% 
%   o p.4, l.79. The formulation "estimated coefficients are biased" is somewhat
%     sloppy, bias is a property of the estimator.
% 
%   o p.6, l.127 / p.10, l.198. The symbol T is used for different quantities
%     here.
% 
%   o p.12, eq.63. One Sigma is lacking a hat. Now M is used for the number
%     of equations, before it was G.
% 
%   o p.13, l.285. Rather than calling the methods by their full name, it
%     is preferred to refer to the predict or coef method for "systemfit"
%     objects.
% 
%   o p. 13, l.289. The remarks about why not the full code is presented
%     sound very clumsy and should be avoided.
% 
%   o p.13, l.299. "interest" -> "interested"
% 
%   o p.13, l.321. "deviding" -> "dividing"
% 
%   o p.14, l.334. values to arguments should also be set in verbatim,
%     e.g., "``OLS''" -> "\code{"OLS"}"
% 
%   o p.14, l.340. "tilde" -> "\verb/~/"
% 
%   o p.14, l.341. Are only "+" signs allowed or the full standard
%     formula language in R?
% 
%   o p.14, eq.68/69/71. "*" -> "\cdot" (and not in verbatim)
% 
%   o p.14, l.346. "dependant" -> "dependent"
% 
%   o p.17, l.387. The creation of tx should be simplified, e.g., the
%     first six lines could be replaced by
%       tx <- cbind(rbind(diag(5), 0, 0), 0)
% 
%   o p.17, l.405. "large" -> "larger", also say that maxiter = 1 is the
%     default.
% 
%   o p.18, Tab.1. I find it a bit confusing that the argument ist called 
%     "rcovformula" but does not expect a formula (in the R sense). Also,
%     type 0 should have an alternative string specification. A similar
%     argument holds for "formula3sls".
%

% \bibliographystyle{}
% \bibliography{}

\end{document}

### 2006-08-28 paper/trunk/JSS/response_B.tex
JSS/response_B.tex: added template for respionse to reviewer B
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@683 591db2bd-e0a0-4d22-8c19-2192dbb7da31

\documentclass[12pt,english]{scrartcl}
\input{response_commands}

\begin{document}
\begin{center}
{\Large\textbf{
systemfit: A Package to Estimate
Simultaneous Equation Systems in R (JSS~224)}}

\bigskip

{\Large\textbf{Responses to Reviewer A}}
\end{center}

We thank this reviewer for a very helpful set of comments and suggestions
for improving the paper.
The reviewer has made one overall comment, 9 specific comments and
several smaller comments.
We have structured our reply accordingly.

\section{Overall Comment}
% Review of the manuscript 'systemfit: A package to estimate
% simultaneous equation systems in R'
% This manuscript describes a new R package systemfit for inference in multi-equation models. The
% main parts are section 2 and 4. Section 2 gives the statistical background for the analysis tools
% available in systemfit while section 4 describes how to apply systemfit. The theory is adequately
% explained and the syntax in systemfit is nice and logical allowing complicated models to be quite
% easily fitted.
% My only real concern is that systemfit relies on classic econometric (least squares) developed 40
% years ago. Since then computational power has increased dramatically and superior methods may be
% available. First of all, full likelihood methods are available in existing software for structural
% equation modelling. In particular, R has a package called SEM. I am not sure what is gained with
% systemfit compared to this. Also it is a weakness that systemfit only allows analysis of complete
% cases.

\section{Specific comments}
% Specific comments

\subsection{Instrumental variable estimation}
% Instrumental variable estimation has received much attention lately in biostatistics. Maybe it would
% help 'sell' systemfit if this was emphasized at bit more. For instance, instrumental variables could
% be mentioned directly in the introduction and included as a key word.

\subsection{Section 2}
% Section 2 gives a brief overview of the statistical background. This may be a bit hard to follow for
% people outside econometrics. It would help to include more references, especially in section 2.1.

\subsection{Number of observations}
% Page 2: 'observations' should be explained earlier together with the presentation of the data
% structure. The number of observations should be stated.

\subsection{x and y variables}
% Page 2: It should be clarified whether y-variables in one equation can be x-variables in another.

\subsection{'may' should be 'must'}
% Page 4, line 82: 'may' should be 'must'. Here the difference is extremely important. Maybe the text
% could emphasize more clearly what is required of a variable if it is to be used as an instrumental
% variable. What is a good instrumental variable and what happens if the requirements are not
% fulfilled?

\subsection{Equation 20 and 23}
% The difference between equation 20 and 23 is so smaller that maybe it should be noted in the text
% what it is.

\subsection{Sigma}
% Line 193-194: Language is confusing. It should be explained that calculation of some of the
% estimators earlier introduced require Sigma to be known. Often Sigma is not known so an estimator
% is used instead. However, then statistical properties may changed compared to the theory described
% earlier.

\subsection{First step OLS}
% Line 196: It should be explained what is meant by a first step OLS.

\subsection{Efficiency and maximum likelihood}
% Section 2.3: Somewhere in this section or (2.1) it should be noted that the efficiency results of
% section 2.1 depend on Sigma being known. And that a full maximum likelihood analysis may be
% superior.
%

\subsection{Alternative approach}
% Line 215-219: Alternative approach should be explained more clearly.

\subsection{Syntax for instrumental variables}
% Line 375-379: Syntax on programming of instrumental variables should be made more clear. What
% is a one-sided formula? Does the code given imply that the sum of income, farmPrice and trend was
% used as and instrumental variable? If so, it should be stated in the text.

\subsection{Section 2.7}
% Section 2.7: Given the theory of the previous sections, I was confused about the Hausman test.
% Under the alternative hypothesis, independent variables are correlated with error terms. In this
% section it is stated that this leads to inconsistency of 3SLS. However, on page 5 (top) the opposite is
% stated? Would it not make more sense to compare SUR to 3SLS?

\section{Smaller comments}
% Smaller corrections:
% Sometimes ':' is used just before an equation other times not.
% Page 7, line 147: change 'doesn't' to 'does not'.
% Line 135: '\beta^*' should be '\widehat{\beta^*}'
% Line 209: first \widehat{\Sigma} should be \Sigma
% Equation 63: '\Sigma' should be '\widehat{\Sigma}
% Equation 65: Why two hats?
% Line 299: 'interest' should be 'interested'
% Page 14: 'dependant' should be 'dependent'

% \bibliographystyle{}
% \bibliography{}

\end{document}

### 2006-08-29 paper/trunk/systemfit_code.tex
corrected the paper according to all smaller comments of reviewer B except for the first point
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@684 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 The \code{systemfit} function returns many objects
-that users might be interest in.
+that users might be interested in.
 A complete list is available in the documentation of this function

### 2006-08-29 paper/trunk/systemfit_other.tex
corrected the paper according to all smaller comments of reviewer B except for the first point
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@684 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 \end{eqnarray} 
-This yields an unbiased estimator for all elements of $\widehat{\Sigma}$,
+This yields an unbiased estimator for all elements of $\Sigma$,
 but even if $\widehat{\Sigma}$ is an unbiased estimator of $\Sigma$, 
@@
 }{
-   \hat{u}' ( \Sigma \otimes I )^{-1} \hat{u} /
+   \hat{u}' ( \widehat{\Sigma} \otimes I )^{-1} \hat{u} /
    ( M \cdot T - K )
@@
 
-The LR-statistic for systems of equations is
+The LR statistic for systems of equations is
 \begin{equation}
 LR = T \cdot \left(
-   log \left| \hat{ \hat{ \Sigma } }_r \right|
-   - log \left| \hat{ \hat{ \Sigma } }_u \right|
+   log \left| \hat{ \Sigma }_r \right|
+   - log \left| \hat{ \Sigma }_u \right|
    \right)
@@
 where $T$ is the number of observations per equation, and 
-$\hat{\hat{\Sigma}}_r$ and $\hat{\hat{\Sigma}}_u$ are
+$\hat{\Sigma}_r$ and $\hat{\Sigma}_u$ are
 the residual covariance matrices calculated by formula (\ref{eq:rcov0})

### 2006-08-29 paper/trunk/systemfit_restrictions.tex
rorganized the sections and subsections: there now a section called "Statistical Background" that includes the subsections (former sections) "Estimation", "Restrictions" and "Other issues" (a section of the "statistical background" is suggested by the JSS
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@614 591db2bd-e0a0-4d22-8c19-2192dbb7da31
corrected the paper according to all smaller comments of reviewer B except for the first point
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@684 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 (equations~\ref{eq:ols}--\ref{eq:3slsEViews}).
-This results in the linear independent parameter estimates $\beta^*$ and
+This results in the linear independent parameter estimates $\widehat{\beta}^*$ and
 their covariance matrix.

### 2006-08-29 paper/trunk/systemfit_other.tex
all "\hat"s are "\widehat"s now
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@685 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 F = \frac{
-   ( R \hat{\beta} - q )'
-   ( R ( X' ( \hat{\Sigma} \otimes I )^{-1} X )^{-1} R' )^{-1}
-   ( R \hat{\beta} - q ) /
+   ( R \widehat{\beta} - q )'
+   ( R ( X' ( \widehat{\Sigma} \otimes I )^{-1} X )^{-1} R' )^{-1}
+   ( R \widehat{\beta} - q ) /
    j
 }{
-   \hat{u}' ( \widehat{\Sigma} \otimes I )^{-1} \hat{u} /
+   \widehat{u}' ( \widehat{\Sigma} \otimes I )^{-1} \widehat{u} /
    ( M \cdot T - K )
@@
 $K$ is the total number of estimated coefficients, and
-$\hat{\Sigma}$ is the estimated residual covariance matrix
+$\widehat{\Sigma}$ is the estimated residual covariance matrix
 used in the estimation.
@@
 W =
-   ( R \hat{\beta} - q )'
-   ( R \widehat{Cov} [ \hat{\beta} ] R' )^{-1}
-   ( R \hat{\beta} - q )
+   ( R \widehat{\beta} - q )'
+   ( R \widehat{Cov} [ \widehat{\beta} ] R' )^{-1}
+   ( R \widehat{\beta} - q )
 \end{equation}
@@
 LR = T \cdot \left(
-   log \left| \hat{ \Sigma }_r \right|
-   - log \left| \hat{ \Sigma }_u \right|
+   log \left| \widehat{ \Sigma }_r \right|
+   - log \left| \widehat{ \Sigma }_u \right|
    \right)
@@
 where $T$ is the number of observations per equation, and 
-$\hat{\Sigma}_r$ and $\hat{\Sigma}_u$ are
+$\widehat{\Sigma}_r$ and $\widehat{\Sigma}_u$ are
 the residual covariance matrices calculated by formula (\ref{eq:rcov0})
@@
 \begin{equation}
-  m = \left( \hat{\beta}_{2SLS} - \hat{\beta}_{3SLS} \right)^{'}
-      \left( \Cov \left[ \hat{\beta}_{2SLS} \right] -
-             \Cov \left[ \hat{\beta}_{3SLS} \right] \right)
-      \left( \hat{\beta}_{2SLS} - \hat{\beta}_{3SLS} \right)
+  m = \left( \widehat{\beta}_{2SLS} - \widehat{\beta}_{3SLS} \right)^{'}
+      \left( \Cov \left[ \widehat{\beta}_{2SLS} \right] -
+             \Cov \left[ \widehat{\beta}_{3SLS} \right] \right)
+      \left( \widehat{\beta}_{2SLS} - \widehat{\beta}_{3SLS} \right)
 \label{eq:hausman}
 \end{equation}
-where $\hat{\beta}_{2SLS}$ and $\Cov \left[ \hat{\beta}_{2SLS} \right]$ are the estimated
+where $\widehat{\beta}_{2SLS}$ and $\Cov \left[ \widehat{\beta}_{2SLS} \right]$ are the estimated
 coefficient and covariance matrix from 2SLS estimation, and
-$\hat{\beta}_{3SLS}$ and $\Cov \left[ \hat{\beta}_{3SLS} \right]$ are the estimated
+$\widehat{\beta}_{3SLS}$ and $\Cov \left[ \widehat{\beta}_{3SLS} \right]$ are the estimated
 coefficients and covariance matrix from 3SLS estimation.

### 2006-08-29 paper/trunk/systemfit_code.tex
corrected / harmonized punctuation marks around equations (first small comment of reviewer B)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@686 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 adding up some of these submatrices, and
-finally putting the submatrices together:
+finally putting the submatrices together, so that
 \begin{equation}
@@
    \sigma^{Gi} {X^G} \\
-\end{array} \right]'
+\end{array} \right]' ,
 \end{equation}

### 2006-08-29 paper/trunk/systemfit_estimation.tex
corrected / harmonized punctuation marks around equations (first small comment of reviewer B)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@686 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 \begin{equation}
-   \widehat{\beta}_{OLS} = \left( X'X \right)^{-1} X'y
+   \widehat{\beta}_{OLS} = \left( X'X \right)^{-1} X'y .
    \label{eq:ols}
@@
 \begin{equation}
-   \Cov \left[ \widehat{\beta}_{OLS} \right] = \sigma^2 \left( X'X \right)^{-1}
+   \Cov \left[ \widehat{\beta}_{OLS} \right] = \sigma^2 \left( X'X \right)^{-1} ,
    \label{eq:olsCovSameSigma}
 \end{equation}
-with $\sigma^2 = E \left( u' u \right)$.
+where $\sigma^2 = E \left( u' u \right)$.
 This assumes that the disturbances of all equations have the
@@
 \begin{equation}
-   \Cov \left[ \widehat{\beta}_{OLS} \right] = \left( X' \Omega^{-1} X \right)^{-1}
+   \Cov \left[ \widehat{\beta}_{OLS} \right] = \left( X' \Omega^{-1} X \right)^{-1} ,
    \label{eq:olsCovSingleSigma}
 \end{equation}
-with $\Omega = \Sigma \otimes I$,
+where $\Omega = \Sigma \otimes I$,
 $\sigma_{ij} = 0 \; \forall \; i \neq j$ and
@@
 \begin{equation}
-   \widehat{\beta}_{WLS} = \left( X' \Omega^{-1} X \right)^{-1} X' \Omega^{-1} y
+   \widehat{\beta}_{WLS} = \left( X' \Omega^{-1} X \right)^{-1} X' \Omega^{-1} y ,
 \end{equation}
-with $\Omega = \Sigma \otimes I$, 
+where $\Omega = \Sigma \otimes I$,
 $\sigma_{ij} = 0 \; \forall \; i \neq j$ and
@@
 \begin{equation}
-   \Cov \left[ \widehat{\beta}_{WLS} \right] = \left( X' \Omega^{-1} X \right)^{-1}
+   \Cov \left[ \widehat{\beta}_{WLS} \right] = \left( X' \Omega^{-1} X \right)^{-1} .
 \end{equation}
@@
 Least Squares (GLS) estimation leads to efficient parameter estimates.
-In this case, the GLS is generally called "Seemingly Unrelated Regression"
+In this case, the GLS is generally called °Seemingly Unrelated Regression°
 (SUR) \citep{zellner62}.
@@
 
-The SUR estimator can be obtained by:
+The SUR estimator can be obtained by
 \begin{equation}
-   \widehat{\beta}_{SUR} = \left( X' \Omega^{-1} X \right)^{-1} X' \Omega^{-1} y
+   \widehat{\beta}_{SUR} = \left( X' \Omega^{-1} X \right)^{-1} X' \Omega^{-1} y ,
    \label{eq:sur}
 \end{equation}
-with $\Omega = \Sigma \otimes I$ and
+where $\Omega = \Sigma \otimes I$ and
 $\sigma_{ij} = E \left( u_i' u_j \right)$.
@@
 \begin{equation}
-   \Cov \left[ \widehat{\beta}_{SUR} \right] = \left( X' \Omega^{-1} X \right)^{-1}
+   \Cov \left[ \widehat{\beta}_{SUR} \right] = \left( X' \Omega^{-1} X \right)^{-1} .
    \label{eq:surCov}
@@
 \begin{equation}
-   \widehat{X_i} = H_i \left( H_i' H_i \right)^{-1} H_i' X
+   \widehat{X_i} = H_i \left( H_i' H_i \right)^{-1} H_i' X .
 \end{equation}
 At the second stage the unbiased two-stage least squares estimates
-of $\beta$ are obtained by:
+of $\beta$ are obtained by
 \begin{equation}
    \widehat{\beta}_{2SLS} = \left( \widehat{X}' \widehat{X} \right)^{-1} 
-   \widehat{X}' y 
+   \widehat{X}' y .
    \label{eq:beta2sls}
@@
    \Cov \left[ \widehat{\beta}_{2SLS} \right] = \sigma^2 \left( \widehat{X}'
-   \widehat{X} \right)^{-1}
+   \widehat{X} \right)^{-1} ,
    \label{eq:2slsCovSameSigma}
 \end{equation}
-with $\sigma^2 = E \left( u' u \right)$.
+where $\sigma^2 = E \left( u' u \right)$.
 If the disturbance terms of the individual equations
@@
    \Cov \left[ \widehat{\beta}_{2SLS} \right] = \left( \widehat{X}' \Omega^{-1}
-   \widehat{X} \right)^{-1}
+   \widehat{X} \right)^{-1} ,
    \label{eq:2slsCovSingleSigma}
 \end{equation}
-with $\Omega = \Sigma \otimes I$, 
+where $\Omega = \Sigma \otimes I$,
 $\sigma_{ij} = 0 \; \forall \; i \neq j$ and
@@
    \widehat{\beta}_{W2SLS} = \left( \widehat{X}' \Omega^{-1} \widehat{X} 
-   \right)^{-1} \widehat{X}' \Omega^{-1} y
+   \right)^{-1} \widehat{X}' \Omega^{-1} y ,
 \end{equation}
-with $\Omega = \Sigma \otimes I$, 
+where $\Omega = \Sigma \otimes I$,
 $\sigma_{ij} = 0 \; \forall \; i \neq j$ and
@@
    \Cov \left[ \widehat{\beta}_{W2SLS} \right] = \left( \widehat{X}' \Omega^{-1}
-   \widehat{X} \right)^{-1}
+   \widehat{X} \right)^{-1} .
 \end{equation}
@@
 
-The standard 3SLS estimator can be obtained by:
+The standard 3SLS estimator can be obtained by
 \begin{equation}
    \widehat{\beta}_{3SLS} = \left( \widehat{X}' \Omega^{-1} \widehat{X} 
-   \right)^{-1} \widehat{X}' \Omega^{-1} y
+   \right)^{-1} \widehat{X}' \Omega^{-1} y ,
    \label{eq:3slsGls}
 \end{equation}
-with $\Omega = \Sigma \otimes I$ and
+where $\Omega = \Sigma \otimes I$ and
 $\sigma_{ij} = E \left( u_i' u_j \right)$.
-Its covariance matrix is:
+Its covariance matrix is
 \begin{equation}
    \Cov \left[ \widehat{\beta}_{3SLS} \right] = \left( \widehat{X}' \Omega^{-1}
-   \widehat{X} \right)^{-1}
+   \widehat{X} \right)^{-1} .
    \label{eq:cov3sls}
@@
       0 & 0 & \cdots & H_G
-   \end{array}\right]
+   \end{array}\right] .
 \end{equation}
 Since there might be occasions where this cannot be avoided,
-\cite{schmidt90} analyses other approaches to obtain 3SLS estimators:
-
-One of these approaches is based on instrumental variable estimation
-(3SLS-IV):
+\cite{schmidt90} analyses other approaches to obtain 3SLS estimators.
+One of these approaches based on instrumental variable estimation
+(3SLS-IV) is
 \begin{equation}
    \widehat{\beta}_{3SLS-IV} = \left( \widehat{X}' \Omega^{-1} X
-   \right)^{-1} \widehat{X}' \Omega^{-1} y
+   \right)^{-1} \widehat{X}' \Omega^{-1} y .
    \label{eq:3slsIv}
 \end{equation}
-The covariance matrix of this 3SLS-IV estimator is:
+The covariance matrix of this 3SLS-IV estimator is
 \begin{equation}
    \Cov \left[ \widehat{\beta}_{3SLS-IV} \right] = \left( \widehat{X}' \Omega^{-1}
-   X \right)^{-1}
+   X \right)^{-1} .
 \end{equation}
-Another approach is based on the Generalized Method of Moments (GMM)
-estimator (3SLS-GMM):
+Another approach based on the Generalized Method of Moments (GMM)
+estimator (3SLS-GMM) is
 \begin{equation}
    \widehat{\beta}_{3SLS-GMM} = \left( X' H \left( H' \Omega H \right)^{-1}
-   H' X \right)^{-1} X' H \left( H' \Omega H \right)^{-1} H' y
+   H' X \right)^{-1} X' H \left( H' \Omega H \right)^{-1} H' y .
    \label{eq:3slsGmm}
 \end{equation}
-The covariance matrix of the 3SLS-GMM estimator is:
+The covariance matrix of the 3SLS-GMM estimator is
 \begin{equation}
    \Cov \left[ \widehat{\beta}_{3SLS-GMM} \right] =
-   \left( X' H \left( H' \Omega H \right)^{-1} H' X \right)^{-1}
+   \left( X' H \left( H' \Omega H \right)^{-1} H' X \right)^{-1} .
 \end{equation}
-A fourth approach developed by \cite{schmidt90} himself is:
+A fourth approach developed by \cite{schmidt90} himself is
 \begin{equation}
@@
    \right)^{-1} \widehat{X}' \Omega^{-1} 
-   H \left( H' H \right)^{-1} H' y 
+   H \left( H' H \right)^{-1} H' y .
    \label{eq:3slsSchmidt}
 \end{equation}
-The covariance matrix of this estimator is:
+The covariance matrix of this estimator is
 \begin{equation}
@@
    H \left( H' H \right)^{-1} H' \Omega^{-1} \widehat{X}
-   \left( \widehat{X}' \Omega^{-1}  \widehat{X} \right)^{-1}
+   \left( \widehat{X}' \Omega^{-1}  \widehat{X} \right)^{-1} .
 \end{equation}
-The econometrics software EViews uses following approach:
+The econometrics software EViews uses
 \begin{equation}
@@
    \left( \widehat{X}' \Omega^{-1} \widehat{X} \right)^{-1} 
-   \widehat{X}' \Omega^{-1} \left( y - X \widehat{\beta}_{2SLS} \right)
+   \widehat{X}' \Omega^{-1} \left( y - X \widehat{\beta}_{2SLS} \right) ,
    \label{eq:3slsEViews}

### 2006-08-29 paper/trunk/systemfit_other.tex
corrected / harmonized punctuation marks around equations (first small comment of reviewer B)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@686 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 can be calculated from the residuals of a first-step OLS or 2SLS estimation.
-The following formula is often applied:
+In many cases, the formula
 \begin{equation}
@@
 \end{equation}
-where $T$ is the number of observations in each equation.
+is applied, where $T$ is the number of observations in each equation.
 However, in finite samples this estimator is biased,
@@
    \widehat{\sigma}_{ij} = \frac{ \widehat{u}_i' \widehat{u}_j }
-   { T - \max \left( K_i , K_j \right) }
+   { T - \max \left( K_i , K_j \right) } \; ,
    \label{eq:rcov3}
@@
 A further approach to obtain the estimated residual covariance
-matrix is \citep[p.\ 309]{zellner62c}
+matrix is
 \begin{eqnarray}
@@
 \end{eqnarray} 
+\citep[p.\ 309]{zellner62c}.
 This yields an unbiased estimator for all elements of $\Sigma$,
@@
 The goodness of fit of each single equation can be measured by the
-traditional $R^2$ values:
+traditional $R^2$ values
 \begin{equation}
    R_i^2 = 1 - \frac{ \widehat{u}_i' \widehat{u}_i }
-   { ( y_i - \overline{y_i} )' ( y_i - \overline{y_i} ) }
+   { ( y_i - \overline{y_i} )' ( y_i - \overline{y_i} ) } \; ,
 \end{equation}
@@
 The goodness of fit of the whole system can be measured by the
-McElroy's $R^2$ value \citep{mcelroy77}: 
+McElroy's $R^2$ value 
 % also: \citep[p.\ 345]{greene03}
@@
 \end{equation}
+\citep{mcelroy77},
 where $T$ is the number of observations in each equation,
@@
    ( M \cdot T - K )
-}
+} \; ,
 \end{equation}
@@
    ( R \widehat{Cov} [ \widehat{\beta} ] R' )^{-1}
-   ( R \widehat{\beta} - q )
+   ( R \widehat{\beta} - q ) .
 \end{equation}
@@
    - log \left| \widehat{ \Sigma }_u \right|
-   \right)
+   \right) ,
 \end{equation}
@@
 but the 3SLS estimator is inconsistent.
-The Hausman test statistic is,
+The Hausman test statistic is
 \begin{equation}
@@
              \Cov \left[ \widehat{\beta}_{3SLS} \right] \right)
-      \left( \widehat{\beta}_{2SLS} - \widehat{\beta}_{3SLS} \right)
+      \left( \widehat{\beta}_{2SLS} - \widehat{\beta}_{3SLS} \right) ,
 \label{eq:hausman}

### 2006-08-29 paper/trunk/systemfit_restrictions.tex
corrected / harmonized punctuation marks around equations (first small comment of reviewer B)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@686 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 \begin{equation}
-   \beta = T \cdot \beta^* \label{eq:T-restr} 
+   \beta = T \cdot \beta^* \label{eq:T-restr} ,
 \end{equation}
@@
 To impose these restrictions, the $X$ matrix is
-(post-)\hspace{0pt}multiplied by this $T$ matrix.
+(post-)\hspace{0pt}multiplied by this $T$ matrix, so that
 \begin{equation}
-    X^* = X \cdot T
+    X^* = X \cdot T .
 \end{equation}
@@
 and the covariance matrix of the original parameters
-can be obtained by:
+can be obtained by
 \begin{equation}
-   \Cov \left[ \widehat{\beta} \right] = T \cdot \Cov \left[ \widehat{\beta}^* \right] \cdot T'
+   \Cov \left[ \widehat{\beta} \right]
+   = T \cdot \Cov \left[ \widehat{\beta}^* \right] \cdot T' .
 \end{equation}
@@
 \begin{equation}
-   R \beta^0 = q
+   R \beta^0 = q ,
 \end{equation}
@@
 imposed on the linear independent parameters due to the restrictions
-imposed using the first method:
+imposed using the first method, so that
 \begin{equation}
-   R \beta^{*0} = q
+   R \beta^{*0} = q ,
 \end{equation}
@@
       X' y \\ q 
-   \end{array} \right]
+   \end{array} \right] ,
 \end{equation}
@@
       R & 0
-   \end{array} \right]^{-1}
+   \end{array} \right]^{-1} ,
 \end{equation}
-with $\sigma^2 = E \left( u' u \right)$.
+where $\sigma^2 = E \left( u' u \right)$.
 If the disturbance terms of the individual equations
@@
       R & 0
-   \end{array} \right]^{-1}
+   \end{array} \right]^{-1} ,
 \end{equation}
-with $\Omega = \Sigma \otimes I$, 
+where $\Omega = \Sigma \otimes I$,
 $\sigma_{ij} = 0 \; \forall \; i \neq j$ and
@@
       X' \Omega^{-1} y \\ q 
-   \end{array} \right]
+   \end{array} \right] ,
 \end{equation}
-with $\Omega = \Sigma \otimes I$, 
+where $\Omega = \Sigma \otimes I$,
 $\sigma_{ij} = 0 \; \forall \; i \neq j$ and
@@
       R & 0
-   \end{array} \right]^{-1}
+   \end{array} \right]^{-1} .
 \end{equation}
@@
       X' \Omega^{-1} y \\ q 
-   \end{array} \right]
+   \end{array} \right] ,
 \end{equation}
-with $\Omega = \Sigma \otimes I$ and
+where $\Omega = \Sigma \otimes I$ and
 $\sigma_{ij} = E \left( u_i' u_j \right)$.
@@
       \widehat{X}' y \\ q 
-   \end{array} \right]
+   \end{array} \right] .
    \label{eq:beta2SLSr}
@@
       \widehat{X}' \Omega^{-1} y \\ q 
-   \end{array} \right]
+   \end{array} \right] ,
 \end{equation}
-with $\Omega = \Sigma \otimes I$, 
+where $\Omega = \Sigma \otimes I$,
 $\sigma_{ij} = 0 \; \forall \; i \neq j$ and
@@
       \widehat{X}' \Omega^{-1} y \\ q 
-   \end{array} \right]
+   \end{array} \right] ,
    \label{eq:3slsGlsR}
 \end{equation}
-with $\Omega = \Sigma \otimes I$ and
+where $\Omega = \Sigma \otimes I$ and
 $\sigma_{ij} = E \left( u_i' u_j \right)$.
@@
       R & 0
-   \end{array} \right]^{-1}
+   \end{array} \right]^{-1} .
    \label{eq:cov3slsr}
@@
       \widehat{X}' \Omega^{-1} y \\ q 
-   \end{array} \right]
+   \end{array} \right] ,
    \label{eq:3slsIvR}
 \end{equation}
-with
+where
 \begin{equation}
@@
       X' H \left( H \Omega H \right)^{-1} H' y \\ q 
-   \end{array} \right]
+   \end{array} \right] ,
    \label{eq:3slsGmmR}
 \end{equation}
-with
+where
 \begin{equation}
@@
       R & 0
-   \end{array} \right]^{-1}
+   \end{array} \right]^{-1} .
 \end{equation}
 The restricted 3SLS estimator based on the suggestion of
-\cite{schmidt90} is:
+\cite{schmidt90} is
 \begin{equation}
@@
       \widehat{X}' \Omega^{-1} H \left( H' H \right)^{-1} H' y \\ q 
-   \end{array} \right]
+   \end{array} \right] ,
    \label{eq:3slsSchmidtR}
 \end{equation}
-with
+where
 \begin{eqnarray}
@@
       R & 0
-   \end{array} \right]^{-1}
+   \end{array} \right]^{-1} .
    \nonumber
 \end{eqnarray}
-The econometrics software EViews calculates the restricted 3SLS estimator by:
+The econometrics software EViews calculates the restricted 3SLS estimator by
 \begin{equation}
@@
       \\ q 
-   \end{array} \right]
+   \end{array} \right] ,
    \label{eq:3slsEViewsR}

### 2006-08-29 paper/trunk/systemfit_statistics.tex
gotta get my changes in...
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@645 591db2bd-e0a0-4d22-8c19-2192dbb7da31
corrected / harmonized punctuation marks around equations (first small comment of reviewer B)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@686 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 \begin{equation}
-   y_{i} = X_i \beta_i + u_i, \quad i = 1, 2, \ldots, G
+   y_{i} = X_i \beta_i + u_i, \quad i = 1, 2, \ldots, G ,
 \end{equation}
@@
 \begin{equation}
-   y = X \beta + u 
+   y = X \beta + u .
 \end{equation}
@@
 We assume that there is no correlation of the disturbance terms 
-across observations:
+across observations, so that
 \begin{equation}
    E \left( u_{it} \, u_{jt^*} \right) = 0 
-   \; \forall \; t \neq t^*
+   \; \forall \; t \neq t^* ,
 \end{equation}
@@
 
-However, we explicitly allow for contemporaneous correlation:
+However, we explicitly allow for contemporaneous correlation, i.e.\
 \begin{equation}
-   E \left( u_{it} \, u_{jt} \right) = \sigma_{ij}
+   E \left( u_{it} \, u_{jt} \right) = \sigma_{ij} .
 \end{equation}
@@
 \begin{equation}
-   E \left( u \, u' \right) = \Omega = \Sigma \otimes I
+   E \left( u \, u' \right) = \Omega = \Sigma \otimes I ,
 \end{equation}

### 2006-08-29 paper/trunk/systemfit_code.tex
corrected the paper according to all minor comments of reviewer A except for "p.3, l.45", the second part of "p.3, l.52" and "p.18, Tab.1".
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@687 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 to estimate systems of equations
-(\code{systemfit}, \code{systemfitClassic})
-and to test hypotheses in these systems
-(\code{ftest.systemfit}, \code{wald\-test.sys\-tem\-fit},
-\code{lrtest.systemfit}, \code{hausman.systemfit}).
-Furthermore, this package provides some helper functions
-e.g. to extract the estimated coefficients (\code{coef.systemfit})
-or to calculate predicted values (\code{predict.systemfit}).
+(\code{systemfit}, \code{systemfitClassic}).
+Furthermore, this package provides some methods
+to test hypotheses
+(\code{ftest}, \code{wald\-test},
+\code{lrtest}, \code{hausman}),
+to extract the estimated coefficients (\code{coef})
+and to calculate predicted values (\code{predict}).
 
-The source code of the \pkg{systemfit} is publicly available
+The source code of \pkg{systemfit} is publicly available
 for download from °CRAN° (The Comprehensive R Archive Network,
 \url{http://cran.r-project.org/src/contrib/Descriptions/systemfit.html}).
-Since the whole package has more than 2,100 lines of code,
-it is not presented in this article.
-Furthermore, the code corresponds exactly to the procedures and formulas
+It corresponds exactly to the procedures and formulas
 described in the previous section.

### 2006-08-29 paper/trunk/systemfit_estimation.tex
corrected the paper according to all minor comments of reviewer A except for "p.3, l.45", the second part of "p.3, l.52" and "p.18, Tab.1".
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@687 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 \end{equation}
-where $\sigma^2 = E \left( u' u \right)$.
+where $\sigma^2 = E \left( u_{it}^2 \right)$.
 This assumes that the disturbances of all equations have the
@@
 $\sigma_{ij} = 0 \; \forall \; i \neq j$ and
-$\sigma_{ii} = E \left( u_i' u_i \right)$.
+$\sigma_{ii} = \sigma_i^2 = E \left( u_{it}^2 \right)$.
 
@@
 $\sigma_{ij} = 0 \; \forall \; i \neq j$ and
-$\sigma_{ii} = E \left( u_i' u_i \right)$.
+$\sigma_{ii} = \sigma_i^2 = E \left( u_{it}^2 \right)$.
 Like the OLS estimates these estimates are only efficient
@@
 Least Squares (GLS) estimation leads to efficient parameter estimates.
-In this case, the GLS is generally called °Seemingly Unrelated Regression°
-(SUR) \citep{zellner62}.
+In this case, the GLS estimator is generally called °Seemingly Unrelated Regression°
+(SUR) estimator \citep{zellner62}.
 It should be noted that while an unbiased OLS or WLS estimation requires only that 
@@
 equation are uncorrelated
-$( E \left[ u_i | X_i \right] = 0 \; \forall \; i )$,
+$( E \left[ u_i ' X_i \right] = 0 \; \forall \; i )$,
 a consistent SUR estimation requires that all disturbance terms and all 
 regressors are uncorrelated
-$( E \left[ u | X \right] = 0 )$.
+$( E \left[ u_i ' X_j \right] = 0 \; \forall \; i, j )$.
 
@@
 where $\Omega = \Sigma \otimes I$ and
-$\sigma_{ij} = E \left( u_i' u_j \right)$.
+$\sigma_{ij} = E \left( u_{it} u_{jt} \right)$.
 And the covariance matrix of the estimated parameters is
@@
 If the regressors of one or more equations are correlated 
-with the disturbances ($E \left( u_i | X_i \right) \neq 0$), 
-the estimated coefficients are biased.
+with the disturbances ($E \left( u_i' X_i \right) \neq 0$),
+OLS estimates are biased.
 This can be circumvented by an instrumental variable (IV)
@@
 the disturbance terms of the corresponding equation 
-($E \left( u_i | H_i \right) = 0$).
+($E \left( u_i' H_i \right) = 0$).
 
@@
 \end{equation}
-where $\sigma^2 = E \left( u' u \right)$.
+where $\sigma^2 = E \left( u_{it}^2 \right)$.
 If the disturbance terms of the individual equations
@@
 $\sigma_{ij} = 0 \; \forall \; i \neq j$ and
-$\sigma_{ii} = E \left( u_i' u_i \right)$.
+$\sigma_{ii} = \sigma_i^2 = E \left( u_{it}^2 \right)$.
 The covariance matrix of the estimated parameters is
@@
 If the regressors are correlated with the disturbances 
-($E \left( u | X \right) \neq 0$) and 
+($E \left( u_i' X_i \right) \neq 0$) and
 the disturbances are contemporaneously correlated, 
@@
 where $\Omega = \Sigma \otimes I$ and
-$\sigma_{ij} = E \left( u_i' u_j \right)$.
+$\sigma_{ij} = E \left( u_{it} u_{jt} \right)$.
 Its covariance matrix is
@@
 equation are uncorrelated
-$( E \left[ u_i | H_i \right]) = 0 \; \forall \; i )$,
+$( E \left[ u_i' H_i \right]) = 0 \; \forall \; i )$,
 \cite{schmidt90} points out that this estimator is only consistent 
 if all disturbance terms and all instrumental variables are uncorrelated
-$( E \left[ u | H \right]) = 0 )$
+$( E \left[ u_i' H_j \right]) = 0 \; \forall \; i, j )$
 with

### 2006-08-29 paper/trunk/systemfit_intro.tex
corrected the paper according to all minor comments of reviewer A except for "p.3, l.45", the second part of "p.3, l.52" and "p.18, Tab.1".
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@687 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 
-Many theoretical models that are econometrically estimated
-consist of more than one equation.
-The disturbance terms of these equations are likely
+Many theoretical models consist of more than one equation.
+If these models are econometrically estimated,
+the disturbance terms of these equations are likely
 to be contemporaneously correlated,

### 2006-08-29 paper/trunk/systemfit_other.tex
corrected the paper according to all minor comments of reviewer A except for "p.3, l.45", the second part of "p.3, l.52" and "p.18, Tab.1".
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@687 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
    \widehat{u}' ( \widehat{\Sigma} \otimes I )^{-1} \widehat{u} /
-   ( M \cdot T - K )
+   ( G \cdot T - K )
 } \; ,
@@
 where $j$ is the number of restrictions,
-$M$ is the number of equations,
+$G$ is the number of equations,
 $T$ is the number of observations per equation,
@@
 Under the null hypothesis, $F$ has an $F$ distribution
-with $j$ and $M \cdot T - K$ degrees of freedom
+with $j$ and $G \cdot T - K$ degrees of freedom
 \citep[p.\ 314]{theil71}.

### 2006-08-29 paper/trunk/systemfit_restrictions.tex
corrected the paper according to all minor comments of reviewer A except for "p.3, l.45", the second part of "p.3, l.52" and "p.18, Tab.1".
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@687 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 There are two ways to impose linear parameter restrictions.
-First, a matrix $T$ can be specified that
+First, a matrix $M$ can be specified that
 \begin{equation}
-   \beta = T \cdot \beta^* \label{eq:T-restr} ,
+   \beta = M \cdot \beta^* \label{eq:T-restr} ,
 \end{equation}
 where $\beta^*$ is a vector of restricted (linear independent) coefficients,
-and $T$ is a matrix with the number of rows equal to the number of
+and $M$ is a matrix with the number of rows equal to the number of
 unrestricted coefficients ($\beta$) and
@@
 ($\beta^*$).
-$T$ can be used to map each unrestricted coefficient to one or more
+$M$ can be used to map each unrestricted coefficient to one or more
 restricted coefficients.
@@
 To impose these restrictions, the $X$ matrix is
-(post-)\hspace{0pt}multiplied by this $T$ matrix, so that
+(post-)\hspace{0pt}multiplied by this $M$ matrix, so that
 \begin{equation}
-    X^* = X \cdot T .
+    X^* = X \cdot M .
 \end{equation}
@@
    \Cov \left[ \widehat{\beta} \right]
-   = T \cdot \Cov \left[ \widehat{\beta}^* \right] \cdot T' .
+   = M \cdot \Cov \left[ \widehat{\beta}^* \right] \cdot M' .
 \end{equation}
@@
 where $\Omega = \Sigma \otimes I$ and
-$\sigma_{ij} = E \left( u_i' u_j \right)$.
+$\sigma_{ij} = E \left( u_{it} u_{jt} \right)$.
 The covariance matrix of the estimated parameters is
@@
 where $\Omega = \Sigma \otimes I$ and
-$\sigma_{ij} = E \left( u_i' u_j \right)$.
+$\sigma_{ij} = E \left( u_{it} u_{jt} \right)$.
 The covariance matrix of this estimator is

### 2006-08-30 trunk/R/calcRCov.R
argument rcovformula can now be 'noDfCor' (as synonym for '0')
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@206 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@ -19,7 +19,7 @@
    }
    for( i in 1:nEq ) {
       for( j in ifelse( diag, i, 1 ):ifelse( diag, i, nEq ) ) {
-         if( methodRCov == 0 ) {
+         if( methodRCov == 0 || methodRCov == "noDfCor" ) {
             result[ i, j ] <- sum( residi[[i]] * residi[[j]] ) / nObsEq[i]
          } else if( methodRCov == 1 || methodRCov == "geomean" ) {
             result[ i, j ] <- sum( residi[[i]] * residi[[j]] ) /
@@ -42,7 +42,7 @@
             result[ i, j ] <- sum( residi[[i]] * residi[[j]] ) /
                ( nObsEq[i] - max( nCoefEq[i], nCoefEq[j] ) )
          } else {
-            stop( paste( "Argument 'methodRCov' must be either 0, 1,",
+            stop( paste( "Argument 'methodRCov' must be either 0, 'noDfCor', 1,",
                   "'geomean', 2, 'Theil', 3 or 'max'." ) )
          }
       }
@@ -56,14 +56,14 @@
 
 ## Calculate Sigma squared
 .calcSigma2 <- function( resids, methodRCov, nObs, nCoef ) {
-   if( methodRCov == 0 ) {
+   if( methodRCov == 0 || methodRCov == "noDfCor" ) {
       result <- sum( resids^2 ) / nObs
    } else if( methodRCov == 1 || methodRCov == "geomean" ||
       methodRCov == 3 || methodRCov == "max") {
       result <- sum( resids^2 )/ ( nObs - nCoef )
    } else {
       stop( paste( "Sigma^2 can only be calculated if argument",
-         "'methodRCov' is either 0, 1, 'geomean', 3 or 'max'" ) )
+         "'methodRCov' is either 0, 'noDfCor', 1, 'geomean', 3 or 'max'" ) )
    }
 }
 

### 2006-08-30 trunk/R/calcRCov.R
argument rcovformula can no longer have numeric values -> only strings 'noDfCor', 'geomean', 'max', and 'Theil' are allowed
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@207 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@ -19,12 +19,12 @@
    }
    for( i in 1:nEq ) {
       for( j in ifelse( diag, i, 1 ):ifelse( diag, i, nEq ) ) {
-         if( methodRCov == 0 || methodRCov == "noDfCor" ) {
+         if( methodRCov == "noDfCor" ) {
             result[ i, j ] <- sum( residi[[i]] * residi[[j]] ) / nObsEq[i]
-         } else if( methodRCov == 1 || methodRCov == "geomean" ) {
+         } else if( methodRCov == "geomean" ) {
             result[ i, j ] <- sum( residi[[i]] * residi[[j]] ) /
                sqrt( ( nObsEq[i] - nCoefEq[i] ) * ( nObsEq[j] - nCoefEq[j] ) )
-         } else if( methodRCov == 2 || methodRCov == "Theil" ) {
+         } else if( methodRCov == "Theil" ) {
             #result[ i, j ] <- sum( residi[[i]] * residi[[j]] ) /
             #   ( nObsEq[i] - nCoefEq[i] - nCoefEq[j] + sum( diag(
             #   xEq[[i]] %*% solve( crossprod( xEq[[i]] ), tol=solvetol ) %*%
@@ -38,12 +38,12 @@
                solve( crossprod( xEq[[j]] ), tol=solvetol ) %*%
                crossprod( xEq[[j]], xEq[[i]] ) ) ) )
 
-         } else if( methodRCov == 3 || methodRCov == "max" ) {
+         } else if( methodRCov == "max" ) {
             result[ i, j ] <- sum( residi[[i]] * residi[[j]] ) /
                ( nObsEq[i] - max( nCoefEq[i], nCoefEq[j] ) )
          } else {
-            stop( paste( "Argument 'methodRCov' must be either 0, 'noDfCor', 1,",
-                  "'geomean', 2, 'Theil', 3 or 'max'." ) )
+            stop( paste( "Argument 'methodRCov' must be either 'noDfCor',",
+                  "'geomean', 'max', or 'Theil'." ) )
          }
       }
    }
@@ -56,14 +56,13 @@
 
 ## Calculate Sigma squared
 .calcSigma2 <- function( resids, methodRCov, nObs, nCoef ) {
-   if( methodRCov == 0 || methodRCov == "noDfCor" ) {
+   if( methodRCov == "noDfCor" ) {
       result <- sum( resids^2 ) / nObs
[…]

### 2006-08-30 trunk/R/systemfit.R

@@ -32,7 +32,7 @@ systemfit <- function( method,
                         TX=NULL,
                         maxiter=1,
                         tol=1e-5,
-                        methodRCov=1,
+                        methodRCov="geomean",
                         centerResiduals = FALSE,
                         method3sls="GLS",
                         probdfsys=!(is.null(R.restr) & is.null(TX)),
@@ -620,7 +620,7 @@ systemfit <- function( method,
       # first formula from Greene (2003, p. 345) (numerator modified to save memory)
       rtOmega <- .calcXtOmegaInv( x = resids, sigma = rcov, nObsEq = n,
          solvetol = solvetol )
-      yCov <- .calcRCov( Y, methodRCov = 0, nObsEq = n, centered = TRUE,
+      yCov <- .calcRCov( Y, methodRCov = "noDfCor", nObsEq = n, centered = TRUE,
          solvetol = solvetol )
       residCovInv <- solve( rcov, tol = solvetol )
       denominator <- 0
@@ -631,10 +631,10 @@ systemfit <- function( method,
       }
       mcelr2 <- 1 - ( rtOmega %*% resids ) / denominator
 #       # second formula from Greene (2003, p. 345)
-#        yCov <- sum(diag(.calcRCov( Y, methodRCov = 0, nObsEq = n, centered = TRUE,
+#        yCov <- sum(diag(.calcRCov( Y, methodRCov = "noDfCor", nObsEq = n, centered = TRUE,
 #           solvetol = solvetol )))
 #        yCov <- drop( t(Y-mean(Y)) %*% (Y-mean(Y)) / sum(n) )
-#       yCov <- .calcRCov( Y, methodRCov = 1, nObsEq = n, nCoefEq=rep(1,G),
+#       yCov <- .calcRCov( Y, methodRCov = "geomean", nObsEq = n, nCoefEq=rep(1,G),
 #          centered = TRUE, solvetol = solvetol )
 #        mcelr2 <- 1 - G / sum( diag( solve( rcov, tol = solvetol ) * yCov ) )
   } else {

### 2006-08-30 trunk/tests/test_2sls.R

@@ -35,8 +35,8 @@ restr3q[1,1] <-  0.5
 
 # It is not possible to estimate 2SLS with systemfit exactly
 # as EViews does, because EViews uses
-# methodRCov == 1 for the coefficient covariance matrix and
-# methodRCov == 0 for the residual covariance matrix.
+# methodRCov == "geomean" for the coefficient covariance matrix and
+# methodRCov == "noDfCor" for the residual covariance matrix.
 # systemfit uses always the same formulas for both calculations.
 
 ## *************** 2SLS estimation ************************
@@ -57,15 +57,15 @@ fit2sls1p <- systemfit( "2SLS", system, labels, data = Kmenta, inst = inst,
 print( summary( fit2sls1p ) )
 print( round( fit2sls1p$bcov, digits = 6 ) )
 
-## ********************* 2SLS (methodRCov = 0) *****************
+## ********************* 2SLS (methodRCov = "noDfCor" ) *****************
 fit2sls1r <- systemfit( "2SLS", system, labels, data = Kmenta, inst = inst,
-   methodRCov = 0 )
+   methodRCov = "noDfCor" )
 print( summary( fit2sls1r ) )
 print( round( fit2sls1r$bcov, digits = 6 ) )
 
-## *************** 2SLS (methodRCov=0, single.eq.sigma=F) *************
+## *************** 2SLS (methodRCov="noDfCor", single.eq.sigma=F) *************
 fit2sls1rs <- systemfit( "2SLS", system, labels, data = Kmenta, inst = inst,
-   methodRCov = 0, single.eq.sigma = FALSE )
+   methodRCov = "noDfCor", single.eq.sigma = FALSE )
 print( summary( fit2sls1rs ) )
 print( round( fit2sls1rs$bcov, digits = 6 ) )
 
@@ -88,28 +88,28 @@ fit2sls2p <- systemfit( "2SLS", system, labels, data = Kmenta, R.restr = restrm,
 print( summary( fit2sls2p ) )
 print( round( fit2sls2p$bcov, digits = 6 ) )
 
-## ********************* 2SLS with restriction (methodRCov = 0) **************
+## ********************* 2SLS with restriction (methodRCov = "noDfCor") **************
[…]

### 2006-08-30 trunk/tests/test_3sls.R

@@ -50,13 +50,13 @@ for( i in seq( along = formulas ) ) {
 
    print( "********************* 3SLS EViews-like *****************" )
    fit3sls[[ i ]]$e1e <- systemfit( "3SLS", system, labels, data = Kmenta,
-      inst = inst, methodRCov = 0, probdfsys = TRUE, method3sls = formulas[ i ] )
+      inst = inst, methodRCov = "noDfCor", probdfsys = TRUE, method3sls = formulas[ i ] )
    print( summary( fit3sls[[ i ]]$e1e ) )
    print( round( fit3sls[[ i ]]$e1e$bcov, digits = 6 ) )
 
-   print( "********************* 3SLS with methodRCov = 2 *****************" )
+   print( "********************* 3SLS with methodRCov = Theil *****************" )
    fit3sls[[ i ]]$e1c <- systemfit( "3SLS", system, labels, data = Kmenta,
-      inst = inst, methodRCov = 2, probdfsys = TRUE, method3sls = formulas[ i ] )
+      inst = inst, methodRCov = "Theil", probdfsys = TRUE, method3sls = formulas[ i ] )
    print( summary( fit3sls[[ i ]]$e1c ) )
    print( round( fit3sls[[ i ]]$e1c$bcov, digits = 6 ) )
 
@@ -68,7 +68,7 @@ for( i in seq( along = formulas ) ) {
 
    print( "************** 3SLS with restriction (EViews-like) *****************" )
    fit3sls[[ i ]]$e2e <- systemfit( "3SLS", system, labels, data = Kmenta,
-      inst = inst, methodRCov = 0, probdfsys = TRUE, R.restr = restrm,
+      inst = inst, methodRCov = "noDfCor", probdfsys = TRUE, R.restr = restrm,
       method3sls = formulas[ i ] )
    print( summary( fit3sls[[ i ]]$e2e ) )
    print( round( fit3sls[[ i ]]$e2e$bcov, digits = 6 ) )
@@ -81,7 +81,7 @@ for( i in seq( along = formulas ) ) {
 
    print( "*************** 3SLS with restriction via TX (EViews-like) *******" )
    fit3sls[[ i ]]$e3e <- systemfit( "3SLS", system, labels, data = Kmenta,
-      inst = inst, methodRCov = 0, probdfsys = TRUE, TX = tc,
[…]

### 2006-08-30 trunk/tests/test_ols.R

@@ -32,8 +32,8 @@ restr3q[1,1] <-  0.5
 
 # It is not possible to estimate OLS with systemfit
 # exactly as EViews does, because EViews uses
-# methodRCov == 1 for the coefficient covariance matrix and
-# methodRCov == 0 for the residual covariance matrix, while
+# methodRCov == "geomean" for the coefficient covariance matrix and
+# methodRCov == "noDfCor" for the residual covariance matrix, while
 # systemfit uses always the same formulas for both calculations.
 
 ## *************** OLS estimation ************************
@@ -54,33 +54,33 @@ fitols1p <- systemfit( "OLS", system, labels, data = Kmenta,
 print( summary( fitols1p ) )
 print( round( fitols1p$bcov, digits = 6 ) )
 
-## ****************  OLS (methodRCov=0) ***********************
+## ****************  OLS (methodRCov="noDfCor") ***********************
 fitols1r <- systemfit( "OLS", system, labels, data = Kmenta,
-   methodRCov = 0 )
+   methodRCov = "noDfCor" )
 print( summary( fitols1r ) )
 print( round( fitols1r$bcov, digits = 6 ) )
 
-## ********  OLS (methodRCov=0, single.eq.sigma=F) ***********
+## ********  OLS (methodRCov="noDfCor", single.eq.sigma=F) ***********
 fitols1rs <- systemfit( "OLS", system, labels, data = Kmenta,
-   methodRCov = 0, single.eq.sigma = FALSE )
+   methodRCov = "noDfCor", single.eq.sigma = FALSE )
 print( summary( fitols1rs ) )
 print( round( fitols1rs$bcov, digits = 6 ) )
 
-## ****************  OLS (methodRCov=2 ) ***********************
+## ****************  OLS (methodRCov="Theil" ) ***********************
 fitols1r <- systemfit( "OLS", system, labels, data = Kmenta,
-   methodRCov = 2 )
+   methodRCov = "Theil" )
 print( summary( fitols1r ) )
 print( round( fitols1r$bcov, digits = 6 ) )
 
-## ****************  OLS (methodRCov=3) ***********************
+## ****************  OLS (methodRCov="max") ***********************
[…]

### 2006-08-30 trunk/tests/test_sur.R

@@ -42,24 +42,24 @@ print( summary( fitsur1 ) )
 print( round( fitsur1$bcov, digits = 6 ) )
 
 ## ********************* SUR (EViews-like) *****************
-fitsur1e <- systemfit( "SUR", system, labels, data = Kmenta, methodRCov = 0,
+fitsur1e <- systemfit( "SUR", system, labels, data = Kmenta, methodRCov = "noDfCor",
    probdfsys = TRUE )
 print( summary( fitsur1e ) )
 print( round( fitsur1e$bcov, digits = 6 ) )
 
-## ********************* SUR (methodRCov=2) *****************
-fitsur1c <- systemfit( "SUR", system, labels, data = Kmenta, methodRCov = 2 )
+## ********************* SUR (methodRCov="Theil") *****************
+fitsur1c <- systemfit( "SUR", system, labels, data = Kmenta, methodRCov = "Theil" )
 print( summary( fitsur1c ) )
 print( round( fitsur1c$bcov, digits = 6 ) )
 
-## *************** SUR (methodRCov=2, probdfsys = TRUE ) ***************
-fitsur1cp <- systemfit( "SUR", system, labels, data = Kmenta, methodRCov = 2,
+## *************** SUR (methodRCov="Theil", probdfsys = TRUE ) ***************
+fitsur1cp <- systemfit( "SUR", system, labels, data = Kmenta, methodRCov = "Theil",
    probdfsys = TRUE )
 print( summary( fitsur1cp ) )
 print( round( fitsur1cp$bcov, digits = 6 ) )
 
-## ********************* SUR (methodRCov=3) *****************
-fitsur1c <- systemfit( "SUR", system, labels, data = Kmenta, methodRCov = 3 )
+## ********************* SUR (methodRCov="max") *****************
+fitsur1c <- systemfit( "SUR", system, labels, data = Kmenta, methodRCov = "max" )
 print( summary( fitsur1c ) )
 print( round( fitsur1c$bcov, digits = 6 ) )
 
@@ -70,7 +70,7 @@ print( round( fitsur2$bcov, digits = 6 ) )
 
 ## *************** SUR with cross-equation restriction (EViews-like) **
 fitsur2e <- systemfit( "SUR", system, labels, data = Kmenta, R.restr = restrm,
-   methodRCov = 0 )
+   methodRCov = "noDfCor" )
 print( summary( fitsur2e ) )
[…]

### 2006-08-30 trunk/tests/test_w2sls.R

@@ -41,7 +41,7 @@ print( round( fitw2sls1$bcov, digits = 6 ) )
 
 ## ********************* W2SLS (EViews-like) *****************
 fitw2sls1e <- systemfit( "W2SLS", system, labels, data = Kmenta, inst = inst,
-   methodRCov = 0, probdfsys = TRUE )
+   methodRCov = "noDfCor", probdfsys = TRUE )
 print( summary( fitw2sls1e ) )
 print( round( fitw2sls1e$bcov, digits = 6 ) )
 
@@ -53,7 +53,7 @@ print( round( fitw2sls2$bcov, digits = 6 ) )
 
 ## ********************* W2SLS with restriction (EViews-like) **************
 fitw2sls2e <- systemfit( "W2SLS", system, labels, data = Kmenta, R.restr = restrm,
-   inst = inst, methodRCov = 0, probdfsys = TRUE )
+   inst = inst, methodRCov = "noDfCor", probdfsys = TRUE )
 print( summary( fitw2sls2e ) )
 print( round( fitw2sls2e$bcov, digits = 6 ) )
 
@@ -64,7 +64,7 @@ print( round( fitw2sls3$bcov, digits = 6 ) )
 
 ## ********************* W2SLS with restriction via TX (EViews-like) **************
 fitw2sls3e <- systemfit( "W2SLS", system, labels, data = Kmenta, TX = tc, inst = inst,
-   methodRCov = 0, probdfsys = TRUE )
+   methodRCov = "noDfCor", probdfsys = TRUE )
 print( summary( fitw2sls3e ) )
 print( round( fitw2sls3e$bcov, digits = 6 ) )
 
@@ -76,7 +76,7 @@ print( round( fitw2sls4$bcov, digits = 6 ) )
 
 ## ***************** W2SLS with 2 restrictions (EViews-like) **************
 fitw2sls4e <- systemfit( "W2SLS", system, labels, data = Kmenta, R.restr = restr2m,
-   q.restr = restr2q, inst = inst, methodRCov = 0, probdfsys = TRUE )
+   q.restr = restr2q, inst = inst, methodRCov = "noDfCor", probdfsys = TRUE )
 print( summary( fitw2sls4e ) )
 print( round( fitw2sls4e$bcov, digits = 6 ) )
 
@@ -88,7 +88,7 @@ print( round( fitw2sls5$bcov, digits = 6 ) )
 
[…]

### 2006-08-30 trunk/tests/test_wls.R

@@ -37,7 +37,7 @@ print( summary( fitwls1 ) )
 print( round( fitwls1$bcov, digits = 6 ) )
 
 ## *************** WLS estimation (EViews-like) ************************
-fitwls1e <- systemfit( "WLS", system, labels, data = Kmenta, methodRCov = 0,
+fitwls1e <- systemfit( "WLS", system, labels, data = Kmenta, methodRCov = "noDfCor",
    probdfsys = TRUE )
 print( summary( fitwls1e ) )
 print( round( fitwls1e$bcov, digits = 6 ) )
@@ -49,7 +49,7 @@ print( round( fitwls2$bcov, digits = 6 ) )
 
 ## ************** WLS with cross-equation restriction (EViews-like) *******
 fitwls2e <- systemfit( "WLS", system, labels, data = Kmenta, R.restr = restrm,
-   methodRCov = 0 )
+   methodRCov = "noDfCor" )
 print( summary( fitwls2e ) )
 print( round( fitwls2e$bcov, digits = 6 ) )
 
@@ -60,7 +60,7 @@ print( round( fitwls3$bcov, digits = 6 ) )
 
 ## ******* WLS with cross-equation restriction via TX (EViews-like) *****
 fitwls3e <- systemfit("WLS", system, labels, data = Kmenta, TX = tc,
-   methodRCov = 0 )
+   methodRCov = "noDfCor" )
 print( summary( fitwls3e ) )
 print( round( fitwls3e$bcov, digits = 6 ) )
 
@@ -71,7 +71,7 @@ print( summary( fitwls4 ) )
 print( round( fitwls4$bcov, digits = 6 ) )
 
 ## ***** WLS with 2 cross-equation restrictions (EViews-like) **********
-fitwls4e <- systemfit("WLS", system, labels, data = Kmenta, methodRCov = 0,
+fitwls4e <- systemfit("WLS", system, labels, data = Kmenta, methodRCov = "noDfCor",
    R.restr = restr2m, q.restr = restr2q )
 print( summary( fitwls4e ) )
 print( round( fitwls4e$bcov, digits = 6 ) )
@@ -83,7 +83,7 @@ print( summary( fitwls5 ) )
 print( round( fitwls5$bcov, digits = 6 ) )
 
 ## *********** WLS with 2 cross-equation restrictions via R and TX (EViews-like)
[…]

### 2006-08-30 paper/trunk/systemfit_other.tex
argument rcovformula is no longer allowed to have numeric values -> only strings 'noDfCor', 'geomean', 'max', and 'Theil' are allowed (rev. 207)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@689 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
    \widehat{\sigma}_{ij} = \frac{ \widehat{u}_i' \widehat{u}_j }{ T }
-   \label{eq:rcov0}
+   \label{eq:rcovNoDfCor}
 \end{equation}
@@
    { \sqrt{ \left( T - K_i \right) \cdot \left( T - K_j \right) } }
-   \label{eq:rcov1}
+   \label{eq:rcovGeomean}
 \end{equation}
@@
    { T - \max \left( K_i , K_j \right) } \; ,
-   \label{eq:rcov3}
+   \label{eq:rcovMax}
 \end{equation}
@@
    X_i' X_j \left( X_j' X_j \right)^{-1} X_j' \right] }
-   \label{eq:rcov2} \\
+   \label{eq:rcovTheil} \\
    & = &
@@
 of $\Sigma^{-1}$ \citep[p.\ 322]{theil71}.
-Furthermore, the covariance matrix calculated by (\ref{eq:rcov2})
+Furthermore, the covariance matrix calculated by (\ref{eq:rcovTheil})
 is not necessarily positive semidefinite \citep[p.\ 322]{theil71}. 
 Hence, °it is doubtful whether [this formula] is really superior to 
-[(\ref{eq:rcov0})]° \citep[p.\ 322]{theil71}.
+[(\ref{eq:rcovNoDfCor})]° \citep[p.\ 322]{theil71}.
 
@@
 from the residuals of the previous iteration.
-If equation (\ref{eq:rcov0}) is applied to calculate the estimated
+If equation (\ref{eq:rcovNoDfCor}) is applied to calculate the estimated
 residual covariance matrix,
@@
 and substituting $\widehat{u}_i - \overline{ \widehat{u} }_i$
-for $\widehat{u}_i$ in (\ref{eq:rcov0}--\ref{eq:rcov2}).
+for $\widehat{u}_i$ in (\ref{eq:rcovNoDfCor}--\ref{eq:rcovTheil}).
 
@@
 $\widehat{\Sigma}_r$ and $\widehat{\Sigma}_u$ are
-the residual covariance matrices calculated by formula (\ref{eq:rcov0})
+the residual covariance matrices calculated by formula (\ref{eq:rcovNoDfCor})
 of the restricted and unrestricted estimation, respectively.

### 2006-08-30 paper/trunk/systemfit.tex
added footer for nested docs
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@632 591db2bd-e0a0-4d22-8c19-2192dbb7da31
changes according to the first major comment of reviewer A: modified title and a few sentences (package to estimate -> package for estimating; equation system -> system of equations)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@691 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 
-\title{systemfit: A Package to Estimate\\
-       Simultaneous Equation Systems in \proglang{R}}
+\title{systemfit: A Package for Estimating\\
+       Systems of Simultaneous Equations in \proglang{R}}
 
@@
 \Plainauthor{Arne Henningsen, Jeff D.\ Hamann} %% comma-separated
-\Plaintitle{systemfit: A Package to Estimate Simultaneous Equation Systems 
+\Plaintitle{systemfit: A Package for Estimating Systems of Simultaneous Equations
             in R} %% without formatting
@@
 %% an abstract and keywords
-\Abstract{ 
-Many statistical analyses are based on models containing systems of
-structurally related equations.
-In cases where cross-equation disturbances are correlated,
-full information methods are required \citep{zellner62}.
-If exogenous variables are stochastically dependent
-on the disturbances in the system,
-then instrumental variable estimation methods should be used
-\citep{zellner62b}
-The package \pkg{systemfit} provides the capability
-to estimate systems of linear equations
-within the \proglang{R} programming environment.
-}
-
-\Keywords{R, simultaneous equations systems,
-  seemingly unrelated regression, two-stage least squares,
-  three-stage least squares, parameter restriction}
-
-%% at least one keyword must be supplied
+\input{systemfit_abstract.tex}
 
@@
 %% \Month{September}
-%% \Year{2004}
-%% \Submitdate{2004-09-29}
-%% \Acceptdate{2004-09-29}
+%% \Year{2006}
+%% \Submitdate{2006-02-28}
+%% \Acceptdate{2006-09-29}
 
@@
 
-% \section{CVS Revision Log}
-% \begin{verbatim}
-% $Log$
-% Revision 1.19  2005/11/08 06:31:48  hamannj
-% made seemingly massive changes... but hopefully not for the worse...
-%
-% Revision 1.18  2005/11/07 14:08:56  henningsena
-% a few minor changes (predominantly adding a "\ " after a period (.) that does not indicate an end of a sentence)
-%
-% Revision 1.17  2005/11/07 10:21:59  henningsena
-% just changed some line breaks (deleted new ones and restored deleted ones) to make it easier to find the changes that Jeff committed in the last revision(s)
-%
-% Revision 1.16  2005/11/02 18:31:45  hamannj
-% small last changes
-%
-% Revision 1.15  2005/11/02 18:15:40  hamannj
-% small changes
-
-% Revision 1.14  2005/10/26 21:16:53  hamannj
-% checked out and updated local copy before edits
-
-% Revision 1.13  2005/09/26 20:34:12  henningsena
-% using now package csquotes to have uniform and correct quotation marks (BibTeX-Style jss.bst has to be modified)
-
-% Revision 1.12  2005/09/26 11:24:05  henningsena
-% some minor changes
-
-% Revision 1.11  2005/09/01 12:27:55  henningsena
-% revised structure of the document (sections, subsections, subsubsections) due to removal of nlsystemfit; moved each section into a separate file.
-
-% Revision 1.10  2005/09/01 11:49:02  henningsena
-% moved section about non-linear estimation from systemfit.tex into new file nlsystemfit.tex, because this section should not be included in (this version of) the article
-
-% Revision 1.9  2004/11/16 10:59:09  henningsena
-% some minor changes and some comments on nlsystemfit
-
-% Revision 1.8  2004/11/16 08:55:53  henningsena
-% merged file and resolved conflict due to some minor changes I did while Jeff was adding nlsystemfit.
-
-% Revision 1.7  2004/11/16 05:34:47  hamannj
-% question about nlsystemfit code section (example)
-
-% Revision 1.6  2004/11/16 05:32:46  hamannj
-% made small changes because my code section wouldn't compile. argh.
-
-% Revision 1.5  2004/11/16 05:31:19  hamannj
-% added cvs header and log section. to be removed when submitted.
-
-% \end{verbatim}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \end{document}
+
+%%% Local Variables: 
+%%% mode: latex
+%%% TeX-master: t
+%%% End:

### 2006-08-30 paper/trunk/systemfit_intro.tex
changes according to the first major comment of reviewer A: modified title and a few sentences (package to estimate -> package for estimating; equation system -> system of equations)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@691 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 °Seemingly Unrelated Regression° (SUR) \citep{zellner62}.
-Another reason to estimate an equation system simultaneously are
+Another reason to estimate a system of equations simultaneously are
 cross-equation parameter restrictions.%
@@
 The \pkg{systemfit} package provides the capability to estimate
-linear equation systems in \proglang{R}
+systems of linear equations in \proglang{R}
 \citep{r-project}.
-Although linear equation systems can be estimated
+Although systems of linear equations can be estimated
 with several other statistical and econometric software packages

### 2006-08-30 paper/trunk/systemfit_summary.tex
corrected a typo in the summary
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@690 591db2bd-e0a0-4d22-8c19-2192dbb7da31
changes according to the first major comment of reviewer A: modified title and a few sentences (package to estimate -> package for estimating; equation system -> system of equations)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@691 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 In this article, we have described some of the basic features of the
-\pkg{systemfit} package for estimation of linear systems of
+\pkg{systemfit} package for estimating of systems of linear
 equations.
@@
 Finally, the \pkg{systemfit} package provides a function to estimate
-systems of non-linear estimations.
+systems of non-linear equations.
 However, the function \code{nlsystemfit} is currently under

### 2006-08-30 trunk/R/systemfitClassic.R
changed order of arguments in systemfit() and systemfitClassic() and added default value for argument "method": method, eqns|formula ->  eqns|formula, method = "OLS" (suggested by of reviewer A of the JSS paper)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@208 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@ -1,4 +1,4 @@
-systemfitClassic <- function( method, formula, eqnVar, timeVar, data,
+systemfitClassic <- function( formula, method = "OLS", eqnVar, timeVar, data,
    pooled = FALSE, ... ) {
 
    eqnLabels <- levels( as.factor( data[[ eqnVar ]] ) )
@@ -34,7 +34,7 @@ systemfitClassic <- function( method, formula, eqnVar, timeVar, data,
       }
    }
 
-   result <- systemfit( method = method, eqns = eqnSystem,
+   result <- systemfit( eqns = eqnSystem, method = method,
       eqnlabels = eqnLabels, data = wideData, TX = TX, ... )
 
    return( result )

### 2006-08-30 trunk/tests/test_3sls.R

@@ -43,72 +43,72 @@ for( i in seq( along = formulas ) ) {
    print( "***************************************************" )
    print( paste( "3SLS formula:", formulas[ i ] ) )
    print( "************* 3SLS *********************************" )
-   fit3sls[[ i ]]$e1 <- systemfit( "3SLS", system, labels, data = Kmenta,
+   fit3sls[[ i ]]$e1 <- systemfit( system, "3SLS", labels, data = Kmenta,
       inst = inst, method3sls = formulas[ i ] )
    print( summary( fit3sls[[ i ]]$e1 ) )
    print( round( fit3sls[[ i ]]$e1$bcov, digits = 6 ) )
 
    print( "********************* 3SLS EViews-like *****************" )
-   fit3sls[[ i ]]$e1e <- systemfit( "3SLS", system, labels, data = Kmenta,
+   fit3sls[[ i ]]$e1e <- systemfit( system, "3SLS", labels, data = Kmenta,
       inst = inst, methodRCov = "noDfCor", probdfsys = TRUE, method3sls = formulas[ i ] )
    print( summary( fit3sls[[ i ]]$e1e ) )
    print( round( fit3sls[[ i ]]$e1e$bcov, digits = 6 ) )
 
    print( "********************* 3SLS with methodRCov = Theil *****************" )
-   fit3sls[[ i ]]$e1c <- systemfit( "3SLS", system, labels, data = Kmenta,
+   fit3sls[[ i ]]$e1c <- systemfit( system, "3SLS", labels, data = Kmenta,
       inst = inst, methodRCov = "Theil", probdfsys = TRUE, method3sls = formulas[ i ] )
    print( summary( fit3sls[[ i ]]$e1c ) )
    print( round( fit3sls[[ i ]]$e1c$bcov, digits = 6 ) )
 
    print( "*************** 3SLS with restriction *****************" )
-   fit3sls[[ i ]]$e2 <- systemfit( "3SLS", system, labels, data = Kmenta,
+   fit3sls[[ i ]]$e2 <- systemfit( system, "3SLS", labels, data = Kmenta,
       inst = inst, R.restr = restrm, method3sls = formulas[ i ] )
    print( summary( fit3sls[[ i ]]$e2 ) )
    print( round( fit3sls[[ i ]]$e2$bcov, digits = 6 ) )
 
[…]

### 2006-08-30 trunk/tests/test_ols.R

@@ -38,48 +38,48 @@ restr3q[1,1] <-  0.5
 
 ## *************** OLS estimation ************************
 ## ********** OLS estimation (default) ********************
-fitols1 <- systemfit( "OLS", system, labels, data = Kmenta )
+fitols1 <- systemfit( system, "OLS", labels, data = Kmenta )
 print( summary( fitols1 ) )
 print( round( fitols1$bcov, digits = 6 ) )
 
 ## ********** OLS estimation (no single.eq sigma=F) ******************
-fitols1s <- systemfit( "OLS", system, labels, data = Kmenta,
+fitols1s <- systemfit( system, "OLS", labels, data = Kmenta,
    single.eq.sigma = FALSE )
 print( summary( fitols1s ) )
 print( round( fitols1s$bcov, digits = 6 ) )
 
 ## ****************  OLS (probdfsys=T) ***********************
-fitols1p <- systemfit( "OLS", system, labels, data = Kmenta,
+fitols1p <- systemfit( system, "OLS", labels, data = Kmenta,
    probdfsys = TRUE )
 print( summary( fitols1p ) )
 print( round( fitols1p$bcov, digits = 6 ) )
 
 ## ****************  OLS (methodRCov="noDfCor") ***********************
-fitols1r <- systemfit( "OLS", system, labels, data = Kmenta,
+fitols1r <- systemfit( system, "OLS", labels, data = Kmenta,
    methodRCov = "noDfCor" )
 print( summary( fitols1r ) )
 print( round( fitols1r$bcov, digits = 6 ) )
 
 ## ********  OLS (methodRCov="noDfCor", single.eq.sigma=F) ***********
-fitols1rs <- systemfit( "OLS", system, labels, data = Kmenta,
+fitols1rs <- systemfit( system, "OLS", labels, data = Kmenta,
    methodRCov = "noDfCor", single.eq.sigma = FALSE )
 print( summary( fitols1rs ) )
 print( round( fitols1rs$bcov, digits = 6 ) )
 
 ## ****************  OLS (methodRCov="Theil" ) ***********************
-fitols1r <- systemfit( "OLS", system, labels, data = Kmenta,
+fitols1r <- systemfit( system, "OLS", labels, data = Kmenta,
    methodRCov = "Theil" )
 print( summary( fitols1r ) )
 print( round( fitols1r$bcov, digits = 6 ) )
 
[…]

### 2006-08-30 trunk/tests/test_sur.R

@@ -37,175 +37,175 @@ system2 <- list( demand, supply2 )
 
 
 ## *************** SUR estimation ************************
-fitsur1 <- systemfit( "SUR", system, labels, data = Kmenta )
+fitsur1 <- systemfit( system, "SUR", labels, data = Kmenta )
 print( summary( fitsur1 ) )
 print( round( fitsur1$bcov, digits = 6 ) )
 
 ## ********************* SUR (EViews-like) *****************
-fitsur1e <- systemfit( "SUR", system, labels, data = Kmenta, methodRCov = "noDfCor",
+fitsur1e <- systemfit( system, "SUR", labels, data = Kmenta, methodRCov = "noDfCor",
    probdfsys = TRUE )
 print( summary( fitsur1e ) )
 print( round( fitsur1e$bcov, digits = 6 ) )
 
 ## ********************* SUR (methodRCov="Theil") *****************
-fitsur1c <- systemfit( "SUR", system, labels, data = Kmenta, methodRCov = "Theil" )
+fitsur1c <- systemfit( system, "SUR", labels, data = Kmenta, methodRCov = "Theil" )
 print( summary( fitsur1c ) )
 print( round( fitsur1c$bcov, digits = 6 ) )
 
 ## *************** SUR (methodRCov="Theil", probdfsys = TRUE ) ***************
-fitsur1cp <- systemfit( "SUR", system, labels, data = Kmenta, methodRCov = "Theil",
+fitsur1cp <- systemfit( system, "SUR", labels, data = Kmenta, methodRCov = "Theil",
    probdfsys = TRUE )
 print( summary( fitsur1cp ) )
 print( round( fitsur1cp$bcov, digits = 6 ) )
 
 ## ********************* SUR (methodRCov="max") *****************
-fitsur1c <- systemfit( "SUR", system, labels, data = Kmenta, methodRCov = "max" )
+fitsur1c <- systemfit( system, "SUR", labels, data = Kmenta, methodRCov = "max" )
 print( summary( fitsur1c ) )
 print( round( fitsur1c$bcov, digits = 6 ) )
 
 ## *************** SUR with cross-equation restriction **************
-fitsur2 <- systemfit( "SUR", system, labels, data = Kmenta, R.restr = restrm )
+fitsur2 <- systemfit( system, "SUR", labels, data = Kmenta, R.restr = restrm )
 print( summary( fitsur2 ) )
[…]

### 2006-08-31 trunk/R/calcGLS.R
renamed some arguments of .calcXtOmegaInv() and .calcGLS(): y -> yVec, x -> xMat, x2 -> xMat2
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@222 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@ -1,42 +1,42 @@
-.calcXtOmegaInv <- function( x, sigma, nObsEq, solvetol = 1e-5, invertSigma = TRUE ){
+.calcXtOmegaInv <- function( xMat, sigma, nObsEq, solvetol = 1e-5, invertSigma = TRUE ){
    nEq <- length( nObsEq )
    if( invertSigma ) {
       sigmaInv <- solve( sigma, tol = solvetol )
    } else {
       sigmaInv <- sigma
    }
-   eqSelect <- rep( 0, nrow( x ) )
+   eqSelect <- rep( 0, nrow( xMat ) )
    for( i in 1:nEq ) {
       eqSelect[ ( sum( nObsEq[ 0:( i - 1 ) ] ) + 1 ):sum( nObsEq[ 1:i ] ) ] <- i
    }
-   result <- matrix( 0, nrow = ncol( x ), ncol = nrow( x ) )
+   result <- matrix( 0, nrow = ncol( xMat ), ncol = nrow( xMat ) )
    for( i in 1:nEq ) {
       for( j in 1:nEq ) {
          result[ , eqSelect == i ] <- result[ , eqSelect == i ] +
-            t( x )[ , eqSelect == j ] * sigmaInv[ i, j ]
+            t( xMat )[ , eqSelect == j ] * sigmaInv[ i, j ]
       }
    }
    return( result )
 }
 
-.calcGLS <- function( x, y = NULL, x2 = x, R.restr = NULL, q.restr = NULL, 
+.calcGLS <- function( xMat, yVec = NULL, xMat2 = xMat, R.restr = NULL, q.restr = NULL,
       sigma, nObsEq, solvetol = 1e-5 ){
-   xtOmegaInv <- .calcXtOmegaInv( x = x, sigma = sigma, nObsEq = nObsEq, 
+   xtOmegaInv <- .calcXtOmegaInv( xMat = xMat, sigma = sigma, nObsEq = nObsEq,
       solvetol = solvetol )
    if( is.null( R.restr ) ) {
-      result <- solve( xtOmegaInv %*% x2, tol = solvetol )
-      if( !is.null( y ) ) {
-         result <- result %*% xtOmegaInv %*% y
+      result <- solve( xtOmegaInv %*% xMat2, tol = solvetol )
+      if( !is.null( yVec ) ) {
+         result <- result %*% xtOmegaInv %*% yVec
       }
    } else {
-      W <- rbind( cbind( xtOmegaInv %*% x2, t(R.restr) ),
[…]

### 2006-08-31 trunk/R/systemfit.R

@@ -195,7 +195,7 @@ systemfit <- function(  eqns,
       rcov <- .calcRCov( resids, methodRCov = methodRCov, nObsEq = nObsEq,
          nCoefEq = nExogLiEq, xEq = xMatEq, diag = TRUE, centered = centerResiduals,
          solvetol = solvetol )               # residual covariance matrix
-      bcov <- .calcGLS( x = xMatAll, R.restr = R.restr, q.restr = q.restr,
+      bcov <- .calcGLS( xMat = xMatAll, R.restr = R.restr, q.restr = q.restr,
          sigma = rcov, nObsEq = nObsEq, solvetol = solvetol )
                     # coefficient covariance matrix
     } else {
@@ -223,11 +223,11 @@ systemfit <- function(  eqns,
       rcov <- .calcRCov( resids, methodRCov = methodRCov, nObsEq = nObsEq,
          nCoefEq = nExogLiEq, xEq = xMatEq, diag = TRUE, centered = centerResiduals,
          solvetol = solvetol )
-      b  <- .calcGLS( x = xMatAll, y = yVecAll, R.restr = R.restr, q.restr = q.restr,
+      b  <- .calcGLS( xMat = xMatAll, yVec = yVecAll, R.restr = R.restr, q.restr = q.restr,
          sigma = rcov, nObsEq = nObsEq, solvetol = solvetol ) # coefficients
       bdif <- b-bl # difference of coefficients between this and previous step
     }
-    bcov <- .calcGLS( x = xMatAll, R.restr = R.restr, q.restr = q.restr,
+    bcov <- .calcGLS( xMat = xMatAll, R.restr = R.restr, q.restr = q.restr,
        sigma = rcov, nObsEq = nObsEq, solvetol = solvetol )
        # final step coefficient covariance matrix
     resids <- yVecAll - xMatAll %*% b                        # residuals
@@ -246,11 +246,11 @@ systemfit <- function(  eqns,
       rcov <- .calcRCov( resids, methodRCov = methodRCov, nObsEq = nObsEq,
          nCoefEq = nExogLiEq, xEq = xMatEq, centered = centerResiduals,
          solvetol = solvetol )
-      b <- .calcGLS( x = xMatAll, y = yVecAll, R.restr = R.restr, q.restr = q.restr,
[…]

### 2006-09-29 trunk/R/confint.systemfit.R
renamed variable for coefficients returned by systemfit: x$coef -> x$coefficients, x$eq[[ i ]]$coef -> x$eq[[ i ]]$coefficients
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@332 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@ -10,8 +10,8 @@ confint.systemfit <- function( object, parm = NULL, level = 0.95,
    probLower <- ( 1 - level ) / 2
    probBoth <- c( probLower, 1 - probLower )
    pct <- paste( round( 100 * probBoth, 1 ), "%" )
-   ci <- matrix( NA, length( object$coef ), 2,
-            dimnames = list( names( object$coef ), pct ) )
+   ci <- matrix( NA, length( object$coefficients ), 2,
+            dimnames = list( names( object$coefficients ), pct ) )
    j <- 1
    for( i in 1:length( object$eq ) ) {
       object$eq[[i]]$df.residual.sys <- object$df.residual
@@ -35,8 +35,8 @@ confint.systemfit.equation <- function( object, parm = NULL, level = 0.95,
    probLower <- ( 1 - level ) / 2
    probBoth <- c( probLower, 1 - probLower )
    pct <- paste( round( 100 * probBoth, 1 ), "%" )
-   ci <- matrix( NA, length( object$coef ), 2,
-            dimnames = list( names( object$coef ), pct ) )
+   ci <- matrix( NA, length( object$coefficients ), 2,
+            dimnames = list( names( object$coefficients ), pct ) )
    if( useDfSys ) {
       fac <- qt( probBoth, object$df.residual.sys )
    } else {

### 2006-09-29 trunk/R/fiml.R

@@ -59,9 +59,9 @@
    mlResult$optim <- optim( startCoef, .systemfitFimlLik, method = "BFGS",
       control = list( fnscale = -1 ), mlVars = mlVars )
    # methods: "BFGS" "Nelder-Mead"
-   mlResult$coef <- mlResult$optim$par
-   mlResult$coefCov <- diag( 1, length( mlResult$coef ) )
-   mlResult$resids  <- yVec - xMat %*% mlResult$coef
+   mlResult$coefficients <- mlResult$optim$par
+   mlResult$coefCov <- diag( 1, length( mlResult$coefficients ) )
+   mlResult$resids  <- yVec - xMat %*% mlResult$coefficients
 
    return( mlResult )
 }

### 2006-09-29 trunk/R/predict.systemfit.R

@@ -46,7 +46,7 @@ predict.systemfit.equation <- function( object, newdata = NULL,
    }
 
    # fitted values
-   predicted <- data.frame( fit = drop( xMat %*% object$coef ) )
+   predicted <- data.frame( fit = drop( xMat %*% object$coefficients ) )
 
    # calculate variance covariance matrices
    if( se.fit | interval == "confidence" ) {

### 2006-09-29 trunk/R/print.systemfit.R

@@ -20,7 +20,7 @@ print.systemfit <- function( x, digits=6,... ) {
       }
    }
    cat( "Coefficients:\n" )
-   print( x$coef )
+   print( x$coefficients )
    invisible( x )
 }
 
@@ -43,6 +43,6 @@ print.systemfit.equation <- function( x, digits=6, ... ) {
    }
 
    cat("\nCoefficients:")
-   print( x$coef )
+   print( x$coefficients )
    invisible( x )
 }

### 2006-09-29 trunk/R/smallMethods.R

@@ -40,7 +40,7 @@ se.ratio.systemfit <- function( resultsi, resultsj, eqni ) {
 
 ## return all coefficients
 coef.systemfit <- function( object, ... ) {
-   object$coef
+   object$coefficients
 }
 
 ## return all coefficients, std.errors, t-values and p-values
@@ -50,12 +50,12 @@ coef.summary.systemfit <- function( object, ... ) {
 
 ## return the coefficients of a single equation
 coef.systemfit.equation <- function( object, ... ) {
-   object$coef
+   object$coefficients
 }
 
 ## return coefficients, std.errors, t-values and p-values of a single equation
 coef.summary.systemfit.equation <- function( object, ... ) {
-   object$coef
+   object$coefficients
 }
 
 ## return all residuals

### 2006-09-29 trunk/R/summary.systemfit.R

@@ -25,7 +25,7 @@ summary.systemfit <- function( object, useDfSys = NULL, ... ) {
 
    # coefficients, standard errors, ... 
    result$coefCov <- object$bcov
-   coef <- object$coef
+   coef <- object$coefficients
    stdEr <- diag( result$coefCov )^0.5  # standard errors
    tStat <- coef / stdEr                # t-statistic
    if( useDfSys ) {             # p-values
@@ -153,7 +153,7 @@ summary.systemfit.equation <- function( object, useDfSys = NULL, ... ) {
 
    # coefficients, standard errors, ... 
    result$coefCov <- object$bcov
-   coef <- object$coef
+   coef <- object$coefficients
    stdEr <- diag( result$coefCov )^0.5  # standard errors
    tStat <- coef / stdEr                # t-statistic
    if( useDfSys ) {             # p-values

### 2006-09-29 trunk/R/systemfit.R

@@ -395,7 +395,7 @@ systemfit <- function(  eqns,
       nCoefEq = nCoefLiEq, yVec = yVecAll, xMat = xMatAll, xEq = xMatEq, methodRCov = control$methodRCov,
       centerResiduals = control$centerResiduals, solvetol = control$solvetol )
     #print( fimlResult )
-    coef <- fimlResult$coef
+    coef <- fimlResult$coefficients
     bcov <- fimlResult$coefCov
     resids <- fimlResult$resids
   }
@@ -444,7 +444,7 @@ systemfit <- function(  eqns,
     resulti$df.residual  <- df[i]           # degrees of freedom of residuals
     resulti$df.residual.sys  <- nObsAll- nCoefLiAll
        # degrees of freedom of residuals of the whole system
-    resulti$coef         <- coefEqI         # estimated coefficients
+    resulti$coefficients <- coefEqI         # estimated coefficients
     resulti$bcov         <- bcovi           # covariance matrix of estimated coefficients
     if( control$returnResponse ){
       resulti$response     <- yVecEq[[i]]     # vector of endogenous variables
@@ -540,7 +540,7 @@ systemfit <- function(  eqns,
      # number of linear independent coefficients in each equation
   results$df.residual <- nObsAll - nCoefLiAll
      # degrees of freedom of the whole system
-  results$coef    <- coef           # all estimated coefficients
+  results$coefficients <- coef           # all estimated coefficients
   results$bt      <- bt             # transformed vector of estimated coefficients
   results$bcov    <- bcov           # coefficients covariance matrix
   results$btcov   <- btcov          # covariance matrix for transformed coeff. vector

### 2006-09-29 paper/trunk/systemfit_compare_lm.tex
added appendix (table) to compare values returned by systemfit() and values returned by lm()
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@695 591db2bd-e0a0-4d22-8c19-2192dbb7da31

\section{Comparison of systemfit and lm}

\begin{table}[htbp]
\caption{Values returned by systemfit and lm}
\begin{tabular}{lll}
\hline
lm & systemfit & systemfit.equation \\
\hline
coefficients & coefficients & coefficients \\
residuals & & residuals \\
fitted.values & & fitted.values \\
rank & & \\
weights & & \\
df.residual & df.residual & df.residual \\
 & & df.residual.sys \\
call & call & \\
terms & & terms \\
contrasts & & \\
xlevels & & \\
offset & & \\
y** & & response* \\
x** & & modelMatrix* \\
 & & instMatrix* \\
model* & & modelFrame* \\
 & method & \\
 & eqnLabels & eqnLabel \\
 & nObs & nObs \\
 & nCoefAll & nCoefAll \\
 & nCoefEq & nCoef \\
 & nCoefLiAll & nCoefLiAll \\
 & nCoefLiEq & nCoefLi \\
 & iter & \\
 & bt & \\
 & bcov & covb \\
 & btcov & \\
 & rcov & \\
 & drcov & \\
 & rcovest & \\
 & olsr2 & r2 \\
 &  & adjr2 \\
 & mcelr2 & \\
 & R.restr & \\
 & q.restr & \\
 & TX & \\
 & control & \\
 & eq & \\
 & & inst \\
 & & ssr \\
 & & sigma \\
 & & \\
 & & \\
 & & \\
\hline
\end{tabular}
\end{table}

### 2006-10-01 trunk/R/ftest.R
systemfit() and summary.systemfit.equation() no longer return the number of observations ( x$nObs, x$eq[[ i ]]$nObs, s$eq[[ i ]]$nObs )
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@349 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@ -5,7 +5,7 @@ ftest.systemfit <- function( object, R.restr,
    vcov <- vcov( object )
    resid <- unlist( residuals( object ) )
    nEq   <- length( object$eq )
-   nObsEq <- object$nObs / nEq
+   nObsPerEq <- nrow( residuals( object ) )
    if( is.null( object$rcovest ) ) {
       rcov <- diag( nEq )
    } else {
@@ -22,7 +22,7 @@ ftest.systemfit <- function( object, R.restr,
       ( R.restr %*% coef - q.restr )
 
    denominator <- t( resid ) %*%
-      ( solve( rcov ) %x% diag( nObsEq ) ) %*%
+      ( solve( rcov ) %x% diag( nObsPerEq ) ) %*%
       resid
    #print( denominator )
 

### 2006-10-01 trunk/R/lrtest.R

@@ -3,7 +3,7 @@ lrtest.systemfit <- function( resultc, resultu ) {
   lrtest <- list()
   if( resultc$method %in% c( "SUR", "WSUR" ) &
       resultu$method %in% c( "SUR", "WSUR" ) ) {
-    nObs <- resultu$nObs / length( resultu$eq )
+    nObsPerEq <- nrow( residuals( resultu ) )
     lrtest$nRestr  <- resultu$rank - resultc$rank
     if( resultc$control$methodRCov != resultu$control$methodRCov ) {
       stop( paste( "both estimations must use the same formula to calculate",
@@ -11,7 +11,7 @@ lrtest.systemfit <- function( resultc, resultu ) {
     }
    residc <- as.matrix( residuals( resultc ) )
    residu <- as.matrix( residuals( resultu ) )
-   lrtest$statistic <- nObs * ( log( det( (t(residc) %*% residc)) ) -
+   lrtest$statistic <- nObsPerEq * ( log( det( (t(residc) %*% residc)) ) -
                          log( det( (t(residu) %*% residu))))
     lrtest$p.value <- 1 - pchisq( lrtest$statistic, lrtest$nRestr )
   }

### 2006-10-01 trunk/R/smallMethods.R

@@ -14,7 +14,7 @@ correlation.systemfit <- function( results, eqni, eqnj ) {
                       (1+sum(nCoefEq[1:eqnj])-nCoefEq[eqnj]):(sum(nCoefEq[1:eqnj]))]
   rij <- NULL
 
-  for(i in 1:results$eq[[1]]$nObs ) {
+  for( i in 1:nrow( residuals( results ) ) ) {
     xik    <- results$eq[[eqni]]$modelMatrix[i,]
     xjk    <- results$eq[[eqnj]]$modelMatrix[i,]
     top    <- xik %*% cij %*% xjk
@@ -30,7 +30,7 @@ correlation.systemfit <- function( results, eqni, eqnj ) {
 ## of the values, so you can examine the range, mean, etc
 se.ratio.systemfit <- function( resultsi, resultsj, eqni ) {
   ratio <- NULL
-  for(i in 1:resultsi$eq[[eqni]]$nObs ) {
+  for( i in 1:nrow( residuals( resultsi ) ) ) {
     xik    <- resultsi$eq[[eqni]]$modelMatrix[i,]
     top    <- sqrt( xik %*% resultsi$eq[[eqni]]$bcov %*% xik )
     bottom <- sqrt( xik %*% resultsj$eq[[eqni]]$bcov %*% xik )

### 2006-10-01 trunk/R/summary.systemfit.R

@@ -6,6 +6,13 @@ summary.systemfit <- function( object, useDfSys = NULL, ... ) {
          # TRUE if there are restrictions imposed
    }
 
+   # number of equations
+   nEq <- length( object$eq )
+   # number of observations per equation
+   nObsPerEq <- nrow( residuals( object ) )
+   # total number of observations
+   nObs <- nObsPerEq * nEq
+
    # preparing objects that will be returned
    result <- list()
    result$call <- object$call
@@ -44,15 +51,14 @@ summary.systemfit <- function( object, useDfSys = NULL, ... ) {
    result$coefficients <- cbind( coef, stdEr, tStat, pVal )
    colnames( result$coefficients ) <- c( "Estimate", "Std. Error",
       "t value", "Pr(>|t|)" )
-   result$df <- c( length( coef( object ) ), object$nObs - length( coef( object ) ) )
+   result$df <- c( length( coef( object ) ), nObs - length( coef( object ) ) )
 
    # R^2 values
-   nObsEq <- NULL
+   nObsEq <- rep( nObsPerEq, nEq )
    resid <- NULL
    response <- NULL
    responseMinusMean <- NULL
    for( i in 1:length( object$eq ) ) {
-      nObsEq <- c( nObsEq, object$eq[[ i ]]$nObs )
       resid <- c( resid, residuals( object$eq[[ i ]] ) )
       responseEqI <- fitted( object$eq[[ i ]] ) + residuals( object$eq[[ i ]] )
       response <- c( response, responseEqI )
@@ -80,7 +86,7 @@ summary.systemfit <- function( object, useDfSys = NULL, ... ) {
       for( i in 1:length( object$eq ) ) {
          for( j in 1:length( object$eq ) ) {
             denominator <- denominator + residCovInv[ i, j ] * yCov[ i, j ] *
-               object$eq[[ 1 ]]$nObs
+               nObsPerEq
          }
       }
       result$mcelroy.r.squared <- drop( 1 - ( rtOmega %*% resid ) / denominator )
@@ -116,7 +122,7 @@ print.summary.systemfit <- function( x, digits=6,... ) {
   }
   for(i in 1:length( x$eq ) ) {
     row <- NULL
-    row <- cbind( round( x$eq[[i]]$nObs,  digits ),
+    row <- cbind( round( sum( x$eq[[i]]$df ),  digits ),
                   round( x$eq[[i]]$df[2], digits ),
[…]

### 2006-10-01 trunk/R/systemfit.R

@@ -423,7 +423,6 @@ systemfit <- function(  eqns,
     resulti$i            <- i               # equation number
     resulti$eqnLabel     <- eqnLabels[[i]]
     resulti$terms        <- termsEq[[ i ]]
-    resulti$nObs         <- nObsEq[i]       # number of observations
     resulti$rank         <- nCoefLiEq[i]
       # rank = number of linear independent coefficients
     resulti$nCoef.sys    <- nCoefAll
@@ -474,7 +473,6 @@ systemfit <- function(  eqns,
 
   ## build the "return" structure for the whole system
   results$method  <- method
-  results$nObs    <- nObsAll        # total number of observations of all equations
   results$rank    <- nCoefLiAll
      # rank = total number of linear independent coefficients of all equations
   results$df.residual <- nObsAll - nCoefLiAll

### 2006-10-01 paper/trunk/systemfit_compare_lm.tex
systemfit_compare_lm.tex: updated table of variables returned by systemfit due to revisions 335-349 of the systemfit package
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@698 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 fitted.values & & fitted.values \\
-rank & & \\
+rank & rank & rank \\
+ & & rank.sys \\
+ & & nCoef.sys \\
 weights & & \\
@@
  & method & \\
- & eqnLabels & eqnLabel \\
- & nObs & nObs \\
- & nCoefAll & nCoefAll \\
- & nCoefEq & nCoef \\
- & nCoefLiAll & nCoefLiAll \\
- & nCoefLiEq & nCoefLi \\
+ & & eqnLabel \\
  & iter & \\
@@
  & rcov & \\
- & drcov & \\
  & rcovest & \\
- & olsr2 & r2 \\
- &  & adjr2 \\
- & mcelr2 & \\
  & R.restr & \\
@@
  & & sigma \\
- & & \\
- & & \\
- & & \\
 \hline

### 2006-10-02 paper/trunk/systemfit_abstract.tex
changes according to the first major comment of reviewer A: modified title and a few sentences (package to estimate -> package for estimating; equation system -> system of equations)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@691 591db2bd-e0a0-4d22-8c19-2192dbb7da31
added instrumental variables as suggested by reviewer A.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@701 591db2bd-e0a0-4d22-8c19-2192dbb7da31
added examples of stat. analyses that use IV methods
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@702 591db2bd-e0a0-4d22-8c19-2192dbb7da3

@@
 \Abstract{
-Many statistical analyses are based on models containing systems of
+Many statistical analyses (econometrics, biostatistics and
+experimental design) are based on models containing systems of
 structurally related equations.
@@
 
-\Keywords{R, simultaneous equations systems,
+\Keywords{R, system of simultaneous equations,
   seemingly unrelated regression, two-stage least squares,
-  three-stage least squares}
+  three-stage least squares, instrumental variables}
 

### 2006-10-02 paper/trunk/JSS/response_B.tex
corrected the paper according to all smaller comments of reviewer B except for the first point
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@684 591db2bd-e0a0-4d22-8c19-2192dbb7da31
corrected / harmonized punctuation marks around equations (first small comment of reviewer B)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@686 591db2bd-e0a0-4d22-8c19-2192dbb7da31
started responding to reviewer B comments. have to go to class for now.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@703 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 
+\textbf{[Still working on this section]}
+
+It was our intention to create a package for fitting models that could
+be verified against published results before moving on to include more
+sophisticated fitting methods.
+
+The reasons for doing this were two-fold.
+Our first reasons was to simply introduce a econometrics package where
+the main focus of the package was econometrics (a branch of the R
+packages that was under-represented).
+
+that at the time of the package introduction (2002) was
+
+The second reason for creating the package was to create a "forum" for
+more advanced fitting methods as described in Section 6 (Summary and
+Outlook). It is our intention to impliment these more advanced methods
+over time, but we felt that having a "basic" package that was accepted
+as part of a peer-review process would create an environment that
+would allow for acceptance in the future since the more sophisticated
+methods can be subject to difficulties beyond numerical routines.
+
 \section{Specific comments}
@@
 \subsection{Instrumental variable estimation}
-% Instrumental variable estimation has received much attention lately in biostatistics. Maybe it would
-% help 'sell' systemfit if this was emphasized at bit more. For instance, instrumental variables could
-% be mentioned directly in the introduction and included as a key word.
+% Instrumental variable estimation has received much attention lately
+% in biostatistics. Maybe it would help 'sell' systemfit if this was
+% emphasized at bit more. For instance, instrumental variables could
+% be mentioned directly in the introduction and included as a key
+% word.
+
+We have mentioned "instrumental variables" in the abstract, include
+three examples of topics that use IV methods and have added
+"instrumental variables" to the keywords.
 
 \subsection{Section 2}
-% Section 2 gives a brief overview of the statistical background. This may be a bit hard to follow for
-% people outside econometrics. It would help to include more references, especially in section 2.1.
+% Section 2 gives a brief overview of the statistical background. This
+% may be a bit hard to follow for people outside econometrics. It
+% would help to include more references, especially in section 2.1.
 
@@
 % Page 14: 'dependant' should be 'dependent'
+We have corrected the paper according to all of these comments.
 

### 2006-10-02 paper/trunk/JSS/response_B.tex
addressed some of reviewer b's comments
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@704 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 
-The reasons for doing this were two-fold.
-Our first reasons was to simply introduce a econometrics package where
-the main focus of the package was econometrics (a branch of the R
-packages that was under-represented).
-
-that at the time of the package introduction (2002) was
+The reasons for doing this were two-fold.  Our first reason was to
+simply introduce an econometrics package where the main focus of the
+package was "pure" econometrics and not to have a few specialized
+routines, as is the case with the tsls() function is the SEM
+package. While there are definitely links among structural equation
+models, IV methods and experimental design, the focus of this package
+was econometrics; more specifically simultaneous equation methods.
 
@@
 Outlook). It is our intention to impliment these more advanced methods
-over time, but we felt that having a "basic" package that was accepted
-as part of a peer-review process would create an environment that
-would allow for acceptance in the future since the more sophisticated
-methods can be subject to difficulties beyond numerical routines.
+(unbalanced data sets, FIML, GMM, etc.) over time, but we felt that
+having a "basic" package that was accepted as part of a peer-review
+process would create an environment that would reduce resistance to
+acceptance in the future since the more sophisticated methods can be
+subject to the difficulties beyond simple numerical routines as were
+used in the systemfit package so far.  Perhaps once the package gains
+enough acceptance, the functions can be made part of the core
+functions of R, but attempting to directly include the functions
+presented in the package might prove to be exceedingly difficult with
+the core development team.
+
+At the time the package was originally introduced (2002), there were
+very few econometrics functions for R and only recently have more
+functions and packages been submitted. Adding these functions to an
+existing econometrics package would have made sense, had one been
+available.
 
@@
 
+We have included a brief statement that the number of observations for
+each equation in the system is the same.
+
 \subsection{x and y variables}
 % Page 2: It should be clarified whether y-variables in one equation can be x-variables in another.
+
+This is stated in the introduction on Page one, lines 13-14.
 
@@
 
+The selection of instrument variables is discussed in more detail in
+other references and may quite unique to each field. 
+We chose not to focus on details such as selection of instrument
+variables, variance-covariance structures and other specifics of
+simultaneous equations as this article was written to present the
+basic equations, fitting methods and usage of the software package.
+While we desired to include a wide variety of topics, such as
+structure of the residual variance-covariance matrix, many of these
+were included in the package to compare, and finally match, the
+results from other software packages.
+For guidance on selection of instrument variables, the reader is
+referred to the citations in the article.
+
 \subsection{Equation 20 and 23}
@@
 % what it is.
+
+We have added a note to make the reader aware of the difference
+between the two equations.
 
@@
 
+We have made minor edits to the section to avoid any possible
+confusion. We discovered in the currently existing body of literature
+that there is no well defined presentation of computing the residual
+covariance matrix and have made a brief attempt here, but our attempt
+was by no means exhaustive.
+
 \subsection{First step OLS}
 % Line 196: It should be explained what is meant by a first step OLS.
+
+We have removed the term "first-step" as previously used in the
+manuscript. This term refers to the first-stage of a multi-stage
+estimation process (OLS->SUR) or (2SLS->3SLS).
 
@@
 % superior.
-%
+
+\textbf{[Arne, I'm not sure how to answer this one. I'm not sure this is an
+important issue for the manuscript. We mention that FIML is a more
+sophisticated approach in the summary and outlook section, but we
+don't really mention FIML in the paper. Should we?]}
 
@@
 % Line 215-219: Alternative approach should be explained more clearly.
+
+\textbf{[Arne, can we simply not mention the "alternative method" in the
+paper? I think it should stay as is, but I need a little more
+understanding of the reasoning behind the method and can't defend
+it. I just don't have any exposure to it yet.]}
 
@@
 
+We have included the proper citation for the R manuals which describes
+the syntax for formulas.
+
 \subsection{Section 2.7}
@@
 % stated? Would it not make more sense to compare SUR to 3SLS?
+
+The Hausman test is somewhat unique from other detection tests in that
+the test is designed "backwards" in the sense that the model that is
+being tested is correctly specified and not whether a predetermined 
+value is sufficiently close to the parameter estimate.
+
+\textbf{[I'm tired and will have to get back to this]}
 

### 2007-06-05 paper/trunk/systemfit_summary.tex
added citation of paelinck79 and anselin88
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@706 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 full information maximum likelihood (FIML), generalized methods of
-moments (GMM) and spatial econometric methods.
+moments (GMM) and spatial econometric methods
+\citep{paelinck79,anselin88}.
 

### 2007-06-05 paper/trunk/JSS/response_A.tex
corrected the paper according to all minor comments of reviewer A except for "p.3, l.45", the second part of "p.3, l.52" and "p.18, Tab.1".
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@687 591db2bd-e0a0-4d22-8c19-2192dbb7da31
changes according to the last minor comment of reviewer A: renamed arguments: rcovformula -> methodRCov + formula3sls -> method3sls (rev 205 of systemfit);  argument rcovformula can now be 'noDfCor' (as synonym for '0') (rev 206 of systemfit)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@688 591db2bd-e0a0-4d22-8c19-2192dbb7da31
changes acc

@@
 %
+We have changed the title according to these suggestions.
+
 \subsection{Abstract/Introduction}
@@
 %
+Both, abstract and introduction, now focus much more
+on the capabilities of the \pkg{systemfit} package
+rather than on the general theory of estimating systems of equations.
+
 \subsection{Estimation Methods (2.1)}
@@
 %     there are a lot of single letter elements - more standard naming as in "lm"
-%     objects etc. would be more desirable. Also, computing the partial t tests
+%     objects etc. would be more desirable.
+We have renamed single letter elements (not finished yet).
+
+%     Also, computing the partial t tests
 %     should/could be done in the summary() method, avoiding that certain parameters
@@
 %     as the first argument and to have then some `method' argument.
+We have reversed order of the arguments method and eqns/formula
+in systemfit() and systemfitClassic().
 % 
@@
 \section{Minor Points and Typos}
+We have corrected the paper according to all of these comments
+except for °p.3, l.45° and the second part of °p.3, l.52°.
+
 % MINOR POINTS & TYPOS

### 2007-06-05 paper/trunk/systemfit_abstract.tex
The abstract and the introduction now focus much more on the capabilities of systemfit rather than on the general theory of estimating systems of equations; added note about this in the response to reviewer A
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@709 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 \Abstract{
-Many statistical analyses (econometrics, biostatistics and
+Many statistical analyses (e.g.\ in econometrics, biostatistics and
 experimental design) are based on models containing systems of
 structurally related equations.
-In cases where cross-equation disturbances are correlated,
-full information methods are required \citep{zellner62}.
-If exogenous variables are stochastically dependent
-on the disturbances in the system,
-then instrumental variable estimation methods should be used
-\citep{zellner62b}
-The package \pkg{systemfit} provides the capability
+The \pkg{systemfit} package provides the capability
 to estimate systems of linear equations
 within the \proglang{R} programming environment.
+For instance, this package can be used for estimations
+by ``Ordinary Least Squares'' (OLS),
+``Seemingly Unrelated Regression'' (SUR),
+and the instrumental variable (IV) methods
+``Two-Stage Least Squares'' (2SLS) and
+``Three-Stage Least Squares'' (3SLS),
+where SUR and 3SLS estimations can optionally be iterated.
+Furthermore, the \pkg{systemfit} package provides tools
+for several statistical tests.
+It has been tested on a variety of datasets
+and the reliability of its results has been shown.
 }

### 2007-06-05 paper/trunk/systemfit_intro.tex
The abstract and the introduction now focus much more on the capabilities of systemfit rather than on the general theory of estimating systems of equations; added note about this in the response to reviewer A
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@709 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 leads to inefficient parameter estimates.
-However, estimating all equations simultaneously,
-taking the covariance structure of the residuals into account,
+However, estimating all equations simultaneously
+with a °Generalized Least Squares° (GLS) estimator,
+which takes the covariance structure of the residuals into account,
 leads to efficient estimates.
@@
 \citep{r-project}.
+Currently, the estimation methods
+°Ordinary Least Squares° (OLS),
+°Weighted Least Squares° (WLS),
+°Seemingly Unrelated Regression° (SUR),
+°Two-Stage Least Squares° (2SLS),
+°Weighted Two-Stage Least Squares° (W2SLS), and
+°Three-Stage Least Squares° (3SLS)
+are implemented.%
+\footnote{
+In this context,
+the term °weighted° in °Weighted Least Squares° (WLS) and
+°Weighted Two-Stage Least Squares° (W2SLS) means
+that the \emph{equations} might have different weights
+and \emph{not}
+that the \emph{observations} have different weights.
+}
+The WLS, SUR, W2SLS, and 3SLS estimates can be based
+either on one-step (OLS or 2SLS) (co)variances
+or these estimations can be iterated,
+where the (co)variances are calculated from the estimates of the previous step.
+Furthermore,
+the \pkg{systemfit} package
+provides statistical tests for parameter restrictions
+and for testing the consistency of the 3SLS estimation.
+
 Although systems of linear equations can be estimated

### 2007-06-06 paper/trunk/systemfit_other.tex
substituted "\COV" (defined in jss.cls) for all "\Cov"; removed definition(s) of "\Cov"
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@710 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
   m = \left( \widehat{\beta}_{2SLS} - \widehat{\beta}_{3SLS} \right)^{'}
-      \left( \Cov \left[ \widehat{\beta}_{2SLS} \right] -
-             \Cov \left[ \widehat{\beta}_{3SLS} \right] \right)
+      \left( \COV \left[ \widehat{\beta}_{2SLS} \right] -
+             \COV \left[ \widehat{\beta}_{3SLS} \right] \right)
       \left( \widehat{\beta}_{2SLS} - \widehat{\beta}_{3SLS} \right) ,
@@
 \end{equation}
-where $\widehat{\beta}_{2SLS}$ and $\Cov \left[ \widehat{\beta}_{2SLS} \right]$ are the estimated
+where $\widehat{\beta}_{2SLS}$ and $\COV \left[ \widehat{\beta}_{2SLS} \right]$ are the estimated
 coefficient and covariance matrix from 2SLS estimation, and
-$\widehat{\beta}_{3SLS}$ and $\Cov \left[ \widehat{\beta}_{3SLS} \right]$ are the estimated
+$\widehat{\beta}_{3SLS}$ and $\COV \left[ \widehat{\beta}_{3SLS} \right]$ are the estimated
 coefficients and covariance matrix from 3SLS estimation.

### 2007-06-06 paper/trunk/systemfit_estimation.tex
substituted "\COV" (defined in jss.cls) for all "\Cov"; removed definition(s) of "\Cov"
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@710 591db2bd-e0a0-4d22-8c19-2192dbb7da31
substituted "\E [ . ]" (\E is defined in jss.cls) for all "E ( . )" and "E [ . ]"
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@711 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 \begin{equation}
-   \Cov \left[ \widehat{\beta}_{OLS} \right] = \sigma^2 \left( X'X \right)^{-1} ,
+   \COV \left[ \widehat{\beta}_{OLS} \right] = \sigma^2 \left( X'X \right)^{-1} ,
    \label{eq:olsCovSameSigma}
 \end{equation}
-where $\sigma^2 = E \left( u_{it}^2 \right)$.
+where $\sigma^2 = \E \left[ u_{it}^2 \right]$.
 This assumes that the disturbances of all equations have the
@@
 \begin{equation}
-   \Cov \left[ \widehat{\beta}_{OLS} \right] = \left( X' \Omega^{-1} X \right)^{-1} ,
+   \COV \left[ \widehat{\beta}_{OLS} \right] = \left( X' \Omega^{-1} X \right)^{-1} ,
    \label{eq:olsCovSingleSigma}
@@
 $\sigma_{ij} = 0 \; \forall \; i \neq j$ and
-$\sigma_{ii} = \sigma_i^2 = E \left( u_{it}^2 \right)$.
+$\sigma_{ii} = \sigma_i^2 = \E \left[ u_{it}^2 \right]$.
 
@@
 $\sigma_{ij} = 0 \; \forall \; i \neq j$ and
-$\sigma_{ii} = \sigma_i^2 = E \left( u_{it}^2 \right)$.
+$\sigma_{ii} = \sigma_i^2 = \E \left[ u_{it}^2 \right]$.
 Like the OLS estimates these estimates are only efficient
@@
 \begin{equation}
-   \Cov \left[ \widehat{\beta}_{WLS} \right] = \left( X' \Omega^{-1} X \right)^{-1} .
+   \COV \left[ \widehat{\beta}_{WLS} \right] = \left( X' \Omega^{-1} X \right)^{-1} .
 \end{equation}
@@
 equation are uncorrelated
-$( E \left[ u_i ' X_i \right] = 0 \; \forall \; i )$,
+$( \E \left[ u_i ' X_i \right] = 0 \; \forall \; i )$,
 a consistent SUR estimation requires that all disturbance terms and all 
 regressors are uncorrelated
-$( E \left[ u_i ' X_j \right] = 0 \; \forall \; i, j )$.
+$( \E \left[ u_i ' X_j \right] = 0 \; \forall \; i, j )$.
 
@@
 where $\Omega = \Sigma \otimes I$ and
-$\sigma_{ij} = E \left( u_{it} u_{jt} \right)$.
+$\sigma_{ij} = \E \left[ u_{it} u_{jt} \right]$.
 And the covariance matrix of the estimated parameters is
 \begin{equation}
-   \Cov \left[ \widehat{\beta}_{SUR} \right] = \left( X' \Omega^{-1} X \right)^{-1} .
+   \COV \left[ \widehat{\beta}_{SUR} \right] = \left( X' \Omega^{-1} X \right)^{-1} .
    \label{eq:surCov}
@@
 If the regressors of one or more equations are correlated 
-with the disturbances ($E \left( u_i' X_i \right) \neq 0$),
+with the disturbances ($\E \left[ u_i' X_i \right] \neq 0$),
 OLS estimates are biased.
@@
 the disturbance terms of the corresponding equation 
-($E \left( u_i' H_i \right) = 0$).
+($\E \left[ u_i' H_i \right] = 0$).
 
@@
 \begin{equation}
-   \Cov \left[ \widehat{\beta}_{2SLS} \right] = \sigma^2 \left( \widehat{X}'
+   \COV \left[ \widehat{\beta}_{2SLS} \right] = \sigma^2 \left( \widehat{X}'
    \widehat{X} \right)^{-1} ,
@@
 \end{equation}
-where $\sigma^2 = E \left( u_{it}^2 \right)$.
+where $\sigma^2 = \E \left[ u_{it}^2 \right]$.
 If the disturbance terms of the individual equations
@@
 \begin{equation}
-   \Cov \left[ \widehat{\beta}_{2SLS} \right] = \left( \widehat{X}' \Omega^{-1}
+   \COV \left[ \widehat{\beta}_{2SLS} \right] = \left( \widehat{X}' \Omega^{-1}
    \widehat{X} \right)^{-1} ,
@@
 $\sigma_{ij} = 0 \; \forall \; i \neq j$ and
-$\sigma_{ii} = \sigma_i^2 = E \left( u_{it}^2 \right)$.
+$\sigma_{ii} = \sigma_i^2 = \E \left[ u_{it}^2 \right]$.
 The covariance matrix of the estimated parameters is
 \begin{equation}
-   \Cov \left[ \widehat{\beta}_{W2SLS} \right] = \left( \widehat{X}' \Omega^{-1}
+   \COV \left[ \widehat{\beta}_{W2SLS} \right] = \left( \widehat{X}' \Omega^{-1}
    \widehat{X} \right)^{-1} .
@@
 If the regressors are correlated with the disturbances 
-($E \left( u_i' X_i \right) \neq 0$) and
+($\E \left[ u_i' X_i \right] \neq 0$) and
 the disturbances are contemporaneously correlated, 
@@
 where $\Omega = \Sigma \otimes I$ and
-$\sigma_{ij} = E \left( u_{it} u_{jt} \right)$.
+$\sigma_{ij} = \E \left[ u_{it} u_{jt} \right]$.
 Its covariance matrix is
 \begin{equation}
-   \Cov \left[ \widehat{\beta}_{3SLS} \right] = \left( \widehat{X}' \Omega^{-1}
+   \COV \left[ \widehat{\beta}_{3SLS} \right] = \left( \widehat{X}' \Omega^{-1}
    \widehat{X} \right)^{-1} .
@@
 equation are uncorrelated
-$( E \left[ u_i' H_i \right]) = 0 \; \forall \; i )$,
+$( \E \left[ u_i' H_i \right]) = 0 \; \forall \; i )$,
 \cite{schmidt90} points out that this estimator is only consistent 
 if all disturbance terms and all instrumental variables are uncorrelated
-$( E \left[ u_i' H_j \right]) = 0 \; \forall \; i, j )$
+$( \E \left[ u_i' H_j \right]) = 0 \; \forall \; i, j )$
 with
@@
 \begin{equation}
-   \Cov \left[ \widehat{\beta}_{3SLS-IV} \right] = \left( \widehat{X}' \Omega^{-1}
+   \COV \left[ \widehat{\beta}_{3SLS-IV} \right] = \left( \widehat{X}' \Omega^{-1}
    X \right)^{-1} .
@@
 \begin{equation}
-   \Cov \left[ \widehat{\beta}_{3SLS-GMM} \right] =
+   \COV \left[ \widehat{\beta}_{3SLS-GMM} \right] =
    \left( X' H \left( H' \Omega H \right)^{-1} H' X \right)^{-1} .
@@
 \begin{equation}
-   \Cov \left[ \widehat{\beta}_{3SLS-Schmidt} \right] =
+   \COV \left[ \widehat{\beta}_{3SLS-Schmidt} \right] =
    \left( \widehat{X}' \Omega^{-1}  \widehat{X} \right)^{-1} 

### 2007-06-06 paper/trunk/systemfit_estimation.tex
a few minor improvements
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@712 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 \end{equation}
-where $\sigma^2 = \E \left[ u_{it}^2 \right]$.
+where $\sigma^2 = \E \left[ u_{it}^2 \right] \, \forall \, i$.
 This assumes that the disturbances of all equations have the
@@
 \end{equation}
-The econometrics software EViews uses
+The econometrics software \proglang{EViews} uses
 \begin{equation}
@@
 as defined by (\ref{eq:beta2sls}).
-EViews uses the standard 3SLS formula (\ref{eq:cov3sls}) to
+\proglang{EViews} uses the standard 3SLS formula (\ref{eq:cov3sls}) to
 calculate the covariance matrix of the 3SLS estimator.

### 2007-06-06 paper/trunk/JSS/response_A.tex
added note to our response to reviewer A that gives the reason why we keep the order in the section about 3SLS as it is
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@713 591db2bd-e0a0-4d22-8c19-2192dbb7da31
The paper no longer treats Omega as known in sections 2.1 and 2.2; the fact that Sigma and Omega are unknown and have to be estimated is no longer "hidden" in section 2.3, but is mentioned in sections 2.1 and 2.2 now; added a note in the response to reviewer A that this has been changed in the paper
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@714 591db2bd-e

@@
 %
+The paper no longer treats $\Omega$ as known in sections~2.1 and~2.2.
+The fact that $\Sigma$ and $\Omega$ are unknown and have to be estimated
+is no longer °hidden° in section 2.3,
+but is mentioned in sections~2.1 and~2.2 now.
+
 %     For the varying 3SLS estimators, it would help to point out why/when
@@
 %
+We think that it is better
+to present the review of the varying 3SLS estimators
+at the end rather than at the beginning of the 3SLS section,
+because it is easier for the reader to understand this review
+if the compared estimation methods have been described before.
+
 \subsection{Imposing Linear Restrictions (2.2)}

### 2007-06-06 paper/trunk/systemfit_estimation.tex
The paper no longer treats Omega as known in sections 2.1 and 2.2; the fact that Sigma and Omega are unknown and have to be estimated is no longer "hidden" in section 2.3, but is mentioned in sections 2.1 and 2.2 now; added a note in the response to reviewer A that this has been changed in the paper
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@714 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 If the whole system is treated as one single equation,
-the covariance matrix of the estimated parameters is
+an estimator of the covariance matrix of the estimated parameters is
 \begin{equation}
-   \COV \left[ \widehat{\beta}_{OLS} \right] = \sigma^2 \left( X'X \right)^{-1} ,
+   \COVHat \left[ \widehat{\beta}_{OLS} \right] = \sHat^2 \left( X'X \right)^{-1} ,
    \label{eq:olsCovSameSigma}
 \end{equation}
-where $\sigma^2 = \E \left[ u_{it}^2 \right] \, \forall \, i$.
+where $\sHat^2$ is the estimated variance of the residuals.
 This assumes that the disturbances of all equations have the
-same variance.
+same variance $(\E [ u_{it}^2 ] = \sigma^2 \, \forall \, i)$.
 
@@
 are allowed to have different variances,
-the covariance matrix of the estimated parameters is
+an estimator of the covariance matrix of the estimated parameters is
 \begin{equation}
-   \COV \left[ \widehat{\beta}_{OLS} \right] = \left( X' \Omega^{-1} X \right)^{-1} ,
+   \COVHat \left[ \widehat{\beta}_{OLS} \right] = \left( X' \OHat^{-1} X \right)^{-1} ,
    \label{eq:olsCovSingleSigma}
 \end{equation}
-where $\Omega = \Sigma \otimes I$,
-$\sigma_{ij} = 0 \; \forall \; i \neq j$ and
-$\sigma_{ii} = \sigma_i^2 = \E \left[ u_{it}^2 \right]$.
+where $\OHat = \SHat \otimes I$,
+$\sHat_{ij} = 0 \; \forall \; i \neq j$ and
+$\sHat_{ii} = \sHat_i^2$ is the estimated variance
+of the disturbence term in the $i$th equation.
 
@@
 
-The Weighted Least Squares (WLS) estimator of the system 
-is obtained by
+The Weighted Least Squares (WLS) estimator allows for different variances
+of the disturbance terms in the different equations
+$( \sigma_i^2 \neq \sigma_j^2 \, \forall \, i \neq j$,
+where $\sigma_i^2 = \E [ u_{it}^2 ])$.
+The WLS estimator can be obtained by
 \begin{equation}
-   \widehat{\beta}_{WLS} = \left( X' \Omega^{-1} X \right)^{-1} X' \Omega^{-1} y ,
+   \widehat{\beta}_{WLS} = \left( X' \OHat^{-1} X \right)^{-1} X' \OHat^{-1} y ,
 \end{equation}
-where $\Omega = \Sigma \otimes I$,
-$\sigma_{ij} = 0 \; \forall \; i \neq j$ and
-$\sigma_{ii} = \sigma_i^2 = \E \left[ u_{it}^2 \right]$.
-Like the OLS estimates these estimates are only efficient
+where $\OHat = \SHat \otimes I$,
+$\sHat_{ij} = 0 \; \forall \; i \neq j$ and
+$\sHat_{ii} = \sHat_i^2$ is the estimated variance
+of the disturbance term in the $i$th equation.%
+\footnote{%
+Theoretically, one should use the variances of the true residuals.
+However, they are not known in most empirical applications.
+Therefore, true variances are generally replaced by estimated variances
+that are calculated from the residuals of a first-step OLS estimation.
+}
+Like the OLS estimates, these estimates are only efficient
 if the disturbance terms are not contemporaneously correlated.
-The covariance matrix of the estimated parameters is
+An estimator of the covariance matrix of the estimated parameters is
 \begin{equation}
-   \COV \left[ \widehat{\beta}_{WLS} \right] = \left( X' \Omega^{-1} X \right)^{-1} .
+   \COVHat \left[ \widehat{\beta}_{WLS} \right] = \left( X' \OHat^{-1} X \right)^{-1} .
 \end{equation}
@@
 (SUR) estimator \citep{zellner62}.
+Since the true covariance matrix of the disturbance terms
+is generally unknown,
+a Feasible Generalized Least Squares (FGLS) estimation,
+which is based on an estimated covariance matrix of the disturbance terms,
+is the textbook solution for this problem.
 It should be noted that while an unbiased OLS or WLS estimation requires only that
@@
 \begin{equation}
-   \widehat{\beta}_{SUR} = \left( X' \Omega^{-1} X \right)^{-1} X' \Omega^{-1} y ,
+   \widehat{\beta}_{SUR} = \left( X' \OHat^{-1} X \right)^{-1} X' \OHat^{-1} y ,
    \label{eq:sur}
 \end{equation}
-where $\Omega = \Sigma \otimes I$ and
-$\sigma_{ij} = \E \left[ u_{it} u_{jt} \right]$.
-And the covariance matrix of the estimated parameters is
+where $\OHat = \SHat \otimes I$ and
+$\SHat$ is the estimated covariance matrix of the disturbance terms.
+An estimator of the covariance matrix of the estimated parameters is
 \begin{equation}
-   \COV \left[ \widehat{\beta}_{SUR} \right] = \left( X' \Omega^{-1} X \right)^{-1} .
+   \COVHat \left[ \widehat{\beta}_{SUR} \right] = \left( X' \OHat^{-1} X \right)^{-1} .
    \label{eq:surCov}
@@
 \end{equation}
-At the second stage the unbiased two-stage least squares estimates
+At the second stage, the unbiased two-stage least squares estimates
 of $\beta$ are obtained by
@@
 If the whole system is treated as one single equation,
-the covariance matrix of the estimated parameters is
+an estimator of the covariance matrix of the estimated parameters is
 \begin{equation}
-   \COV \left[ \widehat{\beta}_{2SLS} \right] = \sigma^2 \left( \widehat{X}'
+   \COVHat \left[ \widehat{\beta}_{2SLS} \right] = \sHat^2 \left( \widehat{X}'
    \widehat{X} \right)^{-1} ,
@@
 \end{equation}
-where $\sigma^2 = \E \left[ u_{it}^2 \right]$.
+where $\sHat^2$ is the estimated variance of the disturbance terms.
 If the disturbance terms of the individual equations
 are allowed to have different variances, 
-the covariance matrix of the estimated parameters is
+an estimator of the covariance matrix of the estimated parameters is
 \begin{equation}
-   \COV \left[ \widehat{\beta}_{2SLS} \right] = \left( \widehat{X}' \Omega^{-1}
+   \COVHat \left[ \widehat{\beta}_{2SLS} \right] = \left( \widehat{X}' \OHat^{-1}
    \widehat{X} \right)^{-1} ,
@@
 \end{equation}
-where $\Omega = \Sigma \otimes I$,
-$\sigma_{ij} = 0 \; \forall \; i \neq j$ and
-$\sigma_{ii} = \sigma_i^2 = \E \left[ u_{it}^2 \right]$.
+where $\OHat = \SHat \otimes I$,
+$\sHat_{ij} = 0 \; \forall \; i \neq j$ and
+$\sHat_{ii} = \sHat_i^2$ is the estimated variance
+of the disturbance term in the $i$th equation.
 
@@
 \begin{equation}
-   \widehat{\beta}_{W2SLS} = \left( \widehat{X}' \Omega^{-1} \widehat{X} 
-   \right)^{-1} \widehat{X}' \Omega^{-1} y ,
+   \widehat{\beta}_{W2SLS} = \left( \widehat{X}' \OHat^{-1} \widehat{X}
+   \right)^{-1} \widehat{X}' \OHat^{-1} y ,
 \end{equation}
-where $\Omega = \Sigma \otimes I$,
-$\sigma_{ij} = 0 \; \forall \; i \neq j$ and
-$\sigma_{ii} = \sigma_i^2 = \E \left[ u_{it}^2 \right]$.
-The covariance matrix of the estimated parameters is
+where $\OHat = \SHat \otimes I$,
+$\sHat_{ij} = 0 \; \forall \; i \neq j$ and
+$\sHat_{ii} = \sHat_i^2$ is the estimated variance
+of the disturbance term in the $i$th equation.
+An estimator of the covariance matrix of the estimated parameters is
 \begin{equation}
-   \COV \left[ \widehat{\beta}_{W2SLS} \right] = \left( \widehat{X}' \Omega^{-1}
+   \COVHat \left[ \widehat{\beta}_{W2SLS} \right] = \left( \widehat{X}' \OHat^{-1}
    \widehat{X} \right)^{-1} .
@@
 the disturbances are contemporaneously correlated, 
-a Generalized Least Squares (GLS) version of the two-stage least squares
-estimation leads to consistent and efficient estimates.
+a Feasible Generalized Least Squares (FGLS) version of the two-stage least squares
+estimation leads to consistent and asymptotically efficient estimates.
 This estimation procedure is generally called °Three-stage Least
@@
 \begin{equation}
-   \widehat{\beta}_{3SLS} = \left( \widehat{X}' \Omega^{-1} \widehat{X} 
-   \right)^{-1} \widehat{X}' \Omega^{-1} y ,
+   \widehat{\beta}_{3SLS} = \left( \widehat{X}' \OHat^{-1} \widehat{X}
+   \right)^{-1} \widehat{X}' \OHat^{-1} y ,
    \label{eq:3slsGls}
 \end{equation}
-where $\Omega = \Sigma \otimes I$ and
-$\sigma_{ij} = \E \left[ u_{it} u_{jt} \right]$.
-Its covariance matrix is
+where $\OHat = \SHat \otimes I$ and
+$\SHat$ is the estimated covariance matrix of the disturbance terms.
+An estimator of the covariance matrix of the estimated parameters is
 \begin{equation}
-   \COV \left[ \widehat{\beta}_{3SLS} \right] = \left( \widehat{X}' \Omega^{-1}
+   \COVHat \left[ \widehat{\beta}_{3SLS} \right] = \left( \widehat{X}' \OHat^{-1}
    \widehat{X} \right)^{-1} .
@@
 \begin{equation}
-   \widehat{\beta}_{3SLS-IV} = \left( \widehat{X}' \Omega^{-1} X
-   \right)^{-1} \widehat{X}' \Omega^{-1} y .
+   \widehat{\beta}_{3SLS-IV} = \left( \widehat{X}' \OHat^{-1} X
+   \right)^{-1} \widehat{X}' \OHat^{-1} y .
    \label{eq:3slsIv}
 \end{equation}
-The covariance matrix of this 3SLS-IV estimator is
+An estimator of the covariance matrix of the estimated 3SLS-IV parameters is
 \begin{equation}
-   \COV \left[ \widehat{\beta}_{3SLS-IV} \right] = \left( \widehat{X}' \Omega^{-1}
+   \COVHat \left[ \widehat{\beta}_{3SLS-IV} \right] = \left( \widehat{X}' \OHat^{-1}
    X \right)^{-1} .
@@
 \begin{equation}
-   \widehat{\beta}_{3SLS-GMM} = \left( X' H \left( H' \Omega H \right)^{-1}
-   H' X \right)^{-1} X' H \left( H' \Omega H \right)^{-1} H' y .
+   \widehat{\beta}_{3SLS-GMM} = \left( X' H \left( H' \OHat H \right)^{-1}
+   H' X \right)^{-1} X' H \left( H' \OHat H \right)^{-1} H' y .
    \label{eq:3slsGmm}
 \end{equation}
-The covariance matrix of the 3SLS-GMM estimator is
+An estimator of the covariance matrix of the estimated 3SLS-GMM parameters is
 \begin{equation}
-   \COV \left[ \widehat{\beta}_{3SLS-GMM} \right] =
-   \left( X' H \left( H' \Omega H \right)^{-1} H' X \right)^{-1} .
+   \COVHat \left[ \widehat{\beta}_{3SLS-GMM} \right] =
+   \left( X' H \left( H' \OHat H \right)^{-1} H' X \right)^{-1} .
 \end{equation}
@@
 \begin{equation}
-   \widehat{\beta}_{3SLS-Schmidt} = \left( \widehat{X}' \Omega^{-1} \widehat{X}
-   \right)^{-1} \widehat{X}' \Omega^{-1} 
+   \widehat{\beta}_{3SLS-Schmidt} = \left( \widehat{X}' \OHat^{-1} \widehat{X}
+   \right)^{-1} \widehat{X}' \OHat^{-1} 
    H \left( H' H \right)^{-1} H' y .
@@
 \end{equation}
-The covariance matrix of this estimator is
+An estimator of the covariance matrix of these estimated parameters is
 \begin{equation}
-   \COV \left[ \widehat{\beta}_{3SLS-Schmidt} \right] =
-   \left( \widehat{X}' \Omega^{-1}  \widehat{X} \right)^{-1} 
-   \widehat{X}' \Omega^{-1} H \left( H' H \right)^{-1} H' \Omega 
-   H \left( H' H \right)^{-1} H' \Omega^{-1} \widehat{X}
-   \left( \widehat{X}' \Omega^{-1}  \widehat{X} \right)^{-1} .
+   \COVHat \left[ \widehat{\beta}_{3SLS-Schmidt} \right] =
+   \left( \widehat{X}' \OHat^{-1}  \widehat{X} \right)^{-1}
+   \widehat{X}' \OHat^{-1} H \left( H' H \right)^{-1} H' \OHat
+   H \left( H' H \right)^{-1} H' \OHat^{-1} \widehat{X}
+   \left( \widehat{X}' \OHat^{-1}  \widehat{X} \right)^{-1} .
 \end{equation}
@@
    \widehat{\beta}_{3SLS-EViews} = \widehat{\beta}_{2SLS} + 
-   \left( \widehat{X}' \Omega^{-1} \widehat{X} \right)^{-1} 
-   \widehat{X}' \Omega^{-1} \left( y - X \widehat{\beta}_{2SLS} \right) ,
+   \left( \widehat{X}' \OHat^{-1} \widehat{X} \right)^{-1}
+   \widehat{X}' \OHat^{-1} \left( y - X \widehat{\beta}_{2SLS} \right) ,
    \label{eq:3slsEViews}
@@
 \proglang{EViews} uses the standard 3SLS formula (\ref{eq:cov3sls}) to
-calculate the covariance matrix of the 3SLS estimator.
+calculate an estimator of the covariance matrix of the estimated parameters.
 

### 2007-06-06 paper/trunk/systemfit_restrictions.tex
substituted "\COV" (defined in jss.cls) for all "\Cov"; removed definition(s) of "\Cov"
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@710 591db2bd-e0a0-4d22-8c19-2192dbb7da31
substituted "\E [ . ]" (\E is defined in jss.cls) for all "E ( . )" and "E [ . ]"
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@711 591db2bd-e0a0-4d22-8c19-2192dbb7da31
a few minor improvements
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@712 591db2bd-e0a0-4d22-8c19-2192dbb7da31
The paper no longer treats Omega as known in sections 2.1 and 2.2; the fact that Sigma and Omega

@@
 The original parameters can be obtained by equation (\ref{eq:T-restr})
-and the covariance matrix of the original parameters
+and the estimated covariance matrix of the original parameters
 can be obtained by
 \begin{equation}
-   \Cov \left[ \widehat{\beta} \right]
-   = M \cdot \Cov \left[ \widehat{\beta}^* \right] \cdot M' .
+   \COVHat \left[ \widehat{\beta} \right]
+   = M \cdot \COVHat \left[ \widehat{\beta}^* \right] \cdot M' .
 \end{equation}
@@
 If the whole system is treated as one single equation,
-the covariance matrix of the estimated parameters is
+an estimator of the covariance matrix of the estimated parameters is
 \begin{equation}
-   \Cov 
+   \COVHat
    \left[ \begin{array}{c}
@@
    \end{array} \right] 
-   = \sigma^2 
+   = \sHat^2 
    \left[ \begin{array}{cc}
@@
 \end{equation}
-where $\sigma^2 = E \left( u_{it}^2 \right)$.
+where $\sHat$ is the estimated variance of the disturbance terms.
 If the disturbance terms of the individual equations
 are allowed to have different variances, 
-the covariance matrix of the estimated parameters is
+an estimator of the covariance matrix of the estimated parameters is
 \begin{equation}
-   \Cov 
+   \COVHat
    \left[ \begin{array}{c}
@@
    \left[ \begin{array}{cc}
-      X' \Omega^{-1} X & R' \\ 
+      X' \OHat^{-1} X & R' \\
       R & 0
@@
 \end{equation}
-where $\Omega = \Sigma \otimes I$,
-$\sigma_{ij} = 0 \; \forall \; i \neq j$ and
-$\sigma_{ii} = \sigma_i^2 = E \left( u_{it}^2 \right)$.
+where $\OHat = \SHat \otimes I$,
+$\sHat_{ij} = 0 \; \forall \; i \neq j$ and
+$\sHat_{ii} = \sHat_i^2$ is the estimated variance
+of the disturbance term in the $i$th equation.
 
@@
    \left[ \begin{array}{c}
-      X' \Omega^{-1} y \\ q 
+      X' \OHat^{-1} y \\ q
    \end{array} \right] ,
 \end{equation}
-where $\Omega = \Sigma \otimes I$,
-$\sigma_{ij} = 0 \; \forall \; i \neq j$ and
-$\sigma_{ii} = \sigma_i^2 = E \left( u_{it}^2 \right)$.
-The covariance matrix of the estimated parameters is
+where $\OHat = \SHat \otimes I$,
+$\sHat_{ij} = 0 \; \forall \; i \neq j$ and
+$\sHat_{ii} = \sHat_i^2$ is the estimated variance
+of the disturbance term in the $i$th equation.
+An estimator of the covariance matrix of the estimated parameters is
 \begin{equation}
-   \Cov 
+   \COVHat
    \left[ \begin{array}{c}
@@
    \left[ \begin{array}{c}
-      X' \Omega^{-1} y \\ q 
+      X' \OHat^{-1} y \\ q
    \end{array} \right] ,
 \end{equation}
-where $\Omega = \Sigma \otimes I$ and
-$\sigma_{ij} = E \left( u_{it} u_{jt} \right)$.
-The covariance matrix of the estimated parameters is
+where $\OHat = \SHat \otimes I$ and
+$\SHat$ is the estimated covariance matrix of the disturbance terms.
+An estimator of the covariance matrix of the estimated parameters is
 \begin{equation}
-   \Cov 
+   \COVHat
    \left[ \begin{array}{c}
@@
 \end{equation}
-where $\sigma^2 = E \left( u_{it}^2 \right)$.
+where $\sHat^2$ is the estimated variance of the disturbance terms.
 If the disturbance terms of the individual equations
 are allowed to have different variances, 
-the covariance matrix of the estimated parameters is
+an estimator of the covariance matrix of the estimated parameters is
 \begin{equation}
-   \Cov 
+   \COVHat
    \left[ \begin{array}{c}
@@
    \left[ \begin{array}{cc}
-      \widehat{X}' \Omega^{-1} \widehat{X} & R' \\ 
+      \widehat{X}' \OHat^{-1} \widehat{X} & R' \\
       R & 0
@@
    \left[ \begin{array}{c}
-      \widehat{X}' \Omega^{-1} y \\ q 
+      \widehat{X}' \OHat^{-1} y \\ q
    \end{array} \right] ,
 \end{equation}
-where $\Omega = \Sigma \otimes I$,
-$\sigma_{ij} = 0 \; \forall \; i \neq j$ and
-$\sigma_{ii} = \sigma_i^2 = E \left( u_{it}^2 \right)$.
-The covariance matrix of the estimated parameters is
+where $\OHat = \SHat \otimes I$,
+$\sHat_{ij} = 0 \; \forall \; i \neq j$ and
+$\sHat_{ii} = \sHat_i^2$ is the estimated variance
+of the disturbance term in the $i$th equation.
+An estimator of the covariance matrix of the estimated parameters is
 \begin{equation}
-   \Cov 
+   \COVHat
    \left[ \begin{array}{c}
@@
    \left[ \begin{array}{c}
-      \widehat{X}' \Omega^{-1} y \\ q 
+      \widehat{X}' \OHat^{-1} y \\ q
    \end{array} \right] ,
@@
 \end{equation}
-where $\Omega = \Sigma \otimes I$ and
-$\sigma_{ij} = E \left( u_{it} u_{jt} \right)$.
-The covariance matrix of this estimator is
+where $\OHat = \SHat \otimes I$ and
+$\SHat$ is the estimated covariance matrix of the disturbance terms.
+An estimator of the covariance matrix of the estimated parameters is
 \begin{equation}
-   \Cov 
+   \COVHat
    \left[ \begin{array}{c}
@@
    \left[ \begin{array}{cc}
-      \widehat{X}' \Omega^{-1} X & R' \\ 
+      \widehat{X}' \OHat^{-1} X & R' \\
       R & 0
@@
 \begin{equation}
-   \Cov 
+   \COVHat
    \left[ \begin{array}{c}
@@
    \left[ \begin{array}{cc}
-      X' H \left( H' \Omega H \right)^{-1} H' X & R' \\ 
+      X' H \left( H' \OHat H \right)^{-1} H' X & R' \\
       R & 0
@@
    \left[ \begin{array}{c}
-      X' H \left( H \Omega H \right)^{-1} H' y \\ q 
+      X' H \left( H \OHat H \right)^{-1} H' y \\ q
    \end{array} \right] ,
@@
    \left[ \begin{array}{c}
-      \widehat{X}' \Omega^{-1} H \left( H' H \right)^{-1} H' y \\ q 
+      \widehat{X}' \OHat^{-1} H \left( H' H \right)^{-1} H' y \\ q
    \end{array} \right] ,
@@
 \begin{eqnarray}
-   \Cov 
+   \COVHat
    \left[ \begin{array}{c}
@@
    \left[ \begin{array}{cc}
-      \widehat{X}' \Omega^{-1} H \left( H' H \right)^{-1} H' \Omega
-      H \left( H' H \right)^{-1} H' \Omega^{-1} \widehat{X} & 0' \\ 
+      \widehat{X}' \OHat^{-1} H \left( H' H \right)^{-1} H' \OHat
+      H \left( H' H \right)^{-1} H' \OHat^{-1} \widehat{X} & 0' \\
       0 & 0
@@
 \end{eqnarray}
-The econometrics software EViews calculates the restricted 3SLS estimator by
+The econometrics software \proglang{EViews} calculates the restricted 3SLS estimator by
 \begin{equation}
@@
    \left[ \begin{array}{c}
-      \widehat{X}' \Omega^{-1} \left( y - X \widehat{\beta}^0_{2SLS} \right)
+      \widehat{X}' \OHat^{-1} \left( y - X \widehat{\beta}^0_{2SLS} \right)
       \\ q 
@@
 by equation (\ref{eq:beta2SLSr}). 
-To calculate the covariance matrix
-EViews uses the standard formula of the restricted 3SLS
-estimator~(\ref{eq:cov3slsr}).
+\proglang{EViews} uses the standard formula of the restricted 3SLS
+estimator~(\ref{eq:cov3slsr}) to calculate an estimator
+for the covariance matrix of the estimated parameters.
 

### 2007-06-07 paper/trunk/JSS/response_A.tex
response to reviewer A: explained why we describe each estimation method in a single (subsub)section (rather than shortening this section by exploiting the common properties of the different estimators)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@715 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 % 
+The reviewer is right that it is possible to shorten
+the description of the estimation methods
+by exploiting the common properties of the different estimators.
+However, in our opinion this shorting would considerably worsen
+the readability of this section.
+Furthermore, a major purpose of this section is
+that a user of the \pkg{systemfit} package
+can take a look at the theory (e.g.\ formulas) of a specific estimation method.
+Therefore, we fully describe each estimation method
+in a single (subsub)section.
+
 %     Terminology for the different estimators also varies between various

### 2007-06-07 paper/trunk/JSS/response_A.tex
made clearer that section 2.2 is about estimating coefficients under linear restrictions (and not about inference); explained more concisely, how the two methods to impose linear restrictions are implemented in systemfit, i.e. that the rest of this (sub)section belongs to the second method; renamed restricted coefficients
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@716 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 %
+First, it has been made clearer
+that this section is about estimating coefficients under linear restrictions
+(and not about inference).
+Second, it has been explained more concisely,
+how the two methods to impose linear restrictions are implemented
+in \pkg{systemfit},
+i.e\ that the rest of this (sub)section belongs to the second method
+(equation 31, former equation 33).
+
 \subsection{systemfitClassic (4.3)}

### 2007-06-07 paper/trunk/systemfit_restrictions.tex
made clearer that section 2.2 is about estimating coefficients under linear restrictions (and not about inference); explained more concisely, how the two methods to impose linear restrictions are implemented in systemfit, i.e. that the rest of this (sub)section belongs to the second method; renamed restricted coefficients
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@716 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 
-It is common to perform hypothesis tests by imposing restrictions on
-the parameter estimates.
+In many empirical applications,
+it is desirable to impose restrictions on the parameters
+that are derived from the underlying theoretical model.%
+\footnote{%
+For instance in econometric demand and production analysis,
+it is common to impose (and test) homogeneity and symmetry properties.
+}
 There are two ways to impose linear parameter restrictions.
@@
 \begin{equation}
-   \beta = M \cdot \beta^* \label{eq:T-restr} ,
+   \beta = M \cdot \beta^M \label{eq:T-restr} ,
 \end{equation}
-where $\beta^*$ is a vector of restricted (linear independent) coefficients,
+where $\beta^M$ is a vector of restricted (linear independent) coefficients,
 and $M$ is a matrix with the number of rows equal to the number of
@@
 the number of columns equal to the number of restricted coefficients
-($\beta^*$).
+($\beta^M$).
 $M$ can be used to map each unrestricted coefficient to one or more
 restricted coefficients.
-
-To impose these restrictions, the $X$ matrix is
-(post-)\hspace{0pt}multiplied by this $M$ matrix, so that
-\begin{equation}
-    X^* = X \cdot M .
-\end{equation}
-Then, $X^*$ is substituted for $X$ and a standard estimation as described
-in the previous section is done
-(equations~\ref{eq:ols}--\ref{eq:3slsEViews}).
-This results in the linear independent parameter estimates $\widehat{\beta}^*$ and
-their covariance matrix.
-The original parameters can be obtained by equation (\ref{eq:T-restr})
-and the estimated covariance matrix of the original parameters
-can be obtained by
-\begin{equation}
-   \COVHat \left[ \widehat{\beta} \right]
-   = M \cdot \COVHat \left[ \widehat{\beta}^* \right] \cdot M' .
-\end{equation}
 
@@
 \begin{equation}
-   R \beta^0 = q ,
+   R \beta^R = q ,
+   \label{eq:restr-R}
 \end{equation}
-where $\beta^0$ is the vector of the restricted coefficients, 
+where $\beta^R$ is the vector of the restricted coefficients,
 and $R$ and $q$ are a matrix and vector, respectively,
@@
 These two methods can be combined.
-In this case the restrictions imposed using the latter method are
-imposed on the linear independent parameters due to the restrictions
-imposed using the first method, so that
+In this case, the restrictions imposed using the latter method are
+imposed on the linear independent parameters
+that are restricted by the first method, so that
 \begin{equation}
-   R \beta^{*0} = q ,
+   R \beta^{MR} = q ,
 \end{equation}
-where $\beta^{*0}$ is the vector of the restricted $\beta^*$ coefficients.
+where $\beta^{MR}$ is the vector of the restricted $\beta^M$ coefficients.
+
+\subsubsection[Implementation in systemfit]{Implementation in \pkg{systemfit}}
+
+If the first method (equation~\ref{eq:T-restr}) is chosen
+to impose these restrictions,
+the matrix of regressors $X$ is (post-)\hspace{0pt}multiplied
+by the $M$ matrix,
+so that
+\begin{equation}
+    X^M = X \cdot M .
+\end{equation}
+Then, $X^M$ is substituted for $X$ and a standard estimation as described
+in the previous section is done
+(equations~\ref{eq:ols}--\ref{eq:3slsEViews}).
+This results in the linear independent parameter estimates $\widehat{\beta}^M$ and
+their covariance matrix.
+The original parameters can be obtained by equation (\ref{eq:T-restr})
+and the estimated covariance matrix of the original parameters
+can be obtained by
+\begin{equation}
+   \COVHat \left[ \widehat{\beta} \right]
+   = M \cdot \COVHat \left[ \widehat{\beta}^M \right] \cdot M' .
+\end{equation}
+
+The implementation of the second method to impose linear restrictions
+(equation~\ref{eq:restr-R})
+is described for each estimation method in the following sections.
 
@@
 
-The OLS estimator restricted by $R \beta^0 = q$ can be obtained by
+The OLS estimator restricted by $R \beta^R = q$ can be obtained by
 \begin{equation}
    \left[ \begin{array}{c}
-      \widehat{\beta}^0_{OLS} \\ \widehat{\lambda}
+      \widehat{\beta}^R_{OLS} \\ \widehat{\lambda}
    \end{array} \right]
@@
    \left[ \begin{array}{c}
-      \widehat{\beta}^0_{OLS} \\ \widehat{\lambda}
+      \widehat{\beta}^R_{OLS} \\ \widehat{\lambda}
    \end{array} \right] 
@@
 
-The WLS estimator restricted by $R \beta^0 = q$ can be obtained by
+The WLS estimator restricted by $R \beta^R = q$ can be obtained by
 \begin{equation}
    \left[ \begin{array}{c}
-      \widehat{\beta}^0_{WLS} \\ \widehat{\lambda}
+      \widehat{\beta}^R_{WLS} \\ \widehat{\lambda}
    \end{array} \right]
@@
    \left[ \begin{array}{c}
-      \widehat{\beta}^0_{WLS} \\ \widehat{\lambda}
+      \widehat{\beta}^R_{WLS} \\ \widehat{\lambda}
    \end{array} \right] 
@@
 
-The SUR estimator restricted by $R \beta^0 = q$ can be obtained by
+The SUR estimator restricted by $R \beta^R = q$ can be obtained by
 \begin{equation}
    \left[ \begin{array}{c}
-      \widehat{\beta}^0_{SUR} \\ \widehat{\lambda}
+      \widehat{\beta}^R_{SUR} \\ \widehat{\lambda}
    \end{array} \right]
@@
    \left[ \begin{array}{c}
-      \widehat{\beta}^0_{SUR} \\ \widehat{\lambda}
+      \widehat{\beta}^R_{SUR} \\ \widehat{\lambda}
    \end{array} \right] 
@@
 
-The 2SLS estimator restricted by $R \beta^0 = q$ can be obtained by
+The 2SLS estimator restricted by $R \beta^R = q$ can be obtained by
 \begin{equation}
    \left[ \begin{array}{c}
-      \widehat{\beta}^0_{2SLS} \\ \widehat{\lambda}
+      \widehat{\beta}^R_{2SLS} \\ \widehat{\lambda}
    \end{array} \right]
@@
    \left[ \begin{array}{c}
-      \widehat{\beta}^0_{2SLS} \\ \widehat{\lambda}
+      \widehat{\beta}^R_{2SLS} \\ \widehat{\lambda}
    \end{array} \right] 
@@
 
-The W2SLS estimator restricted by $R \beta^0 = q$ can be obtained by
+The W2SLS estimator restricted by $R \beta^R = q$ can be obtained by
 \begin{equation}
    \left[ \begin{array}{c}
-      \widehat{\beta}^0_{W2SLS} \\ \widehat{\lambda}
+      \widehat{\beta}^R_{W2SLS} \\ \widehat{\lambda}
    \end{array} \right]
@@
    \left[ \begin{array}{c}
-      \widehat{\beta}^0_{W2SLS} \\ \widehat{\lambda}
+      \widehat{\beta}^R_{W2SLS} \\ \widehat{\lambda}
    \end{array} \right] 
@@
 
-The standard 3SLS estimator restricted by $R \beta^0 = q$ can be obtained by
+The standard 3SLS estimator restricted by $R \beta^R = q$ can be obtained by
 \begin{equation}
    \left[ \begin{array}{c}
-      \widehat{\beta}^0_{3SLS} \\ \widehat{\lambda}
+      \widehat{\beta}^R_{3SLS} \\ \widehat{\lambda}
    \end{array} \right]
@@
    \left[ \begin{array}{c}
-      \widehat{\beta}^0_{3SLS} \\ \widehat{\lambda}
+      \widehat{\beta}^R_{3SLS} \\ \widehat{\lambda}
    \end{array} \right] 
@@
 \end{equation}
-The 3SLS-IV estimator restricted by $R \beta^0 = q$ can be obtained by
+The 3SLS-IV estimator restricted by $R \beta^R = q$ can be obtained by
 \begin{equation}
    \left[ \begin{array}{c}
-      \widehat{\beta}^0_{3SLS-IV} \\ \widehat{\lambda}
+      \widehat{\beta}^R_{3SLS-IV} \\ \widehat{\lambda}
    \end{array} \right]
@@
    \left[ \begin{array}{c}
-      \widehat{\beta}^0_{3SLS-IV} \\ \widehat{\lambda}
+      \widehat{\beta}^R_{3SLS-IV} \\ \widehat{\lambda}
    \end{array} \right] 
@@
    \left[ \begin{array}{c}
-      \widehat{\beta}^0_{3SLS-GMM} \\ \widehat{\lambda}
+      \widehat{\beta}^R_{3SLS-GMM} \\ \widehat{\lambda}
    \end{array} \right]
@@
    \left[ \begin{array}{c}
-      \widehat{\beta}^0_{3SLS-GMM} \\ \widehat{\lambda}
+      \widehat{\beta}^R_{3SLS-GMM} \\ \widehat{\lambda}
    \end{array} \right] 
@@
    \left[ \begin{array}{c}
-      \widehat{\beta}^0_{3SLS-Schmidt} \\ \widehat{\lambda}
+      \widehat{\beta}^R_{3SLS-Schmidt} \\ \widehat{\lambda}
    \end{array} \right]
@@
    \left[ \begin{array}{c}
-      \widehat{\beta}^0_{3SLS-Schmidt} \\ \widehat{\lambda}
+      \widehat{\beta}^R_{3SLS-Schmidt} \\ \widehat{\lambda}
    \end{array} \right] 
@@
    \left[ \begin{array}{c}
-      \widehat{\beta}^0_{3SLS-EViews} \\ \widehat{\lambda}
+      \widehat{\beta}^R_{3SLS-EViews} \\ \widehat{\lambda}
    \end{array} \right]
@@
    \left[ \begin{array}{c}
-      \widehat{X}' \OHat^{-1} \left( y - X \widehat{\beta}^0_{2SLS} \right)
+      \widehat{X}' \OHat^{-1} \left( y - X \widehat{\beta}^R_{2SLS} \right)
       \\ q 
@@
 \end{equation}
-where $\widehat{\beta}^0_{2SLS}$ is the restricted 2SLS estimator calculated
+where $\widehat{\beta}^R_{2SLS}$ is the restricted 2SLS estimator calculated
 by equation (\ref{eq:beta2SLSr}). 

### 2007-06-07 paper/trunk/JSS/response_A.tex
corrected small typo
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@717 591db2bd-e0a0-4d22-8c19-2192dbb7da31

--- 2007-06-07 paper/trunk/JSS/response_A.tex
@@
 %
+First, it has been made clearer
+that this section is about estimating coefficients under linear restrictions
+(and not about inference).
+Second, it has been explained more concisely,
+how the two methods to impose linear restrictions are implemented
+in \pkg{systemfit},
+i.e.\ that the rest of this (sub)section belongs to the second method
+(equation 31, former equation 33).
+
 \subsection{systemfitClassic (4.3)}

### 2007-06-07 paper/trunk/systemfit_restrictions.tex
replaced all \widehat{\beta} by the macro \bHat, which is defined as \hat{\beta}
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@719 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 (equations~\ref{eq:ols}--\ref{eq:3slsEViews}).
-This results in the linear independent parameter estimates $\widehat{\beta}^M$ and
+This results in the linear independent parameter estimates $\bHat^M$ and
 their covariance matrix.
@@
 \begin{equation}
-   \COVHat \left[ \widehat{\beta} \right]
-   = M \cdot \COVHat \left[ \widehat{\beta}^M \right] \cdot M' .
+   \COVHat \left[ \bHat \right]
+   = M \cdot \COVHat \left[ \bHat^M \right] \cdot M' .
 \end{equation}
@@
    \left[ \begin{array}{c}
-      \widehat{\beta}^R_{OLS} \\ \widehat{\lambda}
+      \bHat^R_{OLS} \\ \widehat{\lambda}
    \end{array} \right]
@@
    \left[ \begin{array}{c}
-      \widehat{\beta}^R_{OLS} \\ \widehat{\lambda}
+      \bHat^R_{OLS} \\ \widehat{\lambda}
    \end{array} \right] 
@@
    \left[ \begin{array}{c}
-      \widehat{\beta}^R_{WLS} \\ \widehat{\lambda}
+      \bHat^R_{WLS} \\ \widehat{\lambda}
    \end{array} \right]
@@
    \left[ \begin{array}{c}
-      \widehat{\beta}^R_{WLS} \\ \widehat{\lambda}
+      \bHat^R_{WLS} \\ \widehat{\lambda}
    \end{array} \right] 
@@
    \left[ \begin{array}{c}
-      \widehat{\beta}^R_{SUR} \\ \widehat{\lambda}
+      \bHat^R_{SUR} \\ \widehat{\lambda}
    \end{array} \right]
@@
    \left[ \begin{array}{c}
-      \widehat{\beta}^R_{SUR} \\ \widehat{\lambda}
+      \bHat^R_{SUR} \\ \widehat{\lambda}
    \end{array} \right] 
@@
    \left[ \begin{array}{c}
-      \widehat{\beta}^R_{2SLS} \\ \widehat{\lambda}
+      \bHat^R_{2SLS} \\ \widehat{\lambda}
    \end{array} \right]
@@
    \left[ \begin{array}{c}
-      \widehat{\beta}^R_{2SLS} \\ \widehat{\lambda}
+      \bHat^R_{2SLS} \\ \widehat{\lambda}
    \end{array} \right] 
@@
    \left[ \begin{array}{c}
-      \widehat{\beta}^R_{W2SLS} \\ \widehat{\lambda}
+      \bHat^R_{W2SLS} \\ \widehat{\lambda}
    \end{array} \right]
@@
    \left[ \begin{array}{c}
-      \widehat{\beta}^R_{W2SLS} \\ \widehat{\lambda}
+      \bHat^R_{W2SLS} \\ \widehat{\lambda}
    \end{array} \right] 
@@
    \left[ \begin{array}{c}
-      \widehat{\beta}^R_{3SLS} \\ \widehat{\lambda}
+      \bHat^R_{3SLS} \\ \widehat{\lambda}
    \end{array} \right]
@@
    \left[ \begin{array}{c}
-      \widehat{\beta}^R_{3SLS} \\ \widehat{\lambda}
+      \bHat^R_{3SLS} \\ \widehat{\lambda}
    \end{array} \right] 
@@
    \left[ \begin{array}{c}
-      \widehat{\beta}^R_{3SLS-IV} \\ \widehat{\lambda}
+      \bHat^R_{3SLS-IV} \\ \widehat{\lambda}
    \end{array} \right]
@@
    \left[ \begin{array}{c}
-      \widehat{\beta}^R_{3SLS-IV} \\ \widehat{\lambda}
+      \bHat^R_{3SLS-IV} \\ \widehat{\lambda}
    \end{array} \right] 
@@
    \left[ \begin{array}{c}
-      \widehat{\beta}^R_{3SLS-GMM} \\ \widehat{\lambda}
+      \bHat^R_{3SLS-GMM} \\ \widehat{\lambda}
    \end{array} \right]
@@
    \left[ \begin{array}{c}
-      \widehat{\beta}^R_{3SLS-GMM} \\ \widehat{\lambda}
+      \bHat^R_{3SLS-GMM} \\ \widehat{\lambda}
    \end{array} \right] 
@@
    \left[ \begin{array}{c}
-      \widehat{\beta}^R_{3SLS-Schmidt} \\ \widehat{\lambda}
+      \bHat^R_{3SLS-Schmidt} \\ \widehat{\lambda}
    \end{array} \right]
@@
    \left[ \begin{array}{c}
-      \widehat{\beta}^R_{3SLS-Schmidt} \\ \widehat{\lambda}
+      \bHat^R_{3SLS-Schmidt} \\ \widehat{\lambda}
    \end{array} \right] 
@@
    \left[ \begin{array}{c}
-      \widehat{\beta}^R_{3SLS-EViews} \\ \widehat{\lambda}
+      \bHat^R_{3SLS-EViews} \\ \widehat{\lambda}
    \end{array} \right]
@@
    \left[ \begin{array}{c}
-      \widehat{X}' \OHat^{-1} \left( y - X \widehat{\beta}^R_{2SLS} \right)
+      \widehat{X}' \OHat^{-1} \left( y - X \bHat^R_{2SLS} \right)
       \\ q 
@@
 \end{equation}
-where $\widehat{\beta}^R_{2SLS}$ is the restricted 2SLS estimator calculated
+where $\bHat^R_{2SLS}$ is the restricted 2SLS estimator calculated
 by equation (\ref{eq:beta2SLSr}). 

### 2007-06-07 paper/trunk/systemfit_other.tex
replaced all \widehat{\beta} by the macro \bHat now also in subsections 2.[3-7]
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@720 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 F = \frac{
-   ( R \widehat{\beta} - q )'
+   ( R \bHat - q )'
    ( R ( X' ( \widehat{\Sigma} \otimes I )^{-1} X )^{-1} R' )^{-1}
-   ( R \widehat{\beta} - q ) /
+   ( R \bHat - q ) /
    j
@@
 W =
-   ( R \widehat{\beta} - q )'
-   ( R \widehat{Cov} [ \widehat{\beta} ] R' )^{-1}
-   ( R \widehat{\beta} - q ) .
+   ( R \bHat - q )'
+   ( R \widehat{Cov} [ \bHat ] R' )^{-1}
+   ( R \bHat - q ) .
 \end{equation}
@@
 \begin{equation}
-  m = \left( \widehat{\beta}_{2SLS} - \widehat{\beta}_{3SLS} \right)^{'}
-      \left( \COV \left[ \widehat{\beta}_{2SLS} \right] -
-             \COV \left[ \widehat{\beta}_{3SLS} \right] \right)
-      \left( \widehat{\beta}_{2SLS} - \widehat{\beta}_{3SLS} \right) ,
+  m = \left( \bHat_{2SLS} - \bHat_{3SLS} \right)^{'}
+      \left( \COV \left[ \bHat_{2SLS} \right] -
+             \COV \left[ \bHat_{3SLS} \right] \right)
+      \left( \bHat_{2SLS} - \bHat_{3SLS} \right) ,
 \label{eq:hausman}
 \end{equation}
-where $\widehat{\beta}_{2SLS}$ and $\COV \left[ \widehat{\beta}_{2SLS} \right]$ are the estimated
+where $\bHat_{2SLS}$ and $\COV \left[ \bHat_{2SLS} \right]$ are the estimated
 coefficient and covariance matrix from 2SLS estimation, and
-$\widehat{\beta}_{3SLS}$ and $\COV \left[ \widehat{\beta}_{3SLS} \right]$ are the estimated
+$\bHat_{3SLS}$ and $\COV \left[ \bHat_{3SLS} \right]$ are the estimated
 coefficients and covariance matrix from 3SLS estimation.

### 2007-06-07 paper/trunk/systemfit_other.tex
replaced several \widehat{*} in systemfit_other.tex by the respective macros \*Hat (I forgot this in some previous revisions)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@724 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 Generally, the estimated covariance matrix of the residuals
-($\widehat{\Sigma} = \left[ \widehat{\sigma}_{ij} \right]$)
+($\SHat = \left[ \sHat_{ij} \right]$)
 can be calculated from the residuals of a first-step OLS or 2SLS estimation.
@@
 \begin{equation}
-   \widehat{\sigma}_{ij} = \frac{ \widehat{u}_i' \widehat{u}_j }{ T }
+   \sHat_{ij} = \frac{ \widehat{u}_i' \widehat{u}_j }{ T }
    \label{eq:rcovNoDfCor}
@@
 \begin{equation}
-   \widehat{\sigma}_{ij} = \frac{ \widehat{u}_i' \widehat{u}_j }
+   \sHat_{ij} = \frac{ \widehat{u}_i' \widehat{u}_j }
    { \sqrt{ \left( T - K_i \right) \cdot \left( T - K_j \right) } }
@@
 \begin{equation}
-   \widehat{\sigma}_{ij} = \frac{ \widehat{u}_i' \widehat{u}_j }
+   \sHat_{ij} = \frac{ \widehat{u}_i' \widehat{u}_j }
    { T - \max \left( K_i , K_j \right) } \; ,
@@
 \begin{eqnarray}
-   \widehat{\sigma}_{ij} & = & 
+   \sHat_{ij} & = &
    \frac{ \widehat{u}_i' \widehat{u}_j } 
@@
 This yields an unbiased estimator for all elements of $\Sigma$,
-but even if $\widehat{\Sigma}$ is an unbiased estimator of $\Sigma$, 
-its inverse $\widehat{\Sigma}^{-1}$ is not an unbiased estimator 
+but even if $\SHat$ is an unbiased estimator of $\Sigma$,
+its inverse $\SHat^{-1}$ is not an unbiased estimator
 of $\Sigma^{-1}$ \citep[p.\ 322]{theil71}.
@@
 \begin{equation}
-   R_*^2 = 1 - \frac{ \widehat{u}' \widehat{ \Omega }^{-1} \widehat{u} }
-   { y' \left( \widehat{ \Sigma }^{-1} \otimes
+   R_*^2 = 1 - \frac{ \widehat{u}' \OHat^{-1} \widehat{u} }
+   { y' \left( \SHat^{-1} \otimes
    \left( I - \frac{i i'}{T} \right) \right) y }
@@
    ( R \bHat - q )'
-   ( R ( X' ( \widehat{\Sigma} \otimes I )^{-1} X )^{-1} R' )^{-1}
+   ( R ( X' ( \SHat \otimes I )^{-1} X )^{-1} R' )^{-1}
    ( R \bHat - q ) /
@@
 }{
-   \widehat{u}' ( \widehat{\Sigma} \otimes I )^{-1} \widehat{u} /
+   \widehat{u}' ( \SHat \otimes I )^{-1} \widehat{u} /
    ( G \cdot T - K )
@@
 $K$ is the total number of estimated coefficients, and
-$\widehat{\Sigma}$ is the estimated residual covariance matrix
+$\SHat$ is the estimated residual covariance matrix
 used in the estimation.
@@
    ( R \bHat - q )'
-   ( R \widehat{Cov} [ \bHat ] R' )^{-1}
+   ( R \COVHat [ \bHat ] R' )^{-1}
    ( R \bHat - q ) .
@@
 LR = T \cdot \left(
-   log \left| \widehat{ \Sigma }_r \right|
-   - log \left| \widehat{ \Sigma }_u \right|
+   log \left| \SHat_r \right|
+   - log \left| \SHat_u \right|
    \right) ,
@@
 where $T$ is the number of observations per equation, and 
-$\widehat{\Sigma}_r$ and $\widehat{\Sigma}_u$ are
+$\SHat_r$ and $\SHat_u$ are
 the residual covariance matrices calculated by formula (\ref{eq:rcovNoDfCor})

### 2007-06-07 paper/trunk/systemfit_other.tex
replaced all \widehat{u} by the macro \uHat, which is defined as \hat{u}
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@725 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 \begin{equation}
-   \sHat_{ij} = \frac{ \widehat{u}_i' \widehat{u}_j }{ T }
+   \sHat_{ij} = \frac{ \uHat_i' \uHat_j }{ T }
    \label{eq:rcovNoDfCor}
@@
 \begin{equation}
-   \sHat_{ij} = \frac{ \widehat{u}_i' \widehat{u}_j }
+   \sHat_{ij} = \frac{ \uHat_i' \uHat_j }
    { \sqrt{ \left( T - K_i \right) \cdot \left( T - K_j \right) } }
@@
 \begin{equation}
-   \sHat_{ij} = \frac{ \widehat{u}_i' \widehat{u}_j }
+   \sHat_{ij} = \frac{ \uHat_i' \uHat_j }
    { T - \max \left( K_i , K_j \right) } \; ,
@@
    \sHat_{ij} & = &
-   \frac{ \widehat{u}_i' \widehat{u}_j } 
+   \frac{ \uHat_i' \uHat_j }
    { T - K_i - K_j + tr \left[ X_i \left( X_i' X_i \right)^{-1}
@@
    & = &
-   \frac{ \widehat{u}_i' \widehat{u}_j } 
+   \frac{ \uHat_i' \uHat_j }
    { T - K_i - K_j + tr \left[ \left( X_i' X_i \right)^{-1}
@@
 the means of the residuals of each equation are not equal to zero
-$( \overline{ \widehat{u} }_i \neq 0 )$.
+$( \overline{ \uHat }_i \neq 0 )$.
 Therefore, it might be argued
@@
 by subtracting the means from the residuals
-and substituting $\widehat{u}_i - \overline{ \widehat{u} }_i$
-for $\widehat{u}_i$ in (\ref{eq:rcovNoDfCor}--\ref{eq:rcovTheil}).
+and substituting $\uHat_i - \overline{ \uHat }_i$
+for $\uHat_i$ in (\ref{eq:rcovNoDfCor}--\ref{eq:rcovTheil}).
 
@@
 \begin{equation}
-   R_i^2 = 1 - \frac{ \widehat{u}_i' \widehat{u}_i }
+   R_i^2 = 1 - \frac{ \uHat_i' \uHat_i }
    { ( y_i - \overline{y_i} )' ( y_i - \overline{y_i} ) } \; ,
@@
 \begin{equation}
-   R_*^2 = 1 - \frac{ \widehat{u}' \OHat^{-1} \widehat{u} }
+   R_*^2 = 1 - \frac{ \uHat' \OHat^{-1} \uHat }
    { y' \left( \SHat^{-1} \otimes
@@
 }{
-   \widehat{u}' ( \SHat \otimes I )^{-1} \widehat{u} /
+   \uHat' ( \SHat \otimes I )^{-1} \uHat /
    ( G \cdot T - K )

### 2007-06-07 paper/trunk/systemfit_other.tex
a small cosmetic improvement of an equation
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@726 591db2bd-e0a0-4d22-8c19-2192dbb7da31
substituted \COVHat for all \COV in the description of the Hausman test
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@728 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
    ( R \bHat - q )'
-   ( R \COVHat [ \bHat ] R' )^{-1}
+   ( R \, \COVHat [ \bHat ] \, R' )^{-1}
    ( R \bHat - q ) .
@@
   m = \left( \bHat_{2SLS} - \bHat_{3SLS} \right)^{'}
-      \left( \COV \left[ \bHat_{2SLS} \right] -
-             \COV \left[ \bHat_{3SLS} \right] \right)
+      \left( \COVHat \left[ \bHat_{2SLS} \right] -
+             \COVHat \left[ \bHat_{3SLS} \right] \right)
       \left( \bHat_{2SLS} - \bHat_{3SLS} \right) ,
@@
 \end{equation}
-where $\bHat_{2SLS}$ and $\COV \left[ \bHat_{2SLS} \right]$ are the estimated
+where $\bHat_{2SLS}$ and $\COVHat \left[ \bHat_{2SLS} \right]$ are the estimated
 coefficient and covariance matrix from 2SLS estimation, and
-$\bHat_{3SLS}$ and $\COV \left[ \bHat_{3SLS} \right]$ are the estimated
+$\bHat_{3SLS}$ and $\COVHat \left[ \bHat_{3SLS} \right]$ are the estimated
 coefficients and covariance matrix from 3SLS estimation.

### 2007-06-08 paper/trunk/systemfit_code.tex
changed paper according to changes of the package: systemfitClassic does not longer exist, but this kind of analysis can now be done directly in systemfit (if argument data is of class "pdata.frame")
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@730 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 
-The \pkg{systemfit} package includes functions
-to estimate systems of equations
-(\code{systemfit}, \code{systemfitClassic}).
+The most important function in the \pkg{systemfit} package
+is called \code{systemfit}.
+It can be used to estimate systems of linear equations
+by several different estimation methods.
 Furthermore, this package provides some methods
@@
 can be specified by the user.
+If \code{systemfit} is applied to
+a (classical) °Seemingly Unrelated Regression° analysis with panel-like data,
+it calls the hidden internal function \code{.systemfitPanel},
+which reshapes the data, creates the formulas to be estimated,
+and --- if requested --- specifies restrictions
+to ensure that the parameters of all individuals are equal.
+
 The \code{systemfit} function returns many objects
@@
 that is included in the package.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection[Wrapper function systemfitClassic]{The wrapper function
-   \code{systemfitClassic}}
-Furthermore, the \pkg{systemfit} package includes the function
-\code{systemfitClassic}.
-This is a wrapper function for \code{systemfit}
-that can be applied to (classical) panel-like data in long format
-if the regressors are the same for all equations.
-The data are reshaped and the formulas are modified to enable
-an estimation using the standard \code{systemfit} function.
-The user can specify whether the coefficients should be restricted
-to be equal in all equations.
 

### 2007-06-11 paper/trunk/JSS/response_A.tex
added note that we have streamlined the notations of the betas (in rev. 163)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@729 591db2bd-e0a0-4d22-8c19-2192dbb7da31
response to referee A: added answer that the function systemfitClassic() has been removed and that its functionality has been overtaken by systemfit()
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@731 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 (equation 31, former equation 33).
+Finally, we have streamlined the notations of the $\beta$s.
 
@@
 %     using it).
+The function \code{systemfitClassic} has been removed.
+%Its functionality has been taken over by \code{systemfit}.
+A (classical) °Seemingly Unrelated Regression° analysis
+with panel-like data
+can now be done by \code{systemfit} itself.
+If the argument \code{data} is an object of class \code{pdata.frame}
+(°panel data frame°, created with the function \code{pdata.frame}
+from the \proglang{R} package \pkg{plm} \citep{r-plm-0.1-2}),
+\code{systemfit} estimates the provided formula for each
+(cross-section) individual.
+
 %     Second, the interface should also be re-designed: the standard in R is
@@
 
-% \bibliographystyle{}
-% \bibliography{}
+\bibliographystyle{jss}
+\bibliography{../systemfit}
 

### 2007-06-11 paper/trunk/JSS/response_A.tex
response to referee A: added answer that we consider it not desirable to relax the requirement of using the same formula for all equations in (classical) °Seemingly Unrelated Regression° analyses with panel-like data.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@732 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 (cross-section) individual.
+
+In our opinion,
+it is not desirable to relax the requirement
+of using the same formula for all equations
+in (classical) °Seemingly Unrelated Regression° analyses
+with panel-like data.
+First, it would be rather cumbersome for the user to specify,
+which formula should be applied to which (cross-section) individual.
+Second, this type of analysis can be easily done
+by reshaping the data (e.g.\ with \code{reshape})
+and then using (standard) \code{systemfit}.
+Third, we think that this feature would rarely be used.
 

### 2007-06-11 paper/trunk/JSS/response_A.tex
response to referee A: added answer that all remaining concerns regarding systemfitClassic() have been resolved by its removal
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@733 591db2bd-e0a0-4d22-8c19-2192dbb7da31
response to referee A: added answer that the naming of the coefficients has been improved according to the suggestions of this referee
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@734 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 %
+All remaining concerns regarding \code{systemfitClassic}
+(order of arguments;
+argument \code{data} being mandatory;
+arguments \code{eqnVar} and \code{timeVar} being strings)
+have been resolved by its removal.
+
 \subsection{Testing Linear Restrictions (4.4)}
@@
 %     would be preferable.
+The naming of the coefficients has been improved
+according to the suggestions of this referee:
+They are now \code{<eqname>}\verb!_!\code{<varname>},
+where \code{<eqname>} is the equation name,
+which is set to \code{"eq1"}, \code{"eq2"} etc.\
+if no equation names are supplied by the user.
+
 % 

### 2007-06-11 trunk/R/systemfit.R
started to modify systemfit so that the coef names can be improved for panel-like data (not finished)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@368 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@ -45,7 +45,9 @@ systemfit <- function(  eqns,
       stop( "The methods '2SLS', 'W2SLS', '3SLS', and 'W3SLS' need instruments!" )
    }
 
+   panelLike <- FALSE
    if( class( data )[1] == "pdata.frame" ) {
+      panelLike <- TRUE
       if( !is.null( TX ) && pooled ){
          stop( "argument 'TX' cannot be used for pooled estimation",
             " of panel-like data" )
@@ -425,9 +427,10 @@ systemfit <- function(  eqns,
               # covariance matrix of estimated coefficients of equation i
 
     # set names
-    names( coefEqI ) <- colnames( xMatEq[[i]] )
-    colnames( bcovi ) <- colnames( xMatEq[[i]] )
-    rownames( bcovi ) <- colnames( xMatEq[[i]] )
+    xNamesEqI <- colnames( xMatEq[[i]] )
+    names( coefEqI ) <- xNamesEqI
+    colnames( bcovi ) <- xNamesEqI
+    rownames( bcovi ) <- xNamesEqI
 
     ssr    <- sum(residi[[i]]^2)                         # sum of squared residuals
     sigma  <- sqrt( ssr / df[i] ) # estimated standand deviation of residuals

### 2007-06-12 paper/trunk/JSS/response_A.tex
response to referee A: added answer that the naming of the coefficients from an analysis with panel-like data has been improved according to the suggestions of this referee
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@735 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 according to the suggestions of this referee:
-They are now \code{<eqname>}\verb!_!\code{<varname>},
+they are now \code{<eqname>}\verb!_!\code{<varname>},
 where \code{<eqname>} is the equation name,
@@
 if no equation names are supplied by the user.
+In the case of a (classical) °Seemingly Unrelated Regression° analysis
+with panel-like data,
+all redundant parts of the coefficient names have been removed:
+the coefficient names in the vector of all coefficients are
+\code{<individual>}\verb!_!\code{<varname>}
+and the coefficient names in the vector of the coefficients of a single
+individual are just \code{<varname>}.
 

### 2007-06-12 paper/trunk/JSS/response_A.tex
response to referee A: revised answer that we have changed the order of arguments of systemfit (and of systemfitClassic)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@736 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 %     as the first argument and to have then some `method' argument.
-We have reversed order of the arguments method and eqns/formula
-in systemfit() and systemfitClassic().
+We have reversed the order of arguments \code{method} and \code{eqns}
+of \code{systemfit()}.
+It is no longer necessary to reverse the order of the arguments
+of \code{systemfitClassic()},
+because this function has been removed.
+
 % 

### 2007-06-13 trunk/R/linear.hypothesis.R
replaced "waldtest.systemfit" by method "linear.hypothesis" for objects of class "systemfit"; package "car" is loaded automatically now (required package)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@374 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@ -1,0 +1,35 @@
+linear.hypothesis.systemfit <- function( model,
+      hypothesis.matrix, rhs = NULL, test = c( "F", "Chisq" ), ... ){
+
+   thisCall <- match.call()
+
+   result <- car:::linear.hypothesis.default( model,
+      hypothesis.matrix = hypothesis.matrix, rhs = rhs, test = test, ... )
+
+   if( test == "Chisq" ){
+      attributes( result )$heading[ 1 ] <-
+         "Linear hypothesis test (Wald-test)\n\nHypothesis:"
+      if( "model" %in% names( thisCall ) ) {
+         modelPos <- grep( "^Model 1: model\n", attributes( result )$heading )
+         if( class( thisCall$model ) == "name" ) {
+            modelName <- as.character( thisCall$model )
+         } else if( class( thisCall$model ) == "call" ) {
+            modelName <- format( thisCall$model )
+         } else {
+            modelName <- thisCall$model
+         }
+         attributes( result )$heading[ modelPos[ 1 ] ] <-
+            sub( "^Model 1: model\n",
+               paste( "Model 1: ", modelName, "\n", sep = "" ),
+               attributes( result )$heading[ modelPos[ 1 ] ] )
+# attributes( result )$thisCall <- thisCall
+      }
+   } else if ( test == "F" ) {
+      stop( "F-test with linear.hypothesis is not implemented yet" )
+   } else {
+      stop( "unknown test statistic '", test, "'. Please use 'F' or 'Chisq'" )
+   }
+
+   return( result )
+}
+

### 2007-06-13 trunk/R/waldtest.R

@@ -1,28 +1,0 @@
-waldtest.systemfit <- function( object, restrictions,
-   restrict.rhs = rep( 0, nrow( restrictions ) ) ){
-
-   coef <- coef( object )
-   vcov <- vcov( object )
-
-   result <- list()
-
-   result$nRestr <- nrow( restrictions )
-
-   result$statistic <- t( restrictions %*% coef - restrict.rhs ) %*%
-      solve( restrictions %*% vcov %*% t( restrictions ) ) %*%
-      ( restrictions %*% coef - restrict.rhs )
-
-   result$p.value <- 1 - pchisq( result$statistic, result$nRestr )
-
-   class( result ) <- "waldtest.systemfit"
-   return( result )
-}
-
-print.waldtest.systemfit <- function( x, digits = 4, ... ){
-   cat( "\n", "Wald-test for linear parameter restrictions",
-      " in equation systems\n", sep = "" )
-   cat( "Wald-statistic:", formatC( x$statistic, digits = digits ), "\n" )
-   cat( "degrees of freedom:", x$nRestr, "\n" )
-   cat( "p-value:", formatC( x$p.value, digits = digits ), "\n\n" )
-   invisible( x )
-}

### 2007-06-13 trunk/tests/test_2sls.R

@@ -424,21 +424,21 @@ print( ftest.systemfit( fit2sls1, restr2m, restr2q ) )
 
 ## ************** Wald tests ****************
 # testing first restriction
-print( waldtest.systemfit( fit2sls1, restrm ) )
-print( waldtest.systemfit( fit2sls1s, restrm ) )
-print( waldtest.systemfit( fit2sls1p, restrm ) )
-print( waldtest.systemfit( fit2sls1r, restrm ) )
+print( linear.hypothesis( fit2sls1, restrm, test = "Chisq" ) )
+print( linear.hypothesis( fit2sls1s, restrm, test = "Chisq" ) )
+print( linear.hypothesis( fit2sls1p, restrm, test = "Chisq" ) )
+print( linear.hypothesis( fit2sls1r, restrm, test = "Chisq" ) )
 
 # testing second restriction
 # first restriction not imposed
-print( waldtest.systemfit( fit2sls1, restrOnly2m, restrOnly2q ) )
+print( linear.hypothesis( fit2sls1, restrOnly2m, restrOnly2q, test = "Chisq" ) )
 # first restriction imposed
-print( waldtest.systemfit( fit2sls2, restrOnly2m, restrOnly2q ) )
-print( waldtest.systemfit( fit2sls2r, restrOnly2m, restrOnly2q ) )
-print( waldtest.systemfit( fit2sls3, restrOnly2m, restrOnly2q ) )
+print( linear.hypothesis( fit2sls2, restrOnly2m, restrOnly2q, test = "Chisq" ) )
+print( linear.hypothesis( fit2sls2r, restrOnly2m, restrOnly2q, test = "Chisq" ) )
+print( linear.hypothesis( fit2sls3, restrOnly2m, restrOnly2q, test = "Chisq" ) )
 
 # testing both of the restrictions
-print( waldtest.systemfit( fit2sls1, restr2m, restr2q ) )
+print( linear.hypothesis( fit2sls1, restr2m, restr2q, test = "Chisq" ) )
 
 
 ## **************** model frame ************************

### 2007-06-13 trunk/tests/test_3sls.R

@@ -486,29 +486,29 @@ print( ftest.systemfit( fit3slsd[[ 4 ]]$e1e, restr2m, restr2q ) )
 
 ## ************** Wald tests ****************
 # testing first restriction
-print( waldtest.systemfit( fit3sls[[ 1 ]]$e1, restrm ) )
-print( waldtest.systemfit( fit3sls[[ 2 ]]$e1e, restrm ) )
-print( waldtest.systemfit( fit3sls[[ 3 ]]$e1c, restrm ) )
-print( waldtest.systemfit( fit3slsi[[ 4 ]]$e1, restrm ) )
-print( waldtest.systemfit( fit3slsd[[ 5 ]]$e1e, restrm ) )
+print( linear.hypothesis( fit3sls[[ 1 ]]$e1, restrm, test = "Chisq" ) )
+print( linear.hypothesis( fit3sls[[ 2 ]]$e1e, restrm, test = "Chisq" ) )
+print( linear.hypothesis( fit3sls[[ 3 ]]$e1c, restrm, test = "Chisq" ) )
+print( linear.hypothesis( fit3slsi[[ 4 ]]$e1, restrm, test = "Chisq" ) )
+print( linear.hypothesis( fit3slsd[[ 5 ]]$e1e, restrm, test = "Chisq" ) )
 
 # testing second restriction
 # first restriction not imposed
-print( waldtest.systemfit( fit3sls[[ 5 ]]$e1c, restrOnly2m, restrOnly2q ) )
-print( waldtest.systemfit( fit3slsi[[ 1 ]]$e1e, restrOnly2m, restrOnly2q ) )
-print( waldtest.systemfit( fit3slsd[[ 2 ]]$e1, restrOnly2m, restrOnly2q ) )
+print( linear.hypothesis( fit3sls[[ 5 ]]$e1c, restrOnly2m, restrOnly2q, test = "Chisq" ) )
+print( linear.hypothesis( fit3slsi[[ 1 ]]$e1e, restrOnly2m, restrOnly2q, test = "Chisq" ) )
+print( linear.hypothesis( fit3slsd[[ 2 ]]$e1, restrOnly2m, restrOnly2q, test = "Chisq" ) )
 # first restriction imposed
-print( waldtest.systemfit( fit3sls[[ 4 ]]$e2, restrOnly2m, restrOnly2q ) )
[…]

### 2007-06-13 trunk/tests/test_ols.R

@@ -350,20 +350,20 @@ print( ftest.systemfit( fitols1, restr2m, restr2q ) )
 
 ## ************** Wald tests ****************
 # testing first restriction
-print( waldtest.systemfit( fitols1, restrm ) )
-print( waldtest.systemfit( fitols1s, restrm ) )
-print( waldtest.systemfit( fitols1p, restrm ) )
-print( waldtest.systemfit( fitols1r, restrm ) )
+print( linear.hypothesis( fitols1, restrm, test = "Chisq" ) )
+print( linear.hypothesis( fitols1s, restrm, test = "Chisq" ) )
+print( linear.hypothesis( fitols1p, restrm, test = "Chisq" ) )
+print( linear.hypothesis( fitols1r, restrm, test = "Chisq" ) )
 
 # testing second restriction
 # first restriction not imposed
-print( waldtest.systemfit( fitols1, restrOnly2m, restrOnly2q ) )
+print( linear.hypothesis( fitols1, restrOnly2m, restrOnly2q, test = "Chisq" ) )
 # first restriction imposed
-print( waldtest.systemfit( fitols2, restrOnly2m, restrOnly2q ) )
-print( waldtest.systemfit( fitols3, restrOnly2m, restrOnly2q ) )
+print( linear.hypothesis( fitols2, restrOnly2m, restrOnly2q, test = "Chisq" ) )
+print( linear.hypothesis( fitols3, restrOnly2m, restrOnly2q, test = "Chisq" ) )
 
 # testing both of the restrictions
-print( waldtest.systemfit( fitols1, restr2m, restr2q ) )
+print( linear.hypothesis( fitols1, restr2m, restr2q, test = "Chisq" ) )
 
 
 ## ****************** model frame **************************

### 2007-06-13 trunk/tests/test_sur.R

@@ -497,24 +497,24 @@ print( ftest.systemfit( fitsuri1e2, restr2m, restr2q ) )
 
 ## ************** Wald tests ****************
 # testing first restriction
-print( waldtest.systemfit( fitsur1, restrm ) )
-print( waldtest.systemfit( fitsur1r2, restrm ) )
-print( waldtest.systemfit( fitsuri1e2, restrm ) )
-print( waldtest.systemfit( fitsuri1r3, restrm ) )
+print( linear.hypothesis( fitsur1, restrm, test = "Chisq" ) )
+print( linear.hypothesis( fitsur1r2, restrm, test = "Chisq" ) )
+print( linear.hypothesis( fitsuri1e2, restrm, test = "Chisq" ) )
+print( linear.hypothesis( fitsuri1r3, restrm, test = "Chisq" ) )
 
 # testing second restriction
 # first restriction not imposed
-print( waldtest.systemfit( fitsur1e2, restrOnly2m, restrOnly2q ) )
-print( waldtest.systemfit( fitsuri1, restrOnly2m, restrOnly2q ) )
+print( linear.hypothesis( fitsur1e2, restrOnly2m, restrOnly2q, test = "Chisq" ) )
+print( linear.hypothesis( fitsuri1, restrOnly2m, restrOnly2q, test = "Chisq" ) )
 # first restriction imposed
-print( waldtest.systemfit( fitsur2, restrOnly2m, restrOnly2q ) )
-print( waldtest.systemfit( fitsur3, restrOnly2m, restrOnly2q ) )
-print( waldtest.systemfit( fitsuri2e, restrOnly2m, restrOnly2q ) )
-print( waldtest.systemfit( fitsuri3e, restrOnly2m, restrOnly2q ) )
+print( linear.hypothesis( fitsur2, restrOnly2m, restrOnly2q, test = "Chisq" ) )
+print( linear.hypothesis( fitsur3, restrOnly2m, restrOnly2q, test = "Chisq" ) )
+print( linear.hypothesis( fitsuri2e, restrOnly2m, restrOnly2q, test = "Chisq" ) )
+print( linear.hypothesis( fitsuri3e, restrOnly2m, restrOnly2q, test = "Chisq" ) )
 
 # testing both of the restrictions
-print( waldtest.systemfit( fitsur1r3, restr2m, restr2q ) )
-print( waldtest.systemfit( fitsuri1e2, restr2m, restr2q ) )
[…]

### 2007-06-13 trunk/tests/test_w2sls.R

@@ -338,22 +338,22 @@ print( ftest.systemfit( fitw2slsd1, restr2m, restr2q ) )
 
 ## ************** Wald tests ****************
 # testing first restriction
-print( waldtest.systemfit( fitw2sls1, restrm ) )
-print( waldtest.systemfit( fitw2slsd1e, restrm ) )
+print( linear.hypothesis( fitw2sls1, restrm, test = "Chisq" ) )
+print( linear.hypothesis( fitw2slsd1e, restrm, test = "Chisq" ) )
 
 # testing second restriction
 # first restriction not imposed
-print( waldtest.systemfit( fitw2sls1e, restrOnly2m, restrOnly2q ) )
-print( waldtest.systemfit( fitw2slsd1, restrOnly2m, restrOnly2q ) )
+print( linear.hypothesis( fitw2sls1e, restrOnly2m, restrOnly2q, test = "Chisq" ) )
+print( linear.hypothesis( fitw2slsd1, restrOnly2m, restrOnly2q, test = "Chisq" ) )
 # first restriction imposed
-print( waldtest.systemfit( fitw2sls2, restrOnly2m, restrOnly2q ) )
-print( waldtest.systemfit( fitw2sls3, restrOnly2m, restrOnly2q ) )
-print( waldtest.systemfit( fitw2slsd2e, restrOnly2m, restrOnly2q ) )
-print( waldtest.systemfit( fitw2slsd3e, restrOnly2m, restrOnly2q ) )
+print( linear.hypothesis( fitw2sls2, restrOnly2m, restrOnly2q, test = "Chisq" ) )
+print( linear.hypothesis( fitw2sls3, restrOnly2m, restrOnly2q, test = "Chisq" ) )
+print( linear.hypothesis( fitw2slsd2e, restrOnly2m, restrOnly2q, test = "Chisq" ) )
+print( linear.hypothesis( fitw2slsd3e, restrOnly2m, restrOnly2q, test = "Chisq" ) )
 
 # testing both of the restrictions
-print( waldtest.systemfit( fitw2sls1e, restr2m, restr2q ) )
-print( waldtest.systemfit( fitw2slsd1, restr2m, restr2q ) )
+print( linear.hypothesis( fitw2sls1e, restr2m, restr2q, test = "Chisq" ) )
+print( linear.hypothesis( fitw2slsd1, restr2m, restr2q, test = "Chisq" ) )
 
 
 ## ****************** model frame **************************

### 2007-06-13 trunk/tests/test_wls.R

@@ -387,22 +387,22 @@ print( ftest.systemfit( fitwlsi1, restr2m, restr2q ) )
 
 ## ************** Wald tests ****************
 # testing first restriction
-print( waldtest.systemfit( fitwls1, restrm ) )
-print( waldtest.systemfit( fitwlsi1e, restrm ) )
+print( linear.hypothesis( fitwls1, restrm, test = "Chisq" ) )
+print( linear.hypothesis( fitwlsi1e, restrm, test = "Chisq" ) )
 
 # testing second restriction
 # first restriction not imposed
-print( waldtest.systemfit( fitwls1e, restrOnly2m, restrOnly2q ) )
-print( waldtest.systemfit( fitwlsi1, restrOnly2m, restrOnly2q ) )
+print( linear.hypothesis( fitwls1e, restrOnly2m, restrOnly2q, test = "Chisq" ) )
+print( linear.hypothesis( fitwlsi1, restrOnly2m, restrOnly2q, test = "Chisq" ) )
 # first restriction imposed
-print( waldtest.systemfit( fitwls2, restrOnly2m, restrOnly2q ) )
-print( waldtest.systemfit( fitwls3, restrOnly2m, restrOnly2q ) )
-print( waldtest.systemfit( fitwlsi2e, restrOnly2m, restrOnly2q ) )
-print( waldtest.systemfit( fitwlsi3e, restrOnly2m, restrOnly2q ) )
+print( linear.hypothesis( fitwls2, restrOnly2m, restrOnly2q, test = "Chisq" ) )
+print( linear.hypothesis( fitwls3, restrOnly2m, restrOnly2q, test = "Chisq" ) )
+print( linear.hypothesis( fitwlsi2e, restrOnly2m, restrOnly2q, test = "Chisq" ) )
+print( linear.hypothesis( fitwlsi3e, restrOnly2m, restrOnly2q, test = "Chisq" ) )
 
 # testing both of the restrictions
-print( waldtest.systemfit( fitwls1e, restr2m, restr2q ) )
-print( waldtest.systemfit( fitwlsi1, restr2m, restr2q ) )
+print( linear.hypothesis( fitwls1e, restr2m, restr2q, test = "Chisq" ) )
+print( linear.hypothesis( fitwlsi1, restr2m, restr2q, test = "Chisq" ) )
 
 
 ## ****************** model frame **************************

### 2007-06-14 paper/trunk/systemfit_code.tex
replaced "waldtest.systemfit" by "linear.hypothesis.systemfit" or "linear.hypothesis( ... , test = "Chisq" )"
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@740 591db2bd-e0a0-4d22-8c19-2192dbb7da31
changes due to the replacement of "ftest.systemfit" by "linear.hypothesis.systemfit"; explained the implementation and usage of "linear.hypothesis.systemfit" in more detail
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@741 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 \subsection{Statistical tests}
-The statistical tests described in
-sections~\ref{sec:testingRestrictions} and~\ref{sec:hausman}
-are implemented as specified in these sections.
-The functions \code{ftest.systemfit}, \code{waldtest.systemfit} and
-\code{lrtest.systemfit} test linear restrictions on the
-estimated parameters.
+This section explains the implementation of the statistical tests described in
+sections~\ref{sec:testingRestrictions} and~\ref{sec:hausman}.
+The method \code{linear.hypothesis.systemfit}
+can be used to test linear restrictions on the estimated parameters
+by an $F$ test or a Wald/$\chi^2$ test.
+Internally, the $F$ statistic is computed by the hidden function
+\code{.ftest.systemfit}
+and the Wald/$\chi^2$ statistic is computed by the method
+\code{linear.hypothesis.default} of the \pkg{car} package.
+The function \code{lrtest.systemfit}
+computes the Likelihood-ratio (LR) test statistic.
 On the other hand, the function \code{hausman.systemfit}

### 2007-06-21 paper/trunk/JSS/response_A.tex
response to reviewer A: revised introduction to this response; added justification that we kept section 2 (Statistical background) as it is (and do *not* merge it with section "usage")
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@743 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 
-We thank this reviewer for a very helpful set of comments and suggestions
-for improving the paper.
+We thank this reviewer for her/his very helpful comments and suggestions
+that assisted us to considerably improve the software as well as the paper.
 The reviewer has made one overall comment, 9 major comments and
@@
 % description of the theory in a more general and unified way while weaving
-% it with the translation into computational tools. As for the software itself,
+% it with the translation into computational tools.
+We consider the presentation of the theoretical background
+in a separate section
+as an advantage rather than as a drawback.
+As the formulas to calculate the estimators and their covariances
+are not essential to \emph{use} \pkg{systemfit},
+we think that the formulas and the usage should not be presented together.
+Furthermore, our structure allows readers
+who already know the theoretical background
+to skip this entire section
+and go immediately into the section they are interested in
+(e.g.\ section °Using systemfit°).
+Hence, they do not have to scan a large composite section
+just to figure out a few commands of \pkg{systemfit}.
+Readers who have skipped the section °Statistical background°,
+can easily go back to this section
+if they are interested in a particular part of the theory,
+because all parts in section °Using systemfit°
+have references to their theoretical background
+in section °Statistical background°.
+
+% As for the software itself,
 % several modifications are necessary to make it more intuitive for R users

### 2007-06-22 paper/trunk/JSS/response_A.tex
response to reviewer A: revised paragraph that justifies that we keep section 2 (Statistical background) as it is (and do *not* merge it with section "usage")
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@744 591db2bd-e0a0-4d22-8c19-2192dbb7da31

--- 2007-06-12 paper/trunk/JSS/response_A.tex
@@
 % description of the theory in a more general and unified way while weaving
-% it with the translation into computational tools. As for the software itself,
+% it with the translation into computational tools.
+We appreciate this recommendation of the reviewer very much,
+but we consider the presentation of the theoretical background
+in a separate section
+as an advantage rather than as a drawback.
+As the formulas to calculate the estimators and their covariances
+are not essential to \emph{use} \pkg{systemfit},
+we think that the formulas and the usage should be presented separately.
+Furthermore, our structure allows readers
+who already know the theoretical background
+(or who do not want to know it)
+to skip this entire section
+and to go immediately into the section they are interested in
+(e.g.\ section °Using systemfit°).
+Hence, they do not have to scan a large composite section
+just to figure out a few pieces of information
+(e.g.\ how to use the commands of \pkg{systemfit}).
+Readers who have skipped the °Statistical background°
+can easily go back to this section
+if they are interested in a particular part of the theory,
+because all parts in section °Using systemfit°
+have references to their theoretical background
+in section °Statistical background°.
+
+% As for the software itself,
 % several modifications are necessary to make it more intuitive for R users

### 2007-06-22 paper/trunk/JSS/response_A.tex
response to reviewer A: added answer that we have considerably modified the software package systemfit to make it more intuitive for R users
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@745 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 % tools. More details are described below.
+As suggested by this reviewer,
+the software package \pkg{systemfit} has been considerably modified
+to make it more intuitive for \proglang{R} users.
+We tried to make the usage of \pkg{systemfit}
+as close as possible to the primary regression tools in \proglang{R}
+(e.g.\ \code{lm}).
+More details of theses modifications are described below.
+
 % 

### 2007-06-22 paper/trunk/JSS/response_A.tex
improved formatting of titles of the responses to the reviewers; corrected title of response to reviewer B ("reviewer A" was written in the title)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@746 591db2bd-e0a0-4d22-8c19-2192dbb7da31
the title of the responses to the reviewers shows the modified title of the paper now -- with the old title in a footnote
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@747 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 \begin{center}
-{\Large\textbf{
-systemfit: A Package to Estimate
-Simultaneous Equation Systems in R (JSS~224)}}
+{\Large\textbf{\textsf{
+systemfit: A Package for Estimating\\[0.3em]
+Systems of Simultaneous Equations in \proglang{R} (JSS~224)%
+\footnote{
+The original title was: °systemfit: A Package to Estimate
+Simultaneous Equation Systems in \proglang{R}°.}
+}}}
 
@@
 
-{\Large\textbf{Responses to Reviewer A}}
+{\Large\textbf{\textsf{Responses to Reviewer A}}}
 \end{center}

### 2007-06-22 paper/trunk/JSS/response_B.tex
improved formatting of titles of the responses to the reviewers; corrected title of response to reviewer B ("reviewer A" was written in the title)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@746 591db2bd-e0a0-4d22-8c19-2192dbb7da31
the title of the responses to the reviewers shows the modified title of the paper now -- with the old title in a footnote
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@747 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 
-{\Large\textbf{Responses to Reviewer A}}
+{\Large\textbf{\textsf{Responses to Reviewer B}}}
 \end{center}

### 2007-06-22 paper/trunk/JSS/response_A.tex
response to reviewer A: added answer that we have considerably modified the tools for testing linear hypothesis; corrected a small typo
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@748 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 by exploiting the common properties of the different estimators.
-However, in our opinion this shorting would considerably worsen
+However, in our opinion this shortening would considerably worsen
 the readability of this section.
@@
 %     classes for this.
+We have modified the tools for testing linear hypothesis
+according to the suggestions of this reviewer.
+The user can test linear restrictions by a Wald/$\chi^2$ or F test
+using the generic function \code{linear.hypothesis} now.
+The method \code{linear.hypothesis.systemfit} uses
+\code{linear.}\hspace{0pt}\code{hypothesis.}\hspace{0pt}\code{default}
+of the \pkg{car} package
+to compute the Wald/$\chi^2$ statistic.
+The $F$ statistic is computed in the \pkg{systemfit} package internally,
+but the returned object is of the same class and structure
+as the result of \code{linear.hypothesis.default}.
+
+The method \code{lrtest.systemfit} is a wrapper function
+to \code{lrtest.default} of the \pkg{lmtest} package now.
+Its main task is to specify the names for the compared objects.
+
 %

### 2007-06-22 trunk/R/smallMethods.R
in case of an analysis with panel-like data, the column names of the matrix returned by model.matrix.systemfit() are "<eqn>_<var>" rather than "<eqn>_<eqn>_<var>" now
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@398 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@ -148,8 +148,15 @@ model.matrix.systemfit <- function( object, ... ){
          cbind( matrix( 0, nrow( mmi ), ncol( result ) ), mmi ) )
       mmRowNames <- c( mmRowNames,
          paste( object$eq[[ i ]]$eqnLabel, "_", rownames( mmi ), sep = "" ) )
-      mmColNames <- c( mmColNames,
-         paste( object$eq[[ i ]]$eqnLabel, "_", colnames( mmi ), sep = "" ) )
+      for( j in 1:ncol( mmi ) ){
+         cName <- colnames( mmi )[ j ]
+         if( object$panelLike && cName != "(Intercept)" ){
+            mmColNames <- c( mmColNames, cName )
+         } else {
+            mmColNames <- c( mmColNames,
+               paste( object$eq[[ i ]]$eqnLabel, "_", cName, sep = "" ) )
+         }
+      }
    }
    rownames( result ) <- mmRowNames
    colnames( result ) <- mmColNames

### 2007-06-27 paper/trunk/JSS/response_A.tex
response to referee A: added answer that technical control parameters of systemfit() are collected in a single argument "control" now
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@751 591db2bd-e0a0-4d22-8c19-2192dbb7da31
response to referee A: added answer that the names of several arguments have been improved
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@754 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 %     arguments could be passed to systemfit.control() assuring backward compatibility
-%     and facilitating interactive usage. The names of various
+%     and facilitating interactive usage.
+Following the suggestion of this referee,
+the technical control parameters of \code{systemfit()} are collected
+in a single argument \code{control} now.
+This argument is a list that contains the technical control parameters,
+where suitable default values are set by the function
+\code{systemfit.control()}.
+
+%     The names of various
 %     arguments should be re-thought, e.g. "tx", "R.restr" and "q.rests" seem rather
 %     arbitrary. Similar comments hold for other arguments, see also below for some
-%     more comments. As noted above, the standard in R seems to supply the formula(s)
+%     more comments.
+The names of several arguments have been improved.
+The arguments to impose parameter restrictions by $R \beta = q$
+(former \code{R.restr} and \code{q.restr})
+have been renamed to \code{restrict.matrix} and \code{restrict.rhs},
+respectively
+(following the arguments \code{hypothesis.matrix} and \code{rhs}
+of the generic function \code{linear.hypothesis}).
+The argument to impose parameter restrictions
+by modifying (post-multiplying) the regressor matrix
+(fromer \code{TX}) has been renamed to \code{restrict.regMat}.
+Finally, the arguments to specify the methods
+for the 3SLS estimation and to calculated the residual covariance matrix
+(former \code{formula3sls} and \code{rcovformula})
+have been renamed to \code{method3sls} and \code{methodRCov},
+respectively.
+
+%     As noted above, the standard in R seems to supply the formula(s)
 %     as the first argument and to have then some `method' argument.

### 2007-06-27 paper/trunk/JSS/response_A.tex
response to referee A: added answer that we have moved the computation of the partial t statistic from the systemfit() function to its summary() method
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@755 591db2bd-e0a0-4d22-8c19-2192dbb7da31
response to referee A: added answer that the call of systemfit is stored now and that this allows the application of update() to systemfit objects
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@756 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 %     be specified for model fitting.
+Furthermore, we have followed the advice of this referee
+and moved the computation of the partial $t$ statistic
+from the \code{systemfit()} function to its \code{summary()} method.
+
 % 
@@
 %       fitsurRmat <- update(fm, R.restr = Rmat, q.restr = qvec)
-%     The call can easily be captured via match.call(). Similarly, one could think 
+%     The call can easily be captured via match.call().
+The call of \code{systemfit} is stored in the returned object now.
+This allows the application of \code{update()} to \code{systemfit} objects.
+
+%     Similarly, one could think 
 %     about how the terms and formula can be preserved and possibly made available

### 2007-06-27 paper/trunk/JSS/response_A.tex
response to referee A: added answer that we have added several methods for generic functions (terms, formulas, logLik, df.residual)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@757 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 %     in terms() and formula() calls. Other standard methods would also be nice to have,
-%     e.g., df.residual() or logLik(). (By the way: The row/column naming in fitted()
+%     e.g., df.residual() or logLik().
+Following the suggestions of this referee,
+the °model terms° (including the formulas) of each equation are preserved now.
+We have added methods \code{terms.systemfit}, \code{terms.systemfit.equation},
+\code{formula.systemfit}, and \code{formula.systemfit.equation}
+to extract the °model terms° and formulas
+of the system and of a single equation.
+The new method \code{logLik} computes the log-likelihood value
+of a fitted \code{systemfit} object.
+Furthermore, the default method of \code{df.residual}
+can extract the degrees of freedom of the residuals,
+both of the system and of a single equation now.
+
+%     (By the way: The row/column naming in fitted()
 %     and residuals() should match, probably re-using equation and observation names.)

### 2007-06-27 trunk/R/smallMethods.R
fitted.systemfit() returns a data.frame rather than a matrix now
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@418 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@ -128,7 +128,7 @@ fitted.systemfit <- function( object, ... ) {
       fitted.values[ , i ]           <- object$eq[[ i ]]$fitted.values
       colnames( fitted.values )[ i ] <- object$eq[[ i ]]$eqnLabel
    }
-   fitted.values
+   return( as.data.frame( fitted.values ) )
 }
 
 ## return the fitted values of e single euation

### 2007-06-27 paper/trunk/JSS/response_A.tex
response to referee A: corrected a small typo
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@758 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 by modifying (post-multiplying) the regressor matrix
-(fromer \code{TX}) has been renamed to \code{restrict.regMat}.
+(former \code{TX}) has been renamed to \code{restrict.regMat}.
 Finally, the arguments to specify the methods

### 2007-06-28 trunk/R/predict.systemfit.R
renamed elements of systemfit and systemfit.equation objects that store the estimated covariance matrix of the coefficients: bcov -> coefCov
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@425 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@ -50,10 +50,10 @@ predict.systemfit.equation <- function( object, newdata = NULL,
 
    # calculate variance covariance matrices
    if( se.fit | interval == "confidence" ) {
-      yCovConf <- drop( xMat %*% object$bcov %*% t( xMat ) )
+      yCovConf <- drop( xMat %*% object$coefCov %*% t( xMat ) )
    }
    if( se.pred | interval == "prediction" ) {
-      yCovPred <- drop( xMat %*% object$bcov %*% t( xMat ) + object$sigma^2 )
+      yCovPred <- drop( xMat %*% object$coefCov %*% t( xMat ) + object$sigma^2 )
    }
    # standard errors of fitted values
    if( se.fit ) {

### 2007-06-28 trunk/R/smallMethods.R

@@ -6,11 +6,11 @@ correlation.systemfit <- function( results, eqni, eqnj ) {
   for( i in 1:length( results$eq ) ) {
      nCoefEq <- c( nCoefEq, length( coef( results$eq[[ i ]] ) ) )
   }
-  cij <- results$bcov[(1+sum(nCoefEq[1:eqni])-nCoefEq[eqni]):(sum(nCoefEq[1:eqni])),
+  cij <- results$coefCov[(1+sum(nCoefEq[1:eqni])-nCoefEq[eqni]):(sum(nCoefEq[1:eqni])),
                       (1+sum(nCoefEq[1:eqnj])-nCoefEq[eqnj]):(sum(nCoefEq[1:eqnj]))]
-  cii <- results$bcov[(1+sum(nCoefEq[1:eqni])-nCoefEq[eqni]):(sum(nCoefEq[1:eqni])),
+  cii <- results$coefCov[(1+sum(nCoefEq[1:eqni])-nCoefEq[eqni]):(sum(nCoefEq[1:eqni])),
                       (1+sum(nCoefEq[1:eqni])-nCoefEq[eqni]):(sum(nCoefEq[1:eqni]))]
-  cjj <- results$bcov[(1+sum(nCoefEq[1:eqnj])-nCoefEq[eqnj]):(sum(nCoefEq[1:eqnj])),
+  cjj <- results$coefCov[(1+sum(nCoefEq[1:eqnj])-nCoefEq[eqnj]):(sum(nCoefEq[1:eqnj])),
                       (1+sum(nCoefEq[1:eqnj])-nCoefEq[eqnj]):(sum(nCoefEq[1:eqnj]))]
   rij <- NULL
 
@@ -32,8 +32,8 @@ se.ratio.systemfit <- function( resultsi, resultsj, eqni ) {
   ratio <- NULL
   for( i in 1:nrow( residuals( resultsi ) ) ) {
     xik    <- resultsi$eq[[eqni]]$modelMatrix[i,]
-    top    <- sqrt( xik %*% resultsi$eq[[eqni]]$bcov %*% xik )
-    bottom <- sqrt( xik %*% resultsj$eq[[eqni]]$bcov %*% xik )
+    top    <- sqrt( xik %*% resultsi$eq[[eqni]]$coefCov %*% xik )
+    bottom <- sqrt( xik %*% resultsj$eq[[eqni]]$coefCov %*% xik )
     rk     <- top / bottom
     ratio  <- rbind( ratio, rk )
   }
[…]

### 2007-06-28 trunk/R/summary.systemfit.R

@@ -36,7 +36,7 @@ summary.systemfit <- function( object, useDfSys = NULL, ... ) {
    }
 
    # coefficients, standard errors, ... 
-   result$coefCov <- object$bcov
+   result$coefCov <- object$coefCov
    coef <- object$coefficients
    stdEr <- diag( result$coefCov )^0.5  # standard errors
    tStat <- coef / stdEr                # t-statistic
@@ -214,7 +214,7 @@ summary.systemfit.equation <- function( object, useDfSys = NULL, ... ) {
    result$residuals <- object$residuals
 
    # coefficients, standard errors, ... 
-   result$coefCov <- object$bcov
+   result$coefCov <- object$coefCov
    coef <- object$coefficients
    stdEr <- diag( result$coefCov )^0.5  # standard errors
    tStat <- coef / stdEr                # t-statistic

### 2007-06-28 trunk/R/systemfit.R

@@ -604,7 +604,7 @@ systemfit <- function(  eqns,
     resulti$df.residual.sys  <- nObsAll- nCoefLiAll
        # degrees of freedom of residuals of the whole system
     resulti$coefficients <- coefEqI         # estimated coefficients
-    resulti$bcov         <- bcovi           # covariance matrix of estimated coefficients
+    resulti$coefCov      <- bcovi           # covariance matrix of estimated coefficients
     if( control$returnResponse ){
       resulti$response     <- yVecEq[[i]]     # vector of endogenous variables
     }
@@ -649,7 +649,7 @@ systemfit <- function(  eqns,
   results$df.residual <- nObsAll - nCoefLiAll
      # degrees of freedom of the whole system
   results$coefficients <- coef           # all estimated coefficients
-  results$bcov    <- bcov           # coefficients covariance matrix
+  results$coefCov  <- bcov           # coefficients covariance matrix
   results$residCov <- rcov           # residual covarance matrix
   results$iter    <- iter           # residual correlation matrix
   if( method %in% c( "SUR", "3SLS" ) ){

### 2007-06-28 paper/trunk/systemfit_compare_lm.tex
aligned latex code of table that compares the objects returned by lm and systemfit
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@762 591db2bd-e0a0-4d22-8c19-2192dbb7da31
table that compares the objects returned by lm and systemfit: updated elements of systemfit objects due to changes of the systemfit package
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@763 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 model*          &                & modelFrame* \\
- & method & \\
+                & method         & method \\
                 &                & eqnLabel \\
+                &                & eqnNo \\
                 & iter           & \\
- & bt & \\
- & bcov & covb \\
- & btcov & \\
- & rcov & \\
- & rcovest & \\
- & R.restr & \\
- & q.restr & \\
- & TX & \\
+                & coefCov        & coefCov \\
+                & residCov       & \\
+                & residCovEst    & \\
+                & restrict.matrix& \\
+                & restrict.rhs   & \\
+                & restrict.regMat& \\
                 & control        & \\
                 & eq             & \\
+                & panelLike      & \\
                 &                & inst \\

### 2007-06-29 paper/trunk/systemfit_compare_lm.tex
table that compares the objects returned by lm and systemfit: updated elements of systemfit objects due to changes of the systemfit package
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@764 591db2bd-e0a0-4d22-8c19-2192dbb7da31

--- 2006-10-01 paper/trunk/systemfit_compare_lm.tex
@@
 terms           &                & terms \\
+                &                & inst \\
 contrasts       &                & \\
@@
 model*          &                & modelFrame* \\
- & method & \\
+                & method         & method \\
                 &                & eqnLabel \\
+                &                & eqnNo \\
                 & iter           & \\
- & bt & \\
- & bcov & covb \\
- & btcov & \\
- & rcov & \\
- & rcovest & \\
- & R.restr & \\
- & q.restr & \\
- & TX & \\
+                & coefCov        & coefCov \\
+                & residCov       & \\
+                & residCovEst    & \\
+                & restrict.matrix& \\
+                & restrict.rhs   & \\
+                & restrict.regMat& \\
                 & control        & \\
                 & eq             & \\
- & & inst \\
- & & ssr \\
- & & sigma \\
+                & panelLike      & \\
 \hline

### 2007-06-30 paper/trunk/systemfit_compare_lm.tex
revised and extended appendix section (and table) that compares the objects returned by systemfit and lm
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@766 591db2bd-e0a0-4d22-8c19-2192dbb7da31

--- 2006-10-01 paper/trunk/systemfit_compare_lm.tex
@@
-\section{Comparison of systemfit and lm}
+\section[Object returned by systemfit]{Object returned by \code{systemfit}}
+
+\code{systemfit} returns a list of class \code{systemfit}
+that contains all results that belong to the entire system
+of equations.
+One special element of this list called \code{eq}.
+It is a list that contains one object for each estimated equation.
+These objects are of the class \code{systemfit.equation}
+and contain the results that belong only to the regarding equation.
+
+The elements returned by \code{systemfit}
+are similar to those returned by \code{lm},
+the basic tool for linear regressions in \proglang{R}.
+While some counterparts of elements returned by \code{lm}
+can be found directly in objects of class \code{systemfit},
+other counterparts are available for each equation
+in objects of class \code{systemfit.equation}.
+This is demonstrated in table~\ref{tab:compare-lm}
 
 \begin{table}[htbp]
-\caption{Values returned by systemfit and lm}
+\caption{Elements returned by \code{systemfit} and \code{lm}}
+\label{tab:compare-lm}
+\centering
+\setlength{\tabcolsep}{4mm}
 \begin{tabular}{lll}
-\hline
-lm & systemfit & systemfit.equation \\
+\hline  \hline
+\textbf{\code{lm}} & \textbf{\code{systemfit}} & \textbf{\code{systemfit.equation}} \\
 \hline
 coefficients    & coefficients   & coefficients \\
+                & coefCov        & coefCov \\
+fitted.values   &                & fitted.values \\
 residuals       &                & residuals \\
-fitted.values & & fitted.values \\
+                & residCov       & \\
+                & residCovEst    & \\
 rank            & rank           & rank \\
@@
                 &                & nCoef.sys \\
-weights & & \\
 df.residual     & df.residual    & df.residual \\
@@
 terms           &                & terms \\
+                &                & inst \\
+weights         &                & \\
 contrasts       &                & \\
@@
 offset          &                & \\
-y** & & response* \\
-x** & & modelMatrix* \\
- & & instMatrix* \\
 model*          &                & modelFrame* \\
- & method & \\
+y**             &                & response** \\
+x**             &                & modelMatrix** \\
+                &                & instMatrix** \\
+                & iter           & \\
+                & eq             & \\
                 &                & eqnLabel \\
- & iter & \\
- & bt & \\
- & bcov & covb \\
- & btcov & \\
- & rcov & \\
- & rcovest & \\
- & R.restr & \\
- & q.restr & \\
- & TX & \\
+                &                & eqnNo \\
+                & method         & method \\
+                & panelLike      & \\
+                & restrict.matrix& \\
+                & restrict.rhs   & \\
+                & restrict.regMat& \\
                 & control        & \\
- & eq & \\
- & & inst \\
- & & ssr \\
- & & sigma \\
-\hline
+\hline  \hline
 \end{tabular}
+\medskip\\
+Note:
+*) if requested by the user (default is \code{TRUE})\\
+**) if requested by the user (default is \code{FALSE})
 \end{table}

### 2007-06-30 paper/trunk/systemfit_code.tex
modifications, because lrtest.systemfit is now a wrapper function to lrtest.default
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@742 591db2bd-e0a0-4d22-8c19-2192dbb7da31
corrected a small typo
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@749 591db2bd-e0a0-4d22-8c19-2192dbb7da31
revised section about source code of systemfit (mainly changes due to modifications of the systemfit package)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@767 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 by several different estimation methods.
-Furthermore, this package provides some methods
+Furthermore, this package provides methods
 to test hypotheses
-(\code{ftest}, \code{wald\-test},
-\code{lrtest}, \code{hausman}),
-to extract the estimated coefficients (\code{coef})
+(\code{linear.hypothesis},
+\code{lrtest}, \code{hausman}\hspace{0pt}\code{.systemfit}),
+to calculate summary results (\code{summary}),
+to print (summary) results (\code{print})
 and to calculate predicted values (\code{predict}).
+Moreover, there are several methods to extract
+the estimated coefficients (\code{coef}),
+their covariance matrix (\code{vcov}),
+the fitted values (\code{fitted}),
+the residuals (\code{residuals}),
+the formula(s) (\code{formula}),
+the model terms (\code{terms}),
+the model frame (\code{model.frame}),
+the model matrix (\code{model.matrix}),
+and the log likelihood value (\code{logLik}).
+There does not exist a special method to extract
+the degrees of freedom of the residuals from \code{systemfit} objects,
+because the default method of \code{df.residual}
+works for these objects.
 
@@
 
-The \code{systemfit} function returns many objects
-that users might be interested in.
-A complete list is available in the documentation of this function
+The \code{systemfit} function returns an object of class \code{systemfit}
+that contains many elements users might be interested in.
+A complete description is available in the documentation of this function
 that is included in the package.
+A comparison with the elements returned by \code{lm},
+the basic tool for linear regressions in \proglang{R},
+is available in appendix~\ref{sec:returned-object}.
 
@@
 and the Wald/$\chi^2$ statistic is computed by the method
-\code{linear.hypothesis.default} of the \pkg{car} package.
-The function \code{lrtest.systemfit}
-computes the Likelihood-ratio (LR) test statistic.
-On the other hand, the function \code{hausman.systemfit}
+\code{linear.hypothesis.default} of the \pkg{car} package
+\citep{r-car-1.2-1}.
+The method \code{lrtest.systemfit} is a wrapper function
+to \code{lrtest.default} of the \pkg{lmtest} package
+\citep{r-lmtest},
+which computes the Likelihood-ratio (LR) test statistic.
+All these functions return an object of class \code{anova}
+that contains --- amongst others ---
+the empirical test statistic, the degrees of freedom,
+and the corresponding $p$ value.
+
+The function \code{hausman.systemfit}
 tests the consistency of the 3SLS estimator.
-All functions return the empirical test statistic,
-the degree(s) of freedom, and the $p$ value.
+It returns an object of class \code{htest},
+which contains --- amongst others ---
+the empirical test statistic,
+the degrees of freedom, and the $p$ value.
 

### 2007-06-30 paper/trunk/JSS/response_A.tex
response to referee A: added answer that the data frames returned by residuals.systemfit() and fitted.systemfit() have the same row names and column names now
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@759 591db2bd-e0a0-4d22-8c19-2192dbb7da31
response to referee A: fragmented long "\code{}" part to avoid an overfull or underfull vbox
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@760 591db2bd-e0a0-4d22-8c19-2192dbb7da31
response to referee A: revised (completed) answer that we have renamed many elements of a fitted systemfit object to make the naming more stan

@@
 %     objects etc. would be more desirable.
-We have renamed single letter elements (not finished yet).
+We have renamed many elements of a fitted \code{systemfit} object
+to make the naming more standard.
+A comparison with the elements returned by \code{lm}
+is available in the paper (appendix~A).
 
@@
 the °model terms° (including the formulas) of each equation are preserved now.
-We have added methods \code{terms.systemfit}, \code{terms.systemfit.equation},
+We have added methods \code{terms.systemfit},
+\code{terms}\hspace{0pt}\code{.systemfit}\hspace{0pt}\code{.equation},
 \code{formula.systemfit}, and \code{formula.systemfit.equation}
@@
 %     and residuals() should match, probably re-using equation and observation names.)
+The data frames returned by \code{residuals.systemfit} and
+\code{fitted.systemfit} have the same row names and column names now.
+
 % 

### 2007-06-30 trunk/R/hausman.R
renamed argument "modified.reg" of coef.systemfit(), coef.summary.systemfit() and vcov.systemfit() to "modified.regMat" to make this name more consistent with the name of argument "restrict.regMat" of systemfit()
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@442 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@ -12,10 +12,10 @@ hausman.systemfit <- function( results2sls, results3sls ) {
       result$q <- coef( results2sls ) - coef( results3sls )
       result$qVar <- vcov( results2sls ) - vcov( results3sls )
    } else {
-      result$q <- coef( results2sls, modified.reg = TRUE ) -
-         coef( results3sls, modified.reg = TRUE )
-      result$qVar <- vcov( results2sls, modified.reg = TRUE ) -
-         vcov( results3sls, modified.reg = TRUE )
+      result$q <- coef( results2sls, modified.regMat = TRUE ) -
+         coef( results3sls, modified.regMat = TRUE )
+      result$qVar <- vcov( results2sls, modified.regMat = TRUE ) -
+         vcov( results3sls, modified.regMat = TRUE )
    }
 
 #    if( min( eigen( hausman$qVar )$values ) < 0 ) {

### 2007-06-30 trunk/R/smallMethods.R

@@ -42,8 +42,8 @@ se.ratio.systemfit <- function( resultsi, resultsj, eqni ) {
 
 
 ## return all coefficients
-coef.systemfit <- function( object, modified.reg = FALSE, ... ) {
-   if( modified.reg ){
+coef.systemfit <- function( object, modified.regMat = FALSE, ... ) {
+   if( modified.regMat ){
       if( is.null( object$restrict.regMat ) ){
          stop( "coefficients of the modified regressor matrix are not available,",
             " because argument 'restrict.regMat' has not been used in this estimation." )
@@ -57,8 +57,8 @@ coef.systemfit <- function( object, modified.reg = FALSE, ... ) {
 }
 
 ## return all coefficients, std.errors, t-values and p-values
-coef.summary.systemfit <- function( object, modified.reg = FALSE, ... ) {
-   if( modified.reg ){
+coef.summary.systemfit <- function( object, modified.regMat = FALSE, ... ) {
+   if( modified.regMat ){
       if( is.null( object$coefModReg ) ){
          stop( "coefficients of the modified regressor matrix are not available,",
             " because argument 'restrict.regMat' has not been used in this estimation." )
@@ -96,8 +96,8 @@ residuals.systemfit.equation <- function( object, ... ) {
 }
 
 ## return the variance covariance matrix of the coefficients
-vcov.systemfit <- function( object, modified.reg = FALSE, ... ) {
-   if( modified.reg ){
+vcov.systemfit <- function( object, modified.regMat = FALSE, ... ) {
+   if( modified.regMat ){
       if( is.null( object$restrict.regMat ) ){
          stop( "coefficients of the modified regressor matrix",
             " and their covariance matrix are not available,",

### 2007-06-30 trunk/R/summary.systemfit.R

@@ -55,8 +55,8 @@ summary.systemfit <- function( object, useDfSys = NULL, ... ) {
 
    # coefficients of the modified regressor matrix
    if( !is.null( object$restrict.regMat ) ) {
-      coefModReg <- coef( object, modified.reg = TRUE )
-      stdErModReg <- diag( vcov( object, modified.reg = TRUE ) )^0.5  # standard errors
+      coefModReg <- coef( object, modified.regMat = TRUE )
+      stdErModReg <- diag( vcov( object, modified.regMat = TRUE ) )^0.5  # standard errors
       tStatModReg <- coefModReg / stdErModReg    # t-statistic
       if( useDfSys ) {             # p-values
          pValModReg <- 2 * ( 1 - pt( abs( tStatModReg ), object$df.residual ) )

### 2007-06-30 trunk/tests/test_2sls.R

@@ -261,14 +261,14 @@ print( round( coef( fit2sls2p ), digits = 6 ) )
 print( round( coef( fit2sls2p$eq[[ 2 ]] ), digits = 6 ) )
 
 print( round( coef( fit2sls3 ), digits = 6 ) )
-print( round( coef( fit2sls3, modified.reg = TRUE ), digits = 6 ) )
+print( round( coef( fit2sls3, modified.regMat = TRUE ), digits = 6 ) )
 print( round( coef( fit2sls3$eq[[ 1 ]] ), digits = 6 ) )
 
 print( round( coef( fit2sls4s ), digits = 6 ) )
 print( round( coef( fit2sls4s$eq[[ 2 ]] ), digits = 6 ) )
 
 print( round( coef( fit2sls5r ), digits = 6 ) )
-print( round( coef( fit2sls5r, modified.reg = TRUE ), digits = 6 ) )
+print( round( coef( fit2sls5r, modified.regMat = TRUE ), digits = 6 ) )
 print( round( coef( fit2sls5r$eq[[ 2 ]] ), digits = 6 ) )
 
 
@@ -281,7 +281,7 @@ print( round( coef( summary( fit2sls2p$eq[[ 2 ]], useDfSys = FALSE ) ),
    digits = 6 ) )
 
 print( round( coef( summary( fit2sls3 ) ), digits = 6 ) )
-print( round( coef( summary( fit2sls3 ), modified.reg = TRUE ), digits = 6 ) )
+print( round( coef( summary( fit2sls3 ), modified.regMat = TRUE ), digits = 6 ) )
 print( round( coef( summary( fit2sls3$eq[[ 1 ]] ) ), digits = 6 ) )
 
 print( round( coef( summary( fit2sls4s ) ), digits = 6 ) )
@@ -289,7 +289,7 @@ print( round( coef( summary( fit2sls4s$eq[[ 2 ]] ) ), digits = 6 ) )
 
 print( round( coef( summary( fit2sls5r, useDfSys = FALSE ) ), digits = 6 ) )
 print( round( coef( summary( fit2sls5r, useDfSys = FALSE ),
-   modified.reg = TRUE ), digits = 6 ) )
+   modified.regMat = TRUE ), digits = 6 ) )
 print( round( coef( summary( fit2sls5r$eq[[ 2 ]], useDfSys = FALSE ) ),
    digits = 6 ) )
 
[…]

### 2007-06-30 trunk/tests/test_3sls.R

@@ -419,11 +419,11 @@ print( round( coef( fit3sls[[ 2 ]]$e2w ), digits = 6 ) )
 print( round( coef( fit3sls[[ 3 ]]$e2w$eq[[ 1 ]] ), digits = 6 ) )
 
 print( round( coef( fit3slsd[[ 5 ]]$e3e ), digits = 6 ) )
-print( round( coef( fit3slsd[[ 5 ]]$e3e, modified.reg = TRUE ), digits = 6 ) )
+print( round( coef( fit3slsd[[ 5 ]]$e3e, modified.regMat = TRUE ), digits = 6 ) )
 print( round( coef( fit3slsd[[ 1 ]]$e3e$eq[[ 2 ]] ), digits = 6 ) )
 
 print( round( coef( fit3slsd[[ 4 ]]$e3w ), digits = 6 ) )
-print( round( coef( fit3slsd[[ 4 ]]$e3w, modified.reg = TRUE ), digits = 6 ) )
+print( round( coef( fit3slsd[[ 4 ]]$e3w, modified.regMat = TRUE ), digits = 6 ) )
 print( round( coef( fit3slsd[[ 5 ]]$e3w$eq[[ 2 ]] ), digits = 6 ) )
 
 print( round( coef( fit3sls[[ 1 ]]$e4 ), digits = 6 ) )
@@ -433,7 +433,7 @@ print( round( coef( fit3slsi[[ 2 ]]$e4we ), digits = 6 ) )
 print( round( coef( fit3slsi[[ 1 ]]$e4we$eq[[ 1 ]] ), digits = 6 ) )
 
 print( round( coef( fit3slsi[[ 2 ]]$e5e ), digits = 6 ) )
-print( round( coef( fit3slsi[[ 2 ]]$e5e, modified.reg = TRUE ), digits = 6 ) )
+print( round( coef( fit3slsi[[ 2 ]]$e5e, modified.regMat = TRUE ), digits = 6 ) )
 print( round( coef( fit3slsi[[ 3 ]]$e5e$eq[[ 2 ]] ), digits = 6 ) )
 
 
@@ -454,14 +454,14 @@ print( round( coef( summary( fit3slsi[[ 5 ]]$e2$eq[[ 1 ]] ) ), digits = 6 ) )
 print( round( coef( summary( fit3slsd[[ 5 ]]$e3e, useDfSys = FALSE ) ),
    digits = 6 ) )
[…]

### 2007-06-30 trunk/tests/test_ols.R

@@ -220,14 +220,14 @@ print( round( coef( fitols2s ), digits = 6 ) )
 print( round( coef( fitols2s$eq[[ 1 ]] ), digits = 6 ) )
 
 print( round( coef( fitols3p ), digits = 6 ) )
-print( round( coef( fitols3p, modified.reg = TRUE ), digits = 6 ) )
+print( round( coef( fitols3p, modified.regMat = TRUE ), digits = 6 ) )
 print( round( coef( fitols3p$eq[[ 2 ]] ), digits = 6 ) )
 
 print( round( coef( fitols4r ), digits = 6 ) )
 print( round( coef( fitols4r$eq[[ 1 ]] ), digits = 6 ) )
 
 print( round( coef( fitols5 ), digits = 6 ) )
-print( round( coef( fitols5, modified.reg = TRUE ), digits = 6 ) )
+print( round( coef( fitols5, modified.regMat = TRUE ), digits = 6 ) )
 print( round( coef( fitols5$eq[[ 2 ]] ), digits = 6 ) )
 
 
@@ -240,7 +240,7 @@ print( round( coef( summary( fitols2s ) ), digits = 6 ) )
 print( round( coef( summary( fitols2s$eq[[ 1 ]] ) ), digits = 6 ) )
 
 print( round( coef( summary( fitols3p, useDfSys = FALSE ) ), digits = 6 ) )
-print( round( coef( summary( fitols3p, useDfSys = FALSE ), modified.reg = TRUE ),
+print( round( coef( summary( fitols3p, useDfSys = FALSE ), modified.regMat = TRUE ),
    digits = 6 ) )
 print( round( coef( summary( fitols3p$eq[[ 2 ]], useDfSys = FALSE ) ),
    digits = 6 ) )
@@ -250,7 +250,7 @@ print( round( coef( summary( fitols4r$eq[[ 1 ]], useDfSys = FALSE ) ),
    digits = 6 ) )
 
 print( round( coef( summary( fitols5 ) ), digits = 6 ) )
-print( round( coef( summary( fitols5 ), modified.reg = TRUE ), digits = 6 ) )
+print( round( coef( summary( fitols5 ), modified.regMat = TRUE ), digits = 6 ) )
 print( round( coef( summary( fitols5$eq[[ 2 ]] ) ), digits = 6 ) )
 
 
[…]

### 2007-06-30 trunk/tests/test_panel.R

@@ -77,11 +77,11 @@ print( greeneOlsPooled )
 summary( greeneOlsPooled )
 sum( sapply( greeneOlsPooled$eq, function(x){return(summary(x)$ssr)}) )/97 # sigma^2
 coef( greeneOlsPooled )
-coef( greeneOlsPooled, modified.reg = TRUE )
+coef( greeneOlsPooled, modified.regMat = TRUE )
 coef( summary( greeneOlsPooled ) )
-coef( summary( greeneOlsPooled ), modified.reg = TRUE )
+coef( summary( greeneOlsPooled ), modified.regMat = TRUE )
 vcov( greeneOlsPooled )
-vcov( greeneOlsPooled, modified.reg = TRUE )
+vcov( greeneOlsPooled, modified.regMat = TRUE )
 residuals( greeneOlsPooled )
 confint( greeneOlsPooled )
 fitted( greeneOlsPooled )
@@ -119,11 +119,11 @@ greeneSurPooled <- systemfit( formulaGrunfeld, "SUR",
 print( greeneSurPooled )
 summary( greeneSurPooled )
 coef( greeneSurPooled )
-coef( greeneSurPooled, modified.reg = TRUE )
+coef( greeneSurPooled, modified.regMat = TRUE )
 coef( summary( greeneSurPooled ) )
-coef( summary( greeneSurPooled ), modified.reg = TRUE )
+coef( summary( greeneSurPooled ), modified.regMat = TRUE )
 vcov( greeneSurPooled )
-vcov( greeneSurPooled, modified.reg = TRUE )
+vcov( greeneSurPooled, modified.regMat = TRUE )
 residuals( greeneSurPooled )
 confint( greeneSurPooled )
 fitted( greeneSurPooled )

### 2007-06-30 trunk/tests/test_sur.R

@@ -341,18 +341,18 @@ print( round( coef( fitsur2we ), digits = 6 ) )
 print( round( coef( fitsur2we$eq[[ 1 ]] ), digits = 6 ) )
 
 print( round( coef( fitsur3 ), digits = 6 ) )
-print( round( coef( fitsur3, modified.reg = TRUE ), digits = 6 ) )
+print( round( coef( fitsur3, modified.regMat = TRUE ), digits = 6 ) )
 print( round( coef( fitsur3$eq[[ 2 ]] ), digits = 6 ) )
 
 print( round( coef( fitsur4r2 ), digits = 6 ) )
 print( round( coef( fitsur4r2$eq[[ 1 ]] ), digits = 6 ) )
 
 print( round( coef( fitsuri5e ), digits = 6 ) )
-print( round( coef( fitsuri5e, modified.reg = TRUE ), digits = 6 ) )
+print( round( coef( fitsuri5e, modified.regMat = TRUE ), digits = 6 ) )
 print( round( coef( fitsuri5e$eq[[ 2 ]] ), digits = 6 ) )
 
 print( round( coef( fitsur5w ), digits = 6 ) )
-print( round( coef( fitsur5w, modified.reg = TRUE ), digits = 6 ) )
+print( round( coef( fitsur5w, modified.regMat = TRUE ), digits = 6 ) )
 print( round( coef( fitsur5w$eq[[ 1 ]] ), digits = 6 ) )
 
 
@@ -365,11 +365,11 @@ print( round( coef( summary( fitsuri2$eq[[ 1 ]], useDfSys = FALSE ) ),
    digits = 6 ) )
 
 print( round( coef( summary( fitsur3 ) ), digits = 6 ) )
-print( round( coef( summary( fitsur3 ), modified.reg = TRUE ), digits = 6 ) )
+print( round( coef( summary( fitsur3 ), modified.regMat = TRUE ), digits = 6 ) )
 print( round( coef( summary( fitsur3$eq[[ 2 ]] ) ), digits = 6 ) )
 
 print( round( coef( summary( fitsuri3we ) ), digits = 6 ) )
-print( round( coef( summary( fitsuri3we ), modified.reg = TRUE ), digits = 6 ) )
+print( round( coef( summary( fitsuri3we ), modified.regMat = TRUE ), digits = 6 ) )
[…]

### 2007-06-30 trunk/tests/test_w2sls.R

@@ -162,14 +162,14 @@ print( round( coef( fitw2slsd2e ), digits = 6 ) )
 print( round( coef( fitw2slsd2e$eq[[ 1 ]] ), digits = 6 ) )
 
 print( round( coef( fitw2slsd3e ), digits = 6 ) )
-print( round( coef( fitw2slsd3e, modified.reg = TRUE ), digits = 6 ) )
+print( round( coef( fitw2slsd3e, modified.regMat = TRUE ), digits = 6 ) )
 print( round( coef( fitw2slsd3e$eq[[ 2 ]] ), digits = 6 ) )
 
 print( round( coef( fitw2sls4 ), digits = 6 ) )
 print( round( coef( fitw2sls4$eq[[ 1 ]] ), digits = 6 ) )
 
 print( round( coef( fitw2sls5 ), digits = 6 ) )
-print( round( coef( fitw2sls5, modified.reg = TRUE ), digits = 6 ) )
+print( round( coef( fitw2sls5, modified.regMat = TRUE ), digits = 6 ) )
 print( round( coef( fitw2sls5$eq[[ 2 ]] ), digits = 6 ) )
 
 
@@ -183,7 +183,7 @@ print( round( coef( summary( fitw2slsd2e$eq[[ 1 ]] ) ), digits = 6 ) )
 
 print( round( coef( summary( fitw2slsd3e, useDfSys = FALSE ) ), digits = 6 ) )
 print( round( coef( summary( fitw2slsd3e, useDfSys = FALSE ),
-   modified.reg = TRUE ), digits = 6 ) )
+   modified.regMat = TRUE ), digits = 6 ) )
 print( round( coef( summary( fitw2slsd3e$eq[[ 2 ]], useDfSys = FALSE ) ),
    digits = 6 ) )
 
@@ -191,7 +191,7 @@ print( round( coef( summary( fitw2sls4 ) ), digits = 6 ) )
 print( round( coef( summary( fitw2sls4$eq[[ 1 ]] ) ), digits = 6 ) )
 
 print( round( coef( summary( fitw2sls5 ) ), digits = 6 ) )
-print( round( coef( summary( fitw2sls5 ), modified.reg = TRUE ), digits = 6 ) )
+print( round( coef( summary( fitw2sls5 ), modified.regMat = TRUE ), digits = 6 ) )
[…]

### 2007-06-30 trunk/tests/test_wls.R

@@ -189,14 +189,14 @@ print( round( coef( fitwlsi2 ), digits = 6 ) )
 print( round( coef( fitwlsi2$eq[[ 2 ]] ), digits = 6 ) )
 
 print( round( coef( fitwls3e ), digits = 6 ) )
-print( round( coef( fitwls3e, modified.reg = TRUE ), digits = 6 ) )
+print( round( coef( fitwls3e, modified.regMat = TRUE ), digits = 6 ) )
 print( round( coef( fitwls3e$eq[[ 1 ]] ), digits = 6 ) )
 
 print( round( coef( fitwls4 ), digits = 6 ) )
 print( round( coef( fitwls4$eq[[ 2 ]] ), digits = 6 ) )
 
 print( round( coef( fitwlsi5 ), digits = 6 ) )
-print( round( coef( fitwlsi5, modified.reg = TRUE ), digits = 6 ) )
+print( round( coef( fitwlsi5, modified.regMat = TRUE ), digits = 6 ) )
 print( round( coef( fitwlsi5$eq[[ 1 ]] ), digits = 6 ) )
 
 
@@ -208,7 +208,7 @@ print( round( coef( summary( fitwlsi2 ) ), digits = 6 ) )
 print( round( coef( summary( fitwlsi2$eq[[ 2 ]] ) ), digits = 6 ) )
 
 print( round( coef( summary( fitwls3e ) ), digits = 6 ) )
-print( round( coef( summary( fitwls3e ), modified.reg = TRUE ), digits = 6 ) )
+print( round( coef( summary( fitwls3e ), modified.regMat = TRUE ), digits = 6 ) )
 print( round( coef( summary( fitwls3e$eq[[ 1 ]] ) ), digits = 6 ) )
 
 print( round( coef( summary( fitwls4, useDfSys = FALSE ) ), digits = 6 ) )
@@ -217,7 +217,7 @@ print( round( coef( summary( fitwls4$eq[[ 2 ]], useDfSys = FALSE ) ),
 
 print( round( coef( summary( fitwlsi5, useDfSys = FALSE ) ), digits = 6 ) )
 print( round( coef( summary( fitwlsi5, useDfSys = FALSE ),
-   modified.reg = TRUE ), digits = 6 ) )
+   modified.regMat = TRUE ), digits = 6 ) )
[…]

### 2007-07-01 paper/trunk/JSS/response_A.tex
response to referee A: added answer that we have added arguments "printResidCov" and "printEquations" to summary.systemfit() and added arguments "residCov" and "equations" to print.summary.systemfit()
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@769 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 %       summary(sfobject, equation = ..., cor = ...)
+The user has more control about the amount of output of
+\code{summary()} now.
+We have added arguments \code{residCov} and \code{equations}
+to function \code{print.summary.systemfit()}.
+Argument \code{residCov} indicates
+whether the residual correlation matrix,
+the residual covariance matrix, and its determinant are printed;
+argument \code{equations} indicates
+whether summary results of each equation are printed
+(or if just the coefficients are printed).
+As users often use \code{summary()} without \code{print()},
+we have added arguments \code{printResidCov} and \code{printEquations}
+to \code{summary()},
+which set default values for arguments \code{residCov} and \code{equations}
+of function \code{print.summary.systemfit()}.
+
 % 

### 2007-07-01 paper/trunk/JSS/response_A.tex
response to referee A: added answer that we have streamlined the source code of the systemfit package, for instance by replacing all '\code{method == "OLS" | method == "WLS" | method == ...}' by '\code{method %in% c("OLS", "WLS", ...)}'
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@770 591db2bd-e0a0-4d22-8c19-2192dbb7da31
response to referee A: added answer that we have removed the data set "GrunfeldTheil", improved the discussion of different versions of Grunfeld's data set in the documentation of "GrunfeldGreene", added this discussion in the paper, and added a note that a co

@@
 %       method == "OLS" | method == "WLS" | method == ...
+We have streamlined the source code of the \pkg{systemfit} package.
+For instance, we have followed the suggestion of this referee
+and replaced all
+°\code{method == "OLS" | method == "WLS" | method == ...}°
+by °\code{method \%in\% c("OLS", "WLS", ...)}°.
+
 %     Furthermore, various marix computations can be made even more stable. Currently,
@@
 %     available in package "Ecdat".
+The data set \code{GrunfeldTheil} has been removed,
+because it is a subset of the data set \code{GrunfeldGreene}.
+We have improved the discussion of different versions of Grunfeld's data set
+in the documentation of \code{GrunfeldGreene}
+and we have added this discussion in the paper.
+Furthermore, we have added a note
+that a correct version of this data set with five additional firms
+is available as data set \code{Grunfeld} in the \pkg{Ecdat} package.
+
 % 

### 2007-07-02 paper/trunk/JSS/response_A.tex
response to referee A: added answer that we have removed the replicated textbook results and present the commands to replicate these results only
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@775 591db2bd-e0a0-4d22-8c19-2192dbb7da31
response to referee A: added answer that we have added the aims "teaching" and "reproducible research" to further motivate section "Testing reliability" and that we have renamed this section to "Replication of textbook results".
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@777 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 %     I would add replication to reliability) and summarize the most important results
-%     and experiences from the replication. Then, I would pick a `typical' example and
+%     and experiences from the replication.
+As suggested by this referee,
+we have added the aims °teaching° and °reproducible research°
+to further motivate this section.
+As this section no longer focusses exclusively on testing reliability,
+we have renamed it to °Replication of textbook results°.
+
+%     Then, I would pick a `typical' example and
 %     discuss it in some more detail with full R output. For the remaining examples,
@@
 %     from the published numbers.
+To make this section more readable, we have removed the estimation results
+and present the commands to replicate these results only.
+
 % 

### 2007-07-02 paper/trunk/JSS/response_A.tex
response to referee A: added answer that we currently cannot determine the reason for the deviations between systemfit's iterated FGLS results and Greene's ML results
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@778 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 %     wrong model?
+In theory, iterated FGLS should converge to ML results,
+but in practice these results often (more or less) slightly deviate.
+As the \pkg{systemfit} package currently does not provide ML estimations,
+we cannot determine the reason for the deviations
+between \code{systemfit()}'s iterated FGLS results
+and Greene's ML results.
+We will take a look at this issue,
+when we have finished to implement ML estimations in \pkg{systemfit()}.
+
 % 

### 2007-07-02 paper/trunk/systemfit_estimation.tex
replaced all \widehat{\beta} by the macro \bHat, which is defined as \hat{\beta}
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@719 591db2bd-e0a0-4d22-8c19-2192dbb7da31
replaced all \widehat{X} by the macro \XHat
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@721 591db2bd-e0a0-4d22-8c19-2192dbb7da31
made naming of (estimated) residuals and (true) disturbances more consistent
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@779 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 \begin{equation}
-   \widehat{\beta}_{OLS} = \left( X'X \right)^{-1} X'y .
+   \bHat_{OLS} = \left( X'X \right)^{-1} X'y .
    \label{eq:ols}
@@
 \begin{equation}
-   \COVHat \left[ \widehat{\beta}_{OLS} \right] = \sHat^2 \left( X'X \right)^{-1} ,
+   \COVHat \left[ \bHat_{OLS} \right] = \sHat^2 \left( X'X \right)^{-1} ,
    \label{eq:olsCovSameSigma}
 \end{equation}
-where $\sHat^2$ is the estimated variance of the residuals.
+where $\sHat^2$ is the estimated variance of the disturbances.
 This assumes that the disturbances of all equations have the
@@
 \begin{equation}
-   \COVHat \left[ \widehat{\beta}_{OLS} \right] = \left( X' \OHat^{-1} X \right)^{-1} ,
+   \COVHat \left[ \bHat_{OLS} \right] = \left( X' \OHat^{-1} X \right)^{-1} ,
    \label{eq:olsCovSingleSigma}
@@
 $\sHat_{ii} = \sHat_i^2$ is the estimated variance
-of the disturbence term in the $i$th equation.
+of the disturbance term in the $i$th equation.
 
@@
 \begin{equation}
-   \widehat{\beta}_{WLS} = \left( X' \OHat^{-1} X \right)^{-1} X' \OHat^{-1} y ,
+   \bHat_{WLS} = \left( X' \OHat^{-1} X \right)^{-1} X' \OHat^{-1} y ,
 \end{equation}
@@
 $\sHat_{ij} = 0 \; \forall \; i \neq j$ and
-$\sHat_{ii} = \sHat_i^2$ is the estimated variance
-of the disturbance term in the $i$th equation.%
-\footnote{%
-Theoretically, one should use the variances of the true residuals.
+$\sHat_{ii} = \sHat_i^2$ is the variance
+of the (estimated) residuals in the $i$th equation.
+Theoretically, it would be better to use the variances of
+the (true) disturbances $( \sigma_{ii} )$.
 However, they are not known in most empirical applications.
 Therefore, true variances are generally replaced by estimated variances
-that are calculated from the residuals of a first-step OLS estimation.
-}
+$( \sHat_{ii} )$
+that are calculated from the residuals of a first-step OLS estimation
+(see section~\ref{sec:residcov}).
+
 Like the OLS estimates, these estimates are only efficient
@@
 \begin{equation}
-   \COVHat \left[ \widehat{\beta}_{WLS} \right] = \left( X' \OHat^{-1} X \right)^{-1} .
+   \COVHat \left[ \bHat_{WLS} \right] = \left( X' \OHat^{-1} X \right)^{-1} .
 \end{equation}
@@
 \begin{equation}
-   \widehat{\beta}_{SUR} = \left( X' \OHat^{-1} X \right)^{-1} X' \OHat^{-1} y ,
+   \bHat_{SUR} = \left( X' \OHat^{-1} X \right)^{-1} X' \OHat^{-1} y ,
    \label{eq:sur}
@@
 \begin{equation}
-   \COVHat \left[ \widehat{\beta}_{SUR} \right] = \left( X' \OHat^{-1} X \right)^{-1} .
+   \COVHat \left[ \bHat_{SUR} \right] = \left( X' \OHat^{-1} X \right)^{-1} .
    \label{eq:surCov}
@@
 \begin{equation}
-   \widehat{X_i} = H_i \left( H_i' H_i \right)^{-1} H_i' X .
+   \XHat_i = H_i \left( H_i' H_i \right)^{-1} H_i' X .
 \end{equation}
@@
 \begin{equation}
-   \widehat{\beta}_{2SLS} = \left( \widehat{X}' \widehat{X} \right)^{-1} 
-   \widehat{X}' y .
+   \bHat_{2SLS} = \left( \XHat' \XHat \right)^{-1}
+   \XHat' y .
    \label{eq:beta2sls}
@@
 \begin{equation}
-   \COVHat \left[ \widehat{\beta}_{2SLS} \right] = \sHat^2 \left( \widehat{X}'
-   \widehat{X} \right)^{-1} ,
+   \COVHat \left[ \bHat_{2SLS} \right] = \sHat^2 \left( \XHat'
+   \XHat \right)^{-1} ,
    \label{eq:2slsCovSameSigma}
@@
 \begin{equation}
-   \COVHat \left[ \widehat{\beta}_{2SLS} \right] = \left( \widehat{X}' \OHat^{-1}
-   \widehat{X} \right)^{-1} ,
+   \COVHat \left[ \bHat_{2SLS} \right] = \left( \XHat' \OHat^{-1}
+   \XHat \right)^{-1} ,
    \label{eq:2slsCovSingleSigma}
@@
 \begin{equation}
-   \widehat{\beta}_{W2SLS} = \left( \widehat{X}' \OHat^{-1} \widehat{X}
-   \right)^{-1} \widehat{X}' \OHat^{-1} y ,
+   \bHat_{W2SLS} = \left( \XHat' \OHat^{-1} \XHat
+   \right)^{-1} \XHat' \OHat^{-1} y ,
 \end{equation}
@@
 \begin{equation}
-   \COVHat \left[ \widehat{\beta}_{W2SLS} \right] = \left( \widehat{X}' \OHat^{-1}
-   \widehat{X} \right)^{-1} .
+   \COVHat \left[ \bHat_{W2SLS} \right] = \left( \XHat' \OHat^{-1}
+   \XHat \right)^{-1} .
 \end{equation}
@@
 \begin{equation}
-   \widehat{\beta}_{3SLS} = \left( \widehat{X}' \OHat^{-1} \widehat{X}
-   \right)^{-1} \widehat{X}' \OHat^{-1} y ,
+   \bHat_{3SLS} = \left( \XHat' \OHat^{-1} \XHat
+   \right)^{-1} \XHat' \OHat^{-1} y ,
    \label{eq:3slsGls}
@@
 \begin{equation}
-   \COVHat \left[ \widehat{\beta}_{3SLS} \right] = \left( \widehat{X}' \OHat^{-1}
-   \widehat{X} \right)^{-1} .
+   \COVHat \left[ \bHat_{3SLS} \right] = \left( \XHat' \OHat^{-1}
+   \XHat \right)^{-1} .
    \label{eq:cov3sls}
@@
 \begin{equation}
-   \widehat{\beta}_{3SLS-IV} = \left( \widehat{X}' \OHat^{-1} X
-   \right)^{-1} \widehat{X}' \OHat^{-1} y .
+   \bHat_{3SLS-IV} = \left( \XHat' \OHat^{-1} X
+   \right)^{-1} \XHat' \OHat^{-1} y .
    \label{eq:3slsIv}
@@
 \begin{equation}
-   \COVHat \left[ \widehat{\beta}_{3SLS-IV} \right] = \left( \widehat{X}' \OHat^{-1}
+   \COVHat \left[ \bHat_{3SLS-IV} \right] = \left( \XHat' \OHat^{-1}
    X \right)^{-1} .
@@
 \begin{equation}
-   \widehat{\beta}_{3SLS-GMM} = \left( X' H \left( H' \OHat H \right)^{-1}
+   \bHat_{3SLS-GMM} = \left( X' H \left( H' \OHat H \right)^{-1}
    H' X \right)^{-1} X' H \left( H' \OHat H \right)^{-1} H' y .
@@
 \begin{equation}
-   \COVHat \left[ \widehat{\beta}_{3SLS-GMM} \right] =
+   \COVHat \left[ \bHat_{3SLS-GMM} \right] =
    \left( X' H \left( H' \OHat H \right)^{-1} H' X \right)^{-1} .
@@
 \begin{equation}
-   \widehat{\beta}_{3SLS-Schmidt} = \left( \widehat{X}' \OHat^{-1} \widehat{X}
-   \right)^{-1} \widehat{X}' \OHat^{-1} 
+   \bHat_{3SLS-Schmidt} = \left( \XHat' \OHat^{-1} \XHat
+   \right)^{-1} \XHat' \OHat^{-1}
    H \left( H' H \right)^{-1} H' y .
@@
 \begin{equation}
-   \COVHat \left[ \widehat{\beta}_{3SLS-Schmidt} \right] =
-   \left( \widehat{X}' \OHat^{-1}  \widehat{X} \right)^{-1}
-   \widehat{X}' \OHat^{-1} H \left( H' H \right)^{-1} H' \OHat
-   H \left( H' H \right)^{-1} H' \OHat^{-1} \widehat{X}
-   \left( \widehat{X}' \OHat^{-1}  \widehat{X} \right)^{-1} .
+   \COVHat \left[ \bHat_{3SLS-Schmidt} \right] =
+   \left( \XHat' \OHat^{-1}  \XHat \right)^{-1}
+   \XHat' \OHat^{-1} H \left( H' H \right)^{-1} H' \OHat
+   H \left( H' H \right)^{-1} H' \OHat^{-1} \XHat
+   \left( \XHat' \OHat^{-1}  \XHat \right)^{-1} .
 \end{equation}
@@
 \begin{equation}
-   \widehat{\beta}_{3SLS-EViews} = \widehat{\beta}_{2SLS} + 
-   \left( \widehat{X}' \OHat^{-1} \widehat{X} \right)^{-1}
-   \widehat{X}' \OHat^{-1} \left( y - X \widehat{\beta}_{2SLS} \right) ,
+   \bHat_{3SLS-EViews} = \bHat_{2SLS} +
+   \left( \XHat' \OHat^{-1} \XHat \right)^{-1}
+   \XHat' \OHat^{-1} \left( y - X \bHat_{2SLS} \right) ,
    \label{eq:3slsEViews}
 \end{equation}
-where $\widehat{\beta}_{2SLS}$ is the two-stage least squares estimator
+where $\bHat_{2SLS}$ is the two-stage least squares estimator
 as defined by (\ref{eq:beta2sls}).

### 2007-07-02 paper/trunk/systemfit_other.tex
added paragraph that the residual covariance matrix for a restricted WLS, SUR, W2SLS, or 3SLS estimation can be obtained either from a restricted or from an unrestricted first-step estimation; added paragraph that argument "residCovRestricted" can be used to specify this in systemfit()
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@739 591db2bd-e0a0-4d22-8c19-2192dbb7da31
made naming of (estimated) residuals and (true) disturbances more consistent
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@779 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 
-Since the true residuals of the estimated equations are generally not known,
-the true covariance matrix of the residuals cannot be determined.
+Since the (true) disturbances of the estimated equations are generally not known,
+their covariance matrix cannot be determined.
 Thus, this covariance matrix must be calculated from the
-\emph{estimated} residuals. 
-Generally, the estimated covariance matrix of the residuals
+estimated residuals.
+Generally, the residual covariance matrix
 ($\SHat = \left[ \sHat_{ij} \right]$)
@@
 
-A further approach to obtain the estimated residual covariance
+A further approach to obtain a residual covariance
 matrix is
@@
 The WLS, SUR, W2SLS and 3SLS parameter estimates are consistent
-if the estimated residual covariance matrix is calculated
+if the residual covariance matrix is calculated
 using the residuals from a first-step OLS or 2SLS estimation.
@@
 from the residuals of the previous iteration.
-If equation (\ref{eq:rcovNoDfCor}) is applied to calculate the estimated
+If equation (\ref{eq:rcovNoDfCor}) is applied to calculate the
 residual covariance matrix,
@@
 for $\uHat_i$ in (\ref{eq:rcovNoDfCor}--\ref{eq:rcovTheil}).
+
+If any restrictions are imposed on the parameters,
+the residual covariance matrix
+for a WLS, SUR, W2SLS, or 3SLS estimation
+can be obtained either from a restricted or from an unrestricted
+first-step estimation.
 
@@
 $K$ is the total number of estimated coefficients, and
-$\SHat$ is the estimated residual covariance matrix
+$\SHat$ is the (estimated) residual covariance matrix
 used in the estimation.

### 2007-07-02 paper/trunk/systemfit_statistics.tex
corrected the paper according to all minor comments of reviewer A except for "p.3, l.45", the second part of "p.3, l.52" and "p.18, Tab.1".
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@687 591db2bd-e0a0-4d22-8c19-2192dbb7da31
substituted "\E [ . ]" (\E is defined in jss.cls) for all "E ( . )" and "E [ . ]"
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@711 591db2bd-e0a0-4d22-8c19-2192dbb7da31
made naming of (estimated) residuals and (true) disturbances more consistent
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@779 591db2bd-e0a0-4d22-8c19-2192d

@@
       0 & 0 & \cdots & X_G
-   \end{array}\right] \cdot 
+   \end{array}\right]
    \left[ \begin{array}{c}
@@
 \begin{equation}
-   E \left( u_{it} \, u_{jt^*} \right) = 0 
-   \; \forall \; t \neq t^* ,
+   \E \left[ u_{it} \, u_{js} \right] = 0
+   \; \forall \; t \neq s ,
 \end{equation}
 where $i$ and $j$ indicate the equation number 
-and $t$ and $t^*$ denote the observation number.
+and $t$ and $s$ denote the observation number.
 
@@
 \begin{equation}
-   E \left( u_{it} \, u_{jt} \right) = \sigma_{ij} .
+   \E \left[ u_{it} \, u_{jt} \right] = \sigma_{ij} .
 \end{equation}
 
-Thus, the covariance matrix of the total system is
+Thus, the covariance matrix of all disturbances is
 \begin{equation}
-   E \left( u \, u' \right) = \Omega = \Sigma \otimes I ,
+   \E \left[ u \, u' \right] = \Omega = \Sigma \otimes I ,
 \end{equation}
-where $\Sigma = \left[ \sigma_{ij} \right]$ is the residual covariance matrix
-and $I$ is an identity matrix.
+where $\Sigma = \left[ \sigma_{ij} \right]$ is the (contemporaneous)
+disturbance covariance matrix and $I$ is an identity matrix.
 

### 2007-07-02 paper/trunk/JSS/response_B.tex
slightly revised response to referee B
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@783 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 
-We have mentioned "instrumental variables" in the abstract, include
+We have mentioned °instrumental variables° in the abstract, included
 three examples of topics that use IV methods and have added
-"instrumental variables" to the keywords.
+°instrumental variables° to the keywords.
 

### 2007-07-03 paper/trunk/systemfit_statistics.tex
added some references that describe simultaneous equation systems
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@784 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 
-In this section we provide the statistical background
-of the functionality provided by the \pkg{systemfit} package.
+In this section we give a short overview of the statistical background
+on that the \pkg{systemfit} package is based on.
+More detailed descriptions of simultaneous equations systems
+are available for instance in
+\citet[chapter~7]{theil71}
+\citet[part~4]{judge82}
+\citet[part~5]{judge85}
+\citet{srivastava87},
+\citet[chapters 14--15]{greene03}, and
+\citet[chapter~10]{zivot06}.
 After introducing notations and assumptions,

### 2007-07-03 paper/trunk/systemfit_code.tex
added section "Overlap with other functions and packages in R"
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@786 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 that belong to the $i$'s equation.
+
+\subsection[Overlap with other functions and packages in R]
+   {Overlap with other functions and packages in \proglang{R}}
+
+Single-equation models can be fitted in \proglang{R}
+by OLS with function \code{lm} (package \pkg{stats})
+and by 2SLS with function \code{tsls}
+(package \pkg{sem}, \citealp{r-sem-0.9-8}).
+This is also possible with the \code{systemfit} function,
+but \code{systemfit} is specialized in estimating systems of equation,
+i.e.\ more than one equation.
+Its capability to estimate single-equation models is just a side-effect.
+
+Function \code{sem} (package \pkg{sem}, \citealp{r-sem-0.9-8})
+can be used to estimate structural equation models in \proglang{R}
+by Full Information Maximum Likelihood (FIML)
+assuming multinormal errors.
+A special feature of this function is the estimation
+of models with unobserved (°latent°) variables,
+which is not possible with \code{systemfit}.
+While \code{sem} cannot be used to consistently estimate
+systems of simultaneous equations with some endogenous regressors,
+it can be used to estimate systems of equations,
+where all regressors are exogenous.
+However, the latter is rather cumbersome.
+Hence, \code{systemfit} is the only function in \proglang{R}
+that can be used to estimate systems of simultaneous equations
+and it is the most convenient function to estimate
+systems of equations with purely exogenous regressors.

### 2007-07-03 paper/trunk/runSweave.sh
added appendix (section) that shows how to estimate systems of equations with "sem"
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@787 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@ -2,3 +2,5 @@
 echo "Sweave(\"systemfit_usage.Snw\")" | R --no-save --no-restore
 
 echo "Sweave(\"systemfit_reliability.Snw\")" | R --no-save --no-restore
+
+echo "Sweave(\"systemfit_sem.Snw\")" | R --no-save --no-restore

### 2007-07-04 paper/trunk/JSS/response_B.tex
response to referee B: added answer that we have added several references to literature that provides a detailed descriptions of simultaneous equations systems
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@785 591db2bd-e0a0-4d22-8c19-2192dbb7da31
response to referee B: revised answer to the "Overall Comment" (no ML methods, comparison with sem)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@788 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 
-\textbf{[Still working on this section]}
-
-It was our intention to create a package for fitting models that could
+It was our intention to create a package
+for fitting simultaneous equation models that could
 be verified against published results before moving on to include more
@@
 The reasons for doing this were two-fold.  Our first reason was to
-simply introduce an econometrics package where the main focus of the
-package was "pure" econometrics and not to have a few specialized
-routines, as is the case with the tsls() function is the SEM
+simply introduce an econometrics package where the main objective of the
+package was a unified framework for simultaneous equation models
+and not to have a few specialized
+routines, as is the case with the \code{tsls()} function in the \pkg{sem}
 package. While there are definitely links among structural equation
@@
 
-The second reason for creating the package was to create a "forum" for
+The second reason for creating the package was to create a °forum° for
 more advanced fitting methods as described in Section 6 (Summary and
-Outlook). It is our intention to impliment these more advanced methods
+Outlook). It is our intention to implement these more advanced methods
 (unbalanced data sets, FIML, GMM, etc.) over time, but we felt that
-having a "basic" package that was accepted as part of a peer-review
+having a °basic° package that was accepted as part of a peer-review
 process would create an environment that would reduce resistance to
@@
 subject to the difficulties beyond simple numerical routines as were
-used in the systemfit package so far.  Perhaps once the package gains
-enough acceptance, the functions can be made part of the core
-functions of R, but attempting to directly include the functions
-presented in the package might prove to be exceedingly difficult with
-the core development team.
+used in the \pkg{systemfit} package so far.
 
-At the time the package was originally introduced (2002), there were
-very few econometrics functions for R and only recently have more
-functions and packages been submitted. Adding these functions to an
-existing econometrics package would have made sense, had one been
-available.
+Encouraged by the question of this referee,
+we have added a comparison of \code{systemfit()} and \code{sem()}
+(see section~3.4 and appendix~B).
 
@@
 % would help to include more references, especially in section 2.1.
+
+We have added several references to literature (e.g.\ econometrics textbooks)
+that provides a detailed description of simultaneous equations systems.
 

### 2007-07-04 paper/trunk/systemfit_other.tex
improved description of two-step method to obtain FGLS results (suggested by referee B)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@789 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 their covariance matrix cannot be determined.
-Thus, this covariance matrix must be calculated from the
-estimated residuals.
-Generally, the residual covariance matrix
-($\SHat = \left[ \sHat_{ij} \right]$)
-can be calculated from the residuals of a first-step OLS or 2SLS estimation.
-In many cases, the formula
+Therefore, this covariance matrix is generally calculated from
+estimated residuals
+that are obtained from a first-step OLS or 2SLS estimation.
+Then, in a second step, the estimated residual covariance matrix
+can be employed for a WLS, SUR, W2SLS, or 3SLS estimation.
+In many cases, the residual covariance matrix is calculated by
 \begin{equation}
-   \sHat_{ij} = \frac{ \uHat_i' \uHat_j }{ T }
+   \sHat_{ij} = \frac{ \uHat_i' \uHat_j }{ T },
    \label{eq:rcovNoDfCor}
 \end{equation}
-is applied, where $T$ is the number of observations in each equation.
+where $T$ is the number of observations in each equation.
 However, in finite samples this estimator is biased,

### 2007-07-04 paper/trunk/JSS/response_B.tex
response to referee B: revised a few answers
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@790 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 
-This is stated in the introduction on Page one, lines 13-14.
+This is stated in the introduction (page 2, lines 14-15).
 
-\subsection{'may' should be 'must'}
+\subsection{Instrumental variables}
 % Page 4, line 82: 'may' should be 'must'. Here the difference is extremely important. Maybe the text
@@
 We have added a note to make the reader aware of the difference
-between the two equations.
+between these two equations.
 
-\subsection{Sigma}
+\subsection{Residual covariance matrix (Sigma)}
 % Line 193-194: Language is confusing. It should be explained that calculation of some of the
@@
 
-We have made minor edits to the section to avoid any possible
-confusion. We discovered in the currently existing body of literature
-that there is no well defined presentation of computing the residual
-covariance matrix and have made a brief attempt here, but our attempt
-was by no means exhaustive.
+We have described this issue more clearly now.
+For instance in sections~2.1 and~2.2, the residual covariance matrix
+is no longer treated as known.
+Hence, the fact that $\Sigma$ and $\Omega$ are unknown and have to be estimated
+is no longer °hidden° in section 2.3.
 
@@
 
-We have removed the term "first-step" as previously used in the
-manuscript. This term refers to the first-stage of a multi-stage
-estimation process (OLS->SUR) or (2SLS->3SLS).
+We have explained the two-step procedure
+to obtain the estimated residual covariance matrix
+more clearly now.
 

### 2007-07-04 paper/trunk/systemfit_estimation.tex
improved description of two-step method to obtain FGLS results (suggested by referee B)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@789 591db2bd-e0a0-4d22-8c19-2192dbb7da31
minor (but important) correction suggested by referee B (may -> must); added answer to referee B that we have done this
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@791 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 of the (estimated) residuals in the $i$th equation.
-Theoretically, it would be better to use the variances of
+Theoretically, $\sHat_{ii}$ should be the variance of
 the (true) disturbances $( \sigma_{ii} )$.
@@
 can be either different or identical for all equations.
-The instrumental variables of each equation may not be correlated with 
+The instrumental variables of each equation must not be correlated with
 the disturbance terms of the corresponding equation 

### 2007-07-04 paper/trunk/JSS/response_B.tex
minor (but important) correction suggested by referee B (may -> must); added answer to referee B that we have done this
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@791 591db2bd-e0a0-4d22-8c19-2192dbb7da31
made clearer that GLS is efficient, but FGLS is only asymptotically efficient (suggested by referee B); added answer to referee B that we have done this
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@792 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 \subsection{Instrumental variables}
-% Page 4, line 82: 'may' should be 'must'. Here the difference is extremely important. Maybe the text
+% Page 4, line 82: 'may' should be 'must'. Here the difference is extremely important.
+
+According to the comment of this referee,
+we have replaced °may° by °must°.
+
+% Maybe the text
 % could emphasize more clearly what is required of a variable if it is to be used as an instrumental
@@
 
-\textbf{[Arne, I'm not sure how to answer this one. I'm not sure this is an
-important issue for the manuscript. We mention that FIML is a more
-sophisticated approach in the summary and outlook section, but we
-don't really mention FIML in the paper. Should we?]}
+We have noted that the GLS estimator (with known $\Sigma$) is efficient,
+but the FGLS estimator (with unknown $\Sigma$)
+is only \emph{asymptotically} efficient.
+Since the iterated SUR (FGLS) estimator converges to FIML results,
+the FIML method is not more efficient than iterated SUR (FGLS).
 

### 2007-07-04 paper/trunk/systemfit_estimation.tex
made clearer that GLS is efficient, but FGLS is only asymptotically efficient (suggested by referee B); added answer to referee B that we have done this
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@792 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 
-Like the OLS estimates, these estimates are only efficient
+These estimates are (asymptotically) efficient
 if the disturbance terms are not contemporaneously correlated.
@@
 (SUR) estimator \citep{zellner62}.
-Since the true covariance matrix of the disturbance terms
-is generally unknown,
-a Feasible Generalized Least Squares (FGLS) estimation,
-which is based on an estimated covariance matrix of the disturbance terms,
-is the textbook solution for this problem.
+However, the true covariance matrix of the disturbance terms
+is generally unknown.
+The textbook solution for this problem is
+a Feasible Generalized Least Squares (FGLS) estimation.
+As the FGLS estimator is based on
+an estimated covariance matrix of the disturbance terms,
+it is only asymptotically efficient.
 It should be noted that while an unbiased OLS or WLS estimation requires only that

### 2007-07-04 paper/trunk/JSS/response_B.tex
explained the SUR approach of LIMDEP's TSCS function (OLS -> WLS -> SUR) more clearly now (suggested by referee B); added answer to referee B that we have done this
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@793 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 
-\textbf{[Arne, can we simply not mention the "alternative method" in the
-paper? I think it should stay as is, but I need a little more
-understanding of the reasoning behind the method and can't defend
-it. I just don't have any exposure to it yet.]}
+We have explained the °alternative approach° more clearly now.
 

### 2007-07-04 paper/trunk/systemfit_other.tex
explained the SUR approach of LIMDEP's TSCS function (OLS -> WLS -> SUR) more clearly now (suggested by referee B); added answer to referee B that we have done this
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@793 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 using the residuals from a first-step OLS or 2SLS estimation.
-There exists also an alternative slightly different approach.%
+There exists also an alternative slightly different approach
+that consists of three steps.%
 \footnote{
@@
 }
-This alternative approach uses the residuals of a first-step OLS or 2SLS estimation
-to apply a WLS or W2SLS estimation on a second step.
-Then, it calculates the residual covariance matrix
-from the residuals of the second-step estimation
-to estimates the model by SUR or 3SLS in a third step.
+In a first step, an OLS or 2SLS estimation is applied to obtain residuals
+to calculate a (first-step) residual covariance matrix.
+In a second step, the first-step residual covariance matrix
+is used to estimate the model by WLS or W2SLS
+and new residuals are obtained to calculate
+a (second-step) residual covariance matrix.
+Finally, in the third step,
+the second-step residual covariance matrix is used
+to estimate the model by SUR or 3SLS.
 If no cross-equation restrictions are imposed,

### 2007-07-04 paper/trunk/JSS/response_B.tex
explained the syntax for instrumental variables (one-sided formula) more clearly now (suggested by referee B); added answer to referee B that we have done this
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@794 591db2bd-e0a0-4d22-8c19-2192dbb7da31
corrected this mistake in the description of the null hypothesis of the Hausman test (pointed out by referee B) and explained the alternative hypothesis in more detail; added answer to referee B that we have done this
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@795 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 
-We have included the proper citation for the R manuals which describes
-the syntax for formulas.
+We have explained the syntax for instrumental variables (one-sided formula)
+more clearly now.
 
-\subsection{Section 2.7}
+\subsection{Hausman test (section 2.7)}
 % Section 2.7: Given the theory of the previous sections, I was confused about the Hausman test.
@@
 
-The Hausman test is somewhat unique from other detection tests in that
-the test is designed "backwards" in the sense that the model that is
-being tested is correctly specified and not whether a predetermined 
-value is sufficiently close to the parameter estimate.
-
-\textbf{[I'm tired and will have to get back to this]}
+Indeed, there was a mistake in the description of the null hypothesis
+of the Hausman test.
+We have corrected this mistake
+and explain the alternative hypothesis in more detail now.
 

### 2007-07-04 paper/trunk/systemfit_other.tex
corrected this mistake in the description of the null hypothesis of the Hausman test (pointed out by referee B) and explained the alternative hypothesis in more detail; added answer to referee B that we have done this
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@795 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 \citet{hausman78} developed a test for misspecification.
-The null hypothesis of the test is that all exogenous variables are
-uncorrelated with all disturbance terms.
+The null hypothesis of the test is
+that the instrumental variables of each equation are
+uncorrelated with the disturbance terms of all other equations
+($\E \left[ u_i' H_j \right] = 0 \, \forall \, i \neq j$).
 Under this hypothesis both the 2SLS and the 3SLS estimator are consistent
@@
 Under the alternative hypothesis the 2SLS estimator is consistent
-but the 3SLS estimator is inconsistent.
+but the 3SLS estimator is inconsistent,
+i.e.\ the instrumental variables of each equation are
+uncorrelated with the disturbances of the same equation
+($\E \left[ u_i' H_i \right] = 0 \, \forall \, i$),
+but the instrumental variables of at least one equation
+are correlated with the disturbances of another equation
+($\E \left[ u_i' H_j \right] \neq 0 \, \exists \, i \neq j$).
 The Hausman test statistic is

### 2007-07-04 trunk/R/calcResidCov.R
renamed internal function ".calcRCov()" to ".calcResidCov()"
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@448 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@ -1,5 +1,5 @@
 ## Calculate the residual covariance matrix
-.calcRCov <- function( resids, methodResidCov, nObsEq = NULL, nCoefEq = NULL, xEq = NULL,
+.calcResidCov <- function( resids, methodResidCov, nObsEq = NULL, nCoefEq = NULL, xEq = NULL,
       diag = FALSE, centered = FALSE, solvetol = .Machine$double.eps ) {
 
    eqNames <- NULL

### 2007-07-04 trunk/R/fiml.R

@@ -9,7 +9,7 @@
          length( mlCoef ), " but must have length ", ncol( mlVars$xMat ) )
    }
    mlResids <- mlVars$yVec - mlVars$xMat %*% mlCoef
-   mlSigma <- .calcRCov( resids = mlResids, methodResidCov = mlVars$methodResidCov,
+   mlSigma <- .calcResidCov( resids = mlResids, methodResidCov = mlVars$methodResidCov,
       nObsEq = mlVars$nObsEq, nCoefEq = mlVars$nCoefEq, xEq = mlVars$xEq,
       centered = mlVars$centerResiduals, diag = FALSE,
       solvetol = mlVars$solvetol )

### 2007-07-04 trunk/R/logLik.R

@@ -1,7 +1,7 @@
 logLik.systemfit <- function( object, ... ){
 
    resid <- residuals( object )
-   residCov <- .calcRCov( resid, "noDfCor" )
+   residCov <- .calcResidCov( resid, "noDfCor" )
    nObsPerEq <- nrow( resid )
    nEq <- ncol( resid )
 

### 2007-07-04 trunk/R/summary.systemfit.R

@@ -94,7 +94,7 @@ summary.systemfit <- function( object, useDfSys = NULL,
       rtOmega <- .calcXtOmegaInv( xMat = matrix( resid, ncol = 1 ),
          sigma = object$residCov, nObsEq = nObsEq,
          solvetol = object$control$solvetol )
-      yCov <- .calcRCov( response, methodResidCov = "noDfCor",
+      yCov <- .calcResidCov( response, methodResidCov = "noDfCor",
          nObsEq = nObsEq, centered = TRUE,
          solvetol = object$control$solvetol )
       residCovInv <- solve( object$residCov, tol = object$control$solvetol )

### 2007-07-04 trunk/R/systemfit.R

@@ -317,7 +317,7 @@ systemfit <- function(  eqns,
   if(method=="OLS") {
     resids <- yVecAll - xMatAll %*% coef                                        # residuals
     if(control$single.eq.sigma) {
-      rcov <- .calcRCov( resids, methodResidCov = control$methodResidCov, nObsEq = nObsEq,
+      rcov <- .calcResidCov( resids, methodResidCov = control$methodResidCov, nObsEq = nObsEq,
          nCoefEq = nCoefLiEq, xEq = xMatEq, diag = TRUE, centered = control$centerResiduals,
          solvetol = control$solvetol )               # residual covariance matrix
       coefCov <- .calcGLS( xMat = xMatAll, R.restr = R.restr, q.restr = q.restr,
@@ -346,7 +346,7 @@ systemfit <- function(  eqns,
       iter  <- iter+1
       bl    <- coef                # coefficients of previous step
       resids <- yVecAll - xMatAll %*% coef     # residuals
-      rcov <- .calcRCov( resids, methodResidCov = control$methodResidCov, nObsEq = nObsEq,
+      rcov <- .calcResidCov( resids, methodResidCov = control$methodResidCov, nObsEq = nObsEq,
          nCoefEq = nCoefLiEq, xEq = xMatEq, diag = TRUE, centered = control$centerResiduals,
          solvetol = control$solvetol )
       coef  <- .calcGLS( xMat = xMatAll, yVec = yVecAll, R.restr = R.restr, q.restr = q.restr,
@@ -368,7 +368,7 @@ systemfit <- function(  eqns,
       iter  <- iter+1
       bl    <- coef                           # coefficients of previous step
       resids <- yVecAll-xMatAll%*%coef                     # residuals
-      rcov <- .calcRCov( resids, methodResidCov = control$methodResidCov, nObsEq = nObsEq,
+      rcov <- .calcResidCov( resids, methodResidCov = control$methodResidCov, nObsEq = nObsEq,
          nCoefEq = nCoefLiEq, xEq = xMatEq, centered = control$centerResiduals,
          solvetol = control$solvetol )
       coef <- .calcGLS( xMat = xMatAll, yVec = yVecAll, R.restr = R.restr, q.restr = q.restr,
@@ -404,7 +404,7 @@ systemfit <- function(  eqns,
   if(method=="2SLS") {
     resids <- yVecAll - xMatAll %*% coef                        # residuals
[…]

### 2007-07-04 paper/trunk/JSS/response_A.tex
response to referee A: removed note that some of the minor points have not (yet) been corrected
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@780 591db2bd-e0a0-4d22-8c19-2192dbb7da31
changes due to renaming argument "methodRCov" to "methodResidCov" in the systemfit package
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@796 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 (former \code{formula3sls} and \code{rcovformula})
-have been renamed to \code{method3sls} and \code{methodRCov},
+have been renamed to \code{method3sls} and \code{methodResidCov},
 respectively.
@@
 \section{Minor Points and Typos}
-We have corrected the paper according to all of these comments
-except for °p.3, l.45° and the second part of °p.3, l.52°.
+We have corrected the paper according to all of these comments.
 

### 2007-07-12 paper/trunk/systemfit_code.tex
added appendix (section) that shows how to estimate systems of equations with "sem"
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@787 591db2bd-e0a0-4d22-8c19-2192dbb7da31
added paragraph that (and how) we use the "Matrix" package for the GLS estimations
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@803 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 First, if a formula contains an inverse of a matrix
-that is post-multiplied by a vector,
+that is post-multiplied by a vector or matrix,
 we use \code{solve(A,b)} instead of \code{solve(A) \%*\% b}.
@@
 \end{array} \right]' ,
+\label{eq:omegaInv}
 \end{equation}
@@
 that belong to the $i$'s equation.
+
+Since version 1.0, \pkg{systemfit} uses the \pkg{Matrix} package
+\citep{r-matrix-07} by default.
+This package provides classes for different types of matrices.
+For instance, we choose class \code{dgeMatrix}
+(°real matrices in general storage mode°),
+for matrices $X_i$ in~(\ref{eq:model-array}),
+class \code{dgCMatrix} (°general, numeric, sparse matrices
+in the (sorted) compressed sparse column format°)
+for matrix $X$ in~(\ref{eq:model-matrices}),
+and class \code{dspMatrix}
+(°symmetric real matrices in packed storage (one triangle only)°)
+for the residual covariance matrix $\SHat$.
+If the \pkg{Matrix} package is used,
+the possibly huge matrix $\Omega^{-1}$ is no longer a problem,
+because it is a sparse matrix
+that can be stored in a compressed format (class \code{dgCMatrix}).
+Hence, we no longer need the algorithm in equation~(\ref{eq:omegaInv}).
+We have tried different ways to calculate a GLS estimator
+like in (\ref{eq:sur})
+and we found
+that the following code is the fastest:
+\begin{Schunk}
+\begin{Sinput}
+> sigmaInv <- solve( residCov )
+> xtOmegaInv <- t( xMat ) %*% kronecker( sigmaInv, Diagonal( nObs ) )
+> coef <- solve( xtOmegaInv %*% xMat, xtOmegaInv %*%  yVec )
+\end{Sinput}
+\end{Schunk}
+In this code snippet,
+\code{residCov} is the residual covariance matrix $\SHat$,
+\code{xMat} is the matrix $X$ and
+\code{yVec} is the vector $y$ in~(\ref{eq:sur}).
 
@@
 where all regressors are exogenous.
-However, the latter is rather cumbersome.
+However, the latter is rather cumbersome
+(see appendix~\ref{sec:sem}).
 Hence, \code{systemfit} is the only function in \proglang{R}

### 2007-07-22 paper/trunk/systemfit_abstract.tex
made small changes to attempt updating the repos
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@814 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 within the \proglang{R} programming environment.
-For instance, this package can be used for estimations
-by ``Ordinary Least Squares'' (OLS),
+For instance, this package can be used for 
+``Ordinary Least Squares'' (OLS),
 ``Seemingly Unrelated Regression'' (SUR),

### 2007-07-23 paper/trunk/systemfit_intro.tex
renamed section "Testing reliability" to "Replication of textbook results"; added more goals of this section (teaching + reproducible research)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@776 591db2bd-e0a0-4d22-8c19-2192dbb7da31
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@817 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 The implementation of the statistical procedures in \proglang{R} is
-shortly explained in Section~\ref{sec:code}.
+briefly explained in Section~\ref{sec:code}.
 Section~\ref{sec:Usage} demonstrates how to run
 \pkg{systemfit} and how some of the features presented in the previous
-section can be utilized.
-In Section~\ref{sec:reliability} the reliability of the results from
-\pkg{systemfit} are presented.
+section can be used.
+In Section~\ref{sec:reliability} we replicate several textbook results
+with the \pkg{systemfit} package.
 Finally, a summary and outlook are presented in

### 2007-07-23 paper/trunk/systemfit_restrictions.tex
replaced all \widehat{X} by the macro \XHat
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@721 591db2bd-e0a0-4d22-8c19-2192dbb7da31
replaced all \widehat{\lambda} by the macro \lHat, which has been defined as \hat{\lambda}
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@722 591db2bd-e0a0-4d22-8c19-2192dbb7da31
increased space between rows in some arrays
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@723 591db2bd-e0a0-4d22-8c19-2192dbb7da31
renamed (subsub)section that describes the calculation of restricted estimators (method 1 and 2) -> Calculation

@@
 }
-There are two ways to impose linear parameter restrictions.
+There are two methods to impose linear parameter restrictions.
 First, a matrix $M$ can be specified that
@@
 
-The second way to impose linear parameter restrictions
+The second method to impose linear parameter restrictions
 can be formulated by
@@
 
-The first way is less flexible than this latter one%
+The first method is less flexible than the second%
 \footnote{
@@
 }, 
-but the first way is preferable if equality constraints for coefficients
+and is preferable if equality constraints for coefficients
 across many equations of the system are imposed. 
@@
 
-\subsubsection[Implementation in systemfit]{Implementation in \pkg{systemfit}}
+\subsubsection{Calculation of restricted estimators}
 
@@
    \left[ \begin{array}{c}
-      \bHat^R_{OLS} \\ \widehat{\lambda}
+      \bHat^R_{OLS} \\[0.2em] \lHat
    \end{array} \right]
@@
    \left[ \begin{array}{c}
-      \bHat^R_{OLS} \\ \widehat{\lambda}
+      \bHat^R_{OLS} \\[0.2em] \lHat
    \end{array} \right] 
@@
    \left[ \begin{array}{c}
-      \bHat^R_{WLS} \\ \widehat{\lambda}
+      \bHat^R_{WLS} \\[0.2em] \lHat
    \end{array} \right]
@@
    \left[ \begin{array}{c}
-      \bHat^R_{WLS} \\ \widehat{\lambda}
+      \bHat^R_{WLS} \\[0.2em] \lHat
    \end{array} \right] 
@@
    \left[ \begin{array}{c}
-      \bHat^R_{SUR} \\ \widehat{\lambda}
+      \bHat^R_{SUR} \\[0.2em] \lHat
    \end{array} \right]
@@
    \left[ \begin{array}{c}
-      \bHat^R_{SUR} \\ \widehat{\lambda}
+      \bHat^R_{SUR} \\[0.2em] \lHat
    \end{array} \right] 
@@
    \left[ \begin{array}{c}
-      \bHat^R_{2SLS} \\ \widehat{\lambda}
+      \bHat^R_{2SLS} \\[0.2em] \lHat
    \end{array} \right]
@@
    \left[ \begin{array}{cc}
-      \widehat{X}' \widehat{X} & R' \\ 
+      \XHat' \XHat & R' \\
       R & 0
@@
    \left[ \begin{array}{c}
-      \widehat{X}' y \\ q 
+      \XHat' y \\ q
    \end{array} \right] .
@@
    \left[ \begin{array}{c}
-      \bHat^R_{2SLS} \\ \widehat{\lambda}
+      \bHat^R_{2SLS} \\[0.2em] \lHat
    \end{array} \right] 
@@
    \left[ \begin{array}{cc}
-      \widehat{X}' \OHat^{-1} \widehat{X} & R' \\
+      \XHat' \OHat^{-1} \XHat & R' \\
       R & 0
@@
    \left[ \begin{array}{c}
-      \bHat^R_{W2SLS} \\ \widehat{\lambda}
+      \bHat^R_{W2SLS} \\[0.2em] \lHat
    \end{array} \right]
@@
    \left[ \begin{array}{c}
-      \widehat{X}' \OHat^{-1} y \\ q
+      \XHat' \OHat^{-1} y \\ q
    \end{array} \right] ,
@@
    \left[ \begin{array}{c}
-      \bHat^R_{W2SLS} \\ \widehat{\lambda}
+      \bHat^R_{W2SLS} \\[0.2em] \lHat
    \end{array} \right] 
@@
    \left[ \begin{array}{c}
-      \bHat^R_{3SLS} \\ \widehat{\lambda}
+      \bHat^R_{3SLS} \\[0.2em] \lHat
    \end{array} \right]
@@
    \left[ \begin{array}{c}
-      \bHat^R_{3SLS} \\ \widehat{\lambda}
+      \bHat^R_{3SLS} \\[0.2em] \lHat
    \end{array} \right] 
@@
    \left[ \begin{array}{c}
-      \bHat^R_{3SLS-IV} \\ \widehat{\lambda}
+      \bHat^R_{3SLS-IV} \\[0.2em] \lHat
    \end{array} \right]
@@
    \left[ \begin{array}{cc}
-      \widehat{X}' \OHat^{-1} X & R' \\
+      \XHat' \OHat^{-1} X & R' \\
       R & 0
@@
    \left[ \begin{array}{c}
-      \bHat^R_{3SLS-IV} \\ \widehat{\lambda}
+      \bHat^R_{3SLS-IV} \\[0.2em] \lHat
    \end{array} \right] 
@@
    \left[ \begin{array}{c}
-      \bHat^R_{3SLS-GMM} \\ \widehat{\lambda}
+      \bHat^R_{3SLS-GMM} \\[0.2em] \lHat
    \end{array} \right]
@@
    \left[ \begin{array}{c}
-      \bHat^R_{3SLS-GMM} \\ \widehat{\lambda}
+      \bHat^R_{3SLS-GMM} \\[0.2em] \lHat
    \end{array} \right] 
@@
    \left[ \begin{array}{c}
-      \bHat^R_{3SLS-Schmidt} \\ \widehat{\lambda}
+      \bHat^R_{3SLS-Schmidt} \\[0.2em] \lHat
    \end{array} \right]
@@
    \left[ \begin{array}{c}
-      \widehat{X}' \OHat^{-1} H \left( H' H \right)^{-1} H' y \\ q
+      \XHat' \OHat^{-1} H \left( H' H \right)^{-1} H' y \\ q
    \end{array} \right] ,
@@
    \left[ \begin{array}{c}
-      \bHat^R_{3SLS-Schmidt} \\ \widehat{\lambda}
+      \bHat^R_{3SLS-Schmidt} \\[0.2em] \lHat
    \end{array} \right] 
@@
    \left[ \begin{array}{cc}
-      \widehat{X}' \OHat^{-1} H \left( H' H \right)^{-1} H' \OHat
-      H \left( H' H \right)^{-1} H' \OHat^{-1} \widehat{X} & 0' \\
+      \XHat' \OHat^{-1} H \left( H' H \right)^{-1} H' \OHat
+      H \left( H' H \right)^{-1} H' \OHat^{-1} \XHat & 0' \\
       0 & 0
@@
    \left[ \begin{array}{c}
-      \bHat^R_{3SLS-EViews} \\ \widehat{\lambda}
+      \bHat^R_{3SLS-EViews} \\[0.2em] \lHat
    \end{array} \right]
@@
    \left[ \begin{array}{c}
-      \widehat{X}' \OHat^{-1} \left( y - X \bHat^R_{2SLS} \right)
+      \XHat' \OHat^{-1} \left( y - X \bHat^R_{2SLS} \right)
       \\ q 

### 2007-07-23 paper/trunk/systemfit_summary.tex
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@817 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 equations may reduce the number of observations considerably, and
-thus, the information utilized in the estimation.
+thus, the information utilized in estimation of system parameters.
 Hence, it is our intention to include the capability for estimations

### 2007-07-24 paper/trunk/systemfit_estimation.tex
some small editings and corrections in sections 1-3
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@819 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 can be either different or identical for all equations.
-The instrumental variables of each equation must not be correlated with
+They must not be correlated with
 the disturbance terms of the corresponding equation 
@@
 a Feasible Generalized Least Squares (FGLS) version of the two-stage least squares
-estimation leads to consistent and asymptotically efficient estimates.
+estimation leads to consistent and asymptotically more efficient estimates.
 This estimation procedure is generally called °Three-stage Least
@@
 if all disturbance terms and all instrumental variables are uncorrelated
-$( \E \left[ u_i' H_j \right]) = 0 \; \forall \; i, j )$
-with
-\begin{equation}
-   H =
-   \left[ \begin{array}{cccc}
-      H_1 & 0 & \cdots & 0\\
-      0 & H_2 & \cdots & 0\\
-      \vdots & \vdots & \ddots & \vdots\\
-      0 & 0 & \cdots & H_G
-   \end{array}\right] .
-\end{equation}
+$( \E \left[ u_i' H_j \right]) = 0 \; \forall \; i, j )$.
+
 Since there might be occasions where this cannot be avoided,
@@
    \bHat_{3SLS-GMM} = \left( X' H \left( H' \OHat H \right)^{-1}
-   H' X \right)^{-1} X' H \left( H' \OHat H \right)^{-1} H' y .
+   H' X \right)^{-1} X' H \left( H' \OHat H \right)^{-1} H' y
    \label{eq:3slsGmm}
+\end{equation}
+with
+\begin{equation}
+   H =
+   \left[ \begin{array}{cccc}
+      H_1 & 0 & \cdots & 0\\
+      0 & H_2 & \cdots & 0\\
+      \vdots & \vdots & \ddots & \vdots\\
+      0 & 0 & \cdots & H_G
+   \end{array}\right] .
 \end{equation}
@@
 An estimator of the covariance matrix of these estimated parameters is
-\begin{equation}
-   \COVHat \left[ \bHat_{3SLS-Schmidt} \right] =
+\begin{align}
+   \COVHat \left[ \bHat_{3SLS-Schmidt} \right] = &
    \left( \XHat' \OHat^{-1}  \XHat \right)^{-1}
-   \XHat' \OHat^{-1} H \left( H' H \right)^{-1} H' \OHat
-   H \left( H' H \right)^{-1} H' \OHat^{-1} \XHat
+   \XHat' \OHat^{-1} H \left( H' H \right)^{-1} H' \OHat H \\
+   & \left( H' H \right)^{-1} H' \OHat^{-1} \XHat
    \left( \XHat' \OHat^{-1}  \XHat \right)^{-1} .
-\end{equation}
+   \nonumber
+\end{align}
 The econometrics software \proglang{EViews} uses

### 2007-07-24 paper/trunk/systemfit_intro.tex
some small editings and corrections in sections 1-3
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@819 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 parameter estimates.
-This simultaneity bias can be corrected for in each equation by
-applying a °Two-Stage Least Squares° (2SLS) method or for all
-equations simultaneously when combined with SUR resulting in a
-°Three-Stage Least Squares° (3SLS) estimation of the system of
-equations.
+This simultaneity bias can be corrected for by
+applying a °Two-Stage Least Squares° (2SLS) estimation
+to each equation.
+Combining this estimation method with the SUR method results
+in a simultaneous estimation of the system of equations
+by the °Three-Stage Least Squares° (3SLS) method
+\citep{zellner62b}.
 

### 2007-07-27 paper/tags/JSS_20070727/systemfit_other.tex
created tag for revised version submitted to the JSS
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@831 591db2bd-e0a0-4d22-8c19-2192dbb7da31

--- 2007-07-04 paper/trunk/systemfit_other.tex
@@
 
-Since the (true) disturbances of the estimated equations are generally not known,
+Since the (true) disturbances ($u_i$) of the estimated equations
+are generally not known,
 their covariance matrix cannot be determined.
 Therefore, this covariance matrix is generally calculated from
-estimated residuals
+estimated residuals ($\uHat_i$)
 that are obtained from a first-step OLS or 2SLS estimation.
@@
 each single equation,
-because this yields the same $p$ values as the equation-wise
+because this yields the same $P$ values as the equation-wise
 OLS estimation.
@@
 used in the estimation.
-Under the null hypothesis, $F$ has an $F$ distribution
+Under the null hypothesis, $F$ has an approximate $F$ distribution
 with $j$ and $G \cdot T - K$ degrees of freedom
-\citep[p.\ 314]{theil71}.
+\citep[p.~314]{theil71}.
 
@@
 
-The LR statistic for systems of equations is
+The Likelihood-Ratio (LR) statistic for systems of equations is
 \begin{equation}
@@
 \citet{hausman78} developed a test for misspecification.
-The null hypothesis of the test is that all exogenous variables are
-uncorrelated with all disturbance terms.
-Under this hypothesis both the 2SLS and the 3SLS estimator are consistent
-but only the 3SLS estimator is (asymptotically) efficient.
-Under the alternative hypothesis the 2SLS estimator is consistent
-but the 3SLS estimator is inconsistent.
+The null hypothesis of the test is
+that the instrumental variables of each equation are
+uncorrelated with the disturbance terms of all other equations
+($\E \left[ u_i' H_j \right] = 0 \, \forall \, i \neq j$).
+Under this null hypothesis, both the 2SLS and the 3SLS estimator are consistent,
+but the 3SLS estimator is (asymptotically) more efficient.
+Under the alternative hypothesis, the 2SLS estimator is consistent
+but the 3SLS estimator is inconsistent,
+i.e.\ the instrumental variables of each equation are
+uncorrelated with the disturbances of the same equation
+($\E \left[ u_i' H_i \right] = 0 \, \forall \, i$),
+but the instrumental variables of at least one equation
+are correlated with the disturbances of another equation
+($\E \left[ u_i' H_j \right] \neq 0 \, \exists \, i \neq j$).
 The Hausman test statistic is
@@
 where $\bHat_{2SLS}$ and $\COVHat \left[ \bHat_{2SLS} \right]$ are the estimated
-coefficient and covariance matrix from 2SLS estimation, and
+coefficient and covariance matrix from a 2SLS estimation, and
 $\bHat_{3SLS}$ and $\COVHat \left[ \bHat_{3SLS} \right]$ are the estimated
-coefficients and covariance matrix from 3SLS estimation.
-Under the null hypothesis this test statistic has a
+coefficients and covariance matrix from a 3SLS estimation.
+Under the null hypothesis, this test statistic has a
 $\chi^2$ distribution with degrees of freedom equal to the number of

### 2007-07-27 paper/tags/JSS_20070727/runSweave.sh

@@ -1,0 +1,8 @@
+#!/bin/sh
+echo "Sweave(\"systemfit_code.Snw\")" | R --no-save --no-restore
+
+echo "Sweave(\"systemfit_usage.Snw\")" | R --no-save --no-restore
+
+echo "Sweave(\"systemfit_reliability.Snw\")" | R --no-save --no-restore
+
+echo "Sweave(\"systemfit_sem.Snw\")" | R --no-save --no-restore

### 2007-07-27 paper/tags/JSS_20070727/test_gls.R

@@ -1,0 +1,138 @@
+library( MASS )
+set.seed( 20070705 )
+
+nEq <- 6
+nExog <- 10
+nObs <- 2500
+
+xMat   <- NULL
+xMatEq <- list()
+for( eqNo in 1:nEq ){
+   xMatEq[[ eqNo ]] <- matrix( 1, nrow = nObs, ncol = 1 )
+   for( exogNo in 1:nExog ){
+      xMatEq[[ eqNo ]] <- cbind( xMatEq[[ eqNo ]], rnorm( nObs ) )
+   }
+   if( eqNo == 1 ){
+      xMat <- xMatEq[[ eqNo ]]
+   } else {
+      xMat <- rbind(
+         cbind( xMat, matrix( 0, nrow = nrow( xMat ), ncol = nExog + 1 ) ),
+         cbind( matrix( 0, nrow = nObs, ncol = ncol( xMat ) ), xMatEq[[ eqNo ]] ) )
+   }
+}
+myCoef <- c( 10:( 9 + nEq * ( nExog + 1 ) ) ) / 10
+
+sigma <- diag( c( 1:nEq ) ) * 20
+sigma[ upper.tri( sigma ) ] <- sigma[ lower.tri( sigma ) ] <- 12
+
+disturbances <- mvrnorm( nObs, rep( 0, nEq ), sigma )
+
+yVec <- xMat %*% myCoef + c( disturbances )
+
+## OLS
+# Naive
+system.time(
+   olsNaive <- solve( t( xMat ) %*% xMat ) %*% t( xMat ) %*%  yVec
+)
+
+# with solve( , )
+system.time(
+   olsSolve <- solve( t( xMat ) %*% xMat, t( xMat ) %*%  yVec )
+)
+all.equal( olsNaive, olsSolve )
+
+# with crossprod
+system.time(
+   olsCross <- solve( crossprod( xMat ) ) %*% crossprod( xMat, yVec )
+)
+all.equal( olsNaive, olsCross )
+
+#with
+system.time(
+   olsCrossSolve <- solve( crossprod( xMat ), crossprod( xMat, yVec ) )
+)
+all.equal( olsNaive, olsCrossSolve )
+
+
+## residuals
+residVec <- yVec - xMat %*% olsNaive
+residMat <- matrix( residVec, nrow = nObs, ncol = nEq )
+residCov <- crossprod( residMat ) / nObs
+
+
+## SUR ( GLS )
+# Naive
+system.time( {
[…]

### 2007-07-27 tags/JSS_20070727/R/linear.hypothesis.R
created tag for version that was submitted to the JSS with the revised paper
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@476 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@ -1,0 +1,80 @@
+linear.hypothesis.systemfit <- function( model,
+      hypothesis.matrix, rhs = NULL, test = c( "F", "Chisq" ),
+      vcov. = NULL, ... ){
+
+   thisCall <- match.call()
+   test <- match.arg( test )
+
+   if( "model" %in% names( thisCall ) ) {
+      if( class( thisCall$model ) == "name" ) {
+         modelName <- as.character( thisCall$model )
+      } else if( class( thisCall$model ) == "call" ) {
+         modelName <- format( thisCall$model )
+      } else {
+         modelName <- thisCall$model
+      }
+   }
+   if( test == "Chisq" ){
+      result <- car:::linear.hypothesis.default( model,
+         hypothesis.matrix = hypothesis.matrix, rhs = rhs, test = test,
+         vcov. = vcov., ... )
+
+      attributes( result )$heading[ 1 ] <-
+         "Linear hypothesis test (Wald-test)\n\nHypothesis:"
+
+      modelPos <- grep( "^Model 1: .*Model 2:", attributes( result )$heading )
+      attributes( result )$heading[ modelPos[ 1 ] ] <-
+         sub( "^Model 1: .*Model 2:",
+            paste( "Model 1: ", modelName, "\nModel 2:", sep = "" ),
+            attributes( result )$heading[ modelPos[ 1 ] ] )
+
+   } else if ( test == "F" ) {
+      if( is.character( hypothesis.matrix ) ) {
+         R.restr <- car:::makeHypothesis( names( coef( model ) ),
+            hypothesis.matrix, rhs )
+         if( is.null( dim( R.restr ) ) ){
+            R.restr <- t( R.restr )
+         }
+         q.restr <- R.restr[ , ncol( R.restr ), drop = FALSE ]
+         R.restr <- R.restr[ , -ncol( R.restr ), drop = FALSE ]
+         rownames( R.restr ) <- hypothesis.matrix
+      } else {
+         R.restr <- hypothesis.matrix
+         if( is.null( rhs ) ) {
+            q.restr <- rep( 0, nrow( hypothesis.matrix ) )
+         } else {
+            q.restr <- rhs
+         }
+      }
+
+      result <- as.data.frame( matrix( NA, nrow = 2, ncol = 4 ) )
+      names( result ) <- c( "Res.Df", "Df", "F", "Pr(>F)" )
+      rownames( result ) <- c( 1, 2 )
+
[…]

### 2007-07-27 tags/JSS_20070727/R/lrtest.R

@@ -1,0 +1,50 @@
+## Likelihood Ratio Test
+lrtest.systemfit <- function( object, ... ) {
+
+   thisCall <- match.call()
+
+   if( "object" %in% names( thisCall ) ) {
+      if( class( thisCall$object ) == "name" ) {
+         object$name <- as.character( thisCall$object )
+      } else if( class( thisCall$object ) == "call" ) {
+         object$name <- format( thisCall$object )
+      } else {
+         object$name <- thisCall$object
+      }
+   }
+
+   extractName <- function( object ){
+      if( !exists( ".lrtestSystemfitNameNumber" ) ) {
+         .lrtestSystemfitNameNumber <<- 1
+      } else {
+         .lrtestSystemfitNameNumber <<- .lrtestSystemfitNameNumber + 1
+      }
+      objectName <- object$name
+      if( is.null( objectName ) ) {
+         objectName <- paste( "object", .lrtestSystemfitNameNumber, sep = "_" )
+      }
+      return( objectName )
+   }
+
+   result <- lrtest.default( object = object, ..., name = extractName )
+
+   for( i in 2:nrow( result ) ){
+      if( ( result[ i, "#Df" ] - result[ i - 1, "#Df" ] ) *
+            ( result[ i, "LogLik" ] - result[ i - 1, "LogLik" ] ) < 0 ) {
+         if( result[ i, "LogLik" ] > result[ i - 1, "LogLik" ] ) {
+            compareLikelihood <- "larger"
+            compareDf <- "less"
+         } else {
+            compareLikelihood <- "smaller"
+            compareDf <- "more"
+         }
+         warning( "model '", i, "' has a ", compareLikelihood,
+            " log-likelihood value than the ", compareDf,
+            " restricted model '", i - 1, "'" )
+      }
+   }
+
+   rm( .lrtestSystemfitNameNumber, inherits = TRUE )
+
+   return( result )
+}

### 2007-07-27 tags/JSS_20070727/R/smallMethods.R

@@ -1,0 +1,204 @@
+## this function returns a vector of the
+## cross-equation corrlations between eq i and eq j
+## from the results set for equation ij
+correlation.systemfit <- function( results, eqni, eqnj ) {
+  nCoefEq <- NULL
+  for( i in 1:length( results$eq ) ) {
+     nCoefEq <- c( nCoefEq, length( coef( results$eq[[ i ]] ) ) )
+  }
+  cij <- vcov( results )[(1+sum(nCoefEq[1:eqni])-nCoefEq[eqni]):(sum(nCoefEq[1:eqni])),
+                      (1+sum(nCoefEq[1:eqnj])-nCoefEq[eqnj]):(sum(nCoefEq[1:eqnj]))]
+  cii <- vcov( results )[(1+sum(nCoefEq[1:eqni])-nCoefEq[eqni]):(sum(nCoefEq[1:eqni])),
+                      (1+sum(nCoefEq[1:eqni])-nCoefEq[eqni]):(sum(nCoefEq[1:eqni]))]
+  cjj <- vcov( results )[(1+sum(nCoefEq[1:eqnj])-nCoefEq[eqnj]):(sum(nCoefEq[1:eqnj])),
+                      (1+sum(nCoefEq[1:eqnj])-nCoefEq[eqnj]):(sum(nCoefEq[1:eqnj]))]
+  rij <- NULL
+
+  for( i in 1:nrow( residuals( results ) ) ) {
+    xik    <- model.matrix( results$eq[[eqni]] )[i,]
+    xjk    <- model.matrix( results$eq[[eqnj]] )[i,]
+    top    <- xik %*% cij %*% xjk
+    bottom <- sqrt( ( xik %*% cii %*% xik ) * ( xjk %*% cjj %*% xjk ) )
+    rijk   <- top / bottom
+    rij    <- rbind( rij, rijk )
+  }
+  rij
+}
+
+## determines the improvement of resultsj (3sls) over
+## resultsi (2sls) for equation i and returns a matrix
+## of the values, so you can examine the range, mean, etc
+se.ratio.systemfit <- function( resultsi, resultsj, eqni ) {
+  ratio <- NULL
+  for( i in 1:nrow( residuals( resultsi ) ) ) {
+    xik    <- model.matrix( resultsi$eq[[eqni]] )[i,]
+    top    <- sqrt( xik %*% vcov( resultsi$eq[[eqni]] ) %*% xik )
[…]

### 2007-07-27 tags/JSS_20070727/R/summary.systemfit.R

@@ -1,0 +1,311 @@
+## prepare summary results that belong to the whole system
+summary.systemfit <- function( object, useDfSys = NULL,
+      printResidCov = TRUE, printEquations = TRUE, ... ) {
+
+   if( is.null( useDfSys ) ) {
+      useDfSys <- length( coef( object ) ) != object$rank
+         # TRUE if there are restrictions imposed
+   }
+
+   # number of equations
+   nEq <- length( object$eq )
+   # number of observations per equation
+   nObsPerEq <- nrow( residuals( object ) )
+   # total number of observations
+   nObs <- nObsPerEq * nEq
+
+   # preparing objects that will be returned
+   result <- list()
+   result$call <- object$call
+   result$method <- object$method
+   result$iter <- object$iter
+   result$control <- object$control
+   result$residuals <- residuals( object )
+   result$residCovEst <- object$residCovEst
+   result$residCov <- object$residCov
+   if( !is.null( result$residCovEst ) ) {
+      dimnames( result$residCovEst ) <- dimnames( result$residCov )
+   }
+   result$residCor <- cor( residuals( object ) )
+   dimnames( result$residCor ) <- dimnames( result$residCov )
+   result$detResidCov <- det( object$residCov, tol = object$control$solvetol )
+
+   # now prepare summury results for the individual equations
+   result$eq <- list()
+   for( i in 1:length( object$eq ) ) {
+       result$eq[[ i ]] <- summary( object$eq[[i]], useDfSys = useDfSys )
+   }
+
+   # coefficients, standard errors, ... 
+   result$coefCov <- object$coefCov
+   coef <- object$coefficients
+   stdEr <- diag( result$coefCov )^0.5  # standard errors
+   tStat <- coef / stdEr                # t-statistic
+   if( useDfSys ) {             # p-values
+      pVal <- 2 * ( 1 - pt( abs( tStat ), object$df.residual ) )
+   } else {
+      pVal <- NULL
+      for( i in 1:length( object$eq ) ){
+         pVal <- c( pVal, coef( result$eq[[ i ]] )[ , 4 ] )
+      }
+   }
+   result$coefficients <- cbind( coef, stdEr, tStat, pVal )
+   colnames( result$coefficients ) <- c( "Estimate", "Std. Error",
+      "t value", "Pr(>|t|)" )
[…]

### 2007-07-27 tags/JSS_20070727/R/systemfit.R

@@ -1,0 +1,732 @@
+###   $Id$
+###
+###            Simultaneous Equation Estimation for R
+###
+### Copyright 2002-2004 Jeff D. Hamann <jeff.hamann@forestinformatics.com>
+###                     Arne Henningsen <http://www.arne-henningsen.de>
+###
+### This file is part of the nlsystemfit library for R and related languages.
+### It is made available under the terms of the GNU General Public
+### License, version 2, or at your option, any later version,
+### incorporated herein by reference.
+###
+### This program is distributed in the hope that it will be
+### useful, but WITHOUT ANY WARRANTY; without even the implied
+### warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR
+### PURPOSE.  See the GNU General Public License for more
+### details.
+###
+### You should have received a copy of the GNU General Public
+### License along with this program; if not, write to the Free
+### Software Foundation, Inc., 59 Temple Place - Suite 330, Boston,
+### MA 02111-1307, USA
+
+
+systemfit <- function(  eqns,
+                        method = "OLS",
+                        inst=NULL,
+                        data=list(),
+                        restrict.matrix = NULL,
+                        restrict.rhs = NULL,
+                        restrict.regMat = NULL,
+                        pooled = FALSE,
+                        control = systemfit.control( ... ),
+                        ... )
+{
+
+   ## some tests
+   if(!( method %in% c( "OLS", "WLS", "SUR", "2SLS", "W2SLS", "3SLS",
+         "LIML", "FIML" ) ) ){
+      stop( "The method must be 'OLS', 'WLS', 'SUR',",
+         " '2SLS', 'W2SLS', or '3SLS'" )
+   }
+   if( method %in% c( "2SLS", "W2SLS", "3SLS" ) &
+         is.null(inst) ) {
+      stop( "The methods '2SLS', 'W2SLS', and '3SLS' need instruments!" )
+   }
+
+   panelLike <- FALSE
+   if( class( data )[1] == "pdata.frame" ) {
+      panelLike <- TRUE
+      if( !is.null( restrict.regMat ) && pooled ){
+         stop( "argument 'restrict.regMat' cannot be used for pooled estimation",
+            " of panel-like data" )
+      }
[…]

### 2007-07-27 tags/JSS_20070727/R/systemfit.control.R

@@ -1,0 +1,107 @@
+systemfit.control <- function(
+      maxiter = 1,
+      tol = 1e-5,
+      methodResidCov = "geomean",
+      centerResiduals = FALSE,
+      residCovRestricted = TRUE,
+      residCovWeighted = FALSE,
+      method3sls = "GLS",
+      single.eq.sigma = NULL,
+      useMatrix = TRUE,
+      solvetol = .Machine$double.eps,
+      returnModelFrame = TRUE,
+      returnModelMatrix = FALSE,
+      returnInstMatrix = FALSE,
+      returnResponse = FALSE )
+{
+   result <- list()
+
+   ## maxiter
+   if( maxiter <= 0 || round( maxiter ) != maxiter ) {
+      stop( "control parameter 'maxiter' must be a positive integer" )
+   }
+   result$maxiter <- maxiter
+
+   ## tol
+   if( tol <= 0 || !is.numeric( tol ) || length( tol ) != 1 ) {
+      stop( "control parameter 'tol' must be a positive scalar" )
+   }
+   result$tol <- tol
+
+   ## methodResidCov
+   if( !( methodResidCov %in% c( "noDfCor", "geomean", "max", "Theil" ) ) ) {
+      stop( "control parameter 'methodResidCov' must be either",
+         " 'noDfCor', 'geomean', 'max', or 'Theil'" )
+   }
+   result$methodResidCov <- methodResidCov
+
+   ## centerResiduals
+   if( !is.logical( centerResiduals ) || length( centerResiduals ) != 1 ) {
+      stop( "control parameter 'centerResiduals' must be logical" )
+   }
+   result$centerResiduals <- centerResiduals
+
+   ## method3sls
+   if( !( method3sls %in% c( "GLS", "IV", "GMM", "Schmidt", "EViews" ) ) ) {
+      stop( "control parameter 'method3sls' must be either",
+         " 'GLS', 'IV', 'GMM', 'Schmidt', or 'EViews'" )
+   }
+   result$method3sls <- method3sls
+
+   ## single.eq.sigma
+   if( ( !is.logical( single.eq.sigma ) || length( single.eq.sigma ) != 1 )
+         && !is.null( single.eq.sigma ) ) {
+      stop( "control parameter 'single.eq.sigma' must be logical or NULL" )
+   }
+   result$single.eq.sigma <- single.eq.sigma
+
+   ## useMatrix
+   if( !is.logical( useMatrix ) || length( useMatrix ) != 1 ) {
[…]

### 2007-07-27 tags/JSS_20070727/tests/test_2sls.R

@@ -1,0 +1,694 @@
+
+library( systemfit )
+data( "Kmenta" )
+useMatrix <- FALSE
+
+demand <- consump ~ price + income
+supply <- consump ~ price + farmPrice + trend
+inst <- ~ income + farmPrice + trend
+inst1  <- ~ income + farmPrice
+instlist <- list( inst1, inst )
+system <- list( demand = demand, supply = supply )
+restrm <- matrix(0,1,7)  # restriction matrix "R"
+restrm[1,3] <-  1
+restrm[1,7] <- -1
+restrict <- "demand_income - supply_trend = 0"
+restr2m <- matrix(0,2,7)  # restriction matrix "R" 2
+restr2m[1,3] <-  1
+restr2m[1,7] <- -1
+restr2m[2,2] <- -1
+restr2m[2,5] <-  1
+restr2q <- c( 0, 0.5 )  # restriction vector "q" 2
+restrict2 <- c( "demand_income - supply_trend = 0",
+   "- demand_price + supply_price = 0.5" )
+tc <- matrix(0,7,6)
+tc[1,1] <- 1
+tc[2,2] <- 1
+tc[3,3] <- 1
+tc[4,4] <- 1
+tc[5,5] <- 1
+tc[6,6] <- 1
+tc[7,3] <- 1
+restr3m <- matrix(0,1,6)  # restriction matrix "R" 2
+restr3m[1,2] <- -1
+restr3m[1,5] <-  1
+restr3q <- c( 0.5 )  # restriction vector "q" 2
+restrict3 <- "- C2 + C5 = 0.5"
+
+# It is not possible to estimate 2SLS with systemfit exactly
+# as EViews does, because EViews uses
+# methodResidCov == "geomean" for the coefficient covariance matrix and
+# methodResidCov == "noDfCor" for the residual covariance matrix.
+# systemfit uses always the same formulas for both calculations.
+
+## *************** 2SLS estimation ************************
+## ************ 2SLS estimation (default)*********************
+fit2sls1 <- systemfit( system, "2SLS", data = Kmenta, inst = inst,
+   returnModelMatrix = TRUE, useMatrix = useMatrix )
+print( summary( fit2sls1 ) )
+
+## *************** 2SLS estimation (single.eq.sigma=F)*******************
+fit2sls1s <- systemfit( system, "2SLS", data = Kmenta, inst = inst,
+   single.eq.sigma = FALSE, useMatrix = useMatrix )
[…]

### 2007-07-27 tags/JSS_20070727/tests/test_3sls.R

@@ -1,0 +1,1122 @@
+
+library( systemfit )
+data( "Kmenta" )
+useMatrix <- FALSE
+
+demand <- consump ~ price + income
+supply <- consump ~ price + farmPrice + trend
+inst <- ~ income + farmPrice + trend
+inst1  <- ~ income + farmPrice
+instlist <- list( inst1, inst )
+system <- list( demand = demand, supply = supply )
+restrm <- matrix(0,1,7)  # restriction matrix "R"
+restrm[1,3] <-  1
+restrm[1,7] <- -1
+restrict <- "demand_income - supply_trend = 0"
+restr2m <- matrix(0,2,7)  # restriction matrix "R" 2
+restr2m[1,3] <-  1
+restr2m[1,7] <- -1
+restr2m[2,2] <- -1
+restr2m[2,5] <-  1
+restr2q <- c( 0, 0.5 )  # restriction vector "q" 2
+restrict2 <- c( "demand_income - supply_trend = 0",
+   "- demand_price + supply_price = 0.5" )
+tc <- matrix(0,7,6)
+tc[1,1] <- 1
+tc[2,2] <- 1
+tc[3,3] <- 1
+tc[4,4] <- 1
+tc[5,5] <- 1
+tc[6,6] <- 1
+tc[7,3] <- 1
+restr3m <- matrix(0,1,6)  # restriction matrix "R" 2
+restr3m[1,2] <- -1
+restr3m[1,5] <-  1
+restr3q <- c( 0.5 )  # restriction vector "q" 2
+restrict3 <- "- C2 + C5 = 0.5"
+
+
+## *************** 3SLS estimation ************************
+fit3sls <- list()
+formulas <- c( "GLS", "IV", "Schmidt", "GMM", "EViews" )
+for( i in seq( along = formulas ) ) {
+   fit3sls[[ i ]] <- list()
+
+   print( "***************************************************" )
+   print( paste( "3SLS formula:", formulas[ i ] ) )
+   print( "************* 3SLS *********************************" )
+   fit3sls[[ i ]]$e1 <- systemfit( system, "3SLS", data = Kmenta,
+      inst = inst, method3sls = formulas[ i ], useMatrix = useMatrix )
+   print( summary( fit3sls[[ i ]]$e1 ) )
+
+   print( "********************* 3SLS EViews-like *****************" )
+   fit3sls[[ i ]]$e1e <- systemfit( system, "3SLS", data = Kmenta,
[…]

### 2007-07-27 tags/JSS_20070727/tests/test_hausman.R

@@ -1,0 +1,126 @@
+
+library( systemfit )
+data( "Kmenta" )
+useMatrix <- FALSE
+
+eqDemand <- consump ~ price + income
+eqSupply <- consump ~ price + farmPrice + trend
+inst <- ~ income + farmPrice + trend
+eqSystem <- list( demand = eqDemand, supply = eqSupply )
+restrm <- matrix(0,1,7)  # restriction matrix "R"
+restrm[1,3] <-  1
+restrm[1,7] <- -1
+restr2m <- matrix(0,2,7)  # restriction matrix "R" 2
+restr2m[1,3] <-  1
+restr2m[1,7] <- -1
+restr2m[2,2] <- -1
+restr2m[2,5] <-  1
+restr2q <- c( 0, 0.5 )  # restriction vector "q" 2
+tc <- matrix(0,7,6)
+tc[1,1] <- 1
+tc[2,2] <- 1
+tc[3,3] <- 1
+tc[4,4] <- 1
+tc[5,5] <- 1
+tc[6,6] <- 1
+tc[7,3] <- 1
+restr3m <- matrix(0,1,6)  # restriction matrix "R" 2
+restr3m[1,2] <- -1
+restr3m[1,5] <-  1
+restr3q <- c( 0.5 )  # restriction vector "q" 2
+
+
+## ******************* unrestricted estimation *****************
+## ******************** default estimation *********************
+fit2sls1 <- systemfit( eqSystem, "2SLS", data = Kmenta, inst = inst,
+   useMatrix = useMatrix )
+fit3sls1 <- systemfit( eqSystem, "3SLS", data = Kmenta, inst = inst,
+   useMatrix = useMatrix )
+print( hausman.systemfit( fit2sls1, fit3sls1 ) )
+
+## ************** 2SLS estimation with single.eq.sigma = FALSE *****************
+fit2sls1s <- systemfit( eqSystem, "2SLS", data = Kmenta, inst = inst,
+   single.eq.sigma = FALSE, useMatrix = useMatrix )
+print( hausman.systemfit( fit2sls1s, fit3sls1 ) )
+
+## ******************* estimations with methodResidCov = 0 *****************
+fit2sls1r <- systemfit( eqSystem, "2SLS", data = Kmenta, inst = inst,
+   methodResidCov = "noDfCor", useMatrix = useMatrix )
[…]

### 2007-07-27 tags/JSS_20070727/tests/test_ols.R

@@ -1,0 +1,597 @@
+
+library( systemfit )
+data( "Kmenta" )
+useMatrix <- FALSE
+
+demand <- consump ~ price + income
+supply <- consump ~ price + farmPrice + trend
+system <- list( demand = demand, supply = supply )
+restrm <- matrix(0,1,7)  # restriction matrix "R"
+restrm[1,3] <-  1
+restrm[1,7] <- -1
+restrict <- "demand_income - supply_trend = 0"
+restr2m <- matrix(0,2,7)  # restriction matrix "R" 2
+restr2m[1,3] <-  1
+restr2m[1,7] <- -1
+restr2m[2,2] <- -1
+restr2m[2,5] <-  1
+restr2q <- c( 0, 0.5 )    # restriction vector "q" 2
+restrict2 <- c( "demand_income - supply_trend = 0",
+   "- demand_price + supply_price = 0.5" )
+tc <- matrix(0,7,6)
+tc[1,1] <- 1
+tc[2,2] <- 1
+tc[3,3] <- 1
+tc[4,4] <- 1
+tc[5,5] <- 1
+tc[6,6] <- 1
+tc[7,3] <- 1
+restr3m <- matrix(0,1,6)  # restriction matrix "R" 2
+restr3m[1,2] <- -1
+restr3m[1,5] <-  1
+restr3q <- c( 0.5 )       # restriction vector "q" 2
+restrict3 <- "- C2 + C5 = 0.5"
+
+# It is not possible to estimate OLS with systemfit
+# exactly as EViews does, because EViews uses
+# methodResidCov == "geomean" for the coefficient covariance matrix and
+# methodResidCov == "noDfCor" for the residual covariance matrix, while
+# systemfit uses always the same formulas for both calculations.
+
+## *************** OLS estimation ************************
+## ********** OLS estimation (default) ********************
+fitols1 <- systemfit( system, "OLS", data = Kmenta, useMatrix = useMatrix )
+print( summary( fitols1 ) )
+
+## ********** OLS estimation (no single.eq sigma=F) ******************
+fitols1s <- systemfit( system, "OLS", data = Kmenta,
+   single.eq.sigma = FALSE, useMatrix = useMatrix )
+print( summary( fitols1s ) )
+
+## ****************  OLS (useDfSys=T) ***********************
+print( summary( fitols1, useDfSys = TRUE ) )
+
+## ****************  OLS (methodResidCov="noDfCor") ***********************
[…]

### 2007-07-27 tags/JSS_20070727/tests/test_panel.R

@@ -1,0 +1,151 @@
+library( systemfit )
+library( plm )
+useMatrix <- FALSE
+
+## Repeating the OLS and SUR estimations in Theil (1971, pp. 295, 300)
+data( "GrunfeldGreene" )
+GrunfeldTheil <- subset( GrunfeldGreene,
+   firm %in% c( "General Electric", "Westinghouse" ) )
+pdata.frame( GrunfeldTheil, "firm", "year" )
+formulaGrunfeld <- invest ~ value + capital
+
+# OLS
+theilOls <- systemfit( formulaGrunfeld, "OLS",
+   data = GrunfeldTheil, useMatrix = useMatrix )
+print( theilOls )
+summary( theilOls )
+summary( theilOls, useDfSys = TRUE, printResidCov = FALSE,
+   printEquations = FALSE )
+summary( theilOls, printEquations = FALSE )
+coef( theilOls )
+coef( summary(theilOls ) )
+vcov( theilOls )
+residuals( theilOls )
+confint( theilOls )
+fitted(theilOls  )
+logLik( theilOls )
+model.frame( theilOls )
+model.matrix( theilOls )
+formula( theilOls )
+formula( theilOls$eq[[ 1 ]] )
+terms( theilOls )
+terms( theilOls$eq[[ 1 ]] )
+
+# SUR
+theilSur <- systemfit( formulaGrunfeld, "SUR",
+   data = GrunfeldTheil, methodResidCov = "noDfCor", useMatrix = useMatrix )
+print( theilSur )
+summary( theilSur )
+summary( theilSur, useDfSys = TRUE, printEquations = FALSE )
+summary( theilSur, printResidCov = FALSE, printEquations = FALSE )
+coef( theilSur )
+coef( summary( theilSur ) )
+vcov( theilSur )
+residuals( theilSur )
+confint( theilSur )
+fitted( theilSur )
+logLik( theilSur )
+model.frame( theilSur )
+model.matrix( theilSur )
+formula( theilSur )
+formula( theilSur$eq[[ 2 ]] )
+terms( theilSur )
+terms( theilSur$eq[[ 2 ]] )
+
+
+## Repeating the OLS and SUR estimations in Greene (2003, pp. 351)
+pdata.frame( GrunfeldGreene, "firm", "year" )
+formulaGrunfeld <- invest ~ value + capital
+
+# OLS
[…]

### 2007-07-27 tags/JSS_20070727/tests/test_sur.R

@@ -1,0 +1,1171 @@
+
+library( systemfit )
+data( "Kmenta" )
+useMatrix <- FALSE
+
+demand <- consump ~ price + income
+supply <- consump ~ price + farmPrice + trend
+system <- list( demand = demand, supply = supply )
+restrm <- matrix(0,1,7)  # restriction matrix "R"
+restrm[1,3] <-  1
+restrm[1,7] <- -1
+restrict <- "demand_income - supply_trend = 0"
+restr2m <- matrix(0,2,7)  # restriction matrix "R" 2
+restr2m[1,3] <-  1
+restr2m[1,7] <- -1
+restr2m[2,2] <- -1
+restr2m[2,5] <-  1
+restr2q <- c( 0, 0.5 )  # restriction vector "q" 2
+restrict2 <- c( "demand_income - supply_trend = 0",
+   "- demand_price + supply_price = 0.5" )
+restrict2i <- c( "demand_income - supply_trend = 0",
+   "- demand_price + supply_income = 0.5" )
+tc <- matrix(0,7,6)
+tc[1,1] <- 1
+tc[2,2] <- 1
+tc[3,3] <- 1
+tc[4,4] <- 1
+tc[5,5] <- 1
+tc[6,6] <- 1
+tc[7,3] <- 1
+restr3m <- matrix(0,1,6)  # restriction matrix "R" 2
+restr3m[1,2] <- -1
+restr3m[1,5] <-  1
+restr3q <- c( 0.5 )  # restriction vector "q" 2
+restrict3 <- "- C2 + C5 = 0.5"
+
+# the standard equations do not converge and lead to a singular weighting matrix
+# both in R and in EViews, since both equations have the same endogenous variable
+supply2 <- price ~ income + farmPrice + trend
+system2 <- list( demand = demand, supply = supply2 )
+
+
+## *************** SUR estimation ************************
+fitsur1 <- systemfit( system, "SUR", data = Kmenta, useMatrix = useMatrix )
+print( summary( fitsur1 ) )
+
+## ********************* SUR (EViews-like) *****************
+fitsur1e <- systemfit( system, "SUR", data = Kmenta, methodResidCov = "noDfCor",
+   useMatrix = useMatrix )
+print( summary( fitsur1e, useDfSys = TRUE ) )
+
+## ********************* SUR (methodResidCov="Theil") *****************
[…]

### 2007-07-27 tags/JSS_20070727/tests/test_w2sls.R

@@ -1,0 +1,642 @@
+
+library( systemfit )
+data( "Kmenta" )
+useMatrix <- FALSE
+
+demand <- consump ~ price + income
+supply <- consump ~ price + farmPrice + trend
+inst <- ~ income + farmPrice + trend
+inst1  <- ~ income + farmPrice
+instlist <- list( inst1, inst )
+system <- list( demand = demand, supply = supply )
+restrm <- matrix(0,1,7)  # restriction matrix "R"
+restrm[1,3] <-  1
+restrm[1,7] <- -1
+restrict <- "demand_income - supply_trend = 0"
+restr2m <- matrix(0,2,7)  # restriction matrix "R" 2
+restr2m[1,3] <-  1
+restr2m[1,7] <- -1
+restr2m[2,2] <- -1
+restr2m[2,5] <-  1
+restr2q <- c( 0, 0.5 )  # restriction vector "q" 2
+restrict2 <- c( "demand_income - supply_trend = 0",
+   "- demand_price + supply_price = 0.5" )
+tc <- matrix(0,7,6)
+tc[1,1] <- 1
+tc[2,2] <- 1
+tc[3,3] <- 1
+tc[4,4] <- 1
+tc[5,5] <- 1
+tc[6,6] <- 1
+tc[7,3] <- 1
+restr3m <- matrix(0,1,6)  # restriction matrix "R" 2
+restr3m[1,2] <- -1
+restr3m[1,5] <-  1
+restr3q <- c( 0.5 )  # restriction vector "q" 2
+restrict3 <- "- C2 + C5 = 0.5"
+
+
+## ********************* W2SLS *****************
+fitw2sls1 <- systemfit( system, "W2SLS", data = Kmenta, inst = inst,
+   useMatrix = useMatrix )
+print( summary( fitw2sls1 ) )
+
+## ********************* W2SLS (EViews-like) *****************
+fitw2sls1e <- systemfit( system, "W2SLS", data = Kmenta, inst = inst,
+   methodResidCov = "noDfCor", returnModelMatrix = TRUE,
+   useMatrix = useMatrix )
+print( summary( fitw2sls1e, useDfSys = TRUE ) )
+
+## ********************* W2SLS with restriction *******************
+fitw2sls2 <- systemfit( system, "W2SLS", data = Kmenta, restrict.matrix = restrm,
+   inst = inst, useMatrix = useMatrix )
+print( summary( fitw2sls2 ) )
[…]

### 2007-07-27 tags/JSS_20070727/tests/test_wls.R

@@ -1,0 +1,718 @@
+
+library( systemfit )
+data( "Kmenta" )
+useMatrix <- FALSE
+
+demand <- consump ~ price + income
+supply <- consump ~ price + farmPrice + trend
+system <- list( demand = demand, supply = supply )
+restrm <- matrix(0,1,7)  # restriction matrix "R"
+restrm[1,3] <-  1
+restrm[1,7] <- -1
+restrict <- "demand_income - supply_trend = 0"
+restr2m <- matrix(0,2,7)  # restriction matrix "R" 2
+restr2m[1,3] <-  1
+restr2m[1,7] <- -1
+restr2m[2,2] <- -1
+restr2m[2,5] <-  1
+restr2q <- c( 0, 0.5 )  # restriction vector "q" 2
+restrict2 <- c( "demand_income - supply_trend = 0",
+   "- demand_price + supply_price = 0.5" )
+tc <- matrix(0,7,6)
+tc[1,1] <- 1
+tc[2,2] <- 1
+tc[3,3] <- 1
+tc[4,4] <- 1
+tc[5,5] <- 1
+tc[6,6] <- 1
+tc[7,3] <- 1
+restr3m <- matrix(0,1,6)  # restriction matrix "R" 2
+restr3m[1,2] <- -1
+restr3m[1,5] <-  1
+restr3q <- c( 0.5 )  # restriction vector "q" 2
+restrict3 <- "- C2 + C5 = 0.5"
+
+
+## *************** WLS estimation ************************
+fitwls1 <- systemfit( system, "WLS", data = Kmenta, useMatrix = useMatrix )
+print( summary( fitwls1 ) )
+
+## *************** WLS estimation (EViews-like) ************************
+fitwls1e <- systemfit( system, "WLS", data = Kmenta, methodResidCov = "noDfCor",
+   returnModelMatrix = TRUE, useMatrix = useMatrix )
+print( summary( fitwls1e, useDfSys = TRUE ) )
+
+## ************** WLS with cross-equation restriction ***************
+fitwls2 <- systemfit( system, "WLS", data = Kmenta, restrict.matrix = restrm,
+   returnModelMatrix = TRUE, useMatrix = useMatrix )
+print( summary( fitwls2 ) )
+# the same with symbolically specified restrictions
+fitwls2Sym <- systemfit( system, "WLS", data = Kmenta,
+   restrict.matrix = restrict, returnModelMatrix = TRUE,
+   useMatrix = useMatrix )
[…]

### 2007-11-22 paper/trunk/systemfit_estimation.tex
the unifying properties of the estimators are (almost fully) exploited now in subsection "Estimation" (requested by reviewer 2); splitted subsection "Estimation" into "Estimation with only exogenous regressors" (OLS, WLS, SUR) and "Estimation with endogenous regressors" (2SLS, W2SLS, 3SLS); adjusted references to sections and equations in the modified subsections(s)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@832 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Estimation}\label{sec:Estimation}
+\subsection{Estimation with only exogenous regressors}
+\label{sec:Estimation-ols-wls-sur}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+If all regressors are exogenous,
+the system of equations~(\ref{eq:model})
+can be consistently estimated by Ordinary Least Squares (OLS),
+Weighted Least Squares (WLS), and Seemingly Unrelated Regression (SUR).
+These estimators can be obtained by
+\begin{equation}
+   \bHat = \left( X' \OHat^{-1} X \right)^{-1} X' \OHat^{-1} y .
+   \label{eq:ols-wls-sur}
+\end{equation}
+The covariance matrix of these estimators
+can be estimated by
+\begin{equation}
+   \COVHat \left[ \bHat \right] = \left( X' \OHat^{-1} X \right)^{-1} .
+   \label{eq:cov-ols-wls-sur}
+\end{equation}
 
@@
 
-The Ordinary Least Squares (OLS) estimator of the system 
-is obtained by
-\begin{equation}
-   \bHat_{OLS} = \left( X'X \right)^{-1} X'y .
-   \label{eq:ols}
-\end{equation}
-These estimates are efficient only if the disturbance terms are not
-contemporaneously correlated, which means
-$\sigma_{ij} = 0 \; \forall \; i \neq j$.
+The Ordinary Least Squares (OLS) estimator is based on the assumption
+that the disturbance terms are not contemporaneously correlated
+$(\sigma_{ij} = 0 \; \forall \; i \neq j)$
+and have the same variance in each equation
+$( \sigma_i^2 = \sigma_j^2 \, \forall \, i, j$,
+where $\sigma_i^2 = \E [ u_{it}^2 ])$.
+In this case, $\OHat$ in equation~(\ref{eq:ols-wls-sur})
+is equal to $I_{G \cdot T}$ and thus, cancels out.
+The OLS estimator is efficient,
+as long as the disturbances are not contemporaneously correlated.
+
 If the whole system is treated as one single equation,
-an estimator of the covariance matrix of the estimated parameters is
-\begin{equation}
-   \COVHat \left[ \bHat_{OLS} \right] = \sHat^2 \left( X'X \right)^{-1} ,
-   \label{eq:olsCovSameSigma}
-\end{equation}
+$\OHat$ in equation~(\ref{eq:cov-ols-wls-sur}) is $\sHat^2 I_{G \cdot T}$,
 where $\sHat^2$ is the estimated variance of the disturbances.
@@
 same variance $(\E [ u_{it}^2 ] = \sigma^2 \, \forall \, i)$.
-
 If the disturbance terms of the individual equations
 are allowed to have different variances,
-an estimator of the covariance matrix of the estimated parameters is
-\begin{equation}
-   \COVHat \left[ \bHat_{OLS} \right] = \left( X' \OHat^{-1} X \right)^{-1} ,
-   \label{eq:olsCovSingleSigma}
-\end{equation}
-where $\OHat = \SHat \otimes I$,
-$\sHat_{ij} = 0 \; \forall \; i \neq j$ and
+$\OHat$ in equation~(\ref{eq:cov-ols-wls-sur}) is $\SHat \otimes I_T$,
+where $\sHat_{ij} = 0 \; \forall \; i \neq j$ and
 $\sHat_{ii} = \sHat_i^2$ is the estimated variance
@@
 OLS estimation is equal to the covariance matrix obtained
-by equation (\ref{eq:olsCovSingleSigma}).
+by equation~(\ref{eq:cov-ols-wls-sur})
+with $\OHat$ equal to $\SHat \otimes I_T$.
 
@@
 of the disturbance terms in the different equations
-$( \sigma_i^2 \neq \sigma_j^2 \, \forall \, i \neq j$,
-where $\sigma_i^2 = \E [ u_{it}^2 ])$.
-The WLS estimator can be obtained by
-\begin{equation}
-   \bHat_{WLS} = \left( X' \OHat^{-1} X \right)^{-1} X' \OHat^{-1} y ,
-\end{equation}
-where $\OHat = \SHat \otimes I$,
-$\sHat_{ij} = 0 \; \forall \; i \neq j$ and
-$\sHat_{ii} = \sHat_i^2$ is the variance
-of the (estimated) residuals in the $i$th equation.
+$( \sigma_i^2 \neq \sigma_j^2 \, \forall \, i \neq j)$.
+In this case, $\OHat$ in equations~(\ref{eq:ols-wls-sur})
+and~(\ref{eq:cov-ols-wls-sur})
+is $\SHat \otimes I_T$,
+where $\sHat_{ij} = 0 \; \forall \; i \neq j$ and
+$\sHat_{ii} = \sHat_i^2$ is the estimated variance
+of the disturbance terms in the $i$th equation.
 Theoretically, $\sHat_{ii}$ should be the variance of
@@
 
-These estimates are (asymptotically) efficient
+The WLS estimator is (asymptotically) efficient only
 if the disturbance terms are not contemporaneously correlated.
-An estimator of the covariance matrix of the estimated parameters is
-\begin{equation}
-   \COVHat \left[ \bHat_{WLS} \right] = \left( X' \OHat^{-1} X \right)^{-1} .
-\end{equation}
 If no cross-equation parameter restrictions are imposed,
-the parameter estimates are equal to the OLS estimates.
+the parameter estimates are equal to OLS estimates.
 
@@
 
-When the disturbances are contemporaneously correlated, a Generalized 
+If the disturbances are contemporaneously correlated, a Generalized
 Least Squares (GLS) estimation leads to efficient parameter estimates.
@@
 it is only asymptotically efficient.
+In case of an SUR estimator, $\OHat$ in equations~(\ref{eq:ols-wls-sur})
+and~(\ref{eq:cov-ols-wls-sur})
+is $\SHat \otimes I_T$,
+where $\SHat$ is the estimated covariance matrix of the disturbance terms.
+
 It should be noted that while an unbiased OLS or WLS estimation requires only that
@@
 
-The SUR estimator can be obtained by
-\begin{equation}
-   \bHat_{SUR} = \left( X' \OHat^{-1} X \right)^{-1} X' \OHat^{-1} y ,
-   \label{eq:sur}
-\end{equation}
-where $\OHat = \SHat \otimes I$ and
-$\SHat$ is the estimated covariance matrix of the disturbance terms.
-An estimator of the covariance matrix of the estimated parameters is
-\begin{equation}
-   \COVHat \left[ \bHat_{SUR} \right] = \left( X' \OHat^{-1} X \right)^{-1} .
-   \label{eq:surCov}
-\end{equation}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Two-stage least squares (2SLS)}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Estimation with endogenous regressors}
+\label{sec:Estimation-2sls-w2sls-3sls}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 with the disturbances ($\E \left[ u_i' X_i \right] \neq 0$),
-OLS estimates are biased.
-This can be circumvented by an instrumental variable (IV)
-two-stage least squares (2SLS) estimation.
+OLS, WLS, and SUR estimates are biased.
+This can be circumvented by a
+two-stage least squares (2SLS),
+weighted two-stage least squares (W2SLS),
+or a three-stage least squares (3SLS) estimation
+with instrumental variables (IV).
 The instrumental variables for each equation $H_i$ 
@@
 \end{equation}
-At the second stage, the unbiased two-stage least squares estimates
-of $\beta$ are obtained by
+Then, these 'fitted' regressors are used
+to obtain unbiased 2SLS, W2SLS, or 3SLS estimates of $\beta$ by
 \begin{equation}
-   \bHat_{2SLS} = \left( \XHat' \XHat \right)^{-1}
-   \XHat' y .
-   \label{eq:beta2sls}
+   \bHat
+   = \left( \XHat' \OHat^{-1} \XHat \right)^{-1} \XHat' \OHat^{-1} y .
+   \label{eq:2sls-w2sls-3sls}
 \end{equation}
-If the whole system is treated as one single equation,
-an estimator of the covariance matrix of the estimated parameters is
+An estimator of the covariance matrix of the estimated parameters is
 \begin{equation}
-   \COVHat \left[ \bHat_{2SLS} \right] = \sHat^2 \left( \XHat'
-   \XHat \right)^{-1} ,
-   \label{eq:2slsCovSameSigma}
+   \COVHat \left[ \bHat \right]
+   = \left( \XHat' \OHat^{-1} \XHat \right)^{-1} .
+   \label{eq:cov-2sls-w2sls-3sls}
 \end{equation}
-where $\sHat^2$ is the estimated variance of the disturbance terms.
-If the disturbance terms of the individual equations
-are allowed to have different variances, 
-an estimator of the covariance matrix of the estimated parameters is
-\begin{equation}
-   \COVHat \left[ \bHat_{2SLS} \right] = \left( \XHat' \OHat^{-1}
-   \XHat \right)^{-1} ,
-   \label{eq:2slsCovSingleSigma}
-\end{equation}
-where $\OHat = \SHat \otimes I$,
-$\sHat_{ij} = 0 \; \forall \; i \neq j$ and
-$\sHat_{ii} = \sHat_i^2$ is the estimated variance
-of the disturbance term in the $i$th equation.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Two-stage least squares (2SLS)}
+
+The Two-Stage Least Squares (2SLS) estimator
+is based on the same assumptions as the OLS estimator.
+Accordingly, $\OHat$ in equation~(\ref{eq:2sls-w2sls-3sls})
+is equal to $I_{G \cdot T}$ and thus, cancels out.
+Like for the OLS estimator,
+the whole system can be treated either as one single equation
+with $\OHat$ in equation~(\ref{eq:cov-2sls-w2sls-3sls})
+equal to $\sHat^2 I_{G \cdot T}$,
+or the disturbance terms of the individual equations
+are allowed to have different variances
+with $\OHat$ in equation~(\ref{eq:cov-2sls-w2sls-3sls})
+equal to $\SHat \otimes I_T$,
+where $\sHat_{ij} = 0 \; \forall \; i \neq j$ and
+$\sHat_{ii} = \sHat_i^2$.
 
@@
 
-The Weighted Two-Stage Least Squares (W2SLS) estimator of the system 
-is obtained by
-\begin{equation}
-   \bHat_{W2SLS} = \left( \XHat' \OHat^{-1} \XHat
-   \right)^{-1} \XHat' \OHat^{-1} y ,
-\end{equation}
-where $\OHat = \SHat \otimes I$,
-$\sHat_{ij} = 0 \; \forall \; i \neq j$ and
-$\sHat_{ii} = \sHat_i^2$ is the estimated variance
-of the disturbance term in the $i$th equation.
-An estimator of the covariance matrix of the estimated parameters is
-\begin{equation}
-   \COVHat \left[ \bHat_{W2SLS} \right] = \left( \XHat' \OHat^{-1}
-   \XHat \right)^{-1} .
-\end{equation}
+The Weighted Two-Stage Least Squares (W2SLS) estimator allows
+for different variances of the disturbance terms in the different equations.
+Hence, $\OHat$ in equations~(\ref{eq:2sls-w2sls-3sls})
+and~(\ref{eq:cov-2sls-w2sls-3sls})
+is $\SHat \otimes I_T$,
+where $\sHat_{ij} = 0 \; \forall \; i \neq j$ and
+$\sHat_{ii} = \sHat_i^2$.
+If no cross-equation parameter restrictions are imposed,
+the parameter estimates are equal to the 2SLS estimates.
 
@@
 
-If the regressors are correlated with the disturbances 
-($\E \left[ u_i' X_i \right] \neq 0$) and
-the disturbances are contemporaneously correlated, 
+If the disturbances are contemporaneously correlated,
 a Feasible Generalized Least Squares (FGLS) version of the two-stage least squares
@@
 Squares° (3SLS) \citep{zellner62b}.
+The standard 3SLS estimator and its covariance matrix are obtained by
+equations~(\ref{eq:2sls-w2sls-3sls}) and~(\ref{eq:cov-2sls-w2sls-3sls})
+with $\OHat$ equal to $\SHat \otimes I_T$,
+where $\SHat$ is the estimated covariance matrix of the disturbance terms.
 
-The standard 3SLS estimator can be obtained by
-\begin{equation}
-   \bHat_{3SLS} = \left( \XHat' \OHat^{-1} \XHat
-   \right)^{-1} \XHat' \OHat^{-1} y ,
-   \label{eq:3slsGls}
-\end{equation}
-where $\OHat = \SHat \otimes I$ and
-$\SHat$ is the estimated covariance matrix of the disturbance terms.
-An estimator of the covariance matrix of the estimated parameters is
-\begin{equation}
-   \COVHat \left[ \bHat_{3SLS} \right] = \left( \XHat' \OHat^{-1}
-   \XHat \right)^{-1} .
-   \label{eq:cov3sls}
-\end{equation}
 While an unbiased 2SLS or W2SLS estimation requires only that
@@
 where $\bHat_{2SLS}$ is the two-stage least squares estimator
-as defined by (\ref{eq:beta2sls}).
-\proglang{EViews} uses the standard 3SLS formula (\ref{eq:cov3sls}) to
+as defined above.
+\proglang{EViews} uses the standard 3SLS formula~(\ref{eq:cov-2sls-w2sls-3sls}) to
 calculate an estimator of the covariance matrix of the estimated parameters.

### 2007-11-22 paper/trunk/systemfit_other.tex
the unifying properties of the estimators are (almost fully) exploited now in subsection "Estimation" (requested by reviewer 2); splitted subsection "Estimation" into "Estimation with only exogenous regressors" (OLS, WLS, SUR) and "Estimation with endogenous regressors" (2SLS, W2SLS, 3SLS); adjusted references to sections and equations in the modified subsections(s)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@832 591db2bd-e0a0-4d22-8c19-2192dbb7da31

--- 2007-07-04 paper/trunk/systemfit_other.tex
@@
 the covariance matrix of the parameters is calculated
-by~(\ref{eq:olsCovSingleSigma}),
+with $\OHat$ in equation~(\ref{eq:cov-ols-wls-sur})
+equal to $\SHat \otimes I_T$,
 the estimated parameters and their standard errors are identical
@@
 the covariance matrix of an OLS estimation is calculated
-by~(\ref{eq:olsCovSameSigma}),
+with $\OHat$ in equation~(\ref{eq:cov-ols-wls-sur})
+equal to $\sHat^2 I_{G \cdot T}$,
 the system estimation is similar to a single equation estimation.

### 2007-11-22 paper/trunk/systemfit_restrictions.tex
the unifying properties of the estimators are (almost fully) exploited now in subsection "Estimation" (requested by reviewer 2); splitted subsection "Estimation" into "Estimation with only exogenous regressors" (OLS, WLS, SUR) and "Estimation with endogenous regressors" (2SLS, W2SLS, 3SLS); adjusted references to sections and equations in the modified subsections(s)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@832 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 }, 
-and is preferable if equality constraints for coefficients
+but is preferable if equality constraints for coefficients
 across many equations of the system are imposed. 
@@
 in the previous section is done
-(equations~\ref{eq:ols}--\ref{eq:3slsEViews}).
+(equations~\ref{eq:ols-wls-sur}--\ref{eq:3slsEViews}).
 This results in the linear independent parameter estimates $\bHat^M$ and

### 2007-11-22 paper/trunk/systemfit_statistics.tex
added paragraph that (and how) we use the "Matrix" package for the GLS estimations
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@803 591db2bd-e0a0-4d22-8c19-2192dbb7da31
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@817 591db2bd-e0a0-4d22-8c19-2192dbb7da31
revised some of Jeff's modifications
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@818 591db2bd-e0a0-4d22-8c19-2192dbb7da31
some small editings and corrections in sections 1-3
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@819 591db2bd-e0a0-4d22-8c19-2192dbb7da31
the unifying pro

@@
 are available for instance in
-\citet[chapter~7]{theil71}
-\citet[part~4]{judge82}
-\citet[part~5]{judge85}
+\citet[chapter~7]{theil71},
+\citet[part~4]{judge82},
+\citet[part~5]{judge85},
 \citet{srivastava87},
@@
 \citet[chapter~10]{zivot06}.
+
 After introducing notations and assumptions,
@@
    y_{i} = X_i \beta_i + u_i, \quad i = 1, 2, \ldots, G ,
+   \label{eq:model}
 \end{equation}
@@
    \end{array} \right]
+   \label{eq:model-array}
 \end{equation}
@@
    y = X \beta + u .
+   \label{eq:model-matrices}
 \end{equation}
@@
 where $i$ and $j$ indicate the equation number 
-and $t$ and $s$ denote the observation number.
+and $t$ and $s$ denote the observation number,
+where the number of observations is the same for all equations.
 
@@
 \begin{equation}
-   \E \left[ u \, u' \right] = \Omega = \Sigma \otimes I ,
+   \E \left[ u \, u' \right] = \Omega = \Sigma \otimes I_T ,
 \end{equation}
 where $\Sigma = \left[ \sigma_{ij} \right]$ is the (contemporaneous)
-disturbance covariance matrix and $I$ is an identity matrix.
+disturbance covariance matrix,
+$\otimes$ is the Kronecker product,
+$I_T$ is an identity matrix of dimension $T$,
+and $T$ is the number of observations in each equation.
 

### 2007-11-22 paper/trunk/systemfit_restrictions.tex
the unifying properties of the restricted estimators are (almost fully) exploited now in subsection "Imposing linear restrictions" (requested by reviewer 2); adjusted references to sections and equations in the modified subsections(s)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@833 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Restricted OLS estimation}
+\subsubsection{Restricted OLS, WLS, and SUR estimation}
 
-The OLS estimator restricted by $R \beta^R = q$ can be obtained by
-\begin{equation}
-   \left[ \begin{array}{c}
-      \bHat^R_{OLS} \\[0.2em] \lHat
-   \end{array} \right]
-   =
-   \left[ \begin{array}{cc}
-      X' X & R' \\ 
-      R & 0
-   \end{array} \right]^{-1}
-   \cdot
-   \left[ \begin{array}{c}
-      X' y \\ q 
-   \end{array} \right] ,
-\end{equation}
-where $\lambda$ is a vector of the Lagrangean multipliers of the restrictions.
-If the whole system is treated as one single equation,
-an estimator of the covariance matrix of the estimated parameters is
-\begin{equation}
-   \COVHat
-   \left[ \begin{array}{c}
-      \bHat^R_{OLS} \\[0.2em] \lHat
-   \end{array} \right] 
-   = \sHat^2 
-   \left[ \begin{array}{cc}
-      X' X & R' \\ 
-      R & 0
-   \end{array} \right]^{-1} ,
-\end{equation}
-where $\sHat$ is the estimated variance of the disturbance terms.
-If the disturbance terms of the individual equations
-are allowed to have different variances, 
-an estimator of the covariance matrix of the estimated parameters is
-\begin{equation}
-   \COVHat
-   \left[ \begin{array}{c}
-      \bHat^R_{OLS} \\[0.2em] \lHat
-   \end{array} \right] 
-   = 
-   \left[ \begin{array}{cc}
-      X' \OHat^{-1} X & R' \\
-      R & 0
-   \end{array} \right]^{-1} ,
-\end{equation}
-where $\OHat = \SHat \otimes I$,
-$\sHat_{ij} = 0 \; \forall \; i \neq j$ and
-$\sHat_{ii} = \sHat_i^2$ is the estimated variance
-of the disturbance term in the $i$th equation.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Restricted WLS estimation}
-
-The WLS estimator restricted by $R \beta^R = q$ can be obtained by
-\begin{equation}
-   \left[ \begin{array}{c}
-      \bHat^R_{WLS} \\[0.2em] \lHat
-   \end{array} \right]
-   =
-   \left[ \begin{array}{cc}
-      X' \OHat^{-1} X & R' \\
-      R & 0
-   \end{array} \right]^{-1}
-   \cdot
-   \left[ \begin{array}{c}
-      X' \OHat^{-1} y \\ q
-   \end{array} \right] ,
-\end{equation}
-where $\OHat = \SHat \otimes I$,
-$\sHat_{ij} = 0 \; \forall \; i \neq j$ and
-$\sHat_{ii} = \sHat_i^2$ is the estimated variance
-of the disturbance term in the $i$th equation.
-An estimator of the covariance matrix of the estimated parameters is
-\begin{equation}
-   \COVHat
-   \left[ \begin{array}{c}
-      \bHat^R_{WLS} \\[0.2em] \lHat
-   \end{array} \right] 
-   = 
-   \left[ \begin{array}{cc}
-      X' \OHat^{-1} X & R' \\
-      R & 0
-   \end{array} \right]^{-1} .
-\end{equation}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Restricted SUR estimation}
-
-The SUR estimator restricted by $R \beta^R = q$ can be obtained by
+The OLS, WLS, and SUR estimators restricted by $R \beta^R = q$ can be obtained by
 \begin{equation}
@@
    \end{array} \right] ,
+   \label{eq:ols-wls-sur-r}
 \end{equation}
-where $\OHat = \SHat \otimes I$ and
-$\SHat$ is the estimated covariance matrix of the disturbance terms.
+where $\lambda$ is a vector of the Lagrangean multipliers of the restrictions
+and $\OHat$ is defined as in section~\ref{sec:Estimation-ols-wls-sur}.
 An estimator of the covariance matrix of the estimated parameters is
@@
    \end{array} \right]^{-1} .
+   \label{eq:cov-ols-wls-sur-r}
 \end{equation}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Restricted 2SLS estimation}
+\subsubsection{Restricted 2SLS, W2SLS, and 3SLS estimation}
 
-The 2SLS estimator restricted by $R \beta^R = q$ can be obtained by
-\begin{equation}
-   \left[ \begin{array}{c}
-      \bHat^R_{2SLS} \\[0.2em] \lHat
-   \end{array} \right]
-   =
-   \left[ \begin{array}{cc}
-      \XHat' \XHat & R' \\
-      R & 0
-   \end{array} \right]^{-1}
-   \cdot
-   \left[ \begin{array}{c}
-      \XHat' y \\ q
-   \end{array} \right] .
-   \label{eq:beta2SLSr}
-\end{equation}
-If the whole system is treated as one single equation,
-an estimator of the covariance matrix of the estimated parameters is
-\begin{equation}
-   \COVHat
-   \left[ \begin{array}{c}
-      \bHat^R_{2SLS} \\[0.2em] \lHat
-   \end{array} \right] 
-   = \sHat^2 
-   \left[ \begin{array}{cc}
-      \XHat' \XHat & R' \\
-      R & 0
-   \end{array} \right]^{-1} ,
-\end{equation}
-where $\sHat^2$ is the estimated variance of the disturbance terms.
-If the disturbance terms of the individual equations
-are allowed to have different variances, 
-an estimator of the covariance matrix of the estimated parameters is
-\begin{equation}
-   \COVHat
-   \left[ \begin{array}{c}
-      \bHat^R_{2SLS} \\[0.2em] \lHat
-   \end{array} \right] 
-   = 
-   \left[ \begin{array}{cc}
-      \XHat' \OHat^{-1} \XHat & R' \\
-      R & 0
-   \end{array} \right]^{-1} ,
-\end{equation}
-where $\OHat = \SHat \otimes I$,
-$\sHat_{ij} = 0 \; \forall \; i \neq j$ and
-$\sHat_{ii} = \sHat_i^2$ is the estimated variance
-of the disturbance term in the $i$th equation.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Restricted W2SLS estimation}
-
-The W2SLS estimator restricted by $R \beta^R = q$ can be obtained by
-\begin{equation}
-   \left[ \begin{array}{c}
-      \bHat^R_{W2SLS} \\[0.2em] \lHat
-   \end{array} \right]
-   =
-   \left[ \begin{array}{cc}
-      \XHat' \OHat^{-1} \XHat & R' \\
-      R & 0
-   \end{array} \right]^{-1}
-   \cdot
-   \left[ \begin{array}{c}
-      \XHat' \OHat^{-1} y \\ q
-   \end{array} \right] ,
-\end{equation}
-where $\OHat = \SHat \otimes I$,
-$\sHat_{ij} = 0 \; \forall \; i \neq j$ and
-$\sHat_{ii} = \sHat_i^2$ is the estimated variance
-of the disturbance term in the $i$th equation.
-An estimator of the covariance matrix of the estimated parameters is
-\begin{equation}
-   \COVHat
-   \left[ \begin{array}{c}
-      \bHat^R_{W2SLS} \\[0.2em] \lHat
-   \end{array} \right] 
-   = 
-   \left[ \begin{array}{cc}
-      \XHat' \OHat^{-1} \XHat & R' \\
-      R & 0
-   \end{array} \right]^{-1} .
-\end{equation}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Restricted 3SLS estimation}
-
-The standard 3SLS estimator restricted by $R \beta^R = q$ can be obtained by
+The 2SLS, W2SLS, and standard 3SLS estimators
+restricted by $R \beta^R = q$ can be obtained by
 \begin{equation}
@@
    \end{array} \right] ,
-   \label{eq:3slsGlsR}
+   \label{eq:2sls-w2sls-3sls-r}
 \end{equation}
-where $\OHat = \SHat \otimes I$ and
-$\SHat$ is the estimated covariance matrix of the disturbance terms.
+where $\OHat$ is defined as in section~\ref{sec:Estimation-2sls-w2sls-3sls}.
 An estimator of the covariance matrix of the estimated parameters is
@@
    \end{array} \right]^{-1} .
-   \label{eq:cov3slsr}
+   \label{eq:cov-2sls-w2sls-3sls-r}
 \end{equation}
+
 The 3SLS-IV estimator restricted by $R \beta^R = q$ can be obtained by
@@
 where $\bHat^R_{2SLS}$ is the restricted 2SLS estimator calculated
-by equation (\ref{eq:beta2SLSr}). 
+by equation (\ref{eq:2sls-w2sls-3sls-r}).
 \proglang{EViews} uses the standard formula of the restricted 3SLS
-estimator~(\ref{eq:cov3slsr}) to calculate an estimator
+estimator~(\ref{eq:cov-2sls-w2sls-3sls-r}) to calculate an estimator
 for the covariance matrix of the estimated parameters.

### 2007-11-23 paper/trunk/systemfit_intro.tex
modified second (and first) sentence of introduction (requested by reviewer 1)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@834 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 
-Many theoretical models consist of more than one equation.
-If these models are econometrically estimated,
-the disturbance terms of these equations are likely
+Many theoretical models that are econometrically estimated
+consist of more than one equation.
+The disturbance terms of these equations are likely
 to be contemporaneously correlated,
@@
 that influence the disturbance term in one equation
-probably influence the disturbance terms in other equations.
+probably influence the disturbance terms in other equations, too.
 Ignoring this contemporaneous correlation

### 2007-11-23 paper/trunk/systemfit_estimation.tex
added notes that the formulas to calculate 2SLS, W2SLS, and 3SLS estimators and their covariances can be obtained from the formulas to calculate OLS, WLS, and SUR estimators and their covariances by substituting fitted regressors for the original regressors
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@838 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 \end{equation}
-Then, these 'fitted' regressors are used
+Then, these 'fitted' regressors are substituted for the original regressors
+in equation~(\ref{eq:ols-wls-sur})
 to obtain unbiased 2SLS, W2SLS, or 3SLS estimates of $\beta$ by
@@
 \end{equation}
-An estimator of the covariance matrix of the estimated parameters is
+An estimator of the covariance matrix of the estimated parameters
+can be obtained from equation~(\ref{eq:cov-ols-wls-sur}) analogously.
+Hence, we get
 \begin{equation}

### 2007-11-26 paper/trunk/systemfit_compare_lm.tex
revised section about source code of systemfit (mainly changes due to modifications of the systemfit package)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@767 591db2bd-e0a0-4d22-8c19-2192dbb7da31
added a missing full stop
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@812 591db2bd-e0a0-4d22-8c19-2192dbb7da31
Using now typewriter markup for the code elements in table 3 (comparison of elements returned by systemfit and lm); noted this in the response to the referees
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@840 591db2bd-e0a0-4d22-8c19-2192dbb7d

@@
 \section[Object returned by systemfit]{Object returned by \code{systemfit}}
+\label{sec:returned-object}
 
@@
 
-The elements returned by \code{systemfit}
-are similar to those returned by \code{lm},
-the basic tool for linear regressions in \proglang{R}.
-While some counterparts of elements returned by \code{lm}
-can be found directly in objects of class \code{systemfit},
-other counterparts are available for each equation
-in objects of class \code{systemfit.equation}.
-This is demonstrated in table~\ref{tab:compare-lm}
-
 \begin{table}[htbp]
@@
 \setlength{\tabcolsep}{4mm}
+{\ttfamily
 \begin{tabular}{lll}
@@
 \end{tabular}
+}
 \medskip\\
 Note:
-*) if requested by the user (default is \code{TRUE})\\
-**) if requested by the user (default is \code{FALSE})
+\texttt{*)} if requested by the user (default is \code{TRUE})\\
+\texttt{**)} if requested by the user (default is \code{FALSE})
 \end{table}
+
+The elements returned by \code{systemfit}
+are similar to those returned by \code{lm},
+the basic tool for linear regressions in \proglang{R}.
+While some counterparts of elements returned by \code{lm}
+can be found directly in objects of class \code{systemfit},
+other counterparts are available for each equation
+in objects of class \code{systemfit.equation}.
+This is demonstrated in table~\ref{tab:compare-lm}.

### 2007-11-26 trunk/R/linear.hypothesis.R
replaced hyphen in "F-test" and "Wald-test" by a blank (requested by the second reviewer in the second round)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@496 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@ -20,7 +20,7 @@ linear.hypothesis.systemfit <- function( model,
          vcov. = vcov., ... )
 
       attributes( result )$heading[ 1 ] <-
-         "Linear hypothesis test (Wald-test)\n\nHypothesis:"
+         "Linear hypothesis test (Wald test)\n\nHypothesis:"
 
       modelPos <- grep( "^Model 1: .*Model 2:", attributes( result )$heading )
       attributes( result )$heading[ modelPos[ 1 ] ] <-
@@ -60,7 +60,7 @@ linear.hypothesis.systemfit <- function( model,
       result[ 2, 3 ] <- ftest$statistic
       result[ 2, 4 ] <- ftest$p.value
 
-      title <- "Linear hypothesis test (F-test)\n\nHypothesis:"
+      title <- "Linear hypothesis test (F test)\n\nHypothesis:"
       topnote <- paste( "Model 1: ", modelName,
          "\nModel 2: restricted model", sep = "" )
       if( is.null( vcov. ) ){

### 2007-11-26 trunk/R/summary.systemfit.R
we now use printCoefmat() to print the matrix of coefficients, standard errors, t-values, and p-values in print.summary.systemfit() and print.summary.systemfit.equation(); updated test files
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@497 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@ -206,7 +206,7 @@ print.summary.systemfit <- function( x,
       }
    } else {
       cat( "\nCoefficients:\n" )
-      print( coef( x ) ) 
+      printCoefmat( coef( x ), digits = digits )
    }
 
   invisible( x )
@@ -278,20 +278,7 @@ print.summary.systemfit.equation <- function( x,
   }
   cat("\n")
 
-  Signif <- symnum(x$coefficients[,4], corr = FALSE, na = FALSE,
-                   cutpoints = c( 0, 0.001, 0.01, 0.05, 0.1, 1 ),
-                   symbols   = c( "***", "**", "*", "." ," " ))
-
-  table <- cbind(round( x$coefficients,  digits ),
-                 Signif)
-
-  rownames(table) <- rownames(x$coefficients)
-  colnames(table) <- c("Estimate","Std. Error","t value","Pr(>|t|)","")
-
-  ##print.matrix(table, quote = FALSE, right = TRUE )
-  ##prmatrix(table, quote = FALSE, right = TRUE )
-  print(table, quote = FALSE, right = TRUE, digits = digits )
-  cat("---\nSignif. codes: ",attr(Signif,"legend"),"\n")
+  printCoefmat( x$coefficients, digits = digits )
 
   cat(paste("\nResidual standard error:", round( x$sigma, digits ),
             "on", x$df[ 2 ], "degrees of freedom\n" ))

### 2007-11-29 trunk/R/createSystemfitModel.R
renamed the first (and mandatory) argument of systemfit(): eqns -> formula; renamed one element returned by createSystemfitModel(): eqns -> formula
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@499 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@ -25,10 +25,10 @@ createSystemfitModel <- function( nEq, nRegEq, nObs, coef = NULL, sigma = NULL )
 
    disturbances <- mvrnorm( nObs, rep( 0, nEq ), sigma )
    result$data <- data.frame( obsNo = c( 1:nObs ) )
-   result$eqns <- list()
+   result$formula <- list()
 
    for( eqNo in 1:nEq ){
-      result$eqns[[ eqNo ]] <- as.formula( paste( "y.", eqNo, " ~ ",
+      result$formula[[ eqNo ]] <- as.formula( paste( "y.", eqNo, " ~ ",
          paste( paste( "x.", eqNo, ".", sep = "" ),
          c( 1:nRegEq ), sep = "", collapse = " + " ), sep = "" ) )
       for( regNo in 1:nRegEq ){
@@ -36,7 +36,7 @@ createSystemfitModel <- function( nEq, nRegEq, nObs, coef = NULL, sigma = NULL )
             rnorm( nObs )
       }
       result$data[[ paste( "y", eqNo, sep = "." ) ]] <- 0
-      xMatEq <- model.matrix( result$eqns[[ eqNo ]], result$data )
+      xMatEq <- model.matrix( result$formula[[ eqNo ]], result$data )
       result$data[[ paste( "y", eqNo, sep = "." ) ]] <-
          drop( xMatEq %*% coef[ ( ( eqNo - 1 ) * ( nRegEq + 1 ) + 1 ):
             ( eqNo * ( nRegEq + 1 ) ) ] ) + disturbances[ , eqNo ]

### 2007-11-29 trunk/R/systemfit.R

@@ -22,7 +22,7 @@
 ### MA 02111-1307, USA
 
 
-systemfit <- function(  eqns,
+systemfit <- function(  formula,
                         method = "OLS",
                         inst=NULL,
                         data=list(),
@@ -38,23 +38,23 @@ systemfit <- function(  eqns,
    panelLike <- class( data )[1] == "pdata.frame"
 
    ## some tests
-   # argument 'eqns'
+   # argument 'formula'
    if( panelLike ){
-      if( class( eqns ) != "formula" ){
-         stop( "argument 'eqns' must be an object of class 'formula'",
+      if( class( formula ) != "formula" ){
+         stop( "argument 'formula' must be an object of class 'formula'",
             " for panel-like models" )
       }
    } else {
       # single-equation models
-      if( class( eqns ) == "formula" ){
-         eqns <- list( eqns )
+      if( class( formula ) == "formula" ){
+         formula <- list( formula )
       }
-      if( class( eqns ) != "list" ){
-         stop( "argument 'eqns' must be an object of class 'formula'",
+      if( class( formula ) != "list" ){
+         stop( "argument 'formula' must be an object of class 'formula'",
             " or a list of objects of class 'formula'" )
       }
-      if( !all( lapply( eqns, class ) == "formula" ) ){
-         stop( "the list of argument 'eqns' must",
+      if( !all( lapply( formula, class ) == "formula" ) ){
+         stop( "the list of argument 'formula' must",
             " contain only objects of class 'formula'" )
       }
    }
@@ -76,10 +76,10 @@ systemfit <- function(  eqns,
          stop( "argument 'restrict.regMat' cannot be used for pooled estimation",
             " of panel-like data" )
       }
-      result <- .systemfitPanel( formula = eqns,
+      result <- .systemfitPanel( formula = formula,
          data = data, pooled = pooled )
       data <- result$wideData
-      eqns <- result$eqnSystem
+      formula <- result$eqnSystem
       if( pooled ){
          restrict.regMat <- result$restrict.regMat
       }
@@ -93,14 +93,14 @@ systemfit <- function(  eqns,
   results <- list()               # results to be returned
   results$eq <- list()            # results for the individual equations
   iter    <- NULL                 # number of iterations
[…]

### 2007-11-29 trunk/R/systemfit.R
added several tests for argument 'inst'
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@500 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@ -65,9 +65,34 @@ systemfit <- function(  formula,
       stop( "The method must be 'OLS', 'WLS', 'SUR',",
          " '2SLS', 'W2SLS', or '3SLS'" )
    }
-   if( method %in% c( "2SLS", "W2SLS", "3SLS" ) &
-         is.null(inst) ) {
-      stop( "The methods '2SLS', 'W2SLS', and '3SLS' need instruments!" )
+
+   # argument 'inst'
+   if( method %in% c( "2SLS", "W2SLS", "3SLS" ) ){
+      if( is.null( inst ) ) {
+         stop( "The methods '2SLS', 'W2SLS', and '3SLS' need instruments" )
+      } else if( class( inst ) == "formula" ){
+         if( length( inst ) != 2 ){
+            stop( "argument 'inst' must be a one-sided formula" )
+         }
+      } else if( class( inst ) == "list" ){
+         if( !all( lapply( inst, class ) == "formula" ) ){
+            stop( "the list of argument 'inst' must",
+               " contain only objects of class 'formula'" )
+         }
+         if( !all( lapply( inst, length ) == 2 ) ){
+            stop( "the list of argument 'inst' must",
+               " contain only one-sided formulas" )
+         }
+      } else {
+         stop( "argument 'inst' must be an object of class 'formula'",
+            " or a list of objects of class 'formula'" )
+      }
+   } else {
+      if( !is.null( inst ) ) {
+         warning( "The methods 'OLS', 'WLS', and 'SUR' do not need instruments;",
+            " ignoring argument 'inst'" )
+         inst <- NULL
+      }
    }
 
    # prepare model and data for panel-like models

### 2007-11-29 paper/trunk/runSweave.sh
moved R examples illustrating the different computation times to the appendix (requested by the second referee); noted this in the response to the referees
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@857 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@ -5,4 +5,6 @@ echo "Sweave(\"systemfit_usage.Snw\")" | R --no-save --no-restore
 
 echo "Sweave(\"systemfit_reliability.Snw\")" | R --no-save --no-restore
 
+echo "Sweave(\"systemfit_timings.Snw\")" | R --no-save --no-restore
+
 echo "Sweave(\"systemfit_sem.Snw\")" | R --no-save --no-restore

### 2007-11-29 trunk/R/systemfit.R
added some further tests for argument 'inst'
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@505 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@ -90,6 +90,17 @@ systemfit <- function(  formula,
          }
          inst <- lapply( c( 1:length( formula ) ), function(x) inst )
       } else if( class( inst ) == "list" ){
+         if( length( inst ) != length( formula ) ){
+            stop( "if different instruments are specified for each equation,",
+               " the length of argument 'inst' must be equal to the number",
+               " of equations" )
+         }
+         if( !is.null( names( inst ) ) && !is.null( names( formula ) ) ){
+            if( names( inst ) != names( formula ) ){
+               warning( "names of formulas for instruments (argument 'inst')",
+                  " are not equal to names of equations (argument 'formula')" )
+            }
+         }
          if( !all( lapply( inst, class ) == "formula" ) ){
             stop( "the list of argument 'inst' must",
                " contain only objects of class 'formula'" )

### 2007-11-30 paper/trunk/systemfit_estimation.tex
slightly improved description of the different estimation methods (mainly streamlined notation of the different Omegas, which was suggested by the second referee); noted this in the response to the referees
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@863 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 and have the same variance in each equation
-$( \sigma_i^2 = \sigma_j^2 \, \forall \, i, j$,
-where $\sigma_i^2 = \E [ u_{it}^2 ])$.
+$( \sigma_i^2 = \sigma_j^2 \, \forall \, i, j)$.
 In this case, $\OHat$ in equation~(\ref{eq:ols-wls-sur})
@@
 $\OHat$ in equation~(\ref{eq:cov-ols-wls-sur}) is $\sHat^2 I_{G \cdot T}$,
-where $\sHat^2$ is the estimated variance of the disturbances.
-This assumes that the disturbances of all equations have the
-same variance $(\E [ u_{it}^2 ] = \sigma^2 \, \forall \, i)$.
+where $\sHat^2$ is an estimator for the variance of all disturbances
+$(\sigma^2 = \E [ u_{it}^2 ])$.
 If the disturbance terms of the individual equations
@@
 of the disturbance terms in the different equations
-$( \sigma_i^2 \neq \sigma_j^2 \, \forall \, i \neq j)$.
+$( \sigma_i^2 \neq \sigma_j^2 \, \forall \, i \neq j)$,
+but assumes
+that the disturbance terms are not contemporaneously correlated.
 In this case, $\OHat$ in equations~(\ref{eq:ols-wls-sur})
@@
 it is only asymptotically efficient.
-In case of an SUR estimator, $\OHat$ in equations~(\ref{eq:ols-wls-sur})
+In case of a SUR estimator, $\OHat$ in equations~(\ref{eq:ols-wls-sur})
 and~(\ref{eq:cov-ols-wls-sur})
@@
 The Two-Stage Least Squares (2SLS) estimator
-is based on the same assumptions as the OLS estimator.
+is based on the same assumptions about the disturbance terms as the OLS estimator.
 Accordingly, $\OHat$ in equation~(\ref{eq:2sls-w2sls-3sls})

### 2007-11-30 paper/trunk/systemfit_estimation.tex
The coefficients of the model (beta) are consistently called 'coefficients' now (and no longer 'parameters' every now and then); noted this in the response to the referees
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@864 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 
-If no cross-equation parameter restrictions are imposed, the simultaneous 
-OLS estimation of the system leads to the same parameter estimates 
+If no cross-equation restrictions on the coefficients are imposed, the simultaneous
+OLS estimation of the system leads to the same estimated coefficients
 as an equation-wise OLS estimation.
-The covariance matrix of the parameters from an equation-wise
+The covariance matrix of the coefficients from an equation-wise
 OLS estimation is equal to the covariance matrix obtained
@@
 if the disturbance terms are not contemporaneously correlated.
-If no cross-equation parameter restrictions are imposed,
-the parameter estimates are equal to OLS estimates.
+If no cross-equation restrictions on the coefficients are imposed,
+the estimated coefficients are equal to OLS estimates.
 
@@
 If the disturbances are contemporaneously correlated, a Generalized
-Least Squares (GLS) estimation leads to efficient parameter estimates.
+Least Squares (GLS) estimation leads to an efficient estimator for the coefficients.
 In this case, the GLS estimator is generally called °Seemingly Unrelated Regression°
@@
 \end{equation}
-An estimator of the covariance matrix of the estimated parameters
+An estimator of the covariance matrix of the estimated coefficients
 can be obtained from equation~(\ref{eq:cov-ols-wls-sur}) analogously.
@@
 $\sHat_{ii} = \sHat_i^2$.
-If no cross-equation parameter restrictions are imposed,
-the parameter estimates are equal to the 2SLS estimates.
+If no cross-equation restrictions on the coefficients are imposed,
+the estimated coefficients are equal to 2SLS estimates.
 
@@
 \end{equation}
-An estimator of the covariance matrix of the estimated 3SLS-IV parameters is
+An estimator of the covariance matrix of the estimated 3SLS-IV coefficients is
 \begin{equation}
@@
 \end{equation}
-An estimator of the covariance matrix of the estimated 3SLS-GMM parameters is
+An estimator of the covariance matrix of the estimated 3SLS-GMM coefficients is
 \begin{equation}
@@
 \end{equation}
-An estimator of the covariance matrix of these estimated parameters is
+An estimator of the covariance matrix of these estimated coefficients is
 \begin{align}
@@
 \proglang{EViews} uses the standard 3SLS formula~(\ref{eq:cov-2sls-w2sls-3sls}) to
-calculate an estimator of the covariance matrix of the estimated parameters.
+calculate an estimator of the covariance matrix of the estimated coefficients.
 
@@
 ($H_1 = H_2 = \ldots = H_G$), 
-all the above mentioned approaches lead to identical parameter estimates.
+all the above mentioned approaches lead to identical estimates.
 However, if this is not the case, the results depend on the 

### 2007-11-30 paper/trunk/systemfit_intro.tex
The coefficients of the model (beta) are consistently called 'coefficients' now (and no longer 'parameters' every now and then); noted this in the response to the referees
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@864 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 and estimating these equations separately
-leads to inefficient parameter estimates.
+leads to inefficient estimates of the coefficients.
 However, estimating all equations simultaneously
@@
 Another reason to estimate a system of equations simultaneously are
-cross-equation parameter restrictions.%
+cross-equation restrictions on the coefficients.%
 \footnote{
-Especially the economic theory suggests many cross-equation parameter
-restrictions (e.g.\ the symmetry restriction in demand models).
+Especially the economic theory suggests many cross-equation
+restrictions on the coefficients (e.g.\ the symmetry restriction in demand models).
 }
@@
 Ignoring the endogeneity of these variables can lead to inconsistent
-parameter estimates.
+estimates.
 This simultaneity bias can be corrected for by
@@
 the \pkg{systemfit} package
-provides statistical tests for parameter restrictions
+provides statistical tests for restrictions on the coefficients
 and for testing the consistency of the 3SLS estimation.

### 2007-11-30 paper/trunk/systemfit_other.tex
The coefficients of the model (beta) are consistently called 'coefficients' now (and no longer 'parameters' every now and then); noted this in the response to the referees
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@864 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 
-The WLS, SUR, W2SLS and 3SLS parameter estimates are consistent
+The WLS, SUR, W2SLS and 3SLS coefficient estimates are consistent
 if the residual covariance matrix is calculated
@@
 If no cross-equation restrictions are imposed,
-the parameter estimates of OLS and WLS as well as 2SLS and W2SLS are identical.
+the coefficient estimates of OLS and WLS as well as 2SLS and W2SLS are identical.
 Hence, in this case both approaches generate the same results.
@@
 
-If any restrictions are imposed on the parameters,
+If any restrictions are imposed on the coefficients,
 the residual covariance matrix
@@
 To our knowledge the question about how to determine the degrees
-of freedom for single-parameter $t$ tests is not comprehensively
+of freedom for single-coefficient $t$ tests is not comprehensively
 discussed in the literature.
@@
 (total number of observations in all equations minus
-total number of estimated parameters)
+total number of estimated coefficients)
 are applied,
@@
 (number of observations in the equations minus
-number of estimated parameters in the equation)
+number of estimated coefficients in the equation)
 are used.
@@
 If a system of equations is estimated by an unrestricted OLS and
-the covariance matrix of the parameters is calculated
+the covariance matrix of the coefficients is calculated
 with $\OHat$ in equation~(\ref{eq:cov-ols-wls-sur})
 equal to $\SHat \otimes I_T$,
-the estimated parameters and their standard errors are identical
+the estimated coefficients and their standard errors are identical
 to an equation-wise OLS estimation.
@@
 $\chi^2$ distribution with degrees of freedom equal to the number of
-estimated parameters.
+estimated coefficients.
 

### 2007-11-30 paper/trunk/systemfit_restrictions.tex
The coefficients of the model (beta) are consistently called 'coefficients' now (and no longer 'parameters' every now and then); noted this in the response to the referees
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@864 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 In many empirical applications,
-it is desirable to impose restrictions on the parameters
+it is desirable to impose restrictions on the coefficients
 that are derived from the underlying theoretical model.%
@@
 }
-There are two methods to impose linear parameter restrictions.
+There are two methods to impose linear restrictions on the coefficients.
 First, a matrix $M$ can be specified that
@@
 
-The second method to impose linear parameter restrictions
+The second method to impose linear restrictions on the coefficients
 can be formulated by
@@
 In this case, the restrictions imposed using the latter method are
-imposed on the linear independent parameters
+imposed on the linear independent coefficients
 that are restricted by the first method, so that
@@
 (equations~\ref{eq:ols-wls-sur}--\ref{eq:3slsEViews}).
-This results in the linear independent parameter estimates $\bHat^M$ and
+This results in the linear independent coefficient estimates $\bHat^M$ and
 their covariance matrix.
-The original parameters can be obtained by equation (\ref{eq:T-restr})
-and the estimated covariance matrix of the original parameters
+The original coefficients can be obtained by equation (\ref{eq:T-restr})
+and the estimated covariance matrix of the original coefficients
 can be obtained by
@@
 and $\OHat$ is defined as in section~\ref{sec:Estimation-ols-wls-sur}.
-An estimator of the covariance matrix of the estimated parameters is
+An estimator of the covariance matrix of the estimated coefficients is
 \begin{equation}
@@
 where $\OHat$ is defined as in section~\ref{sec:Estimation-2sls-w2sls-3sls}.
-An estimator of the covariance matrix of the estimated parameters is
+An estimator of the covariance matrix of the estimated coefficients is
 \begin{equation}
@@
 estimator~(\ref{eq:cov-2sls-w2sls-3sls-r}) to calculate an estimator
-for the covariance matrix of the estimated parameters.
+for the covariance matrix of the estimated coefficients.
 
@@
 ($H_1 = H_2 = \ldots = H_G$), 
-all the above mentioned approaches lead to identical parameter estimates
-and identical covariance matrices of the estimated parameters.
+all the above mentioned approaches lead to identical coefficient estimates
+and identical covariance matrices of the estimated coefficients.
 

### 2007-11-30 paper/trunk/systemfit_statistics.tex
corrected "background on that" -> "background that" (pointed out by the second referee); noted this in the response to the referees
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@861 591db2bd-e0a0-4d22-8c19-2192dbb7da31
The coefficients of the model (beta) are consistently called 'coefficients' now (and no longer 'parameters' every now and then); noted this in the response to the referees
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@864 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 In this section we give a short overview of the statistical background
-on that the \pkg{systemfit} package is based on.
+that the \pkg{systemfit} package is based on.
 More detailed descriptions of simultaneous equations systems
@@
 we provide the formulas to estimate systems of linear equations.
-We then demonstrate how to impose linear restrictions on parameters.
+We then demonstrate how to impose linear restrictions on the coefficients.
 Finally, we present additional relevant issues about estimation of

### 2007-11-30 paper/trunk/systemfit_estimation.tex
clarified notations by avoiding the ambiguous word 'impose'; instead, we now clearly describe whether we mean 'estimating coefficients under restrictions' or 'testing restrictions on the coefficients'; noted this in the response to the referees
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@866 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 
-If no cross-equation restrictions on the coefficients are imposed, the simultaneous
+If the estimated coefficients are not constrained by cross-equation restrictions,
+the simultaneous
 OLS estimation of the system leads to the same estimated coefficients
@@
 if the disturbance terms are not contemporaneously correlated.
-If no cross-equation restrictions on the coefficients are imposed,
-the estimated coefficients are equal to OLS estimates.
+If the estimated coefficients are not constrained by cross-equation restrictions,
+they are equal to OLS estimates.
 
@@
 $\sHat_{ii} = \sHat_i^2$.
-If no cross-equation restrictions on the coefficients are imposed,
-the estimated coefficients are equal to 2SLS estimates.
+If the estimated coefficients are not constrained by cross-equation restrictions,
+they are equal to 2SLS estimates.
 

### 2007-11-30 paper/trunk/systemfit_other.tex
clarified notations by avoiding the ambiguous word 'impose'; instead, we now clearly describe whether we mean 'estimating coefficients under restrictions' or 'testing restrictions on the coefficients'; noted this in the response to the referees
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@866 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 to estimate the model by SUR or 3SLS.
-If no cross-equation restrictions are imposed,
-the coefficient estimates of OLS and WLS as well as 2SLS and W2SLS are identical.
+If the estimated coefficients are not constrained by cross-equation restrictions,
+OLS and WLS estimates as well as 2SLS and W2SLS estimates are identical.
 Hence, in this case both approaches generate the same results.
@@
 
-If any restrictions are imposed on the coefficients,
+If the coefficients are estimated under any restrictions,
 the residual covariance matrix

### 2007-11-30 paper/trunk/systemfit_restrictions.tex
clarified notations by avoiding the ambiguous word 'impose'; instead, we now clearly describe whether we mean 'estimating coefficients under restrictions' or 'testing restrictions on the coefficients'; noted this in the response to the referees
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@866 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Imposing linear restrictions}\label{sec:Restrictions}
+\subsection{Estimation under linear restrictions on the coefficients}
+\label{sec:Restrictions}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 In many empirical applications,
-it is desirable to impose restrictions on the coefficients
-that are derived from the underlying theoretical model.%
-\footnote{%
-For instance in econometric demand and production analysis,
-it is common to impose (and test) homogeneity and symmetry properties.
-}
-There are two methods to impose linear restrictions on the coefficients.
+it is desirable to estimate the coefficients under linear restrictions.
+For instance, in econometric demand and production analysis,
+it is common to estimate the coefficients
+under homogeneity and symmetry restrictions
+that are derived from the underlying theoretical model.
+
+There are two different methods to estimate the coefficients
+under linear restrictions.
 First, a matrix $M$ can be specified that
@@
 
-The second method to impose linear restrictions on the coefficients
-can be formulated by
+The second method to estimate the coefficients under linear restrictions
+constrains the coefficients by
 \begin{equation}
@@
 and $R$ and $q$ are a matrix and vector, respectively,
-to impose the restrictions \citep[see][p.\ 100]{greene03}.
+that specify the restrictions \citep[see][p.~100]{greene03}.
 Each linear independent restriction is represented by one row of $R$
@@
 \footnote{
-While restrictions like $\beta_1 = 2 \beta_2$ can be imposed by
+While restrictions like $\beta_1 = 2 \beta_2$ can be specified by
 both methods,
-restrictions like $\beta_1 + \beta_2 = 4$ can be imposed only
+restrictions like $\beta_1 + \beta_2 = 4$ can be specified only
 by the second method.
 }, 
-but is preferable if equality constraints for coefficients
-across many equations of the system are imposed. 
-Of course, these restrictions can be also imposed using
+but is preferable if the coefficients are estimated
+under many equality constraints across different equations of the system.
+Of course, these restrictions can be also specified using
 the latter method.
@@
 These two methods can be combined.
-In this case, the restrictions imposed using the latter method are
+In this case, the restrictions specified using the latter method are
 imposed on the linear independent coefficients
@@
 If the first method (equation~\ref{eq:T-restr}) is chosen
-to impose these restrictions,
+to estimate the coefficients under these restrictions,
 the matrix of regressors $X$ is (post-)\hspace{0pt}multiplied
@@
 
-The implementation of the second method to impose linear restrictions
-(equation~\ref{eq:restr-R})
+The implementation of the second method to estimate the coefficients
+under linear restrictions (equation~\ref{eq:restr-R})
 is described for each estimation method in the following sections.

### 2007-12-03 paper/trunk/systemfit_compare_lm.tex
a minor improvement and a minor correction in the appendix comparing the objects returned by systemfit and lm
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@868 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 \code{systemfit} returns a list of class \code{systemfit}
-that contains all results that belong to the entire system
+that contains the results that belong to the entire system
 of equations.
-One special element of this list called \code{eq}.
+One special element of this list is called \code{eq}.
 It is a list that contains one object for each estimated equation.

### 2007-12-05 paper/trunk/systemfit_compare_lm.tex
updated paper, because arguments 'returnModelFrame', 'returnModelMatrix', 'returnResponse', and 'returnInstMatrix' of 'systemfit() have been renamed as 'model', 'x', 'y', and 'z', respectively, and elements 'modelFrame', 'modelMatrix', 'response', and 'instMatrix' returned by 'systemfit()' have been renamed as 'model', 'x', 'y', and 'z', respectively (revision 506--513 of the systemfit package); noted this in the response to the referees
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@874 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 offset          &                & \\
-model*          &                & modelFrame* \\
-y**             &                & response** \\
-x**             &                & modelMatrix** \\
-                &                & instMatrix** \\
+model*          &                & model* \\
+y**             &                & y** \\
+x**             &                & x** \\
+                &                & z** \\
                 & iter           & \\

### 2007-12-05 paper/trunk/systemfit_compare_lm.tex
changed order of elements returned by systemfit() in the table comparing the returned elements of systemfit() and lm(): 'model', 'x', 'y', and 'z' are in alphabetical order now
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@876 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 model*          &                & model* \\
+x**             &                & x** \\
 y**             &                & y** \\
-x**             &                & x** \\
                 &                & z** \\

### 2007-12-05 trunk/R/systemfit.R
renamed internal variables for instrumental variables: hMatEq -> zMatEq, hMatAll -> zMatAll, HtOmega -> ZtOmega
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@516 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@ -256,7 +256,7 @@ systemfit <- function(  formula,
       # evaluated model frame of instruments
       evalModelFrameInst <- list()
       # list for matrices of instruments in each equation
-      hMatEq  <- list()
+      zMatEq  <- list()
       # fitted values of regressors
       xMatHatEq <- list()
       # prepare data for individual equations
@@ -267,11 +267,11 @@ systemfit <- function(  formula,
          modelFrameInst[[ i ]]$formula <- inst[[ i ]]
          evalModelFrameInst[[ i ]] <- eval( modelFrameInst[[ i ]] )
          termsInst[[ i ]] <- attr( evalModelFrameInst[[ i ]], "terms" )
-         hMatEq[[i]] <- model.matrix( termsInst[[ i ]], evalModelFrameInst[[ i ]] )
+         zMatEq[[i]] <- model.matrix( termsInst[[ i ]], evalModelFrameInst[[ i ]] )
          if( control$useMatrix ){
-            hMatEq[[ i ]] <- as( hMatEq[[ i ]], "dgeMatrix" )
+            zMatEq[[ i ]] <- as( zMatEq[[ i ]], "dgeMatrix" )
          }
-         if( nrow( hMatEq[[ i ]] ) != nrow( xMatAll[ rowsEq, ] ) ) {
+         if( nrow( zMatEq[[ i ]] ) != nrow( xMatAll[ rowsEq, ] ) ) {
             stop( paste( "The instruments and the regressors of equation",
                as.character( i ), "have different numbers of observations." ) )
          }
@@ -280,12 +280,12 @@ systemfit <- function(  formula,
          if( control$useMatrix ){
             xMatAllThisEq <- as( xMatAllThisEq, "dgeMatrix" )
          }
-         xMatHatEq[[ i ]] <- hMatEq[[i]] %*% 
-            solve( crossprod( hMatEq[[i]] ),
-            crossprod( hMatEq[[i]], xMatAllThisEq ), tol=control$solvetol )
+         xMatHatEq[[ i ]] <- zMatEq[[i]] %*%
+            solve( crossprod( zMatEq[[i]] ),
+            crossprod( zMatEq[[i]], xMatAllThisEq ), tol=control$solvetol )
       }
       # stacked matrices of all instruments
-      hMatAll <- .stackMatList( hMatEq, way = "diag",
+      zMatAll <- .stackMatList( zMatEq, way = "diag",
[…]

### 2007-12-05 paper/trunk/systemfit_estimation.tex
renamed matrices for instrumental variables: H_i -> Z_i, H -> Z
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@877 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 with instrumental variables (IV).
-The instrumental variables for each equation $H_i$ 
+The instrumental variables for each equation $Z_i$ 
 can be either different or identical for all equations.
@@
 the disturbance terms of the corresponding equation 
-($\E \left[ u_i' H_i \right] = 0$).
+($\E \left[ u_i' Z_i \right] = 0$).
 
@@
 \begin{equation}
-   \XHat_i = H_i \left( H_i' H_i \right)^{-1} H_i' X .
+   \XHat_i = Z_i \left( Z_i' Z_i \right)^{-1} Z_i' X .
 \end{equation}
@@
 equation are uncorrelated
-$( \E \left[ u_i' H_i \right]) = 0 \; \forall \; i )$,
+$( \E \left[ u_i' Z_i \right]) = 0 \; \forall \; i )$,
 \cite{schmidt90} points out that this estimator is only consistent 
 if all disturbance terms and all instrumental variables are uncorrelated
-$( \E \left[ u_i' H_j \right]) = 0 \; \forall \; i, j )$.
+$( \E \left[ u_i' Z_j \right]) = 0 \; \forall \; i, j )$.
 
@@
 \begin{equation}
-   \bHat_{3SLS-GMM} = \left( X' H \left( H' \OHat H \right)^{-1}
-   H' X \right)^{-1} X' H \left( H' \OHat H \right)^{-1} H' y
+   \bHat_{3SLS-GMM} = \left( X' Z \left( Z' \OHat Z \right)^{-1}
+   Z' X \right)^{-1} X' Z \left( Z' \OHat Z \right)^{-1} Z' y
    \label{eq:3slsGmm}
@@
 \begin{equation}
-   H =
+   Z =
    \left[ \begin{array}{cccc}
-      H_1 & 0 & \cdots & 0\\
-      0 & H_2 & \cdots & 0\\
+      Z_1 & 0 & \cdots & 0\\
+      0 & Z_2 & \cdots & 0\\
       \vdots & \vdots & \ddots & \vdots\\
-      0 & 0 & \cdots & H_G
+      0 & 0 & \cdots & Z_G
    \end{array}\right] .
@@
    \COVHat \left[ \bHat_{3SLS-GMM} \right] =
-   \left( X' H \left( H' \OHat H \right)^{-1} H' X \right)^{-1} .
+   \left( X' Z \left( Z' \OHat Z \right)^{-1} Z' X \right)^{-1} .
 \end{equation}
@@
    \right)^{-1} \XHat' \OHat^{-1}
-   H \left( H' H \right)^{-1} H' y .
+   Z \left( Z' Z \right)^{-1} Z' y .
    \label{eq:3slsSchmidt}
@@
    \left( \XHat' \OHat^{-1}  \XHat \right)^{-1}
-   \XHat' \OHat^{-1} H \left( H' H \right)^{-1} H' \OHat H \\
-   & \left( H' H \right)^{-1} H' \OHat^{-1} \XHat
+   \XHat' \OHat^{-1} Z \left( Z' Z \right)^{-1} Z' \OHat Z \\
+   & \left( Z' Z \right)^{-1} Z' \OHat^{-1} \XHat
    \left( \XHat' \OHat^{-1}  \XHat \right)^{-1} .
@@
 If the same instrumental variables are used in all equations 
-($H_1 = H_2 = \ldots = H_G$), 
+($Z_1 = Z_2 = \ldots = Z_G$),
 all the above mentioned approaches lead to identical estimates.

### 2007-12-05 paper/trunk/systemfit_other.tex
renamed matrices for instrumental variables: H_i -> Z_i, H -> Z
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@877 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 uncorrelated with the disturbance terms of all other equations
-($\E \left[ u_i' H_j \right] = 0 \, \forall \, i \neq j$).
+($\E \left[ u_i' Z_j \right] = 0 \, \forall \, i \neq j$).
 Under this null hypothesis, both the 2SLS and the 3SLS estimator are consistent,
@@
 uncorrelated with the disturbances of the same equation
-($\E \left[ u_i' H_i \right] = 0 \, \forall \, i$),
+($\E \left[ u_i' Z_i \right] = 0 \, \forall \, i$),
 but the instrumental variables of at least one equation
 are correlated with the disturbances of another equation
-($\E \left[ u_i' H_j \right] \neq 0 \, \exists \, i \neq j$).
+($\E \left[ u_i' Z_j \right] \neq 0 \, \exists \, i \neq j$).
 The Hausman test statistic is

### 2007-12-05 paper/trunk/systemfit_restrictions.tex
renamed matrices for instrumental variables: H_i -> Z_i, H -> Z
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@877 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
    \left[ \begin{array}{cc}
-      X' H \left( H' \OHat H \right)^{-1} H' X & R' \\
+      X' Z \left( Z' \OHat Z \right)^{-1} Z' X & R' \\
       R & 0
@@
    \left[ \begin{array}{c}
-      X' H \left( H \OHat H \right)^{-1} H' y \\ q
+      X' Z \left( Z \OHat Z \right)^{-1} Z' y \\ q
    \end{array} \right] ,
@@
    \left[ \begin{array}{c}
-      \XHat' \OHat^{-1} H \left( H' H \right)^{-1} H' y \\ q
+      \XHat' \OHat^{-1} Z \left( Z' Z \right)^{-1} Z' y \\ q
    \end{array} \right] ,
@@
    \left[ \begin{array}{cc}
-      \XHat' \OHat^{-1} H \left( H' H \right)^{-1} H' \OHat
-      H \left( H' H \right)^{-1} H' \OHat^{-1} \XHat & 0' \\
+      \XHat' \OHat^{-1} Z \left( Z' Z \right)^{-1} Z' \OHat
+      Z \left( Z' Z \right)^{-1} Z' \OHat^{-1} \XHat & 0' \\
       0 & 0
@@
 If the same instrumental variables are used in all equations 
-($H_1 = H_2 = \ldots = H_G$), 
+($Z_1 = Z_2 = \ldots = Z_G$),
 all the above mentioned approaches lead to identical coefficient estimates

### 2007-12-06 paper/trunk/systemfit_other.tex
revised (sub)section "Testing linear restrictions": added derivation of F test and Wald tests, the description of these tests is more clear, more elaborate and fully correct now
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@884 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 
-Linear restrictions can be tested by an $F$ test, Wald test or
-likelihood ratio (LR) test.
+Linear restrictions can be tested by an $F$ test, two Wald tests and
+a likelihood ratio (LR) test.
 
@@
 F = \frac{
+   ( R \bHat - q )'
+   ( R ( X' ( \Sigma \otimes I )^{-1} X )^{-1} R' )^{-1}
+   ( R \bHat - q ) /
+   j
+}{
+   \uHat' ( \Sigma \otimes I )^{-1} \uHat /
+   ( G \cdot T - K )
+} \; ,
+\label{eq:f-test}
+\end{equation}
+where $j$ is the number of restrictions,
+$K$ is the total number of estimated coefficients,
+and all other variables are as defined before
+\citep[p.~314]{theil71}.
+Under the null hypothesis, $F$ is $F$ distributed
+with $j$ and $G \cdot T - K$ degrees of freedom.
+
+However, $F$ in equation~(\ref{eq:f-test}) cannot be computed,
+because $\Sigma$ is generally unknown.
+As a solution, \citet[p.~314]{theil71} proposes
+to replace the unknown $\Sigma$ in~(\ref{eq:f-test})
+by the estimated covariance matrix $\SHat$.
+\begin{equation}
+\hat{F} = \frac{
    ( R \bHat - q )'
@@
    ( G \cdot T - K )
-} \; ,
+}
+\label{eq:f-test-theil}
 \end{equation}
-where $j$ is the number of restrictions,
-$G$ is the number of equations,
-$T$ is the number of observations per equation,
-$K$ is the total number of estimated coefficients, and
-$\SHat$ is the (estimated) residual covariance matrix
-used in the estimation.
-Under the null hypothesis, $F$ has an approximate $F$ distribution
-with $j$ and $G \cdot T - K$ degrees of freedom
-\citep[p.~314]{theil71}.
+Asymptotically, $\hat{F}$ has the same distribution as $F$
+in equation~(\ref{eq:f-test}),
+because the numerator of~(\ref{eq:f-test-theil})
+converges in probability to the numerator of~(\ref{eq:f-test})
+and the denominator of~(\ref{eq:f-test-theil})
+converges in probability to the denominator of~(\ref{eq:f-test})
+\citep[p.~402]{theil71}.
 
-The Wald statistic for systems of equations is
+Furthermore, the denominators of both~(\ref{eq:f-test})
+and~(\ref{eq:f-test-theil}) converge in probability to~$1$.
+Taking this into account and applying equation~(\ref{eq:cov-ols-wls-sur}),
+we obtain the usual $F$~statistic of the Wald test.
+\begin{equation}
+\hat{\hat{F}} = \frac{
+   ( R \bHat - q )'
+   ( R \, \COVHat \left[ \bHat \right] R' )^{-1}
+   ( R \bHat - q )
+}{
+   j
+}
+\label{eq:f-test-wald}
+\end{equation}
+Under the null hypotheses, also $\hat{\hat{F}}$ is
+asymptotically $F$ distributed
+with $j$ and $G \cdot T - K$ degrees of freedom.
+
+Multiplying equation~(\ref{eq:f-test-wald}) with $j$,
+we obtain the usual $\chi^2$ statistic for the Wald test
 \begin{equation}

### 2007-12-10 paper/trunk/systemfit_estimation.tex
several corrections and a few small improvements
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@901 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 that are calculated from the residuals of a first-step OLS estimation
-(see section~\ref{sec:residcov}).
+(see section~\ref{sec:residcov}).%
+\footnote{%
+Note that $\OHat$ in equation~(\ref{eq:ols-wls-sur})
+is not the same $\OHat$ as in equation~(\ref{eq:cov-ols-wls-sur}).
+The first is calculated from the residuals of a first-step OLS estimation;
+the second is calculated from the residuals of this WLS estimation.
+The same applies to the SUR, W2SLS, and 3SLS estimations
+described in the following sections.
+}
 
@@
 \begin{equation}
-   \XHat_i = Z_i \left( Z_i' Z_i \right)^{-1} Z_i' X .
+   \XHat_i = Z_i \left( Z_i' Z_i \right)^{-1} Z_i' X_i .
 \end{equation}
@@
    \bHat
-   = \left( \XHat' \OHat^{-1} \XHat \right)^{-1} \XHat' \OHat^{-1} y .
+   = \left( \XHat' \OHat^{-1} \XHat \right)^{-1} \XHat' \OHat^{-1} y ,
    \label{eq:2sls-w2sls-3sls}
+\end{equation}
+where
+\begin{equation}
+   \XHat =
+   \left[ \begin{array}{cccc}
+      \XHat_1 & 0       & \cdots & 0\\
+      0       & \XHat_2 & \cdots & 0\\
+      \vdots  & \vdots  & \ddots & \vdots\\
+      0       & 0       & \cdots & \XHat_G
+   \end{array}\right] .
 \end{equation}
@@
 $( \E \left[ u_i' Z_i \right]) = 0 \; \forall \; i )$,
-\cite{schmidt90} points out that this estimator is only consistent 
+\cite{schmidt90} points out that the 3SLS estimator is only consistent
 if all disturbance terms and all instrumental variables are uncorrelated
 $( \E \left[ u_i' Z_j \right]) = 0 \; \forall \; i, j )$.
-
 Since there might be occasions where this cannot be avoided,
 \cite{schmidt90} analyses other approaches to obtain 3SLS estimators.
+
 One of these approaches based on instrumental variable estimation

### 2007-12-10 paper/trunk/systemfit_intro.tex
clarified notations by avoiding the ambiguous word 'impose'; instead, we now clearly describe whether we mean 'estimating coefficients under restrictions' or 'testing restrictions on the coefficients'; noted this in the response to the referees
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@866 591db2bd-e0a0-4d22-8c19-2192dbb7da31
several corrections and a few small improvements
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@901 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 }
-These restrictions can be tested and/or imposed only in a simultaneous
-estimation approach.
+Estimating the coefficients under cross-equation restrictions
+and testing these restrictions
+requires a simultaneous estimation approach.
 
@@
 Section~\ref{sec:Usage} demonstrates how to run
-\pkg{systemfit} and how some of the features presented in the previous
+\pkg{systemfit} and how some of the features presented in the second
 section can be used.

### 2007-12-10 paper/trunk/systemfit_summary.tex
The coefficients of the model (beta) are consistently called 'coefficients' now (and no longer 'parameters' every now and then); noted this in the response to the referees
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@864 591db2bd-e0a0-4d22-8c19-2192dbb7da31
several corrections and a few small improvements
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@901 591db2bd-e0a0-4d22-8c19-2192dbb7da31

--- 2007-06-05 paper/trunk/systemfit_summary.tex
@@
 In this article, we have described some of the basic features of the
-\pkg{systemfit} package for estimating of systems of linear
+\pkg{systemfit} package for estimating systems of linear
 equations.
@@
 Furthermore, the package provides some statistical tests
-for parameter restrictions and consistency of 3SLS estimation.
+for restrictions on the coefficients and consistency of 3SLS estimation.
 It has been tested on a variety of datasets and has produced satisfactory

### 2007-12-10 paper/trunk/systemfit.tex
added appendix (table) to compare values returned by systemfit() and values returned by lm()
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@695 591db2bd-e0a0-4d22-8c19-2192dbb7da31
updated path to our department's BibTeX data base
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@708 591db2bd-e0a0-4d22-8c19-2192dbb7da31
substituted "\COV" (defined in jss.cls) for all "\Cov"; removed definition(s) of "\Cov"
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@710 591db2bd-e0a0-4d22-8c19-2192dbb7da31
The paper no longer treats Omega as known in sections 2.1 an

@@
 \documentclass[article]{jss}
-\usepackage[pdftex]{thumbpdf}
-%   run: pdflatex *.tex -> thumbpdf *.tex -> pdflatex *.tex
 \usepackage{csquotes}
@@
 
-\usepackage{lineno}
+% \usepackage{lineno}
+% 
+% \linenumbers
 
-\linenumbers
-
-\newcommand{\Cov}{\text{Cov}}
+\newcommand{\bHat}{\hat{\beta}}
+\newcommand{\COVHat}{\widehat{\COV}}
+\newcommand{\lHat}{\hat{\lambda}}
+\newcommand{\OHat}{\widehat{\Omega}}
+\newcommand{\SHat}{\widehat{\Sigma}}
+\newcommand{\sHat}{\hat{\sigma}}
+\newcommand{\uHat}{\hat{u}}
+\newcommand{\XHat}{\widehat{X}}
 \clubpenalty=10000
 \widowpenalty=10000
+\setlength{\emergencystretch}{3em}
+% \setlength{\overfullrule}{5pt}
+
+\newcommand{\codeD}[2]{\code{#1}\hspace{0pt}\code{.#2}}
+\newcommand{\codeDD}[3]{\code{#1}\hspace{0pt}\code{.#2}\hspace{0pt}\code{.#3}}
+\newcommand{\codeDDD}[4]{\code{#1}\hspace{0pt}\code{.#2}\hspace{0pt}%
+\code{.#3}\hspace{0pt}\code{.#4}}
 
@@
 
+\section*{Acknowledgments}
+We thank Achim Zeileis, John Fox, Ott Toomet, William H.\ Greene,
+two anonymous referees
+and several users of \pkg{systemfit}
+for their comments and suggestions
+that helped us to improve the \pkg{systemfit} package as well as this paper.
+Of course, any remaining errors are the authors'.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \bibliography{systemfit} % a subset of my big bibtex file
-%\bibliography{/home/suapm095/Documents/Literatur/arne}
+%\bibliography{agrarpol}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\appendix
+\input{systemfit_compare_lm.tex}
+
+\input{systemfit_timings.tex}
+
+\input{systemfit_sem.tex}
 

### 2007-12-27 paper/trunk/systemfit.tex
using now (hopefully) always \proglang, \pkg, and \code (also in titles, references, ...)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@908 591db2bd-e0a0-4d22-8c19-2192dbb7da31
changed \Shorttitle: "systemfit" -> "\pkg{systemfit}: A Package for Estimating Systems of Simultaneous Equations in \proglang{R}"
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@909 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 
-\title{systemfit: A Package for Estimating\\
+\title{\pkg{systemfit}: A Package for Estimating\\
        Systems of Simultaneous Equations in \proglang{R}}
@@
             in R} %% without formatting
-\Shorttitle{systemfit} %% a short title (if necessary)
+\Shorttitle{\pkg{systemfit}: A Package for Estimating
+   Systems of Simultaneous Equations in \proglang{R}}
+   %% a short title (if necessary)
 

### 2007-12-27 paper/trunk/systemfit_abstract.tex
made minor updates to the abstract for readability
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@815 591db2bd-e0a0-4d22-8c19-2192dbb7da31
changed last sentence from shown to demonstrated.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@816 591db2bd-e0a0-4d22-8c19-2192dbb7da31
revised some of Jeff's modifications
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@818 591db2bd-e0a0-4d22-8c19-2192dbb7da31
expressions like "Generalized Least Squares" are no longer capitalized
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@910 591db2bd-e0a0-4d

@@
 For instance, this package can be used for 
-``Ordinary Least Squares'' (OLS),
-``Seemingly Unrelated Regression'' (SUR),
+``ordinary least squares'' (OLS),
+``seemingly unrelated regression'' (SUR),
 and the instrumental variable (IV) methods
-``Two-Stage Least Squares'' (2SLS) and
-``Three-Stage Least Squares'' (3SLS),
+``two-stage least squares'' (2SLS) and
+``three-stage least squares'' (3SLS),
 where SUR and 3SLS estimations can optionally be iterated.
@@
 It has been tested on a variety of datasets
-and the reliability of its results has been shown.
+and its reliability is demonstrated.
 }

### 2007-12-27 paper/trunk/systemfit_estimation.tex
expressions like "Generalized Least Squares" are no longer capitalized
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@910 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 the system of equations~(\ref{eq:model})
-can be consistently estimated by Ordinary Least Squares (OLS),
-Weighted Least Squares (WLS), and Seemingly Unrelated Regression (SUR).
+can be consistently estimated by ordinary least squares (OLS),
+weighted least squares (WLS), and seemingly unrelated regression (SUR).
 These estimators can be obtained by
@@
 
-The Ordinary Least Squares (OLS) estimator is based on the assumption
+The ordinary least squares (OLS) estimator is based on the assumption
 that the disturbance terms are not contemporaneously correlated
@@
 
-The Weighted Least Squares (WLS) estimator allows for different variances
+The weighted least squares (WLS) estimator allows for different variances
 of the disturbance terms in the different equations
@@
 
-If the disturbances are contemporaneously correlated, a Generalized
-Least Squares (GLS) estimation leads to an efficient estimator for the coefficients.
-In this case, the GLS estimator is generally called °Seemingly Unrelated Regression°
+If the disturbances are contemporaneously correlated, a generalized
+least squares (GLS) estimation leads to an efficient estimator for the coefficients.
+In this case, the GLS estimator is generally called °seemingly unrelated regression°
 (SUR) estimator \citep{zellner62}.
@@
 The textbook solution for this problem is
-a Feasible Generalized Least Squares (FGLS) estimation.
+a feasible generalized least squares (FGLS) estimation.
 As the FGLS estimator is based on
@@
 
-The Two-Stage Least Squares (2SLS) estimator
+The two-stage least squares (2SLS) estimator
 is based on the same assumptions about the disturbance terms as the OLS estimator.
@@
 
-The Weighted Two-Stage Least Squares (W2SLS) estimator allows
+The weighted two-stage least squares (W2SLS) estimator allows
 for different variances of the disturbance terms in the different equations.
@@
 If the disturbances are contemporaneously correlated,
-a Feasible Generalized Least Squares (FGLS) version of the two-stage least squares
+a feasible generalized least squares (FGLS) version of the two-stage least squares
 estimation leads to consistent and asymptotically more efficient estimates.
-This estimation procedure is generally called °Three-stage Least
-Squares° (3SLS) \citep{zellner62b}.
+This estimation procedure is generally called °three-stage least
+squares° (3SLS) \citep{zellner62b}.
 The standard 3SLS estimator and its covariance matrix are obtained by
@@
 \end{equation}
-Another approach based on the Generalized Method of Moments (GMM)
+Another approach based on the generalized method of moments (GMM)
 estimator (3SLS-GMM) is

### 2007-12-27 paper/trunk/systemfit_intro.tex
expressions like "Generalized Least Squares" are no longer capitalized
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@910 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 However, estimating all equations simultaneously
-with a °Generalized Least Squares° (GLS) estimator,
+with a °generalized least squares° (GLS) estimator,
 which takes the covariance structure of the residuals into account,
@@
 This estimation procedure is generally called
-°Seemingly Unrelated Regression° (SUR) \citep{zellner62}.
+°seemingly unrelated regression° (SUR) \citep{zellner62}.
 Another reason to estimate a system of equations simultaneously are
@@
 This simultaneity bias can be corrected for by
-applying a °Two-Stage Least Squares° (2SLS) estimation
+applying a °two-stage least squares° (2SLS) estimation
 to each equation.
@@
 in a simultaneous estimation of the system of equations
-by the °Three-Stage Least Squares° (3SLS) method
+by the °three-stage least squares° (3SLS) method
 \citep{zellner62b}.
@@
 Currently, the estimation methods
-°Ordinary Least Squares° (OLS),
-°Weighted Least Squares° (WLS),
-°Seemingly Unrelated Regression° (SUR),
-°Two-Stage Least Squares° (2SLS),
-°Weighted Two-Stage Least Squares° (W2SLS), and
-°Three-Stage Least Squares° (3SLS)
+°ordinary least squares° (OLS),
+°weighted least squares° (WLS),
+°seemingly unrelated regression° (SUR),
+°two-stage least squares° (2SLS),
+°weighted two-stage least squares° (W2SLS), and
+°three-stage least squares° (3SLS)
 are implemented.%
@@
 In this context,
-the term °weighted° in °Weighted Least Squares° (WLS) and
-°Weighted Two-Stage Least Squares° (W2SLS) means
+the term °weighted° in °weighted least squares° (WLS) and
+°weighted two-stage least squares° (W2SLS) means
 that the \emph{equations} might have different weights

### 2007-12-27 paper/trunk/systemfit_other.tex
added two \labels for referencing these equations
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@886 591db2bd-e0a0-4d22-8c19-2192dbb7da31
several corrections and a few small improvements
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@901 591db2bd-e0a0-4d22-8c19-2192dbb7da31
expressions like "Generalized Least Squares" are no longer capitalized
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@910 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
    { y' \left( \SHat^{-1} \otimes
-   \left( I - \frac{i i'}{T} \right) \right) y }
+   \left( I_T - \frac{\iota \iota'}{T} \right) \right) y } ,
 \end{equation}
-\citep{mcelroy77},
-where $T$ is the number of observations in each equation,
-$I$ is an $T \times T$ identity matrix and 
-$i$ is a column vector of $T$ ones.
+where $\iota$ is a column vector of $T$ ones \citep{mcelroy77}.
 
@@
    ( R \bHat - q ) .
+\label{eq:chi2-test-wald}
 \end{equation}
@@
 
-The Likelihood-Ratio (LR) statistic for systems of equations is
+The likelihood-ratio (LR) statistic for systems of equations is
 \begin{equation}
@@
    \right) ,
+\label{eq:lr-test}
 \end{equation}

### 2007-12-27 paper/trunk/systemfit_other.tex
using \proglang for "LIMDEP" now
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@911 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 For instance, this approach is applied by
-the command °TSCS° of the software LIMDEP that carries out SUR estimations
+the command °TSCS° of the software \proglang{LIMDEP} that carries out SUR estimations
 in which all coefficient vectors are constrained to be equal

### 2007-12-27 paper/trunk/systemfit_estimation.tex
changed transpose symbol: ' -> ^\top
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@913 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 \begin{equation}
-   \bHat = \left( X' \OHat^{-1} X \right)^{-1} X' \OHat^{-1} y .
+   \bHat = \left( X^\top \OHat^{-1} X \right)^{-1} X^\top \OHat^{-1} y .
    \label{eq:ols-wls-sur}
@@
 \begin{equation}
-   \COVHat \left[ \bHat \right] = \left( X' \OHat^{-1} X \right)^{-1} .
+   \COVHat \left[ \bHat \right] = \left( X^\top \OHat^{-1} X \right)^{-1} .
    \label{eq:cov-ols-wls-sur}
@@
 equation are uncorrelated
-$( \E \left[ u_i ' X_i \right] = 0 \; \forall \; i )$,
+$( \E \left[ u_i^\top X_i \right] = 0 \; \forall \; i )$,
 a consistent SUR estimation requires that all disturbance terms and all 
 regressors are uncorrelated
-$( \E \left[ u_i ' X_j \right] = 0 \; \forall \; i, j )$.
+$( \E \left[ u_i^\top X_j \right] = 0 \; \forall \; i, j )$.
 
@@
 If the regressors of one or more equations are correlated
-with the disturbances ($\E \left[ u_i' X_i \right] \neq 0$),
+with the disturbances ($\E \left[ u_i^\top X_i \right] \neq 0$),
 OLS, WLS, and SUR estimates are biased.
@@
 the disturbance terms of the corresponding equation 
-($\E \left[ u_i' Z_i \right] = 0$).
+($\E \left[ u_i^\top Z_i \right] = 0$).
 
@@
 \begin{equation}
-   \XHat_i = Z_i \left( Z_i' Z_i \right)^{-1} Z_i' X_i .
+   \XHat_i = Z_i \left( Z_i^\top Z_i \right)^{-1} Z_i^\top X_i .
 \end{equation}
@@
    \bHat
-   = \left( \XHat' \OHat^{-1} \XHat \right)^{-1} \XHat' \OHat^{-1} y ,
+   = \left( \XHat^\top \OHat^{-1} \XHat \right)^{-1} \XHat^\top \OHat^{-1} y ,
    \label{eq:2sls-w2sls-3sls}
@@
    \COVHat \left[ \bHat \right]
-   = \left( \XHat' \OHat^{-1} \XHat \right)^{-1} .
+   = \left( \XHat^\top \OHat^{-1} \XHat \right)^{-1} .
    \label{eq:cov-2sls-w2sls-3sls}
@@
 equation are uncorrelated
-$( \E \left[ u_i' Z_i \right]) = 0 \; \forall \; i )$,
+$( \E \left[ u_i^\top Z_i \right]) = 0 \; \forall \; i )$,
 \cite{schmidt90} points out that the 3SLS estimator is only consistent
 if all disturbance terms and all instrumental variables are uncorrelated
-$( \E \left[ u_i' Z_j \right]) = 0 \; \forall \; i, j )$.
+$( \E \left[ u_i^\top Z_j \right]) = 0 \; \forall \; i, j )$.
 Since there might be occasions where this cannot be avoided,
@@
 \begin{equation}
-   \bHat_{3SLS-IV} = \left( \XHat' \OHat^{-1} X
-   \right)^{-1} \XHat' \OHat^{-1} y .
+   \bHat_{3SLS-IV} = \left( \XHat^\top \OHat^{-1} X
+   \right)^{-1} \XHat^\top \OHat^{-1} y .
    \label{eq:3slsIv}
@@
 \begin{equation}
-   \COVHat \left[ \bHat_{3SLS-IV} \right] = \left( \XHat' \OHat^{-1}
+   \COVHat \left[ \bHat_{3SLS-IV} \right] = \left( \XHat^\top \OHat^{-1}
    X \right)^{-1} .
@@
 \begin{equation}
-   \bHat_{3SLS-GMM} = \left( X' Z \left( Z' \OHat Z \right)^{-1}
-   Z' X \right)^{-1} X' Z \left( Z' \OHat Z \right)^{-1} Z' y
+   \bHat_{3SLS-GMM} = \left( X^\top Z \left( Z^\top \OHat Z \right)^{-1}
+   Z^\top X \right)^{-1} X^\top Z \left( Z^\top \OHat Z \right)^{-1} Z^\top y
    \label{eq:3slsGmm}
@@
    \COVHat \left[ \bHat_{3SLS-GMM} \right] =
-   \left( X' Z \left( Z' \OHat Z \right)^{-1} Z' X \right)^{-1} .
+   \left( X^\top Z \left( Z^\top \OHat Z \right)^{-1} Z^\top X \right)^{-1} .
 \end{equation}
@@
 \begin{equation}
-   \bHat_{3SLS-Schmidt} = \left( \XHat' \OHat^{-1} \XHat
-   \right)^{-1} \XHat' \OHat^{-1}
-   Z \left( Z' Z \right)^{-1} Z' y .
+   \bHat_{3SLS-Schmidt} = \left( \XHat^\top \OHat^{-1} \XHat
+   \right)^{-1} \XHat^\top \OHat^{-1}
+   Z \left( Z^\top Z \right)^{-1} Z^\top y .
    \label{eq:3slsSchmidt}
@@
    \COVHat \left[ \bHat_{3SLS-Schmidt} \right] = &
-   \left( \XHat' \OHat^{-1}  \XHat \right)^{-1}
-   \XHat' \OHat^{-1} Z \left( Z' Z \right)^{-1} Z' \OHat Z \\
-   & \left( Z' Z \right)^{-1} Z' \OHat^{-1} \XHat
-   \left( \XHat' \OHat^{-1}  \XHat \right)^{-1} .
+   \left( \XHat^\top \OHat^{-1}  \XHat \right)^{-1}
+   \XHat^\top \OHat^{-1} Z \left( Z^\top Z \right)^{-1} Z^\top \OHat Z \\
+   & \left( Z^\top Z \right)^{-1} Z^\top \OHat^{-1} \XHat
+   \left( \XHat^\top \OHat^{-1}  \XHat \right)^{-1} .
    \nonumber
@@
    \bHat_{3SLS-EViews} = \bHat_{2SLS} +
-   \left( \XHat' \OHat^{-1} \XHat \right)^{-1}
-   \XHat' \OHat^{-1} \left( y - X \bHat_{2SLS} \right) ,
+   \left( \XHat^\top \OHat^{-1} \XHat \right)^{-1}
+   \XHat^\top \OHat^{-1} \left( y - X \bHat_{2SLS} \right) ,
    \label{eq:3slsEViews}

### 2007-12-27 paper/trunk/systemfit_other.tex
changed transpose symbol: ' -> ^\top
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@913 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 \begin{equation}
-   \sHat_{ij} = \frac{ \uHat_i' \uHat_j }{ T },
+   \sHat_{ij} = \frac{ \uHat_i^\top \uHat_j }{ T },
    \label{eq:rcovNoDfCor}
@@
 \begin{equation}
-   \sHat_{ij} = \frac{ \uHat_i' \uHat_j }
+   \sHat_{ij} = \frac{ \uHat_i^\top \uHat_j }
    { \sqrt{ \left( T - K_i \right) \cdot \left( T - K_j \right) } }
@@
 \begin{equation}
-   \sHat_{ij} = \frac{ \uHat_i' \uHat_j }
+   \sHat_{ij} = \frac{ \uHat_i^\top \uHat_j }
    { T - \max \left( K_i , K_j \right) } \; ,
@@
    \sHat_{ij} & = &
-   \frac{ \uHat_i' \uHat_j }
-   { T - K_i - K_j + tr \left[ X_i \left( X_i' X_i \right)^{-1}
-   X_i' X_j \left( X_j' X_j \right)^{-1} X_j' \right] }
+   \frac{ \uHat_i^\top \uHat_j }
+   { T - K_i - K_j + tr \left[ X_i \left( X_i^\top X_i \right)^{-1}
+   X_i^\top X_j \left( X_j^\top X_j \right)^{-1} X_j^\top \right] }
    \label{eq:rcovTheil} \\
    & = &
-   \frac{ \uHat_i' \uHat_j }
-   { T - K_i - K_j + tr \left[ \left( X_i' X_i \right)^{-1}
-   X_i' X_j \left( X_j' X_j \right)^{-1} X_j' X_i \right] }
+   \frac{ \uHat_i^\top \uHat_j }
+   { T - K_i - K_j + tr \left[ \left( X_i^\top X_i \right)^{-1}
+   X_i^\top X_j \left( X_j^\top X_j \right)^{-1} X_j^\top X_i \right] }
 \end{eqnarray} 
@@
 \begin{equation}
-   R_i^2 = 1 - \frac{ \uHat_i' \uHat_i }
-   { ( y_i - \overline{y_i} )' ( y_i - \overline{y_i} ) } \; ,
+   R_i^2 = 1 - \frac{ \uHat_i^\top \uHat_i }
+   { ( y_i - \overline{y_i} )^\top ( y_i - \overline{y_i} ) } \; ,
 \end{equation}
@@
 \begin{equation}
-   R_*^2 = 1 - \frac{ \uHat' \OHat^{-1} \uHat }
-   { y' \left( \SHat^{-1} \otimes
-   \left( I_T - \frac{\iota \iota'}{T} \right) \right) y } ,
+   R_*^2 = 1 - \frac{ \uHat^\top \OHat^{-1} \uHat }
+   { y^\top \left( \SHat^{-1} \otimes
+   \left( I_T - \frac{\iota \iota^\top}{T} \right) \right) y } ,
 \end{equation}
@@
 F = \frac{
-   ( R \bHat - q )'
-   ( R ( X' ( \Sigma \otimes I )^{-1} X )^{-1} R' )^{-1}
+   ( R \bHat - q )^\top
+   ( R ( X^\top ( \Sigma \otimes I )^{-1} X )^{-1} R^\top )^{-1}
    ( R \bHat - q ) /
@@
 }{
-   \uHat' ( \Sigma \otimes I )^{-1} \uHat /
+   \uHat^\top ( \Sigma \otimes I )^{-1} \uHat /
    ( G \cdot T - K )
@@
 \hat{F} = \frac{
-   ( R \bHat - q )'
-   ( R ( X' ( \SHat \otimes I )^{-1} X )^{-1} R' )^{-1}
+   ( R \bHat - q )^\top
+   ( R ( X^\top ( \SHat \otimes I )^{-1} X )^{-1} R^\top )^{-1}
    ( R \bHat - q ) /
@@
 }{
-   \uHat' ( \SHat \otimes I )^{-1} \uHat /
+   \uHat^\top ( \SHat \otimes I )^{-1} \uHat /
    ( G \cdot T - K )
@@
 \hat{\hat{F}} = \frac{
-   ( R \bHat - q )'
-   ( R \, \COVHat \left[ \bHat \right] R' )^{-1}
+   ( R \bHat - q )^\top
+   ( R \, \COVHat \left[ \bHat \right] R^\top )^{-1}
    ( R \bHat - q )
@@
 W =
-   ( R \bHat - q )'
-   ( R \, \COVHat [ \bHat ] \, R' )^{-1}
+   ( R \bHat - q )^\top
+   ( R \, \COVHat [ \bHat ] \, R^\top )^{-1}
    ( R \bHat - q ) .
@@
 uncorrelated with the disturbance terms of all other equations
-($\E \left[ u_i' Z_j \right] = 0 \, \forall \, i \neq j$).
+($\E \left[ u_i^\top Z_j \right] = 0 \, \forall \, i \neq j$).
 Under this null hypothesis, both the 2SLS and the 3SLS estimator are consistent,
@@
 uncorrelated with the disturbances of the same equation
-($\E \left[ u_i' Z_i \right] = 0 \, \forall \, i$),
+($\E \left[ u_i^\top Z_i \right] = 0 \, \forall \, i$),
 but the instrumental variables of at least one equation
 are correlated with the disturbances of another equation
-($\E \left[ u_i' Z_j \right] \neq 0 \, \exists \, i \neq j$).
+($\E \left[ u_i^\top Z_j \right] \neq 0 \, \exists \, i \neq j$).
 The Hausman test statistic is
 \begin{equation}
-  m = \left( \bHat_{2SLS} - \bHat_{3SLS} \right)^{'}
+  m = \left( \bHat_{2SLS} - \bHat_{3SLS} \right)^\top
       \left( \COVHat \left[ \bHat_{2SLS} \right] -

### 2007-12-27 paper/trunk/systemfit_estimation.tex
when referring to equations, we now always use "Equation~\ref{}"
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@914 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 If all regressors are exogenous,
-the system of equations~(\ref{eq:model})
+the system of equations (Equation~\ref{eq:model})
 can be consistently estimated by ordinary least squares (OLS),
@@
 $( \sigma_i^2 = \sigma_j^2 \, \forall \, i, j)$.
-In this case, $\OHat$ in equation~(\ref{eq:ols-wls-sur})
+In this case, $\OHat$ in Equation~\ref{eq:ols-wls-sur}
 is equal to $I_{G \cdot T}$ and thus, cancels out.
@@
 If the whole system is treated as one single equation,
-$\OHat$ in equation~(\ref{eq:cov-ols-wls-sur}) is $\sHat^2 I_{G \cdot T}$,
+$\OHat$ in Equation~\ref{eq:cov-ols-wls-sur} is $\sHat^2 I_{G \cdot T}$,
 where $\sHat^2$ is an estimator for the variance of all disturbances
@@
 are allowed to have different variances,
-$\OHat$ in equation~(\ref{eq:cov-ols-wls-sur}) is $\SHat \otimes I_T$,
+$\OHat$ in Equation~\ref{eq:cov-ols-wls-sur} is $\SHat \otimes I_T$,
 where $\sHat_{ij} = 0 \; \forall \; i \neq j$ and
@@
 OLS estimation is equal to the covariance matrix obtained
-by equation~(\ref{eq:cov-ols-wls-sur})
+by Equation~\ref{eq:cov-ols-wls-sur}
 with $\OHat$ equal to $\SHat \otimes I_T$.
@@
 that the disturbance terms are not contemporaneously correlated.
-In this case, $\OHat$ in equations~(\ref{eq:ols-wls-sur})
-and~(\ref{eq:cov-ols-wls-sur})
+In this case, $\OHat$ in Equations~\ref{eq:ols-wls-sur}
+and~\ref{eq:cov-ols-wls-sur}
 is $\SHat \otimes I_T$,
@@
 \footnote{%
-Note that $\OHat$ in equation~(\ref{eq:ols-wls-sur})
-is not the same $\OHat$ as in equation~(\ref{eq:cov-ols-wls-sur}).
+Note that $\OHat$ in Equation~\ref{eq:ols-wls-sur}
+is not the same $\OHat$ as in Equation~\ref{eq:cov-ols-wls-sur}.
 The first is calculated from the residuals of a first-step OLS estimation;
@@
 it is only asymptotically efficient.
-In case of a SUR estimator, $\OHat$ in equations~(\ref{eq:ols-wls-sur})
-and~(\ref{eq:cov-ols-wls-sur})
+In case of a SUR estimator, $\OHat$ in Equations~\ref{eq:ols-wls-sur}
+and~\ref{eq:cov-ols-wls-sur}
 is $\SHat \otimes I_T$,
@@
 Then, these 'fitted' regressors are substituted for the original regressors
-in equation~(\ref{eq:ols-wls-sur})
+in Equation~\ref{eq:ols-wls-sur}
 to obtain unbiased 2SLS, W2SLS, or 3SLS estimates of $\beta$ by
@@
 An estimator of the covariance matrix of the estimated coefficients
-can be obtained from equation~(\ref{eq:cov-ols-wls-sur}) analogously.
+can be obtained from Equation~\ref{eq:cov-ols-wls-sur} analogously.
 Hence, we get
@@
 is based on the same assumptions about the disturbance terms as the OLS estimator.
-Accordingly, $\OHat$ in equation~(\ref{eq:2sls-w2sls-3sls})
+Accordingly, $\OHat$ in Equation~\ref{eq:2sls-w2sls-3sls}
 is equal to $I_{G \cdot T}$ and thus, cancels out.
@@
 the whole system can be treated either as one single equation
-with $\OHat$ in equation~(\ref{eq:cov-2sls-w2sls-3sls})
+with $\OHat$ in Equation~\ref{eq:cov-2sls-w2sls-3sls}
 equal to $\sHat^2 I_{G \cdot T}$,
@@
 are allowed to have different variances
-with $\OHat$ in equation~(\ref{eq:cov-2sls-w2sls-3sls})
+with $\OHat$ in Equation~\ref{eq:cov-2sls-w2sls-3sls}
 equal to $\SHat \otimes I_T$,
@@
 for different variances of the disturbance terms in the different equations.
-Hence, $\OHat$ in equations~(\ref{eq:2sls-w2sls-3sls})
-and~(\ref{eq:cov-2sls-w2sls-3sls})
+Hence, $\OHat$ in Equations~\ref{eq:2sls-w2sls-3sls}
+and~\ref{eq:cov-2sls-w2sls-3sls}
 is $\SHat \otimes I_T$,
@@
 The standard 3SLS estimator and its covariance matrix are obtained by
-equations~(\ref{eq:2sls-w2sls-3sls}) and~(\ref{eq:cov-2sls-w2sls-3sls})
+Equations~\ref{eq:2sls-w2sls-3sls} and~\ref{eq:cov-2sls-w2sls-3sls}
 with $\OHat$ equal to $\SHat \otimes I_T$,
@@
 as defined above.
-\proglang{EViews} uses the standard 3SLS formula~(\ref{eq:cov-2sls-w2sls-3sls}) to
+\proglang{EViews} uses the standard 3SLS formula
+(Equation~\ref{eq:cov-2sls-w2sls-3sls}) to
 calculate an estimator of the covariance matrix of the estimated coefficients.
@@
 \citep{schmidt90}.
-In this case, only the 3SLS-GMM (\ref{eq:3slsGmm})
+In this case, only the 3SLS-GMM (Equation~\ref{eq:3slsGmm})
 and the 3SLS estimator developed by \cite{schmidt90} 
-(\ref{eq:3slsSchmidt}) are consistent.
+(Equation~\ref{eq:3slsSchmidt}) are consistent.
 

### 2007-12-27 paper/trunk/systemfit_other.tex
when referring to equations, we now always use "Equation~\ref{}"
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@914 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 of $\Sigma^{-1}$ \citep[p.\ 322]{theil71}.
-Furthermore, the covariance matrix calculated by (\ref{eq:rcovTheil})
+Furthermore, the covariance matrix calculated by Equation~\ref{eq:rcovTheil}
 is not necessarily positive semidefinite \citep[p.\ 322]{theil71}. 
 Hence, °it is doubtful whether [this formula] is really superior to 
-[(\ref{eq:rcovNoDfCor})]° \citep[p.\ 322]{theil71}.
+[Equation~\ref{eq:rcovNoDfCor}]° \citep[p.\ 322]{theil71}.
 
@@
 from the residuals of the previous iteration.
-If equation (\ref{eq:rcovNoDfCor}) is applied to calculate the
+If Equation~\ref{eq:rcovNoDfCor} is applied to calculate the
 residual covariance matrix,
@@
 and substituting $\uHat_i - \overline{ \uHat }_i$
-for $\uHat_i$ in (\ref{eq:rcovNoDfCor}--\ref{eq:rcovTheil}).
+for $\uHat_i$ in Equations~\ref{eq:rcovNoDfCor}--\ref{eq:rcovTheil}.
 
@@
 the covariance matrix of the coefficients is calculated
-with $\OHat$ in equation~(\ref{eq:cov-ols-wls-sur})
+with $\OHat$ in Equation~\ref{eq:cov-ols-wls-sur}
 equal to $\SHat \otimes I_T$,
@@
 the covariance matrix of an OLS estimation is calculated
-with $\OHat$ in equation~(\ref{eq:cov-ols-wls-sur})
+with $\OHat$ in Equation~\ref{eq:cov-ols-wls-sur}
 equal to $\sHat^2 I_{G \cdot T}$,
@@
 
-However, $F$ in equation~(\ref{eq:f-test}) cannot be computed,
+However, $F$ in Equation~\ref{eq:f-test} cannot be computed,
 because $\Sigma$ is generally unknown.
 As a solution, \citet[p.~314]{theil71} proposes
-to replace the unknown $\Sigma$ in~(\ref{eq:f-test})
+to replace the unknown $\Sigma$ in Equation~\ref{eq:f-test}
 by the estimated covariance matrix $\SHat$.
@@
 Asymptotically, $\hat{F}$ has the same distribution as $F$
-in equation~(\ref{eq:f-test}),
-because the numerator of~(\ref{eq:f-test-theil})
-converges in probability to the numerator of~(\ref{eq:f-test})
-and the denominator of~(\ref{eq:f-test-theil})
-converges in probability to the denominator of~(\ref{eq:f-test})
+in Equation~\ref{eq:f-test},
+because the numerator of Equation~\ref{eq:f-test-theil}
+converges in probability to the numerator of Equation~\ref{eq:f-test}
+and the denominator of Equation~\ref{eq:f-test-theil}
+converges in probability to the denominator of Equation~\ref{eq:f-test}
 \citep[p.~402]{theil71}.
 
-Furthermore, the denominators of both~(\ref{eq:f-test})
-and~(\ref{eq:f-test-theil}) converge in probability to~$1$.
-Taking this into account and applying equation~(\ref{eq:cov-ols-wls-sur}),
+Furthermore, the denominators of both Equations~\ref{eq:f-test}
+and~\ref{eq:f-test-theil} converge in probability to~$1$.
+Taking this into account and applying Equation~\ref{eq:cov-ols-wls-sur},
 we obtain the usual $F$~statistic of the Wald test.
@@
 
-Multiplying equation~(\ref{eq:f-test-wald}) with $j$,
+Multiplying Equation~\ref{eq:f-test-wald} with $j$,
 we obtain the usual $\chi^2$ statistic for the Wald test
@@
 $\SHat_r$ and $\SHat_u$ are
-the residual covariance matrices calculated by formula (\ref{eq:rcovNoDfCor})
+the residual covariance matrices calculated by Equation~\ref{eq:rcovNoDfCor}
 of the restricted and unrestricted estimation, respectively.

### 2007-12-27 paper/trunk/systemfit_restrictions.tex
several corrections and a few small improvements
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@901 591db2bd-e0a0-4d22-8c19-2192dbb7da31
changed transpose symbol: ' -> ^\top
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@913 591db2bd-e0a0-4d22-8c19-2192dbb7da31
when referring to equations, we now always use "Equation~\ref{}"
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@914 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 
-If the first method (equation~\ref{eq:T-restr}) is chosen
+If the first method (Equation~\ref{eq:T-restr}) is chosen
 to estimate the coefficients under these restrictions,
@@
 in the previous section is done
-(equations~\ref{eq:ols-wls-sur}--\ref{eq:3slsEViews}).
+(Equations~\ref{eq:ols-wls-sur}--\ref{eq:3slsEViews}).
 This results in the linear independent coefficient estimates $\bHat^M$ and
 their covariance matrix.
-The original coefficients can be obtained by equation (\ref{eq:T-restr})
+The original coefficients can be obtained by Equation~\ref{eq:T-restr}
 and the estimated covariance matrix of the original coefficients
@@
    \COVHat \left[ \bHat \right]
-   = M \cdot \COVHat \left[ \bHat^M \right] \cdot M' .
+   = M \cdot \COVHat \left[ \bHat^M \right] \cdot M^\top .
 \end{equation}
@@
 The implementation of the second method to estimate the coefficients
-under linear restrictions (equation~\ref{eq:restr-R})
+under linear restrictions (Equation~\ref{eq:restr-R})
 is described for each estimation method in the following sections.
@@
    \left[ \begin{array}{c}
-      \bHat^R_{SUR} \\[0.2em] \lHat
+      \bHat^R \\[0.2em] \lHat
    \end{array} \right]
@@
    \left[ \begin{array}{cc}
-      X' \OHat^{-1} X & R' \\
+      X^\top \OHat^{-1} X & R^\top \\
       R & 0
@@
    \left[ \begin{array}{c}
-      X' \OHat^{-1} y \\ q
+      X^\top \OHat^{-1} y \\ q
    \end{array} \right] ,
@@
    \left[ \begin{array}{c}
-      \bHat^R_{SUR} \\[0.2em] \lHat
+      \bHat^R \\[0.2em] \lHat
    \end{array} \right] 
@@
    \left[ \begin{array}{c}
-      \bHat^R_{3SLS} \\[0.2em] \lHat
+      \bHat^R \\[0.2em] \lHat
    \end{array} \right]
@@
    \left[ \begin{array}{cc}
-      \XHat' \OHat^{-1} \XHat & R' \\
+      \XHat^\top \OHat^{-1} \XHat & R^\top \\
       R & 0
@@
    \left[ \begin{array}{c}
-      \XHat' \OHat^{-1} y \\ q
+      \XHat^\top \OHat^{-1} y \\ q
    \end{array} \right] ,
@@
    \left[ \begin{array}{c}
-      \bHat^R_{3SLS} \\[0.2em] \lHat
+      \bHat^R \\[0.2em] \lHat
    \end{array} \right] 
@@
    \left[ \begin{array}{cc}
-      \XHat' \OHat^{-1} X & R' \\
+      \XHat^\top \OHat^{-1} X & R^\top \\
       R & 0
@@
    \left[ \begin{array}{cc}
-      \XHat' \OHat^{-1} \XHat & R' \\
+      \XHat^\top \OHat^{-1} X & R^\top \\
       R & 0
@@
    \left[ \begin{array}{cc}
-      X' Z \left( Z' \OHat Z \right)^{-1} Z' X & R' \\
+      X^\top Z \left( Z^\top \OHat Z \right)^{-1} Z^\top X & R^\top \\
       R & 0
@@
    \left[ \begin{array}{c}
-      X' Z \left( Z \OHat Z \right)^{-1} Z' y \\ q
+      X^\top Z \left( Z \OHat Z \right)^{-1} Z^\top y \\ q
    \end{array} \right] ,
@@
    \left[ \begin{array}{c}
-      \XHat' \OHat^{-1} Z \left( Z' Z \right)^{-1} Z' y \\ q
+      \XHat^\top \OHat^{-1} Z \left( Z^\top Z \right)^{-1} Z^\top y \\ q
    \end{array} \right] ,
@@
    \left[ \begin{array}{cc}
-      \XHat' \OHat^{-1} Z \left( Z' Z \right)^{-1} Z' \OHat
-      Z \left( Z' Z \right)^{-1} Z' \OHat^{-1} \XHat & 0' \\
+      \XHat^\top \OHat^{-1} Z \left( Z^\top Z \right)^{-1} Z^\top \OHat
+      Z \left( Z^\top Z \right)^{-1} Z^\top \OHat^{-1} \XHat & 0^\top \\
       0 & 0
@@
    \left[ \begin{array}{c}
-      \XHat' \OHat^{-1} \left( y - X \bHat^R_{2SLS} \right)
+      \XHat^\top \OHat^{-1} \left( y - X \bHat^R_{2SLS} \right)
       \\ q 
@@
 where $\bHat^R_{2SLS}$ is the restricted 2SLS estimator calculated
-by equation (\ref{eq:2sls-w2sls-3sls-r}).
+by Equation~\ref{eq:2sls-w2sls-3sls-r}.
 \proglang{EViews} uses the standard formula of the restricted 3SLS
-estimator~(\ref{eq:cov-2sls-w2sls-3sls-r}) to calculate an estimator
+estimator (Equation~\ref{eq:cov-2sls-w2sls-3sls-r}) to calculate an estimator
 for the covariance matrix of the estimated coefficients.

### 2007-12-27 paper/trunk/systemfit_estimation.tex
replaced single quotes (') around "fitted" by (automatically) directed double quotes (°) (2 times)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@915 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 
-At the first stage new ('fitted') regressors are obtained by
+At the first stage new (°fitted°) regressors are obtained by
 \begin{equation}
@@
 \end{equation}
-Then, these 'fitted' regressors are substituted for the original regressors
+Then, these °fitted° regressors are substituted for the original regressors
 in Equation~\ref{eq:ols-wls-sur}

### 2007-12-27 paper/trunk/systemfit_compare_lm.tex
using now always \codeD{}{} and \codeDD{}{}{} in place of \code{*.*} and \code{*.*.*}, respectively, to allow line breaks behind the dots
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@917 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 It is a list that contains one object for each estimated equation.
-These objects are of the class \code{systemfit.equation}
+These objects are of the class \codeD{systemfit}{equation}
 and contain the results that belong only to the regarding equation.
@@
 other counterparts are available for each equation
-in objects of class \code{systemfit.equation}.
+in objects of class \codeD{systemfit}{equation}.
 This is demonstrated in table~\ref{tab:compare-lm}.

### 2007-12-27 paper/trunk/systemfit_other.tex
improved formatting of some superscripts and subscripts: log -> \log; defined new command \tr for the trace operator (\mathrm); defined new command \LR for the likelihood-ratio statistic (\mathit); subscripts "2SLS", "3SLS", and "3SLS-*" are in \text{}mode now; superscripts "M", "R", and "MR" of \beta are in \text{}mode now; superscript "M" of "X" is in \text{}mode now
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@919 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
    \frac{ \uHat_i^\top \uHat_j }
-   { T - K_i - K_j + tr \left[ X_i \left( X_i^\top X_i \right)^{-1}
+   { T - K_i - K_j + \tr \left[ X_i \left( X_i^\top X_i \right)^{-1}
    X_i^\top X_j \left( X_j^\top X_j \right)^{-1} X_j^\top \right] }
@@
    \frac{ \uHat_i^\top \uHat_j }
-   { T - K_i - K_j + tr \left[ \left( X_i^\top X_i \right)^{-1}
+   { T - K_i - K_j + \tr \left[ \left( X_i^\top X_i \right)^{-1}
    X_i^\top X_j \left( X_j^\top X_j \right)^{-1} X_j^\top X_i \right] }
@@
 \begin{equation}
-LR = T \cdot \left(
-   log \left| \SHat_r \right|
-   - log \left| \SHat_u \right|
+\LR = T \cdot \left(
+   \log \left| \SHat_r \right|
+   - \log \left| \SHat_u \right|
    \right) ,
@@
 of the restricted and unrestricted estimation, respectively.
-Asymptotically, $LR$ has a $\chi^2$
+Asymptotically, $\LR$ has a $\chi^2$
 distribution with $j$ degrees of freedom
@@
 \begin{equation}
-  m = \left( \bHat_{2SLS} - \bHat_{3SLS} \right)^\top
-      \left( \COVHat \left[ \bHat_{2SLS} \right] -
-             \COVHat \left[ \bHat_{3SLS} \right] \right)
-      \left( \bHat_{2SLS} - \bHat_{3SLS} \right) ,
+  m = \left( \bHat_\text{2SLS} - \bHat_\text{3SLS} \right)^\top
+      \left( \COVHat \left[ \bHat_\text{2SLS} \right] -
+             \COVHat \left[ \bHat_\text{3SLS} \right] \right)
+      \left( \bHat_\text{2SLS} - \bHat_\text{3SLS} \right) ,
 \label{eq:hausman}
 \end{equation}
-where $\bHat_{2SLS}$ and $\COVHat \left[ \bHat_{2SLS} \right]$ are the estimated
+where $\bHat_\text{2SLS}$ and $\COVHat \left[ \bHat_\text{2SLS} \right]$ are the estimated
 coefficient and covariance matrix from a 2SLS estimation, and
-$\bHat_{3SLS}$ and $\COVHat \left[ \bHat_{3SLS} \right]$ are the estimated
+$\bHat_\text{3SLS}$ and $\COVHat \left[ \bHat_\text{3SLS} \right]$ are the estimated
 coefficients and covariance matrix from a 3SLS estimation.

### 2007-12-27 paper/trunk/systemfit_restrictions.tex
improved formatting of some superscripts and subscripts: log -> \log; defined new command \tr for the trace operator (\mathrm); defined new command \LR for the likelihood-ratio statistic (\mathit); subscripts "2SLS", "3SLS", and "3SLS-*" are in \text{}mode now; superscripts "M", "R", and "MR" of \beta are in \text{}mode now; superscript "M" of "X" is in \text{}mode now
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@919 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 \begin{equation}
-   \beta = M \cdot \beta^M \label{eq:T-restr} ,
+   \beta = M \cdot \beta^\text{M} \label{eq:T-restr} ,
 \end{equation}
-where $\beta^M$ is a vector of restricted (linear independent) coefficients,
+where $\beta^\text{M}$ is a vector of restricted (linear independent) coefficients,
 and $M$ is a matrix with the number of rows equal to the number of
@@
 the number of columns equal to the number of restricted coefficients
-($\beta^M$).
+($\beta^\text{M}$).
 $M$ can be used to map each unrestricted coefficient to one or more
@@
 \begin{equation}
-   R \beta^R = q ,
+   R \beta^\text{R} = q ,
    \label{eq:restr-R}
 \end{equation}
-where $\beta^R$ is the vector of the restricted coefficients,
+where $\beta^\text{R}$ is the vector of the restricted coefficients,
 and $R$ and $q$ are a matrix and vector, respectively,
@@
 \begin{equation}
-   R \beta^{MR} = q ,
+   R \beta^\text{MR} = q ,
 \end{equation}
-where $\beta^{MR}$ is the vector of the restricted $\beta^M$ coefficients.
+where $\beta^\text{MR}$ is the vector of the restricted $\beta^\text{M}$ coefficients.
 
@@
 \begin{equation}
-    X^M = X \cdot M .
+    X^\text{M} = X \cdot M .
 \end{equation}
-Then, $X^M$ is substituted for $X$ and a standard estimation as described
+Then, $X^\text{M}$ is substituted for $X$ and a standard estimation as described
 in the previous section is done
 (Equations~\ref{eq:ols-wls-sur}--\ref{eq:3slsEViews}).
-This results in the linear independent coefficient estimates $\bHat^M$ and
+This results in the linear independent coefficient estimates $\bHat^\text{M}$ and
 their covariance matrix.
@@
    \COVHat \left[ \bHat \right]
-   = M \cdot \COVHat \left[ \bHat^M \right] \cdot M^\top .
+   = M \cdot \COVHat \left[ \bHat^\text{M} \right] \cdot M^\top .
 \end{equation}
@@
 
-The OLS, WLS, and SUR estimators restricted by $R \beta^R = q$ can be obtained by
+The OLS, WLS, and SUR estimators restricted by $R \beta^\text{R} = q$ can be obtained by
 \begin{equation}
    \left[ \begin{array}{c}
-      \bHat^R \\[0.2em] \lHat
+      \bHat^\text{R} \\[0.2em] \lHat
    \end{array} \right]
@@
    \left[ \begin{array}{c}
-      \bHat^R \\[0.2em] \lHat
+      \bHat^\text{R} \\[0.2em] \lHat
    \end{array} \right] 
@@
 The 2SLS, W2SLS, and standard 3SLS estimators
-restricted by $R \beta^R = q$ can be obtained by
+restricted by $R \beta^\text{R} = q$ can be obtained by
 \begin{equation}
    \left[ \begin{array}{c}
-      \bHat^R \\[0.2em] \lHat
+      \bHat^\text{R} \\[0.2em] \lHat
    \end{array} \right]
@@
 
-The 3SLS-IV estimator restricted by $R \beta^R = q$ can be obtained by
+The 3SLS-IV estimator restricted by $R \beta^\text{R} = q$ can be obtained by
 \begin{equation}
    \left[ \begin{array}{c}
-      \bHat^R_{3SLS-IV} \\[0.2em] \lHat
+      \bHat^\text{R}_\text{3SLS-IV} \\[0.2em] \lHat
    \end{array} \right]
@@
    \left[ \begin{array}{c}
-      \bHat^R_{3SLS-IV} \\[0.2em] \lHat
+      \bHat^\text{R}_\text{3SLS-IV} \\[0.2em] \lHat
    \end{array} \right] 
@@
    \left[ \begin{array}{c}
-      \bHat^R_{3SLS-GMM} \\[0.2em] \lHat
+      \bHat^\text{R}_\text{3SLS-GMM} \\[0.2em] \lHat
    \end{array} \right]
@@
    \left[ \begin{array}{c}
-      \bHat^R_{3SLS-GMM} \\[0.2em] \lHat
+      \bHat^\text{R}_\text{3SLS-GMM} \\[0.2em] \lHat
    \end{array} \right] 
@@
    \left[ \begin{array}{c}
-      \bHat^R_{3SLS-Schmidt} \\[0.2em] \lHat
+      \bHat^\text{R}_\text{3SLS-Schmidt} \\[0.2em] \lHat
    \end{array} \right]
@@
    \left[ \begin{array}{c}
-      \bHat^R_{3SLS-Schmidt} \\[0.2em] \lHat
+      \bHat^\text{R}_\text{3SLS-Schmidt} \\[0.2em] \lHat
    \end{array} \right] 
@@
    \left[ \begin{array}{c}
-      \bHat^R_{3SLS-EViews} \\[0.2em] \lHat
+      \bHat^\text{R}_\text{3SLS-EViews} \\[0.2em] \lHat
    \end{array} \right]
@@
    \left[ \begin{array}{c}
-      \XHat^\top \OHat^{-1} \left( y - X \bHat^R_{2SLS} \right)
+      \XHat^\top \OHat^{-1} \left( y - X \bHat^\text{R}_\text{2SLS} \right)
       \\ q 
@@
 \end{equation}
-where $\bHat^R_{2SLS}$ is the restricted 2SLS estimator calculated
+where $\bHat^\text{R}_\text{2SLS}$ is the restricted 2SLS estimator calculated
 by Equation~\ref{eq:2sls-w2sls-3sls-r}.

### 2007-12-27 paper/trunk/systemfit_other.tex
using now always "p.~" instead of "p.\ "
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@924 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 However, these formulas yield unbiased estimators only if $K_i = K_j$
-\citep[p.\ 469]{judge85}. 
+\citep[p.~469]{judge85}.
 % Greene (2003, p. 344) says that the second is unbiased if i = j or K_i = K_j,
@@
 \end{eqnarray} 
-\citep[p.\ 309]{zellner62c}.
+\citep[p.~309]{zellner62c}.
 This yields an unbiased estimator for all elements of $\Sigma$,
@@
 its inverse $\SHat^{-1}$ is not an unbiased estimator
-of $\Sigma^{-1}$ \citep[p.\ 322]{theil71}.
+of $\Sigma^{-1}$ \citep[p.~322]{theil71}.
 Furthermore, the covariance matrix calculated by Equation~\ref{eq:rcovTheil}
-is not necessarily positive semidefinite \citep[p.\ 322]{theil71}. 
+is not necessarily positive semidefinite \citep[p.~322]{theil71}.
 Hence, °it is doubtful whether [this formula] is really superior to 
-[Equation~\ref{eq:rcovNoDfCor}]° \citep[p.\ 322]{theil71}.
+[Equation~\ref{eq:rcovNoDfCor}]° \citep[p.~322]{theil71}.
 
@@
 an iterated SUR estimation converges to maximum
-likelihood \citep[p.\ 345]{greene03}.
+likelihood \citep[p.~345]{greene03}.
 
@@
 McElroy's $R^2$ value 
-% also: \citep[p.\ 345]{greene03}
+% also: \citep[p.~345]{greene03}
 \begin{equation}
@@
 under the null hypothesis
-\citep[p.\ 347]{greene03}.
+\citep[p.~347]{greene03}.
 
@@
 under the null hypothesis
-\citep[p.\ 349]{greene03}.
+\citep[p.~349]{greene03}.
 

### 2007-12-27 paper/trunk/systemfit_estimation.tex
improved formatting of some superscripts and subscripts: log -> \log; defined new command \tr for the trace operator (\mathrm); defined new command \LR for the likelihood-ratio statistic (\mathit); subscripts "2SLS", "3SLS", and "3SLS-*" are in \text{}mode now; superscripts "M", "R", and "MR" of \beta are in \text{}mode now; superscript "M" of "X" is in \text{}mode now
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@919 591db2bd-e0a0-4d22-8c19-2192dbb7da31
When referring to sections, the word "Section" is capitalized now
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemf

@@
 that are calculated from the residuals of a first-step OLS estimation
-(see section~\ref{sec:residcov}).%
+(see Section~\ref{sec:residcov}).%
 \footnote{%
@@
 \begin{equation}
-   \bHat_{3SLS-IV} = \left( \XHat^\top \OHat^{-1} X
+   \bHat_\text{3SLS-IV} = \left( \XHat^\top \OHat^{-1} X
    \right)^{-1} \XHat^\top \OHat^{-1} y .
@@
 \begin{equation}
-   \COVHat \left[ \bHat_{3SLS-IV} \right] = \left( \XHat^\top \OHat^{-1}
+   \COVHat \left[ \bHat_\text{3SLS-IV} \right] = \left( \XHat^\top \OHat^{-1}
    X \right)^{-1} .
@@
 \begin{equation}
-   \bHat_{3SLS-GMM} = \left( X^\top Z \left( Z^\top \OHat Z \right)^{-1}
+   \bHat_\text{3SLS-GMM} = \left( X^\top Z \left( Z^\top \OHat Z \right)^{-1}
    Z^\top X \right)^{-1} X^\top Z \left( Z^\top \OHat Z \right)^{-1} Z^\top y
@@
 \begin{equation}
-   \COVHat \left[ \bHat_{3SLS-GMM} \right] =
+   \COVHat \left[ \bHat_\text{3SLS-GMM} \right] =
    \left( X^\top Z \left( Z^\top \OHat Z \right)^{-1} Z^\top X \right)^{-1} .
@@
 \begin{equation}
-   \bHat_{3SLS-Schmidt} = \left( \XHat^\top \OHat^{-1} \XHat
+   \bHat_\text{3SLS-Schmidt} = \left( \XHat^\top \OHat^{-1} \XHat
    \right)^{-1} \XHat^\top \OHat^{-1}
@@
 \begin{align}
-   \COVHat \left[ \bHat_{3SLS-Schmidt} \right] = &
+   \COVHat \left[ \bHat_\text{3SLS-Schmidt} \right] = &
    \left( \XHat^\top \OHat^{-1}  \XHat \right)^{-1}
@@
 \begin{equation}
-   \bHat_{3SLS-EViews} = \bHat_{2SLS} +
+   \bHat_\text{3SLS-EViews} = \bHat_\text{2SLS} +
    \left( \XHat^\top \OHat^{-1} \XHat \right)^{-1}
-   \XHat^\top \OHat^{-1} \left( y - X \bHat_{2SLS} \right) ,
+   \XHat^\top \OHat^{-1} \left( y - X \bHat_\text{2SLS} \right) ,
    \label{eq:3slsEViews}
 \end{equation}
-where $\bHat_{2SLS}$ is the two-stage least squares estimator
+where $\bHat_\text{2SLS}$ is the two-stage least squares estimator
 as defined above.

### 2007-12-27 paper/trunk/systemfit_restrictions.tex
When referring to sections, the word "Section" is capitalized now
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@931 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 where $\lambda$ is a vector of the Lagrangean multipliers of the restrictions
-and $\OHat$ is defined as in section~\ref{sec:Estimation-ols-wls-sur}.
+and $\OHat$ is defined as in Section~\ref{sec:Estimation-ols-wls-sur}.
 An estimator of the covariance matrix of the estimated coefficients is
@@
 \end{equation}
-where $\OHat$ is defined as in section~\ref{sec:Estimation-2sls-w2sls-3sls}.
+where $\OHat$ is defined as in Section~\ref{sec:Estimation-2sls-w2sls-3sls}.
 An estimator of the covariance matrix of the estimated coefficients is

### 2007-12-29 paper/trunk/systemfit_abstract.tex
modifications of Achim Zeileis (editor of the JSS) just before publication (mainly small cosmetic improvements)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@939 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 \Abstract{
-Many statistical analyses (e.g.\ in econometrics, biostatistics and
+Many statistical analyses (e.g., in econometrics, biostatistics and
 experimental design) are based on models containing systems of
@@
 
-\Keywords{R, system of simultaneous equations,
+\Keywords{\proglang{R}, system of simultaneous equations,
+  seemingly unrelated regression, two-stage least squares,
+  three-stage least squares, instrumental variables}
+\Plainkeywords{R, system of simultaneous equations,
   seemingly unrelated regression, two-stage least squares,

### 2007-12-29 paper/trunk/systemfit_compare_lm.tex
When referring to tables, the word "Table" is capitalized now
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@918 591db2bd-e0a0-4d22-8c19-2192dbb7da31
moved captions of tables *below* the table
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@923 591db2bd-e0a0-4d22-8c19-2192dbb7da31
modifications of Achim Zeileis (editor of the JSS) just before publication (mainly small cosmetic improvements)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@939 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 \begin{table}[htbp]
-\caption{Elements returned by \code{systemfit} and \code{lm}}
-\label{tab:compare-lm}
 \centering
@@
 {\ttfamily
-\begin{tabular}{lll}
-\hline  \hline
+\begin{tabular}{|l|l|l|}
+\hline
 \textbf{\code{lm}} & \textbf{\code{systemfit}} & \textbf{\code{systemfit.equation}} \\
@@
 offset          &                & \\
-model*          &                & model* \\
-x**             &                & x** \\
-y**             &                & y** \\
-                &                & z** \\
+model\textnormal{*} &            & model\textnormal{*} \\
+x\textnormal{**} &               & x\textnormal{**} \\
+y\textnormal{**} &               & y\textnormal{**} \\
+                &                & z\textnormal{**} \\
                 & iter           & \\
@@
                 & control        & \\
-\hline  \hline
+\hline
 \end{tabular}
 }
-\medskip\\
-Note:
-\texttt{*)} if requested by the user (default is \code{TRUE})\\
-\texttt{**)} if requested by the user (default is \code{FALSE})
+\caption{Elements returned by \code{systemfit} and \code{lm}
+(* if requested by the user with default \code{TRUE},
+** if requested by the user with default \code{FALSE}).}
+\label{tab:compare-lm}
 \end{table}
@@
 in objects of class \codeD{systemfit}{equation}.
-This is demonstrated in table~\ref{tab:compare-lm}.
+This is demonstrated in Table~\ref{tab:compare-lm}.

### 2007-12-29 paper/trunk/systemfit_intro.tex
modifications of Achim Zeileis (editor of the JSS) just before publication (mainly small cosmetic improvements)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@939 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 This estimation procedure is generally called
-°seemingly unrelated regression° (SUR) \citep{zellner62}.
+°seemingly unrelated regression° \citep[SUR,][]{zellner62}.
 Another reason to estimate a system of equations simultaneously are
@@
 Especially the economic theory suggests many cross-equation
-restrictions on the coefficients (e.g.\ the symmetry restriction in demand models).
+restrictions on the coefficients (e.g., the symmetry restriction in demand models).
 }
@@
 with several other statistical and econometric software packages
-(e.g.\ \proglang{SAS}, \proglang{EViews}, \proglang{TSP}),
+(e.g., \proglang{SAS}, \proglang{EViews}, \proglang{TSP}),
 \pkg{systemfit} has several advantages.

### 2007-12-29 paper/trunk/systemfit_other.tex
modifications of Achim Zeileis (editor of the JSS) just before publication (mainly small cosmetic improvements)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@939 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 For instance, this approach is applied by
-the command °TSCS° of the software \proglang{LIMDEP} that carries out SUR estimations
+the command \code{TSCS} of the software \proglang{LIMDEP} that carries out SUR estimations
 in which all coefficient vectors are constrained to be equal
@@
 but the 3SLS estimator is inconsistent,
-i.e.\ the instrumental variables of each equation are
+i.e., the instrumental variables of each equation are
 uncorrelated with the disturbances of the same equation

### 2008-03-31 paper/trunk/systemfit_short.tex
added short description of the package/paper for submission to useR! 2008
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@944 591db2bd-e0a0-4d22-8c19-2192dbb7da31

--- 2007-12-29 paper/trunk/systemfit_intro.tex
@@
-%       $Id$
+\documentclass[12pt]{scrartcl}
+\usepackage[utf8]{inputenc}
+\usepackage[T1]{fontenc}
+\usepackage{textcomp}
+\usepackage{lmodern}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Introduction}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\usepackage{geometry}
+\geometry{verbose, a4paper, tmargin=25mm, bmargin=30mm, lmargin=25mm,
+   rmargin=25mm, headheight=0mm, headsep=0mm, footskip=13mm}
+\setlength{\emergencystretch}{2em}
+\usepackage{setspace}
+\onehalfspacing
+
+\usepackage{csquotes}
+\MakeOuterQuote{°}
+
+\usepackage{natbib}
+\bibliographystyle{jss}
+
+\newcommand{\code}[1]{\texttt{#1}}
+\newcommand{\pkg}[1]{\mbox{\textbf{#1}}}
+\newcommand{\proglang}[1]{\mbox{\textsf{#1}}}
+\newcommand{\hz}{\hspace{0pt}}
+
+\author{Arne Henningsen and Jeff D.\ Hamann\\
+   \normalsize University of Kiel and Forest Informatics, Inc.}
+
+\title{\pkg{systemfit}: A Package for Estimating Systems
+of Simultaneous Equations in \proglang{R}}
+
+\date{March 31, 2008}
+
+\begin{document}
+\maketitle
 
@@
 Another reason to estimate a system of equations simultaneously are
-cross-equation restrictions on the coefficients.%
-\footnote{
-Especially the economic theory suggests many cross-equation
-restrictions on the coefficients (e.g., the symmetry restriction in demand models).
-}
+cross-equation restrictions on the coefficients.
 Estimating the coefficients under cross-equation restrictions
@@
 °three-stage least squares° (3SLS)
-are implemented.%
-\footnote{
-In this context,
-the term °weighted° in °weighted least squares° (WLS) and
-°weighted two-stage least squares° (W2SLS) means
-that the \emph{equations} might have different weights
-and \emph{not}
-that the \emph{observations} have different weights.
-}
+are implemented.
 The WLS, SUR, W2SLS, and 3SLS estimates can be based
@@
 
-% This paper is organized as follows:
-In Section~\ref{sec:statistics} we introduce the statistical
-background of estimating equation systems.
-The implementation of the statistical procedures in \proglang{R} is
-briefly explained in Section~\ref{sec:code}.
-Section~\ref{sec:Usage} demonstrates how to run
-\pkg{systemfit} and how some of the features presented in the second
-section can be used.
-In Section~\ref{sec:reliability} we replicate several textbook results
-with the \pkg{systemfit} package.
-Finally, a summary and outlook are presented in
-Section~\ref{sec:Summmary}.
+A special focus of our presentation at the useR! conference
+should be the changes and improvements of the \pkg{systemfit} package
+that were introduced in version~1.0
+(published in the very end of December 2007).
 
-%%% Local Variables: 
-%%% mode: latex
-%%% TeX-master: "systemfit"
-%%% End:
+\bibliography{systemfit}
+
+\end{document}

### 2008-09-12 paper/trunk/runSweave.sh
Sweave and Stangle are called in the "C" locale now (LC_ALL="C")
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@946 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@ -1,18 +1,18 @@
 #!/bin/sh
-echo "Sweave(\"systemfit_code.Snw\")" | R --no-save --no-restore
+echo "Sweave(\"systemfit_code.Snw\")" | LC_ALL="C" R --no-save --no-restore
 
-echo "Sweave(\"systemfit_usage.Snw\")" | R --no-save --no-restore
+echo "Sweave(\"systemfit_usage.Snw\")" | LC_ALL="C" R --no-save --no-restore
 echo "Stangle(\"systemfit_usage.Snw\", annotate=FALSE, split=TRUE,
-   prefix=FALSE)" | R --no-save --no-restore
+   prefix=FALSE)" | LC_ALL="C" R --no-save --no-restore
 
-echo "Sweave(\"systemfit_reliability.Snw\")" | R --no-save --no-restore
+echo "Sweave(\"systemfit_reliability.Snw\")" | LC_ALL="C" R --no-save --no-restore
 echo "Stangle(\"systemfit_reliability.Snw\", annotate=FALSE, split=TRUE,
-   prefix=FALSE)" | R --no-save --no-restore
+   prefix=FALSE)" | LC_ALL="C" R --no-save --no-restore
 
 echo "Sweave(\"systemfit_timings.Snw\")" | R --no-save --no-restore
 echo "Stangle(\"systemfit_timings.Snw\", annotate=FALSE, split=TRUE,
-   prefix=FALSE)" | R --no-save --no-restore
+   prefix=FALSE)" | LC_ALL="C" R --no-save --no-restore
 
 echo "Sweave(\"systemfit_sem.Snw\")" | R --no-save --no-restore
 echo "Stangle(\"systemfit_sem.Snw\", annotate=FALSE, split=TRUE,
-   prefix=FALSE)" | R --no-save --no-restore
+   prefix=FALSE)" | LC_ALL="C" R --no-save --no-restore

### 2008-09-12 paper/trunk/systemfit.Rnw
the master file (systemfit.tex) is an Sweave file now (systemfit.Rnw) (although it not yet contains R code but the subfiles should be moved into this master file soon)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@947 591db2bd-e0a0-4d22-8c19-2192dbb7da31

--- 2007-12-10 paper/trunk/systemfit.tex
@@
 \newcommand{\XHat}{\widehat{X}}
+\newcommand{\tr}{\mathrm{tr}}
+\newcommand{\LR}{\mathit{LR}}
 \clubpenalty=10000
@@
 
-\hyphenation{sys-tem-fit}
+\hyphenation{systemfit}
 
-\author{Arne Henningsen\\University of Kiel\And
-        Jeff D. Hamann\\ Forest Informatics, Inc. }
+\author{Arne Henningsen\\University of Kiel
+   \And Jeff D.\ Hamann\\ Forest Informatics, Inc.}
+\Plainauthor{Arne Henningsen, Jeff D. Hamann}
 
-\title{systemfit: A Package for Estimating\\
-       Systems of Simultaneous Equations in \proglang{R}}
-
-%% for pretty printing and a nice hypersummary also set:
-\Plainauthor{Arne Henningsen, Jeff D.\ Hamann} %% comma-separated
-\Plaintitle{systemfit: A Package for Estimating Systems of Simultaneous Equations
-            in R} %% without formatting
-\Shorttitle{systemfit} %% a short title (if necessary)
+\title{\pkg{systemfit}: A Package for Estimating Systems of Simultaneous Equations in \proglang{R}}
+\Plaintitle{systemfit: A Package for Estimating Systems of Simultaneous Equations in R}
+\Shorttitle{\pkg{systemfit}: Estimating Systems of Simultaneous Equations in \proglang{R}}
 
@@
 %% If it was not (yet) accepted, leave them commented.
-%% \Volume{13}
-%% \Issue{9}
-%% \Month{September}
-%% \Year{2006}
-%% \Submitdate{2006-02-28}
-%% \Acceptdate{2006-09-29}
+\Volume{23}
+\Issue{4}
+\Month{December}
+\Year{2007}
+\Submitdate{2006-03-15}
+\Acceptdate{2007-10-24}
 
@@
    PO Box 1421\\
-   Corvallis, Oregon 97339-1421\\
-   e-mail: \email{jeff.hamann@forestinformatics.com}\\
-   URL: \url{http://www.forestinformatics.com}\\
+   Corvallis, Oregon 97339-1421, United States of America\\
+   E-mail: \email{jeff.hamann@forestinformatics.com}\\
+   URL: \url{http://www.forestinformatics.com/}\\
 }
@@
 
-\appendix
+\clearpage
+
+\begin{appendix}
+
 \input{systemfit_compare_lm.tex}
@@
 \input{systemfit_sem.tex}
+
+\end{appendix}
 

### 2008-09-12 paper/trunk/runSweave.sh

@@ -1,4 +1,6 @@
 #!/bin/sh
+echo "Sweave(\"systemfit.Rnw\")" | LC_ALL="C" R --no-save --no-restore
+
 echo "Sweave(\"systemfit_code.Snw\")" | LC_ALL="C" R --no-save --no-restore
 
 echo "Sweave(\"systemfit_usage.Snw\")" | LC_ALL="C" R --no-save --no-restore

### 2008-09-12 paper/trunk/systemfit.Rnw
moved Abstract (systemfit_abstract.tex) into master file (systemfit.Rnw)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@948 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 %% an abstract and keywords
-\input{systemfit_abstract.tex}
+\Abstract{
+Many statistical analyses (e.g., in econometrics, biostatistics and
+experimental design) are based on models containing systems of
+structurally related equations.
+The \pkg{systemfit} package provides the capability
+to estimate systems of linear equations
+within the \proglang{R} programming environment.
+For instance, this package can be used for 
+``ordinary least squares'' (OLS),
+``seemingly unrelated regression'' (SUR),
+and the instrumental variable (IV) methods
+``two-stage least squares'' (2SLS) and
+``three-stage least squares'' (3SLS),
+where SUR and 3SLS estimations can optionally be iterated.
+Furthermore, the \pkg{systemfit} package provides tools
+for several statistical tests.
+It has been tested on a variety of datasets
+and its reliability is demonstrated.
+}
+
+\Keywords{\proglang{R}, system of simultaneous equations,
+  seemingly unrelated regression, two-stage least squares,
+  three-stage least squares, instrumental variables}
+\Plainkeywords{R, system of simultaneous equations,
+  seemingly unrelated regression, two-stage least squares,
+  three-stage least squares, instrumental variables}
+
+%% at least one keyword must be supplied
 

### 2008-09-12 paper/trunk/systemfit.Rnw
moved Introduction (systemfit_intro.tex) into master file (systemfit.Rnw)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@949 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 
-\input{systemfit_intro.tex}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Introduction}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Many theoretical models that are econometrically estimated
+consist of more than one equation.
+The disturbance terms of these equations are likely
+to be contemporaneously correlated,
+because unconsidered factors
+that influence the disturbance term in one equation
+probably influence the disturbance terms in other equations, too.
+Ignoring this contemporaneous correlation
+and estimating these equations separately
+leads to inefficient estimates of the coefficients.
+However, estimating all equations simultaneously
+with a °generalized least squares° (GLS) estimator,
+which takes the covariance structure of the residuals into account,
+leads to efficient estimates.
+This estimation procedure is generally called
+°seemingly unrelated regression° \citep[SUR,][]{zellner62}.
+Another reason to estimate a system of equations simultaneously are
+cross-equation restrictions on the coefficients.%
+\footnote{
+Especially the economic theory suggests many cross-equation
+restrictions on the coefficients (e.g., the symmetry restriction in demand models).
+}
+Estimating the coefficients under cross-equation restrictions
+and testing these restrictions
+requires a simultaneous estimation approach.
+
+Furthermore, these models can contain variables
+that appear on the left-hand side in one equation
+and on the right-hand side of another equation.
+Ignoring the endogeneity of these variables can lead to inconsistent
+estimates.
+This simultaneity bias can be corrected for by
+applying a °two-stage least squares° (2SLS) estimation
+to each equation.
+Combining this estimation method with the SUR method results
+in a simultaneous estimation of the system of equations
+by the °three-stage least squares° (3SLS) method
+\citep{zellner62b}.
+
+% For all of the methods developed in the package, the disturbances of
+% the individual equations are assumed to be independent and identically
+% distributed (iid).  
+% In the future, we would like to add the ability to fit equations were
+% the disturbances are serially correlated (wikins 1969).
+
+The \pkg{systemfit} package provides the capability to estimate
+systems of linear equations in \proglang{R}
+\citep{r-project}.
+Currently, the estimation methods
+°ordinary least squares° (OLS),
+°weighted least squares° (WLS),
+°seemingly unrelated regression° (SUR),
+°two-stage least squares° (2SLS),
+°weighted two-stage least squares° (W2SLS), and
+°three-stage least squares° (3SLS)
+are implemented.%
+\footnote{
+In this context,
+the term °weighted° in °weighted least squares° (WLS) and
+°weighted two-stage least squares° (W2SLS) means
+that the \emph{equations} might have different weights
+and \emph{not}
+that the \emph{observations} have different weights.
+}
+The WLS, SUR, W2SLS, and 3SLS estimates can be based
+either on one-step (OLS or 2SLS) (co)variances
+or these estimations can be iterated,
+where the (co)variances are calculated from the estimates of the previous step.
+Furthermore,
+the \pkg{systemfit} package
+provides statistical tests for restrictions on the coefficients
+and for testing the consistency of the 3SLS estimation.
+
+Although systems of linear equations can be estimated
+with several other statistical and econometric software packages
+(e.g., \proglang{SAS}, \proglang{EViews}, \proglang{TSP}),
+\pkg{systemfit} has several advantages.
+First, all estimation procedures are publicly available in the source code.
+Second, the estimation algorithms can be easily modified to meet specific
+requirements.
+Third, the (advanced) user can control estimation details generally
+not available in other software packages by overriding reasonable defaults.
+
+% This paper is organized as follows:
+In Section~\ref{sec:statistics} we introduce the statistical
+background of estimating equation systems.
+The implementation of the statistical procedures in \proglang{R} is
+briefly explained in Section~\ref{sec:code}.
+Section~\ref{sec:Usage} demonstrates how to run
+\pkg{systemfit} and how some of the features presented in the second
+section can be used.
+In Section~\ref{sec:reliability} we replicate several textbook results
+with the \pkg{systemfit} package.
+Finally, a summary and outlook are presented in
+Section~\ref{sec:Summmary}.
+
 \input{systemfit_statistics.tex}

### 2008-09-12 paper/trunk/systemfit.Rnw
moved section "Statistical background" (systemfit_statistics.tex) into master file (systemfit.Rnw)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@950 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 
-\input{systemfit_statistics.tex}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Statistical background}\label{sec:statistics}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+In this section we give a short overview of the statistical background
+that the \pkg{systemfit} package is based on.
+More detailed descriptions of simultaneous equations systems
+are available for instance in
+\citet[Chapter~7]{theil71},
+\citet[Part~4]{judge82},
+\citet[Part~5]{judge85},
+\citet{srivastava87},
+\citet[Chapters 14--15]{greene03}, and
+\citet[Chapter~10]{zivot06}.
+
+After introducing notations and assumptions,
+we provide the formulas to estimate systems of linear equations.
+We then demonstrate how to estimate coefficients under linear restrictions.
+Finally, we present additional relevant issues about estimation of
+equation systems.
+
+Consider a system of $G$ equations, where the $i$th equation is of
+the form 
+\begin{equation}
+   y_{i} = X_i \beta_i + u_i, \quad i = 1, 2, \ldots, G ,
+   \label{eq:model}
+\end{equation}
+where $y_i$ is a vector of the dependent variable,
+$X_i$ is a matrix of the exogenous variables,
+$\beta_i$ is the coefficient vector and
+$u_i$ is a vector of the disturbance terms of the $i$th equation.
+
+We can write the °stacked° system as
+\begin{equation}
+   \left[ \begin{array}{c}
+      y_1 \\ y_2\\ \vdots\\ y_G
+   \end{array} \right] = 
+   \left[ \begin{array}{cccc}
+      X_1 & 0 & \cdots & 0\\
+      0 & X_2 & \cdots & 0\\
+      \vdots & \vdots & \ddots & \vdots\\
+      0 & 0 & \cdots & X_G
+   \end{array}\right]
+   \left[ \begin{array}{c}
+      \beta_1 \\ \beta_2 \\ \vdots\\ \beta_G
+   \end{array} \right] +
+   \left[ \begin{array}{c}
+      u_1 \\ u_2 \\ \vdots\\ u_G
+   \end{array} \right]
+   \label{eq:model-array}
+\end{equation}
+
+or more simply as
+\begin{equation}
+   y = X \beta + u .
+   \label{eq:model-matrices}
+\end{equation}
+
+We assume that there is no correlation of the disturbance terms 
+across observations, so that
+\begin{equation}
+   \E \left[ u_{it} \, u_{js} \right] = 0
+   \; \forall \; t \neq s ,
+\end{equation}
+where $i$ and $j$ indicate the equation number 
+and $t$ and $s$ denote the observation number,
+where the number of observations is the same for all equations.
+
+However, we explicitly allow for contemporaneous correlation, i.e.,
+\begin{equation}
+   \E \left[ u_{it} \, u_{jt} \right] = \sigma_{ij} .
+\end{equation}
+
+Thus, the covariance matrix of all disturbances is
+\begin{equation}
+   \E \left[ u \, u^\top \right] = \Omega = \Sigma \otimes I_T ,
+\end{equation}
+where $\Sigma = \left[ \sigma_{ij} \right]$ is the (contemporaneous)
+disturbance covariance matrix,
+$\otimes$ is the Kronecker product,
+$I_T$ is an identity matrix of dimension $T$,
+and $T$ is the number of observations in each equation.
+
+\input{systemfit_estimation.tex}
+\input{systemfit_restrictions.tex}
+\input{systemfit_other.tex}
+
 \input{systemfit_code.tex}

### 2008-09-12 paper/trunk/systemfit.Rnw
moved subsections "Estimation with only exogenous regressors" and "Estimation with endogenous regressors" (systemfit_estimation.tex) into master file (systemfit.Rnw)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@951 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 
-\input{systemfit_estimation.tex}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Estimation with only exogenous regressors}
+\label{sec:Estimation-ols-wls-sur}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+If all regressors are exogenous,
+the system of equations (Equation~\ref{eq:model})
+can be consistently estimated by ordinary least squares (OLS),
+weighted least squares (WLS), and seemingly unrelated regression (SUR).
+These estimators can be obtained by
+\begin{equation}
+   \bHat = \left( X^\top \OHat^{-1} X \right)^{-1} X^\top \OHat^{-1} y .
+   \label{eq:ols-wls-sur}
+\end{equation}
+The covariance matrix of these estimators
+can be estimated by
+\begin{equation}
+   \COVHat \left[ \bHat \right] = \left( X^\top \OHat^{-1} X \right)^{-1} .
+   \label{eq:cov-ols-wls-sur}
+\end{equation}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Ordinary least squares (OLS)}
+
+The ordinary least squares (OLS) estimator is based on the assumption
+that the disturbance terms are not contemporaneously correlated
+$(\sigma_{ij} = 0 \; \forall \; i \neq j)$
+and have the same variance in each equation
+$( \sigma_i^2 = \sigma_j^2 \, \forall \, i, j)$.
+In this case, $\OHat$ in Equation~\ref{eq:ols-wls-sur}
+is equal to $I_{G \cdot T}$ and thus, cancels out.
+The OLS estimator is efficient,
+as long as the disturbances are not contemporaneously correlated.
+
+If the whole system is treated as one single equation,
+$\OHat$ in Equation~\ref{eq:cov-ols-wls-sur} is $\sHat^2 I_{G \cdot T}$,
+where $\sHat^2$ is an estimator for the variance of all disturbances
+$(\sigma^2 = \E [ u_{it}^2 ])$.
+If the disturbance terms of the individual equations
+are allowed to have different variances,
+$\OHat$ in Equation~\ref{eq:cov-ols-wls-sur} is $\SHat \otimes I_T$,
+where $\sHat_{ij} = 0 \; \forall \; i \neq j$ and
+$\sHat_{ii} = \sHat_i^2$ is the estimated variance
+of the disturbance term in the $i$th equation.
+
+If the estimated coefficients are not constrained by cross-equation restrictions,
+the simultaneous
+OLS estimation of the system leads to the same estimated coefficients
+as an equation-wise OLS estimation.
+The covariance matrix of the coefficients from an equation-wise
+OLS estimation is equal to the covariance matrix obtained
+by Equation~\ref{eq:cov-ols-wls-sur}
+with $\OHat$ equal to $\SHat \otimes I_T$.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Weighted least squares (WLS)}
+
+The weighted least squares (WLS) estimator allows for different variances
+of the disturbance terms in the different equations
+$( \sigma_i^2 \neq \sigma_j^2 \, \forall \, i \neq j)$,
+but assumes
+that the disturbance terms are not contemporaneously correlated.
+In this case, $\OHat$ in Equations~\ref{eq:ols-wls-sur}
+and~\ref{eq:cov-ols-wls-sur}
+is $\SHat \otimes I_T$,
+where $\sHat_{ij} = 0 \; \forall \; i \neq j$ and
+$\sHat_{ii} = \sHat_i^2$ is the estimated variance
+of the disturbance terms in the $i$th equation.
+Theoretically, $\sHat_{ii}$ should be the variance of
+the (true) disturbances $( \sigma_{ii} )$.
+However, they are not known in most empirical applications.
+Therefore, true variances are generally replaced by estimated variances
+$( \sHat_{ii} )$
+that are calculated from the residuals of a first-step OLS estimation
+(see Section~\ref{sec:residcov}).%
+\footnote{%
+Note that $\OHat$ in Equation~\ref{eq:ols-wls-sur}
+is not the same $\OHat$ as in Equation~\ref{eq:cov-ols-wls-sur}.
+The first is calculated from the residuals of a first-step OLS estimation;
+the second is calculated from the residuals of this WLS estimation.
+The same applies to the SUR, W2SLS, and 3SLS estimations
+described in the following sections.
+}
+
+The WLS estimator is (asymptotically) efficient only
+if the disturbance terms are not contemporaneously correlated.
+If the estimated coefficients are not constrained by cross-equation restrictions,
+they are equal to OLS estimates.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Seemingly unrelated regression (SUR)}
+
+If the disturbances are contemporaneously correlated, a generalized
+least squares (GLS) estimation leads to an efficient estimator for the coefficients.
+In this case, the GLS estimator is generally called °seemingly unrelated regression°
+(SUR) estimator \citep{zellner62}.
+However, the true covariance matrix of the disturbance terms
+is generally unknown.
+The textbook solution for this problem is
+a feasible generalized least squares (FGLS) estimation.
+As the FGLS estimator is based on
+an estimated covariance matrix of the disturbance terms,
+it is only asymptotically efficient.
+In case of a SUR estimator, $\OHat$ in Equations~\ref{eq:ols-wls-sur}
+and~\ref{eq:cov-ols-wls-sur}
+is $\SHat \otimes I_T$,
+where $\SHat$ is the estimated covariance matrix of the disturbance terms.
+
+It should be noted that while an unbiased OLS or WLS estimation requires only that
+the regressors and the disturbance terms of each single 
+equation are uncorrelated
+$( \E \left[ u_i^\top X_i \right] = 0 \; \forall \; i )$,
+a consistent SUR estimation requires that all disturbance terms and all 
+regressors are uncorrelated
+$( \E \left[ u_i^\top X_j \right] = 0 \; \forall \; i, j )$.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Estimation with endogenous regressors}
+\label{sec:Estimation-2sls-w2sls-3sls}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+If the regressors of one or more equations are correlated
+with the disturbances ($\E \left[ u_i^\top X_i \right] \neq 0$),
+OLS, WLS, and SUR estimates are biased.
+This can be circumvented by a
+two-stage least squares (2SLS),
+weighted two-stage least squares (W2SLS),
+or a three-stage least squares (3SLS) estimation
+with instrumental variables (IV).
+The instrumental variables for each equation $Z_i$ 
+can be either different or identical for all equations.
+They must not be correlated with
+the disturbance terms of the corresponding equation 
+($\E \left[ u_i^\top Z_i \right] = 0$).
+
+At the first stage new (°fitted°) regressors are obtained by
+\begin{equation}
+   \XHat_i = Z_i \left( Z_i^\top Z_i \right)^{-1} Z_i^\top X_i .
+\end{equation}
+Then, these °fitted° regressors are substituted for the original regressors
+in Equation~\ref{eq:ols-wls-sur}
+to obtain unbiased 2SLS, W2SLS, or 3SLS estimates of $\beta$ by
+\begin{equation}
+   \bHat
+   = \left( \XHat^\top \OHat^{-1} \XHat \right)^{-1} \XHat^\top \OHat^{-1} y ,
+   \label{eq:2sls-w2sls-3sls}
+\end{equation}
+where
+\begin{equation}
+   \XHat =
+   \left[ \begin{array}{cccc}
+      \XHat_1 & 0       & \cdots & 0\\
+      0       & \XHat_2 & \cdots & 0\\
+      \vdots  & \vdots  & \ddots & \vdots\\
+      0       & 0       & \cdots & \XHat_G
+   \end{array}\right] .
+\end{equation}
+An estimator of the covariance matrix of the estimated coefficients
+can be obtained from Equation~\ref{eq:cov-ols-wls-sur} analogously.
+Hence, we get
+\begin{equation}
+   \COVHat \left[ \bHat \right]
+   = \left( \XHat^\top \OHat^{-1} \XHat \right)^{-1} .
+   \label{eq:cov-2sls-w2sls-3sls}
+\end{equation}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Two-stage least squares (2SLS)}
+
+The two-stage least squares (2SLS) estimator
+is based on the same assumptions about the disturbance terms as the OLS estimator.
+Accordingly, $\OHat$ in Equation~\ref{eq:2sls-w2sls-3sls}
+is equal to $I_{G \cdot T}$ and thus, cancels out.
+Like for the OLS estimator,
+the whole system can be treated either as one single equation
+with $\OHat$ in Equation~\ref{eq:cov-2sls-w2sls-3sls}
+equal to $\sHat^2 I_{G \cdot T}$,
+or the disturbance terms of the individual equations
+are allowed to have different variances
+with $\OHat$ in Equation~\ref{eq:cov-2sls-w2sls-3sls}
+equal to $\SHat \otimes I_T$,
+where $\sHat_{ij} = 0 \; \forall \; i \neq j$ and
+$\sHat_{ii} = \sHat_i^2$.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Weighted two-stage least squares (W2SLS)}
+
+The weighted two-stage least squares (W2SLS) estimator allows
+for different variances of the disturbance terms in the different equations.
+Hence, $\OHat$ in Equations~\ref{eq:2sls-w2sls-3sls}
+and~\ref{eq:cov-2sls-w2sls-3sls}
+is $\SHat \otimes I_T$,
+where $\sHat_{ij} = 0 \; \forall \; i \neq j$ and
+$\sHat_{ii} = \sHat_i^2$.
+If the estimated coefficients are not constrained by cross-equation restrictions,
+they are equal to 2SLS estimates.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Three-stage least squares (3SLS)}
+
+If the disturbances are contemporaneously correlated,
+a feasible generalized least squares (FGLS) version of the two-stage least squares
+estimation leads to consistent and asymptotically more efficient estimates.
+This estimation procedure is generally called °three-stage least
+squares° \citep[3SLS,][]{zellner62b}.
+The standard 3SLS estimator and its covariance matrix are obtained by
+Equations~\ref{eq:2sls-w2sls-3sls} and~\ref{eq:cov-2sls-w2sls-3sls}
+with $\OHat$ equal to $\SHat \otimes I_T$,
+where $\SHat$ is the estimated covariance matrix of the disturbance terms.
+
+While an unbiased 2SLS or W2SLS estimation requires only that
+the instrumental variables and the disturbance terms of each single 
+equation are uncorrelated
+$( \E \left[ u_i^\top Z_i \right]) = 0 \; \forall \; i )$,
+\cite{schmidt90} points out that the 3SLS estimator is only consistent
+if all disturbance terms and all instrumental variables are uncorrelated
+$( \E \left[ u_i^\top Z_j \right]) = 0 \; \forall \; i, j )$.
+Since there might be occasions where this cannot be avoided,
+\cite{schmidt90} analyses other approaches to obtain 3SLS estimators.
+
+One of these approaches based on instrumental variable estimation
+(3SLS-IV) is
+\begin{equation}
+   \bHat_\text{3SLS-IV} = \left( \XHat^\top \OHat^{-1} X
+   \right)^{-1} \XHat^\top \OHat^{-1} y .
+   \label{eq:3slsIv}
+\end{equation}
+An estimator of the covariance matrix of the estimated 3SLS-IV coefficients is
+\begin{equation}
+   \COVHat \left[ \bHat_\text{3SLS-IV} \right] = \left( \XHat^\top \OHat^{-1}
+   X \right)^{-1} .
+\end{equation}
+Another approach based on the generalized method of moments (GMM)
+estimator (3SLS-GMM) is
+\begin{equation}
+   \bHat_\text{3SLS-GMM} = \left( X^\top Z \left( Z^\top \OHat Z \right)^{-1}
+   Z^\top X \right)^{-1} X^\top Z \left( Z^\top \OHat Z \right)^{-1} Z^\top y
+   \label{eq:3slsGmm}
+\end{equation}
+with
+\begin{equation}
+   Z =
+   \left[ \begin{array}{cccc}
+      Z_1 & 0 & \cdots & 0\\
+      0 & Z_2 & \cdots & 0\\
+      \vdots & \vdots & \ddots & \vdots\\
+      0 & 0 & \cdots & Z_G
+   \end{array}\right] .
+\end{equation}
+An estimator of the covariance matrix of the estimated 3SLS-GMM coefficients is
+\begin{equation}
+   \COVHat \left[ \bHat_\text{3SLS-GMM} \right] =
+   \left( X^\top Z \left( Z^\top \OHat Z \right)^{-1} Z^\top X \right)^{-1} .
+\end{equation}
+A fourth approach developed by \cite{schmidt90} himself is
+\begin{equation}
+   \bHat_\text{3SLS-Schmidt} = \left( \XHat^\top \OHat^{-1} \XHat
+   \right)^{-1} \XHat^\top \OHat^{-1}
+   Z \left( Z^\top Z \right)^{-1} Z^\top y .
+   \label{eq:3slsSchmidt}
+\end{equation}
+An estimator of the covariance matrix of these estimated coefficients is
+\begin{align}
+   \COVHat \left[ \bHat_\text{3SLS-Schmidt} \right] = &
+   \left( \XHat^\top \OHat^{-1}  \XHat \right)^{-1}
+   \XHat^\top \OHat^{-1} Z \left( Z^\top Z \right)^{-1} Z^\top \OHat Z \\
+   & \left( Z^\top Z \right)^{-1} Z^\top \OHat^{-1} \XHat
+   \left( \XHat^\top \OHat^{-1}  \XHat \right)^{-1} .
+   \nonumber
+\end{align}
+The econometrics software \proglang{EViews} uses
+\begin{equation}
+   \bHat_\text{3SLS-EViews} = \bHat_\text{2SLS} +
+   \left( \XHat^\top \OHat^{-1} \XHat \right)^{-1}
+   \XHat^\top \OHat^{-1} \left( y - X \bHat_\text{2SLS} \right) ,
+   \label{eq:3slsEViews}
+\end{equation}
+where $\bHat_\text{2SLS}$ is the two-stage least squares estimator
+as defined above.
+\proglang{EViews} uses the standard 3SLS formula
+(Equation~\ref{eq:cov-2sls-w2sls-3sls}) to
+calculate an estimator of the covariance matrix of the estimated coefficients.
+
+If the same instrumental variables are used in all equations 
+($Z_1 = Z_2 = \ldots = Z_G$),
+all the above mentioned approaches lead to identical estimates.
+However, if this is not the case, the results depend on the 
+method used \citep{schmidt90}.
+The only reason to use different instruments for different equations
+is a correlation of the instruments of one equation with the
+disturbance terms of another equation.
+Otherwise, one could simply use all instruments in every equation
+\citep{schmidt90}.
+In this case, only the 3SLS-GMM (Equation~\ref{eq:3slsGmm})
+and the 3SLS estimator developed by \cite{schmidt90} 
+(Equation~\ref{eq:3slsSchmidt}) are consistent.
+
 \input{systemfit_restrictions.tex}

### 2008-09-12 paper/trunk/systemfit.Rnw
moved subsection "Estimation under linear restrictions on the coefficients" (systemfit_restrictions.tex) into master file (systemfit.Rnw)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@952 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 
-\input{systemfit_restrictions.tex}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Estimation under linear restrictions on the coefficients}
+\label{sec:Restrictions}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+In many empirical applications,
+it is desirable to estimate the coefficients under linear restrictions.
+For instance, in econometric demand and production analysis,
+it is common to estimate the coefficients
+under homogeneity and symmetry restrictions
+that are derived from the underlying theoretical model.
+
+There are two different methods to estimate the coefficients
+under linear restrictions.
+First, a matrix $M$ can be specified that
+\begin{equation}
+   \beta = M \cdot \beta^\text{M} \label{eq:T-restr} ,
+\end{equation}
+where $\beta^\text{M}$ is a vector of restricted (linear independent) coefficients,
+and $M$ is a matrix with the number of rows equal to the number of
+unrestricted coefficients ($\beta$) and
+the number of columns equal to the number of restricted coefficients
+($\beta^\text{M}$).
+$M$ can be used to map each unrestricted coefficient to one or more
+restricted coefficients.
+
+The second method to estimate the coefficients under linear restrictions
+constrains the coefficients by
+\begin{equation}
+   R \beta^\text{R} = q ,
+   \label{eq:restr-R}
+\end{equation}
+where $\beta^\text{R}$ is the vector of the restricted coefficients,
+and $R$ and $q$ are a matrix and vector, respectively,
+that specify the restrictions \citep[see][p.~100]{greene03}.
+Each linear independent restriction is represented by one row of $R$
+and the corresponding element of~$q$.
+
+The first method is less flexible than the second%
+\footnote{
+While restrictions like $\beta_1 = 2 \beta_2$ can be specified by
+both methods,
+restrictions like $\beta_1 + \beta_2 = 4$ can be specified only
+by the second method.
+}, 
+but is preferable if the coefficients are estimated
+under many equality constraints across different equations of the system.
+Of course, these restrictions can be also specified using
+the latter method.
+However, while the latter method increases the dimension of the 
+matrices to be inverted during estimation, the first reduces it. 
+Thus, in some cases the latter way leads to estimation problems
+(e.g., (near) singularity of the matrices to be inverted),
+while the first does not.
+
+These two methods can be combined.
+In this case, the restrictions specified using the latter method are
+imposed on the linear independent coefficients
+that are restricted by the first method, so that
+\begin{equation}
+   R \beta^\text{MR} = q ,
+\end{equation}
+where $\beta^\text{MR}$ is the vector of the restricted $\beta^\text{M}$ coefficients.
+
+\subsubsection{Calculation of restricted estimators}
+
+If the first method (Equation~\ref{eq:T-restr}) is chosen
+to estimate the coefficients under these restrictions,
+the matrix of regressors $X$ is (post-)\hspace{0pt}multiplied
+by the $M$ matrix,
+so that
+\begin{equation}
+    X^\text{M} = X \cdot M .
+\end{equation}
+Then, $X^\text{M}$ is substituted for $X$ and a standard estimation as described
+in the previous section is done
+(Equations~\ref{eq:ols-wls-sur}--\ref{eq:3slsEViews}).
+This results in the linear independent coefficient estimates $\bHat^\text{M}$ and
+their covariance matrix.
+The original coefficients can be obtained by Equation~\ref{eq:T-restr}
+and the estimated covariance matrix of the original coefficients
+can be obtained by
+\begin{equation}
+   \COVHat \left[ \bHat \right]
+   = M \cdot \COVHat \left[ \bHat^\text{M} \right] \cdot M^\top .
+\end{equation}
+
+The implementation of the second method to estimate the coefficients
+under linear restrictions (Equation~\ref{eq:restr-R})
+is described for each estimation method in the following sections.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Restricted OLS, WLS, and SUR estimation}
+
+The OLS, WLS, and SUR estimators restricted by $R \beta^\text{R} = q$ can be obtained by
+\begin{equation}
+   \left[ \begin{array}{c}
+      \bHat^\text{R} \\[0.2em] \lHat
+   \end{array} \right]
+   =
+   \left[ \begin{array}{cc}
+      X^\top \OHat^{-1} X & R^\top \\
+      R & 0
+   \end{array} \right]^{-1}
+   \cdot
+   \left[ \begin{array}{c}
+      X^\top \OHat^{-1} y \\ q
+   \end{array} \right] ,
+   \label{eq:ols-wls-sur-r}
+\end{equation}
+where $\lambda$ is a vector of the Lagrangean multipliers of the restrictions
+and $\OHat$ is defined as in Section~\ref{sec:Estimation-ols-wls-sur}.
+An estimator of the covariance matrix of the estimated coefficients is
+\begin{equation}
+   \COVHat
+   \left[ \begin{array}{c}
+      \bHat^\text{R} \\[0.2em] \lHat
+   \end{array} \right] 
+   = 
+   \left[ \begin{array}{cc}
+      X^\top \OHat^{-1} X & R^\top \\
+      R & 0
+   \end{array} \right]^{-1} .
+   \label{eq:cov-ols-wls-sur-r}
+\end{equation}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Restricted 2SLS, W2SLS, and 3SLS estimation}
+
+The 2SLS, W2SLS, and standard 3SLS estimators
+restricted by $R \beta^\text{R} = q$ can be obtained by
+\begin{equation}
+   \left[ \begin{array}{c}
+      \bHat^\text{R} \\[0.2em] \lHat
+   \end{array} \right]
+   =
+   \left[ \begin{array}{cc}
+      \XHat^\top \OHat^{-1} \XHat & R^\top \\
+      R & 0
+   \end{array} \right]^{-1}
+   \cdot
+   \left[ \begin{array}{c}
+      \XHat^\top \OHat^{-1} y \\ q
+   \end{array} \right] ,
+   \label{eq:2sls-w2sls-3sls-r}
+\end{equation}
+where $\OHat$ is defined as in Section~\ref{sec:Estimation-2sls-w2sls-3sls}.
+An estimator of the covariance matrix of the estimated coefficients is
+\begin{equation}
+   \COVHat
+   \left[ \begin{array}{c}
+      \bHat^\text{R} \\[0.2em] \lHat
+   \end{array} \right] 
+   = 
+   \left[ \begin{array}{cc}
+      \XHat^\top \OHat^{-1} \XHat & R^\top \\
+      R & 0
+   \end{array} \right]^{-1} .
+   \label{eq:cov-2sls-w2sls-3sls-r}
+\end{equation}
+
+The 3SLS-IV estimator restricted by $R \beta^\text{R} = q$ can be obtained by
+\begin{equation}
+   \left[ \begin{array}{c}
+      \bHat^\text{R}_\text{3SLS-IV} \\[0.2em] \lHat
+   \end{array} \right]
+   =
+   \left[ \begin{array}{cc}
+      \XHat^\top \OHat^{-1} X & R^\top \\
+      R & 0
+   \end{array} \right]^{-1}
+   \cdot
+   \left[ \begin{array}{c}
+      \XHat^\top \OHat^{-1} y \\ q
+   \end{array} \right] ,
+   \label{eq:3slsIvR}
+\end{equation}
+where
+\begin{equation}
+   \COVHat
+   \left[ \begin{array}{c}
+      \bHat^\text{R}_\text{3SLS-IV} \\[0.2em] \lHat
+   \end{array} \right] 
+   = 
+   \left[ \begin{array}{cc}
+      \XHat^\top \OHat^{-1} X & R^\top \\
+      R & 0
+   \end{array} \right]^{-1} .
+\end{equation}
+The restricted 3SLS-GMM estimator can be obtained by
+\begin{equation}
+   \left[ \begin{array}{c}
+      \bHat^\text{R}_\text{3SLS-GMM} \\[0.2em] \lHat
+   \end{array} \right]
+   =
+   \left[ \begin{array}{cc}
+      X^\top Z \left( Z^\top \OHat Z \right)^{-1} Z^\top X & R^\top \\
+      R & 0
+   \end{array} \right]^{-1}
+   \cdot
+   \left[ \begin{array}{c}
+      X^\top Z \left( Z \OHat Z \right)^{-1} Z^\top y \\ q
+   \end{array} \right] ,
+   \label{eq:3slsGmmR}
+\end{equation}
+where
+\begin{equation}
+   \COVHat
+   \left[ \begin{array}{c}
+      \bHat^\text{R}_\text{3SLS-GMM} \\[0.2em] \lHat
+   \end{array} \right] 
+   = 
+   \left[ \begin{array}{cc}
+      X^\top Z \left( Z^\top \OHat Z \right)^{-1} Z^\top X & R^\top \\
+      R & 0
+   \end{array} \right]^{-1} .
+\end{equation}
+The restricted 3SLS estimator based on the suggestion of
+\cite{schmidt90} is
+\begin{equation}
+   \left[ \begin{array}{c}
+      \bHat^\text{R}_\text{3SLS-Schmidt} \\[0.2em] \lHat
+   \end{array} \right]
+   =
+   \left[ \begin{array}{cc}
+      \XHat^\top \OHat^{-1} \XHat & R^\top \\
+      R & 0
+   \end{array} \right]^{-1}
+   \cdot
+   \left[ \begin{array}{c}
+      \XHat^\top \OHat^{-1} Z \left( Z^\top Z \right)^{-1} Z^\top y \\ q
+   \end{array} \right] ,
+   \label{eq:3slsSchmidtR}
+\end{equation}
+where
+\begin{eqnarray}
+   \COVHat
+   \left[ \begin{array}{c}
+      \bHat^\text{R}_\text{3SLS-Schmidt} \\[0.2em] \lHat
+   \end{array} \right] 
+   & = & 
+   \left[ \begin{array}{cc}
+      \XHat^\top \OHat^{-1} \XHat & R^\top \\
+      R & 0
+   \end{array} \right]^{-1}
+   \\
+   & & \cdot
+   \left[ \begin{array}{cc}
+      \XHat^\top \OHat^{-1} Z \left( Z^\top Z \right)^{-1} Z^\top \OHat
+      Z \left( Z^\top Z \right)^{-1} Z^\top \OHat^{-1} \XHat & 0^\top \\
+      0 & 0
+   \end{array} \right]^{-1}
+   \nonumber \\
+   & & \cdot
+   \left[ \begin{array}{cc}
+      \XHat^\top \OHat^{-1} \XHat & R^\top \\
+      R & 0
+   \end{array} \right]^{-1} .
+   \nonumber
+\end{eqnarray}
+The econometrics software \proglang{EViews} calculates the restricted 3SLS estimator by
+\begin{equation}
+   \left[ \begin{array}{c}
+      \bHat^\text{R}_\text{3SLS-EViews} \\[0.2em] \lHat
+   \end{array} \right]
+   =
+   \left[ \begin{array}{cc}
+      \XHat^\top \OHat^{-1} \XHat & R^\top \\
+      R & 0
+   \end{array} \right]^{-1}
+   \cdot
+   \left[ \begin{array}{c}
+      \XHat^\top \OHat^{-1} \left( y - X \bHat^\text{R}_\text{2SLS} \right)
+      \\ q 
+   \end{array} \right] ,
+   \label{eq:3slsEViewsR}
+\end{equation}
+where $\bHat^\text{R}_\text{2SLS}$ is the restricted 2SLS estimator calculated
+by Equation~\ref{eq:2sls-w2sls-3sls-r}.
+\proglang{EViews} uses the standard formula of the restricted 3SLS
+estimator (Equation~\ref{eq:cov-2sls-w2sls-3sls-r}) to calculate an estimator
+for the covariance matrix of the estimated coefficients.
+
+If the same instrumental variables are used in all equations 
+($Z_1 = Z_2 = \ldots = Z_G$),
+all the above mentioned approaches lead to identical coefficient estimates
+and identical covariance matrices of the estimated coefficients.
+
 \input{systemfit_other.tex}

### 2008-09-12 paper/trunk/systemfit.Rnw
moved subsections "Residual covariance matrix", "Degrees of freedom", "Goodness of fit", "Testing linear restrictions", and "Hausman test" (systemfit_other.tex) into master file (systemfit.Rnw)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@953 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 
-\input{systemfit_other.tex}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\section{Other issues and tools}\label{sec:Other}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Residual covariance matrix}\label{sec:residcov}
+
+Since the (true) disturbances ($u_i$) of the estimated equations
+are generally not known,
+their covariance matrix cannot be determined.
+Therefore, this covariance matrix is generally calculated from
+estimated residuals ($\uHat_i$)
+that are obtained from a first-step OLS or 2SLS estimation.
+Then, in a second step, the estimated residual covariance matrix
+can be employed for a WLS, SUR, W2SLS, or 3SLS estimation.
+In many cases, the residual covariance matrix is calculated by
+\begin{equation}
+   \sHat_{ij} = \frac{ \uHat_i^\top \uHat_j }{ T },
+   \label{eq:rcovNoDfCor}
+\end{equation}
+where $T$ is the number of observations in each equation.
+However, in finite samples this estimator is biased,
+because it is not corrected for degrees of freedom.
+The usual single-equation procedure to correct for degrees of freedom
+cannot always be applied, because the number of regressors in each equation
+might differ.
+Two alternative approaches to calculate the residual covariance
+matrix are
+\begin{equation}
+   \sHat_{ij} = \frac{ \uHat_i^\top \uHat_j }
+   { \sqrt{ \left( T - K_i \right) \cdot \left( T - K_j \right) } }
+   \label{eq:rcovGeomean}
+\end{equation}
+and
+\begin{equation}
+   \sHat_{ij} = \frac{ \uHat_i^\top \uHat_j }
+   { T - \max \left( K_i , K_j \right) } \; ,
+   \label{eq:rcovMax}
+\end{equation}
+where $K_i$ and $K_j$ are the number of regressors in equation
+$i$ and $j$, respectively.
+However, these formulas yield unbiased estimators only if $K_i = K_j$
+\citep[p.~469]{judge85}.
+% Greene (2003, p. 344) says that the second is unbiased if i = j or K_i = K_j,
+% whereas the first is unbiased only if i = j. 
+% However, if K_i = K_j the first and the second are equal.
+% Why is the first biased if K_i = K_j ???????????
+
+A further approach to obtain a residual covariance
+matrix is
+\begin{eqnarray}
+   \sHat_{ij} & = &
+   \frac{ \uHat_i^\top \uHat_j }
+   { T - K_i - K_j + \tr \left[ X_i \left( X_i^\top X_i \right)^{-1}
+   X_i^\top X_j \left( X_j^\top X_j \right)^{-1} X_j^\top \right] }
+   \label{eq:rcovTheil} \\
+   & = &
+   \frac{ \uHat_i^\top \uHat_j }
+   { T - K_i - K_j + \tr \left[ \left( X_i^\top X_i \right)^{-1}
+   X_i^\top X_j \left( X_j^\top X_j \right)^{-1} X_j^\top X_i \right] }
+\end{eqnarray} 
+\citep[p.~309]{zellner62c}.
+This yields an unbiased estimator for all elements of $\Sigma$,
+but even if $\SHat$ is an unbiased estimator of $\Sigma$,
+its inverse $\SHat^{-1}$ is not an unbiased estimator
+of $\Sigma^{-1}$ \citep[p.~322]{theil71}.
+Furthermore, the covariance matrix calculated by Equation~\ref{eq:rcovTheil}
+is not necessarily positive semidefinite \citep[p.~322]{theil71}.
+Hence, °it is doubtful whether [this formula] is really superior to 
+[Equation~\ref{eq:rcovNoDfCor}]° \citep[p.~322]{theil71}.
+
+The WLS, SUR, W2SLS and 3SLS coefficient estimates are consistent
+if the residual covariance matrix is calculated
+using the residuals from a first-step OLS or 2SLS estimation.
+There exists also an alternative slightly different approach
+that consists of three steps.%
+\footnote{
+For instance, this approach is applied by
+the command \code{TSCS} of the software \proglang{LIMDEP} that carries out SUR estimations
+in which all coefficient vectors are constrained to be equal
+\citep{greene06}.
+}
+In a first step, an OLS or 2SLS estimation is applied to obtain residuals
+to calculate a (first-step) residual covariance matrix.
+In a second step, the first-step residual covariance matrix
+is used to estimate the model by WLS or W2SLS
+and new residuals are obtained to calculate
+a (second-step) residual covariance matrix.
+Finally, in the third step,
+the second-step residual covariance matrix is used
+to estimate the model by SUR or 3SLS.
+If the estimated coefficients are not constrained by cross-equation restrictions,
+OLS and WLS estimates as well as 2SLS and W2SLS estimates are identical.
+Hence, in this case both approaches generate the same results.
+
+It is also possible to iterate WLS, SUR, W2SLS and 3SLS estimations.
+At each iteration the residual covariance matrix is calculated
+from the residuals of the previous iteration.
+If Equation~\ref{eq:rcovNoDfCor} is applied to calculate the
+residual covariance matrix,
+an iterated SUR estimation converges to maximum
+likelihood \citep[p.~345]{greene03}.
+
+In some uncommon cases,
+for instance in pooled estimations,
+where the coefficients are restricted to be equal in all equations,
+the means of the residuals of each equation are not equal to zero
+$( \overline{ \uHat }_i \neq 0 )$.
+Therefore, it might be argued
+that the residual covariance matrix should be calculated
+by subtracting the means from the residuals
+and substituting $\uHat_i - \overline{ \uHat }_i$
+for $\uHat_i$ in Equations~\ref{eq:rcovNoDfCor}--\ref{eq:rcovTheil}.
+
+If the coefficients are estimated under any restrictions,
+the residual covariance matrix
+for a WLS, SUR, W2SLS, or 3SLS estimation
+can be obtained either from a restricted or from an unrestricted
+first-step estimation.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Degrees of freedom}
+\label{sec:degreesOfFreedom}
+
+To our knowledge the question about how to determine the degrees
+of freedom for single-coefficient $t$ tests is not comprehensively
+discussed in the literature.
+While sometimes the degrees of freedom of the entire system
+(total number of observations in all equations minus
+total number of estimated coefficients)
+are applied,
+in other cases the degrees of freedom of each single equation
+(number of observations in the equations minus
+number of estimated coefficients in the equation)
+are used.
+Asymptotically, this distinction does not make a difference.
+However, in many empirical applications, the number of observations
+of each equation is rather small, and
+therefore, it matters.
+
+If a system of equations is estimated by an unrestricted OLS and
+the covariance matrix of the coefficients is calculated
+with $\OHat$ in Equation~\ref{eq:cov-ols-wls-sur}
+equal to $\SHat \otimes I_T$,
+the estimated coefficients and their standard errors are identical
+to an equation-wise OLS estimation.
+In this case, it is reasonable to use the degrees of freedom of
+each single equation,
+because this yields the same $P$ values as the equation-wise
+OLS estimation.
+
+In contrast, if a system of equations is estimated with many
+cross-equation restrictions and
+the covariance matrix of an OLS estimation is calculated
+with $\OHat$ in Equation~\ref{eq:cov-ols-wls-sur}
+equal to $\sHat^2 I_{G \cdot T}$,
+the system estimation is similar to a single equation estimation.
+Therefore, in this case, it seems to be reasonable to use the degrees
+of freedom of the entire system.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Goodness of fit}
+
+The goodness of fit of each single equation can be measured by the
+traditional $R^2$ values
+\begin{equation}
+   R_i^2 = 1 - \frac{ \uHat_i^\top \uHat_i }
+   { ( y_i - \overline{y_i} )^\top ( y_i - \overline{y_i} ) } \; ,
+\end{equation}
+where $R_i^2$ is the $R^2$ value of the $i$th equation
+and $\overline{y_i}$ is the mean value of $y_i$.
+
+The goodness of fit of the whole system can be measured by the
+McElroy's $R^2$ value 
+% also: \citep[p.~345]{greene03}
+\begin{equation}
+   R_*^2 = 1 - \frac{ \uHat^\top \OHat^{-1} \uHat }
+   { y^\top \left( \SHat^{-1} \otimes
+   \left( I_T - \frac{\iota \iota^\top}{T} \right) \right) y } ,
+\end{equation}
+where $\iota$ is a column vector of $T$ ones \citep{mcelroy77}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Testing linear restrictions}
+\label{sec:testingRestrictions}
+
+Linear restrictions can be tested by an $F$ test, two Wald tests and
+a likelihood ratio (LR) test.
+
+The $F$ statistic for systems of equations is
+\begin{equation}
+F = \frac{
+   ( R \bHat - q )^\top
+   ( R ( X^\top ( \Sigma \otimes I )^{-1} X )^{-1} R^\top )^{-1}
+   ( R \bHat - q ) /
+   j
+}{
+   \uHat^\top ( \Sigma \otimes I )^{-1} \uHat /
+   ( G \cdot T - K )
+} \; ,
+\label{eq:f-test}
+\end{equation}
+where $j$ is the number of restrictions,
+$K$ is the total number of estimated coefficients,
+and all other variables are as defined before
+\citep[p.~314]{theil71}.
+Under the null hypothesis, $F$ is $F$ distributed
+with $j$ and $G \cdot T - K$ degrees of freedom.
+
+However, $F$ in Equation~\ref{eq:f-test} cannot be computed,
+because $\Sigma$ is generally unknown.
+As a solution, \citet[p.~314]{theil71} proposes
+to replace the unknown $\Sigma$ in Equation~\ref{eq:f-test}
+by the estimated covariance matrix $\SHat$.
+\begin{equation}
+\hat{F} = \frac{
+   ( R \bHat - q )^\top
+   ( R ( X^\top ( \SHat \otimes I )^{-1} X )^{-1} R^\top )^{-1}
+   ( R \bHat - q ) /
+   j
+}{
+   \uHat^\top ( \SHat \otimes I )^{-1} \uHat /
+   ( G \cdot T - K )
+}
+\label{eq:f-test-theil}
+\end{equation}
+Asymptotically, $\hat{F}$ has the same distribution as $F$
+in Equation~\ref{eq:f-test},
+because the numerator of Equation~\ref{eq:f-test-theil}
+converges in probability to the numerator of Equation~\ref{eq:f-test}
+and the denominator of Equation~\ref{eq:f-test-theil}
+converges in probability to the denominator of Equation~\ref{eq:f-test}
+\citep[p.~402]{theil71}.
+
+Furthermore, the denominators of both Equations~\ref{eq:f-test}
+and~\ref{eq:f-test-theil} converge in probability to~$1$.
+Taking this into account and applying Equation~\ref{eq:cov-ols-wls-sur},
+we obtain the usual $F$~statistic of the Wald test.
+\begin{equation}
+\hat{\hat{F}} = \frac{
+   ( R \bHat - q )^\top
+   ( R \, \COVHat \left[ \bHat \right] R^\top )^{-1}
+   ( R \bHat - q )
+}{
+   j
+}
+\label{eq:f-test-wald}
+\end{equation}
+Under the null hypotheses, also $\hat{\hat{F}}$ is
+asymptotically $F$ distributed
+with $j$ and $G \cdot T - K$ degrees of freedom.
+
+Multiplying Equation~\ref{eq:f-test-wald} with $j$,
+we obtain the usual $\chi^2$ statistic for the Wald test
+\begin{equation}
+W =
+   ( R \bHat - q )^\top
+   ( R \, \COVHat [ \bHat ] \, R^\top )^{-1}
+   ( R \bHat - q ) .
+\label{eq:chi2-test-wald}
+\end{equation}
+Asymptotically, $W$ has a $\chi^2$
+distribution with $j$ degrees of freedom
+under the null hypothesis
+\citep[p.~347]{greene03}.
+
+The likelihood-ratio (LR) statistic for systems of equations is
+\begin{equation}
+\LR = T \cdot \left(
+   \log \left| \SHat_r \right|
+   - \log \left| \SHat_u \right|
+   \right) ,
+\label{eq:lr-test}
+\end{equation}
+where $T$ is the number of observations per equation, and 
+$\SHat_r$ and $\SHat_u$ are
+the residual covariance matrices calculated by Equation~\ref{eq:rcovNoDfCor}
+of the restricted and unrestricted estimation, respectively.
+Asymptotically, $\LR$ has a $\chi^2$
+distribution with $j$ degrees of freedom
+under the null hypothesis
+\citep[p.~349]{greene03}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Hausman test}
+\label{sec:hausman}
+
+\citet{hausman78} developed a test for misspecification.
+The null hypothesis of the test is
+that the instrumental variables of each equation are
+uncorrelated with the disturbance terms of all other equations
+($\E \left[ u_i^\top Z_j \right] = 0 \, \forall \, i \neq j$).
+Under this null hypothesis, both the 2SLS and the 3SLS estimator are consistent,
+but the 3SLS estimator is (asymptotically) more efficient.
+Under the alternative hypothesis, the 2SLS estimator is consistent
+but the 3SLS estimator is inconsistent,
+i.e., the instrumental variables of each equation are
+uncorrelated with the disturbances of the same equation
+($\E \left[ u_i^\top Z_i \right] = 0 \, \forall \, i$),
+but the instrumental variables of at least one equation
+are correlated with the disturbances of another equation
+($\E \left[ u_i^\top Z_j \right] \neq 0 \, \exists \, i \neq j$).
+The Hausman test statistic is
+\begin{equation}
+  m = \left( \bHat_\text{2SLS} - \bHat_\text{3SLS} \right)^\top
+      \left( \COVHat \left[ \bHat_\text{2SLS} \right] -
+             \COVHat \left[ \bHat_\text{3SLS} \right] \right)
+      \left( \bHat_\text{2SLS} - \bHat_\text{3SLS} \right) ,
+\label{eq:hausman}
+\end{equation}
+where $\bHat_\text{2SLS}$ and $\COVHat \left[ \bHat_\text{2SLS} \right]$ are the estimated
+coefficient and covariance matrix from a 2SLS estimation, and
+$\bHat_\text{3SLS}$ and $\COVHat \left[ \bHat_\text{3SLS} \right]$ are the estimated
+coefficients and covariance matrix from a 3SLS estimation.
+Under the null hypothesis, this test statistic has a
+$\chi^2$ distribution with degrees of freedom equal to the number of
+estimated coefficients.
 

### 2008-09-12 paper/trunk/systemfit.Rnw
moved sections "Source code" (systemfit_code.Snw) into master file (systemfit.Rnw)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@954 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 
-\input{systemfit_code.tex}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Source code}\label{sec:code}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+<<echo=FALSE>>=
+options( prompt = "R> " )
+@
+
+The source code of the \pkg{systemfit} package is publicly available
+for download from the Comprehensive \proglang{R} Archive Network (CRAN,
+\url{http://CRAN.R-project.org/}).
+The basic functionality of this package is provided by the
+function \code{systemfit}.
+Moreover, this package provides
+tools for statistical tests,
+functions (methods) to show, extract or calculate results,
+some convenience functions,
+and internal helper functions.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection[The basic function systemfit]{The basic function \code{systemfit}}
+
+The \code{systemfit} function estimates systems of linear equations
+by different estimation methods.
+Where possible,
+the user interface and the returned object of this function
+follow the function \code{lm}
+--- the basic tool for linear regressions in \proglang{R} ---
+to make the usage of \code{systemfit} as easy as possible
+for \proglang{R} users.
+
+The econometric estimation is done by applying
+the formulas in Sections~\ref{sec:Estimation-ols-wls-sur}
+and~\ref{sec:Estimation-2sls-w2sls-3sls}
+or --- if the coefficients are estimate under linear restrictions ---
+by the formulas in Section~\ref{sec:Restrictions}.
+If the restrictions on the coefficients are specified symbolically,
+function \code{makeHypothesis} of the \pkg{car} package \citep{r-car-1.2-1, fox02a}
+is used to create the restriction matrix.
+
+The \code{systemfit} function returns a list of class \code{systemfit}
+that contains the results that belong to the entire system of equations.
+One special element of this list is called \code{eq},
+which is a list that contains one object for each estimated equation.
+These objects are lists of class \codeD{systemfit}{equation}
+and contain the results that belong only to the regarding equation.
+A complete description is available in the documentation of this function
+that is included in the package.
+A comparison of the elements returned by \code{lm} and by \code{systemfit}
+is available in appendix~\ref{sec:returned-object}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Statistical tests}
+
+The \codeD{linear}{hypothesis} and
+\code{lrtest} methods for \code{systemfit} objects
+as well as the function \codeD{hausman}{systemfit}
+apply the statistical tests described in
+Sections~\ref{sec:testingRestrictions} and~\ref{sec:hausman}.
+
+The \codeD{linear}{hypothesis} method for {systemfit} objects
+can be used to test linear restrictions on the estimated coefficients
+by \citeauthor{theil71}'s $F$ test or by usual Wald tests.
+Internally, \citeauthor{theil71}'s $F$ statistic is computed
+by the hidden function \codeD{.ftest}{systemfit}
+and the Wald tests are computed by
+the default \codeD{linear}{hypothesis} method of the \pkg{car} package
+\citep{r-car-1.2-1, fox02a}.
+The \code{lrtest} method for \code{systemfit} objects is a wrapper function
+to the default \code{lrtest} method of the \pkg{lmtest} package
+\citep{r-lmtest},
+which computes the likelihood-ratio (LR) test statistic.
+All these functions return an object of class \code{anova}
+that contains --- amongst others ---
+the empirical test statistic, the degrees of freedom,
+and the corresponding $P$ value.
+
+The function \codeD{hausman}{systemfit}
+tests the consistency of the 3SLS estimator.
+It returns an object of class \code{htest},
+which contains --- amongst others ---
+the empirical test statistic,
+the degrees of freedom, and the $P$ value.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Other methods and functions}
+
+The \pkg{systemfit} package provides several methods
+for objects both of classes \code{systemfit}
+and \codeD{systemfit}{equation}:
+\code{print} methods print the estimation results,
+\code{summary} methods calculate summary results,
+\code{confint} methods compute confidence intervals for the coefficients,
+\code{predict} methods calculate predicted values,
+\code{coef} methods extract the estimated coefficients,
+\code{vcov} methods extract their covariance matrix,
+\code{fitted} methods extract the fitted values,
+\code{residuals} methods extract the residuals,
+\code{formula} methods extract the formula(s),
+\code{terms} methods extract the model terms,
+\codeD{model}{frame} methods extract the model frame,
+and \codeD{model}{matrix} methods extract the model matrix.
+Some methods can be applied to objects of class \code{systemfit} only:
+a \code{correlation} method calculates
+the correlations between the predictions of two equations,
+an \code{se.ratio} method computes the ratios of the standard errors
+of the predictions between two models,
+and a \code{logLik} method extracts the log likelihood value.
+The package provides \code{print} methods
+to print objects of classes
+\codeD{summary}{systemfit}, \codeDD{summary}{systemfit}{equation},
+and \codeD{confint}{systemfit}
+that are returned by the above mentioned \code{summary}
+and \code{confint} methods.
+There exist also two \code{coef} methods
+to extract the estimated coefficients,
+their standard errors, $t$ values, and $P$ values
+from objects of classes
+\codeD{summary}{systemfit} and \codeDD{summary}{systemfit}{equation}.%
+\footnote{%
+There does not exist a special method to extract
+the degrees of freedom of the residuals from \code{systemfit} objects,
+because the default method of \code{df.residual}
+works for these objects.
+}
+
+The convenience function
+\code{createSystemfitModel} creates a model for \code{systemfit}
+by random numbers;
+\codeD{systemfit}{control} sets suitable default values
+for the technical control parameters for \code{systemfit}.
+
+Finally, the package includes some internal (hidden) helper functions:
+\codeD{.prepareData}{systemfit},
+\code{.stackMatList}, and \code{.prepareWmatrix}
+for preparing the data matrices;
+\code{.calcXtOmegaInv} and \code{.calcGLS}
+for calculating the GLS estimator;
+\code{.calcResidCov} and \code{.calcSigma2}
+for calculating the (co)variances of the residuals;
+and \codeD{.ftest}{systemfit}
+for calculating \citeauthor{theil71}'s $F$ statistic.
+If \code{systemfit} is applied to
+a (classical) °seemingly unrelated regression° analysis with panel data,
+it calls the hidden internal function \code{.systemfitPanel},
+which reshapes the data, creates the formulas to be estimated,
+and --- if requested --- specifies restrictions
+to ensure that the coefficients of all individuals are equal.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Efficiency of computations}
+\label{sec:code-efficiency}
+We have followed \cite{bates04} to make the code of \pkg{systemfit}
+faster and more stable.
+First, if a formula contains an inverse of a matrix
+that is post-multiplied by a vector or matrix,
+we use \code{solve(A,b)} instead of \code{solve(A) \%*\% b}.
+Second, we calculate crossproducts
+by \code{crossprod(X)} or \code{crossprod(X,y)}
+instead of \code{t(X) \%*\% X} or \code{t(X) \%*\% y},
+respectively.
+
+The matrix $\Omega^{-1}$ that is used to compute the estimated
+coefficients and their covariance matrix is of size
+$( G \cdot T ) \times ( G \cdot T )$
+(see Sections~\ref{sec:Estimation-ols-wls-sur},
+\ref{sec:Estimation-2sls-w2sls-3sls}, and~\ref{sec:Restrictions}).
+In case of large data sets, $\Omega^{-1}$ becomes computationally infeasible.
+Therefore, we use the following transformation and compute $X^\top \Omega^{-1}$
+by dividing the $X$ matrix into submatrices,
+doing some calculations with these submatrices,
+adding up some of these submatrices, and
+finally putting the submatrices together, so that
+\begin{equation}
+X^\top \Omega^{-1}
+%= X^\top \left( \Sigma^{-1} \otimes I \right)
+= \sum_{i=1} \left[ \begin{array}{c}
+   \sigma^{1i} {X^1} \\
+   \sigma^{2i} {X^2} \\
+   \vdots \\
+   \sigma^{Gi} {X^G} \\
+\end{array} \right]^\top ,
+\label{eq:omegaInv}
+\end{equation}
+where $\sigma^{ij}$ are the elements of the matrix $\Sigma^{-1}$,
+and $X^i$ is a submatrix of $X$ that contains the rows
+that belong to the $i$'s equation.
+This computation is done inside the internal (hidden) function
+\code{.calcXtOmegaInv}.
+
+Since version 1.0, the \code{systemfit} function by default
+uses the \pkg{Matrix} package \citep{r-matrix-07}
+for all computations where matrices are involved.
+The \pkg{Matrix} package provides classes for different types of matrices.
+For instance, we choose class \code{dgeMatrix}
+(°real matrices in general storage mode°),
+for matrices $X_i$ in Equation~\ref{eq:model-array},
+class \code{dgCMatrix} (°general, numeric, sparse matrices
+in the (sorted) compressed sparse column format°)
+for matrix $X$ in Equation~\ref{eq:model-matrices},
+and class \code{dspMatrix}
+(°symmetric real matrices in packed storage (one triangle only)°)
+for the residual covariance matrix $\SHat$.
+If the \pkg{Matrix} package is used,
+the possibly huge matrix $\Omega^{-1}$ is no longer a problem,
+because it is a sparse matrix
+that can be stored in a compressed format (class \code{dgCMatrix}).
+Hence, we no longer need the algorithm in Equation~\ref{eq:omegaInv}.
+We have tested different ways to calculate a GLS estimator
+like in Equation~\ref{eq:ols-wls-sur}
+and we found
+that the following code is the fastest:
+<<eval=FALSE>>=
+sigmaInv <- solve( residCov )
+xtOmegaInv <- crossprod( xMat, kronecker( sigmaInv, Diagonal( nObs ) ) )
+coef <- solve( xtOmegaInv %*% xMat, xtOmegaInv %*%  yVec )
+@
+In this code snippet,
+\code{residCov} is the residual covariance matrix $\SHat$,
+\code{nObs} is the number of observations in each equation $T$,
+\code{xMat} is the matrix $X$ and
+\code{yVec} is the vector $y$ in Equation~\ref{eq:ols-wls-sur}.
+
+By default, the \code{systemfit} function uses the \pkg{Matrix} package
+to perform GLS estimations,
+because using this package considerably decreases
+the computation time for many common models.
+However, the estimation of small models with small data sets
+gets slower by using the \pkg{Matrix} package
+(see appendix~\ref{sec:timings}).
+While this increase in computation time
+is often imperceptible to human beings,
+it might matter in some cases
+such as iterated estimations or Monte Carlo studies.
+Therefore, the user can opt for not using the \pkg{Matrix} package,
+but Equation~\ref{eq:omegaInv} with standard \proglang{R} matrices.
+
+\subsection[Overlap with other functions and packages in R]
+   {Overlap with other functions and packages in \proglang{R}}
+
+Single-equation models can be fitted in \proglang{R}
+by OLS with function \code{lm} (package \pkg{stats})
+and by 2SLS with function \code{tsls}
+(package \pkg{sem}, \citealp{r-sem-0.9-8}).
+This is also possible with the \code{systemfit} function,
+but \code{systemfit} is specialized in estimating systems of equation,
+i.e., more than one equation.
+Its capability to estimate single-equation models is just a side-effect.
+
+Function \code{sem} (package \pkg{sem}, \citealp{r-sem-0.9-8})
+can be used to estimate structural equation models in \proglang{R}
+by limited information maximum likelihood (LIML)
+and full information maximum likelihood (FIML)
+assuming normal or multinormal errors, respectively.
+A special feature of this function is the estimation
+of models with unobserved (°latent°) variables,
+which is not possible with \code{systemfit}.
+While \code{sem} cannot be used to consistently estimate
+systems of simultaneous equations with some endogenous regressors,
+it can be used to estimate systems of equations,
+where all regressors are exogenous.
+However, the latter is rather cumbersome
+(see appendix~\ref{sec:sem}).
+Hence, \code{systemfit} is the only function in \proglang{R}
+that can be used to estimate systems of simultaneous equations
+and it is the most convenient function to estimate
+systems of equations with purely exogenous regressors.
+
 \input{systemfit_usage.tex}

### 2008-09-12 paper/trunk/runSweave.sh

@@ -1,8 +1,6 @@
 #!/bin/sh
 echo "Sweave(\"systemfit.Rnw\")" | LC_ALL="C" R --no-save --no-restore
 
-echo "Sweave(\"systemfit_code.Snw\")" | LC_ALL="C" R --no-save --no-restore
-
 echo "Sweave(\"systemfit_usage.Snw\")" | LC_ALL="C" R --no-save --no-restore
 echo "Stangle(\"systemfit_usage.Snw\", annotate=FALSE, split=TRUE,
    prefix=FALSE)" | LC_ALL="C" R --no-save --no-restore

### 2008-09-12 paper/trunk/systemfit.Rnw
moved section "Using systemfit" (systemfit_usage.Snw) into master file (systemfit.Rnw)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@955 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 
-\input{systemfit_usage.tex}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section[Using systemfit]{Using \pkg{systemfit}}\label{sec:Usage}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\SweaveOpts{prefix.string=trash/usage}
+<<echo=FALSE>>=
+options( prompt = "R> " )
+@
+
+In this section we demonstrate how to use the \pkg{systemfit} package.
+First, we show the standard usage of \code{systemfit} by a simple example.
+Second, several options that can be specified by the user are presented.
+Then, the usage of \code{systemfit}
+for a (classical) °seemingly unrelated regression° analysis
+with panel data is described.
+Finally, we demonstrate how to apply some statistical tests.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection[Standard usage of systemfit]{Standard usage of \code{systemfit}}
+\label{sec:standard-usage}
+
+As described in the previous section,
+systems of equations can be econometrically estimated
+with the function \code{systemfit}.
+The only mandatory argument is \code{formula}.
+Typically, it is a list of formulas to be estimated,
+but it may also be a single formula for estimating a single-equation model.
+Each formula is a standard regression formula in \proglang{R}
+(see documentation of \code{formula}).
+
+The following demonstration uses an example taken from
+\citet[p.~685]{kmenta86}.
+We want to estimate a small model of the US food market:
+\begin{align}
+\texttt{consump} & = \beta_1 + \beta_2 \cdot \texttt{price}
+   + \beta_3 \cdot \texttt{income} \\
+\texttt{consump} & = \beta_4 + \beta_5 \cdot \texttt{price}
+   + \beta_6 \cdot \texttt{farmPrice} + \beta_7 \cdot \texttt{trend}
+\end{align}
+The first equation represents the demand side of the food market.
+Variable \code{consump} (food consumption per capita) is the dependent variable.
+The regressors are \code{price} (ratio of food prices to general consumer prices)
+and \code{income} (disposable income) as well as a constant.
+The second equation specifies the supply side of the food market.
+Variable \code{consump} is the dependent variable of this equation as well.
+The regressors are again \code{price} (ratio of food prices to general 
+consumer prices) and a constant as well as 
+\code{farmPrice} (ratio of preceding year's prices received by farmers
+to general consumer prices) and 
+\code{trend} (a time trend in years).
+These equations can be estimated by OLS in \proglang{R} by
+<<systemfit_usage>>=
+library( "systemfit" )
+data( "Kmenta" )
+attach( Kmenta )
+eqDemand <- consump ~ price + income
+eqSupply <- consump ~ price + farmPrice + trend
+eqSystem <- list( demand = eqDemand, supply = eqSupply )
+fitols <- systemfit( eqSystem )
+print( fitols )
+@
+
+The first line loads the \pkg{systemfit} package.
+The second line loads example data that are included with the package.
+They are attached to the \proglang{R} search path in line three.
+In the fourth and fifth line, the demand and supply equations are specified,
+respectively.%
+\footnote{
+A regression constant is always implied if not explicitly omitted.
+}
+In the sixth line, these equations are concatenated into a list and
+are labeled \code{demand} and \code{supply}, respectively.%
+\footnote{
+If no labels are provided, the equations are numbered consecutively
+( \code{eq1}, \code{eq2}, \ldots ).
+}
+Finally, in the last two lines, the regression is performed
+and the estimation results are printed.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection[User options of systemfit]{User options of \code{systemfit}}
+\label{sec:user-options}
+
+The user can modify the default estimation method by providing
+additional optional arguments,
+e.g., to specify instrumental variables
+or restrictions on the coefficients.
+All optional arguments are described in the following:
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Estimation method}
+
+The optional argument \code{method} is a string
+that determines the estimation method.
+It must be either \code{"OLS"}, \code{"WLS"}, \code{"SUR"},
+\code{"2SLS"}, \code{"W2SLS"}, or \code{"3SLS"}.
+These methods correspond to the estimation methods
+described in Sections~\ref{sec:Estimation-ols-wls-sur},
+\ref{sec:Estimation-2sls-w2sls-3sls}, and~\ref{sec:Restrictions}.
+The following command estimates the model described above as
+°seemingly unrelated regression°.
+<<systemfit_usage>>=
+fitsur <- systemfit( eqSystem, method = "SUR" )
+@
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Instrumental variables}
+The instruments for a 2SLS, W2SLS or 3SLS estimation can be specified
+by the argument \code{inst}.
+If the same instruments should be used for all equations,
+\code{inst} must be a one-sided formula.%
+\footnote{
+A one-sided formula is a standard formula in \proglang{R}
+without a dependent variable.}
+If different instruments should be used for each equation,
+\code{inst} must be a list that contains a one-sided formula
+for each equation.
+The following example uses instrumental variables
+to estimate the model described above
+by °three-stage least squares° (3SLS).
+While the first command specifies the same instruments for all equations,
+the second uses different instruments:
+<<systemfit_usage>>=
+fit3sls  <- systemfit( eqSystem, method = "3SLS",
+   inst = ~ income + farmPrice + trend )
+fit3sls2 <- systemfit( eqSystem, method = "3SLS",
+   inst = list( ~ farmPrice + trend, ~ income + farmPrice + trend ) )
+@
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Data}
+Having all data in the global environment or attached to the search path
+is often inconvenient.
+Therefore, \code{systemfit} has the argument \code{data}
+to specify a data frame
+that contains the variables of the model.
+In the following example, we use this argument to specify
+that the data for the estimation
+should be taken from the data frame \code{Kmenta}.
+Hence, we no longer need to attach this data frame before calling \code{systemfit}:
+<<systemfit_usage>>=
+fitsur <- systemfit( eqSystem, method = "SUR", data = Kmenta )
+@
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Restrictions on the coefficients}
+As outlined in Section~\ref{sec:Restrictions}, restrictions on the coefficients
+can be specified in two ways.
+One way is to use the matrix $R$ and the vector $q$
+(see Section~\ref{sec:Restrictions}).
+These restrictions can be specified symbolically
+by argument \codeD{restrict}{matrix}
+as in the generic function
+\codeD{linear}{hypothesis} of the \pkg{car} package \citep{r-car-1.2-1, fox02a}.
+This argument must be a vector of character strings,
+where each element represents one linear restriction,
+and each element must be either a linear combination of coefficients,
+or a linear equation in the coefficients
+(see documentation of function \codeD{linear}{hypothesis}
+in the \pkg{car} package, \citealp{r-car-1.2-1, fox02a}).
+We illustrate this by estimating the model under
+the restriction $\beta_2 + \beta_6 = 0$.
+Since the name of $\beta_2$
+(coefficient of variable \code{price} in equation \code{demand})
+is \code{demand\_price}
+and the name of $\beta_6$
+(coefficient of variable \code{farmPrice} in equation \code{supply})
+is \code{supply\_farmPrice},
+this restriction can be specified by
+<<systemfit_usage,keep.source=TRUE>>=
+restrict <- "demand_price + supply_farmPrice = 0"
+fitsurRmat <- systemfit(eqSystem, method = "SUR",
+   restrict.matrix = restrict)
+@
+\label{code:Rmat}
+
+Alternatively, the restrictions via matrix $R$ and vector $q$
+can be specified numerically.
+The matrix $R$ can be specified with argument \codeD{restrict}{matrix}
+and the vector $q$ with argument \codeD{restrict}{rhs}.
+<<systemfit_usage,keep.source=TRUE>>=
+Rmat <- matrix(0, nrow = 1, ncol = 7)
+Rmat[1, 2] <- 1
+Rmat[1, 6] <- 1
+qvec <- c(0)
+fitsurRmatNum <- systemfit(eqSystem, method = "SUR",
+   restrict.matrix = Rmat, restrict.rhs = qvec)
+@
+The first line creates a $1 \times 7$ matrix of zeros,
+where 1 is the number of restrictions and
+7 is the number of unrestricted coefficients.
+The following two lines specify this matrix in a way
+that the multiplication with the coefficient vector
+results in $ \beta_2 + \beta_6 $.
+The fourth line creates a vector with a single element
+that contains the right hand side of the restriction, i.e., zero.
+Finally the coefficients are estimated under the restriction
+$\beta_2 + \beta_6 = 0$.
+
+The other way to specify restrictions on the coefficients is to modify the regressor matrix
+by post-multiplying it with a matrix, say $M$
+(see Section~\ref{sec:Restrictions}).
+This kind of restriction can be specified
+by setting argument \codeD{restrict}{regMat} equal to the matrix $M$.
+We convert the restriction specified above to $\beta_2 = - \beta_6$,
+and set $\beta_1 = \beta^\text{M}_1$, \ldots, $\beta_5 = \beta^\text{M}_5$,
+$\beta_6 = - \beta^\text{M}_2$, and $\beta_7 =  \beta^\text{M}_6$.
+We can do this in \proglang{R} by
+<<systemfit_usage,keep.source=TRUE>>=
+modRegMat <- matrix(0, nrow = 7, ncol = 6)
+modRegMat[1:5, 1:5] <- diag(5)
+modRegMat[6, 2] <- -1
+modRegMat[7, 6] <- 1
+fitsurRegMat <- systemfit(eqSystem, method = "SUR",
+   restrict.regMat = modRegMat)
+@
+The first line creates a $7 \times 6$ matrix of zeros,
+where 7 is the number of unrestricted coefficients and
+6 is the number of restricted coefficients.
+The following three lines specify the matrix $M$ (\code{modRegMat})
+as described before.
+Finally the coefficients are estimated under the restriction
+$\beta^\text{M}_2 = \beta_2 = - \beta_6$.
+
+Of course, the estimation results do not depend on the method
+that was used to specify this restriction:
+<<systemfit_usage>>=
+all.equal( coef( fitsurRmat ), coef( fitsurRmatNum ) )
+all.equal( coef( fitsurRmat ), coef( fitsurRegMat ) )
+@
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Iteration control}
+The estimation methods WLS, SUR, W2SLS and 3SLS need a
+covariance matrix of the residuals
+that can be calculated from a first-step OLS or 2SLS estimation
+(see Section~\ref{sec:residcov}).
+This procedure can be iterated and at each iteration the covariance
+matrix is calculated from the previous step estimation.
+This iteration is repeated until the maximum number of iterations
+is reached or the coefficient estimates have converged.
+The maximum number of iterations is specified by argument \code{maxiter}.
+Its default value is one, which means no iteration.
+The convergence criterion is
+\begin{equation}
+   \sqrt{ \frac{ \sum_i (\beta_{i,g} - \beta_{i,g-1})^2 }
+   { \sum_i \beta_{i,g-1}^2 }}
+      < \texttt{tol} ,
+\end{equation}
+where $\beta_{i,g}$ is the $i$th coefficient of the $g$th iteration.
+The default value of the convergence criterion (argument \code{tol})
+is $10^{-5}$.
+
+In the following example,
+we estimate the model described above by iterated SUR:
+<<systemfit_usage>>=
+fitsurit <- systemfit( eqSystem, method = "SUR", maxiter = 500 )
+@
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Residual covariance matrix}
+It was explained in Section~\ref{sec:residcov} that several different
+methods have been proposed to calculate the residual
+covariance matrix.
+The user can specify, which method \code{systemfit} should use.
+Possible values of the argument \code{methodResidCov} are presented in
+Table~\ref{tab:methodResidCov}.
+By default, \code{systemfit} uses Equation~\ref{eq:rcovGeomean}.
+
+\begin{table}[H]
+\centering
+\begin{tabular}{|c|c|}
+\hline
+argument & equation \\
+\code{methodResidCov} & \\
+\hline
+\code{"noDfCor"} & \ref{eq:rcovNoDfCor} \\
+\hline
+\code{"geomean"} & \ref{eq:rcovGeomean} \\
+\hline
+\code{"max"}     & \ref{eq:rcovMax} \\
+\hline
+\code{"Theil"}   & \ref{eq:rcovTheil} \\
+\hline
+\end{tabular}
+\caption{Possible values of argument \code{methodResidCov}}
+\label{tab:methodResidCov}
+\end{table}
+
+Furthermore, the user can specify
+whether the means should be subtracted from the residuals
+before Equations~\ref{eq:rcovNoDfCor}, \ref{eq:rcovGeomean}, \ref{eq:rcovMax},
+or~\ref{eq:rcovTheil} are applied
+to calculate the residual covariance matrix
+(see Section~\ref{sec:residcov}).
+The corresponding argument is called \code{centerResiduals}.
+It must be either \code{TRUE} (subtract the means) or
+\code{FALSE} (take the unmodified residuals).
+The default value of \code{centerResiduals} is \code{FALSE}.
+
+Moreover, if the coefficients are estimated under restrictions,
+the user can use argument \code{residCovRestricted} to specify
+whether the residual covariance matrix
+for a WLS, SUR, W2SLS, or 3SLS estimation
+should be obtained from a restricted or from an unrestricted
+first-step estimation
+(see Section~\ref{sec:residcov}).
+If this argument is \code{TRUE} (the default),
+the residual covariance matrix is obtained from a restricted
+OLS or 2SLS estimation.
+If it is \code{FALSE},
+the residual covariance matrix is obtained from an unrestricted
+first-step estimation.
+
+Finally, argument \code{residCovWeighted} can be used to decide,
+whether the residual covariance matrix for a SUR (3SLS) estimation
+should be obtained from a WLS (W2SLS) estimation
+instead of from an OLS (2SLS) estimation
+(see Section~\ref{sec:residcov}).
+By default, \code{residCovWeighted} is \code{FALSE},
+which means that the residuals of an OLS (2SLS) estimation are used
+to compute the residual covariance matrix.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{3SLS formula}
+As discussed in Sections~\ref{sec:Estimation-2sls-w2sls-3sls} and~\ref{sec:Restrictions},
+there exist several different
+methods to perform a 3SLS estimation.
+The user can specify the method by argument \code{method3sls}.
+Possible values are presented in Table~\ref{tab:method3sls}.
+The default value is \code{"GLS"}.
+
+\begin{table}[H]
+\centering
+\begin{tabular}{|c|c|c|}
+\hline
+argument           & equation       & equation \\
+\code{method3sls} & (unrestricted) & (restricted) \\
+\hline
+\code{"GLS"}     & \ref{eq:2sls-w2sls-3sls} & \ref{eq:2sls-w2sls-3sls-r} \\
+\hline
+\code{"IV"}      & \ref{eq:3slsIv}      & \ref{eq:3slsIvR} \\
+\hline
+\code{"GMM"}     & \ref{eq:3slsGmm}     & \ref{eq:3slsGmmR} \\
+\hline
+\code{"Schmidt"} & \ref{eq:3slsSchmidt} & \ref{eq:3slsSchmidtR} \\
+\hline
+\code{"EViews"}  & \ref{eq:3slsEViews}  & \ref{eq:3slsEViewsR} \\
+\hline
+\end{tabular}
+\caption{Possible values of argument \code{method3sls}}
+\label{tab:method3sls}
+\end{table}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Sigma squared}
+In case of OLS or 2SLS estimations, argument \code{singleEqSigma}
+can be used to specify,
+whether different $\sigma^2$s for each single equation
+or the same $\sigma^2$ for all equations should be used.
+If argument \code{singleEqSigma} is \code{TRUE},
+$\OHat$ in Equation~\ref{eq:cov-ols-wls-sur} or~\ref{eq:cov-2sls-w2sls-3sls}
+is set to $\SHat \otimes I_T$.
+In contrast, if argument \code{singleEqSigma} is \code{FALSE},
+$\OHat$ in Equation~\ref{eq:cov-ols-wls-sur} or~\ref{eq:cov-2sls-w2sls-3sls}
+is set to $\sHat^2 I_{G \cdot T}$.
+In case of an unrestricted regression,
+argument \code{singleEqSigma} is \code{TRUE} by default.
+However, if the coefficients are estimated under restrictions,
+this argument is \code{FALSE} by default.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{System options}
+Furthermore, two options regarding some internal calculations are available.
+First, argument \code{solvetol} specifies the tolerance level
+for detecting linear dependencies when inverting a matrix or
+calculating a determinant (using functions \code{solve} and \code{det}).
+The default value depends on the used computer system and is equal to the
+default tolerance level of \code{solve} and \code{det}.
+
+Second, argument \code{useMatrix} specifies
+whether the \pkg{Matrix} package \citep{r-matrix-07}
+should be used for all computations where matrices are involved
+(see Section~\ref{sec:code-efficiency}).
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Returned data objects}
+
+Finally, the user can decide
+whether \code{systemfit} should return some data objects.
+Argument \code{model} indicates
+whether a data frame with the data of the model
+should be returned.
+Its default value is \code{TRUE},
+i.e., the model frame is returned.
+Arguments \code{x}, \code{y},
+and \code{z} specify
+whether the model matrices ($X_i$),
+the responses ($y_i$), and the matrices of instrumental variables ($Z_i$),
+respectively, should be returned.
+These three arguments are \code{FALSE} by default,
+i.e., these data objects are not returned.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection[Summary results with summary.systemfit]
+   {Summary results with \code{summary.systemfit}}
+
+The \code{summary} method can be used
+to compute and print summary results of objects
+returned by \code{systemfit}.
+<<systemfit_usage>>=
+summary( fitsur )
+@
+First, the estimation method is reported
+and a few summary statistics for the entire system
+and for each equation are given.
+Then, the covariance matrix used for estimation
+and the covariance matrix as well as the correlation matrix
+of the (final) residuals are printed.
+Finally, the estimation results of each equation are reported:
+the formula of the estimated equation,
+the estimated coefficients, their standard errors, $t$ values, $P$ values
+and codes indicating their statistical significance,
+as well as some other statistics like
+the standard error of the residuals and
+the $R^2$ value of the equation.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection[Degrees of freedom for t tests]
+   {Degrees of freedom for $t$ tests}
+
+The \code{summary} method for \code{systemfit} objects has an optional argument \code{useDfSys}.
+It selects the approach
+that is applied by \code{systemfit}
+to determine the degrees of freedom of $t$ tests of the estimated coefficients
+(Section~\ref{sec:degreesOfFreedom}).
+If argument \code{useDfSys} is \code{TRUE},
+the degrees of freedom of the whole system are taken.
+In contrast, if \code{useDfSys} is \code{FALSE},
+the degrees of freedom of the single equation are taken.
+If the coefficients are estimated under restrictions,
+argument \code{useDfSys} is \code{TRUE} by default.
+However, if no restrictions on the coefficient are specified,
+the default value of \code{useDfSys} is \code{FALSE}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Reduce amount of printed output}
+
+The optional arguments \code{residCov} and \code{equations}
+can be used reduce the amount of the printed output.
+Argument \code{residCov} specifies
+whether the covariance matrix and
+the correlation matrix
+of the residuals are printed.
+Argument \code{equations} specifies
+whether summary results of each equation are printed.
+By default, both arguments are \code{TRUE}.
+The following command returns a sparse summary output:
+<<systemfit_usage>>=
+summary( fitsur, residCov = FALSE, equations = FALSE )
+@
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Panel data}
+
+The \code{systemfit} function can also be used
+for a (classical) °seemingly unrelated regression° analysis
+with panel data.
+For this type of analysis,
+the data must be provided in a transformed data frame of class \codeD{plm}{dim}%
+\footnote{
+Generally, panel data can be either in °long format°
+(different individuals are arranged below each other)
+or in °wide format°
+(different individuals are arranged next to each other).
+For this analysis, the data must be in °long format°.
+}
+which can be created with the function \codeD{plm}{data}
+from the \proglang{R} package \pkg{plm}
+\citep{r-plm-0.3-1}.
+In contrast to the previously described usage of \code{systemfit},
+argument \code{formula} must be a single equation (object of class \code{formula}).
+This formula is estimated for all individuals.
+
+We demonstrate the application of \code{systemfit}
+to panel data
+using an example taken from \citet[p.~340]{greene03}
+that is based on \citet{grunfeld58}.
+We want to estimate a model for gross investment of 5 US firms
+in the years 1935--1954:
+\begin{equation}
+\texttt{invest}_{it} = \beta_1 + \beta_2 \cdot \texttt{value}_{it} +
+   \beta_3 \cdot \texttt{capital}_{it}
+\end{equation}
+where \code{invest} is the gross investment of firm $i$ in year $t$,
+\code{value} is the market value of the firm at the end of the
+previous year, and
+\code{capital} is the capital stock of the firm at the end of the
+previous year.
+
+This model can be estimated by
+<<systemfit_usage>>=
+data( "GrunfeldGreene" )
+library( "plm" )
+GGPanel <- plm.data( GrunfeldGreene, c( "firm", "year" ) )
+greeneSur <- systemfit( invest ~ value + capital, method = "SUR",
+   data = GGPanel )
+@
+The first line loads the example data set \code{GrunfeldGreene}
+that is included in the \pkg{systemfit} package.
+The second line loads the \pkg{plm} package
+and the following line specifies a data frame of class \codeD{plm}{dim},
+where the variables \code{firm} and \code{year} indicate
+the individual (cross-section) and time identifier, respectively.
+Finally, a seemingly unrelated regression is performed.
+
+The optional argument \code{pooled} is a logical variable indicating
+whether the coefficients are restricted to be equal for all
+individuals.
+By default, this argument is set to \code{FALSE}.
+The following command does a seemingly unrelated regression
+of the same model as before,
+but with coefficients restricted to be equal for all individuals.
+<<systemfit_usage>>=
+greeneSurPooled <- systemfit( invest ~ value + capital, method = "SUR",
+   data = GGPanel, pooled = TRUE )
+@
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Testing linear restrictions}
+
+As described in Section~\ref{sec:testingRestrictions},
+linear restrictions can be tested by an $F$ test, two Wald tests and
+an LR test.
+The \pkg{systemfit} package provides
+the method \codeD{linear}{hypothesis} for the $F$ and Wald tests
+as well as the method \code{lrtest} for LR tests.
+
+We will now test the restriction $\beta_2 = -\beta_6$
+that was specified by the matrix \code{Rmat} and the vector \code{qvec}
+in the example above (p.~\pageref{code:Rmat}).
+<<systemfit_usage>>=
+linear.hypothesis( fitsur, Rmat, qvec, test = "FT" )
+
+linear.hypothesis( fitsur, Rmat, qvec, test = "F" )
+
+linear.hypothesis( fitsur, Rmat, qvec, test = "Chisq" )
+
+lrtest( fitsurRmat, fitsur )
+@
+The linear restrictions are tested by \citeauthor{theil71}'s $F$ test
+(Equation~\ref{eq:f-test-theil}) first,
+second by the $F$ statistic of a Wald test
+(Equation~\ref{eq:f-test-wald}),
+third by the $\chi^2$ statistic of a Wald test
+(Equation~\ref{eq:chi2-test-wald}),
+and finally by an LR test
+(Equation~\ref{eq:lr-test}).
+
+The first argument of the \codeD{linear}{hypothesis} method for \code{systemfit} objects
+must be an unrestricted regression
+returned by \code{systemfit}.
+The second and third arguments are the restriction matrix $R$ and
+the optional vector $q$, as described in Section~\ref{sec:Restrictions}.
+Analogously to the argument \codeD{restrict}{matrix}
+of the \code{systemfit} function,
+the restrictions can be specified either in matrix form or symbolically.
+The optional argument \code{test} must be
+a character string, \code{"FT"}, \code{"F"}, or \code{"Chisq"},
+specifying whether to compute
+\citeauthor{theil71}'s finite-sample $F$ test
+(with approximate $F$ distribution)
+the finite-sample Wald test (with approximate $F$ distribution)
+or the large-sample Wald test (with asymptotic $\chi^2$ distribution).
+
+All arguments of the \code{lrtest} method for \code{systemfit} objects
+must be fitted model objects returned by \code{systemfit}.
+It consecutively compares all provided fitted model objects.
+
+All tests print a short description of the test
+and the tested model objects first.
+Then, a small table is printed,
+where each row belongs to one (unrestricted or restricted) model.
+The second row reports (amongst others)
+the degree(s) of freedom of the test,
+the empirical test statistic,
+and the marginal level of significance ($P$ value).
+Although all tests check the same hypothesis,
+there is some variation of the $P$ values.
+However, all tests suggest the same decision:
+The null hypothesis $\beta_2 = -\beta_6$ cannot be rejected at any
+reasonable level of significance.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Hausman test}
+A Hausman test, which is described in Section~\ref{sec:hausman},
+can be carried out with following commands:
+<<systemfit_usage>>=
+fit2sls  <- systemfit( eqSystem, method = "2SLS",
+   inst = ~ income + farmPrice + trend, data = Kmenta )
+fit3sls <- systemfit( eqSystem, method = "3SLS",
+   inst = ~ income + farmPrice + trend, data = Kmenta )
+hausman.systemfit( fit2sls, fit3sls )
+@
+<<echo=FALSE>>=
+hausmantest <- hausman.systemfit( fit2sls, fit3sls )
+@
+First of all, the model is estimated by 2SLS and then by 3SLS.
+Finally, in the last line
+the test is carried out by the command \codeD{hausman}{systemfit}.
+This function requires two arguments:
+the result of a 2SLS estimation and
+the result of a 3SLS estimation.
+The Hausman test statistic is
+\Sexpr{round( hausmantest$statistic, digits = 3)},
+which has a $\chi^2$ distribution
+with \Sexpr{hausmantest$parameter} degrees of freedom
+under the null hypothesis.
+The corresponding $P$ value is
+\Sexpr{round( hausmantest$p.value, digits = 3 )}.
+This shows that the null hypothesis is not rejected
+at any reasonable level of significance.
+Hence, we can assume that the 3SLS estimator is consistent.
+
 \input{systemfit_reliability.tex}

### 2008-09-12 paper/trunk/runSweave.sh

@@ -1,8 +1,7 @@
 #!/bin/sh
 echo "Sweave(\"systemfit.Rnw\")" | LC_ALL="C" R --no-save --no-restore
 
-echo "Sweave(\"systemfit_usage.Snw\")" | LC_ALL="C" R --no-save --no-restore
-echo "Stangle(\"systemfit_usage.Snw\", annotate=FALSE, split=TRUE,
+echo "Stangle(\"systemfit.Rnw\", annotate=FALSE, split=TRUE,
    prefix=FALSE)" | LC_ALL="C" R --no-save --no-restore
 
 echo "Sweave(\"systemfit_reliability.Snw\")" | LC_ALL="C" R --no-save --no-restore

### 2008-09-12 paper/trunk/systemfit.Rnw
moved section "Replication of textbook results" (systemfit_reliability.Snw) into master file (systemfit.Rnw)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@956 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 
-\input{systemfit_reliability.tex}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Replication of textbook results}\label{sec:reliability}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\SweaveOpts{prefix.string=trash/replication}
+<<echo=FALSE>>=
+options( prompt = "R> " )
+@
+
+In this section, we reproduce results from several textbook examples
+using the \pkg{systemfit} package for several reasons.
+First, a comparison of \pkg{systemfit}'s results
+with results published in the literature
+confirms the reliability of the \pkg{systemfit} package.
+Second, this section helps teachers and students
+become familiar with using the \pkg{systemfit} package.
+Third, the section encourages reproducible research,
+which should be a general goal in scientific analysis
+\citep{buckheit95,schwab00}.
+For instance, by preparing this section,
+the exact estimation methods of the replicated analyses have been discovered
+and a few errors in \citet{greene03} have been found
+\citep[see][]{greene06a}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection[Kmenta (1986): Example on p. 685 (food market)]
+   {Kmenta (1986): Example on p.~685 (food market)}
+
+First, we reproduce an example taken from \citet[p.~685]{kmenta86}.
+The data are available from Table~13-1 (p.~687),
+and the results are presented in Table~13-2 (p.~712) of this book.
+
+Before starting the estimation,
+we load the data and specify the two formulas of the model
+as well as the instrumental variables.
+Then the equation system is estimated by OLS, 2SLS, 3SLS, and
+iterated 3SLS.
+After each estimation, we provide the commands
+to print the estimated coefficients.
+<<systemfit_replication,echo=FALSE>>=
+library( "systemfit" )
+@
+<<systemfit_replication,results=hide>>=
+data( "Kmenta" )
+eqDemand <- consump ~ price + income
+eqSupply <- consump ~ price + farmPrice + trend
+inst <- ~ income + farmPrice + trend
+system <- list( demand = eqDemand, supply = eqSupply )
+@
+OLS estimation:
+<<systemfit_replication,results=hide>>=
+fitOls <- systemfit( system, data = Kmenta )
+round( coef( summary( fitOls ) ), digits = 4 )
+@
+2SLS estimation:
+<<systemfit_replication,results=hide>>=
+fit2sls <- systemfit( system, method = "2SLS", inst = inst, data = Kmenta )
+round( coef( summary( fit2sls ) ), digits = 4 )
+@
+3SLS estimation:
+<<systemfit_replication,results=hide>>=
+fit3sls <- systemfit( system, method = "3SLS", inst = inst, data = Kmenta )
+round( coef( summary( fit3sls ) ), digits = 4 )
+@
+Iterated 3SLS estimation:
+<<systemfit_replication,results=hide>>=
+fitI3sls <- systemfit( system, method = "3SLS", inst = inst, data = Kmenta,
+   maxit = 250 )
+round( coef( summary( fitI3sls ) ), digits = 4 )
+@
+
+The above commands return exactly the same
+coefficients and standard errors as published in \citet[p.~712]{kmenta86}
+except for two minor exceptions:
+two standard errors of the 2SLS estimation deviate by $0.0001$.
+However, this difference is likely due to rounding errors
+in \code{systemfit} or \citet{kmenta86} and
+is so small that it empirically does not matter.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Greene (2003): Example 15.1 (Klein's model I)}
+\label{sec:KleinsModel}
+
+Second, we try to replicate Klein's °Model I° \citep{klein50}
+that is described in \citet[p.~381]{greene03}.
+The data are available from the online complements to \citet{greene03},
+Table~F15.1
+(\url{http://pages.stern.nyu.edu/~wgreene/Text/tables/TableF15-1.txt}),
+and the estimation results are presented in Table~15.3 (p.~412).
+
+Initially, the data are loaded and three equations
+as well as the instrumental variables are specified.
+As in the example before, the equation system is estimated by
+OLS, 2SLS, 3SLS, and iterated 3SLS, and
+commands to print the estimated coefficients are presented.
+<<systemfit_replication>>=
+data( "KleinI" )
+eqConsump  <- consump ~ corpProf + corpProfLag + wages
+eqInvest   <- invest ~ corpProf + corpProfLag + capitalLag
+eqPrivWage <- privWage ~ gnp + gnpLag + trend
+inst <- ~ govExp + taxes + govWage + trend + capitalLag + corpProfLag + gnpLag
+system <- list( Consumption = eqConsump, Investment = eqInvest,
+   PrivateWages = eqPrivWage )
+@
+OLS estimation:
+<<systemfit_replication,results=hide>>=
+kleinOls <- systemfit( system, data = KleinI )
+round( coef( summary( kleinOls ) ), digits = 3 )
+@
+2SLS estimation:
+<<systemfit_replication,results=hide>>=
+klein2sls <- systemfit( system, method = "2SLS", inst = inst, data = KleinI,
+   methodResidCov = "noDfCor" )
+round( coef( summary( klein2sls ) ), digits = 3 )
+@
+3SLS estimation:
+<<systemfit_replication,results=hide>>=
+klein3sls <- systemfit( system, method = "3SLS", inst = inst, data = KleinI,
+   methodResidCov = "noDfCor" )
+round( coef( summary( klein3sls ) ), digits = 3 )
+@
+iterated 3SLS estimation:
+<<systemfit_replication,results=hide>>=
+kleinI3sls <- systemfit( system, method = "3SLS", inst = inst, data = KleinI,
+   methodResidCov = "noDfCor", maxit = 500 )
+round( coef( summary( kleinI3sls ) ), digits = 3 )
+@
+
+Again, these commands return almost the same results
+as published in \citet{greene03}.%
+\footnote{
+There are two typos in Table~15.3 (p.~412).
+Please take a look at the errata
+\citep{greene06a}.
+}
+There are only two minor deviations,
+where these values differ merely in the last digit.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Greene (2003): Example 14.1 (Grunfeld's investment data)}
+\label{sec:grunfeld-greene}
+
+Third, we reproduce Example~14.1 of \citet[p.~340]{greene03}
+that is based on \citet{grunfeld58}.
+The data are available from the online complements to \citet{greene03},
+Table~F13.1
+(\url{http://pages.stern.nyu.edu/~wgreene/Text/tables/TableF13-1.txt}).
+Several different versions of the °Grunfeld° data set can be found,
+whereas the version of \citet{greene03}
+is considered incorrect \citep{cummins01}.
+However, we use this incorrect version to replicate the results
+in \citet{greene03}, Tables~14.1 and~14.2 (p.~351).%
+\footnote{
+A correct version of this data set with five additional firms
+is available as data set \code{Grunfeld} in the \pkg{Ecdat} package
+\citep{r-Ecdat-0.1-5}.
+}
+
+First, we load the data and the \pkg{plm} package,
+indicate the individual (cross-section) and time identifiers,
+and specify the formula to be estimated.
+Then, the system is estimated by OLS, pooled OLS, SUR, and pooled SUR.
+After each estimation, we show the commands to print the estimated coefficients,
+the $\sigma^2$ values of the OLS estimations, and
+the residual covariance matrix as well as the residual correlation matrix
+of the SUR estimations.
+<<systemfit_replication>>=
+data( "GrunfeldGreene" )
+library( "plm" )
+GGPanel <- plm.data( GrunfeldGreene, c( "firm", "year" ) )
+formulaGrunfeld <- invest ~ value + capital
+@
+OLS estimation (Table 14.2):
+<<systemfit_replication,results=hide>>=
+greeneOls <- systemfit( formulaGrunfeld,
+   data = GGPanel )
+round( coef( summary( greeneOls ) ), digits = 4 )
+round( sapply( greeneOls$eq, function(x){return(summary(x)$ssr/20)} ), digits = 3 )
+@
+pooled OLS (Table 14.2):
+<<systemfit_replication,results=hide>>=
+greeneOlsPooled <- systemfit( formulaGrunfeld,
+   data = GGPanel, pooled = TRUE )
+round( coef( summary( greeneOlsPooled$eq[[1]] ) ), digits = 4 ) #$
+sum( sapply( greeneOlsPooled$eq, function(x){return(summary(x)$ssr)}) )/100
+@
+SUR estimation (Table~14.1):
+<<systemfit_replication,results=hide>>=
+greeneSur <- systemfit( formulaGrunfeld, method = "SUR",
+   data = GGPanel, methodResidCov = "noDfCor" )
+round( coef( summary( greeneSur ) ), digits = 4 )
+round( greeneSur$residCov, digits = 3 ) #$
+round( summary( greeneSur )$residCor, digits = 3 ) #$
+@
+pooled SUR estimation (Table~14.1):
+<<systemfit_replication,results=hide>>=
+greeneSurPooled <- systemfit( formulaGrunfeld, method = "SUR",
+   data = GGPanel, pooled = TRUE, methodResidCov = "noDfCor",
+   residCovWeighted = TRUE )
+round( coef( summary( greeneSurPooled$eq[[1]] ) ), digits = 4 ) #$
+
+round( greeneSurPooled$residCov, digits = 3 ) #$
+round( cov( residuals( greeneSurPooled ) ), digits = 3 )
+round( summary( greeneSurPooled )$residCor, digits = 3 ) #$
+@
+
+These commands return
+nearly the same results as published in \citet{greene03}.%
+\footnote{
+There are several typos and errors in Table~14.1 (p.~412).
+Please take a look at the errata of this book
+\citep{greene06a}.
+}
+We present two different commands to print
+the residual covariance matrix of the pooled SUR estimation.
+The first calculates the covariance matrix without centering the residuals
+(see Section~\ref{sec:residcov});
+the returned values are equal to those published in \citet[p.~351]{greene03}.
+The second command calculates the residual covariance matrix
+after centering the residuals;
+these returned values are equal to those published in the errata
+\citep{greene06a}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection[Theil (1971): Example on p. 295ff (Grunfeld's investment data)]
+   {Theil (1971): Example on p.~295ff (Grunfeld's investment data)}
+
+Finally, we estimate an example taken from \citet[p.~295ff]{theil71}
+that is also based on \citet{grunfeld58}.
+The data are available in Table~7.1 of \citet[p.~296]{theil71}.
+They are a subset of the data set published by \citet{greene03}
+(see Section~\ref{sec:grunfeld-greene}).
+
+After extracting the data from the \code{GrunfeldGreene} data set,
+the individual (cross-section) and time identifiers are indicated.
+Then, the formula is specified,
+and the model is estimated by OLS and SUR.
+Commands to print the estimated coefficients are reported
+after each estimation.
+<<systemfit_replication>>=
+GrunfeldTheil <- subset( GrunfeldGreene,
+   firm %in% c( "General Electric", "Westinghouse" ) )
+GTPanel <- plm.data( GrunfeldTheil, c( "firm", "year" ) )
+formulaGrunfeld <- invest ~ value + capital
+@
+OLS estimation (page 295)
+<<systemfit_replication,results=hide>>=
+theilOls <- systemfit( formulaGrunfeld,
+   data = GTPanel )
+round( coef( summary( theilOls ) ), digits = 3 )
+@
+SUR estimation (page 300)
+<<systemfit_replication,results=hide>>=
+theilSur <- systemfit( formulaGrunfeld, method = "SUR",
+   data = GTPanel, methodResidCov = "noDfCor" )
+round( coef( summary( theilSur ) ), digits = 3 )
+@
+
+These commands
+return exactly the same results as published in
+\citet[pp.~295, 300]{theil71}.
+
+Now, we apply an $F$ test to check whether the slope parameters are equal
+for General Electric and Westinghouse (pages~313--315).
+Then we re-estimate the model under these restrictions on the coefficients.
+
+$F$ test (page 313--315)%
+\footnote{%
+The same restriction can be specified also symbolically by
+\code{RMatrix <- c("General.Electric\_value = Westinghouse\_value",
+"General.Electric\_capital = Westinghouse\_capital")}
+}
+<<systemfit_replication,results=hide>>=
+RMatrix <- rbind( c( 0, 1, 0, 0, -1, 0 ), c( 0, 0, 1, 0, 0, -1 ) )
+linear.hypothesis( theilSur, RMatrix )
+@
+
+Restricted SUR estimation (page~316)
+<<systemfit_replication,results=hide,keep.source=TRUE>>=
+theilSurRestr <- systemfit(formulaGrunfeld, method = "SUR",
+   data = GTPanel, methodResidCov = "noDfCor", restrict.matrix = RMatrix,
+   residCovRestricted = FALSE)
+round(coef(summary(theilSurRestr)), digits = 3)
+@
+
+The method \codeD{linear}{hypothesis} returns the same value of the $F$ statistic
+as published in \citet[p.~315]{theil71}.
+Hence, we arrive at the the same conclusion:
+we accept the null hypothesis (restrictions on the coefficients are true)
+at the 5~percent significance level.
+
+Also the results of the restricted SUR estimation are identical
+to the results published in
+\citet[p.~316]{theil71}.
+
 \input{systemfit_summary.tex}

### 2008-09-12 paper/trunk/runSweave.sh

@@ -4,10 +4,6 @@ echo "Sweave(\"systemfit.Rnw\")" | LC_ALL="C" R --no-save --no-restore
 echo "Stangle(\"systemfit.Rnw\", annotate=FALSE, split=TRUE,
    prefix=FALSE)" | LC_ALL="C" R --no-save --no-restore
 
-echo "Sweave(\"systemfit_reliability.Snw\")" | LC_ALL="C" R --no-save --no-restore
-echo "Stangle(\"systemfit_reliability.Snw\", annotate=FALSE, split=TRUE,
-   prefix=FALSE)" | LC_ALL="C" R --no-save --no-restore
-
 echo "Sweave(\"systemfit_timings.Snw\")" | R --no-save --no-restore
 echo "Stangle(\"systemfit_timings.Snw\", annotate=FALSE, split=TRUE,
    prefix=FALSE)" | LC_ALL="C" R --no-save --no-restore

### 2008-09-12 paper/trunk/systemfit.Rnw
moved section "Summary and Outlook" (systemfit_summary.tex) into master file (systemfit.Rnw)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@957 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 
-\input{systemfit_summary.tex}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Summary and outlook}\label{sec:Summmary}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\nopagebreak
+In this article, we have described some of the basic features of the
+\pkg{systemfit} package for estimating systems of linear
+equations.
+Many details of the estimation can be controlled by the user.
+Furthermore, the package provides some statistical tests
+for restrictions on the coefficients and consistency of 3SLS estimation.
+It has been tested on a variety of datasets and has produced satisfactory
+for a few years.
+While the \pkg{systemfit} package performs the basic fitting methods,
+more sophisticated tools exist.
+We hope to implement missing functionalities
+in the near future.
+% Some of these are discussed in the following.
+
+\subsubsection*{Unbalanced datasets}
+Currently, the \pkg{systemfit} package requires
+that all equations have the same number of observations.
+However, many data sets have unbalanced observations.%
+\footnote{
+For instance,
+forestry datasets typically contain many observations of inexpensive
+variables (stem diameter, tree count) and few expensive variables such
+as stem height or volume.
+}
+Simply dropping data points that do not contain observations for all
+equations may reduce the number of observations considerably, and
+thus, the information utilized in the estimation.
+Hence, it is our intention to include the capability for estimations
+with unbalanced data sets as described in \citet{schmidt77} in future
+releases of \pkg{systemfit}.
+
+\subsubsection*{Serial correlation and heteroscedasticity}
+For all of the methods developed in the package, the disturbances of
+the individual equations are assumed to be independent and identically
+distributed (iid).
+The package could be enhanced by the inclusion of methods to fit
+equations with serially correlated and heteroscedastic disturbances
+\citep{parks67}.
+
+\subsubsection*{Estimation methods}
+In the future, we wish to include more sophisticated estimation
+methods such as limited information maximum likelihood (LIML),
+full information maximum likelihood (FIML), generalized methods of
+moments (GMM) and spatial econometric methods
+\citep{paelinck79,anselin88}.
+
+\subsubsection*{Non-linear estimation}
+Finally, the \pkg{systemfit} package provides a function to estimate
+systems of non-linear equations.
+However, the function \code{nlsystemfit} is currently under
+development and the results are not yet always reliable due to
+convergence difficulties.
 

### 2008-09-12 paper/trunk/systemfit.Rnw
moved appendix section "Object returned by systemfit" (systemfit_compare_lm.tex) into master file (systemfit.Rnw)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@958 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 
-\input{systemfit_compare_lm.tex}
+\section[Object returned by systemfit]{Object returned by \code{systemfit}}
+\label{sec:returned-object}
+
+\code{systemfit} returns a list of class \code{systemfit}
+that contains the results that belong to the entire system
+of equations.
+One special element of this list is called \code{eq}.
+It is a list that contains one object for each estimated equation.
+These objects are of the class \codeD{systemfit}{equation}
+and contain the results that belong only to the regarding equation.
+
+\begin{table}[htbp]
+\centering
+\setlength{\tabcolsep}{4mm}
+{\ttfamily
+\begin{tabular}{|l|l|l|}
+\hline
+\textbf{\code{lm}} & \textbf{\code{systemfit}} & \textbf{\code{systemfit.equation}} \\
+\hline
+coefficients    & coefficients   & coefficients \\
+                & coefCov        & coefCov \\
+fitted.values   &                & fitted.values \\
+residuals       &                & residuals \\
+                & residCov       & \\
+                & residCovEst    & \\
+rank            & rank           & rank \\
+                &                & rank.sys \\
+                &                & nCoef.sys \\
+df.residual     & df.residual    & df.residual \\
+                &                & df.residual.sys \\
+call            & call           & \\
+terms           &                & terms \\
+                &                & inst \\
+weights         &                & \\
+contrasts       &                & \\
+xlevels         &                & \\
+offset          &                & \\
+model\textnormal{*} &            & model\textnormal{*} \\
+x\textnormal{**} &               & x\textnormal{**} \\
+y\textnormal{**} &               & y\textnormal{**} \\
+                &                & z\textnormal{**} \\
+                & iter           & \\
+                & eq             & \\
+                &                & eqnLabel \\
+                &                & eqnNo \\
+                & method         & method \\
+                & panelLike      & \\
+                & restrict.matrix& \\
+                & restrict.rhs   & \\
+                & restrict.regMat& \\
+                & control        & \\
+\hline
+\end{tabular}
+}
+\caption{Elements returned by \code{systemfit} and \code{lm}
+(* if requested by the user with default \code{TRUE},
+** if requested by the user with default \code{FALSE}).}
+\label{tab:compare-lm}
+\end{table}
+
+The elements returned by \code{systemfit}
+are similar to those returned by \code{lm},
+the basic tool for linear regressions in \proglang{R}.
+While some counterparts of elements returned by \code{lm}
+can be found directly in objects of class \code{systemfit},
+other counterparts are available for each equation
+in objects of class \codeD{systemfit}{equation}.
+This is demonstrated in Table~\ref{tab:compare-lm}.
 

### 2008-09-12 paper/trunk/systemfit.Rnw
moved appendix section "Computation times" (systemfit_timings.Snw) into master file (systemfit.Rnw)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@959 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 
-\input{systemfit_timings.tex}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Computation times}\label{sec:timings}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\SweaveOpts{prefix.string=trash/timings}
+<<echo=FALSE>>=
+options( prompt = "R> " )
+@
+
+Theoretically, one would expect
+that the calculations with the \pkg{Matrix} package are
+faster and more robust than calculations with the traditional method.
+To test this hypothesis, we use function \code{createSystemfitModel}
+to create a medium-sized multi-equation model with 8~equations,
+10~regressors in each equation (without constant),
+and 750~observations.
+Then, we estimated this model with and without using the \pkg{Matrix} package.
+Finally, the results are compared.
+<<systemfit_timings>>=
+library( "systemfit" )
+set.seed( 1 )
+systemfitModel <- createSystemfitModel( nEq = 8, nReg = 10, nObs = 750 )
+system.time(
+   fitMatrix <- systemfit( systemfitModel$formula, method = "SUR",
+      data = systemfitModel$data )
+)
+system.time(
+   fitTrad <- systemfit( systemfitModel$formula, method = "SUR",
+      data = systemfitModel$data, useMatrix = FALSE )
+)
+all.equal( fitMatrix, fitTrad )
+@
+The returned computation times clearly show
+that using the \pkg{Matrix} package makes the estimation faster.
+The comparison of the estimation results shows
+that both methods return the same results.
+The only differences between the returned objects
+are --- as expected --- the \code{call}
+and the stored control variable \code{useMatrix}.
+
+However, the estimation of rather small models is much slower
+with the \pkg{Matrix} package than without this package.
+Moreover, the differences in computation time accumulate,
+if the estimation is iterated.
+<<systemfit_timings>>=
+smallModel <- createSystemfitModel( nEq = 3, nReg = 4, nObs = 50 )
+system.time(
+   fitSmallMatrix <- systemfit( smallModel$formula, method = "SUR",
+      data = smallModel$data, maxit = 500 )
+)
+system.time(
+   fitSmallTrad <- systemfit( smallModel$formula, method = "SUR",
+      data = smallModel$data, maxit = 500, useMatrix = FALSE )
+)
+all.equal( fitSmallMatrix, fitSmallTrad )
+@
+As mentioned above, the usage of the \pkg{Matrix} package
+clearly increases the computation times
+for iterated (SUR) estimations of small models with small data sets.
 

### 2008-09-12 paper/trunk/runSweave.sh

@@ -4,10 +4,6 @@ echo "Sweave(\"systemfit.Rnw\")" | LC_ALL="C" R --no-save --no-restore
 echo "Stangle(\"systemfit.Rnw\", annotate=FALSE, split=TRUE,
    prefix=FALSE)" | LC_ALL="C" R --no-save --no-restore
 
-echo "Sweave(\"systemfit_timings.Snw\")" | R --no-save --no-restore
-echo "Stangle(\"systemfit_timings.Snw\", annotate=FALSE, split=TRUE,
-   prefix=FALSE)" | LC_ALL="C" R --no-save --no-restore
-
 echo "Sweave(\"systemfit_sem.Snw\")" | R --no-save --no-restore
 echo "Stangle(\"systemfit_sem.Snw\", annotate=FALSE, split=TRUE,
    prefix=FALSE)" | LC_ALL="C" R --no-save --no-restore

### 2008-09-12 paper/trunk/systemfit.Rnw
moved appendix section "Estimating systems of equations with sem" (systemfit_sem.Snw) into master file (systemfit.Rnw)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@960 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 
-\input{systemfit_sem.tex}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section[Estimating systems of equations with sem]
+   {Estimating systems of equations with \code{sem}}
+\label{sec:sem}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\SweaveOpts{prefix.string=trash/sem}
+<<echo=FALSE>>=
+options( prompt = "R> " )
+options( width = 75 )
+@
+
+This section compares the commands to estimate a system of equations
+by \code{sem} and \code{systemfit}.
+This comparison uses Klein's °Model I° (see Section~\ref{sec:KleinsModel}).
+Before starting the estimation,
+we load the \pkg{sem} and \pkg{systemfit} package
+as well as the required data set.
+<<systemfit_sem>>=
+library( "sem" )
+library( "systemfit" )
+data( "KleinI" )
+@
+
+First, we estimate the system by limited information maximum likelihood
+(LIML) with \code{sem}:
+<<systemfit_sem,keep.source = TRUE>>=
+limlRam <- matrix(c(
+   "consump  <-  corpProf",    "consump_corpProf",    NA,
+   "consump  <-  corpProfLag", "consump_corpProfLag", NA,
+   "consump  <-  wages",       "consump_wages",       NA,
+   "invest   <-  corpProf",    "invest_corpProf",     NA,
+   "invest   <-  corpProfLag", "invest_corpProfLag",  NA,
+   "invest   <-  capitalLag",  "invest_capitalLag",   NA,
+   "privWage <-  gnp",         "privWage_gnp",        NA,
+   "privWage <-  gnpLag",      "privWage_gnpLag",     NA,
+   "privWage <-  trend",       "privWage_trend",      NA,
+   "consump  <-> consump",     "s11", NA,
+   "privWage <-> privWage",    "s22", NA,
+   "invest   <-> invest",      "s33", NA),
+   ncol = 3, byrow = TRUE)
+class(limlRam) <- "mod"
+exogVar <- c("corpProf", "corpProfLag", "wages", "capitalLag", "trend",
+   "gnp", "gnpLag")
+endogVar <- c("consump", "invest", "privWage")
+allVar <- c(exogVar, endogVar)
+
+limlResult <- sem(ram = limlRam, S = cov(KleinI[ -1, allVar ]),
+   N = (nrow(KleinI) - 1), fixed.x = exogVar)
+print(limlResult)
+@
+
+Theoretically, the LIML results should be identical to OLS results.
+Therefore, we re-estimate this model by OLS with \code{systemfit}.
+<<systemfit_sem,keep.source=TRUE>>=
+eqConsump <- consump ~ corpProf + corpProfLag + wages
+eqInvest <- invest ~ corpProf + corpProfLag + capitalLag
+eqPrivWage <- privWage ~ gnp + gnpLag + trend
+system <- list(consump = eqConsump, invest = eqInvest,
+   privWage = eqPrivWage)
+olsResult <- systemfit(system, data = KleinI)
+print(olsResult)
+@
+As expected, the results are identical.
+
+Now, we estimate the system by full information maximum likelihood
+(FIML) with \code{sem}:
+<<systemfit_sem,keep.source = TRUE>>=
+fimlRam <- rbind(limlRam,
+   c("consump  <-> invest",   "s12", NA),
+   c("consump  <-> privWage", "s13", NA),
+   c("privWage <-> invest",   "s23", NA))
+class(fimlRam) <- "mod"
+
+fimlResult <- sem(ram = fimlRam, S = cov(KleinI[ -1, allVar ]),
+   N = (nrow(KleinI) - 1), fixed.x = exogVar)
+print(fimlResult)
+@
+
+Theoretically, results of an iterated SUR estimation should
+converge to FIML results.
+Therefore, we re-estimate this model by iterated SUR with \code{systemfit}.
+<<systemfit_sem>>=
+surResult <- systemfit( system, method = "SUR", data = KleinI,
+   methodResidCov = "noDfCor", maxit = 500 )
+print( surResult )
+@
+As expected, the results are rather similar.
 

### 2008-09-12 paper/trunk/runSweave.sh

@@ -3,7 +3,3 @@ echo "Sweave(\"systemfit.Rnw\")" | LC_ALL="C" R --no-save --no-restore
 
 echo "Stangle(\"systemfit.Rnw\", annotate=FALSE, split=TRUE,
    prefix=FALSE)" | LC_ALL="C" R --no-save --no-restore
-
-echo "Sweave(\"systemfit_sem.Snw\")" | R --no-save --no-restore
-echo "Stangle(\"systemfit_sem.Snw\", annotate=FALSE, split=TRUE,
-   prefix=FALSE)" | LC_ALL="C" R --no-save --no-restore

### 2008-09-12 pkg/inst/doc/systemfit.Rnw
copied the JSS paper into the package as a vignette
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@961 591db2bd-e0a0-4d22-8c19-2192dbb7da31
updated systemfit.bib from my large default BibTeX data base: the \proglang{} the \pkg{} commands are removed, a very few updates
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@962 591db2bd-e0a0-4d22-8c19-2192dbb7da31
added note that this is a slightly modified version of our JSS paper; removed JSS layout ("nojss" option)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@963 591db2bd-e0a0-4d22-8c19-2192dbb7da31

--- 2008-09-12 paper/trunk/systemfit.Rnw
@@
-%       $Id$
-
-\documentclass[article]{jss}
+\documentclass[article,nojss]{jss}
 \usepackage{csquotes}
@@
 \Abstract{
+This introduction to the \proglang{R} package \pkg{systemfit}
+is a slightly modified version of \cite{henningsen07a},
+published in the \emph{Journal of Statistical Software}.
+
 Many statistical analyses (e.g., in econometrics, biostatistics and
@@
 \begin{document}
+
+% initialisation stuff
+\SweaveOpts{engine=R}
+%\VignetteIndexEntry{Systemfit}
+%\VignetteDepends{systemfit,plm,sem}
+%\VignetteKeywords{R, system of simultaneous equations,
+% seemingly unrelated regression, two-stage least squares,
+% three-stage least squares, instrumental variables}
+%\VignettePackage{systemfit}
+<<echo=FALSE>>=
+options( prompt = "R> ", ctinue = "+  " )
+@
 
@@
 systems of linear equations in \proglang{R}
-\citep{r-project}.
+\citep{r-project-07}.
 Currently, the estimation methods

### 2008-09-12 pkg/inst/doc/systemfit.Rnw
removed some "\SweaveOpts{prefix.string=trash/<*>}" to avoid having a "trash" folder
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@964 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\SweaveOpts{prefix.string=trash/usage}
 <<echo=FALSE>>=
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\SweaveOpts{prefix.string=trash/replication}
 <<echo=FALSE>>=
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\SweaveOpts{prefix.string=trash/timings}
 <<echo=FALSE>>=
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\SweaveOpts{prefix.string=trash/sem}
 <<echo=FALSE>>=

### 2008-09-25 paper/trunk/Slides/systemfit_slides.tex
moved /paper/branches/useR2006/Slides to /paper/trunk/Slides/
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@975 591db2bd-e0a0-4d22-8c19-2192dbb7da31

--- 2006-06-15 paper/branches/useR2006/Slides/systemfit_slides.tex
@@
    \textbf{systemfit}\\ Simultaneous Equation Systems in R}
-\author{Arne Henningsen (University of Kiel, Germany)\\[0.4em]
+\author[Arne Henningsen and Jeff D.\ Hamann]{
+   Arne Henningsen (University of Kiel, Germany)\\[0.4em]
    Jeff D.\ Hamann (Forest Informatics, Inc., Corvallis, USA)}
-\date{Vienna, June 16, 2006}
+\date{useR!, Vienna, June 16, 2006}
 
@@
 \item spatial econometric methods
+\item simplify specification of parameter restrictions
 \item improving the function \texttt{nlsystemfit}
@@
    as well as for spatial data.
+\item And we would like to simplify the specification of parameter
+   restrictions.
 \item Finally, we would like to improve the function \texttt{nlsystemfit}
@@
 \begin{itemize}
-\item Therefore, we are consider whether we should reduce the number
+\item Therefore, we consider whether we should reduce the number
    of arguments.
-\item All arguments that are generally not used by unadvanced,
+\item All arguments that are generally not used by unadvanced users,
    could be subsumed into a list of control parameters

### 2008-10-16 pkg/inst/doc/systemfit.Rnw
changed encoding to utf-8 and added the line "\usepackage[utf8]{inputenc}"
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@969 591db2bd-e0a0-4d22-8c19-2192dbb7da31
removed "systemfit" from \VignetteDepends{}, because this gives an error message in R 2.6.2 (and possibly older versions of R) but not in R 2.7.2
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@985 591db2bd-e0a0-4d22-8c19-2192dbb7da31

@@
 \documentclass[article,nojss]{jss}
-\usepackage{csquotes}
-\MakeOuterQuote{°}
+\usepackage[utf8]{inputenc}
 \usepackage{amsmath}
@@
 %\VignetteIndexEntry{Systemfit}
-%\VignetteDepends{systemfit,plm,sem}
+%\VignetteDepends{plm,sem}
 %\VignetteKeywords{R, system of simultaneous equations,
@@
 However, estimating all equations simultaneously
-with a °generalized least squares° (GLS) estimator,
+with a ``generalized least squares'' (GLS) estimator,
 which takes the covariance structure of the residuals into account,
@@
 This estimation procedure is generally called
-°seemingly unrelated regression° \citep[SUR,][]{zellner62}.
+``seemingly unrelated regression'' \citep[SUR,][]{zellner62}.
 Another reason to estimate a system of equations simultaneously are
@@
 This simultaneity bias can be corrected for by
-applying a °two-stage least squares° (2SLS) estimation
+applying a ``two-stage least squares'' (2SLS) estimation
 to each equation.
@@
 in a simultaneous estimation of the system of equations
-by the °three-stage least squares° (3SLS) method
+by the ``three-stage least squares'' (3SLS) method
 \citep{zellner62b}.
@@
 Currently, the estimation methods
-°ordinary least squares° (OLS),
-°weighted least squares° (WLS),
-°seemingly unrelated regression° (SUR),
-°two-stage least squares° (2SLS),
-°weighted two-stage least squares° (W2SLS), and
-°three-stage least squares° (3SLS)
+``ordinary least squares'' (OLS),
+``weighted least squares'' (WLS),
+``seemingly unrelated regression'' (SUR),
+``two-stage least squares'' (2SLS),
+``weighted two-stage least squares'' (W2SLS), and
+``three-stage least squares'' (3SLS)
 are implemented.%
@@
 In this context,
-the term °weighted° in °weighted least squares° (WLS) and
-°weighted two-stage least squares° (W2SLS) means
+the term ``weighted'' in ``weighted least squares'' (WLS) and
+``weighted two-stage least squares'' (W2SLS) means
 that the \emph{equations} might have different weights
@@
 
-We can write the °stacked° system as
+We can write the ``stacked'' system as
 \begin{equation}
@@
 least squares (GLS) estimation leads to an efficient estimator for the coefficients.
-In this case, the GLS estimator is generally called °seemingly unrelated regression°
+In this case, the GLS estimator is generally called ``seemingly unrelated regression''
 (SUR) estimator \citep{zellner62}.
@@
 
-At the first stage new (°fitted°) regressors are obtained by
+At the first stage new (``fitted'') regressors are obtained by
 \begin{equation}
@@
 \end{equation}
-Then, these °fitted° regressors are substituted for the original regressors
+Then, these ``fitted'' regressors are substituted for the original regressors
 in Equation~\ref{eq:ols-wls-sur}
@@
 estimation leads to consistent and asymptotically more efficient estimates.
-This estimation procedure is generally called °three-stage least
-squares° \citep[3SLS,][]{zellner62b}.
+This estimation procedure is generally called ``three-stage least
+squares'' \citep[3SLS,][]{zellner62b}.
 The standard 3SLS estimator and its covariance matrix are obtained by
@@
 is not necessarily positive semidefinite \citep[p.~322]{theil71}.
-Hence, °it is doubtful whether [this formula] is really superior to 
-[Equation~\ref{eq:rcovNoDfCor}]° \citep[p.~322]{theil71}.
+Hence, ``it is doubtful whether [this formula] is really superior to
+[Equation~\ref{eq:rcovNoDfCor}]'' \citep[p.~322]{theil71}.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-<<echo=FALSE>>=
-options( prompt = "R> " )
-@
 
@@
 If \code{systemfit} is applied to
-a (classical) °seemingly unrelated regression° analysis with panel data,
+a (classical) ``seemingly unrelated regression'' analysis with panel data,
 it calls the hidden internal function \code{.systemfitPanel},
@@
 For instance, we choose class \code{dgeMatrix}
-(°real matrices in general storage mode°),
+(``real matrices in general storage mode''),
 for matrices $X_i$ in Equation~\ref{eq:model-array},
-class \code{dgCMatrix} (°general, numeric, sparse matrices
-in the (sorted) compressed sparse column format°)
+class \code{dgCMatrix} (``general, numeric, sparse matrices
+in the (sorted) compressed sparse column format'')
 for matrix $X$ in Equation~\ref{eq:model-matrices},
 and class \code{dspMatrix}
-(°symmetric real matrices in packed storage (one triangle only)°)
+(``symmetric real matrices in packed storage (one triangle only)'')
 for the residual covariance matrix $\SHat$.
@@
 A special feature of this function is the estimation
-of models with unobserved (°latent°) variables,
+of models with unobserved (``latent'') variables,
 which is not possible with \code{systemfit}.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-<<echo=FALSE>>=
-options( prompt = "R> " )
-@
 
@@
 Then, the usage of \code{systemfit}
-for a (classical) °seemingly unrelated regression° analysis
+for a (classical) ``seemingly unrelated regression'' analysis
 with panel data is described.
@@
 The following command estimates the model described above as
-°seemingly unrelated regression°.
+``seemingly unrelated regression''.
 <<systemfit_usage>>=
@@
 to estimate the model described above
-by °three-stage least squares° (3SLS).
+by ``three-stage least squares'' (3SLS).
 While the first command specifies the same instruments for all equations,
@@
 The \code{systemfit} function can also be used
-for a (classical) °seemingly unrelated regression° analysis
+for a (classical) ``seemingly unrelated regression'' analysis
 with panel data.
@@
 \footnote{
-Generally, panel data can be either in °long format°
+Generally, panel data can be either in ``long format''
 (different individuals are arranged below each other)
-or in °wide format°
+or in ``wide format''
 (different individuals are arranged next to each other).
-For this analysis, the data must be in °long format°.
+For this analysis, the data must be in ``long format''.
 }
@@
 
-Second, we try to replicate Klein's °Model I° \citep{klein50}
+Second, we try to replicate Klein's ``Model I'' \citep{klein50}
 that is described in \citet[p.~381]{greene03}.
@@
 (\url{http://pages.stern.nyu.edu/~wgreene/Text/tables/TableF13-1.txt}).
-Several different versions of the °Grunfeld° data set can be found,
+Several different versions of the ``Grunfeld'' data set can be found,
 whereas the version of \citet{greene03}
@@
 <<echo=FALSE>>=
-options( prompt = "R> " )
 options( width = 75 )
@@
 by \code{sem} and \code{systemfit}.
-This comparison uses Klein's °Model I° (see Section~\ref{sec:KleinsModel}).
+This comparison uses Klein's ``Model I'' (see Section~\ref{sec:KleinsModel}).
 Before starting the estimation,
